### 2021-09-06 AncestralPir/AnalysisCode/Manuscript_Github_AncestralPir.Rmd
Add data from Manuscript (berghei, chabaudi 24h, ancestral phylogeny and expression)

---
title: "Manuscript_Github_ancestral_pir_210801"
author: "Timothy Little"
date: "08/01/2021"
output: 
  html_document:
      toc: true
      number_sections: true
      toc_float: true
---

Got the 19 individual orthologs of PBANKA_0100500 from PlasmoDB, did orthology transformation twice to get everything. Then added the PccCB ortholog = 20 genes.
With the 20 genes did a MUSCLE (v3.8.31) alignment, PhyML (v3.1)* building with LG model 100 bootstrap replicates, on SEAVIEW v4.7.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/AncestralPir/AnalysisCode")
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r load_packages}
library('treeio') # Download trees from various formats
#May need to install ggtree from Github directly due to a recent bug: remotes::install_github("YuLab-SMU/ggtree")
library('ggtree') # Draw phylogenetic trees with ggplot
library('ggstance')
library('ggnewscale')
library('plotly')
library('ComplexHeatmap')
library('circlize') #for ColorRamp2 function
library('ggplot2')
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('cowplot')
library('kableExtra')
library('viridis') # color palettes
library('RColorBrewer') # color palettes
library('corrplot')
library('reshape2')
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('readxl') # for reading in Excel spreadsheets
library('readr') # for reading in everything else
# library('tidyverse')
library('dplyr')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r function_uniquify}
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
    name <- paste(vector_names[i],
                  df_num$rep[match(vector_names[i], df_num$names)],
                  sep = '_')
    #Increase rep counter
    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
    new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r tpm_function}
tpm = function(count_data, lengths) {

  #Print an error if the id columns do not match
  # first if statement to test that both have a GeneID column
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'Geneid', 
                                                  #This needs to be Geneid when some of the samples have 'id' in them
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      print('Warning:Count data and length gene ids DO NOT MATCH')
    }
  } else {
    print('Count data and/or length table supplied do not have an id column, 
          make sure they are ordered correctly')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = TRUE))){
    gene_names <- count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
    counts_tpm <- count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
  }else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = TRUE))){
    lengths.table = lengths[,grep(colnames(lengths), pattern = 'length', ignore.case = T)]
  }else{
    lengths.table <- lengths}
  #Making lengths.table a matrix sometimes throws up problems
  lengths.table = as.matrix(lengths.table)
  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  colnames(counts_tpm) <- 'tpm'
  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), 
                                            pattern = 'Geneid', 
                                            ignore.case = T), 
                                      drop = FALSE],
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

```{r function_get_ecdf_percentile}
ecdf_percentile <- function(df_tpm){
  Fn <- ecdf(
    unlist( #To convert into an atomic vector that ecdf will like.
      select_at(
        df_tpm,
        vars(
          matches('med_fpkm|med_tpm|^tpm$') #Default matches behaviour is ignore.case = TRUE
        ) #Could be med_fpkm or med_tpm.
      )
    )
  )
  #Using a lapply to make a list in case there are multiple ancestral genes.
  list_percentile <- lapply(
    df_tpm$Geneid[
      df_tpm$ancestral_ortholog == 1
      ],
    function(ancestor_gene){
      #For debugging
      # ancestor_gene <- "PCHCB_0101200"
      list(
        Fn(
          filter(df_tpm, Geneid %in% ancestor_gene) %>% 
            select_at(
              vars(
                matches('med_fpkm|med_tpm|^tpm$')
              ) #Could be med_fpkm or med_tpm
            ) 
        )
      )
    }
  )
  #Create data.frame of the output
  gene_percentile_df <- data.frame(
    Geneid = df_tpm$Geneid[
      df_tpm$ancestral_ortholog ==1
      ],
    percentile = unlist(
      lapply(
        list_percentile, `[`, 1
      )
    )
  )
  plot(Fn)
  return(gene_percentile_df)
}
# ecdf_percentile(cyno_genes_fpkm)
```

```{r get_phylip_tree}
phylog <- read.tree('PlasmoDBv48_GenesByOrthologs_ancestralpir_withPccCBandPberghei-PhyML_100bootstrap_tree.phylip')
```

```{r removing_unncessary_tips}
phylog <- drop.tip(phylog, 
                   tip = c("PKNOH_S130216000",
                           "YYE_04766",
                           "YYG_04764",
                           # "PY17X_0102100",
                           "PY06119",
                           'PYYM_0101500',
                           'PVL_110036700')
)
```

```{r gene_and_species_names}
gene_names <- phylog$tip.label %>% 
  str_split_fixed(pattern = '_P.', n =2) %>% 
  .[,1]
write(gene_names,
            file = 'gene_names_ancestral_includedintree.txt',
            sep = '\n')
phylog$old.tip.label <- phylog$tip.label

#Download the gene info with species names and add in the PccCB name.
ortho_gene_info_tbl <- rbind(
  read_csv('PlasmoDBv48_ancestralpir_orthologs_GenesByOrthologs_Summary.csv') %>% 
    select(c('Gene ID', Organism)) %>% 
    dplyr::rename(Geneid = 'Gene ID'),
  data.frame(Geneid = 'PCHCB_0101200',
             'Organism' = 'Plasmodium chabaudi chabaudi CB')
)

phylog$tip.label <- ortho_gene_info_tbl$Organism[match(phylog$old.tip.label, 
                                                       ortho_gene_info_tbl$Geneid)] %>% 
  str_replace(pattern = 'Plasmodium ', 
              replacement = 'P. ' ) %>% 
  str_replace(pattern = '(?<!cynomolgi) strain.+| curtisi|(?<=chabaudi) chabaudi|(?<=yoelii) yoelii', 
              replacement = '') %>% 
  str_extract(pattern = 'P. [[:lower:]]+(-like)?( [[:lower:]]+)?( M| B| CB| Sal-1| P01|)') %>% 
  gsub(pattern = 'chabaudi$', 
       replacement = 'chabaudi AS') %>% 
  gsub(pattern = 'strain ',
       replacement = '') %>% 
  uniquify(first_one = TRUE) 
#There shouldn't be any extra ones now
ovale_gene_names <- str_subset(gene_names, 
                               pattern = 'Poc')
```

'A glance of the blood stage transcriptome of a Southeast Asian Plasmodium ovale isolate'

ovale data is in TPM

```{r ovale_transcriptome_data}
ovale_path <- '../other_species_transcriptome_data/Brashear_2019_ovale_transcriptome_journal.pntd.0007850.s008.xlsx'
ovale <- read_xlsx(ovale_path, sheet = 1)
ovale <- ovale %>% 
  mutate(ancestral_ortholog = ifelse(Gene %in% gene_names, 1, 0))
#Sanity test
(ovale$ancestral_ortholog %>% sum) == 9

ovale_anc_v_nonanc <- ovale %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(med_tpm = median(TPM),
                   mean_tpm = mean(TPM),
                   max_tpm = max(TPM),
                   min_tpm = min(TPM))
#Median TPM of non-ancestral pirs
ovale_nonanc_medtpm <- ovale_anc_v_nonanc %>% filter(ancestral_ortholog == 0) %>% .$med_tpm
#Ratio of the TPM of the ancestrals to the median non-ancestrals
ovale_ancestral <- ovale %>% 
  filter(ancestral_ortholog ==1) %>% 
  mutate(Geneid = Gene,
         ratio = TPM/ovale_nonanc_medtpm,
         median_tpm = TPM) %>% 
  select(Geneid, ratio, median_tpm)
```

```{r ovale_percentile}
#Calculate the percentile of the ancestral v other pirs

ovale %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

ovale[ovale$ancestral_ortholog == 1,]

ovale_tpm_collated_pir <- ovale  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo') | ancestral_ortholog == 1) %>% 
  select(c(Gene, Description, ancestral_ortholog, TPM)) %>% 
  dplyr::arrange(desc(TPM))

#ecdf produces a function which can be used to find the percentile that a value is present in in a distribution
Fn_ovale <- ecdf(ovale_tpm_collated_pir$TPM)
plot(Fn_ovale)
list_ovale_percantile <- lapply(ovale_tpm_collated_pir$Gene[ovale_tpm_collated_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_ovale(filter(ovale_tpm_collated_pir, Gene %in% ancestr)$TPM))
                                })
ovale_percentile <- data.frame(Geneid = ovale_tpm_collated_pir$Gene[ovale_tpm_collated_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_ovale_percantile, `[`, 1)))

#Percentile for all genes
# Fn_ovale_all <- ecdf(ovale$TPM)
# plot(Fn_ovale_all)
# list_ovale_percantile_all <- lapply(ovale$Gene[ovale$ancestral_ortholog == 1], 
#                                 function(ancestr){
#                                   list(Fn_ovale_all(filter(ovale, Gene %in% ancestr)$TPM))
#                                 })
# (ovale_percentile_all <- data.frame(Geneid = ovale$Gene[ovale$ancestral_ortholog == 1],
#                                percentile = unlist(lapply(list_ovale_percantile_all, `[`, 1))))

#Wow must genes in the genome are expressed super low (52% expressed under 25 TPM)
# nrow(filter(ovale, TPM < 25)) /72.80

# ecdf_percentile(mutate(ovale, med_tpm = TPM, Geneid = Gene))
ovale_percentile_all <- ecdf_percentile(mutate(ovale, Geneid = Gene))

```

Data from eLife _A comparative transcriptomic analysis of replicating and dormant liver stages of the relapsing malaria parasite Plasmodium cynomolgi_ 2017.

Note: FPKM

```{r cynomolgi_transcriptome_data}
cyno_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'cynomolgi',
                        full.names = TRUE)
cyno <- read_xlsx(cyno_path, sheet = 1, skip = 3)
colnames(cyno) <- make.names(colnames(cyno))
apply(cyno %>% select_at(vars(matches('Sz|Hz'))), 2, sum)
cyno <- cyno %>% 
  mutate(ancestral_ortholog = ifelse(Gene.Id %in% gene_names,1,0))
(cyno$ancestral_ortholog %>% sum)
cyno_anc_v_nonanc <- cyno %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('Sz\\.|Hz\\.')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  #gsub to remove the numbers for easy summarising across the replicates
  mutate(stage = gsub(variable, pattern = '.[[:digit:]]', replacement = '')) %>% 
  group_by(ancestral_ortholog, stage) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
cyno_schiz_ratio <- filter(cyno_anc_v_nonanc, 
                           ancestral_ortholog == 1) %>% 
  filter(stage == 'Sz') %>% 
  .$median / 
  filter(cyno_anc_v_nonanc, ancestral_ortholog == 0) %>% 
  filter(stage == 'Sz') %>% 
  .$median
cyno_hyp_ratio <- filter(cyno_anc_v_nonanc, ancestral_ortholog == 1) %>% 
  filter(stage == 'Hz') %>% 
  .$median / 
  filter(cyno_anc_v_nonanc, ancestral_ortholog == 0) %>% 
  filter(stage == 'Hz') %>% 
  .$median
cyno_data <- data.frame(Geneid = cyno$Gene.Id[which(cyno$ancestral_ortholog == 1)],
                        ratio = cyno_schiz_ratio,
                        median_tpm = filter(cyno_anc_v_nonanc, ancestral_ortholog == 1) %>% 
                          filter(stage == 'Sz') %>% 
                          .$median)
```

```{r cyno_percentile}
#Calculate the percentile of the ancestral v other pirs
cyno_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PlasmoDBv46_PcynoM',
                        full.names = TRUE)
cyno_genes <- read_csv(cyno_gene_path) %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))

cyno_genes %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

cyno_genes %>% 
  filter(str_detect(Description, pattern = 'pseuodgene'))

cyno_genes_fpkm <- cyno %>% 
  select_at(vars(matches('^Gene|Average'))) %>% 
  melt(id.vars = 'Gene.Id') %>% 
  group_by(Gene.Id) %>% 
  dplyr::summarise(med_tpm = median(value),
                   sd_tpm = sd(value)) %>% 
  mutate(Description = cyno_genes$Description[match(Gene.Id, cyno_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Gene.Id %in% gene_names,1,0)) %>% 
  mutate(Geneid = Gene.Id)

cyno_pir_fpkm <- cyno_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_tpm)) %>% 
  dplyr::arrange(desc(med_tpm))
cyno_pir_fpkm[which(cyno_pir_fpkm$Geneid %in% gene_names),]
cyno_pir_fpkm[which(cyno_pir_fpkm$ancestral_ortholog == 1),]

#The percentile among the pir genes
Fn_cyno <- ecdf(cyno_pir_fpkm$med_tpm)
plot(Fn_cyno)
list_cyno_percantile <- lapply(cyno_pir_fpkm$Geneid[cyno_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_cyno(filter(cyno_pir_fpkm, Geneid %in% ancestr)$med_tpm))
                                })
cyno_percentile <- data.frame(Geneid = cyno_pir_fpkm$Geneid[cyno_pir_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_cyno_percantile, `[`, 1)))

#The percentile among all genes
Fn_cyno_all <- ecdf(cyno_genes_fpkm$med_tpm)
plot(Fn_cyno_all)
list_cyno_percentile_all <- lapply(cyno_genes_fpkm$Geneid[cyno_genes_fpkm$ancestral_ortholog == 1], 
                                   function(ancestr){
                                     list(Fn_cyno_all(filter(cyno_genes_fpkm, Geneid %in% ancestr)$med_tpm))
                                   })
cyno_percentile_all <- data.frame(Geneid = cyno_genes_fpkm$Geneid[cyno_genes_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_cyno_percentile_all, `[`, 1)))

ecdf_percentile(cyno_genes_fpkm)

```

Vivax data from Siegal et al 2020 SciRep 'Analysis of Plasmodium vivax schizont transcriptomes from field isolates reveals heterogeneity of expression of genes involved in host-parasite interactions'.

* FPKM data. 
* Clinical isolates.

```{r}
vivax_genes_fpkm <- read_xlsx('../other_species_transcriptome_data/Siegal2020_vivax-schizont-rnaseq_41598_2020_73562_MOESM2_ESM.xlsx',
          sheet = 3,
          skip = 1,
          col_names = TRUE) %>% 
  mutate(Geneid = gene_id,
         rpkm = `mean RPKM`,
         ancestral_ortholog = ifelse(gene_id %in% gene_names,1,0)) %>% 
  select(Geneid, 
         rpkm,
         ancestral_ortholog)

# sum(vivax_genes_fpkm$ancestral_ortholog) == 1

group_by(vivax_genes_fpkm, 
         ancestral_ortholog) %>% 
  dplyr::summarise(median = median(rpkm),
                   standev = sd(rpkm))

#The percentile among all genes
Fn_vivax_all <- ecdf(vivax_genes_fpkm$rpkm)
plot(Fn_vivax_all)
list_vivax_percentile_all <- lapply(vivax_genes_fpkm$Geneid[vivax_genes_fpkm$ancestral_ortholog == 1], 
                                   function(ancestr){
                                     list(Fn_vivax_all(filter(vivax_genes_fpkm, Geneid %in% ancestr)$rpkm))
                                   })
vivax_percentile_all <- data.frame(Geneid = vivax_genes_fpkm$Geneid[vivax_genes_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_vivax_percentile_all, `[`, 1)))

```

P. chabaudi chabaudi CB data from Lim et al, 2019.

```{r cb_transcriptome_data}
cb_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PccCB_rnaseq',
                        full.names = TRUE)
cb <- read_delim(cb_path, delim = '\t')
apply(cb[,-1], 2, sum) #FPKM, do not add up to 1e6, although they are close
cb <- cb %>% 
  mutate(Geneid = id) %>% 
  select(-id) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(cb$ancestral_ortholog %>% sum)
cb_anc_v_nonanc <- cb %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('CB')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
cb_data <- data.frame(Geneid = cb$Geneid[which(cb$ancestral_ortholog == 1)],
                      ratio = cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 1)]/
                        cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 0)],
                      median_tpm = cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 1)])
```

Used ftp://ftp.sanger.ac.uk/pub/project/pathogens/Plasmodium/chabaudi/CB_v2/ for the gene names

```{r cb_percentile}
#Calculate the percentile of the ancestral v other pirs
cb_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PccCB_gene',
                        full.names = TRUE)

read_delim(cb_gene_path, delim = '\t')[!(apply((apply(read_delim(cb_gene_path, delim = '\t'),
             1,
             str_extract,
             pattern = 'pseudo') %>%
         is.na),
      2, all)),] #Are there any pseudo columns? Yes : 'attributes'

cb_genes <- read_delim(cb_gene_path, delim = '\t', 
                       col_names = c('Geneid', 'Description'))

cb_genes %>% 
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

# cb_genes %>% 
#   group_by(Description) %>% 
#   dplyr::summarise(num = n()) %>% 
#   dplyr::arrange(desc(num))

cb_genes_fpkm <- cb  %>% 
  select_at(vars(matches('TCB|Geneid'))) %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_fpkm = median(value),
                   sd_fpkm = sd(value)) %>% 
  mutate(Description = cb_genes$Description[match(Geneid, cb_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

cb_pir_fpkm <- cb_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'CIR|PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_fpkm)) %>% 
  dplyr::arrange(desc(med_fpkm))
cb_pir_fpkm[which(cb_pir_fpkm$Geneid %in% gene_names),]

Fn_cb <- ecdf(cb_pir_fpkm$med_fpkm)
plot(Fn_cb)
list_cb_percantile <- lapply(cb_pir_fpkm$Geneid[cb_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_cb(filter(cb_pir_fpkm, Geneid %in% ancestr)$med_fpkm))
                                })
cb_percentile <- data.frame(Geneid = cb_pir_fpkm$Geneid[cb_pir_fpkm$ancestral_ortholog == 1],
                            percentile = unlist(lapply(list_cb_percantile, `[`, 1)))

ecdf_percentile(cb_pir_fpkm)
cb_percentile_all <- ecdf_percentile(cb_genes_fpkm)

```

vinckei data from Ramaprasad 2019 'A fast and cost-effective microsampling protocol incorporating reduced animal usage for time-series transcriptomics in rodent malaria parasites'

PlasmoDB v48 is not updated enough to have the vinckei assembly that the paper used. By BLASTing the PlasmoDB sequence against the PvvCY assembly on https://www.ncbi.nlm.nih.gov/assembly/GCA_900681995.1 and using the gff file from ftp://ftp.ncbi.nlm.nih.gov/genomes/all/GCA/900/681/995/GCA_900681995.1_PVVCY_v1 my guess is that it is this gene:

LR215057.1	EMBL	gene	71066	72501	.	-	.	ID=gene-PVVCY_0100190;Name=PVVCY_0100190;gbkey=Gene;gene_biotype=protein_coding;locus_tag=PVVCY_0100190

corresponding CDS data:
\>lcl|LR215057.1_cds_VEV54408.1_19 \[locus_tag=PVVCY_0100190\] \[protein=PIR protein CIR protein\] \[protein_id=VEV54408.1\] [location=complement(join(71066..71140,71320..72291,72493..72501))] \[gbkey=CDS\]

Also have downloaded the features table from the same ftp site as above.

Is this TPM or FPKM? FPKM

```{r vinck_data}
vinck_path <- list.files('../other_species_transcriptome_data/', 
                      #Use glob2rx to get the regex we want
                      pattern = glob2rx('*vinckei*xlsx*'),
                      full.names = TRUE)
vinck <- read_xlsx(vinck_path, skip =1)
select_at(vinck, vars(matches('[[:digit:]]h'))) %>% apply(., 2, sum)
#Definitely FPKM
vinck <- vinck %>% 
  mutate(Geneid = gene_id) %>% 
  select(-gene_id) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(vinck$ancestral_ortholog %>% sum)
vinck_anc_v_nonanc <- vinck %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('[[:digit:]]h')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
vinck_data <- data.frame(Geneid = vinck$Geneid[which(vinck$ancestral_ortholog == 1)],
                        ratio = vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 1)]/
                        vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r vinckei_percentile}
#Calculate the percentile of the ancestral v other pirs
vinck_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'vinckei_feature',
                        full.names = TRUE)

read_delim(vinck_gene_path, delim = '\t')[!(apply((apply(read_delim(vinck_gene_path, delim = '\t'),
             1,
             str_extract,
             pattern = 'pseudo') %>%
         is.na),
      2, all)),] #Are there any pseudo columns? Yes : 'attributes'

vinck_genes <- read_delim(vinck_gene_path, delim = '\t') %>% 
  mutate(Geneid = locus_tag,
         Description = ifelse(is.na(attributes),
                              name,
                              paste(name, attributes))) %>% 
  select(c(Geneid, Description)) %>% 
  filter(!is.na(Description))

vinck_genes %>% filter(Geneid %in% 'PVVCY_0200040')

vinck_genes %>% 
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

vinck_genes_fpkm <- vinck %>% 
  select_at(vars(matches('^[[:digit:]]+h|Geneid'))) %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_fpkm = median(value),
                   sd_fpkm = sd(value)) %>% 
  mutate(Description = vinck_genes$Description[match(Geneid, vinck_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

vinck_pir_fpkm <- vinck_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_fpkm)) %>% 
  dplyr::arrange(desc(med_fpkm))
vinck_pir_fpkm[which(vinck_pir_fpkm$Geneid %in% gene_names),]

Fn_vinck <- ecdf(vinck_pir_fpkm$med_fpkm)
plot(Fn_vinck)
list_vinck_percantile <- lapply(vinck_pir_fpkm$Geneid[vinck_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_vinck(filter(vinck_pir_fpkm, Geneid %in% ancestr)$med_fpkm))
                                })
vinck_percentile <- data.frame(Geneid = vinck_pir_fpkm$Geneid[vinck_pir_fpkm$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_vinck_percantile, `[`, 1)))
ecdf_percentile(vinck_pir_fpkm)
vinck_percentile_all <- ecdf_percentile(vinck_genes_fpkm)

```

This distribution of gene expression is different compared to that of the other samples..

##Coatneyi data

'Distinct amino acid and lipid perturbations characterize acute versus chronic malaria'

from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE103259

Want to convert the counts from this file into TPM with the transcript lengths.

_P. coatneyi_ _ancestral pir_ is not annotated as a _pir_ !

```{r coatneyi_data}
coat_path <- list.files('../other_species_transcriptome_data/', 
                      pattern = '2019_coatneyi_rawcounts',
                      full.names = TRUE)
coat <- read_xlsx(coat_path, skip =3, sheet  =2)
colnames(coat)[1:2] <- c('gene', 'Geneid')
coat <- coat %>% 
  select(-gene) %>% 
  arrange(Geneid) 
#apply(coat[,-1], 2, sum)
plot(apply(coat[,-1], 2, sum), apply(coat[,-1], 2, function(x) sum(x > 0)),
     main = 'coatneyi data features detected vs total counts',
     xlab = 'Total counts of sample',
     ylab = 'Total detected genes (>0 counts) per sample')
#Great ranges in the sum of features counted in each sample, should filter out low ones
coat <- coat[,c(TRUE, apply(coat[,-1], 2, sum) > 1e7)]

coat_transcript_lengths_path <- list.files('../../genomes_genes_annotations/', 
                      pattern = 'PlasmoDBv48_Pcoatneyi',
                      full.names = TRUE)
coat_transcript_lengths <- read_csv(coat_transcript_lengths_path) %>% 
  mutate(Geneid = `Gene ID`) %>% 
  select(c(Geneid, `Transcript Length`))

coat_tpm <- tpm(coat, 
                coat_transcript_lengths)

apply(coat_tpm[,-1], 
      2, 
      sum)

coat_tpm <- coat_tpm %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

(coat_tpm$ancestral_ortholog %>% sum)
coat_anc_v_nonanc <- coat_tpm %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('tpm')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
                   standev = sd(value))
coat_data <- data.frame(Geneid = coat_tpm$Geneid[which(coat_tpm$ancestral_ortholog == 1)],
                        ratio = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)]/
                          coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r coatneyi_percentile}
#Calculate the percentile of the ancestral v other pirs
coat_gene_info <- read_csv(coat_transcript_lengths_path) %>% 
  mutate(Geneid = `Gene ID`) %>% 
  select(Geneid, `Product Description`)

coat_gene_info %>% 
  filter(str_detect(`Product Description`, pattern = 'KIR')) %>% 
  group_by(`Product Description`) %>% 
  dplyr::summarise(n())

coat_tpm_collated <- coat_tpm %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(value),
                   sd_tpm = sd(value)) %>% 
  mutate(Product = coat_gene_info$`Product Description`[match(Geneid, coat_gene_info$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

coat_tpm_collated_pir <- coat_tpm_collated  %>% 
  filter(str_detect(Product, pattern = 'KIR') | ancestral_ortholog == 1) %>% 
  dplyr::arrange(desc(med_tpm))

Fn_coat <- ecdf(coat_tpm_collated_pir$med_tpm)

plot(Fn_coat)

list_coat_percantile <- lapply(coat_tpm_collated_pir$Geneid[coat_tpm_collated_pir$ancestral_ortholog == 1], 
                               function(ancestr){
                                 list(Fn_coat(filter(coat_tpm_collated_pir, Geneid %in% ancestr)$med_tpm))
                               })
coat_percentile <- data.frame(Geneid = coat_tpm_collated_pir$Geneid[coat_tpm_collated_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_coat_percantile, `[`, 1)))

ecdf_percentile(coat_tpm_collated_pir)
coat_percentile_all <- ecdf_percentile(coat_tpm_collated)
```

berghei data - using Brugat et al., 2017

Highest expressed pir gene of any kind is 'PBANKA_0201250' - a pseudogene.

```{r berghei_data}
berg_path <- list.files('../other_species_transcriptome_data/',
                        pattern = 'Expt02',
                        recursive = TRUE,
                        full.names = TRUE)
berg <- read_csv(berg_path) %>%  
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(berg$ancestral_ortholog %>% sum)
berg_anc_v_nonanc <- berg %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(-matches('Geneid')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
berg_data <- data.frame(Geneid = berg$Geneid[which(berg$ancestral_ortholog == 1)],
                        ratio = berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 1)]/
                          berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r berghei_percentile}

berg_descr <- read_csv('../../standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv') %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))
berg_pir <- berg %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(value),
            sd_tpm = sd(value)) %>% 
  mutate(Description = berg_descr$Description[match(Geneid, berg_descr$Geneid)]) %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  filter(str_detect(Description, pattern = 'pseudo|fragment', negate = TRUE)) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0)) %>% 
  dplyr::arrange(desc(med_tpm))

Fn_bir <- ecdf(berg_pir$med_tpm)
plot(Fn_bir)
list_bir_percentile <- lapply(berg_pir$Geneid[berg_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_bir(filter(berg_pir, Geneid %in% ancestr)$med_tpm))
                                })
bir_percentile <- data.frame(Geneid = berg_pir$Geneid[berg_pir$ancestral_ortholog == 1],
                             percentile = unlist(lapply(list_bir_percentile, `[`, 1)))

berg_percentile_all <- ecdf_percentile(
  berg %>% 
    melt(id.vars = 'Geneid') %>% 
    group_by(Geneid) %>% 
    dplyr::summarise(med_tpm = median(value),
              sd_tpm = sd(value)) %>% 
    mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
)
```

Py17XNL data from gametocytes (Hart ... Lindner et al 2019) 'Plasmodium male gametocyte development and transmission are critically regulated by the two putative deadenylases of the CAF1/CCR4/NOT complex'

```{r yoelii_data}
yoel_path <- list.files('..',
                        pattern = 'Hart_2019_Pyeolii',
                        recursive = TRUE,
                        full.names = TRUE)
#Something in the excel doc upsets read_xlsx so don't worry about the huge number of warnings..
yoel_counts_full <- read_xlsx(yoel_path, 
                         sheet = 2) %>% 
  mutate(Geneid = GeneID...1,
         Description = `[Product Description]`,
         Counts = `489 Average`) %>% 
  select(c(Geneid, Description, Counts)) %>% 
  arrange(Geneid)

yoel_counts_full %>%
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

yoel_counts <- select(yoel_counts_full, -Description)

yoel_lengths_path <- list.files('..',
                                pattern = 'PlasmoDBv46_Pyoelii17X_',
                                recursive = TRUE,
                                full.names = TRUE)
yoel_lengths <- read_csv(yoel_lengths_path) %>% 
  mutate(Geneid = `Gene ID`,
         length = `Transcript Length`) %>% 
  select(c(Geneid, length))
#One gene 'PY17X_0902700' has two different length transcripts - take median
yoel_lengths %>% group_by(Geneid) %>% dplyr::summarise(n = n()) %>% dplyr::arrange(desc(n))

#This gives 5074.5 so need to make sure that not having an integer count messes up things but it shouldn't
yoel_lengths <- yoel_lengths %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(length = median(length))
nrow(yoel_lengths)
nrow(yoel_counts)

yoel_tpm <- tpm(count_data = yoel_counts, lengths = yoel_lengths)
yoel_tpm <- yoel_tpm %>% 
  mutate(Description = yoel_counts_full$Description[match(Geneid, 
                                                          yoel_counts_full$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

yoel_tpm_pir <- yoel_tpm %>% 
  filter(str_detect(Description, pattern = 'PIR|YIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  dplyr::arrange(desc(tpm))

Fn_yoel <- ecdf(yoel_tpm_pir$tpm)

plot(Fn_yoel)

list_yoel_percantile <- lapply(yoel_tpm_pir$Geneid[yoel_tpm_pir$ancestral_ortholog == 1], 
                               function(ancestr){
                                 list(Fn_yoel(filter(yoel_tpm_pir, Geneid %in% ancestr)$tpm))
                               })
yoel_percentile <- data.frame(Geneid = yoel_tpm_pir$Geneid[yoel_tpm_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_yoel_percantile, `[`, 1))) 
yoel_percentile

yoel_percentile_all <- ecdf_percentile(yoel_tpm)

```

```{r chabaudi_percentile}

chab_info <- read_csv('../../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv') %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))

chab_info %>% 
  filter(str_detect(Description, pattern = 'IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

chab_tpm <- read_xlsx('../other_species_transcriptome_data/Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
                      sheet = 'chab_TPM_avg_rMT') %>% 
  melt(variable.name = 'stage', value.name = 'tpm') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(tpm),
                sd_tpm = sd(tpm)) %>% 
  mutate(Description = chab_info$Description[match(Geneid, 
                                                       chab_info$Geneid)]) %>% 
      mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

chab_pir <- chab_tpm %>% 
  filter(str_detect(Description, pattern = 'PIR|YIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>%
  dplyr::arrange(desc(med_tpm))

Fn_chab <- ecdf(chab_pir$med_tpm)
plot(Fn_chab)
list_chab_percentile <- lapply(chab_pir$Geneid[chab_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_chab(filter(chab_pir, Geneid %in% ancestr)$med_tpm))
                                })
chab_percentile <- data.frame(Geneid = chab_pir$Geneid[chab_pir$ancestral_ortholog == 1],
                            percentile = unlist(lapply(list_chab_percentile, `[`, 1)))

chab_percentile_all <- ecdf_percentile(chab_tpm)
```

## Phylogenetic Tree / Heatmap of percentile compared to all genes

```{r tree_heatmap_all-genes}

ancestral_df <- bind_rows(cyno_percentile_all,
                          ovale_percentile_all,
                          cb_percentile_all,
                          vinck_percentile_all,
                          coat_percentile_all,
                          yoel_percentile_all,
                          berg_percentile_all,
                          chab_percentile_all,
                          vivax_percentile_all)
ancestral_data <- lapply(phylog$old.tip.label, function(id){
  #If any of the pattern is in the string one will be TRUE
  ratio <- if(any(str_detect(string = paste(id), 
                             pattern = ancestral_df$Geneid))){
    index_id <- str_detect(string = paste(id), 
                           pattern = ancestral_df$Geneid)
    ancestral_df$percentile[index_id]*100
  }else{
    NA
  }
  return(list(ratio = ratio))
}) %>% 
  bind_rows
ancestral_data <- as.data.frame(ancestral_data)

rownames(ancestral_data) <- phylog$tip.label

breaks <- seq(0, 100, by = 10)

(tr <- ggtree(phylog, 
            branch.length = 'none',
            layout = 'circular') +
  geom_tiplab2(aes(angle=angle, x = x + 2.5), 
               cex = 3, 
               color='black',
               font = 3) +
  ggplot2::xlim(0,20) 
)
```

```{r heatmap_with_tree_figure6}

p <- gheatmap(tr, 
              ancestral_data, 
              offset=.2, 
              width=.15, 
              colnames=FALSE) +
  scale_fill_viridis_c(option='C', 
                       name=substitute(
                         atop(
                           '',
                           atop(
                             '', 
                             atop(
                               atop(
                                 paste(
                                   "Percentile of ",
                                   italic('ancestral pir')
                                 ), 
                                 " expression compared"
                               ), 
                               "to all genes"
                             )
                           )
                         )
                       ), 
                       #Use substitute to ensure that the italics function is passed to the name argument
                       #atop ensures that the first line is above the word 'expression'
                       #Two atops used for every two lines 
                       # as a hacky way to ensure that the lines are closer together
                       na.value = 'white', 
                       breaks = breaks) +
  #Then alter theme as hacky way to make text size back to normal and increase margin so the empty line will fit
  theme(legend.title = element_text(size = 15,
                                    margin = margin(t = -2, 
                                                    r = 0, 
                                                    b = 0, 
                                                    l = 0,
                                                    unit = "mm")))

p
```

### 2021-09-06 Pchabaudi24h/Manuscript_Github_Pchabaudi_calculateTPM.Rmd
Add data from Manuscript (berghei, chabaudi 24h, ancestral phylogeny and expression)

---
title: "Manuscript_Github_Pchabaudi_calculateTPM"
author: "Timothy Little"
date: "22/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

knitr::opts_knit$set(root.dir = 
                       ".")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

library('readr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('DESeq2')
library("pheatmap")
library("RColorBrewer")
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots and colours
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('readxl') # for reading in Excel spreadsheets
library('plotly')
library('scales')
library('processx') # for downloading the plotly plots
library('cowplot')
library('ComplexHeatmap')
library('ggbiplot')
library('ggpattern')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select
```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r order_vec}

order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info <- read.csv('../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)

chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID) %>% 
  mutate(Geneid = Gene.ID) %>% 
  select(Geneid, Transcript.Length)

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)

```

```{r cir_info}

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')
```

```{r times_stages_labels}
#Use the rMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r calc_tpm}

chab_count_table <- read.csv('Manuscript_Suppinfo_chabaudi_24h_counts.csv',
                             row.names = 1)

twentyfour_ct_comb <- tpm(count_data = chab_count_table, 
                          lengths = chabaudi_transcript_lengths)
```

```{r tpm_rescale}

#Setting twentyfour_tpm_ct core dataset - remove pseudogene and re-scaled

twentyfour_tpm_nopseudo <- twentyfour_ct_comb %>% 
  filter(!Geneid %in% 'PCHAS_0420800')

twentyfour_tpm_rescale_nopseudo <- lapply(colnames(twentyfour_tpm_nopseudo), 
                                          function(column_name){
                                            column <- getElement(twentyfour_tpm_nopseudo, 
                                                                 paste(column_name))
                                            if(is.numeric(column)){
                                              #is.numeric to avoid trying to scale the Geneid column
                                              tpm_scale_factor <- 1e6 / sum(column)
                                              tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                            }else{
                                              tpm_scaled <- column
                                            }
                                            return(tpm_scaled)
                                          }) %>% 
  bind_cols

colnames(twentyfour_tpm_rescale_nopseudo) <- colnames(twentyfour_tpm_nopseudo)
# apply(twentyfour_tpm_rescale_nopseudo[,-1], 2, sum)

#Setting the re-scaled version as the default from now on
#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo
```

```{r clean_up_and_average}
# Change the names to remove superfluous 'rMT'

twentyfour_ct_comb <- rename_with(twentyfour_ct_comb, 
                                  ~ str_extract(.x,pattern = '(?<=rMT_).+'),
                                  matches('rMT'))

twentyfour_rMT_avg <- melt(twentyfour_ct_comb, 
                           variable.name = 'sample',
                           value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
  group_by(Geneid, stage) %>% 
  summarise(mean_tpm = mean(tpm)) %>% 
  dcast(Geneid ~ stage)
```

```{r save}
write_csv(twentyfour_rMT,
          file = 'Pchabaudi_TPM_allBioReps.csv')

write_csv(twentyfour_rMT_avg,
          file = 'Pchabaudi_TPM_average.csv')
```

### 2021-09-06 Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Add data from Manuscript (berghei, chabaudi 24h, ancestral phylogeny and expression)

---
title: "Manuscript_Github_Pchabaudi_pir"
author: "Timothy Little"
date: "22/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

knitr::opts_knit$set(root.dir = 
                       ".")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

library('readr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('DESeq2')
library("pheatmap")
library("RColorBrewer")
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots and colours
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('readxl') # for reading in Excel spreadsheets
library('plotly')
library('scales')
library('processx') # for downloading the plotly plots
library('cowplot')
library('ComplexHeatmap')
library('ggbiplot')
library('ggpattern')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

#Global save var - set to TRUE if you want to save figures.
global_save_var <- TRUE
```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r ordering_vec}
order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info <- read.csv('../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)

chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID)

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)

```

```{r cir_info}

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')
```

```{r times_stages_labels}
#Use the RMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r read_TPM}
#Need to use the calculateTPM script first.

twentyfour_rMT <- read_csv('Pchabaudi_TPM_allBioReps.csv')

twentyfour_avgtpm <- list(
  'RMT' = list(
    tpm = read_csv('Pchabaudi_TPM_average.csv'),
    experiment = 'RMT'
  )
)

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9', #L
        '#E69F00', #S
        '#999999'),
  sort(
    unique(cir_info$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(cir_info$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$subfam)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(cir_info$subfam == clade_name),
                                   ')'
                                 )
                               })

```

## Figure 3c - Highest expressed all _pir_ heatmap

```{r highest_expressed_stage_heatmap_fig3c}

num_SorL <- cir_info %>% 
  group_by(SorL) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
                        num_SorL$SorL)

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm), 
       function(list_name){

         combined_data_pir <- getElement(twentyfour_avgtpm, 
                                     list_name)$tpm %>% 
           filter(Geneid %in% cir_id) %>% 
           column_to_rownames(var = 'Geneid')

         #Ordering that a more understandable order

         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                            order_vector)),
                                                drop = FALSE]

         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                         means = apply(combined_data_pir, 1, mean),
                                         subfam = cir_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
                                                                            cir_info$Geneid)]) %>% 
           filter(means > 1) %>% 
           arrange(subfam, 
                   desc(means)) %>% 
           .$Geneid

         #Order the heatmap data by the order defined above (SorL, then mean expression).
         combined_data_pir_ht <- combined_data_pir[
           order(
             match(rownames(combined_data_pir),
                   combined_data_ordered),
             na.last = NA #Remove NAs
           ),
         ]

         #Change the colnames to the times
         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                                                           pattern = '[[:digit:]]{2}h')

         #Order by time point
         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                      drop = FALSE] 
         #Add a column for the SorL
         combined_data_pir_ht$SorL <- cir_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
                                                                cir_info$Geneid)]

         return(combined_data_pir_ht)
       })
names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)

maxcol <- ceiling(
  max(
    unlist(
      lapply(combined_data_pir_ht_list,
             function(list_element){
               max(
                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
               )
             }
      )
    )
  )
)

lapply(names(combined_data_pir_ht_list),
       function(list_name_to_order){

         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
                                            list_name_to_order)

         pir_highest_tpm_order <- rownames(combined_data_pir_ht)

         #Sub-family annotation
         clade_info <- combined_data_pir_ht$SorL
         annotation_col <- SorL_col

         #Subfamily annotation

         clade_ha <- rowAnnotation(
           'sub-family' = anno_simple(
             clade_info,
             col = annotation_col,
           ),
           annotation_name_gp = gpar(fontsize = 10)
         )

         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
                                                 labels = 'ancestral',
                                                 labels_gp = gpar(fontsize = 7),
                                                 link_width = unit(3, "mm")))

         #lapply across the datasets now, for each order
         ht_list <- lapply(names(combined_data_pir_ht_list),
                           function(list_name){

                             combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
                                                                list_name)$tpm %>% 
                               column_to_rownames(var = 'Geneid')

                             #Get the times instead of the stage names.
                             colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                                                           pattern = '[[:digit:]]{2}h')

                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                                          drop = FALSE]

                             Heatmap(
                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
                               cluster_rows = FALSE,
                               cluster_columns = FALSE,
                               show_row_names = FALSE,
                               name = 'log(TPM+1)', 
                               col = colorRamp2(c(0,
                                                  maxcol/4,
                                                  maxcol/2,
                                                  3*maxcol/4,
                                                  maxcol),
                                                viridis(5)),
                               na_col = 'white', 
                               border = TRUE,
                               border_gp = gpar(lwd = 0.1),
                               column_names_gp = gpar(fontsize = 10),
                               row_dend_width = unit(3, "cm"),
                               width = unit(4.5, 'cm'),
                               height = unit(7,'cm'),
                               row_title = NULL,
                               left_annotation = if(
                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
                               ){
                                 clade_ha
                               }else{
                                 NULL
                               },
                               right_annotation = if(
                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
                               ){
                                 anc_ha
                               }else{
                                 NULL
                               },
                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
                               #                                        cir_info$Geneid)],
                               column_title = paste(list_name),
                               heatmap_legend_param = list(
                                 labels_gp = gpar(fontsize = 10),
                                 title_gp = gpar(fontsize = 10,
                                                 fontface = 'bold'),
                                 at = c(0,
                                        maxcol/4,
                                        maxcol/2,
                                        3*maxcol/4,
                                        maxcol)
                               )
                             )
                           }
         )
         ht_list <- Reduce(`+`, ht_list)
         clade_lgd <- Legend(
           title = 'sub-family',
           at = labels_SorL,
           legend_gp = gpar(fill = SorL_col)
         )

         if(global_save_var){        
           pdf('Manuscript_figure3c_heatmap.pdf'
               , width = 10
               , height = 4
           )
           draw(ht_list, 
                row_title_side = 'left',
                column_title_side = 'bottom',
                heatmap_legend_side = 'right',
                annotation_legend_side = 'right',
                annotation_legend_list = list(clade_lgd),
                ht_gap = unit(.35,'cm'),
                merge_legend = TRUE)
           dev.off()
         }

         return(
           draw(ht_list, 
                row_title_side = 'left',
                column_title_side = 'bottom',
                heatmap_legend_side = 'right',
                annotation_legend_side = 'right',
                annotation_legend_list = list(clade_lgd),
                ht_gap = unit(.35,'cm'),
                merge_legend = TRUE
           ))
       })

```

## 2D plot of PC1 v PC2 - Supp figure 4A

```{r pca_calculation, results = TRUE}
twentyfour_tpm <- twentyfour_rMT
rownames_geneid <- twentyfour_tpm$Geneid
twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
rownames(twentyfour_tpm_forpca) <- rownames_geneid
twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
#Need to look into what exactly scale and center are doing
```

```{r 2d_pca, results = TRUE}
twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)

twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,1])

colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
colornames <- setNames(colornames, 
                       sort(unique(twentyfour_tpm_pca_coords$time)))

pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)

(
  plot <- ggplot(twentyfour_tpm_pca_coords,
                 aes(
                   x = PC1,
                   y = PC2,
                   col = time
                 )) +
    geom_point(size = 3,
               alpha = 0.8) +
    scale_color_manual(values = colornames) +
    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
    guides(shape=guide_legend(ncol=2),
           col = guide_legend(ncol=2)) +
    theme_classic() +
    theme(legend.key.size = unit(4, 'mm'),
          legend.text = element_text(size = 7))
)

if(global_save_var){
  ggsave2(filename = 'Manuscript_SuppFigure3_chabaudi24h_pca_nopseudo.png',
          plot = plot,
          dpi = 300,
          width = 15,
          height = 10,
          units = 'cm')
}

```

# Bar chart of ancestral _pir_ expression.

```{r}
  cir_plot <- twentyfour_rMT %>% 
    filter(Geneid %in% 'PCHAS_0101200') %>% 
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h')) %>% 
    group_by(sample, time) %>% 
    dplyr::summarise(sum_tpm = sum(tpm))

y_scale <- seq(0, 
               signif(ceiling(max(cir_plot$sum_tpm)), 
                      digits = 2), 
               length.out = 5)

(
  p <- ggplot(data = cir_plot, 
              aes(x = time, 
                  y = sum_tpm)) +
    stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge',
                 fill = 'lightblue') +
    scale_y_continuous(breaks = y_scale) +
    geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
    ylab(as.expression(bquote('Total'~italic(.('PCHAS_0101200'))~'TPM'))) +
    xlab('Asexual blood stage development cycle') +
    theme_classic() +
    theme(axis.text.x=element_text(angle = 0, 
                                   vjust = 0.5,
                                   size = 13))
)

if(global_save_var){
  ggsave2(filename = 'Manuscript_ancestralPir_BarChart.pdf',
          plot = p,
          dpi = 300,
          width = 15,
          height = 10,
          units = 'cm')
}

```

# Calculate and cluster with z-scores {.tabset}

z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:

1. Convert the count/RPKM values of each gene into log values.
2. Calculate the mean and standard deviation of X gene log values in each sample
3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.

* Need to make sure that the z-score is calculated within each transmission method!

```{r zscores}
tpm_to_zscore = function(input){

  #If the input does not have a TPM element then assume that the input itself is the TPM object
  if(any(names(input) %in% 'tpm')){
    tpm_object <- input$tpm
  }else{
    tpm_object <- input
    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
  }

  #Want all of them so will just ignore Geneid
  input_logged <- tpm_object %>%
    mutate_at(vars(-matches('Geneid')), 
              log1p)

  gene_mean_scores = input_logged %>% 
    melt(id.vars = 'Geneid') %>% 
    group_by(Geneid) %>% 
    dplyr::summarise(mean = mean(value),
                     std_dev = sd(value))

  gene_vector = as.character(input_logged$Geneid)

  input_z = lapply(gene_vector, function(gene){
    mean_gene = filter(gene_mean_scores, 
                       Geneid == paste(gene)) %>% 
      select(mean)
    std_gene = filter(gene_mean_scores, 
                      Geneid == paste(gene)) %>% 
      select(std_dev)
    gene_scores = input_logged %>% 
      filter(Geneid == paste(gene)) %>% 
      select(-Geneid)
    if(std_gene != 0){
      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
    }else{
      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
    }
    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
  })

  #Use reduce and rbind to turn the list into one df
  input_z_df = purrr::reduce(input_z, 
                             rbind)
  return(list(zscores = input_z_df,
              tpm = tpm_object,
              experiment = input$experiment))
}

#Make and save z-scores
twentyfour_z <- lapply(twentyfour_avgtpm,
                       tpm_to_zscore)

###
```

# Phaseograms of all the genes.

```{r row_order_phaseograms}

#Getting the row order for each phaseogram for each experiment
# Want the data ordered by the highest zscore for each gene in each stage.
phaseogram_row_order <- lapply(twentyfour_z, 
                               function(list_elements){
                                 zscore_df <- list_elements$zscores
                                 zscore_ordered <- melt(zscore_df, 
                                                        id = 'Geneid',
                                                        variable.name = 'stage',
                                                        value.name = 'zscore') %>% 
                                   group_by(Geneid) %>% 
                                   top_n(zscore, n = 1) %>% 
                                   arrange(stage, 
                                           desc(zscore)) %>%
                                   distinct(Geneid, 
                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
                                   filter(zscore > 0) #For some reason some remain with zero zscore
                                 # zscore_ordered <- zscore_ordered[order(
                                 #   match(
                                 #     zscore_ordered$stage, 
                                 #     order_vector
                                 #   )
                                 # ),]
                                 return(c(list_elements, 
                                          list(zscore_order = zscore_ordered)))
                               })

phaseogram_time_order <- lapply(names(phaseogram_row_order), 
                                function(list_element){
                                  transmission_list <- getElement(phaseogram_row_order,
                                                                  list_element)
                                  order_by_time <- transmission_list$zscore_order %>% 
                                    mutate(stage = times_stages_table$time[match(stage,
                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
                                    arrange(stage)
                                  return(list(zscore_order = order_by_time))
                                }
)
names(phaseogram_time_order) <- names(phaseogram_row_order)

```

```{r function_draw_heatmap_zscore}

draw_heatmap_zscore <- function(zscore_list,
                                title = '',
                                gene_ordering_vector = NULL,
                                filter_genes_vector = NULL,
                                extra_title = NULL,
                                annotation_clade = NULL,
                                annotation_col,
                                threshold_tpm = NULL,
                                use_times = FALSE,
                                use_order_vector = TRUE,
                                tpm_mat = TRUE,
                                tpm_breaks = NULL,
                                ...){
  if(is.null(tpm_breaks) & tpm_mat){
    #Need the generate the breaks in the data for the heatmap legend if not provided.
    max_tpm <- lapply(zscore_list, 
                      function(list_element){
                        #list_element <- zscore_list[[1]]
                        heatmap_data <- as.data.frame(list_element$zscores)
                        title <- list_element$experiment

                        if(!'Geneid' %in% colnames(heatmap_data)){
                          stop('No Geneid column in the heatmap data.frame')
                        }

                        if(!is.null(filter_genes_vector)){
                          heatmap_data <- as.data.frame(
                            filter(heatmap_data, 
                                   Geneid %in% filter_genes_vector)
                          )
                        }

                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                        if(!is.null(gene_ordering_vector)){
                          #Remove genes not in the main data frame but in the ordering vector, 
                          # and then order based on the vector
                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          heatmap_data <- heatmap_data[na.omit(
                            match(gene_ordering_vector, 
                                  rownames(heatmap_data))
                          ),,
                          drop=FALSE]
                        }

                        geneids <- rownames(heatmap_data)

                        #Get max/min TPM scores for the side annotation heatmap                       
                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                          list_element$tpm$Geneid),] %>% 
                          select(-Geneid) %>% 
                          as.matrix %>% 
                          log1p
                        rownames(tpm_mat) <- geneids

                        tpm_min <- rowMins(tpm_mat)
                        names(tpm_min) <- geneids
                        tpm_max <- rowMaxs(tpm_mat)
                        names(tpm_max) <- geneids
                        return(list(tpm_max = tpm_max,
                                    tpm_min = tpm_min))
                      })

    tpm_max_all <- mround(
      max(
        unlist(
          lapply(max_tpm, 
                 getElement,
                 'tpm_max')
        )
      ), 
      base = 1)

    tpm_breaks <- round(
      c(0,
        tpm_max_all/4,
        tpm_max_all/2,
        tpm_max_all*(3/4),
        tpm_max_all),
      digits = 2
    )
  }

  #Generate the colour function from the TPM legend breaks for the TPM annotation
  if(tpm_mat){
    tpm_col_fun <- colorRamp2(
      breaks = tpm_breaks,
      colors = viridis(length(tpm_breaks))
    )
  }

  if(is.numeric(threshold_tpm)){
    #Apply a threshold to the TPM in order to remove low expressed genes.
    # Remember to add log1p here! 
    # Threshold uses the max value, so it ony needs to be above thresh 
    #   in one stage avg in one transmission method.
    #Will alter the gene_filter vector in order to only select the genes above threshold

    gene_maxs <- bind_rows(
      lapply(max_tpm, 
             getElement,
             'tpm_max')
    )

    remove_genes <- names(gene_maxs)[apply(
      gene_maxs, 
      2, 
      function(maxs){
        all(maxs < log2(1+threshold_tpm))
      }
    )]

    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]

  }

  if(is.character(annotation_clade)){
    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
    clade_info <- getElement(cir_info,
                             annotation_clade)[match(order_cir, 
                                                     cir_info$Geneid)]
    names(clade_info) <- str_split_fixed(order_cir, 
                                         pattern = '_', 
                                         n = 2)[,2]
    annotation_col <- annotation_col
  }

  heatmaps <- lapply(zscore_list, 
                     function(list_element){
                       #list_element <- zscore_list[[1]]
                       heatmap_data <- as.data.frame(list_element$zscores)
                       title <- list_element$experiment

                       if(!'Geneid' %in% colnames(heatmap_data)){
                         stop('No Geneid column in the heatmap data.frame')
                       }

                       if(!is.null(filter_genes_vector)){
                         heatmap_data <- as.data.frame(
                           filter(heatmap_data, 
                                  Geneid %in% filter_genes_vector)
                         )
                       }

                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                       if(!is.null(gene_ordering_vector)){
                         #Remove genes not in the main data frame but in the ordering vector,
                         #  and then order based on the vector
                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                         heatmap_data <- heatmap_data[na.omit(
                           match(gene_ordering_vector, 
                                 rownames(heatmap_data))
                         ),,
                         drop=FALSE]
                       }

                       if(tpm_mat){
                         #Get max/min TPM scores for the side annotation heatmap                       
                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                           list_element$tpm$Geneid),] %>%
                           select(-Geneid) %>% 
                           as.matrix %>% 
                           log1p

                         tpm_min <- rowMins(tpm_mat)
                         tpm_max <- rowMaxs(tpm_mat)

                         names(tpm_min) = rownames(heatmap_data)
                         names(tpm_max) = rownames(heatmap_data)

                         tpm_ha <- rowAnnotation(
                           min.max = anno_simple(cbind(tpm_min, 
                                                       tpm_max), 
                                                 col = tpm_col_fun,
                                                 gp = gpar(col = 'black',
                                                           lwd = 0.1),
                                                 border = TRUE,
                                                 simple_anno_size = unit(0.2, "cm")),
                           show_annotation_name = FALSE)
                       }

                       #Draw the left annotation is needed, only for the first element.

                       if(
                         identical(list_element,
                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
                       ){
                         clade_ha <- rowAnnotation(
                           'subfamily' = anno_simple(
                             clade_info,
                             col = annotation_col
                           )
                         )
                       }

                       #Change rownames to remove PCHAS_
                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
                                                                pattern = '_', 
                                                                n = 2)[,2]

                       if(use_times){
                         colnames(heatmap_data) <- str_extract(colnames(heatmap_data), 
                                                               pattern = '[[:digit:]]{2}h')
                       }

                       if(use_order_vector){
                         #Order the columns by ordering_vector
                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
                                                                   order_vector)),
                                                      drop = FALSE]
                       }else{
                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
                                                      drop = FALSE] 
                       }

                       h = Heatmap(heatmap_data,
                                   name = paste0('z-score',
                                                 extra_title),
                                   col = c(cbPalette[6],
                                           'white', 
                                           cbPalette[7]),
                                   row_names_gp = gpar(fontsize = ifelse(
                                     nrow(heatmap_data) > 13, 
                                     12*(13/nrow(heatmap_data)), 
                                     12)
                                   ),
                                   column_names_gp = gpar(fontsize = 10),
                                   row_dend_width = unit(3, "cm"),
                                   right_annotation = if(tpm_mat){
                                     tpm_ha
                                   }else{
                                     NULL
                                   },
                                   left_annotation = 
                                     if(identical(list_element,
                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
                                       clade_ha
                                     }else{
                                       NULL
                                     }, 
                                   #Only want the clade annotation when it is first heatmap
                                   width = unit(4.5, 'cm'),
                                   cluster_columns = FALSE,
                                   height = unit(7,'cm'),
                                   row_title = NULL,
                                   cluster_rows = FALSE,
                                   column_title = paste(title),
                                   heatmap_legend_param = list(
                                     at = c(3,2,1,0,-1,-2,-3),
                                     labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                   ),
                                   ...
                       )
                       return(h)
                     })

  ht_list <- Reduce('+', heatmaps) #To add the list elements together

  if(tpm_mat){
    #Draw the legend
    tpm_lgd <- Legend(
      title = "Min/Max log(TPM + 1)", 
      col = tpm_col_fun, 
      at = tpm_breaks, 
      labels = tpm_breaks,
      title_gp = gpar(fontsize = 7)
    )
  }

  if(is.character(annotation_clade)){
    clade_lgd <- Legend(
      title = 'sub-family',
      at = names(annotation_col),
      legend_gp = gpar(fill = annotation_col)
    )
  }

  draw(ht_list, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       heatmap_legend_side = 'left',
       #Only include the annotation legends for the annotations we actually have.
       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
         list(tpm_lgd,
              clade_lgd)
       }else if(tpm_mat & !is.character(annotation_clade) ){
         list(tpm_lgd)
       } else if(!tpm_mat & is.character(annotation_clade) ){
         list(clade_lgd)
       })
}
```

## Supplementary figure 5 - Phaseograms of the entire genome.

```{r phaseograms_whole_genome}
if(global_save_var){
  pdf(file = 'Manuscript_SuppFig5_phaseogram_allgenes.pdf',
      height = 5,
      width = 5)
  draw_heatmap_zscore(zscore_list = twentyfour_z,
                      gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                      use_times = TRUE,
                      use_order_vector = FALSE,
                      show_row_names = FALSE,
                      tpm_mat = FALSE)
  dev.off()
}

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

```

##  Supplementary figure 6 - Phaseograms of the _cir_ genes with clade annotations

```{r pir_gene_phaseogram_breaks}
  max_tpm <- lapply(twentyfour_z, 
                    function(list_element){
                      #list_element <- zscore_list[[1]]
                      heatmap_data <- as.data.frame(list_element$zscores)
                      title <- list_element$experiment

                      if(!'Geneid' %in% colnames(heatmap_data)){
                        stop('No Geneid column in the heatmap data.frame')
                      }

                            heatmap_data <- as.data.frame(
                              filter(heatmap_data, 
                                     Geneid %in% cir_id)
                            )

                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
                          # 
                          # if(!is.null(gene_ordering_vector)){
                          #   #Remove genes not in the main data frame but in the ordering vector, 
                          #   # and then order based on the vector
                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          #   heatmap_data <- heatmap_data[na.omit(
                          #     match(gene_ordering_vector, 
                          #           rownames(heatmap_data))
                          #   ),,
                          #   drop=FALSE]
                          # }

                          geneids <- rownames(heatmap_data)

                          #Get max/min TPM scores for the side annotation heatmap                       
                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                            list_element$tpm$Geneid),] %>% 
                            select(-Geneid) %>% 
                            as.matrix %>% 
                            log1p
                          rownames(tpm_mat) <- geneids

                          tpm_min <- rowMins(tpm_mat)
                          names(tpm_min) <- geneids
                          tpm_max <- rowMaxs(tpm_mat)
                          names(tpm_max) <- geneids
                          return(list(tpm_max = tpm_max,
                                      tpm_min = tpm_min))
                        })

  tpm_max_all <- mround(
    max(
      unlist(
        lapply(max_tpm, 
               getElement,
               'tpm_max')
      )
    ), 
    base = 1)

  tpm_breaks <- round(
    c(0,
      tpm_max_all/4,
      tpm_max_all/2,
      tpm_max_all*(3/4),
      tpm_max_all),
    digits = 2
  )
```

```{r phaseograms_cir}

if(global_save_var){
 pdf(file = 'Manuscript_SuppFig6_phaseogram_cirgenes.pdf',
      height = 5,
      width = 5)
  draw_heatmap_zscore(zscore_list = twentyfour_z,
                      gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                      filter_genes_vector = cir_id,
                      annotation_clade = 'SorL',
                      annotation_col = SorL_col,
                      threshold_tpm = 1,
                      use_times = TRUE,
                      use_order_vector = FALSE,
                      show_row_names = FALSE,
                      tpm_mat = FALSE)
  dev.off()
}

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

```

## Figure 4 and supplementary figure 7

_pir_ genes selected by being expressed above 1 TPM in all samples.

Adding in housekeeping genes for comparison:
HSP70 putative  PCHAS_0721000
MSP1 PCHAS_0831300
AMA1 PCHAS_0931000

```{r pirs_side-by-side_figure4}

housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
                                 'MSP1' = 'PCHAS_0831300',
                                 'AMA1' = 'PCHAS_0931000')

cir_housekeeping_id <- c(cir_id, 
                         'PCHAS_0721000',
                         'PCHAS_0831300',
                         'PCHAS_0931000')

cir_plot <- twentyfour_rMT %>% 
  filter(Geneid %in% cir_housekeeping_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown'))) %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a')) %>% 
  mutate(subfam =
           forcats::fct_shift(
             relevel(
               as.factor(subfam),
               ref = 'ref')
           )
  )

#Need to add the transmission name to the subfam so that they are ordered correctly.
# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.

cir_plot$Geneid <- ifelse(
  cir_plot$Geneid %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
  cir_plot$Geneid
)

limits_cir <- c(0, 
                mround(max(cir_plot$tpm), 
                       base = 100))

cir_plot_order <- cir_plot %>% 
  filter(time == '14h') %>% 
  arrange(desc(tpm))

cir_plot <- mutate(cir_plot, 
                   Geneid = factor(Geneid, 
                                   levels = unique(cir_plot_order$Geneid)))

transmissionSep_cir_plot <- function(time_vec = '14h',
                                     log_it = TRUE,
                                     limits = limits_cir,
                                     point_size = 1,
                                     text_size = 10,
                                     threshold_min = 100,
                                     threshold_max = Inf){

  cir_plot_filter <- group_by(cir_plot, 
                              Geneid) %>% 
    filter(
      time == time_vec,
      !Geneid %in% cir_id | any(tpm >= threshold_min),
      !Geneid %in% cir_id | all(tpm <= threshold_max)
    )
  #Filter for pirs expressed above or below the thresholds, have the OR so that the ref genes aren't removed.
  # e.g. Either 'Not a cir' OR 'Any TPM is above the minimum'.

  # cir_plot_filter %>% group_by(Geneid) %>% summarise(n())
  if(log_it){
    cir_plot_filter$tpm <- log2(1+cir_plot_filter$tpm)
    limits = log2(1+limits)
    breaks <- round(seq.int(0, 
                            max(limits), 
                            length.out = round(max(limits)) + 1, 
                            digits = 0), 
                    digits = 0)
  }else{
    breaks <- seq.int(0, max(limits), length.out = 11)
  }

  cir_plot_filter$Geneid <- factor(
    str_remove(cir_plot_filter$Geneid, 
               pattern = 'PCHAS_'),
    levels = str_remove(levels(cir_plot_filter$Geneid), 
                        pattern = 'PCHAS_')
  )

  y <- ggplot(
    data = cir_plot_filter,
    mapping = aes(
      x = Geneid,
      y = tpm
    )
  ) +
    stat_summary(fun.y = median, 
                 fun.ymin = median, 
                 fun.ymax = median,
                 geom = "bar",
                 fill = 'lightblue',
                 width = 0.75, 
                 size = 0.25, 
                 position = 'dodge') +
    geom_point(stat = 'identity', 
               position = position_dodge(width = 0.75),
               size = point_size) +
    scale_y_continuous(breaks = breaks,
                       limits = limits) +
    facet_grid(.~subfam, 
               scales = "free_x",
               space = "free_x",
               labeller = labeller(subfam = function(string) {
                 str_remove_all(string, 
                                pattern = '(RMT|SBP)_')
               })) +
    theme_classic() +
    ylab('log(TPM + 1)')+
    ggtitle(paste(unique(cir_plot_filter$time)))

  if(grepl(time_vec, pattern = '20|23')) {
    y <- y + theme(
      axis.text.x = element_text(size = text_size,
                                 angle = 90,
                                 vjust = 0.5)
    )
  } else {
    y <- y + theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    ) 
  }

           #This bit is to change the colour, and text colour, of the facet boxes

           g <- ggplot_gtable(ggplot_build(y))

           # g$widths[5] = 2*g$widths[5]

           strip_t <- which(grepl('strip-t', g$layout$name))
           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(RMT|SBP)_')
           fills <- case_when(str_detect(subfam_vec, 
                                         pattern = 'a') ~ clade_col['ancestral'],
                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                              str_detect(subfam_vec, 
                                         pattern = 'ref') ~ 'white')
           k <- 1
           for (i in strip_t) {
             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
             k <- k+1
           }

           #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
           # manually make them a bit wider.

           short_panels <- which(str_detect(subfam_vec, pattern = 'a|S1'))

           for(i in short_panels){
             num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
             g$widths[num] = 1.75*g$widths[num]
             #1.75 times the width of the right panels.
             # print(num)
           }

           if(global_save_var){
             ggsave2(
               paste0('Pchabaudi_side-by-side_barchart_facetbySubfam_',
                      time_vec,
                      '.png'),
               plot = g,
               dpi = 300,
               width = 28,
               height = 8,
               units = 'cm'
             )
           }

           grid.newpage()
           grid.draw(g)

}

lapply(
  sort(unique(cir_plot$time)),
  transmissionSep_cir_plot,
  threshold_min = 50,
  threshold_max = Inf,
  point_size = 1,
  text_size = 7.5
)
```

# Supp figure 9 - Pir expression of AAPLs and ChAPLs.

```{r Supp_fig_AapL-ChaAPl_expression}

cir_plot <- twentyfour_rMT %>% 
  filter(Geneid %in% cir_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
    chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
                                                       cir_info$Geneid)],
                        levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
                                                                            cir_info$Geneid)])))) %>% 
  filter(!(time %in% c('sch', 
                       'Unknown')),
         !is.na(chapl_aapl) ) %>% 
  group_by(Geneid) %>% 
    arrange(desc(tpm))

  cir_plot_order <- cir_plot %>% 
    filter(time == '14h') %>% 
    arrange(desc(tpm))

  cir_plot <- mutate(cir_plot, 
                     Geneid = factor(Geneid, 
                                     levels = unique(cir_plot_order$Geneid)))
  cir_plot_chapl_aapl_supp_info <- cir_plot

  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))

  breaks = seq(0, max_aapl_chapl, by = 1)
    limits = c(0, max_aapl_chapl)
    point_size = 0.25

  lapply(
    sort(unique(cir_plot$time)),
    function(time_vec){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

      cir_plot$tpm <- log2(1+cir_plot$tpm)

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge',
                   fill = 'lightblue') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      ylab('log(TPM +1)') +
      facet_grid(.~chapl_aapl, 
                 scales = "free_x",
                 space = "free_x") +
      theme_classic() +
      theme(
        axis.text.x = element_text(angle = 90,
                                   size = 5)
      ) +
      ggtitle(paste(unique(cir_plot$time)))

    if(global_save_var){
      ggsave2(
        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
               time_vec,
               '.png'),
        plot = y,
        dpi = 300,
        width = 25,
        height = 10,
        units = 'cm'
      )
    }

    return(y)
    })
```

### 2021-09-06 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_CalculateTPM.Rmd
Add data from Manuscript (berghei, chabaudi 24h, ancestral phylogeny and expression)

---
title: "P.berghei pir calculate TPM"
author: "Timothy Little"
date: "31/08/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r load_packages, include = FALSE, message = FALSE}
library('readxl')
library('readr')
library('dplyr')
library('purrr')
library('ggplot2')
library('DESeq2')
library("RColorBrewer")
library('colorspace')
library('data.table')
library('gridExtra')
library('tibble')
library('scales')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('plotly')
library('ggpattern')
library('ComplexHeatmap')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, select, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r ordering_vectors}
order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
                 'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')

r_order_vector <- str_replace_all(order_vector, 
                                  pattern = '-', 
                                  replacement = '\\.')
```

```{r pir_id, message=FALSE, results='hide'}

# Getting the pir gene names and transcript lengths needed for calculating the transcript-per-million

# First we need to get the _pir gene names. These have been downloaded using a search of PlasmoDB P. berghei ANKA genome v3 using  "\*pir\*" search and manual removal of genes which clearly were not pir.

allgene.info.file <- '../standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv'
allgene.info <- as_tibble(
  read_csv(allgene.info.file)
) %>%
  mutate(Geneid = `Gene ID`, 
         Description = `Product Description`,
         is_pseudo = `Is Pseudo`,
         Transcript.Length = `Transcript Length`,
         paralog_count = `Paralog count`) %>% 
  select(Geneid, 
         Description, 
         is_pseudo, 
         Transcript.Length, 
         paralog_count)

pir_info <- allgene.info %>% 
  filter(str_detect(Description, 
                    pattern = 'PIR(?!.+pseud)'), 
         is_pseudo == 'No') %>% 
  filter(!Geneid %in% 'PBANKA_0524600')
#Removing the pir-like gene.

pir_id <- pir_info$Geneid

#pir

pir_genes_list <- lapply(
  unique(pir_info$Description), 
  function(pir_group){
    select(
      filter(
        pir_info, 
        Description %in% pir_group
      ),
      Geneid
    )
  }
)
names(pir_genes_list) <- unique(pir_info$Description)

#How many pirs are present on each chromosome?

pir_tbl = as_tibble(data.frame(Geneid = pir_id))
pir_chrom_tbl = pir_tbl %>% 
  mutate(chrom = unlist( #Getting the chromosomes by a complicated set of lapply and sapplys...
    sapply(
      lapply(
        sapply(
          as.character(pir_tbl$Geneid), 
          strsplit,
          split = ''
        ), 
        '[', 
        8:9
      ), 
      paste, 
      collapse = ''
    )
  )
  ) 
pir_chrom_tbl %>% 
  group_by(chrom) %>% 
  dplyr::summarise(number_of_pirs_in_each_chromosome = n(),
                   percentage_of_all_pirs = round(n()/nrow(pir_tbl), 3)*100) %>% 
  dplyr::arrange(desc(percentage_of_all_pirs))

#Getting the loci information
pir_loci <- data.frame(
  Geneid = pir_id,
  chrom = str_extract(pir_id, 
                      pattern = '(?<=_)[[:digit:]]{2}'),
  number = as.numeric(str_extract(pir_id, 
                                  pattern = '(?<=_[[:digit:]]{2})[[:digit:]]+'))
) %>% 
  filter(!chrom %in% '00') %>% #Remove chrom00 (contigs)
  group_by(chrom) %>% 
  mutate(jump = ifelse(lag(number, 
                           default = 1001) < 1000 & number - lag(number, 
                                                                 default = 0) > 1000,
                       1,
                       0)) %>% 
  mutate(
    jump = ifelse(cumsum(jump) == 1 | jump == 1, 1, 0)
  ) %>% 
  mutate(
    locus = paste0(
      chrom,
      ifelse(jump == 0 & number < 1000, 
             'Left',
             'Right')
    )
  )

# Detect 'jumps' of the gene number over 1000 within each chromosome, setting the lag default as 1001 in the
#   first part of the ifelse statement so that chrom2 doesn't have an erroneous jump as it starts at >1000
# Later cumsum is do that every gene after the 'jump' is included allowing us to define telomere ends.

pir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
                        delim = '\t')
pir_clade <- mutate(pir_clade, 
                    locus = pir_loci$locus[match(Geneid, pir_loci$Geneid)])
pir_id <- pir_clade$Geneid

pir_clade_list <- lapply(unique(pir_clade$clade), function(clade_name){
  pir_clade$Geneid[pir_clade$clade %in% clade_name]
})
names(pir_clade_list) <- as.character(unique(pir_clade$clade))
```

```{r transcript_redundancy, message = FALSE, results='hide'}

#Note that there is some redundancy
two.transcript <- allgene.info %>% 
  dplyr::count(Geneid) %>% 
  filter(n > 1)

#Nine genes have two transcripts - need to give them one average of the transcripts since featurecounts has only
# calculated one count for the locus
two.transcript.info <- allgene.info[allgene.info$Geneid %in% two.transcript$Geneid,]
two.transcript.mean = aggregate(Transcript.Length~Geneid, data=two.transcript.info, 
                                FUN=function(x) c(mean=mean(x)))

#Remove the first match
allgene.info <- allgene.info[-(match(two.transcript.mean$Geneid, 
                                     allgene.info$Geneid)),]

#For the second match replace the transcript value with the mean calculated with aggregate
allgene.info$Transcript.Length <- as.double(allgene.info$Transcript.Length)
allgene.info[match(two.transcript.mean$Geneid, 
                   allgene.info$Geneid),]$Transcript.Length <- two.transcript.mean$Transcript.Length
lengths.transcript <- subset(allgene.info, select = c(Geneid, Transcript.Length), drop = F)
```

## Read in counts and calculate TPM

```{r read_files_calc_TPM}
#Reading in the counts from file
berghei_counts <- read.csv('tl20-05_berghei_counts.csv',
                               row.names = 1)

berghei_tpm <- tpm(count_data = berghei_counts,
                   lengths = lengths.transcript)

apply(berghei_tpm[-1], 2, sum)
```

```{r check_tpm}

#Used `dput` function to create a df of randomly chosen part of what I assume to be the true values.
check_tpm <- structure(list(E03_Gam_1 = c(578.506468410824, 43.8844462052653, 
                                          93.7895540428778, 55.9251999927489, 28.3887997796848, 49.7689021855188
), 
E03_Gam_2 = c(491.460414371885, 38.6239066701, 119.30578922894, 
              148.452713617397, 38.8846967267782, 25.7641457887425), 
E03_Gam_3 = c(824.81294645121, 
              48.1865616211443, 92.5270154766448, 74.1594147167308, 34.8902736305123, 
              74.9586138010564), 
E03_Ook_1 = c(340.238340611313, 206.852529849517, 
              43.9667720316821, 26.3982623497325, 14.9136412690561, 6.22485349746656
), 
E03_Ook_2 = c(713.689689765822, 178.051382530752, 58.5009548960498, 
              52.6871822865333, 18.8623926433347, 12.0840592930611), 
E03_Ook_3 = c(409.621633449974, 
              104.471439607932, 38.7705779604298, 31.3685913631761, 20.1312670544009, 
              14.7857267550803)), 
row.names = 3345:3350, 
class = "data.frame")

all.equal(berghei_tpm[3345:3350,23:28],check_tpm)

```

```{r save_tpm_oldNames_allSamples}
#Save the 'raw' TPM before any samples have been removed or sample names combined.

write.csv(berghei_tpm,
          'berghei_tpm_oldNamesAllSamples.csv')

```

```{r remove_E06-6h}
#Remove the Caldelari (E06) 6h samples, as they did in their paper.
berghei_tpm[,str_subset(colnames(berghei_tpm), 'E06_Liv.6h')] <- NULL
```

```{r hclusts, message = FALSE}
experiment_id <- unique(na.omit(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}')))

lapply(experiment_id, 
       function(experiment_name){
         tpm_data <- select(berghei_tpm, matches(experiment_name))
         if(ncol(tpm_data) > 3){
           dists = dist(t(tpm_data))
           par(cex=0.7)
           plot(hclust(dists),
                main = experiment_name)
         }
       })

```

Going to combine certain samples based on their similar profiles, e.g. <4h, >48h liver stages.

```{r combine_similar_samples_byNames}
col_stage <- colnames(berghei_tpm)

#Combine all Liver samples 4h or before into LiverPre4h
col_stage <- str_replace_all(col_stage,
                             pattern = 'Liv\\.(2|4)h_[[:digit:]]+',
                             replacement = 'Liv.pre4h')
#Combine all Liver samples 48h or onwards into LiverPost48h
col_stage <- str_replace_all(col_stage,
                             pattern = 'Liv\\.(4[8-9]|[5-9][0-9])h_[[:digit:]]+',
                             replacement = 'Liv.post48h')
#Combine all blood meal samples 3h or before into BldMlPre3h
col_stage <- str_replace_all(col_stage,
                             pattern = '(1h|3h)_[[:digit:]]+', 
                             replacement = 'Pre3h')
#Uniquify the name names
col_stage[grep(col_stage, 
               perl = TRUE,
               invert = TRUE,
               pattern = '_[[:digit:]]{1,}$')] <- uniquify(col_stage[grep(col_stage, 
                                                                          perl = TRUE,
                                                                          invert = TRUE,
                                                                          pattern = '_[[:digit:]]{1,}$')]) 
```

```{r remove_badly_clustering_samples}
berghei_tpm$`E08_Liv.pre4h_1` <- NULL
berghei_tpm$`E03_Asex.Schiz_1` <- NULL
berghei_tpm$`E03_Asex.Schiz_3` <- NULL
```

## Saving the TPM

```{r save_tpm}
write.csv(berghei_tpm,
          'berghei_tpm.csv')
```

### 2021-09-06 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
Add data from Manuscript (berghei, chabaudi 24h, ancestral phylogeny and expression)

---
title: "P.berghei pir analysis"
author: "Timothy Little"
date: "31/08/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r load_packages, include = FALSE, message = FALSE}
library('readxl')
library('readr')
library('dplyr')
library('purrr')
library('ggplot2')
library('DESeq2')
library("RColorBrewer")
library('colorspace')
library('data.table')
library('gridExtra')
library('tibble')
library('scales')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('plotly')
library('ggpattern')
library('ComplexHeatmap')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, select, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r ordering_vectors}
order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
                 'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')

r_order_vector <- str_replace_all(order_vector, 
                                  pattern = '-', 
                                  replacement = '\\.')
```

```{r pir_id, message=FALSE, results='hide'}

# Getting the pir gene names and transcript lengths needed for calculating the transcript-per-million

# First we need to get the _pir gene names. These have been downloaded using a search of PlasmoDB P. berghei ANKA genome v3 using  "\*pir\*" search and manual removal of genes which clearly were not pir.

allgene.info.file <- '../standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv'
allgene.info <- as_tibble(
  read_csv(allgene.info.file)
) %>%
  mutate(Geneid = `Gene ID`, 
         Description = `Product Description`,
         is_pseudo = `Is Pseudo`,
         transcript_length = `Transcript Length`,
         paralog_count = `Paralog count`) %>% 
  select(Geneid, 
         Description, 
         is_pseudo, 
         transcript_length, 
         paralog_count)

pir_info <- allgene.info %>% 
  filter(str_detect(Description, 
                    pattern = 'PIR(?!.+pseud)'), 
         is_pseudo == 'No') %>% 
  filter(!Geneid %in% 'PBANKA_0524600')
#Removing the pir-like gene.

pir_id <- pir_info$Geneid

#pir

pir_genes_list <- lapply(
  unique(pir_info$Description), 
  function(pir_group){
    select(
      filter(
        pir_info, 
        Description %in% pir_group
      ),
      Geneid
    )
  }
)
names(pir_genes_list) <- unique(pir_info$Description)

#How many pirs are present on each chromosome?

pir_tbl = as_tibble(data.frame(Geneid = pir_id))
pir_chrom_tbl = pir_tbl %>% 
  mutate(chrom = unlist( #Getting the chromosomes by a complicated set of lapply and sapplys...
    sapply(
      lapply(
        sapply(
          as.character(pir_tbl$Geneid), 
          strsplit,
          split = ''
        ), 
        '[', 
        8:9
      ), 
      paste, 
      collapse = ''
    )
  )
  ) 
pir_chrom_tbl %>% 
  group_by(chrom) %>% 
  dplyr::summarise(number_of_pirs_in_each_chromosome = n(),
                   percentage_of_all_pirs = round(n()/nrow(pir_tbl), 3)*100) %>% 
  dplyr::arrange(desc(percentage_of_all_pirs))

#Getting the loci information
pir_loci <- data.frame(
  Geneid = pir_id,
  chrom = str_extract(pir_id, 
                      pattern = '(?<=_)[[:digit:]]{2}'),
  number = as.numeric(str_extract(pir_id, 
                                  pattern = '(?<=_[[:digit:]]{2})[[:digit:]]+'))
) %>% 
  filter(!chrom %in% '00') %>% #Remove chrom00 (contigs)
  group_by(chrom) %>% 
  mutate(jump = ifelse(lag(number, 
                           default = 1001) < 1000 & number - lag(number, 
                                                                 default = 0) > 1000,
                       1,
                       0)) %>% 
  mutate(
    jump = ifelse(cumsum(jump) == 1 | jump == 1, 1, 0)
  ) %>% 
  mutate(
    locus = paste0(
      chrom,
      ifelse(jump == 0 & number < 1000, 
             'Left',
             'Right')
    )
  )

# Detect 'jumps' of the gene number over 1000 within each chromosome, setting the lag default as 1001 in the
#   first part of the ifelse statement so that chrom2 doesn't have an erroneous jump as it starts at >1000
# Later cumsum is do that every gene after the 'jump' is included allowing us to define telomere ends.

pir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
                        delim = '\t')
pir_clade <- mutate(pir_clade, 
                    locus = pir_loci$locus[match(Geneid, pir_loci$Geneid)])
pir_id <- pir_clade$Geneid

pir_clade_list <- lapply(unique(pir_clade$clade), function(clade_name){
  pir_clade$Geneid[pir_clade$clade %in% clade_name]
})
names(pir_clade_list) <- as.character(unique(pir_clade$clade))
```

## Saving the TPM

```{r save_tpm}
berghei_tpm <- read.csv('berghei_tpm.csv', 
                        row.names = 1)
```

```{r expt_tpms}
#Get a list of df for each experiment.

expt_nams <- unique(
  na.omit(
    str_extract(colnames(berghei_tpm), 
                pattern = 'E[[:digit:]]{2}')
  )
)

expt_tpms <- lapply(
  expt_nams,
  function(expt_nam){
    expt_df <- select(
      berghei_tpm,
      matches(paste0('Geneid|', expt_nam))
    ) %>% 
      rename_with(~ str_remove(.x, 'E[[:digit:]]{2}_'),
                  matches('E[[:digit:]]{2}'))
    list(tpm = expt_df,
         experiment = expt_nam)
  }
)
names(expt_tpms) <- expt_nams
```

```{r melt_df}
berghei_tpm_melt <- berghei_tpm %>% 
    #Melt the data frame for using dplyr
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    mutate(experiment = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1)),
           stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,2)))

berghei_tpm_avg_byexpt <- berghei_tpm_melt %>% 
  group_by(Geneid, experiment, stage) %>%  
  dplyr::summarise(median_tpm = median(tpm),
                   mean_tpm = mean(tpm), 
                   min_tpm = min(tpm), 
                   max_tpm = max(tpm), 
                   sd_tpm = sd(tpm)) %>% 
  mutate(expt_stage = paste0(experiment, '_', stage))
```

## The highest expressed _pir_ genes across the samples

```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
top_n_genes <- function(tpm_avg, 
                        n = 5, 
                        threshold = 0){
  #Calculating the index used to make the kable table

  kable_index = rep(n, times = length(unique(tpm_avg$stage)))
  names(kable_index) = unique(tpm_avg$stage)

  tpm_avg <- tpm_avg %>% 
    ungroup %>% 
    group_by(stage)

  all_genes = tpm_avg %>% 
    slice_max(order_by = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue")))
  all_genes$short_descr = allgene.info$handy_descriptions[match(all_genes$Geneid, 
                                                                allgene.info$Geneid)]
  all_genes$description = allgene.info$Description[match(all_genes$Geneid, 
                                                         allgene.info$Geneid)]
  all_genes = all_genes %>% 
    kable(format = 'html', 
          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed total genes'), 
          escape = FALSE) %>%
    kable_styling() %>% 
    pack_rows(index = kable_index)

  pir_genes = tpm_avg %>% 
    filter(Geneid %in% pir_id) %>% 
    slice_max(order_by = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue"))) %>% 
    mutate(Geneid = as.character(Geneid)) %>% 
    mutate(Description = allgene.info$handy_descriptions[match(Geneid, 
                                                               allgene.info$Geneid)]) %>% 
    mutate(Geneid = ifelse(grepl(Geneid,
                                 pattern = 'PBANKA_0100500'),
                           cell_spec(Geneid, bold = TRUE),
                           Geneid)) %>%
    kable(format = 'html', 
          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed pir genes'), 
          escape = FALSE) %>% 
    pack_rows(index = kable_index) %>% 
    kable_styling()

  # print(paste0('Completed analysis of ', tpm_input$experiment))
  list = list(all_genes, pir_genes, unique(tpm_avg$experiment))
  names(list) = c('all_genes', 'pir_genes', 'experiment')
  return(list)
}

### Generate and save top genes - core dataset all_top

all_top <- lapply(unique(berghei_tpm_avg_byexpt$experiment), 
                  function(expt) top_n_genes(berghei_tpm_avg_byexpt[berghei_tpm_avg_byexpt$experiment == expt,]))

all_top[[1]]

```

##Supp figure 1A - Pearson correlation plot of collated samples.

```{r all_samples_collated_heatmap}

## Collated across the stages for a more easily understood heatmap
big_top <- berghei_tpm_avg_byexpt

big_top_recast = dcast(big_top, Geneid ~ expt_stage, value.var = 'mean_tpm')
big_top_recast <- column_to_rownames(big_top_recast, var = 'Geneid')
# big_top_recast = big_top_recast[,-1,drop=FALSE]

# apply(big_top_recast, 2, sum)

pdf(file = 'Manuscript_suppfig1A_corrplot_all_experiments_rnaseq.pdf',
    height = 5,
    width = 5)
corrplot(cor(big_top_recast), method = 'color', 
         order = 'hclust', hclust.method = 'ward.D2',
         cl.lim = c(-0.036,1), tl.cex = 0.50,
         col = c(viridis(100), magma(115)), addrect = 6)
dev.off()

corrplot(cor(big_top_recast), method = 'color', 
         order = 'hclust', hclust.method = 'ward.D2',
         cl.lim = c(-0.036,1), tl.cex = 0.50,
         col = c(viridis(100), magma(115)), addrect = 6)
```

## Supp figure 1B - PCA of all the samples.

```{r supp_figure_PCA_allgenes}

big_tpm_pca <-  prcomp(berghei_tpm[,-1], 
                       center = TRUE, 
                       scale. = TRUE)

big_tpm_pca_var <- get_pca_var(big_tpm_pca)
big_tpm_pca_coords <- data.frame(sample = rownames(big_tpm_pca_var$coord),
                                 big_tpm_pca_var$coord)
big_tpm_pca_coords$stage_experiment <- str_extract(
  big_tpm_pca_coords$sample,
  pattern = '.+(?=_[[:digit:]]+$)'
)
big_tpm_pca_coords$stage <- str_extract(big_tpm_pca_coords$sample, 
                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
big_tpm_pca_coords$experiment <- str_extract(big_tpm_pca_coords$sample, 
                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')

colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
                               length.out = length(unique(big_tpm_pca_coords$stage))),
                       unique(big_tpm_pca_coords$stage))

pca_importance <- as.data.frame(summary(big_tpm_pca)$importance)

plot <- ggplot(big_tpm_pca_coords,
       aes(
         x = Dim.1,
         y = Dim.2,
         col = stage,
         shape = experiment
       )) +
  geom_point() +
  scale_fill_manual(values = colornames) +
  scale_shape_manual(
    values = setNames(
      c(1:10),
      unique(big_tpm_pca_coords$experiment)
    )
  ) +
  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
  guides(shape=guide_legend(ncol=2),
         col = guide_legend(ncol=2)) +
  theme_classic() +
  theme(legend.key.size = unit(5, 'mm'),
        legend.text = element_text(size = 6))

plot

```

#Supp figure 1C - PCA using just the pir genes.

```{r supp_figure_PCA_pir}
big_tpm_pir <- berghei_tpm[berghei_tpm$Geneid %in% pir_id,-1,drop = FALSE]

big_tpm_pir <- big_tpm_pir[apply(big_tpm_pir, 2, sum) >= 50]

big_tpm_pir_pca <-  prcomp(big_tpm_pir, 
                       center = TRUE, 
                       scale. = TRUE)

big_tpm_pir_pca_var <- get_pca_var(big_tpm_pir_pca)
big_tpm_pir_pca_coords <- data.frame(sample = rownames(big_tpm_pir_pca_var$coord),
                                 big_tpm_pir_pca_var$coord)
big_tpm_pir_pca_coords$stage_experiment <- str_extract(
  big_tpm_pir_pca_coords$sample,
  pattern = '.+(?=_[[:digit:]]+$)'
)
big_tpm_pir_pca_coords$stage <- str_extract(big_tpm_pir_pca_coords$sample, 
                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
big_tpm_pir_pca_coords$experiment <- str_extract(big_tpm_pir_pca_coords$sample, 
                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')

colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
                               length.out = length(unique(big_tpm_pir_pca_coords$stage))),
                       unique(big_tpm_pir_pca_coords$stage))

pca_importance <- as.data.frame(summary(big_tpm_pir_pca)$importance)

plot_pir <- ggplot(big_tpm_pir_pca_coords,
       aes(
         x = Dim.1,
         y = Dim.2,
         col = stage,
         shape = experiment
       )) +
  geom_point() +
  scale_fill_manual(values = colornames) +
  scale_shape_manual(
    values = setNames(
      c(1:10),
      unique(big_tpm_pca_coords$experiment)
    )
  ) +
  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
  guides(shape=guide_legend(ncol=2),
         col = guide_legend(ncol=2)) +
  theme_classic() +
  theme(legend.key.size = unit(5, 'mm'),
        legend.text = element_text(size = 6))

plot_pir
```
Combining the data across the experiments.

```{r combining_stages_data}
combined_data <- berghei_tpm %>% 
  melt(value.name = 'tpm') %>% 
  mutate(stage = str_extract(variable, 
                             pattern = '(?<=_).+(?=_)')) %>% 
  group_by(Geneid, stage) %>% 
  dplyr::summarise(tpm_mean = mean(tpm),
                   tpm_med = median(tpm),
                   tpm_sd = sd(tpm),
                   tpm_max = max(tpm),
                   tpm_min = min(tpm))

combined_data <- dcast(combined_data, 
                       Geneid ~ stage, 
                       value.var = 'tpm_mean') %>% 
  column_to_rownames(var = 'Geneid')

#Ordering by life cycle progression
combined_data <- combined_data[,
                               ordering(colnames(combined_data), 
                                        order_vector = gsub(order_vector,
                                                            pattern = '-',
                                                            replacement = '\\.')
                               )]

#Check that they still sum to 1e6

apply(combined_data, 2, sum)

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9',
        '#E69F00'),
  sort(
    unique(pir_clade %>% 
             mutate(
               SorL = str_extract(clade, 
                                  pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent')
             ) %>% 
             .$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(pir_clade$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9',
    '#009E73',
    '#0072B2',
    '#E69F00',
    '#D55E00',
    '#CC79A7',
    '#999999',
    'black',
    'red'),
  sort(
   unique(pir_clade$clade)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(pir_clade$clade == clade_name),
                                   ')'
                                 )
                               })
```

# pir expression, S or L-form, stacked bar chart, stack for each gene

```{r SorL_stacked_eachGene_fig2a}
  tpm_data <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = Geneid) %>% 
    mutate(tpm = ifelse(tpm > 1, tpm, 0)) %>% #Filtering step for low value pir genes.
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

tpm_data <- mutate(tpm_data,
                   order = row_number())

tpm_data_SorL_stackgenes <- mutate(tpm_data,
                        SorL = str_extract(clade, 
                                           pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent'))
tpm_data_SorL_stackgenes <- group_by(tpm_data_SorL_stackgenes,
                                     stage, 
                                     SorL,
                                     stack_group) %>% 
  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
  #Include the number of genes in each fill category for the labels
  mutate(SorL = paste0(SorL, 
                       ' (n = ', 
                       n_distinct(stack_group),
                       ')')) %>% 
#Order by SorL and then by total_tpm, then convert stack_group to a factor so ggplot orders correctly
  arrange(SorL, 
          total_tpm) %>% 
  mutate(stage_num = as.integer(stage))
tpm_data_SorL_stackgenes$stage %>% as.integer %>% unique
tpm_data_SorL_stackgenes$stage %>% unique

tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))
```

```{r stacked_bar_chart_figure2b}
#Calculate the range for the y value to make it look nicer
stacked_max <- tpm_data %>% 
  ungroup %>% 
  group_by(stage) %>% 
  dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
  .$max_tpm %>% 
  max

max_tpm <- c(
  seq(
    0,
    mround(
      stacked_max,
      500),
    length.out = 11)
)

stack_label <- 'sub-family'

write_csv(x = tpm_data_SorL_stackgenes,
          path = 'tl20-05_SorL-subfamily_pir.csv')

stage_vec <- unique(
  levels(tpm_data$stage)
)

stage_vec <- tpm_data_SorL_stackgenes$stage
col_group <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'zzAsex' ,
                   str_detect(stage_vec, pattern = 'Liv') ~ 'zzLiv',
                   str_detect(stage_vec, pattern = 'Gam') ~ 'zzGam',
                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'zzMos')
tpm_data_SorL_stackgenes$stage_col <- col_group
fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') #Note that this order is influenced by sort(col_group)

y <- ggplot(data = tpm_data_SorL_stackgenes, 
            aes(
              x = stage_num,
              y = total_tpm, 
              fill = SorL,
              # group = stage_num
            )) +
  geom_rect(
    aes(
      xmin = stage_num - .7,
      xmax = stage_num + .7,
      ymin = -Inf,
      ymax = Inf,
      fill = stage_col
    ),
    alpha = .05
  ) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) + 
  scale_fill_manual(
    breaks = unique(tpm_data_SorL_stackgenes$SorL),
    values = c(
      SorL_col_num,
      fills
    )
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  # scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) +
  scale_x_continuous(
    breaks = tpm_data_SorL_stackgenes$stage_num %>% unique(),
    labels = tpm_data_SorL_stackgenes$stage %>% unique(),
    expand = c(0, 0)
  ) +
    theme(
      axis.text.x = element_text(angle = 90, 
                                 vjust = .5,
                                 size = 10),
      axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    ) +
    xlab('stage') + 
    facet_grid(.~stage, 
               scales = "free_x",
               space = "free_x") +
    scale_x_discrete(expand=c(0,0)) +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x = element_blank(),
      # axis.line.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=8, 
                                  face = 'bold', 
                                  angle = 90),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      ),
      panel.spacing.x = unit(0, "null")
    )

g <- ggplot_gtable(ggplot_build(y))
strip_t <- which(grepl('strip-t', g$layout$name))
stage_vec <- unique(
  levels(tpm_data$stage)
)
fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
k <- 1
for (i in strip_t) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}

grid.newpage()
grid.draw(g)

```

# figure2binset

```{r stack_bar_graph_SorL_lowexpression-figure2binset}
#Include all stages expressed lower than a certain amount

low_express <- tpm_data_SorL_stackgenes %>% 
  group_by(stage) %>% 
  summarise(sum_tpm = sum(total_tpm)) %>% 
  filter(sum_tpm < 350)

#Calculate the range for the y value to make it look nicer

stacked_max <- max(low_express$sum_tpm)

max_tpm <- c(
  seq(
    0,
    mround(
      stacked_max,
      100),
    length.out = 11)
)

tpm_data_SorL_stackgenes_low <- filter(tpm_data_SorL_stackgenes,
                                       stage %in% low_express$stage)

y <- ggplot(data = tpm_data_SorL_stackgenes_low, 
            aes(
              x = stage,
              y = total_tpm, 
              fill = SorL,
              # group = stage_num
            )) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) + 
  scale_fill_manual(
    breaks = unique(tpm_data_SorL_stackgenes$SorL),
    values = c(
      SorL_col_num,
      fills
    )
  ) +
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 15),
    axis.title.y = element_text(size = 8),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 5),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=6, 
                                face = 'bold', 
                                angle = 90),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )

#This bit is to change the colour of the facet boxes

g <- ggplot_gtable(ggplot_build(y))
strip_t <- which(grepl('strip-t', g$layout$name))
stage_vec <- unique(
  levels(tpm_data_SorL_stackgenes_low$stage)[levels(tpm_data_SorL_stackgenes_low$stage) %in% as.character(tpm_data_SorL_stackgenes_low$stage)]
)
fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
k <- 1
for (i in strip_t) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid.newpage()
grid.draw(g)
```

```{r table_of_SorL_highest_supptable}
#For a neat table
SorL_kable_index <- rep.int(2, 
                            times = length(unique(tpm_data_SorL_stackgenes$stage)))
names(SorL_kable_index) <- levels(tpm_data_SorL_stackgenes$stage)

tpm_data_SorL_stackgenes_top_table <- tpm_data_SorL_stackgenes %>% 
  mutate(Geneid = stack_group) %>% 
  filter(str_detect(SorL, 
                    pattern = 'ancestral', 
                    negate = TRUE)) %>% 
  group_by(stage) %>% 
  mutate(sum_tpm = sum(total_tpm)) %>% 
  group_by(stage, 
           SorL) %>% 
  slice_max(order_by = total_tpm,
            n = 1) %>% 
  mutate(total_tpm = round(total_tpm, digits = 2),
         prop_tpm = round(total_tpm/sum_tpm, digits = 2)) %>% 
  select(stage, Geneid, SorL, total_tpm, prop_tpm)

tpm_data_SorL_stackgenes_top_table %>% 
  kable() %>% 
  kable_styling() %>% 
  pack_rows(index = SorL_kable_index) %>% 
  scroll_box(width = "100%", height = "500px")
```

## Figure 2B - sub-family proportions

```{r combined_tpm_stacked_df}

  tpm_data_subfam <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = clade) %>% 
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

tpm_data_subfam <- mutate(tpm_data_subfam,
                   order = row_number())

  # #Apply ordering
  # tpm_data_subfam$stage = factor(tpm_data_subfam$stage, 
  #                         levels = ordering(tpm_data_subfam$stage,
  #                                           order_vector))

  #Calculate the range for the y value to make it look nicer
  stacked_max <- tpm_data_subfam %>% 
    ungroup %>% 
    group_by(stage) %>% 
    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
    .$max_tpm %>% 
    max

  max_tpm <- c(
    seq(
      0,
      mround(
        stacked_max,
        500),
      length.out = 11)
  )
```

```{r combined_tpm_stacked_plot}

  stack_label <- 'sub-family'

y <- ggplot(data = tpm_data_subfam, 
            aes(
              x = stage,
              y = total_tpm, 
              fill = clade,
              group = order
            )) +
  geom_bar(
    # position = position_stack(),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1) +
  scale_fill_manual(
    values = clade_col_num
  ) +
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = seq(0,1,0.1)) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('Proportion of '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=8, 
                                face = 'bold', 
                                angle = 90),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )

  #This bit is to change the colour of the facet boxes

  g <- ggplot_gtable(ggplot_build(y))
  strip_t <- which(grepl('strip-t', g$layout$name))
  stage_vec <- unique(

    levels(tpm_data_subfam$stage)
  )
  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
  k <- 1
  for (i in strip_t) {
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }

  grid.newpage()
  grid.draw(g)
```

## Gametocyte marker gene expression - supp figure 2

```{r}
list_genes <- c(
  list('P28' = 'PBANKA_0514900'),
  list('nek4' = 'PBANKA_0616700'),
  list('map2' = 'PBANKA_0933700'),
  list('hap2' = 'PBANKA_1212600')
)
gam_markers <- unlist(list_genes)

list_tpms_line <- lapply(expt_tpms, 
                         function(input){
                           tpm_data <- input$tpm %>% 
                             filter(Geneid %in% gam_markers) %>% 
                             mutate(Geneid = names(gam_markers)[match(Geneid, gam_markers)]) %>% 
                             melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
                             mutate(stage = unlist(lapply(strsplit(as.character(sample), 
                                                                   split = '_'),
                                                          `[`,
                                                          1))) %>% 
                             group_by(stage, 
                                      sample,
                                      Geneid) %>% 
                             dplyr::summarise(total_tpm = sum(tpm))

                           #Make stages with no cirs in the sample give a zero 
                           if(!all(colnames(input$tpm) %in% tpm_data$sample)){
                             missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
                             missing_stages_data_zero = data.frame(lapply(missing_stages, function(missing_st){
                               data.frame(stage = unlist(lapply(strsplit(as.character(missing_st), split = '_'),`[`,2)),
                                          sample = missing_st,
                                          total_tpm = 0)
                             }))
                             tpm_data <- bind_rows(tpm_data, missing_stages_data_zero)
                           }

                           #Add experiment column
                           tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
                           max_tpm = if(max(tpm_data$total_tpm) < 10){
                             c(seq(0,mround(max(tpm_data$total_tpm),1),1))
                           }else{
                             c(seq(0,mround(max(tpm_data$total_tpm),5),5))
                           }

                           #Get the average data for the line graph
                           tpm_data_line <- tpm_data %>% 
                             ungroup %>% 
                             group_by(stage, 
                                      experiment,
                                      Geneid) %>% 
                             dplyr::summarise(total_tpm_median = median(total_tpm),
                                              Number_of_replicates = n(),
                                              total_tpm_max = max(total_tpm),
                                              total_tpm_min = min(total_tpm)) %>% 
                             arrange(stage)

                           return(tpm_data_line)
                         })

tpm_data_line <- bind_rows(list_tpms_line)

tpm_data_line$stage <- factor(tpm_data_line$stage, 
                             levels = ordering(tpm_data_line$stage,
                                               order_vector))

#Calculate the range for the y value to make it look nicer
if(max(
  tpm_data_line$total_tpm_max
) >= 500){
  max_tpm = c(
    seq(
      0,
      mround(
        max(
          tpm_data_line$total_tpm_max
        ),
        500),
      length.out = 11)
  )
}else{
  max_tpm = c(
    seq(
      0,
      mround(
        max(
          tpm_data_line$total_tpm_max
        ),
        10),
      length.out = 11)
  )
}
```

super hacky way to change the scales in each facet from:
https://fishandwhistle.net/post/2018/modifying-facet-scales-in-ggplot2/

```{r}
scale_override <- function(which, scale) {
  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
    stop("which must be an integer of length 1")
  }

  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
    stop("scale must be an x or y position scale")
  }

  structure(list(which = which, scale = scale), class = "scale_override")
}

CustomFacetWrap <- ggproto(
  "CustomFacetWrap", FacetWrap,
  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
    # make the initial x, y scales list
    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)

    if(is.null(params$scale_overrides)) return(scales)

    max_scale_x <- length(scales$x)
    max_scale_y <- length(scales$y)

    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
    for(scale_override in params$scale_overrides) {
      which <- scale_override$which
      scale <- scale_override$scale

      if("x" %in% scale$aesthetics) {
        if(!is.null(scales$x)) {
          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
          scales$x[[which]] <- scale$clone()
        }
      } else if("y" %in% scale$aesthetics) {
        if(!is.null(scales$y)) {
          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
          scales$y[[which]] <- scale$clone()
        }
      } else {
        stop("Invalid scale")
      }
    }

    # return scales
    scales
  }
)

facet_wrap_custom <- function(..., scale_overrides = NULL) {
  # take advantage of the sanitizing that happens in facet_wrap
  facet_super <- facet_wrap(...)

  # sanitize scale overrides
  if(inherits(scale_overrides, "scale_override")) {
    scale_overrides <- list(scale_overrides)
  } else if(!is.list(scale_overrides) || 
            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) {
    stop("scale_overrides must be a scale_override object or a list of scale_override objects")
  }

  facet_super$params$scale_overrides <- scale_overrides

  ggproto(NULL, CustomFacetWrap,
    shrink = facet_super$shrink,
    params = facet_super$params
  )
}

(
  y <- ggplot(data = tpm_data_line, 
              aes(x = stage, 
                  y = total_tpm_median, 
                  fill = experiment, 
                  group = experiment)) +
    geom_bar(position = position_dodge2(preserve = 'single'),
             stat = 'identity') +
    scale_fill_manual(values = c(cbPalette, 'red', 'purple')[1:(length(unique(tpm_data_line$experiment)))]) +
    geom_errorbar(aes(x = stage,
                      ymax = total_tpm_max, 
                      ymin = total_tpm_min),
                  position = position_dodge2(preserve = 'single'),
                  size = 0.25) +
    scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_text(angle = 90, 
                                   size = 6, 
                                   vjust = 0.5),
          axis.text.y = element_text(size = 5),
          axis.title.y = element_text(size = 8),
          legend.title = element_text(size = 7),
          legend.text = element_text(size = 5),
          legend.key.size = unit(3,'mm'),
          legend.key.width = unit(2,'mm')) +
    scale_size(name = 'Number of genes detected over threshold', 
               range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) +
    facet_wrap_custom(~Geneid, 
                      scales = 'free_y',
                      scale_overrides = list(
                        scale_override(3, scale_y_continuous(breaks = seq(0,650,length.out = 11))),
                        scale_override(2, scale_y_continuous(breaks = seq(0,900,length.out = 11))),
                        scale_override(1, scale_y_continuous(breaks = seq(0,310,length.out = 11)))
                      )) +
    annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)
)
#Could also use functions from a package called 'lemon' to add the lines to the top two facets.
```

## Lower expression of _pir_ genes versus the rest of the genome

```{r low_pir_expression}
set.seed(4908)

gene_expression_df <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(pir_gene = ifelse(Geneid %in% pir_id,
                           'pir',
                           'non-pir')) %>% 
  group_by(stage, pir_gene) %>% 
  slice_sample(n = length(pir_id)) %>%
  summarise(mean_tpm = mean(tpm),
            med_tpm = round(median(tpm),
                            digits = 6))

g <- ggplot(gene_expression_df,
       aes(x = pir_gene,
           y = log2(1+med_tpm))) +
  xlab('Gene group')+
  ylab('l0g(Median TPM + 1)')+
  # geom_violin()+
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               # binwidth = 1,
               # dotsize = 2,
               method = 'dotdensity') +
  geom_path(aes(group = stage,
                col = stage)) +
  theme_classic() +
  theme(
    legend.text = element_text(size = 9),
    legend.key.size = unit(4, 'mm')
  )

g
```

_pir_ expression on average ranges from minimum mean `r min(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$mean_tpm)` TPM and median `r min(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$med_tpm)` TPM to maximum mean `r max(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$mean_tpm)` TPM and median `r max(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$med_tpm)` TPM. Non-\ _pir_ expression ranges from minimum mean `r min(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$mean_tpm)` TPM and median `r min(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$med_tpm)` TPM to maximum mean `r max(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$mean_tpm)` TPM and median `r max(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$med_tpm)` TPM.

### 2021-09-06 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Add data from Manuscript (berghei, chabaudi 24h, ancestral phylogeny and expression)

---
title: "tl20-05_berghei_pir_DESEQ"
author: "Timothy Little"
date: "25/01/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

Using the P. berghei systematic analysis of pir gene expression data we want to perform statistical tests for enrichment of individual pirs in certain stages of the malaria life cycle, using the package DESEQ2.

Notes:
* Have to remove Cook's distance cut off when looking at gene PBANKA_0600031 because otherwise DESEQ says that it is an outlier, from the data itself I'm not sure if I agree.

Future:
* Would be worth looking at http://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#levels-without-samples to see if we can incorporate the experiment values into this too.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)

# Set the working directory for the document where the data is located.
knitr::opts_knit$set(root.dir = 
                       ".")

# Change options so that numbers are displayed with all the zeros (e,g, as 1000000 not 1e6)
options(scipen = 999)

set.seed(8)
```

```{r load_packages}
library('dplyr')
library('readr')
library('readxl')
library('writexl')
library('data.table')
library('tibble')
library('ggplot2')
library('stringr')
library('RColorBrewer')
library('DESeq2')
library('apeglm')
```

Download the data, and remove heterogeneous life cycle stages like 'Gam' and 'BldMl'. Will keep Asex.Mixed since we want to be able to compare that to other stages, although for each Asex stage we should remove the individual stages (like Rings) when looking at mixed, and the Mixed stages when looking at the inidvidual stages.

Note that I'm using the count data with names updated by how I combined similar stages e.g. Liver.post48h.

```{r read_count_data}
berg_counts <- read_xlsx('Manuscript_Frontiers_suppinfo_bergheiCounts_UpdatedNames.xlsx')

#Filter low expressed genes (less than 1 count in all samples), convert to a matrix, add 1 for the log transformation, then transform to log2
berg_counts <- column_to_rownames(berg_counts, var = 'Geneid') %>% 
  select(!matches('BldMl|Gam_'))
berg_counts.mat <- berg_counts[!rowSums(berg_counts) <= 10,] %>% 
  as.matrix

berg_counts.mat[1:5,1:5]
```

```{r generate_genesets}
pir_info <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
                      sheet = 1)

pir_geneset <- list(pirs = pir_info$Geneid)

```

```{r colData}
( coldata <- data.frame(
  row.names = colnames(berg_counts.mat),
  experiment = str_extract(colnames(berg_counts.mat), pattern = '^E[[:digit:]]{2}'),
  stage = str_extract(colnames(berg_counts.mat), pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]{1,2})')
) )
```

```{r dds}
dds <- DESeqDataSetFromMatrix(countData = berg_counts.mat,
                              colData = coldata,
                              design = ~ stage)
dds
```

```{r deseq-ing}
dds.deseq <- DESeq(dds)
resultsNames(dds.deseq)
```

Individual _pir_ genes which we use as examples of being highly upregulated in individual stages.

```{r gamMale_v_asexMix}
coef <- "stage_Gam.Male_vs_Asex.Mixed"
dds.deseq.gamMale_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  type = "apeglm")
gamMale_v_asexMixed_df <- as.data.frame(
  dds.deseq.gamMale_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0317181'),]
) %>% 
  mutate(contrast = coef) %>% 
  rownames_to_column(var = 'Geneid')
```

```{r}
coef <- "stage_Liv.Meroz_vs_Asex.Mixed"
res <- results(dds.deseq,
               name = 'stage_Liv.Meroz_vs_Asex.Mixed',
               cooksCutoff = FALSE)
dds.deseq.livMeroz_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  res = res,
                                                  type = "apeglm")

livMeroz_v_asexMixed_df <- as.data.frame(
  dds.deseq.livMeroz_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0100500',
                                                                                              'PBANKA_0600031'),]) %>%  
  mutate(contrast = coef) %>% 
  rownames_to_column(var = 'Geneid')
```

```{r asexRing_v_asexMixed}
coef <- 'stage_Asex.Ring_vs_Asex.Mixed'

dds.deseq.asexRing_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  type = "apeglm")
asexRing_v_asexMixed_df <- as.data.frame(
  dds.deseq.asexRing_v_asexMixed.lfcShr[rownames(dds.deseq.asexRing_v_asexMixed.lfcShr) %in% c('PBANKA_1300200'),]
) %>% 
  mutate(contrast = coef) %>% 
    rownames_to_column(var = 'Geneid')
```

```{r}
dds$stage <- relevel(dds$stage, ref = 'Gam-Male')
dds.maleGam.deseq <- DESeq(dds)
resultsNames(dds.maleGam.deseq)
```

```{r}
coef <- "stage_Liv.Meroz_vs_Gam.Male"
res <- results(dds.maleGam.deseq,
               name = 'stage_Liv.Meroz_vs_Gam.Male',
               cooksCutoff = FALSE)
dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr <- lfcShrink(dds.maleGam.deseq, 
                                                         coef = coef,
                                                         res= res,
                                                         type = "apeglm")

(livMeroz_v_gamMale_df <- as.data.frame(
  dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr[rownames(dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr) %in% c('PBANKA_0317181',
                                                                                                           'PBANKA_0100500',
                                                                                                           'PBANKA_0600031'),]
  ) %>%  
  mutate(contrast = coef) %>% 
    rownames_to_column(var = 'Geneid'))
```

```{r}
(deseq_stats <- bind_rows(livMeroz_v_gamMale_df, 
          as.data.frame(gamMale_v_asexMixed_df),
          as.data.frame(livMeroz_v_asexMixed_df)) %>% 
  mutate(FDR = p.adjust(pvalue, 
                        method = 'fdr')) %>% 
  select(-c(lfcSE, 
            padj, 
            pvalue)))
```

```{r save_stats_df}
write_xlsx(deseq_stats, 
           'SupplementaryInformation4_Stats_DESEQ_berghei.xlsx')
write.csv(deseq_stats, 
           'SupplementaryInformation4_Stats_DESEQ_berghei.csv')
```

### 2021-09-06 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Add data from Manuscript (berghei, chabaudi 24h, ancestral phylogeny and expression)

---
title: "tl20-05_berghei_pir_GeneSetEnrichmentQusage"
author: "Timothy Little"
date: "15/01/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

Using the P. berghei systematic analysis of pir gene expression data we want to perform statistical tests for enrichment of sets of pirs in certain stages of the malaria life cycle, using the package Qusage.

Notes:

* Do I want to remove the other Asexual samples, or the other Gam samples, when looking into these stages? Probably.
* Probably want to remove samples which express almost no pirs. E.g. Sporoz is signif. downregulated for everything which is no surprise since it basically doesn't express pir genes.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)

# Set the working directory for the document where the data is located.
knitr::opts_knit$set(root.dir = 
                       ".")

# Change options so that numbers are displayed with all the zeros (e,g, as 1000000 not 1e6)
options(scipen = 999)

set.seed(89)
```

```{r load_packages}
library('dplyr')
library('readr')
library('readxl')
library('data.table')
library('tibble')
library('qusage')
library('ggplot2')
library('stringr')
library('RColorBrewer')
library('writexl')
```

Download the data, and remove heterogeneous life cycle stages like 'Gam' and 'BldMl'. Will keep Asex.Mixed since we want to be able to compare that to other stages, although for each Asex stage we should remove the individual stages (like Rings) when looking at mixed, and the Mixed stages when looking at the individual stages.

```{r read_count_data}

berg_tpm <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
                      sheet = 'berg_TPM_allExpts_bioReps') %>% 
  dplyr::rename_with(.fn = ~ str_replace(.x, 
                                         pattern = 'Expt', 
                                         replacement = 'E'))

#Filter low expressed genes (greater than 10 tpm in at least one sample), convert to a matrix, add 1 for the log transformation, then transform to log2

berg_tpm <- column_to_rownames(berg_tpm, var = 'Geneid') %>% 
  select(!matches('BldMl|Gam_'))

berg_tpm <- berg_tpm[apply(berg_tpm, 1, function(x) {any(x >= 10)}),]

berg_tpm_log2plus1.mat <- berg_tpm %>% 
  as.matrix %>% 
  `+`(.+1) %>% 
  log(base = 2)

berg_tpm_log2plus1.mat[1:5,1:5]

```

```{r generate_labels}
#Use make.names to remove the '-' from the labels
colnames(berg_tpm_log2plus1.mat) <- make.names(colnames(berg_tpm_log2plus1.mat))

labels <- colnames(berg_tpm_log2plus1.mat)

#We want the life cycle stages to be the different groups for now.

lifeCycle_labels <- str_extract(labels, 
                                pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]{1,2})')
unique(lifeCycle_labels) %>% sort

```

```{r generate_genesets}
pir_info <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
                      sheet = 1)

pir_geneset <- list(pirs = pir_info$Geneid)

subfam_list <- lapply(unique(pir_info$clade), 
                     function(clade_nam) filter(pir_info, 
                                                clade == clade_nam)$Geneid)
names(subfam_list) <- unique(pir_info$clade)
subfam_list <- subfam_list[unlist(lapply(subfam_list, function(x) length(x) > 3))]

ShortLong_list <- 
  lapply(unique(pir_info$SorL), 
                     function(SorL_nam) filter(pir_info, 
                                                SorL == SorL_nam)$Geneid)
names(ShortLong_list) <- unique(pir_info$SorL)
ShortLong_list <- ShortLong_list[unlist(lapply(ShortLong_list, function(x) length(x) > 3))]

```

What is the question that I actually want to answer? I have a lot of life stages and I want to know if each life stage contains an enrichment for transcription of certain groups of pirs compared to the others. So does this mean that I want to do a pairwise comparison, for each life cycle stage, versus every other identified stage? i.e. lapply for each stage with the 'labels' set as that stage and the rest all set to 'other'?

```{r running_qusage_malevfemale}
maleVfemale <- lapply(
  list(pir_geneset), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Gam.Female',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)

```

```{r lateliverstages}

post48livvearlyliv <- lapply(
  list(pir_geneset,
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Liv.post48h-Liv.24h',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)

```

```{r}
livmero_livpost48.results <- lapply(
  list(pir_geneset,
       ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Liv.Meroz-Liv.post48h',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)

```

```{r}
asexRing_asexMix.results <- lapply(
  list(pir_geneset,
       subfam_list,
       ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Asex.Ring-Asex.Mixed',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
names(asexRing_asexMix.results) <- c('all', 'subfam', 'SorL')
asexRing_asexMix.results$subfam$qstable
asexRing_asexMix.results$SorL$qstable
```

```{r}
maleGam_v_Asex.Mix.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Asex.Mixed',
                                gene_list,
                                n.points = 2^18)

    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
    #                                 FDR < 0.05)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
names(maleGam_v_Asex.Mix.results) <- c('all', 'subfam', 'SorL')

plotGeneSetDistributions(maleGam_v_Asex.Mix.results$SorL$results, path.index =  'S')
plotGeneSetDistributions(maleGam_v_Asex.Mix.results$SorL$results, path.index = 'L')
plotGeneSetDistributions(maleGam_v_Asex.Mix.results$subfam$results, path.index = 'S4')
mutate(maleGam_v_Asex.Mix.results$subfam$qstable , adjust = p.adjust(p.Value, method = 'fdr'))
```

```{r}
maleGam_v_Liv.Meroz.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Liv.Meroz',
                                gene_list,
                                n.points = 2^18)

    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
    #                                 FDR < 0.05)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
```

```{r}
Asex.Mix_v_Liv.Meroz.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Asex.Mixed-Liv.Meroz',
                                gene_list,
                                n.points = 2^18)

    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
    #                                 FDR < 0.05)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
names(Asex.Mix_v_Liv.Meroz.results) <- c('all', 'subfam', 'SorL')

plotGeneSetDistributions(Asex.Mix_v_Liv.Meroz.results$SorL$results, 
                         path.index = 'L')
plotGeneSetDistributions(Asex.Mix_v_Liv.Meroz.results$SorL$results, 
                         path.index = 'S')

bind_rows(lapply(Asex.Mix_v_Liv.Meroz.results, 
                 getElement, 
                 'qstable')) %>% 
  mutate(adjust = p.adjust(p.Value)) %>% 
  arrange(adjust)
```

```{r}
berghei_comparison_stats_df <- bind_rows(
  lapply(c(maleGam_v_Asex.Mix.results, 
           maleGam_v_Liv.Meroz.results, 
           Asex.Mix_v_Liv.Meroz.results,
           maleVfemale, 
           post48livvearlyliv, 
           livmero_livpost48.results),
         getElement,
         'qstable')
) %>% 
  select(contrast, everything()) %>% 
  mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
  filter(FDR < 0.05) %>% 
  arrange(contrast)

max_logfold <- ceiling(max(max(berghei_comparison_stats_df$log.fold.change), 
                           abs(min(berghei_comparison_stats_df$log.fold.change))))

(
x <- ggplot(berghei_comparison_stats_df, 
            aes(x = contrast, 
                y = reorder(pathway.name, dplyr::desc(pathway.name)))) +
  xlab('Comparison') + ylab('Geneset') + 
  # scale_x_discrete(labels= c(1,2,3,4,6)) +
  geom_count(aes(
                 colour = log.fold.change), 
             na.rm = TRUE) +
  labs(colour = 'log2 Fold Change') +
  scale_color_gradientn(
    colours = rev(brewer.pal(11, "RdBu")),
    na.value="transparent",
    limits = c(-max_logfold,max_logfold),
    # limits = c(-0.5,1.25)
  ) +
  scale_size_area(max_size = 6) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = 'white'),
        plot.background = element_blank(),
        legend.background = element_rect(fill = 'transparent'),
        axis.text.x = element_text(angle = 90))
  )
```

```{r save_xlsx}
to_save <- bind_rows(
  lapply(c(maleGam_v_Asex.Mix.results, 
           maleGam_v_Liv.Meroz.results, 
           Asex.Mix_v_Liv.Meroz.results,
           maleVfemale, 
           post48livvearlyliv, 
           livmero_livpost48.results),
         getElement,
         'qstable')
) %>% 
  select(contrast, everything()) %>% 
  mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
  arrange(contrast, pathway.name)

write_xlsx(to_save,
           path = 'SupplementaryInformation_Stats_Qusage_berghei.xlsx')
write.csv(to_save,
          'SupplementaryInformation_Stats_Qusage_berghei.csv')
```

### 2021-09-06 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_CalculateTPM.Rmd
Save hclust plots from berg

@@
 library('ComplexHeatmap')
+library('ggdendro') #Drawing ggplot dendrograms
 
@@
 
-all.equal(berghei_tpm[3345:3350,23:28],check_tpm)
+all.equal(berghei_tpm[3345:3350,colnames(check_tpm)],check_tpm)
 
@@
 
+# Hierarchical clustering plots.
+
 ```{r hclusts, message = FALSE}
@@
            par(cex=0.7)
-           plot(hclust(dists),
-                main = experiment_name)
+           p <- ggdendrogram(hclust(dists)) +
+             ggtitle(experiment_name)
+           ggsave(filename = paste0('thesis_chapter3_figure_hclust_', 
+                                    experiment_name, 
+                                    '.png'),
+                  plot = p,
+                  path = 'plots')
+           return(p)
          }
-       })
-
+       }
+)
 ```

### 2021-09-06 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
density plots for berg and r fun for sourcing

@@
 
+## Density plots
+
+```{r density_plots}
+# Compute the average bandwidth for all the samples so that we have the same value for each figure.
+list_bw = lapply(expt_tpms, function(expt){ 
+  lapply(colnames(expt$tpm)[-1], 
+         function(sample){
+           data = as.numeric(unlist(select(expt$tpm, 
+                                           paste(sample))))
+           dens = density(log1p(data))
+           dens$bw
+         })
+})
+bw_mean = mean(unlist(list_bw))
+bw_range = max(unlist(unlist(list_bw))) - min(unlist(unlist(list_bw)))
+print(paste('Bandwidth range = ', bw_range))
+
+source('../rfunctions/density_samples_fun.R')
+lapply(expt_tpms, density_samples, save = TRUE)
+```
+
 ```{r melt_df}
@@
 
-##Supp figure 1A - Pearson correlation plot of collated samples.
+##Pearson correlation plot of collated samples.
 

### 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
add berg biplot

@@
 source('../rfunctions/density_samples_fun.R')
-lapply(expt_tpms, density_samples, save = TRUE)
+lapply(expt_tpms, 
+       density_samples, 
+       save = TRUE)
 ```
@@
 ```{r supp_figure_PCA_allgenes}
-
-big_tpm_pca <-  prcomp(berghei_tpm[,-1], 
+#Set the Geneids as the berghei_tpm rownames so that the PCA function retains them as loading variables
+big_tpm_pca <-  prcomp(column_to_rownames(berghei_tpm, var = 'Geneid'), 
                        center = TRUE, 
@@
 
+# Biplot of the PCA using all genes.
+
+```{r biplot}
+pca_object_loadings <- data.frame(Geneid = rownames(big_tpm_pca$x),
+                                    big_tpm_pca$x)
+pca_object_loadings$Geneid <- paste0(
+  pca_object_loadings$Geneid,
+  ' ',
+  allgene.info$Description[match(pca_object_loadings$Geneid,
+                                 allgene.info$Geneid)]
+)
+
+plot +
+  geom_point(data = pca_object_loadings,
+             aes(x = PC1*0.005, y = PC2*0.005),
+             inherit.aes = FALSE, size = 0.75)
+```
+
 #Supp figure 1C - PCA using just the pir genes.
@@
 ```
+
 Combining the data across the experiments.

### 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
Fix density function and add biplot labels

@@
        density_samples, 
-       save = TRUE)
+       save = FALSE)
 ```
@@
 
-pdf(file = 'Manuscript_suppfig1A_corrplot_all_experiments_rnaseq.pdf',
-    height = 5,
-    width = 5)
+# pdf(file = 'Manuscript_suppfig1A_corrplot_all_experiments_rnaseq.pdf',
+#     height = 5,
+#     width = 5)
 corrplot(cor(big_top_recast), method = 'color', 
@@
          col = c(viridis(100), magma(115)), addrect = 6)
-dev.off()
+# dev.off()
 
@@
 ```{r biplot}
-pca_object_loadings <- data.frame(Geneid = rownames(big_tpm_pca$x),
-                                    big_tpm_pca$x)
-pca_object_loadings$Geneid <- paste0(
-  pca_object_loadings$Geneid,
-  ' ',
-  allgene.info$Description[match(pca_object_loadings$Geneid,
+library('ggrepel')
+set.seed(982)
+
+pca_object_loadings <- data.frame(big_tpm_pca$x)
+
+#Remove the large PCA object
+rm(big_tpm_pca)
+
+#Calculate the highest loadings, only want these labels
+
+pca_object_loadings$Geneid <- ''
+
+highest_loadings <- pca_object_loadings[rev(
+    order(
+      apply(
+        abs(
+          data.frame(pca_object_loadings$PC1, 
+                     pca_object_loadings$PC2)),
+        1,
+        sum)
+    )
+  )[1:10],]
+
+pca_object_loadings[rev(
+  order(
+    apply(
+      abs(
+        data.frame(pca_object_loadings$PC1, 
+                   pca_object_loadings$PC2)),
+      1,
+      sum)
+  )
+)[1:10],]$Geneid <- 
+  allgene.info$Description[match(rownames(highest_loadings),
   allgene.info$Geneid)]
-)
 
@@
              aes(x = PC1*0.005, y = PC2*0.005),
-             inherit.aes = FALSE, size = 0.75)
+             inherit.aes = FALSE, size = 0.75) +
+  geom_label_repel(data = pca_object_loadings,
+             aes(x = PC1*0.005, y = PC2*0.005,
+                 label = Geneid),
+             inherit.aes = FALSE, size = 3,
+            nudge_x = 0.4,
+            force = 100)
 ```

### 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
save biplot

@@
 
-plot +
+loadings_scale <- 0.005
+
+bp <- plot +
   geom_point(data = pca_object_loadings,
-             aes(x = PC1*0.005, y = PC2*0.005),
+             aes(x = PC1*loadings_scale, y = PC2*loadings_scale),
              inherit.aes = FALSE, size = 0.75) +
   geom_label_repel(data = pca_object_loadings,
-             aes(x = PC1*0.005, y = PC2*0.005,
+             aes(x = PC1*loadings_scale, y = PC2*loadings_scale,
                  label = Geneid),
@@
             force = 100)
+
+ggsave('thesis_chapter3_figure_biplot.png',
+       bp)
+
 ```

### 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
save biplot in plots dir
improve biplot

--- 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
@@
 #Calculate the highest loadings, only want these labels
-
+top_loadings <- 15
 pca_object_loadings$Geneid <- ''
@@
     )
-  )[1:10],]
-
-pca_object_loadings[rev(
-  order(
-    apply(
-      abs(
-        data.frame(pca_object_loadings$PC1, 
-                   pca_object_loadings$PC2)),
-      1,
-      sum)
-  )
-)[1:10],]$Geneid <- 
-  allgene.info$Description[match(rownames(highest_loadings),
-  allgene.info$Geneid)]
-
-plot +
-  geom_point(data = pca_object_loadings,
-             aes(x = PC1*0.005, y = PC2*0.005),
+)[1:top_loadings],]
+
+highest_loadings$Geneid <- 
+  gsub(
+    gsub(allgene.info$Description[match(rownames(highest_loadings),
+                                      allgene.info$Geneid)], 
+       pattern = 'putative',
+       replacement = 'put.'),
+    pattern = 'early transcribed membrane protein',
+    replacement = 'ETRAMP')
+
+loadings_scale <- 0.005
+
+(bp <- plot +
+  geom_point(data = highest_loadings,
+             aes(x = PC1*loadings_scale, y = PC2*loadings_scale),
              inherit.aes = FALSE, size = 0.75) +
-  geom_label_repel(data = pca_object_loadings,
-             aes(x = PC1*0.005, y = PC2*0.005,
+  geom_text_repel(data = highest_loadings,
+             aes(x = PC1*loadings_scale, y = PC2*loadings_scale,
                  label = Geneid),
-             inherit.aes = FALSE, size = 3,
-            nudge_x = 0.4,
-            force = 100)
+             inherit.aes = FALSE, 
+             size = 3,
+             min.segment.length = 0,
+            force = 100
+            ) )
+
+ggsave('plots/thesis_chapter3_figure_biplot.png',
+       width = 7,
+       height = 5,
+       units = "in",
+       bp)
+
 ```

### 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_CalculateTPM.Rmd
Fix consolidated similar stages and gam marker plot ordering

@@
          tpm_data <- select(berghei_tpm, matches(experiment_name))
-         if(ncol(tpm_data) > 3){
+         if(ncol(tpm_data) > 4){ 
+           #No point doing hclusts for any of the expts with under four samples
            dists = dist(t(tpm_data))
@@
                                                                           pattern = '_[[:digit:]]{1,}$')])
+
+colnames(berghei_tpm) <- col_stage
 ```

### 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
Fix consolidated similar stages and gam marker plot ordering

@@
                              levels = ordering(tpm_data_line$stage,
-                                               order_vector))
+                                               r_order_vector))
 
@@
 #Could also use functions from a package called 'lemon' to add the lines to the top two facets.
+
+ggsave('plots/thesis_chapter3_figure_gametocyteMarkers.png',
+       width = 7,
+       height = 5,
+       units = "in",
+       y)
 ```
@@
   xlab('Gene group')+
-  ylab('l0g(Median TPM + 1)')+
+  ylab('log(Median TPM + 1)')+
   # geom_violin()+

### 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
Save pca and corrplot

@@
 library('ComplexHeatmap')
+library('ggrepel')
 
@@
 
-# pdf(file = 'Manuscript_suppfig1A_corrplot_all_experiments_rnaseq.pdf',
-#     height = 5,
-#     width = 5)
+png(file = 'plots/thesis_chapter3_figure_corrplot.png',
+    height = 5,
+    width = 5,
+    units = "in",
+    res = 330)
 corrplot(cor(big_top_recast), method = 'color', 
@@
          col = c(viridis(100), magma(115)), addrect = 6)
-# dev.off()
+dev.off()
 
@@
 
+(
   plot <- ggplot(big_tpm_pca_coords,
@@
         legend.text = element_text(size = 6))
-
-plot
+  )
+
+ggsave('plots/thesis_chapter3_figure_pca_allGenes.png',
+       width = 7,
+       height = 5,
+       units = "in",
+       plot)
 
@@
 ```{r biplot}
-library('ggrepel')
 set.seed(982)
@@
        bp)
-
+#Note that CSP is the 140 highest laodings gene.
 ```

### 2021-09-08 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
Fix density plots with two column legend location

@@
 
-```{r density_plots}
+```{r bandwidth}
 # Compute the average bandwidth for all the samples so that we have the same value for each figure.
@@
 print(paste('Bandwidth range = ', bw_range))
-
+```
+
+```{r density_plots}
 source('../rfunctions/density_samples_fun.R')
@@
        density_samples, 
-       save = FALSE)
+       save = TRUE)
 ```

### 2021-09-09 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
Split PCA/hclust/topGenes file

--- 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
@@
 library('ggplot2')
-library('DESeq2')
 library("RColorBrewer")
@@
 library('stringr')
-library('kableExtra')
 library('viridis')
@@
 library('circlize') # for circular plots
-library('cluster')    # clustering algorithms
-library('factoextra') # clustering algorithms & visualization
-library('ggfortify') # so ggplot understands prcomp objects
-library('ape') # for phylogeny plotting
-library('dendextend') # for dendrogram plotting
 library('plotly')
@@
 
-## Saving the TPM
-
-```{r save_tpm}
+## Reading the TPM
+
+```{r read_tpm}
 berghei_tpm <- read.csv('berghei_tpm.csv', 
@@
 names(expt_tpms) <- expt_nams
-```
-
-## Density plots
-
-```{r density_plots}
-# Compute the average bandwidth for all the samples so that we have the same value for each figure.
-list_bw = lapply(expt_tpms, function(expt){ 
-  lapply(colnames(expt$tpm)[-1], 
-         function(sample){
-           data = as.numeric(unlist(select(expt$tpm, 
-                                           paste(sample))))
-           dens = density(log1p(data))
-           dens$bw
-         })
-})
-bw_mean = mean(unlist(list_bw))
-bw_range = max(unlist(unlist(list_bw))) - min(unlist(unlist(list_bw)))
-print(paste('Bandwidth range = ', bw_range))
-
-source('../rfunctions/density_samples_fun.R')
-lapply(expt_tpms, 
-       density_samples, 
-       save = FALSE)
-```
-
-```{r melt_df}
-berghei_tpm_melt <- berghei_tpm %>% 
-    #Melt the data frame for using dplyr
-    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-    mutate(experiment = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1)),
-           stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,2)))
-
-berghei_tpm_avg_byexpt <- berghei_tpm_melt %>% 
-  group_by(Geneid, experiment, stage) %>%  
-  dplyr::summarise(median_tpm = median(tpm),
-                   mean_tpm = mean(tpm), 
-                   min_tpm = min(tpm), 
-                   max_tpm = max(tpm), 
-                   sd_tpm = sd(tpm)) %>% 
-  mutate(expt_stage = paste0(experiment, '_', stage))
-```
-
-## The highest expressed _pir_ genes across the samples
-
-```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
-top_n_genes <- function(tpm_avg, 
-                        n = 5, 
-                        threshold = 0){
-  #Calculating the index used to make the kable table
-
-  kable_index = rep(n, times = length(unique(tpm_avg$stage)))
-  names(kable_index) = unique(tpm_avg$stage)
-
-  tpm_avg <- tpm_avg %>% 
-    ungroup %>% 
-    group_by(stage)
-
-  all_genes = tpm_avg %>% 
-    slice_max(order_by = mean_tpm, n = n) %>% 
-    dplyr::arrange(stage, desc(mean_tpm)) %>% 
-    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
-    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
-                                color = ifelse(mean_tpm > threshold, "red", "blue")))
-  all_genes$short_descr = allgene.info$handy_descriptions[match(all_genes$Geneid, 
-                                                                allgene.info$Geneid)]
-  all_genes$description = allgene.info$Description[match(all_genes$Geneid, 
-                                                         allgene.info$Geneid)]
-  all_genes = all_genes %>% 
-    kable(format = 'html', 
-          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed total genes'), 
-          escape = FALSE) %>%
-    kable_styling() %>% 
-    pack_rows(index = kable_index)
-
-  pir_genes = tpm_avg %>% 
-    filter(Geneid %in% pir_id) %>% 
-    slice_max(order_by = mean_tpm, n = n) %>% 
-    dplyr::arrange(stage, desc(mean_tpm)) %>% 
-    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
-    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
-                                color = ifelse(mean_tpm > threshold, "red", "blue"))) %>% 
-    mutate(Geneid = as.character(Geneid)) %>% 
-    mutate(Description = allgene.info$handy_descriptions[match(Geneid, 
-                                                               allgene.info$Geneid)]) %>% 
-    mutate(Geneid = ifelse(grepl(Geneid,
-                                 pattern = 'PBANKA_0100500'),
-                           cell_spec(Geneid, bold = TRUE),
-                           Geneid)) %>%
-    kable(format = 'html', 
-          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed pir genes'), 
-          escape = FALSE) %>% 
-    pack_rows(index = kable_index) %>% 
-    kable_styling()
-
-  # print(paste0('Completed analysis of ', tpm_input$experiment))
-  list = list(all_genes, pir_genes, unique(tpm_avg$experiment))
-  names(list) = c('all_genes', 'pir_genes', 'experiment')
-  return(list)
-}
-
-### Generate and save top genes - core dataset all_top
-
-all_top <- lapply(unique(berghei_tpm_avg_byexpt$experiment), 
-                  function(expt) top_n_genes(berghei_tpm_avg_byexpt[berghei_tpm_avg_byexpt$experiment == expt,]))
-
-all_top[[1]]
-
-```
-
-##Pearson correlation plot of collated samples.
-
-```{r all_samples_collated_heatmap}
-
-## Collated across the stages for a more easily understood heatmap
-big_top <- berghei_tpm_avg_byexpt
-
-big_top_recast = dcast(big_top, Geneid ~ expt_stage, value.var = 'mean_tpm')
-big_top_recast <- column_to_rownames(big_top_recast, var = 'Geneid')
-# big_top_recast = big_top_recast[,-1,drop=FALSE]
-
-# apply(big_top_recast, 2, sum)
-
-# pdf(file = 'Manuscript_suppfig1A_corrplot_all_experiments_rnaseq.pdf',
-#     height = 5,
-#     width = 5)
-corrplot(cor(big_top_recast), method = 'color', 
-         order = 'hclust', hclust.method = 'ward.D2',
-         cl.lim = c(-0.036,1), tl.cex = 0.50,
-         col = c(viridis(100), magma(115)), addrect = 6)
-# dev.off()
-
-corrplot(cor(big_top_recast), method = 'color', 
-         order = 'hclust', hclust.method = 'ward.D2',
-         cl.lim = c(-0.036,1), tl.cex = 0.50,
-         col = c(viridis(100), magma(115)), addrect = 6)
-```
-
-## Supp figure 1B - PCA of all the samples.
-
-```{r supp_figure_PCA_allgenes}
-#Set the Geneids as the berghei_tpm rownames so that the PCA function retains them as loading variables
-big_tpm_pca <-  prcomp(column_to_rownames(berghei_tpm, var = 'Geneid'), 
-                       center = TRUE, 
-                       scale. = TRUE)
-
-big_tpm_pca_var <- get_pca_var(big_tpm_pca)
-big_tpm_pca_coords <- data.frame(sample = rownames(big_tpm_pca_var$coord),
-                                 big_tpm_pca_var$coord)
-big_tpm_pca_coords$stage_experiment <- str_extract(
-  big_tpm_pca_coords$sample,
-  pattern = '.+(?=_[[:digit:]]+$)'
-)
-big_tpm_pca_coords$stage <- str_extract(big_tpm_pca_coords$sample, 
-                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
-big_tpm_pca_coords$experiment <- str_extract(big_tpm_pca_coords$sample, 
-                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')
-
-colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
-                               length.out = length(unique(big_tpm_pca_coords$stage))),
-                       unique(big_tpm_pca_coords$stage))
-
-pca_importance <- as.data.frame(summary(big_tpm_pca)$importance)
-
-plot <- ggplot(big_tpm_pca_coords,
-       aes(
-         x = Dim.1,
-         y = Dim.2,
-         col = stage,
-         shape = experiment
-       )) +
-  geom_point() +
-  scale_fill_manual(values = colornames) +
-  scale_shape_manual(
-    values = setNames(
-      c(1:10),
-      unique(big_tpm_pca_coords$experiment)
-    )
-  ) +
-  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-  guides(shape=guide_legend(ncol=2),
-         col = guide_legend(ncol=2)) +
-  theme_classic() +
-  theme(legend.key.size = unit(5, 'mm'),
-        legend.text = element_text(size = 6))
-
-plot
-
-```
-
-# Biplot of the PCA using all genes.
-
-```{r biplot}
-library('ggrepel')
-set.seed(982)
-
-pca_object_loadings <- data.frame(big_tpm_pca$x)
-
-#Remove the large PCA object
-rm(big_tpm_pca)
-
-#Calculate the highest loadings, only want these labels
-top_loadings <- 15
-pca_object_loadings$Geneid <- ''
-
-highest_loadings <- pca_object_loadings[rev(
-    order(
-      apply(
-        abs(
-          data.frame(pca_object_loadings$PC1, 
-                     pca_object_loadings$PC2)),
-        1,
-        sum)
-    )
-)[1:top_loadings],]
-
-highest_loadings$Geneid <- 
-  gsub(
-    gsub(allgene.info$Description[match(rownames(highest_loadings),
-                                      allgene.info$Geneid)], 
-       pattern = 'putative',
-       replacement = 'put.'),
-    pattern = 'early transcribed membrane protein',
-    replacement = 'ETRAMP')
-
-loadings_scale <- 0.005
-
-(bp <- plot +
-  geom_point(data = highest_loadings,
-             aes(x = PC1*loadings_scale, y = PC2*loadings_scale),
-             inherit.aes = FALSE, size = 0.75) +
-  geom_text_repel(data = highest_loadings,
-             aes(x = PC1*loadings_scale, y = PC2*loadings_scale,
-                 label = Geneid),
-             inherit.aes = FALSE, 
-             size = 3,
-             min.segment.length = 0,
-            force = 100
-            ) )
-
-ggsave('plots/thesis_chapter3_figure_biplot.png',
-       width = 7,
-       height = 5,
-       units = "in",
-       bp)
-
-```
-
-#Supp figure 1C - PCA using just the pir genes.
-
-```{r supp_figure_PCA_pir}
-big_tpm_pir <- berghei_tpm[berghei_tpm$Geneid %in% pir_id,-1,drop = FALSE]
-
-big_tpm_pir <- big_tpm_pir[apply(big_tpm_pir, 2, sum) >= 50]
-
-big_tpm_pir_pca <-  prcomp(big_tpm_pir, 
-                       center = TRUE, 
-                       scale. = TRUE)
-
-big_tpm_pir_pca_var <- get_pca_var(big_tpm_pir_pca)
-big_tpm_pir_pca_coords <- data.frame(sample = rownames(big_tpm_pir_pca_var$coord),
-                                 big_tpm_pir_pca_var$coord)
-big_tpm_pir_pca_coords$stage_experiment <- str_extract(
-  big_tpm_pir_pca_coords$sample,
-  pattern = '.+(?=_[[:digit:]]+$)'
-)
-big_tpm_pir_pca_coords$stage <- str_extract(big_tpm_pir_pca_coords$sample, 
-                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
-big_tpm_pir_pca_coords$experiment <- str_extract(big_tpm_pir_pca_coords$sample, 
-                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')
-
-colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
-                               length.out = length(unique(big_tpm_pir_pca_coords$stage))),
-                       unique(big_tpm_pir_pca_coords$stage))
-
-pca_importance <- as.data.frame(summary(big_tpm_pir_pca)$importance)
-
-plot_pir <- ggplot(big_tpm_pir_pca_coords,
-       aes(
-         x = Dim.1,
-         y = Dim.2,
-         col = stage,
-         shape = experiment
-       )) +
-  geom_point() +
-  scale_fill_manual(values = colornames) +
-  scale_shape_manual(
-    values = setNames(
-      c(1:10),
-      unique(big_tpm_pca_coords$experiment)
-    )
-  ) +
-  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-  guides(shape=guide_legend(ncol=2),
-         col = guide_legend(ncol=2)) +
-  theme_classic() +
-  theme(legend.key.size = unit(5, 'mm'),
-        legend.text = element_text(size = 6))
-
-plot_pir
 ```

### 2021-09-09 chapter3_results_bergheiPir/Pberghei_pir_samples_hclust-PCA-topGenes.Rmd
Density plot function and PCA/hclust/topGenes file

---
title: "P. berghei pir - Sample clustering/PCA"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r load_packages, include = FALSE, message = FALSE}
library('readxl')
library('readr')
library('dplyr')
library('purrr')
library('ggplot2')
library('DESeq2')
library("RColorBrewer")
library('colorspace')
library('data.table')
library('gridExtra')
library('tibble')
library('scales')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('plotly')
library('ggpattern')
library('ComplexHeatmap')
library('ggrepel')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, select, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

#Save plots?
global_save_var <- FALSE

```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r pir_id, message=FALSE, results='hide'}

# Getting the pir gene names and transcript lengths needed for calculating the transcript-per-million

# First we need to get the _pir gene names. These have been downloaded using a search of PlasmoDB P. berghei ANKA genome v3 using  "\*pir\*" search and manual removal of genes which clearly were not pir.

allgene.info.file <- '../standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv'
allgene.info <- as_tibble(
  read_csv(allgene.info.file)
) %>%
  mutate(Geneid = `Gene ID`, 
         Description = `Product Description`,
         is_pseudo = `Is Pseudo`,
         transcript_length = `Transcript Length`,
         paralog_count = `Paralog count`) %>% 
  select(Geneid, 
         Description, 
         is_pseudo, 
         transcript_length, 
         paralog_count)

pir_info <- allgene.info %>% 
  filter(str_detect(Description, 
                    pattern = 'PIR(?!.+pseud)'), 
         is_pseudo == 'No') %>% 
  filter(!Geneid %in% 'PBANKA_0524600')
#Removing the pir-like gene.

pir_id <- pir_info$Geneid
```

## Reading the TPM

```{r read_tpm}
berghei_tpm <- read.csv('berghei_tpm.csv', 
                        row.names = 1)
```

## Density plots

```{r bandwidth}
# Compute the average bandwidth for all the samples so that we have the same value for each figure.
list_bw = lapply(expt_tpms, function(expt){ 
  lapply(colnames(expt$tpm)[-1], 
         function(sample){
           data = as.numeric(unlist(select(expt$tpm, 
                                           paste(sample))))
           dens = density(log1p(data))
           dens$bw
         })
})
bw_mean = mean(unlist(list_bw))
bw_range = max(unlist(unlist(list_bw))) - min(unlist(unlist(list_bw)))
print(paste('Bandwidth range = ', bw_range))
```

```{r density_plots}
source('../rfunctions/density_samples_fun.R')
lapply(expt_tpms, 
       density_samples, 
       save = global_save_var)
```

```{r melt_df}
berghei_tpm_melt <- berghei_tpm %>% 
    #Melt the data frame for using dplyr
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    mutate(experiment = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1)),
           stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,2)))

berghei_tpm_avg_byexpt <- berghei_tpm_melt %>% 
  group_by(Geneid, experiment, stage) %>%  
  dplyr::summarise(median_tpm = median(tpm),
                   mean_tpm = mean(tpm), 
                   min_tpm = min(tpm), 
                   max_tpm = max(tpm), 
                   sd_tpm = sd(tpm)) %>% 
  mutate(expt_stage = paste0(experiment, '_', stage))
```

## The highest expressed _pir_ genes across the samples

```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
top_n_genes <- function(tpm_avg, 
                        n = 5, 
                        threshold = 0){
  #Calculating the index used to make the kable table

  kable_index = rep(n, times = length(unique(tpm_avg$stage)))
  names(kable_index) = unique(tpm_avg$stage)

  tpm_avg <- tpm_avg %>% 
    ungroup %>% 
    group_by(stage)

  all_genes = tpm_avg %>% 
    slice_max(order_by = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue")))
  all_genes$short_descr = allgene.info$handy_descriptions[match(all_genes$Geneid, 
                                                                allgene.info$Geneid)]
  all_genes$description = allgene.info$Description[match(all_genes$Geneid, 
                                                         allgene.info$Geneid)]
  all_genes = all_genes %>% 
    kable(format = 'html', 
          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed total genes'), 
          escape = FALSE) %>%
    kable_styling() %>% 
    pack_rows(index = kable_index)

  pir_genes = tpm_avg %>% 
    filter(Geneid %in% pir_id) %>% 
    slice_max(order_by = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue"))) %>% 
    mutate(Geneid = as.character(Geneid)) %>% 
    mutate(Description = allgene.info$handy_descriptions[match(Geneid, 
                                                               allgene.info$Geneid)]) %>% 
    mutate(Geneid = ifelse(grepl(Geneid,
                                 pattern = 'PBANKA_0100500'),
                           cell_spec(Geneid, bold = TRUE),
                           Geneid)) %>%
    kable(format = 'html', 
          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed pir genes'), 
          escape = FALSE) %>% 
    pack_rows(index = kable_index) %>% 
    kable_styling()

  # print(paste0('Completed analysis of ', tpm_input$experiment))
  list = list(all_genes, pir_genes, unique(tpm_avg$experiment))
  names(list) = c('all_genes', 'pir_genes', 'experiment')
  return(list)
}

### Generate and save top genes - core dataset all_top

all_top <- lapply(unique(berghei_tpm_avg_byexpt$experiment), 
                  function(expt) top_n_genes(berghei_tpm_avg_byexpt[berghei_tpm_avg_byexpt$experiment == expt,]))

```

##Pearson correlation plot of collated samples.

```{r all_samples_collated_heatmap}

## Collated across the stages for a more easily understood heatmap
big_top <- berghei_tpm_avg_byexpt

big_top_recast = dcast(big_top, Geneid ~ expt_stage, value.var = 'mean_tpm')
big_top_recast <- column_to_rownames(big_top_recast, var = 'Geneid')
# big_top_recast = big_top_recast[,-1,drop=FALSE]

# apply(big_top_recast, 2, sum)

if(global_save_var){
  png(file = 'plots/thesis_chapter3_figure_corrplot.png',
      height = 5,
      width = 5,
      units = "in",
      res = 330)
}
corrplot(cor(big_top_recast), method = 'color', 
         order = 'hclust', hclust.method = 'ward.D2',
         cl.lim = c(-0.036,1), tl.cex = 0.50,
         col = c(viridis(100), magma(115)), addrect = 6)
if(global_save_var){
  dev.off()
}
```

## Supp figure 1B - PCA of all the samples.

```{r supp_figure_PCA_allgenes}
#Set the Geneids as the berghei_tpm rownames so that the PCA function retains them as loading variables
big_tpm_pca <-  prcomp(column_to_rownames(berghei_tpm, var = 'Geneid'), 
                       center = TRUE, 
                       scale. = TRUE)

big_tpm_pca_var <- get_pca_var(big_tpm_pca)
big_tpm_pca_coords <- data.frame(sample = rownames(big_tpm_pca_var$coord),
                                 big_tpm_pca_var$coord)
big_tpm_pca_coords$stage_experiment <- str_extract(
  big_tpm_pca_coords$sample,
  pattern = '.+(?=_[[:digit:]]+$)'
)
big_tpm_pca_coords$stage <- str_extract(big_tpm_pca_coords$sample, 
                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
big_tpm_pca_coords$experiment <- str_extract(big_tpm_pca_coords$sample, 
                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')

colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
                               length.out = length(unique(big_tpm_pca_coords$stage))),
                       unique(big_tpm_pca_coords$stage))

pca_importance <- as.data.frame(summary(big_tpm_pca)$importance)

(
  plot <- ggplot(big_tpm_pca_coords,
       aes(
         x = Dim.1,
         y = Dim.2,
         col = stage,
         shape = experiment
       )) +
  geom_point() +
  scale_fill_manual(values = colornames) +
  scale_shape_manual(
    values = setNames(
      c(1:10),
      unique(big_tpm_pca_coords$experiment)
    )
  ) +
  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
  guides(shape=guide_legend(ncol=2),
         col = guide_legend(ncol=2)) +
  theme_classic() +
  theme(legend.key.size = unit(5, 'mm'),
        legend.text = element_text(size = 6))
)

if(global_save_var){
  ggsave('plots/thesis_chapter3_figure_pca_allGenes.png',
         width = 7,
         height = 5,
         units = "in",
         plot)
}

```

# Biplot of the PCA using all genes.

```{r biplot}
set.seed(982)

pca_object_loadings <- data.frame(big_tpm_pca$x)

#Remove the large PCA object
rm(big_tpm_pca)

#Calculate the highest loadings, only want these labels
top_loadings <- 15
pca_object_loadings$Geneid <- ''

highest_loadings <- pca_object_loadings[rev(
    order(
      apply(
        abs(
          data.frame(pca_object_loadings$PC1, 
                     pca_object_loadings$PC2)),
        1,
        sum)
    )
)[1:top_loadings],]

highest_loadings$Geneid <- 
  gsub(
    gsub(allgene.info$Description[match(rownames(highest_loadings),
                                      allgene.info$Geneid)], 
       pattern = 'putative',
       replacement = 'put.'),
    pattern = 'early transcribed membrane protein',
    replacement = 'ETRAMP')

loadings_scale <- 0.005

(bp <- plot +
    geom_point(data = highest_loadings,
               aes(x = PC1*loadings_scale, y = PC2*loadings_scale),
               inherit.aes = FALSE, size = 0.75) +
    geom_text_repel(data = highest_loadings,
                    aes(x = PC1*loadings_scale, y = PC2*loadings_scale,
                        label = Geneid),
                    inherit.aes = FALSE, 
                    size = 3,
                    min.segment.length = 0,
                    force = 100
    ) )

if(global_save_var){
  ggsave('plots/thesis_chapter3_figure_biplot.png',
         width = 7,
         height = 5,
         units = "in",
         bp)
}

#Note that CSP is the 140 highest laodings gene.
```

#Supp figure 1C - PCA using just the pir genes.

```{r supp_figure_PCA_pir}
big_tpm_pir <- berghei_tpm[berghei_tpm$Geneid %in% pir_id,-1,drop = FALSE]

big_tpm_pir <- big_tpm_pir[apply(big_tpm_pir, 2, sum) >= 50]

big_tpm_pir_pca <-  prcomp(big_tpm_pir, 
                       center = TRUE, 
                       scale. = TRUE)

big_tpm_pir_pca_var <- get_pca_var(big_tpm_pir_pca)
big_tpm_pir_pca_coords <- data.frame(sample = rownames(big_tpm_pir_pca_var$coord),
                                 big_tpm_pir_pca_var$coord)
big_tpm_pir_pca_coords$stage_experiment <- str_extract(
  big_tpm_pir_pca_coords$sample,
  pattern = '.+(?=_[[:digit:]]+$)'
)
big_tpm_pir_pca_coords$stage <- str_extract(big_tpm_pir_pca_coords$sample, 
                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
big_tpm_pir_pca_coords$experiment <- str_extract(big_tpm_pir_pca_coords$sample, 
                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')

colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
                               length.out = length(unique(big_tpm_pir_pca_coords$stage))),
                       unique(big_tpm_pir_pca_coords$stage))

pca_importance <- as.data.frame(summary(big_tpm_pir_pca)$importance)

(
  plot_pir <- ggplot(big_tpm_pir_pca_coords,
                     aes(
                       x = Dim.1,
                       y = Dim.2,
                       col = stage,
                       shape = experiment
                     )) +
    geom_point() +
    scale_fill_manual(values = colornames) +
    scale_shape_manual(
      values = setNames(
        c(1:10),
        unique(big_tpm_pca_coords$experiment)
      )
    ) +
    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
    guides(shape=guide_legend(ncol=2),
           col = guide_legend(ncol=2)) +
    theme_classic() +
    theme(legend.key.size = unit(5, 'mm'),
          legend.text = element_text(size = 6)) 
)

if(global_save_var){
    ggsave('plots/thesis_chapter3_figure_PCA_pir.png',
         width = 7,
         height = 5,
         units = "in",
         plot_pir)
}
```

### 2021-09-09 rfunctions/density_samples_fun.R

@@ -1,0 +1,58 @@
+density_samples <- function(input, 
+                            save = FALSE,
+                            line_val = 1){
+  #Get some colours for the lines
+  colours = rep_len(cbPalette, length.out = length(colnames(input$tpm)[-1]))
+  list_dens = lapply(colnames(input$tpm)[-1], function(sample){
+    data = as.numeric(unlist(select(input$tpm, paste(sample))))
+    dens = density(log1p(data), bw = bw_mean)
+    #Use the index of the column to get the colour
+    colouring = colours[grep(colnames(input$tpm)[-1], pattern = paste(sample))]
+    list_return = list(dens, colouring)
+    names(list_return) = c('dens','colouring')
+    list_return
+  })
+  max_dens = max(unlist(lapply(list_dens, function(x){max(x$dens$y)})))
+  max_x = max(unlist(lapply(list_dens, function(x){max(x$dens$x)})))
+  line_width = 1
+  par(new=FALSE)
+  if(save){
+    png(file = paste0('plots/thesis_chapter3_figure_',
+                      input$experiment, 
+                      '_densityplot_fixedbw.png'), 
+        width = 15/2.54, 
+        height = 11.5/2.54,
+        units = "in",
+        res = 330)
+  }
+  lapply(seq_along(list_dens), function(index){
+    x = list_dens[[index]]
+    #Change lty value every 9 samples (i.e. when the colours reset)
+    # %/% integer division gives the integer result, e.g. 8%/%9 = 0, 9%/%9 = 1, 18%/%9 = 2.
+    lty <- ((index-1L) %/% 9L) + 1L
+    plot(x$dens, 
+         xlim = c(0, max_x), 
+         ylim = c(0,max_dens), 
+         type = 'l', 
+         col = x$colouring,
+         main = paste(input$experiment),
+         lwd = line_width,
+         lty = lty)
+    if(is.numeric(line_val)){
+      abline(v = log1p(line_val))
+    }
+    legend(max_x-4, 
+           max_dens, 
+           legend = paste(colnames(input$tpm)[-1]), 
+           col = colours,
+           lty = c(rep(1, times = 9), 
+                   rep(2, times = 9), 
+                   rep(3, times = 9)),
+           cex = 0.5,
+           ncol = ifelse(length(list_dens) > 15, 2, 1),
+           xjust = ifelse(length(list_dens) > 15, 0.25, 0))
+    par(new=TRUE)})
+  if(save){
+    dev.off()
+  }
+}

### 2021-09-09 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
renv and remove expt_tpms

@@
                         row.names = 1)
-```
-
-```{r expt_tpms}
-#Get a list of df for each experiment.
-
-expt_nams <- unique(
-  na.omit(
-    str_extract(colnames(berghei_tpm), 
-                pattern = 'E[[:digit:]]{2}')
-  )
-)
-
-expt_tpms <- lapply(
-  expt_nams,
-  function(expt_nam){
-    expt_df <- select(
-      berghei_tpm,
-      matches(paste0('Geneid|', expt_nam))
-    ) %>% 
-      rename_with(~ str_remove(.x, 'E[[:digit:]]{2}_'),
-                  matches('E[[:digit:]]{2}'))
-    list(tpm = expt_df,
-         experiment = expt_nam)
-  }
-)
-names(expt_tpms) <- expt_nams
 ```

### 2021-09-09 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
Fix bug with newer ggplot - revert to 3.3.3, renv snapshot update

@@
 library('plotly')
-library('ggpattern')
 library('ComplexHeatmap')
-library('ggrepel')
 
@@
 fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') #Note that this order is influenced by sort(col_group)
+```
+
+```{r draw_stacked_bar_chart}
+# Note that this code does not work with ggplot 3.3.5, but it does work with 3.3.3. 
+#  Some issue with the colours for the geom_rect background.
 
@@
     ) 
+y
+```
+
+```{r change_facet_cols}
 

### 2021-09-09 chapter3_results_bergheiPir/Pberghei_pir_samples_hclust-PCA-topGenes.Rmd
Remove unneeded package loads, get packages to work with ggplot2 3.3.3

@@
 library('dendextend') # for dendrogram plotting
-library('plotly')
-library('ggpattern')
-library('ComplexHeatmap')
 library('ggrepel')

### 2021-09-13 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
Remove unneeded package loads, get packages to work with ggplot2 3.3.3
Add error bars to smaller inset of stacked bar chart

--- 2021-09-09 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
@@
 
+global_save_var <- FALSE
+
 ```
@@
                  'Liv-24h' ,'Liv-36h','48h','54h','60h',
-                 'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')
+                 'Mero','Asex','Ring','Troph','Schiz',
+                 'Gam','Female','Male',
+                 '-Pre3h','BldMl-24h','Ook')
 
@@
                              pattern = '(?<=_).+(?=_)')) %>% 
-  group_by(Geneid, stage) %>% 
+  group_by(Geneid, 
+           stage) %>% 
   dplyr::summarise(tpm_mean = mean(tpm),
@@
 ```{r SorL_stacked_eachGene_fig2a}
-  tpm_data <- combined_data %>% 
-    rownames_to_column(var = 'Geneid') %>% 
-    filter(Geneid %in% pir_id) %>% 
-    melt(id.vars = 'Geneid', 
-         variable.name = 'stage', 
-         value.name = 'tpm') %>% 
+melted_tpm <- filter(berghei_tpm, 
+                     Geneid %in% pir_id) %>% 
   mutate(clade = pir_clade$clade[match(Geneid, 
@@
   ) %>% 
+  mutate(SorL = str_extract(clade, 
+                            pattern = '((L|S)(?=[[:digit:]]))|ancestral')) %>% 
+  melt(id.vars = c('Geneid','clade', 'SorL'), 
+       variable.name = 'sample', 
+       value.name = 'tpm') %>% 
   group_by(clade) %>% 
   mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
-    mutate(stack_group = Geneid) %>% 
-    mutate(tpm = ifelse(tpm > 1, tpm, 0)) %>% #Filtering step for low value pir genes.
-    group_by(stage, 
-             stack_group,
-             clade) %>% 
-    dplyr::summarise(total_tpm = sum(tpm)) %>%
-  arrange(
-          desc(clade),
-          total_tpm
-)
-
-tpm_data <- mutate(tpm_data,
-                   order = row_number())
-
-tpm_data_SorL_stackgenes <- mutate(tpm_data,
-                        SorL = str_extract(clade, 
-                                           pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent'))
-tpm_data_SorL_stackgenes <- group_by(tpm_data_SorL_stackgenes,
+  group_by(SorL) %>% 
+  mutate(
+    SorL = paste0(SorL, 
+                  ' (n = ', 
+                  n_distinct(Geneid),
+                  ')'),
+    stage = str_extract(sample, 
+                        pattern = '(?<=_).+(?=_)'))
+
+```
+
+```{r stacked_barchart_fun}
+stacked_barchart <- function(melted_tpm,
+                             threshold = 1,
+                             stack_group_name = 'Geneid',
+                             sd_bars = FALSE){
+  #Have to use sym and !! in order to get mutate to work 
+  # (converting the stack_group_name to a symbol and then unquoting in the mutate function)
+  stack_group_name <- rlang::sym(paste(stack_group_name))
+
+  tpm_data <- mutate(melted_tpm, 
+                     stack_group = (!!stack_group_name)) %>% 
+    mutate(tpm = ifelse(tpm > threshold, tpm, 0)) %>% #Filtering step for low value pir genes.
+    group_by(stage, SorL, stack_group, sample) %>% 
+    summarise(sum_tpm = sum(tpm)) %>% 
+    group_by(SorL, stack_group, stage) %>%
+    summarise(mean_sum_tpm = mean(sum_tpm),
+              sd_sum_tpm = sd(sum_tpm)) %>% 
+    arrange(desc(SorL),
+            mean_sum_tpm) %>% 
+    ungroup() %>% 
+    mutate(order = row_number())
+
+  tpm_data_SorL_stackgenes <- 
+    #Order by SorL and then by mean_sum_tpm, then convert stack_group to a factor so ggplot orders correctly
+    arrange(tpm_data,
             stage,
-                                     SorL,
-                                     stack_group) %>% 
-  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
-  #Include the number of genes in each fill category for the labels
-  mutate(SorL = paste0(SorL, 
-                       ' (n = ', 
-                       n_distinct(stack_group),
-                       ')')) %>% 
-#Order by SorL and then by total_tpm, then convert stack_group to a factor so ggplot orders correctly
-  arrange(SorL, 
-          total_tpm) %>% 
-  mutate(stage_num = as.integer(stage))
-tpm_data_SorL_stackgenes$stage %>% as.integer %>% unique
-tpm_data_SorL_stackgenes$stage %>% unique
+            desc(SorL), 
+            mean_sum_tpm) %>% 
+    mutate(stage = factor(stage,
+                          levels = ordering(unique(tpm_data$stage),
+                                            order_vector = r_order_vector)),
+           stage_num = as.integer(
+             factor(stage,
+                    levels = ordering(unique(tpm_data$stage),
+                                      order_vector = r_order_vector))
+           )) %>% 
+    group_by(stage) %>% 
+    mutate(stacked_mean = cumsum(mean_sum_tpm),
+           sd_upper = cumsum(mean_sum_tpm) + cumsum(sd_sum_tpm),
+           sd_lower = ifelse(cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm) < 0,
+                             0, 
+                             cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm)))
 
   tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))
-```
-
-```{r stacked_bar_chart_figure2b}
-#Calculate the range for the y value to make it look nicer
+
   stacked_max <- tpm_data %>% 
@@
     group_by(stage) %>% 
-  dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
+    dplyr::summarise(max_tpm = sum(mean_sum_tpm)) %>% 
     .$max_tpm %>% 
@@
 
-write_csv(x = tpm_data_SorL_stackgenes,
-          path = 'tl20-05_SorL-subfamily_pir.csv')
-
-stage_vec <- unique(
-  levels(tpm_data$stage)
-)
-
   stage_vec <- tpm_data_SorL_stackgenes$stage
@@
   tpm_data_SorL_stackgenes$stage_col <- col_group
-fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') #Note that this order is influenced by sort(col_group)
+  fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') 
+  #Note that this order is influenced by sort(col_group)
+
+  # Note that this code does not work with ggplot 3.3.5, but it does work with 3.3.3. 
+  #  Some issue with the colours for the geom_rect background.
 
@@
                 x = stage_num,
-              y = total_tpm, 
-              fill = SorL,
-              # group = stage_num
+                y = mean_sum_tpm, 
+                fill = SorL
               )) +
@@
       ),
-    alpha = .05
+      alpha = .5
     ) +
@@
              size = 0.1) +
+    {if(sd_bars)list(
+      geom_errorbar(aes(ymax = sd_upper,
+                        ymin = sd_lower),
+                    position = position_dodge(width = .85),
+                    size = .25),
+        geom_point(aes(y = stacked_mean), 
+                   position = position_dodge(width = .85),
+                   show.legend = FALSE,
+                   size = .25)
+    )}+ 
     scale_fill_manual(
@@
     scale_color_manual('black') + 
-  # labs(fill = stack_label) +
     scale_y_continuous(breaks = max_tpm) +
     theme_classic() +
-  # scale_size(range = c(0,3)) +
     ylab(
@@
       axis.ticks.x = element_blank(),
-      # axis.line.x = element_blank(),
       axis.text.y = element_text(size = 10),
@@
   strip_t <- which(grepl('strip-t', g$layout$name))
-stage_vec <- unique(
-  levels(tpm_data$stage)
-)
-fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
-                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
-                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
-                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
+  stage_vec_strip <- sort(unique(tpm_data_SorL_stackgenes$stage))
+  fills <- case_when(str_detect(stage_vec_strip, pattern = 'Asex') ~ 'coral1' ,
+                     str_detect(stage_vec_strip, pattern = 'Liv') ~ 'plum2',
+                     str_detect(stage_vec_strip, pattern = 'Gam') ~ 'lightskyblue1',
+                     str_detect(stage_vec_strip, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
   k <- 1
@@
   grid.draw(g)
+  return(g)
+}
+```
+
+```{r stacked_barchart}
+sd_bars <- TRUE
+g <- stacked_barchart(melted_tpm,
+                      stack_group_name = 'SorL',
+                      sd_bars = sd_bars)
+
+if(global_save_var){
+ggsave('plots/thesis_chapter3_figure_stackedBarChartTPM_SorL_standevError.png',
+       width = 7,
+       height = 5,
+       units = "in",
+       g)
+}
 
@@
   group_by(stage) %>% 
-  summarise(sum_tpm = sum(total_tpm)) %>% 
+  summarise(sum_tpm = sum(mean_sum_tpm)) %>% 
   filter(sum_tpm < 350)
@@
               aes(
-              x = stage,
-              y = total_tpm, 
-              fill = SorL,
-              # group = stage_num
+                x = stage_num,
+                y = mean_sum_tpm, 
+                fill = SorL
               )) +
@@
     scale_color_manual('black') + 
-  labs(fill = stack_label) +
     scale_y_continuous(breaks = max_tpm) +
     theme_classic() +
-  scale_size(range = c(0,3)) +
     ylab(
@@
     ) +
+    scale_x_continuous(
+      breaks = tpm_data_SorL_stackgenes$stage_num %>% unique(),
+      labels = tpm_data_SorL_stackgenes$stage %>% unique(),
+      expand = c(0, 0)
+    ) +
+    theme(
+      axis.text.x = element_text(angle = 90, 
+                                 vjust = .5,
+                                 size = 10),
+      axis.text.y = element_text(size = 10),
+      axis.title.y = element_text(size = 10),
+      legend.title = element_text(size = 12),
+      legend.text = element_text(size = 10),
+      legend.key.size = unit(3,'mm'),
+      legend.key.width = unit(2,'mm'),
+    ) +
+    xlab('stage') + 
     facet_grid(.~stage, 
@@
       axis.ticks.x = element_blank(),
-    # axis.line.x = element_blank(),
-    axis.text.y = element_text(size = 15),
-    axis.title.y = element_text(size = 8),
-    legend.title = element_text(size = 7),
-    legend.text = element_text(size = 5),
+      axis.text.y = element_text(size = 10),
+      axis.title.y = element_text(size = 10),
+      legend.title = element_text(size = 12),
+      legend.text = element_text(size = 10),
       legend.key.size = unit(3,'mm'),
       legend.key.width = unit(2,'mm'),
-    strip.text.x = element_text(size=6, 
+      strip.text.x = element_text(size=8, 
                                   face = 'bold', 

### 2021-09-13 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
source stacked_barchart function

@@
 source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
+source('../rfunctions/stacked_barchart_fun.R')
 ```
@@
 
-```
-
-```{r stacked_barchart_fun}
-stacked_barchart <- function(melted_tpm,
-                             threshold = 1,
-                             stack_group_name = 'Geneid',
-                             sd_bars = FALSE){
-  #Have to use sym and !! in order to get mutate to work 
-  # (converting the stack_group_name to a symbol and then unquoting in the mutate function)
-  stack_group_name <- rlang::sym(paste(stack_group_name))
-
-  tpm_data <- mutate(melted_tpm, 
-                     stack_group = (!!stack_group_name)) %>% 
-    mutate(tpm = ifelse(tpm > threshold, tpm, 0)) %>% #Filtering step for low value pir genes.
-    group_by(stage, SorL, stack_group, sample) %>% 
-    summarise(sum_tpm = sum(tpm)) %>% 
-    group_by(SorL, stack_group, stage) %>%
-    summarise(mean_sum_tpm = mean(sum_tpm),
-              sd_sum_tpm = sd(sum_tpm)) %>% 
-    arrange(desc(SorL),
-            mean_sum_tpm) %>% 
-    ungroup() %>% 
-    mutate(order = row_number())
-
-  tpm_data_SorL_stackgenes <- 
-    #Order by SorL and then by mean_sum_tpm, then convert stack_group to a factor so ggplot orders correctly
-    arrange(tpm_data,
-            stage,
-            desc(SorL), 
-            mean_sum_tpm) %>% 
-    mutate(stage = factor(stage,
-                          levels = ordering(unique(tpm_data$stage),
-                                            order_vector = r_order_vector)),
-           stage_num = as.integer(
-             factor(stage,
-                    levels = ordering(unique(tpm_data$stage),
-                                      order_vector = r_order_vector))
-           )) %>% 
-    group_by(stage) %>% 
-    mutate(stacked_mean = cumsum(mean_sum_tpm),
-           sd_upper = cumsum(mean_sum_tpm) + cumsum(sd_sum_tpm),
-           sd_lower = ifelse(cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm) < 0,
-                             0, 
-                             cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm)))
-
-  tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))
-
-  stacked_max <- tpm_data %>% 
-    ungroup %>% 
-    group_by(stage) %>% 
-    dplyr::summarise(max_tpm = sum(mean_sum_tpm)) %>% 
-    .$max_tpm %>% 
-    max
-
-  max_tpm <- c(
-    seq(
-      0,
-      mround(
-        stacked_max,
-        500),
-      length.out = 11)
-  )
-
-  stack_label <- 'sub-family'
-
-  stage_vec <- tpm_data_SorL_stackgenes$stage
-  col_group <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'zzAsex' ,
-                         str_detect(stage_vec, pattern = 'Liv') ~ 'zzLiv',
-                         str_detect(stage_vec, pattern = 'Gam') ~ 'zzGam',
-                         str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'zzMos')
-  tpm_data_SorL_stackgenes$stage_col <- col_group
-  fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') 
-  #Note that this order is influenced by sort(col_group)
-
-  # Note that this code does not work with ggplot 3.3.5, but it does work with 3.3.3. 
-  #  Some issue with the colours for the geom_rect background.
-
-  y <- ggplot(data = tpm_data_SorL_stackgenes, 
-              aes(
-                x = stage_num,
-                y = mean_sum_tpm, 
-                fill = SorL
-              )) +
-    geom_rect(
-      aes(
-        xmin = stage_num - .7,
-        xmax = stage_num + .7,
-        ymin = -Inf,
-        ymax = Inf,
-        fill = stage_col
-      ),
-      alpha = .5
-    ) +
-    geom_bar(position = position_stack(),
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-    {if(sd_bars)list(
-      geom_errorbar(aes(ymax = sd_upper,
-                        ymin = sd_lower),
-                    position = position_dodge(width = .85),
-                    size = .25),
-        geom_point(aes(y = stacked_mean), 
-                   position = position_dodge(width = .85),
-                   show.legend = FALSE,
-                   size = .25)
-    )}+ 
-    scale_fill_manual(
-      breaks = unique(tpm_data_SorL_stackgenes$SorL),
-      values = c(
-        SorL_col_num,
-        fills
-      )
-    ) +
-    scale_color_manual('black') + 
-    scale_y_continuous(breaks = max_tpm) +
-    theme_classic() +
-    ylab(
-      bquote('total '~italic(.('pir'))~' TPM')
-    ) +
-    scale_x_continuous(
-      breaks = tpm_data_SorL_stackgenes$stage_num %>% unique(),
-      labels = tpm_data_SorL_stackgenes$stage %>% unique(),
-      expand = c(0, 0)
-    ) +
-    theme(
-      axis.text.x = element_text(angle = 90, 
-                                 vjust = .5,
-                                 size = 10),
-      axis.text.y = element_text(size = 10),
-      axis.title.y = element_text(size = 10),
-      legend.title = element_text(size = 12),
-      legend.text = element_text(size = 10),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-    ) +
-    xlab('stage') + 
-    facet_grid(.~stage, 
-               scales = "free_x",
-               space = "free_x") +
-    scale_x_discrete(expand=c(0,0)) +
-    theme(
-      axis.title.x=element_blank(),
-      axis.text.x=element_blank(),
-      axis.ticks.x = element_blank(),
-      axis.text.y = element_text(size = 10),
-      axis.title.y = element_text(size = 10),
-      legend.title = element_text(size = 12),
-      legend.text = element_text(size = 10),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=8, 
-                                  face = 'bold', 
-                                  angle = 90),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      ),
-      panel.spacing.x = unit(0, "null")
-    )
-
-  g <- ggplot_gtable(ggplot_build(y))
-  strip_t <- which(grepl('strip-t', g$layout$name))
-  stage_vec_strip <- sort(unique(tpm_data_SorL_stackgenes$stage))
-  fills <- case_when(str_detect(stage_vec_strip, pattern = 'Asex') ~ 'coral1' ,
-                     str_detect(stage_vec_strip, pattern = 'Liv') ~ 'plum2',
-                     str_detect(stage_vec_strip, pattern = 'Gam') ~ 'lightskyblue1',
-                     str_detect(stage_vec_strip, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
-  k <- 1
-  for (i in strip_t) {
-    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-    k <- k+1
-  }
-
-  grid.newpage()
-  grid.draw(g)
-  return(g)
-}
 ```

### 2021-09-13 rfunctions/stacked_barchart_fun.R

@@ -1,0 +1,177 @@
+stacked_barchart <- function(melted_tpm,
+                             threshold = 1,
+                             stack_group_name = 'Geneid',
+                             sd_bars = FALSE){
+  #Have to use sym and !! in order to get mutate to work 
+  # (converting the stack_group_name to a symbol and then unquoting in the mutate function)
+  stack_group_name <- rlang::sym(paste(stack_group_name))
+  
+  tpm_data <- mutate(melted_tpm, 
+                     stack_group = (!!stack_group_name)) %>% 
+    mutate(tpm = ifelse(tpm > threshold, tpm, 0)) %>% #Filtering step for low value pir genes.
+    group_by(stage, SorL, stack_group, sample) %>% 
+    summarise(sum_tpm = sum(tpm)) %>% 
+    group_by(SorL, stack_group, stage) %>%
+    summarise(mean_sum_tpm = mean(sum_tpm),
+              sd_sum_tpm = sd(sum_tpm)) %>% 
+    arrange(desc(SorL),
+            mean_sum_tpm) %>% 
+    ungroup() %>% 
+    mutate(order = row_number())
+  
+  tpm_data_SorL_stackgenes <- 
+    #Order by SorL and then by mean_sum_tpm, then convert stack_group to a factor so ggplot orders correctly
+    arrange(tpm_data,
+            stage,
+            desc(SorL), 
+            mean_sum_tpm) %>% 
+    mutate(stage = factor(stage,
+                          levels = ordering(unique(tpm_data$stage),
+                                            order_vector = r_order_vector)),
+           stage_num = as.integer(
+             factor(stage,
+                    levels = ordering(unique(tpm_data$stage),
+                                      order_vector = r_order_vector))
+           )) %>% 
+    group_by(stage) %>% 
+    mutate(stacked_mean = cumsum(mean_sum_tpm),
+           sd_upper = cumsum(mean_sum_tpm) + cumsum(sd_sum_tpm),
+           sd_lower = ifelse(cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm) < 0,
+                             0, 
+                             cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm)))
+  
+  tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))
+  
+  stacked_max <- tpm_data %>% 
[…]

### 2021-09-13 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
expand stacked_barchart functionality

@@
 library('ComplexHeatmap')
+library('kableExtra')
 
@@
 sd_bars <- TRUE
-g <- stacked_barchart(melted_tpm,
+g_list <- stacked_barchart(melted_tpm,
                            stack_group_name = 'SorL',
-                      sd_bars = sd_bars)
+                           sd_bars = sd_bars,
+                           inset = TRUE)
 
@@
          units = "in",
-       g)
+         g_list$main)
+  ggsave('plots/thesis_chapter3_figure_stackedBarChartTPM_SorL_standevError_inset.png',
+         width = 7,
+         height = 5,
+         units = "in",
+         g_list$inset)
 }
@@
 
-# figure2binset
+```{r table_of_SorL_highest_supptable}
+tpm_data_SorL_stackgenes <- g_list$data
 
-```{r stack_bar_graph_SorL_lowexpression-figure2binset}
-#Include all stages expressed lower than a certain amount
-
-low_express <- tpm_data_SorL_stackgenes %>% 
-  group_by(stage) %>% 
-  summarise(sum_tpm = sum(mean_sum_tpm)) %>% 
-  filter(sum_tpm < 350)
-
-#Calculate the range for the y value to make it look nicer
-
-stacked_max <- max(low_express$sum_tpm)
-
-max_tpm <- c(
-  seq(
-    0,
-    mround(
-      stacked_max,
-      100),
-    length.out = 11)
-)
-
-tpm_data_SorL_stackgenes_low <- filter(tpm_data_SorL_stackgenes,
-                                       stage %in% low_express$stage)
-
-y <- ggplot(data = tpm_data_SorL_stackgenes_low, 
-              aes(
-                x = stage_num,
-                y = mean_sum_tpm, 
-                fill = SorL
-              )) +
-    geom_bar(position = position_stack(),
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-    {if(sd_bars)list(
-      geom_errorbar(aes(ymax = sd_upper,
-                        ymin = sd_lower),
-                    position = position_dodge(width = .85),
-                    size = .25),
-        geom_point(aes(y = stacked_mean), 
-                   position = position_dodge(width = .85),
-                   show.legend = FALSE,
-                   size = .25)
-    )}+ 
-    scale_fill_manual(
-      breaks = unique(tpm_data_SorL_stackgenes$SorL),
-      values = c(
-        SorL_col_num,
-        fills
-      )
-    ) +
-    scale_color_manual('black') + 
-    scale_y_continuous(breaks = max_tpm) +
-    theme_classic() +
-    ylab(
-      bquote('total '~italic(.('pir'))~' TPM')
-    ) +
-    scale_x_continuous(
-      breaks = tpm_data_SorL_stackgenes$stage_num %>% unique(),
-      labels = tpm_data_SorL_stackgenes$stage %>% unique(),
-      expand = c(0, 0)
-    ) +
-    theme(
-      axis.text.x = element_text(angle = 90, 
-                                 vjust = .5,
-                                 size = 10),
-      axis.text.y = element_text(size = 10),
-      axis.title.y = element_text(size = 10),
-      legend.title = element_text(size = 12),
-      legend.text = element_text(size = 10),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-    ) +
-    xlab('stage') + 
-    facet_grid(.~stage, 
-               scales = "free_x",
-               space = "free_x") +
-      scale_x_discrete(expand=expand_scale(add=1)) +
-    theme(
-      axis.title.x=element_blank(),
-      axis.text.x=element_blank(),
-      axis.ticks.x = element_blank(),
-      axis.text.y = element_text(size = 10),
-      axis.title.y = element_text(size = 10),
-      legend.title = element_text(size = 12),
-      legend.text = element_text(size = 10),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=8, 
-                                  face = 'bold', 
-                                  angle = 90),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      ),
-      panel.spacing.x = unit(0, "null")
-    )
-
-#This bit is to change the colour of the facet boxes
-
-g <- ggplot_gtable(ggplot_build(y))
-strip_t <- which(grepl('strip-t', g$layout$name))
-stage_vec <- unique(
-  levels(tpm_data_SorL_stackgenes_low$stage)[levels(tpm_data_SorL_stackgenes_low$stage) %in% as.character(tpm_data_SorL_stackgenes_low$stage)]
-)
-fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
-                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
-                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
-                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
-k <- 1
-for (i in strip_t) {
-  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-  k <- k+1
-}
-grid.newpage()
-grid.draw(g)
-```
-
-```{r table_of_SorL_highest_supptable}
 #For a neat table
@@
   group_by(stage) %>% 
-  mutate(sum_tpm = sum(total_tpm)) %>% 
+  mutate(sum_tpm = sum(mean_sum_tpm)) %>% 
   group_by(stage, 
            SorL) %>% 
-  slice_max(order_by = total_tpm,
+  slice_max(order_by = mean_sum_tpm,
             n = 1) %>% 
-  mutate(total_tpm = round(total_tpm, digits = 2),
-         prop_tpm = round(total_tpm/sum_tpm, digits = 2)) %>% 
-  select(stage, Geneid, SorL, total_tpm, prop_tpm)
+  mutate(mean_sum_tpm = round(mean_sum_tpm, digits = 2),
+         prop_tpm = round(mean_sum_tpm/sum_tpm, digits = 2)) %>% 
+  select(stage, Geneid, SorL, mean_sum_tpm, prop_tpm)
 

### 2021-09-13 rfunctions/stacked_barchart_fun.R

@@ -1,7 +1,8 @@
 stacked_barchart <- function(melted_tpm,
                              threshold = 1,
                              stack_group_name = 'Geneid',
-                             sd_bars = FALSE){
+                             sd_bars = FALSE,
+                             inset = TRUE){
   #Have to use sym and !! in order to get mutate to work 
   # (converting the stack_group_name to a symbol and then unquoting in the mutate function)
   stack_group_name <- rlang::sym(paste(stack_group_name))
@@ -107,7 +108,8 @@ stacked_barchart <- function(melted_tpm,
       values = c(
         SorL_col_num,
         fills
-      )
+      ),
+      name = 'Sub-family'
     ) +
     scale_color_manual('black') + 
     scale_y_continuous(breaks = max_tpm) +
@@ -173,5 +175,133 @@ stacked_barchart <- function(melted_tpm,
   
   grid.newpage()
   grid.draw(g)
-  return(g)
+  
+  #Draw inset of the smaller value samples
+  if(inset){
+    #Include all stages expressed lower than a certain amount
+    
+    low_express <- tpm_data_SorL_stackgenes %>% 
+      group_by(stage) %>% 
+      summarise(sum_tpm = sum(mean_sum_tpm),
+                max_stan_dev = max(sd_upper)) %>% 
+      filter(sum_tpm < 350)
+    
+    #Calculate the range for the y value to make it look nicer
+    
+    stacked_max <- max(low_express$max_stan_dev)
+    
+    max_tpm <- c(
+      seq(
+        0,
+        mround(
+          stacked_max,
+          100),
+        length.out = 11)
+    )
+    
+    tpm_data_SorL_stackgenes_low <- filter(tpm_data_SorL_stackgenes,
+                                           stage %in% low_express$stage)
+    
+    y <- ggplot(data = tpm_data_SorL_stackgenes_low, 
+                aes(
+                  x = stage_num,
+                  y = mean_sum_tpm, 
+                  fill = SorL
+                )) +
+      geom_bar(position = position_stack(),
+               stat = 'identity',
+               colour = 'black',
+               size = 0.1) +
+      {if(sd_bars)list(
+        geom_errorbar(aes(ymax = sd_upper,
+                          ymin = sd_lower),
+                      position = position_dodge(width = .85),
+                      size = .25),
+        geom_point(aes(y = stacked_mean), 
[…]

### 2021-09-14 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
change pir proportion Pberg graph by facetting for sub-fam and including genome composition

@@
   mutate(clade = pir_clade$clade[match(Geneid, 
+                                       pir_clade$Geneid)],
+         SorL = pir_clade$SorL[match(Geneid, 
                                        pir_clade$Geneid)]
   ) %>% 
-  group_by(clade) %>% 
+  group_by(SorL, clade) %>% 
   mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
-    mutate(stack_group = clade) %>% 
   group_by(stage,
-             stack_group,
+           SorL,
            clade) %>% 
   dplyr::summarise(total_tpm = sum(tpm)) %>%
+  group_by(stage) %>% 
+  mutate(prop_tpm = total_tpm/sum(total_tpm)) %>% 
   arrange(
@@
 
-tpm_data_subfam <- mutate(tpm_data_subfam,
-                   order = row_number())
+clade_composition <- pir_clade %>% 
+  group_by(SorL, clade) %>% 
+  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
+  summarise(prop_tpm = n_distinct(Geneid)/nrow(.)) %>% 
+  mutate(stage = "Genome") %>% 
+  melt(id.vars = c('stage', 'SorL', 'clade'), 
+       value.name = 'prop_tpm') %>% 
+  select(-variable)
 
-  # #Apply ordering
-  # tpm_data_subfam$stage = factor(tpm_data_subfam$stage, 
-  #                         levels = ordering(tpm_data_subfam$stage,
-  #                                           order_vector))
+tpm_data_subfam <- bind_rows(tpm_data_subfam, clade_composition) %>% 
+  mutate(stage = factor(stage, 
+                        levels = ordering(unique(c(as.character(tpm_data_subfam$stage), 
+                                                   clade_composition$stage)), 
+                                          order_vector = c(r_order_vector, 
+                                                           'composition'))))
 
@@
   group_by(stage) %>% 
-    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
+  dplyr::summarise(max_tpm = sum(prop_tpm)) %>% 
   .$max_tpm %>% 
@@
 )
-```
-
-```{r combined_tpm_stacked_plot}
 
@@
               x = stage,
-              y = total_tpm, 
-              fill = clade,
-              group = order
+              y = prop_tpm, 
+              fill = clade
             )) +
   geom_bar(
-    # position = position_stack(),
-    position = 'fill',
+    position = position_stack(),
+    # position = 'fill',
     stat = 'identity',
@@
   labs(fill = stack_label) +
-  scale_y_continuous(breaks = seq(0,1,0.1)) +
+  scale_y_continuous(breaks = seq(0,1,0.2)) +
   theme_classic() +
@@
   ) + 
-  facet_grid(.~stage, 
+  facet_grid(cols = vars(stage),
+             rows = vars(SorL),
              scales = "free_x",

### 2021-09-14 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_CalculateTPM.Rmd
Split PCA/hclust/topGenes file
berg: stats comparison mero-asex-malegam heatmap and save proportional bar chart

@@
 
+Need to look more closely as to what I should and shouldn't remove from the hclust.
+E03_Asex.Schiz_3 looks like a well clustering sample, some of the 54 and 60h Liver samples do not look very similar to one another.
+
 ```{r remove_badly_clustering_samples}
@@
 ```
+
+## Saving the counts with new names/samples removed
+
+```{r save_new_counts}
+berghei_counts[,str_subset(colnames(berghei_counts), 'E06_Liv.6h')] <- NULL
+
+col_stage <- colnames(berghei_counts)
+
+#Combine all Liver samples 4h or before into LiverPre4h
+col_stage <- str_replace_all(col_stage,
+                             pattern = 'Liv\\.(2|4)h_[[:digit:]]+',
+                             replacement = 'Liv.pre4h')
+#Combine all Liver samples 48h or onwards into LiverPost48h
+col_stage <- str_replace_all(col_stage,
+                             pattern = 'Liv\\.(4[8-9]|[5-9][0-9])h_[[:digit:]]+',
+                             replacement = 'Liv.post48h')
+#Combine all blood meal samples 3h or before into BldMlPre3h
+col_stage <- str_replace_all(col_stage,
+                             pattern = '(1h|3h)_[[:digit:]]+', 
+                             replacement = 'Pre3h')
+#Uniquify the name names
+col_stage[grep(col_stage, 
+               perl = TRUE,
+               invert = TRUE,
+               pattern = '_[[:digit:]]{1,}$')] <- uniquify(col_stage[grep(col_stage, 
+                                                                          perl = TRUE,
+                                                                          invert = TRUE,
+                                                                          pattern = '_[[:digit:]]{1,}$')])
+
+colnames(berghei_counts) <- col_stage
+
+berghei_counts$`E08_Liv.pre4h_1` <- NULL
+berghei_counts$`E03_Asex.Schiz_1` <- NULL
+berghei_counts$`E03_Asex.Schiz_3` <- NULL
+
+write.csv(berghei_counts,
+          'berghei_counts_newNames.csv')
+```

### 2021-09-14 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
berg: stats comparison mero-asex-malegam heatmap and save proportional bar chart

@@
 ```{r read_count_data}
-
-berg_tpm <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
-                      sheet = 'berg_TPM_allExpts_bioReps') %>% 
-  dplyr::rename_with(.fn = ~ str_replace(.x, 
-                                         pattern = 'Expt', 
-                                         replacement = 'E'))
+berg_tpm <- read.csv('berghei_counts_newNames.csv',
+                     row.names = 1)
 
@@
 
-berg_tpm <- column_to_rownames(berg_tpm, var = 'Geneid') %>% 
-  select(!matches('BldMl|Gam_'))
+berg_tpm <-  select(berg_tpm, 
+                    !matches('BldMl|Gam_')) %>% 
+  column_to_rownames(var = 'Geneid')
 
@@
                       sheet = 1)
+
+pir_info <- read.table('../standard_files/Pberghei_pir_clades.txt',
+                    header = TRUE)
 
@@
 
+lapply(post48livvearlyliv, function(x) `$`(x,qstable))
 ```
@@
   list(pir_geneset,
-       ShortLong_list), 
+       ShortLong_list,
+       subfam_list), 
   function(gene_list){
@@
 )
-
+lapply(livmero_livpost48.results, function(x) `$`(x,qstable))
 ```
@@
 
-```{r}
+```{r maleGam_v_Asex.Mix}
 maleGam_v_Asex.Mix.results <- lapply(
@@
 
-```{r}
+```{r maleGam_v_Liv.Meroz}
 maleGam_v_Liv.Meroz.results <- lapply(
@@
 
-```{r}
+```{r Asex.Mix_v_Liv.Meroz}
 Asex.Mix_v_Liv.Meroz.results <- lapply(
@@
                                 n.points = 2^18)
-
-    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
-    #                                 FDR < 0.05)
 
@@
 
-```{r}
+```{r fdr}
 berghei_comparison_stats_df <- bind_rows(
@@
   mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
-  filter(FDR < 0.05) %>% 
   arrange(contrast)
+```
 
+```{r}
 max_logfold <- ceiling(max(max(berghei_comparison_stats_df$log.fold.change), 
@@
 (
-x <- ggplot(berghei_comparison_stats_df, 
+  x <- ggplot(filter(berghei_comparison_stats_df,
+                     FDR < 0.05), 
               aes(x = contrast, 
@@
 
+```{r mero_asex_malegam}
+mero_asex_malegam_stats_df <- filter(berghei_comparison_stats_df,
+                                     contrast %in% c('Asex.Mixed-Liv.Meroz',
+                                                     'Gam.Male-Asex.Mixed',
+                                                     'Gam.Male-Liv.Meroz'))
+
+order_path.names <- ordering(unique(mero_asex_malegam_stats_df$pathway.name), 
+                             order_vector = c('pir', 
+                                              'L', 'L[[:digit:]]', 
+                                              'S', 'S[[:digit:]]'))
+
+mero_asex_malegam_stats_df <- mero_asex_malegam_stats_df[order(match(mero_asex_malegam_stats_df$pathway.name,order_path.names)),]
+
+max_logfold <- ceiling(max(max(mero_asex_malegam_stats_df$log.fold.change), 
+                           abs(min(mero_asex_malegam_stats_df$log.fold.change))))
+
+mero_asex_malegam_stats_df_cast <- dcast(mero_asex_malegam_stats_df,
+      contrast ~ pathway.name,
+      value.var = 'log.fold.change') %>% 
+  column_to_rownames(var = 'contrast')
+
+mero_asex_malegam_stats_df_cast <- mero_asex_malegam_stats_df_cast[,order_path.names]
+
+Heatmap(mero_asex_malegam_stats_df_cast,
+        name = 'logFoldChange',
+        cluster_rows = FALSE,
+        cluster_columns = FALSE,
+        rect_gp = gpar(col = "white", lwd = 2),
+        cell_fun = function(j, i, x, y, width, height, fill){
+          fdr_val <- mero_asex_malegam_stats_df[
+            mero_asex_malegam_stats_df$contrast == rownames(mero_asex_malegam_stats_df_cast[i,]),
+          ]$FDR[j]
+          if(fdr_val<= 0.05){
+            grid.text(
+              sprintf("%.3f", #One significant figure
+                      fdr_val),
+              x,
+              y,
+              gp = gpar(fontsize = 10)
+            )
+          }
+        })
+
+```
+
 ```{r save_xlsx}

### 2021-09-14 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Use TPM for qusagegit push

@@
 
-```{r read_count_data}
-berg_tpm <- read.csv('berghei_counts_newNames.csv',
+```{r read_tpm_data}
+berg_tpm <- read.csv('berghei_tpm.csv',
                      row.names = 1)
@@
 ```{r generate_genesets}
-pir_info <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
-                      sheet = 1)
-
 pir_info <- read.table('../standard_files/Pberghei_pir_clades.txt',
@@
 
+lapply(maleVfemale, getElement, 'qstable')
 ```
 
-```{r lateliverstages}
+```{r post48livvearlyliv}
 
@@
 
-```{r}
+```{r livmero_livpost48}
 livmero_livpost48.results <- lapply(
@@
 
-```{r}
+```{r asexRing_asexMix}
 asexRing_asexMix.results <- lapply(
@@
           ]$FDR[j]
-          if(fdr_val<= 0.05){
+          if(fdr_val<= 0.1){
             grid.text(
@@
               y,
-              gp = gpar(fontsize = 10)
+              gp = gpar(fontsize = 10,
+                        fontface = 'bold')
             )

### 2021-09-15 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
berg: stats comparison mero-asex-malegam heatmap and save proportional bar chart
Fix stacked_barchart_fun standev error bars and stat comp heatmaps

@@
 sd_bars <- TRUE
-g_list <- stacked_barchart(melted_tpm,
+g_list <- stacked_barchart(melted_tpm = melted_tpm,
                            stack_group_name = 'SorL',
@@
 grid.draw(g)
+
+ggsave('plots/thesis_chapter3_figure_propBarChart_subFam.png',
+       width = 7,
+       height = 5,
+       units = "in",
+       g)
 ```

### 2021-09-15 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Fix stacked_barchart_fun standev error bars and stat comp heatmaps

@@
 lapply for each stage with the 'labels' set as that stage and the rest all set to 'other'?
-
-```{r running_qusage_malevfemale}
-maleVfemale <- lapply(
-  list(pir_geneset), 
-  function(gene_list){
-    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
-                                lifeCycle_labels,
-                                'Gam.Male-Gam.Female',
-                                gene_list,
-                                n.points = 2^18)
-
-    list(results = qs.results.subfam, 
-         qstable = qsTable(qs.results.subfam) %>% 
-           mutate(contrast = qs.results.subfam$contrast))
-  }
-)
-
-lapply(maleVfemale, getElement, 'qstable')
-```
 
@@
 
-```{r}
-max_logfold <- ceiling(max(max(berghei_comparison_stats_df$log.fold.change), 
-                           abs(min(berghei_comparison_stats_df$log.fold.change))))
-
-(
-  x <- ggplot(filter(berghei_comparison_stats_df,
-                     FDR < 0.05), 
-              aes(x = contrast, 
-                  y = reorder(pathway.name, dplyr::desc(pathway.name)))) +
-    xlab('Comparison') + ylab('Geneset') + 
-    # scale_x_discrete(labels= c(1,2,3,4,6)) +
-    geom_count(aes(
-      colour = log.fold.change), 
-      na.rm = TRUE) +
-    labs(colour = 'log2 Fold Change') +
-    scale_color_gradientn(
-      colours = rev(brewer.pal(11, "RdBu")),
-      na.value="transparent",
-      limits = c(-max_logfold,max_logfold),
-      # limits = c(-0.5,1.25)
-    ) +
-    scale_size_area(max_size = 6) +
-    theme_bw() +
-    theme(panel.grid.minor = element_blank(),
-          panel.background = element_rect(fill = 'white'),
-          plot.background = element_blank(),
-          legend.background = element_rect(fill = 'transparent'),
-          axis.text.x = element_text(angle = 90))
-)
-```
-
 ```{r mero_asex_malegam}
@@
 
-order_path.names <- ordering(unique(mero_asex_malegam_stats_df$pathway.name), 
+order_path.names <- ordering(sort(unique(mero_asex_malegam_stats_df$pathway.name)), 
                              order_vector = c('pir', 
@@
 
+```{r running_qusage_stage_by_stage}
+ordered_stages <- ordering(unique(lifeCycle_labels),
+         order_vector = r_order_vector)
+
+comparisons <- c(
+  unlist(
+    lapply(seq_along(ordered_stages),
+           function(i){ if(i < length(ordered_stages)){ paste0(ordered_stages[i],'-',ordered_stages[i+1]) } }
+    )
+  ),
+  'Asex.Schiz-Gam.Male',
+  'Ook-Sporoz'
+)
+
+#Remove some comparisons
+comparisons <- comparisons[!grepl(comparisons, pattern = 'Asex\\.Mixed-Asex\\.Ring')]
+
+all_comparisons <- lapply(comparisons, 
+                          function(comparison){
+                            list_comp <- lapply(
+                              list(pir_geneset,
+                                   ShortLong_list,
+                                   subfam_list), 
+                              function(gene_list){
+                                qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
+                                                            lifeCycle_labels,
+                                                            comparison,
+                                                            gene_list,
+                                                            n.points = 2^18)
+                                list(results = qs.results.subfam, 
+                                     qstable = qsTable(qs.results.subfam) %>% 
+                                       mutate(contrast = qs.results.subfam$contrast))
+                              })
+                            names(list_comp) <- c('pir', 'shortlong', 'subfam')
+                            list_comp$combined <- bind_rows(lapply(list_comp, 
+                                                                   getElement, 
+                                                                   'qstable'))
+                            return(list_comp)
+                          })
+names(all_comparisons) <- comparisons
+
+all_stats_df <- lapply(all_comparisons, getElement, 'combined') %>% 
+  bind_rows %>% 
+  mutate(fdr_r = p.adjust(p.Value, method = 'fdr')) %>% 
+  arrange(fdr_r)
+
+all_stats_df <- all_stats_df[order(match(all_stats_df$pathway.name,order_path.names)),]
+
+all_stats_df_cast <- dcast(all_stats_df,
+                           contrast ~ pathway.name,
+                           value.var = 'log.fold.change') %>% 
+  column_to_rownames(var = 'contrast')
+
+all_stats_df_cast <- all_stats_df_cast[ordering(rownames(all_stats_df_cast),
+                                                r_order_vector),
+                                       order_path.names]
+
+Heatmap(all_stats_df_cast,
+        name = 'logFoldChange',
+        cluster_rows = FALSE,
+        cluster_columns = FALSE,
+        rect_gp = gpar(col = "white", lwd = 2),
+        cell_fun = function(j, i, x, y, width, height, fill){
+          fdr_val <- all_stats_df[
+            all_stats_df$contrast == rownames(all_stats_df_cast[i,]),
+          ]$fdr_r[j]
+          if(fdr_val<= 0.05){
+            grid.text(
+              sprintf("%.3f", #One significant figure
+                      fdr_val),
+              x,
+              y,
+              gp = gpar(fontsize = 10,
+                        fontface = 'bold')
+            )
+          }
+        })
+```
+
 ```{r save_xlsx}

### 2021-09-15 rfunctions/stacked_barchart_fun.R

@@ -35,11 +35,12 @@ stacked_barchart <- function(melted_tpm,
                                       order_vector = r_order_vector))
            )) %>% 
     group_by(stage) %>% 
-    mutate(stacked_mean = cumsum(mean_sum_tpm),
-           sd_upper = cumsum(mean_sum_tpm) + cumsum(sd_sum_tpm),
-           sd_lower = ifelse(cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm) < 0,
-                             0, 
-                             cumsum(mean_sum_tpm) - cumsum(sd_sum_tpm)))
+    mutate(stacked_mean = cumsum(mean_sum_tpm)) %>% 
+    mutate(sd_upper = stacked_mean + sd_sum_tpm,
+           sd_lower = ifelse(sd_sum_tpm > mean_sum_tpm,
+                             lag(stacked_mean, default = 0), 
+                             stacked_mean - sd_sum_tpm)
+    )
   
   tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))
   

### 2021-09-15 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_pir.Rmd
move stacked_barchart sourcing fun

@@
 source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
-source('../rfunctions/stacked_barchart_fun.R')
 ```
@@
 ```{r stacked_barchart}
+source('../rfunctions/stacked_barchart_fun.R')
 sd_bars <- TRUE

### 2021-09-16 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Add three peak stat comparison to main heatmap

@@
 
-```{r mero_asex_malegam}
-mero_asex_malegam_stats_df <- filter(berghei_comparison_stats_df,
-                                     contrast %in% c('Asex.Mixed-Liv.Meroz',
-                                                     'Gam.Male-Asex.Mixed',
-                                                     'Gam.Male-Liv.Meroz'))
-
-order_path.names <- ordering(sort(unique(mero_asex_malegam_stats_df$pathway.name)), 
-                             order_vector = c('pir', 
-                                              'L', 'L[[:digit:]]', 
-                                              'S', 'S[[:digit:]]'))
-
-mero_asex_malegam_stats_df <- mero_asex_malegam_stats_df[order(match(mero_asex_malegam_stats_df$pathway.name,order_path.names)),]
-
-max_logfold <- ceiling(max(max(mero_asex_malegam_stats_df$log.fold.change), 
-                           abs(min(mero_asex_malegam_stats_df$log.fold.change))))
-
-mero_asex_malegam_stats_df_cast <- dcast(mero_asex_malegam_stats_df,
-      contrast ~ pathway.name,
-      value.var = 'log.fold.change') %>% 
-  column_to_rownames(var = 'contrast')
-
-mero_asex_malegam_stats_df_cast <- mero_asex_malegam_stats_df_cast[,order_path.names]
-
-Heatmap(mero_asex_malegam_stats_df_cast,
-        name = 'logFoldChange',
-        cluster_rows = FALSE,
-        cluster_columns = FALSE,
-        rect_gp = gpar(col = "white", lwd = 2),
-        cell_fun = function(j, i, x, y, width, height, fill){
-          fdr_val <- mero_asex_malegam_stats_df[
-            mero_asex_malegam_stats_df$contrast == rownames(mero_asex_malegam_stats_df_cast[i,]),
-          ]$FDR[j]
-          if(fdr_val<= 0.1){
-            grid.text(
-              sprintf("%.3f", #One significant figure
-                      fdr_val),
-              x,
-              y,
-              gp = gpar(fontsize = 10,
-                        fontface = 'bold')
-            )
-          }
-        })
-
-```
-
 ```{r running_qusage_stage_by_stage}
@@
   'Asex.Schiz-Gam.Male',
-  'Ook-Sporoz'
+  'Ook-Sporoz',
+  'Gam.Male-Asex.Mixed',
+  'Gam.Male-Liv.Meroz'
 )

### 2021-09-16 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Save pirvnonpir TPM comparison graph

@@
                                        order_path.names]
-
+png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_geneGroups.png",
+    width = 7,
+    height = 5,
+    units = "in",
+    res = 330)
 Heatmap(all_stats_df_cast,
@@
         })
+dev.off()
 ```

### 2021-09-16 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Update DESEQ file

@@
 ```{r read_count_data}
-berg_counts <- read_xlsx('Manuscript_Frontiers_suppinfo_bergheiCounts_UpdatedNames.xlsx')
+berg_counts <- read.csv('berghei_counts_newNames.csv',
+                        row.names = 2) %>% 
+  select(!matches('BldMl|Gam_|^X$'))
 
 #Filter low expressed genes (less than 1 count in all samples), convert to a matrix, add 1 for the log transformation, then transform to log2
-berg_counts <- column_to_rownames(berg_counts, var = 'Geneid') %>% 
-  select(!matches('BldMl|Gam_'))
 berg_counts.mat <- berg_counts[!rowSums(berg_counts) <= 10,] %>% 
@@
 ```{r generate_genesets}
-pir_info <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
-                      sheet = 1)
+pir_info <- read.table('../standard_files/Pberghei_pir_clades.txt',
+                    header = TRUE)
 

### 2021-09-16 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Add Liv24h-Livpost48h comp to heatmap and add ordering sourcing into scripts

@@
 
+```{r sourcing_functions}
+source('../rfunctions/uniquify_fun.R')
+source('../rfunctions/ordering_fun.R')
+source('../rfunctions/tpm_fun.R')
+source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
+```
+
+```{r ordering_vectors}
+order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
+                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
+                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
+                 'Mero','Asex','Ring','Troph','Schiz',
+                 'Gam','Female','Male',
+                 '-Pre3h','BldMl-24h','Ook')
+
+r_order_vector <- str_replace_all(order_vector, 
+                                  pattern = '-', 
+                                  replacement = '\\.')
+```
+
 Download the data, and remove heterogeneous life cycle stages like 'Gam' and 'BldMl'.
@@
   experiment = str_extract(colnames(berg_counts.mat), pattern = '^E[[:digit:]]{2}'),
-  stage = str_extract(colnames(berg_counts.mat), pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]{1,2})')
-) )
+  stage = str_extract(colnames(berg_counts.mat), pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]{1,2})')) 
+  )
 ```
@@
                                                   type = "apeglm")
+
+dds.deseq.gamMale_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% pir_info$Geneid,] %>% 
+  as.data.frame() %>% 
+  mutate(subfam = pir_info$clade[match(rownames(.), pir_info$Geneid)]) %>% 
+  arrange(padj)
+
 gamMale_v_asexMixed_df <- as.data.frame(
@@
 ```
+
+```{r}
+ordered_stages <- ordering(unique(coldata$stage), order_vector = r_order_vector)
+comparisons <- c(
+  unlist(
+    lapply(seq_along(ordered_stages),
+           function(i){ if(i < length(ordered_stages)){ paste0(ordered_stages[i],'-',ordered_stages[i+1]) } }
+    )
+  ),
+  'Liv.post48h-Liv.24h',
+  'Asex.Schiz-Gam.Male',
+  'Ook-Sporoz',
+  'Gam.Male-Asex.Mixed',
+  'Gam.Male-Liv.Meroz'
+)
+```

### 2021-09-16 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Add Liv24h-Livpost48h comp to heatmap and add ordering sourcing into scripts

@@
 library('writexl')
+```
+
+```{r sourcing_functions}
+source('../rfunctions/uniquify_fun.R')
+source('../rfunctions/ordering_fun.R')
+source('../rfunctions/tpm_fun.R')
+source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
+```
+
+```{r ordering_vectors}
+order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
+                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
+                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
+                 'Mero','Asex','Ring','Troph','Schiz',
+                 'Gam','Female','Male',
+                 '-Pre3h','BldMl-24h','Ook')
+
+r_order_vector <- str_replace_all(order_vector, 
+                                  pattern = '-', 
+                                  replacement = '\\.')
 ```
@@
 
-```{r livmero_livpost48}
-livmero_livpost48.results <- lapply(
-  list(pir_geneset,
-       ShortLong_list,
-       subfam_list), 
-  function(gene_list){
-    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
-                                lifeCycle_labels,
-                                'Liv.Meroz-Liv.post48h',
-                                gene_list,
-                                n.points = 2^18)
-
-    list(results = qs.results.subfam, 
-         qstable = qsTable(qs.results.subfam) %>% 
-           mutate(contrast = qs.results.subfam$contrast))
-  }
-)
-lapply(livmero_livpost48.results, function(x) `$`(x,qstable))
-```
-
-```{r asexRing_asexMix}
-asexRing_asexMix.results <- lapply(
-  list(pir_geneset,
-       subfam_list,
-       ShortLong_list), 
-  function(gene_list){
-    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
-                                lifeCycle_labels,
-                                'Asex.Ring-Asex.Mixed',
-                                gene_list,
-                                n.points = 2^18)
-
-    list(results = qs.results.subfam, 
-         qstable = qsTable(qs.results.subfam) %>% 
-           mutate(contrast = qs.results.subfam$contrast))
-  }
-)
-names(asexRing_asexMix.results) <- c('all', 'subfam', 'SorL')
-asexRing_asexMix.results$subfam$qstable
-asexRing_asexMix.results$SorL$qstable
-```
-
-```{r maleGam_v_Asex.Mix}
-maleGam_v_Asex.Mix.results <- lapply(
-  list(pir_geneset,
-    subfam_list, 
-    ShortLong_list), 
-  function(gene_list){
-    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
-                                lifeCycle_labels,
-                                'Gam.Male-Asex.Mixed',
-                                gene_list,
-                                n.points = 2^18)
-
-    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
-    #                                 FDR < 0.05)
-
-    list(results = qs.results.subfam, 
-         qstable = qsTable(qs.results.subfam) %>% 
-           mutate(contrast = qs.results.subfam$contrast))
-  }
-)
-names(maleGam_v_Asex.Mix.results) <- c('all', 'subfam', 'SorL')
-
-plotGeneSetDistributions(maleGam_v_Asex.Mix.results$SorL$results, path.index =  'S')
-plotGeneSetDistributions(maleGam_v_Asex.Mix.results$SorL$results, path.index = 'L')
-plotGeneSetDistributions(maleGam_v_Asex.Mix.results$subfam$results, path.index = 'S4')
-mutate(maleGam_v_Asex.Mix.results$subfam$qstable , adjust = p.adjust(p.Value, method = 'fdr'))
-```
-
-```{r maleGam_v_Liv.Meroz}
-maleGam_v_Liv.Meroz.results <- lapply(
-  list(pir_geneset,
-    subfam_list, 
-    ShortLong_list), 
-  function(gene_list){
-    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
-                                lifeCycle_labels,
-                                'Gam.Male-Liv.Meroz',
-                                gene_list,
-                                n.points = 2^18)
-
-    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
-    #                                 FDR < 0.05)
-
-    list(results = qs.results.subfam, 
-         qstable = qsTable(qs.results.subfam) %>% 
-           mutate(contrast = qs.results.subfam$contrast))
-  }
-)
-```
-
-```{r Asex.Mix_v_Liv.Meroz}
-Asex.Mix_v_Liv.Meroz.results <- lapply(
-  list(pir_geneset,
-    subfam_list, 
-    ShortLong_list), 
-  function(gene_list){
-    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
-                                lifeCycle_labels,
-                                'Asex.Mixed-Liv.Meroz',
-                                gene_list,
-                                n.points = 2^18)
-
-    list(results = qs.results.subfam, 
-         qstable = qsTable(qs.results.subfam) %>% 
-           mutate(contrast = qs.results.subfam$contrast))
-  }
-)
-names(Asex.Mix_v_Liv.Meroz.results) <- c('all', 'subfam', 'SorL')
-
-plotGeneSetDistributions(Asex.Mix_v_Liv.Meroz.results$SorL$results, 
-                         path.index = 'L')
-plotGeneSetDistributions(Asex.Mix_v_Liv.Meroz.results$SorL$results, 
-                         path.index = 'S')
-
-bind_rows(lapply(Asex.Mix_v_Liv.Meroz.results, 
-                 getElement, 
-                 'qstable')) %>% 
-  mutate(adjust = p.adjust(p.Value)) %>% 
-  arrange(adjust)
-```
-
-```{r fdr}
-berghei_comparison_stats_df <- bind_rows(
-  lapply(c(maleGam_v_Asex.Mix.results, 
-           maleGam_v_Liv.Meroz.results, 
-           Asex.Mix_v_Liv.Meroz.results,
-           maleVfemale, 
-           post48livvearlyliv, 
-           livmero_livpost48.results),
-         getElement,
-         'qstable')
-) %>% 
-  select(contrast, everything()) %>% 
-  mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
-  arrange(contrast)
-```
-
 ```{r running_qusage_stage_by_stage}
@@
   ),
+  'Liv.post48h-Liv.24h',
   'Asex.Schiz-Gam.Male',

### 2021-09-17 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Heatmap of individual gene comps

@@
 library('apeglm')
+library('ashr')
 ```
@@
                               design = ~ stage)
-dds
 ```
@@
 
+Having looked into independent filter cut-offs, no genes are removed even by ludicrous 'alpha' values, so I assume that our dataset contains few low expressed genes (remember that we filtered low counts out earlier) or outliars of concern.
+
 ```{r}
@@
 comparisons <- c(
-  unlist(
   lapply(seq_along(ordered_stages),
-           function(i){ if(i < length(ordered_stages)){ paste0(ordered_stages[i],'-',ordered_stages[i+1]) } }
+         function(i){ if(i < length(ordered_stages)){ c(ordered_stages[i],ordered_stages[i+1]) } }
+  ),
+  list(
+    c('Liv.post48h','Liv.24h'),
+    c('Asex.Schiz','Gam.Male'),
+    c('Ook','Sporoz'),
+    c('Gam.Male','Asex.Mixed'),
+    c('Gam.Male','Liv.Meroz')
   )
-  ),
-  'Liv.post48h-Liv.24h',
-  'Asex.Schiz-Gam.Male',
-  'Ook-Sporoz',
-  'Gam.Male-Asex.Mixed',
-  'Gam.Male-Liv.Meroz'
 )
+comparisons <- comparisons[-which(sapply(comparisons, is.null))]
+#Remove some comparisons
+comparisons <- comparisons[!grepl(comparisons, pattern = 'Asex\\.Mixed-Asex\\.Ring')]
+
+shrunk.list <- lapply(comparisons, 
+                      function(comparison){
+                        res <- results(
+                          dds.deseq,
+                          contrast = c("stage",comparison[1],comparison[2])
+                        )
+                        shrunk <- lfcShrink(dds.deseq, 
+                                            res = res,
+                                            type = "ashr") %>% 
+                          as.data.frame %>% 
+                          rownames_to_column(var = 'Geneid') %>% 
+                          mutate(comp = paste0(comparison[1],'_',comparison[2])) %>% 
+                          filter(Geneid %in% pir_id,
+                                 !is.na(padj))
+                        return(shrunk)
+                      })
+names(shrunk.list) <- unlist(lapply(comparisons, 
+                                    function(comparison) paste0(comparison[1],'_',comparison[2])))
+sig_pirs <- bind_rows(shrunk.list) %>% 
+  mutate(fdr = p.adjust(pvalue)) %>% 
+  arrange(subfam, Geneid) %>% 
+  filter(fdr < 0.05)
+sig_pirs_order <- unique(sig_pirs$Geneid)
+sig_pirs_cast <- dcast(sig_pirs,
+                  comp ~ Geneid,
+                  value.var = 'log2FoldChange') %>% 
+  column_to_rownames(var = 'comp')
+sig_pirs_cast <- sig_pirs_cast[ordering(rownames(sig_pirs_cast),
+                              r_order_vector),
+                     sig_pirs_order]
+
+Heatmap(sig_pirs_cast,
+        name = 'logFoldChange',
+        cluster_rows = FALSE,
+        cluster_columns = FALSE,
+        # rect_gp = gpar(col = "white", lwd = 1),
+        border_gp = gpar(col = "black", lty = 1),
+        na_col = 'white',
+        show_column_names = FALSE,
+        column_split = gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
+                            pattern = 'ancestral',
+                            replacement = 'a')
+        )
+
 ```

### 2021-09-17 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Remove redundant comparisons and Change minReplicatesForReplace in DESEQ

@@
 ```{r deseq-ing}
-dds.deseq <- DESeq(dds)
+dds.deseq <- DESeq(dds,
+                   minReplicatesForReplace = 3)
 resultsNames(dds.deseq)
@@
 Individual _pir_ genes which we use as examples of being highly upregulated in individual stages.
-
-```{r gamMale_v_asexMix}
-coef <- "stage_Gam.Male_vs_Asex.Mixed"
-dds.deseq.gamMale_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
-                                                  coef = coef,
-                                                  type = "apeglm")
-
-dds.deseq.gamMale_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% pir_info$Geneid,] %>% 
-  as.data.frame() %>% 
-  mutate(subfam = pir_info$clade[match(rownames(.), pir_info$Geneid)]) %>% 
-  arrange(padj)
-
-gamMale_v_asexMixed_df <- as.data.frame(
-  dds.deseq.gamMale_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0317181'),]
-) %>% 
-  mutate(contrast = coef) %>% 
-  rownames_to_column(var = 'Geneid')
-```
-
-```{r}
-coef <- "stage_Liv.Meroz_vs_Asex.Mixed"
-res <- results(dds.deseq,
-               name = 'stage_Liv.Meroz_vs_Asex.Mixed',
-               cooksCutoff = FALSE)
-dds.deseq.livMeroz_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
-                                                  coef = coef,
-                                                  res = res,
-                                                  type = "apeglm")
-
-livMeroz_v_asexMixed_df <- as.data.frame(
-  dds.deseq.livMeroz_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0100500',
-                                                                                              'PBANKA_0600031'),]) %>%  
-  mutate(contrast = coef) %>% 
-  rownames_to_column(var = 'Geneid')
-```
-
-```{r asexRing_v_asexMixed}
-coef <- 'stage_Asex.Ring_vs_Asex.Mixed'
-
-dds.deseq.asexRing_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
-                                                  coef = coef,
-                                                  type = "apeglm")
-asexRing_v_asexMixed_df <- as.data.frame(
-  dds.deseq.asexRing_v_asexMixed.lfcShr[rownames(dds.deseq.asexRing_v_asexMixed.lfcShr) %in% c('PBANKA_1300200'),]
-) %>% 
-  mutate(contrast = coef) %>% 
-    rownames_to_column(var = 'Geneid')
-```
-
-```{r}
-dds$stage <- relevel(dds$stage, ref = 'Gam-Male')
-dds.maleGam.deseq <- DESeq(dds)
-resultsNames(dds.maleGam.deseq)
-```
-
-```{r}
-coef <- "stage_Liv.Meroz_vs_Gam.Male"
-res <- results(dds.maleGam.deseq,
-               name = 'stage_Liv.Meroz_vs_Gam.Male',
-               cooksCutoff = FALSE)
-dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr <- lfcShrink(dds.maleGam.deseq, 
-                                                         coef = coef,
-                                                         res= res,
-                                                         type = "apeglm")
-
-(livMeroz_v_gamMale_df <- as.data.frame(
-  dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr[rownames(dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr) %in% c('PBANKA_0317181',
-                                                                                                           'PBANKA_0100500',
-                                                                                                           'PBANKA_0600031'),]
-  ) %>%  
-  mutate(contrast = coef) %>% 
-    rownames_to_column(var = 'Geneid'))
-```
-
-```{r}
-(deseq_stats <- bind_rows(livMeroz_v_gamMale_df, 
-          as.data.frame(gamMale_v_asexMixed_df),
-          as.data.frame(livMeroz_v_asexMixed_df)) %>% 
-  mutate(FDR = p.adjust(pvalue, 
-                        method = 'fdr')) %>% 
-  select(-c(lfcSE, 
-            padj, 
-            pvalue)))
-```
 
@@
 sig_pirs <- bind_rows(shrunk.list) %>% 
-  mutate(fdr = p.adjust(pvalue)) %>% 
+  mutate(fdr = p.adjust(pvalue),
+         subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
   arrange(subfam, Geneid) %>% 
@@
                      sig_pirs_order]
+
+sig_pirs %>% filter(Geneid %in% 'PBANKA_0600031')
 

### 2021-09-17 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Save deseq heatmap and fix remove asex.mix v ring comp

@@
 #Remove some comparisons
-comparisons <- comparisons[!grepl(comparisons, pattern = 'Asex\\.Mixed-Asex\\.Ring')]
+comparisons <- comparisons[!(grepl(comparisons, pattern = 'Asex\\.Mixed') & grepl(comparisons, pattern = 'Asex\\.Ring'))]
 
@@
 
+sig_pirs %>% filter(Geneid %in% 'PBANKA_0317181')
 sig_pirs %>% filter(Geneid %in% 'PBANKA_0600031')
 
+png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_inidvidualGenes.png",
+    width = 7,
+    height = 5,
+    units = "in",
+    res = 330)
 Heatmap(sig_pirs_cast,
@@
 )
-
+dev.off()
 ```

### 2021-09-21 AncestralPir/AnalysisCode/Manuscript_Github_AncestralPir.Rmd
update renv snapshot and update ancestral pir code to source functions

@@
 knitr::opts_chunk$set(echo = FALSE, results = FALSE, message = FALSE)
-knitr::opts_knit$set(root.dir = 
-                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/AncestralPir/AnalysisCode")
 options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
@@
 
-```{r function_uniquify}
-uniquify <- function(vector_names, first_one = TRUE){
-  new_vector <- c()
-  df_num <- data.frame(names = unique(vector_names), rep = 1)
-  if(first_one){
-    duplicated_names <- unique(vector_names[duplicated(vector_names)])
-  }else{
-    duplicated_names <- c()
-  }
-  for(i in 1:length(vector_names)){
-    if(vector_names[i] %in% duplicated_names|!first_one){
-    name <- paste(vector_names[i],
-                  df_num$rep[match(vector_names[i], df_num$names)],
-                  sep = '_')
-    #Increase rep counter
-    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
-    new_vector[i] <- name
-    }else{
-      new_vector[i] <- vector_names[i]
-    }
-  }
-  new_vector
-}
-```
-
-```{r tpm_function}
-tpm = function(count_data, lengths) {
-
-  #Print an error if the id columns do not match
-  # first if statement to test that both have a GeneID column
-  if(any(grepl(colnames(lengths), 
-               pattern = 'id', 
-               ignore.case = T)) == T & any(grepl(colnames(count_data), 
-                                                  pattern = 'Geneid', 
-                                                  #This needs to be Geneid when some of the samples have 'id' in them
-                                                  ignore.case = T)) == T){
-    if(all(count_data[grep(colnames(count_data), 
-                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
-                                                                             pattern = 'id', ignore.case = T)])) {
-      print('Count data and length gene ids both match')
-    } else {
-      print('Warning:Count data and length gene ids DO NOT MATCH')
-    }
-  } else {
-    print('Count data and/or length table supplied do not have an id column, 
-          make sure they are ordered correctly')
-  }
-
-  #Remove the id column from supplied count_data
-  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = TRUE))){
-    gene_names <- count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
-    counts_tpm <- count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
-  }else{
-    counts_tpm <- count_data}
-
-  #Remove the id column from supplied lengths
-  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = TRUE))){
-    lengths.table = lengths[,grep(colnames(lengths), pattern = 'length', ignore.case = T)]
-  }else{
-    lengths.table <- lengths}
-  #Making lengths.table a matrix sometimes throws up problems
-  lengths.table = as.matrix(lengths.table)
-  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop
-
-  #Now for the actual calculation
-  rpm <- counts_tpm
-  for (i in 1:dim(counts_tpm)[2]) {
-    for (j in 1:dim(counts_tpm)[1]) {
-      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
-    }
-    scale.factor <- sum(rpm[,i])/(1E6)
-    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
-  }
-  colnames(counts_tpm) <- 'tpm'
-  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
-    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), 
-                                            pattern = 'Geneid', 
-                                            ignore.case = T), 
-                                      drop = FALSE],
-                           counts_tpm)
-  } 
-  return(counts_tpm)
-}
+```{r sourcing_functions}
+source('../../rfunctions/uniquify_fun.R')
+source('../../rfunctions/ordering_fun.R')
+source('../../rfunctions/tpm_fun.R')
+source('../../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
+source('../../rfunctions/ecdf_percentile_fun.R')
 ```
@@
   mutate(Geneid = `Gene ID`) %>% 
-  select(c(Geneid, `Transcript Length`))
+  rename(Transcript.Length = `Transcript Length`) %>% 
+  select(c(Geneid, Transcript.Length ))
 
@@
   group_by(ancestral_ortholog) %>% 
-  dplyr::summarise_at(vars(matches('tpm')), 
+  dplyr::summarise_at(vars(matches('WB')), 
                       list(~median(.))) %>% 
@@
   group_by(Geneid) %>% 
-  dplyr::summarise(length = median(length))
+  dplyr::summarise(Transcript.Length  = median(length))
 nrow(yoel_lengths)
@@
                                                           yoel_counts_full$Geneid)]) %>% 
-  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
+  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0)) %>% 
+  rename(tpm = Counts)
 

### 2021-09-21 AncestralPir/AnalysisCode/Manuscript_Github_AncestralPir.Rmd
Save phylo anc plot
Delete superfluous phylo tree

@@
 
-```{r tree_heatmap_all-genes}
-
-ancestral_df <- bind_rows(cyno_percentile_all,
-                          ovale_percentile_all,
-                          cb_percentile_all,
-                          vinck_percentile_all,
-                          coat_percentile_all,
-                          yoel_percentile_all,
-                          berg_percentile_all,
-                          chab_percentile_all,
-                          vivax_percentile_all)
-ancestral_data <- lapply(phylog$old.tip.label, function(id){
-  #If any of the pattern is in the string one will be TRUE
-  ratio <- if(any(str_detect(string = paste(id), 
-                             pattern = ancestral_df$Geneid))){
-    index_id <- str_detect(string = paste(id), 
-                           pattern = ancestral_df$Geneid)
-    ancestral_df$percentile[index_id]*100
-  }else{
-    NA
-  }
-  return(list(ratio = ratio))
-}) %>% 
-  bind_rows
-ancestral_data <- as.data.frame(ancestral_data)
-
-rownames(ancestral_data) <- phylog$tip.label
-
-breaks <- seq(0, 100, by = 10)
-
-(tr <- ggtree(phylog, 
-            branch.length = 'none',
-            layout = 'circular') +
-  geom_tiplab2(aes(angle=angle, x = x + 2.5), 
-               cex = 3, 
-               color='black',
-               font = 3) +
-  ggplot2::xlim(0,20) 
-)
-```
-
 ```{r heatmap_with_tree_figure6}
@@
 
-p
+if(global_save_var){
+  ggsave('plots/thesis_chapter3_figure_ancPhyloHeatmap.png',
+         width = 7,
+         height = 5,
+         units = "in",
+         p)
+}
 ```

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Heatmap of individual gene comps
Reverse order of comparisons in Qusage heatmap

@@
 
+order_path.names <- ordering(c('pirs', unique(pir_info$SorL), unique(pir_info$clade)), 
+                             order_vector = c('pir','L','S'))
+
 ```
@@
 
-```{r post48livvearlyliv}
-
-post48livvearlyliv <- lapply(
-  list(pir_geneset,
-    ShortLong_list), 
-  function(gene_list){
-    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
-                                lifeCycle_labels,
-                                'Liv.post48h-Liv.24h',
-                                gene_list,
-                                n.points = 2^18)
-
-    list(results = qs.results.subfam, 
-         qstable = qsTable(qs.results.subfam) %>% 
-           mutate(contrast = qs.results.subfam$contrast))
-  }
-)
-
-lapply(post48livvearlyliv, function(x) `$`(x,qstable))
-```
+Edit to not do FDR again for all comparisons together, and reversing the order of labels/logFoldchanges to better match actual transitions.
 
@@
     lapply(seq_along(ordered_stages),
-           function(i){ if(i < length(ordered_stages)){ paste0(ordered_stages[i],'-',ordered_stages[i+1]) } }
+           function(i){ if(i < length(ordered_stages)){ paste0(ordered_stages[i+1],'-',ordered_stages[i]) } }
     )
@@
 all_stats_df <- lapply(all_comparisons, getElement, 'combined') %>% 
-  bind_rows %>% 
-  mutate(fdr_r = p.adjust(p.Value, method = 'fdr')) %>% 
-  arrange(fdr_r)
-
-all_stats_df <- all_stats_df[order(match(all_stats_df$pathway.name,order_path.names)),]
+  bind_rows
 
@@
                                                 r_order_vector),
-                                       order_path.names]
+                                       order(match(colnames(all_stats_df_cast), 
+                                                   order_path.names))]
+
 png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_geneGroups.png",
@@
         cell_fun = function(j, i, x, y, width, height, fill){
-          fdr_val <- all_stats_df[
-            all_stats_df$contrast == rownames(all_stats_df_cast[i,]),
-          ]$fdr_r[j]
+          stat_df <- all_stats_df[all_stats_df$contrast == rownames(all_stats_df_cast[i,]),]
+          fdr_val <- stat_df$FDR[j]
+          logfold <- stat_df$log.fold.change[j]
           if(fdr_val<= 0.05){
@@
               gp = gpar(fontsize = 10,
-                        fontface = 'bold')
+                        fontface = 'bold',
+                        col = ifelse(logfold > 2, 'white', 'black'))
             )

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Reverse order of labels and fix sub-fam order

@@
 
-order_path.names <- ordering(c('pirs', unique(pir_info$SorL), unique(pir_info$clade)), 
+order_path.names <- ordering(c('pirs', unique(pir_info$SorL), sort(unique(pir_info$clade))), 
                              order_vector = c('pir','L','S'))
@@
   'Liv.post48h-Liv.24h',
-  'Asex.Schiz-Gam.Male',
-  'Ook-Sporoz',
+  'Gam.Male-Asex.Schiz',
+  'Sporoz-Ook',
   'Gam.Male-Asex.Mixed',
@@
 all_stats_df_cast <- all_stats_df_cast[ordering(rownames(all_stats_df_cast),
-                                                r_order_vector),
+                                                c(r_order_vector,
+                                                  'Gam.Male-Asex.Mixed',
+                                                  'Gam.Male-Liv.Meroz')),
                                                 order(match(colnames(all_stats_df_cast), 

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_CalculateTPM.Rmd
Add newName counts file and add is.equal test for TPM

@@
 library('factoextra') # clustering algorithms & visualization
-library('ggfortify') # so ggplot understands prcomp objects
-library('ape') # for phylogeny plotting
 library('dendextend') # for dendrogram plotting
-library('plotly')
-library('ggpattern')
-library('ComplexHeatmap')
 library('ggdendro') #Drawing ggplot dendrograms
@@
 
-apply(berghei_tpm[-1], 2, sum)
+#Do they all sum to 1e6? Note that all.equal allows for a tiny bit of difference which is okay.
+all.equal(apply(berghei_tpm[-1], 2, sum), 
+          rep(1e6, times = dim(berghei_tpm[-1])[2]), 
+          check.names = FALSE)
 ```

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Add newName counts file and add is.equal test for TPM

@@
 
-```{r save_stats_df}
-write_xlsx(deseq_stats, 
-           'SupplementaryInformation4_Stats_DESEQ_berghei.xlsx')
-write.csv(deseq_stats, 
-           'SupplementaryInformation4_Stats_DESEQ_berghei.csv')
-```
-
 Having looked into independent filter cut-offs, no genes are removed even by ludicrous 'alpha' values, so I assume that our dataset contains few low expressed genes (remember that we filtered low counts out earlier) or outliars of concern.
@@
   lapply(seq_along(ordered_stages),
-         function(i){ if(i < length(ordered_stages)){ c(ordered_stages[i],ordered_stages[i+1]) } }
+         function(i){ if(i < length(ordered_stages)){ c(ordered_stages[i+1],ordered_stages[i]) } }
   ),
   list(
-    c('Liv.post48h','Liv.24h'),
-    c('Asex.Schiz','Gam.Male'),
-    c('Ook','Sporoz'),
+    c('Liv.24h','Liv.post48h'),
+    c('Gam.Male','Asex.Schiz'),
+    c('Sporoz','Ook'),
     c('Gam.Male','Asex.Mixed'),
@@
 ```
+
+```{r save_stats_df}
+sig_pirs_cast
+
+write_xlsx(deseq_stats, 
+           'SupplementaryInformation4_Stats_DESEQ_berghei.xlsx')
+write.csv(deseq_stats, 
+           'SupplementaryInformation4_Stats_DESEQ_berghei.csv')
+# deseq_stats <- read.csv('SupplementaryInformation4_Stats_DESEQ_berghei.csv',
+#                         row.names = 1)
+```

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Fix heatmap labels Qusage which were wrong

@@
 library('writexl')
+library('ComplexHeatmap')
 ```
@@
         cell_fun = function(j, i, x, y, width, height, fill){
-          stat_df <- all_stats_df[all_stats_df$contrast == rownames(all_stats_df_cast[i,]),]
-          fdr_val <- stat_df$FDR[j]
-          logfold <- stat_df$log.fold.change[j]
-          if(fdr_val<= 0.05){
+          stat_df <- all_stats_df[all_stats_df$contrast == rownames(all_stats_df_cast)[i],]
+          stat_df <- stat_df[stat_df$pathway.name == colnames(all_stats_df_cast)[j],]
+          fdr_val <- stat_df$FDR
+          logfold <- stat_df$log.fold.change
+          if(fdr_val <= 0.05){
             grid.text(

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Less adjustment DESeq and Label height change to fit

@@
 
-Individual _pir_ genes which we use as examples of being highly upregulated in individual stages.
+Having looked into independent filter cut-offs, no genes are removed even by ludicrous 'alpha' values, so I assume that our dataset contains few low expressed genes (remember that we filtered low counts out earlier) or outliers of concern.
 
-Having looked into independent filter cut-offs, no genes are removed even by ludicrous 'alpha' values, so I assume that our dataset contains few low expressed genes (remember that we filtered low counts out earlier) or outliars of concern.
-
-```{r}
+```{r comparisons_ordering}
 ordered_stages <- ordering(unique(coldata$stage), order_vector = r_order_vector)
@@
   list(
-    c('Liv.24h','Liv.post48h'),
+    c('Liv.post48h','Liv.24h'),
     c('Gam.Male','Asex.Schiz'),
@@
 comparisons <- comparisons[!(grepl(comparisons, pattern = 'Asex\\.Mixed') & grepl(comparisons, pattern = 'Asex\\.Ring'))]
+```
 
+```{r lfcShrink}
 shrunk.list <- lapply(comparisons, 
@@
 sig_pirs <- bind_rows(shrunk.list) %>% 
-  mutate(fdr = p.adjust(pvalue),
-         subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
+  mutate(subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
   arrange(subfam, Geneid) %>% 
-  filter(fdr < 0.05)
+  filter(padj < 0.05)
 sig_pirs_order <- unique(sig_pirs$Geneid)
@@
 sig_pirs_cast <- sig_pirs_cast[ordering(rownames(sig_pirs_cast),
-                              r_order_vector),
+                                        c(r_order_vector,
+                                          'Gam.Male_Asex.Mixed',
+                                          'Gam.Male_Liv.Meroz')),
                                sig_pirs_order]
@@
 sig_pirs %>% filter(Geneid %in% 'PBANKA_0600031')
+```
+
+```{r heatmap}
+#Add a new-line to the S6 label so that it doesn't overlap with others
+pir_column_split <- gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
+                         pattern = 'ancestral',
+                         replacement = 'a')
+pir_column_split <- ifelse(pir_column_split %in% c('S6', 'L2'), 
+                           paste0(pir_column_split, '\n'),
+                           pir_column_split)
 
@@
         border_gp = gpar(col = "black", lty = 1),
+        column_title_gp = gpar(fontsize = 12),
         na_col = 'white',
         show_column_names = FALSE,
-        column_split = gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
-                            pattern = 'ancestral',
-                            replacement = 'a')
+        column_split = pir_column_split,
+        heatmap_height = unit(10, "cm"),
+        #Shrink heatmap a little so that the S6 label is visible
+        column_gap = unit(2, 'mm')
 )

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Save deseq stats

@@
 ```{r save_stats_df}
-sig_pirs_cast
+deseq_stats <- bind_rows(shrunk.list) %>% 
+  mutate(subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
+  arrange(subfam, Geneid)
 
@@
            'SupplementaryInformation4_Stats_DESEQ_berghei.csv')
-# deseq_stats <- read.csv('SupplementaryInformation4_Stats_DESEQ_berghei.csv',
-#                         row.names = 1)
+
 ```

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Use cool grep technique to append '\n' to column names

@@
 #Add a new-line to the S6 label so that it doesn't overlap with others
-pir_column_split <- gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
+pir_column_split <- gsub(
+  gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
        pattern = 'ancestral',
-                         replacement = 'a')
-pir_column_split <- ifelse(pir_column_split %in% c('S6', 'L2'), 
-                           paste0(pir_column_split, '\n'),
-                           pir_column_split)
+       replacement = 'a'),
+  pattern = '(S6|L2)',
+  replacement = '\\1\n'
+)
 

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Save qusage

@@
 names(all_comparisons) <- comparisons
+```
+
+```{r heatmap}
 

### 2021-10-01 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Save qusage again

@@
 ```{r save_xlsx}
-to_save <- bind_rows(
-  lapply(c(maleGam_v_Asex.Mix.results, 
-           maleGam_v_Liv.Meroz.results, 
-           Asex.Mix_v_Liv.Meroz.results,
-           maleVfemale, 
-           post48livvearlyliv, 
-           livmero_livpost48.results),
-         getElement,
-         'qstable')
-) %>% 
-  select(contrast, everything()) %>% 
-  mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
-  arrange(contrast, pathway.name)
-
-write_xlsx(to_save,
+write_xlsx(all_stats_df,
            path = 'SupplementaryInformation_Stats_Qusage_berghei.xlsx')
-write.csv(to_save,
+write.csv(all_stats_df,
           'SupplementaryInformation_Stats_Qusage_berghei.csv')

### 2021-10-04 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Add colored labels to Qusage heatmap

@@
 library('ashr')
+library('ComplexHeatmap')
+library('gridtext') #For heatmap labels
 ```
@@
                           mutate(comp = paste0(comparison[1],'_',comparison[2])) %>% 
-                          filter(Geneid %in% pir_id,
+                          filter(Geneid %in% pir_geneset$pirs,
                                  !is.na(padj))
@@
 sig_pirs <- bind_rows(shrunk.list) %>% 
-  mutate(subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
+  mutate(subfam = pir_info$clade[match(Geneid, pir_info$Geneid)]) %>% 
   arrange(subfam, Geneid) %>% 
@@
 pir_column_split <- gsub(
-  gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
+  gsub(pir_info$clade[match(colnames(sig_pirs_cast), pir_info$Geneid)],
        pattern = 'ancestral',
@@
   replacement = '\\1\n'
+)
+
+#Make nice colour labels
+row_labels <- gt_render(
+  unlist(
+    lapply(rownames(sig_pirs_cast), 
+           function(comparison) {
+             comp_sep <- unlist(
+               strsplit(comparison, 
+                        split = '_')
+             )
+             txt.colour <- sapply(comp_sep,
+                                  function(stage_name){
+                                    case_when(
+                                      grepl(stage_name, pattern = 'Asex') ~ 'coral1' ,
+                                      grepl(stage_name, pattern = 'Liv') ~ '#BD65B7',
+                                      grepl(stage_name, pattern = 'Gam') ~ '#6998F4',
+                                      grepl(stage_name, pattern  = 'Bld|Sporo|Ook') ~ '#629C57')
+                                  })
+             paste0("<span style='color:",
+                    txt.colour[1],
+                    "'>**",
+                    comp_sep[1],
+                    "** </span><span style='color:",
+                    txt.colour[2],
+                    "'>**",
+                    comp_sep[2],
+                    "**</span>")
+           })
+  )
 )
@@
         cluster_columns = FALSE,
+        row_labels = row_labels,
         # rect_gp = gpar(col = "white", lwd = 1),
@@
 deseq_stats <- bind_rows(shrunk.list) %>% 
-  mutate(subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
+  mutate(subfam = pir_info$clade[match(Geneid, pir_info$Geneid)]) %>% 
   arrange(subfam, Geneid)

### 2021-10-04 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Remove Asex.Ring v Mix comparison Qusage
Add colored labels to Qusage heatmap

@@
 #Remove some comparisons
-comparisons <- comparisons[!grepl(comparisons, pattern = 'Asex\\.Mixed-Asex\\.Ring')]
+comparisons <- comparisons[!grepl(comparisons, pattern = 'Asex\\.Ring-Asex\\.Mixed')]
 
@@
 
+#Make nice colour labels
+row_labels <- gt_render(
+  unlist(
+    lapply(rownames(all_stats_df_cast), 
+           function(comparison) {
+             comp_sep <- unlist(
+               strsplit(comparison, 
+                        split = '-')
+             )
+             txt.colour <- sapply(comp_sep,
+                                  function(stage_name){
+                                    case_when(
+                                      grepl(stage_name, pattern = 'Asex') ~ 'coral1' ,
+                                      grepl(stage_name, pattern = 'Liv') ~ '#BD65B7',
+                                      grepl(stage_name, pattern = 'Gam') ~ '#6998F4',
+                                      grepl(stage_name, pattern  = 'Bld|Sporo|Ook') ~ '#629C57')
+                                  })
+             paste0("<span style='color:",
+                    txt.colour[1],
+                    "'>**",
+                    comp_sep[1],
+                    "** </span><span style='color:",
+                    txt.colour[2],
+                    "'>**",
+                    comp_sep[2],
+                    "**</span>")
+           })
+  )
+)
+
 png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_geneGroups.png",
@@
         cluster_columns = FALSE,
+        row_labels = row_labels,
         rect_gp = gpar(col = "white", lwd = 2),
@@
           'SupplementaryInformation_Stats_Qusage_berghei.csv')
+# all_stats_df <- read.csv('SupplementaryInformation_Stats_Qusage_berghei.csv',
+#                          row.names = 1)
 ```

### 2021-10-04 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Save stats and read in stats and print specific stats

@@
                                     function(comparison) paste0(comparison[1],'_',comparison[2])))
-sig_pirs <- bind_rows(shrunk.list) %>% 
+
+deseq_stats <- bind_rows(shrunk.list) %>% 
   mutate(subfam = pir_info$clade[match(Geneid, pir_info$Geneid)]) %>% 
-  arrange(subfam, Geneid) %>% 
-  filter(padj < 0.05)
+  arrange(subfam, Geneid)
+
+sig_pirs <- filter(deseq_stats, padj < 0.05)
 sig_pirs_order <- unique(sig_pirs$Geneid)
@@
 
-sig_pirs %>% filter(Geneid %in% 'PBANKA_0317181')
-sig_pirs %>% filter(Geneid %in% 'PBANKA_0600031')
+sig_pirs %>% filter(Geneid %in% 'PBANKA_0317181', grepl(comp, pattern = 'Male'))
+sig_pirs %>% filter(Geneid %in% 'PBANKA_0600031', grepl(comp, pattern = 'Meroz'))
 ```
@@
 ```{r save_stats_df}
-deseq_stats <- bind_rows(shrunk.list) %>% 
-  mutate(subfam = pir_info$clade[match(Geneid, pir_info$Geneid)]) %>% 
-  arrange(subfam, Geneid)
-
 write_xlsx(deseq_stats, 
@@
            'SupplementaryInformation4_Stats_DESEQ_berghei.csv')
-
+deseq_stats <- read.csv('SupplementaryInformation4_Stats_DESEQ_berghei.csv',
+                        row.names = 1)
 ```

### 2021-12-03 chapter3_results_bergheiPir/Pberghei_pir_samples_hclust-PCA-topGenes.Rmd
Create expt_tpms object and change function to use it

@@
                         row.names = 1)
+expt_tpms <- lapply(paste0('E0', seq(1,10)),
+                    function(Enum){
+                      df <- berghei_tpm[,str_subset(colnames(berghei_tpm), 
+                                              pattern = paste0('Geneid|',Enum))]
+                    })
+names(expt_tpms) <- paste0('E0', seq(1,10))
 ```
@@
 list_bw = lapply(expt_tpms, function(expt){ 
-  lapply(colnames(expt$tpm)[-1], 
+  lapply(colnames(expt)[-1], 
          function(sample){
-           data = as.numeric(unlist(select(expt$tpm, 
+           data = as.numeric(unlist(select(expt, 
                                            paste(sample))))

### 2021-12-03 rfunctions/density_samples_fun.R

@@ -2,12 +2,12 @@ density_samples <- function(input,
                             save = FALSE,
                             line_val = 1){
   #Get some colours for the lines
-  colours = rep_len(cbPalette, length.out = length(colnames(input$tpm)[-1]))
-  list_dens = lapply(colnames(input$tpm)[-1], function(sample){
-    data = as.numeric(unlist(select(input$tpm, paste(sample))))
+  colours = rep_len(cbPalette, length.out = length(colnames(input)[-1]))
+  list_dens = lapply(colnames(input)[-1], function(sample){
+    data = as.numeric(unlist(select(input, paste(sample))))
     dens = density(log1p(data), bw = bw_mean)
     #Use the index of the column to get the colour
-    colouring = colours[grep(colnames(input$tpm)[-1], pattern = paste(sample))]
+    colouring = colours[grep(colnames(input)[-1], pattern = paste(sample))]
     list_return = list(dens, colouring)
     names(list_return) = c('dens','colouring')
     list_return
@@ -43,7 +43,7 @@ density_samples <- function(input,
     }
     legend(max_x-4, 
            max_dens, 
-           legend = paste(colnames(input$tpm)[-1]), 
+           legend = paste(colnames(input)[-1]), 
            col = colours,
            lty = c(rep(1, times = 9), 
                    rep(2, times = 9), 

### 2021-12-03 chapter3_results_bergheiPir/Pberghei_pir_samples_hclust-PCA-topGenes.Rmd
Reverse previous changes, need to save name of experiment in expt_tpms object

--- 2021-09-09 chapter3_results_bergheiPir/Pberghei_pir_samples_hclust-PCA-topGenes.Rmd
@@
                         row.names = 1)
+expt_tpms <- lapply(na.omit(unique(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}'))),
+                    function(Enum){
+                      df <- berghei_tpm[,str_subset(colnames(berghei_tpm), 
+                                              pattern = paste0('Geneid|',Enum))]
+                      list(experiment = Enum,
+                           tpm = df)
+                    })
+names(expt_tpms) <- na.omit(unique(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}')))
 ```

### 2021-12-03 rfunctions/density_samples_fun.R

@@ -2,12 +2,12 @@ density_samples <- function(input,
                             save = FALSE,
                             line_val = 1){
   #Get some colours for the lines
-  colours = rep_len(cbPalette, length.out = length(colnames(input)[-1]))
-  list_dens = lapply(colnames(input)[-1], function(sample){
-    data = as.numeric(unlist(select(input, paste(sample))))
+  colours = rep_len(cbPalette, length.out = length(colnames(input$tpm)[-1]))
+  list_dens = lapply(colnames(input$tpm)[-1], function(sample){
+    data = as.numeric(unlist(select(input$tpm, paste(sample))))
     dens = density(log1p(data), bw = bw_mean)
     #Use the index of the column to get the colour
-    colouring = colours[grep(colnames(input)[-1], pattern = paste(sample))]
+    colouring = colours[grep(colnames(input$tpm)[-1], pattern = paste(sample))]
     list_return = list(dens, colouring)
     names(list_return) = c('dens','colouring')
     list_return
@@ -43,7 +43,7 @@ density_samples <- function(input,
     }
     legend(max_x-4, 
            max_dens, 
-           legend = paste(colnames(input)[-1]), 
+           legend = paste(colnames(input$tpm)[-1]), 
            col = colours,
            lty = c(rep(1, times = 9), 
                    rep(2, times = 9), 

### 2022-07-21 chapter3_results_bergheiPir/Pberghei_pir_samples_hclust-PCA-topGenes.Rmd
Start Pchabaudi R analysis

--- 2021-09-09 chapter3_results_bergheiPir/Pberghei_pir_samples_hclust-PCA-topGenes.Rmd
@@
                         row.names = 1)
+expt_tpms <- lapply(na.omit(unique(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}'))),
+                    function(Enum){
+                      df <- berghei_tpm[,str_subset(colnames(berghei_tpm), 
+                                              pattern = paste0('Geneid|',Enum))]
+                      colnames(df) <- str_remove( colnames(df) ,pattern = paste0(Enum, '_'))
+                      list(experiment = Enum,
+                           tpm = df)
+                    })
+names(expt_tpms) <- na.omit(unique(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}')))
 ```

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Start Pchabaudi R analysis

--- 2021-09-06 Pchabaudi24h/Manuscript_Github_Pchabaudi_calculateTPM.Rmd
@@
 library('dplyr')
-library('tidyr')
 library('ggplot2')
-library('DESeq2')
-library("pheatmap")
+# library("pheatmap")
 library("RColorBrewer")
@@
 library('viridis')
-library('corrplot')
-library('reshape2')
-library('circlize') # for circular plots and colours
-library('cluster')    # clustering algorithms
-library('factoextra') # clustering algorithms & visualization
-library('ggfortify') # so ggplot understands prcomp objects
-library('ape') # for phylogeny plotting
-library('dendextend') # for dendrogram plotting
-library('readxl') # for reading in Excel spreadsheets
-library('plotly')
-library('scales')
-library('processx') # for downloading the plotly plots
-library('cowplot')
-library('ComplexHeatmap')
-library('ggbiplot')
-library('ggpattern')
+# library('corrplot')
+# library('reshape2')
+# library('circlize') # for circular plots and colours
+# library('cluster')    # clustering algorithms
+# library('factoextra') # clustering algorithms & visualization
+# library('ggfortify') # so ggplot understands prcomp objects
+# library('ape') # for phylogeny plotting
+# library('dendextend') # for dendrogram plotting
+# library('readxl') # for reading in Excel spreadsheets
+# library('plotly')
+# library('scales')
+# library('processx') # for downloading the plotly plots
+# library('cowplot')
+# library('ComplexHeatmap')
+# library('ggpattern')
 

### 2022-07-21 rfunctions/density_samples_fun.R

@@ -37,7 +37,8 @@ density_samples <- function(input,
          col = x$colouring,
          main = paste(input$experiment),
          lwd = line_width,
-         lty = lty)
+         lty = lty,
+         xlab = '')
     if(is.numeric(line_val)){
       abline(v = log1p(line_val))
     }

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Read in the correct data

@@
 cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-show_col(cbPalette)
+#show_col(cbPalette)
 
@@
 ```{r calc_tpm}
-
-chab_count_table <- read.csv('Manuscript_Suppinfo_chabaudi_24h_counts.csv',
-                             row.names = 1)
-
+#Read in the count data and remove one sample which is unknown and had no description.
+chab_count_table <- read.csv('data/tl19-07_twentyfour_counts_samples.csv') %>% 
+  select(-rMT_Unknown)
+#Calculate the TPM
 twentyfour_ct_comb <- tpm(count_data = chab_count_table, 

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Change name of twentyfour_ct_comb to chab_tpm.df and check TPM all add to 1e6

@@
 #Calculate the TPM
-twentyfour_ct_comb <- tpm(count_data = chab_count_table, 
+chab_tpm.df <- tpm(count_data = chab_count_table, 
                           lengths = chabaudi_transcript_lengths)
+
+#Do they all sum to 1e6? Note that all.equal allows for a tiny bit of difference which is okay.
+all.equal(apply(twentyfour_ct_comb[-1], 2, sum), 
+          rep(1e6, times = dim(twentyfour_ct_comb[-1])[2]), 
+          check.names = FALSE)
 ```
@@
 
-twentyfour_tpm_nopseudo <- twentyfour_ct_comb %>% 
+twentyfour_tpm_nopseudo <- chab_tpm.df %>% 
   filter(!Geneid %in% 'PCHAS_0420800')
@@
 #Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
-twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo
+chab_tpm.df <- twentyfour_tpm_rescale_nopseudo
 ```
@@
 
-twentyfour_ct_comb <- rename_with(twentyfour_ct_comb, 
+chab_tpm.df <- rename_with(chab_tpm.df, 
                                   ~ str_extract(.x,pattern = '(?<=rMT_).+'),
@@
 
-twentyfour_rMT_avg <- melt(twentyfour_ct_comb, 
+twentyfour_rMT_avg <- melt(chab_tpm.df, 
                            variable.name = 'sample',

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Change more names to chab_tpm base and save pseudogene image

@@
 #Do they all sum to 1e6? Note that all.equal allows for a tiny bit of difference which is okay.
-all.equal(apply(twentyfour_ct_comb[-1], 2, sum), 
-          rep(1e6, times = dim(twentyfour_ct_comb[-1])[2]), 
+all.equal(apply(chab_tpm.df[-1], 2, sum), 
+          rep(1e6, times = dim(chab_tpm.df[-1])[2]), 
           check.names = FALSE)
+```
+
+```{r super-pseudogene}
+
+chab_tpm_melt.df <- reshape2::melt(chab_tpm.df,
+                                   variable.name = 'sample',
+                                   value.name = 'tpm') %>% 
+  mutate(time = str_extract(sample, 
+                            pattern = '([[:digit:]]{2}h_.+|sch.culture)(?=_[[:digit:]])'),
+         transmission = str_extract(sample,
+                                    pattern = 'rMT|SBP'))
+
+(plot <- ggplot(data = filter(chab_tpm_melt.df, 
+                     Geneid %in% 'PCHAS_0420800'), 
+       aes(x = time, 
+           y = log1p(tpm),
+           fill = transmission)) +
+  stat_summary(fun = median, fun.min = median, fun.max = median,
+               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
+  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
+  scale_fill_manual(values = cbPalette) +
+  theme_classic() +
+  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)))
+
+ggsave(filename = 'plots/pseudogene_tpm_barchart.png',
+       plot)
+
 ```
@@
 
-#Setting twentyfour_tpm_ct core dataset - remove pseudogene and re-scaled
+#Setting chab_tpm_ct core dataset - remove pseudogene and re-scaled
 
-twentyfour_tpm_nopseudo <- chab_tpm.df %>% 
+chab_tpm_nopseudo <- chab_tpm.df %>% 
   filter(!Geneid %in% 'PCHAS_0420800')
 
-twentyfour_tpm_rescale_nopseudo <- lapply(colnames(twentyfour_tpm_nopseudo), 
+chab_tpm_rescale_nopseudo <- lapply(colnames(chab_tpm_nopseudo), 
                                           function(column_name){
-                                            column <- getElement(twentyfour_tpm_nopseudo, 
+                                            column <- getElement(chab_tpm_nopseudo, 
                                                                  paste(column_name))
@@
 
-colnames(twentyfour_tpm_rescale_nopseudo) <- colnames(twentyfour_tpm_nopseudo)
-# apply(twentyfour_tpm_rescale_nopseudo[,-1], 2, sum)
+colnames(chab_tpm_rescale_nopseudo) <- colnames(chab_tpm_nopseudo)
+# apply(chab_tpm_rescale_nopseudo[,-1], 2, sum)
 
 #Setting the re-scaled version as the default from now on
-#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
-chab_tpm.df <- twentyfour_tpm_rescale_nopseudo
+#Setting chab_tpm_ct core dataset - removed pseudogene and re-scaled
+# chab_tpm.df <- chab_tpm_rescale_nopseudo
 ```
@@
 ```{r clean_up_and_average}
-# Change the names to remove superfluous 'rMT'
-
-chab_tpm.df <- rename_with(chab_tpm.df, 
-                                  ~ str_extract(.x,pattern = '(?<=rMT_).+'),
-                                  matches('rMT'))
-
-twentyfour_rMT_avg <- melt(chab_tpm.df, 
+chab_tpm_avg.df <- reshape2::melt(chab_tpm.df, 
                            variable.name = 'sample',
@@
   summarise(mean_tpm = mean(tpm)) %>% 
-  dcast(Geneid ~ stage)
+  reshape2::dcast(Geneid ~ stage)
 ```

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Check the TPM results in the future with 'dput' output of a random part of the data frame)

@@
 
+```{r check_tpm}
+
+#Used `dput` function to create a df of randomly chosen part of what I assume to be the true values.
+
+check_tpm <- structure(list(Geneid = c("PCHAS_1236900", "PCHAS_1237000", "PCHAS_1237100", 
+"PCHAS_1237200", "PCHAS_1237300", "PCHAS_1237400"), rMT_17h_Early.Rings_1 = c(9.76166246727995, 
+2.53080138040591, 10.9126974974354, 197.447300616447, 27.0792042284135, 
+55.9778949394867), rMT_sch.culture_1 = c(4.82209836798507, 39.9498029457425, 
+4.36161304058411, 192.21404227623, 255.05821864658, 252.312788862482
+), rMT_20h_Early.Mid.Rings_1 = c(17.0998539251424, 7.33599453376897, 
+12.8493972283173, 213.998967387733, 11.0784616329143, 60.1572274917456
+), rMT_05h_Rings.Troph.conversion_1 = c(38.3647026622499, 6.90734779639621, 
+2.62130156899195, 194.546775103218, 3.78235443023746, 130.068871216885
+), rMT_sch.culture_2 = c(2.90683176639546, 22.597581594288, 4.18468334820306, 
+368.645503106066, 438.239482969298, 330.318886377763), rMT_05h_Rings.Troph.conversion_2 = c(32.6510468524069, 
+7.16075860604918, 1.31010772565867, 164.370211763169, 2.96069000666786, 
+109.168242066326), rMT_11h_Mid.Trophs_1 = c(26.147595680817, 
+27.0999659114182, 2.16752322103163, 541.644380621737, 2.53144630329207, 
+120.735400671386), rMT_14h_Late.Trophs_1 = c(11.4079837593492, 
+22.856112916092, 10.5413074421478, 951.44994461046, 73.4132553701819, 
+131.681439654117), rMT_02h_Late.Rings_1 = c(31.2811649503316, 
+6.57534976948559, 3.85345590947847, 150.532386139873, 3.9545205115986, 
+94.4792421254667), rMT_08h_Early.Trophs_1 = c(20.157812092634, 
+12.1280059532542, 0.650734564771486, 170.309456869319, 1.59767139279832, 
+83.2278478943328), rMT_17h_Early.Rings_2 = c(17.403862005927, 
+3.69172830428755, 20.6004897213014, 250.122760155978, 17.2424791957208, 
+59.8810845288676), rMT_23h_Late.Mid.Rings_1 = c(29.5379734077001, 
+5.22489884006776, 3.67411111636422, 170.523646451821, 4.16641042383719, 
+82.6825661834616), rMT_23h_Late.Mid.Rings_2 = c(32.6253792704758, 
+5.07505899762957, 3.70087210432098, 163.457320547547, 2.80131367804736, 
+95.4154100528915), rMT_20h_Early.Mid.Rings_2 = c(9.98436534398298, 
+8.86560981918151, 15.9408661685613, 168.519812840241, 18.4953322269757, 
+67.1081583940078), rMT_11h_Mid.Trophs_2 = c(16.8514270116997, 
+15.0710551161147, 1.34774236993737, 283.638146433073, 2.88781264078723, 
+104.469069579326), rMT_17h_Early.Rings_3 = c(9.80685483532316, 
+7.8009072553707, 12.9140266989638, 244.004094798698, 32.6454833873949, 
+61.5795685612941), rMT_05h_Rings.Troph.conversion_3 = c(44.6914009986881, 
+7.31985380744853, 2.09226783944416, 202.848558910157, 3.38742465612069, 
+108.563983207627), rMT_14h_Late.Trophs_2 = c(16.1568208764581, 
+25.859815232017, 5.9401546141338, 849.643405800935, 7.56694623416709, 
+138.133764925281), rMT_08h_Early.Trophs_2 = c(27.5234173956934, 
+13.6584098001981, 2.54054386371032, 341.123228542418, 2.10140240005804, 
+132.057622012364), rMT_20h_Early.Mid.Rings_3 = c(20.2528241267272, 
+5.33303205844855, 11.6247077258001, 252.801844519901, 11.208737218013, 
+96.11509755057), rMT_sch.culture_3 = c(7.02704565010117, 31.941116591369, 
+10.1000976836731, 459.609650763128, 291.376205366797, 211.18569154735
+), rMT_02h_Late.Rings_2 = c(41.5030181049848, 8.57199496745699, 
+6.57168269134606, 260.300795278229, 8.65645647459549, 115.867220110965
+), rMT_11h_Mid.Trophs_3 = c(23.668981121663, 17.3111675994721, 
+2.40864504534792, 403.270229427547, 3.31051807265358, 107.600469742135
+), SBP_02h_Late.Rings_1 = c(32.1039867447089, 4.18890973084741, 
+1.83235045534882, 253.342646772667, 3.24971327904687, 147.599200658419
+), SBP_05h_Rings.Troph.conversion_1 = c(69.9933903732369, 5.56765605241657, 
+3.51252089834496, 253.482911605055, 3.94202779803947, 149.365010977175
+), SBP_08h_Early.Trophs_1 = c(67.4596701130964, 8.04708690737168, 
+1.29888044701547, 280.113721889105, 2.32961719999114, 168.821435727764
+), SBP_11h_Mid.Trophs_1 = c(39.0509461034118, 20.3431512112609, 
+2.45091379976465, 870.791722784726, 2.13865338832938, 218.047682451552
+), SBP_14h_Late.Trophs_1 = c(31.0258753755895, 24.8362618680581, 
+8.02873672984643, 1248.50319033523, 8.08944053716688, 165.499381325518
+), SBP_17h_Early.Rings_1 = c(21.5890368368948, 23.7391769164685, 
+7.41687066698247, 827.388691990966, 97.5224751506972, 187.900653468842
+), SBP_20h_Early.Mid.Rings_1 = c(4.91288997903853, 6.15026465334183, 
+7.00690865862872, 183.461124643034, 40.6941010432805, 94.2171898730868
+), SBP_23h_Late.Mid.Rings_1 = c(17.0960581250041, 4.46989348422346, 
+2.59223414646292, 226.408675568144, 8.83071196883189, 130.839085546676
+), SBP_02h_Late.Rings_2 = c(36.1370878333684, 2.84224286329863, 
+3.76479788403891, 193.540809015887, 4.86746227785013, 102.682045137645
+), SBP_05h_Rings.Troph.conversion_2 = c(55.9703217173144, 5.09338376368401, 
+1.6164154903665, 326.007322410062, 2.84115197098826, 154.530538517825
+), SBP_08h_Early.Trophs_2 = c(59.1292861603963, 6.89801752788374, 
+1.62611489888877, 302.374971383912, 2.28208571581952, 159.285259970324
+), SBP_05h_Rings.Troph.conversion_3 = c(64.0277646953059, 4.36881423377574, 
+5.38898632230727, 265.098915160974, 6.74434224459245, 115.619635926338
+), SBP_11h_Mid.Trophs_2 = c(56.2225636049175, 14.0211061570003, 
+1.53632625772768, 440.630839066675, 2.67215322318625, 180.373213104462
+), SBP_14h_Late.Trophs_2 = c(30.676698797827, 27.1931566057931, 
+6.91186772412193, 1614.99745495943, 22.3622898426203, 156.841577995753
+), SBP_17h_Early.Rings_2 = c(14.7757154238053, 15.693441999666, 
+10.0058832290506, 688.261040899689, 103.363777591419, 166.096400179785
+), SBP_20h_Early.Mid.Rings_2 = c(7.1021602822451, 4.3088000205048, 
+5.66250275226056, 111.85389541494, 18.1995988898424, 79.5127602899565
+), SBP_23h_Late.Mid.Rings_2 = c(27.6455730547316, 2.04450518969984, 
+3.11285673303378, 173.335869764135, 3.75799617308546, 111.076770881732
+), SBP_08h_Early.Trophs_3 = c(63.9441758034038, 8.783379046428, 
+1.36572101399292, 296.461875716015, 2.58960184385856, 183.132522675328
+), SBP_11h_Mid.Trophs_3 = c(70.3465549733047, 11.856861649251, 
+1.25726103732839, 357.299957208987, 2.81495469320334, 192.931412027969
+), SBP_14h_Late.Trophs_3 = c(24.5234329455145, 29.1881967487647, 
+7.8567780739673, 1772.21118398556, 22.3035208013784, 168.207454664877
+), SBP_17h_Early.Rings_3 = c(15.8780527250301, 21.5556594570105, 
+7.02241483912262, 644.958506567788, 190.906286209725, 213.781884925216
+), SBP_20h_Early.Mid.Rings_3 = c(5.4041703258959, 3.85169957772944, 
+5.4692409481873, 114.947181076967, 19.3388591170259, 85.532444920705
+), SBP_23h_Late.Mid.Rings_3 = c(20.6721444393214, 2.08063791434729, 
+6.25771469400429, 131.683650813693, 7.05321807501424, 100.520179346894
+), SBP_02h_Late.Rings_3 = c(31.6880147939969, 3.29026677329356, 
+4.84759449553743, 215.013874125238, 3.49951253434249, 134.56799075684
+)), row.names = 3731:3736, class = "data.frame")
+
+all.equal(chab_tpm.df[3731:3736,colnames(check_tpm)],check_tpm)
+```
+
 ```{r super-pseudogene}

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Save the tpm files

@@
 ```{r clean_up_and_average}
-chab_tpm_avg.df <- reshape2::melt(chab_tpm.df, 
+chab_tpm_avg_nopseudo.df <- reshape2::melt(chab_tpm_rescale_nopseudo, 
                            variable.name = 'sample',
@@
 ```{r save}
-write_csv(twentyfour_rMT,
-          file = 'Pchabaudi_TPM_allBioReps.csv')
+write_csv(chab_tpm.df,
+          file = 'data/Pchabaudi_TPM_all_withPseudo.csv')
 
-write_csv(twentyfour_rMT_avg,
-          file = 'Pchabaudi_TPM_average.csv')
+write_csv(chab_tpm_rescale_nopseudo,
+          file = 'data/Pchabaudi_TPM_all.csv')
+
+write_csv(chab_tpm_avg_nopseudo.df,
+          file = 'data/Pchabaudi_TPM_average.csv')
 ```

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

--- 2021-09-06 Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
@@
 library('dplyr')
-library('tidyr')
 library('ggplot2')
-library('DESeq2')
-library("pheatmap")
 library("RColorBrewer")
 library('data.table')
-library('gridExtra')
 library('tibble')
@@
 library('viridis')
-library('corrplot')
 library('reshape2')
-library('circlize') # for circular plots and colours
-library('cluster')    # clustering algorithms
-library('factoextra') # clustering algorithms & visualization
-library('ggfortify') # so ggplot understands prcomp objects
-library('ape') # for phylogeny plotting
-library('dendextend') # for dendrogram plotting
-library('readxl') # for reading in Excel spreadsheets
-library('plotly')
-library('scales')
-library('processx') # for downloading the plotly plots
-library('cowplot')
 library('ComplexHeatmap')
-library('ggbiplot')
-library('ggpattern')
+library('circlize') # for circular plots and colour
+
+# library('DESeq2')
+# library("pheatmap")
+# 
+# library('gridExtra')
+# 
+# library('corrplot')
+# 
+# library('cluster')    # clustering algorithms
+# library('factoextra') # clustering algorithms & visualization
+# library('ggfortify') # so ggplot understands prcomp objects
+# library('ape') # for phylogeny plotting
+# library('dendextend') # for dendrogram plotting
+# library('readxl') # for reading in Excel spreadsheets
+# library('plotly')
+# library('scales')
+# library('processx') # for downloading the plotly plots
+# library('cowplot')
+# 
+# library('ggbiplot')
+# library('ggpattern')
 
@@
 cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-show_col(cbPalette)
 
@@
 #Global save var - set to TRUE if you want to save figures.
-global_save_var <- TRUE
+global_save_var <- FALSE
 ```
@@
 
-twentyfour_rMT <- read_csv('Pchabaudi_TPM_allBioReps.csv')
-
-twentyfour_avgtpm <- list(
+chab_tpm.df <- read_csv('data/Pchabaudi_TPM_all.csv')
+
+chab_tpm_avg.df <- read_csv('data/Pchabaudi_TPM_average.csv')
+
+chab_tpm_avg.list <- list(
   'RMT' = list(
-    tpm = read_csv('Pchabaudi_TPM_average.csv'),
+    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
+                                pattern = 'rMT|Geneid')],
     experiment = 'RMT'
+  ),
+  'SBP' = list(
+    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
+                                pattern = 'SBP|Geneid')],
+    experiment = 'SBP'
   )
@@
 
-combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm), 
+combined_data_pir_ht_list <- lapply(names(chab_tpm_avg.list), 
        function(list_name){
@@
 names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)
+
+#Need to reach into the data to find the maximum value of the log(TPM+1) for 
+# the colouring in order to maintain consistency.
 
@@
                                row_title = NULL,
+                               split = clade_info,
                                left_annotation = if(

### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

--- 2021-09-06 Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
@@
 library('viridis')
-library('corrplot')
 library('reshape2')
-library('circlize') # for circular plots and colours
-library('cluster')    # clustering algorithms
-library('factoextra') # clustering algorithms & visualization
-library('ggfortify') # so ggplot understands prcomp objects
-library('ape') # for phylogeny plotting
-library('dendextend') # for dendrogram plotting
-library('readxl') # for reading in Excel spreadsheets
-library('plotly')
-library('scales')
-library('processx') # for downloading the plotly plots
-library('cowplot')
-library('ComplexHeatmap')
-library('ggbiplot')
-library('ggpattern')
+
+# library('DESeq2')
+# library("pheatmap")
+# 
+# library('gridExtra')
+# 
+# library('corrplot')
+# 
+# library('circlize') # for circular plots and colours
+# library('cluster')    # clustering algorithms
+# library('factoextra') # clustering algorithms & visualization
+# library('ggfortify') # so ggplot understands prcomp objects
+# library('ape') # for phylogeny plotting
+# library('dendextend') # for dendrogram plotting
+# library('readxl') # for reading in Excel spreadsheets
+# library('plotly')
+# library('scales')
+# library('processx') # for downloading the plotly plots
+# library('cowplot')
+# library('ComplexHeatmap')
+# library('ggbiplot')
+# library('ggpattern')
 

### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

---
title: "Manuscript_Github_Pchabaudi_pir_TPMHeatmaps"
author: "Timothy Little"
date: "20/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

knitr::opts_knit$set(root.dir = 
                       ".")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

library('readr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('DESeq2')
library("pheatmap")
library("RColorBrewer")
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots and colours
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('readxl') # for reading in Excel spreadsheets
library('plotly')
library('scales')
library('processx') # for downloading the plotly plots
library('cowplot')
library('ComplexHeatmap')
library('ggbiplot')
library('ggpattern')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

#Global save var - set to TRUE if you want to save figures.
global_save_var <- FALSE
```

```{r source_fun}
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r order_vec}
order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info <- read.csv('../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)

chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID)

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)

```

```{r cir_info}

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')

#Add chapl_aapl/subfam numbers for easier heatmap drawing later.
cir_info <- cir_info %>% 
  group_by(chapl_aapl_name) %>% 
  mutate(chapl_aapl_name_num = n_distinct(Geneid)) %>% 
  mutate(
    chapl_aapl_name_num = paste0(
      chapl_aapl_name, 
      ' (n = ',
      chapl_aapl_name_num,
      ')'
    )) %>% 
  group_by(subfam) %>% 
  mutate(subfam_num = n_distinct(Geneid)) %>% 
  mutate(
    subfam_num = paste0(
      subfam, 
      ' (n = ',
      subfam_num,
      ')'
    )
  )
```

```{r times_stages_labels}
#Use the RMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r read_TPM}
#Need to use the calculateTPM script first.

twentyfour_rMT <- read_csv('Pchabaudi_TPM_allBioReps.csv')

twentyfour_avgtpm <- list(
  'RMT' = list(
    tpm = read_csv('Pchabaudi_TPM_average.csv'),
    experiment = 'RMT'
  )
)

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9', #L
        '#E69F00', #S
        '#999999'),
  sort(
    unique(cir_info$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(cir_info$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$subfam)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(cir_info$subfam == clade_name),
                                   ')'
                                 )
                               })

chapl_aapl_name_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$chapl_aapl_name)
  )
)

chapl_aapl_name_col_num <- chapl_aapl_name_col
names(chapl_aapl_name_col_num) <- sapply(
  names(chapl_aapl_name_col), 
  function(clade_name) {
    paste0(
      clade_name, 
      ' (n = ',
      sum(cir_info$chapl_aapl_name == clade_name, 
          na.rm = TRUE),
      ')'
    )
  })
```

_pir_ genes selected by being expressed above 1 TPM in at least one sample.

Adding in housekeeping genes for comparison:
HSP70 putative  PCHAS_0721000
MSP1 PCHAS_0831300
AMA1 PCHAS_0931000

```{r highest_expressed_stage_heatmap_fig3b}

housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
                                 'MSP1' = 'PCHAS_0831300',
                                 'AMA1' = 'PCHAS_0931000')

cir_housekeeping_id <- c(cir_id, 
                         'PCHAS_0721000',
                         'PCHAS_0831300',
                         'PCHAS_0931000')

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>% 
  column_to_rownames(var = 'Geneid') %>% 
  filter_all(
      any_vars(. > 10)
  )

#Ordering that a more understandable order

combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                means = apply(combined_data_pir, 1, mean),
                                subfam = factor(
                                  ifelse(rownames(combined_data_pir) %in% cir_id,
                                         cir_info$subfam[match(rownames(combined_data_pir), 
                                                               cir_info$Geneid)],
                                         'ref')
                                )) %>% 
  mutate(subfam =
           forcats::fct_shift(
             relevel(
               as.factor(subfam),
               ref = 'ref')
           )
  ) %>% 
  # filter(means > 1) %>% 
  arrange(subfam, 
          desc(means)) %>% 
  .$Geneid

#Order the heatmap data by the order defined above (Subfam, then mean expression).
combined_data_pir_ht <- combined_data_pir[
  order(
    match(rownames(combined_data_pir),
          combined_data_ordered),
    na.last = NA #Remove NAs
  ),
]

#Change the colnames to the times
colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                              pattern = '[[:digit:]]{2}h')

#Order by time point
combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                             drop = FALSE]

#Add a column for the subfam (ref for the reference genes)
combined_data_pir_ht$subfam <- ifelse(rownames(combined_data_pir_ht) %in% cir_id,
                                      cir_info$subfam_num[match(rownames(combined_data_pir_ht), 
                                                            cir_info$Geneid)],
                                      'ref')

#Get the ref gene actual names instead of the Geneid
rownames(combined_data_pir_ht) <- ifelse(
  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
  rownames(combined_data_pir_ht)
)

#Get the colour range
maxcol <- ceiling(
  max(
    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'subfam'])
  ))

#Sub-family annotation
clade_info <- combined_data_pir_ht$subfam
annotation_col <- c(clade_col, 'ref' = 'black')

#Subfamily annotation

clade_ha <- rowAnnotation(
  'sub-family' = anno_simple(
    clade_info,
    col = annotation_col,
  ),
  annotation_name_gp = gpar(fontsize = 10)
)

#Subfamily name labels
num_subfam <- cir_info %>% 
  group_by(subfam) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_subfam <- c(setNames(paste0(num_subfam$subfam, ' (n = ', num_subfam$num, ')'),
                            num_subfam$subfam),
                   'ref' = 'ref')

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title_rot = 0,
  row_title_gp = gpar(fontsize = 10),
  row_split = forcats::fct_shift(
    relevel(
      as.factor(combined_data_pir_ht$subfam),
      ref = 'ref')
  ),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

clade_lgd <- Legend(
  title = 'sub-family',
  at = c(labels_subfam),
  legend_gp = gpar(fill = annotation_col)
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     # heatmap_legend_side = 'right',
     # annotation_legend_side = 'right',
     # annotation_legend_list = list(clade_lgd),
     merge_legend = TRUE)

if(global_save_var){
  pdf(file = 'Manuscript_Fig_subfam_heatmap.pdf',
      height = 5,
      width = 5)
draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     # annotation_legend_side = 'right',
     # annotation_legend_list = list(clade_lgd),
     merge_legend = TRUE)
  dev.off()
}
```

Add chapl/aapl annotation to above

```{r chapl_aapl_ann}

chapl_aapl_name_vec <- ifelse(grepl(cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), 
                                                                                     cir_info$Geneid)], 
                                                  pattern = 'NA'),
                                            NA,
                                            cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), cir_info$Geneid)])

#chapl_aapl_name annotation

annotation_col <- chapl_aapl_name_col_num[ordering(chapl_aapl_name_vec,
                                                   order_vector = 'ChAPL')]

chapl_aapl_name_ha <- rowAnnotation(
  'CHAPL/AAPL' = anno_simple(
    chapl_aapl_name_vec,
    col = annotation_col,
    na_col = 'white'
  ),
  annotation_name_gp = gpar(fontsize = 10)
)

chapl_aapl_name_lgd <- Legend(
  title = 'CHAPL/AAPL',
  at = names(annotation_col)
,
  legend_gp = gpar(fill = na.omit(annotation_col))
)

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  # row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  row_title_rot = 0,
  row_title_gp = gpar(fontsize = 10),
  row_split = forcats::fct_shift(
    relevel(
      as.factor(combined_data_pir_ht$subfam),
      ref = 'ref')
  ),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)

if(global_save_var){
  pdf(file = 'Interest_SubfamSplit_ChAPL-AAPLAnnotate_heatmap.pdf',
      height = 5,
      width = 5)
draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)
  dev.off()
}

```

# Supp figure 9 - Pir expression of AAPLs and ChAPLs.

_pir_ genes selected by being expressed above 1 TPM in at least one sample.

```{r chapl_appl_heatmap_suppfig9}

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  mutate(chapl_aapl_name = cir_info$chapl_aapl_name_num[match(Geneid, cir_info$Geneid)]) %>% 
  filter(!str_detect(chapl_aapl_name, pattern = 'NA')) %>% 
  column_to_rownames(var = 'Geneid') %>% 
  filter_all(
      any_vars(. > 1)
  )

combined_data_ordered <- tibble(Geneid = rownames(combined_data_pir),
                                means = apply(combined_data_pir[,
                                                                str_detect(colnames(combined_data_pir), 
                                                                           'chapl_aapl',
                                                                           negate = TRUE)],
                                              1,
                                              mean),
                                chapl_aapl_name = factor(
                                  combined_data_pir$chapl_aapl_name
                                )) %>% 
  arrange(chapl_aapl_name, 
          desc(means)) %>% 
  .$Geneid

#Order the heatmap data by the order defined above (chapl_aapl_name, then mean expression).
combined_data_pir_ht <- combined_data_pir[
  order(
    match(rownames(combined_data_pir),
          combined_data_ordered),
    na.last = NA #Remove NAs
  ),
]

#Change the colnames to the times
colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                              pattern = '[[:digit:]]{2}h|chapl_aapl_name')

#Order by time point
combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                             drop = FALSE]

#Get the ref gene actual names instead of the Geneid
rownames(combined_data_pir_ht) <- ifelse(
  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
  rownames(combined_data_pir_ht)
)

#Get the colour range
maxcol <- ceiling(
  max(
    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name'])
  ))

#chapl_aapl_name name labels
num_chapl_aapl_name <- cir_info %>% 
  group_by(chapl_aapl_name) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_chapl_aapl_name <- setNames(paste0(num_chapl_aapl_name$chapl_aapl_name, 
                                          ' (n = ', num_chapl_aapl_name$num, ')'),
                                   num_chapl_aapl_name$chapl_aapl_name)

#loci annotation
chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
annotation_col <- chapl_aapl_name_col_num[ordering(labels_chapl_aapl_name,
                                                    order_vector = 'ChAPL')]

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  # row_title = NULL,
  row_title_rot = 0,
  row_title_gp = gpar(fontsize = 7),
  row_split = 
    factor(combined_data_pir_ht$chapl_aapl_name,
              levels = ordering(combined_data_pir_ht$chapl_aapl_name,
                                order_vector = 'ChAPL')),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  ),
  row_gap = unit(c(rep.int(1, times = 2),
                   3,
                   rep.int(1, times = 4)), "mm")
)

if(global_save_var){
  pdf(file = 'Manuscript_SuppFig_ChAPL-AAPL_heatmap.pdf',
      height = 5,
      width = 5)
  draw(ht, 
       row_title_side = 'left',
       column_title_side = 'bottom'
  )
  dev.off()
}

  draw(ht, 
       row_title_side = 'left',
       column_title_side = 'bottom'
  )
```

```{r}
chapl_aapl_name_ha <- rowAnnotation(
  'CHAPL/AAPL' = anno_simple(
    chapl_aapl_name_info,
    col = annotation_col,
  ),
  annotation_name_gp = gpar(fontsize = 10)
)

chapl_aapl_name_lgd <- Legend(
  title = 'CHAPL/AAPL',
  at = str_subset(
    sort(
      factor(labels_chapl_aapl_name,
             ordering(labels_chapl_aapl_name,
                      order_vector = 'ChAPL'))
    ),
    pattern = 'NA',
    negate = TRUE
  ),
  legend_gp = gpar(fill = na.omit(annotation_col))
)

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  row_km = 9,
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)

if(global_save_var){
  pdf(file = 'Interest_chapl-aapl_clustered_heatmap.pdf',
      height = 5,
      width = 5)
draw(ht, 
     row_title_side = 'left',
     column_title_side = 'bottom',
     heatmap_legend_side = 'right',
     annotation_legend_side = 'right',
     annotation_legend_list = list(chapl_aapl_name_lgd),
     merge_legend = TRUE)
  dev.off()
}
```

### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/fourier_transform.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

---
title: "fourier_transform_inR"
author: "Timothy Little"
date: "03/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Using fft to compute the fourier transform

Let's start with a simple sine wave.

```{r}
#Vector of a sine curve
sine_curve_simpl <- sin(
  seq(0, 
      2*pi, 
      0.01)
)
num_values_simpl <- length(sine_curve)
plot(sine_curve_simpl)
```

Now here is a more complicated wave.

```{r}
sine_curve_compl <- sin(
  seq(0, 
      2*pi, 
      0.01)
) + sin(
  7*seq(0, 
      2*pi, 
      0.01)
  )
num_values_compl <- length(sine_curve)
plot(sine_curve_compl)
```
Even this more complicated curve can be produced from a combination of sine waves. This is the case for any linear, regular wave. The fourier transform can look at a signal (e.g. the values that define the waveform, such as sine_curve_compl) and derive the sine wave components.

```{r}
sine_curve_compl_fourier <- fft(sine_curve_compl)
data.frame( 
  harmonic = 0:9,
  fourier_transform = head(
    round(sine_curve_compl_fourier, digits = 2),
    n = 10)
)
```

The output of the fourier transform is a series of complex numbers, each of which represents a sine wave function. Each one sequentially represents different harmonics, the first value is the 0th harmonic or the sine wave with 0Hz frequency. Next harmonic has 1Hz frequency etc. The 0Hz value doesn't have a periodicity, it's like a constant value added to the sine wave. The sine_curve_compl has tiny values for everything except the 1st and 7th harmonic. This is because, as you can see from the equation we used to generate the values for the curve, it is constructed from a 1Hz sine wave (i.e. a normal sine wave from 0 to 2pi, y = sin(x) ) and a 7Hz sine wave ( y = sin(7x) ).

The inverse fourier transform can get back the values of the curve from the fourier output. We need to normalise it by the length of the initial vector.

```{r}
sine_curve_compl_ifft <- fft(sine_curve_compl_fourier, 
                             inverse = TRUE)
#These still have a complex element so we should remove that so plot doesn't mess up and plot Real v Imaginary.
sine_curve_compl_ifft <- Re(sine_curve_compl_ifft)
plot(sine_curve_compl_ifft)
sine_curve_compl_ifft_norm <- sine_curve_compl_ifft/num_values_compl
plot(sine_curve_compl_ifft_norm)
```

The non-normalised plot has the same shape but the actual values are much larger.

# Spectral analysis.

Using an example from [this website](https://ms.mcmaster.ca/~bolker/eeid/2010/Ecology/Spectral.pdf) here.

```{r}
del<-0.1 # sampling interval
t <- seq(0,
         200,
         by=del)
x <- 2 + cos(2*pi*t/3) + 2*sin(2*pi*t/5)
plot(x, 
     type = 'l')
```

To create a 'periodogram' of this data we can use the function *spectrum()*.

This is like a graph of the strength of each harmonic with a line connected to each point. The strength of a harmonic is the modulus of the complex number that the fourier transform uses to define the sine wave.

```{r}
spectrum(x)
```

All the little blips don't mean much, so normally you smooth this by using the 'span' argument.

```{r}
spectrum(x, span = 10)
```

We can draw a more intuitive plot by turning off the log axis

```{r}
x.spec <- spectrum(x,
                   log="no",
                   span=10,
                   plot=FALSE)
spx <- x.spec$freq/del # Divide frequency by the sampling interval time - gets cycles per unit time.
spy <- 2*x.spec$spec # Multiply by 2 so that the area under curve equals variance
plot(spy~spx,
     xlab="frequency",
     ylab="spectral density",
     type="l",
     xlim = c(0,1))
```

Compute the phase shift of the harmonics.

```{r}
plot(x,
     type = 'l')
x.fft <- fft(x)/length(x)
row_num <- 150
data.frame(
  harmonic =seq(0,
                row_num - 1,
                1),
  fourier = x.fft[0:row_num],
  # real = Re(x.fft[0:row_num]),
  # imag = Im(x.fft[0:row_num]),
  amp = Mod(x.fft[0:row_num]),
  phase_times_pi = atan2(Im(x.fft[0:row_num]),
                       Re(x.fft[0:row_num]))/pi
) %>% 
  round(digits = 2) %>% 
  arrange(desc(amp))
```

```{r}
phase <- cos(0.5*t*pi + pi/2)

plot(phase,
     type = 'l')

spectrum(phase)

phase.spec <- spectrum(phase,
                   log="no",
                   span=10,
                   plot=FALSE)
spx<- phase.spec$freq/del # Divide frequency by the sampling interval time - gets cycles per unit time.
spy <- 2*phase.spec$spec # Multiply by 2 so that the area under curve equals variance
plot(spy~spx,
     xlab="frequency",
     ylab="spectral density",
     type="l",
     xlim = c(0,2))

phase.fft <- fft(phase)/length(phase)
row_num <- 1000
data.frame(
  harmonic =seq(0,
                row_num - 1,
                1),
  fourier = phase.fft[0:row_num],
  # real = Re(phase.fft[0:row_num]),
  # imag = Im(phase.fft[0:row_num]),
  amp = Mod(phase.fft[0:row_num]),
  phase_times_pi = atan2(Im(phase.fft[0:row_num]),
                         Re(phase.fft[0:row_num]))/pi
) %>% 
  round(digits = 5) %>% 
  arrange(desc(amp))
```

```{r}
t <- seq(0,200,by=0.1)
curve_vec <- sin(2*pi*t/16) + 0.4*sin(2*pi*t/5) + 1.5*sin(3*pi*t/10)
plot(curve_vec, type = 'l')
spectrum(curve_vec)
```

```{r}
plot.frequency.spectrum <- function(X.k, xlimits=c(0,length(X.k))) {
  plot.data  <- cbind(0:(length(X.k)-1), Mod(X.k))

  # TODO: why this scaling is necessary?
  plot.data[2:length(X.k),2] <- 2*plot.data[2:length(X.k),2]

  plot(plot.data, t="h", lwd=2, main="", 
       xlab="Frequency (Hz)", ylab="Strength", 
       xlim=xlimits, ylim=c(0,max(Mod(plot.data[,2]))))
}
```

### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_audreybrugat24hchabaudi.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

---
title: "TL19-07_Audrey_24h_chabaudi"
date: "17/06/2020"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

#for server
# knitr::opts_knit$set(root.dir = '~/data/tl19-07_chabaudi_24h_audreyMTbrugatSBP')
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/tl19-07_audreybrugat24hchabaudi")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

To take into account:

- SBP is more synchronous than MT, do we have counts?

- MT and SBP may seem very different due to differences in the amount of material for sequencing.

Suggestions:

- PCA of SBP on its own, and MT on its own, UMAP maybe too?

- Use DESeq2 to get the diff expr genes between time points and then use this to plot a PCA.

- z-score plots: Have a max/min expression heatmap instead of showing the median logTPM.

- Just remove the Unknown sample since it contributes nothing without an identity.

- The value of the expression threshold (variable: min_median) is very high (over 4 TPM) - is this reasonable?

Top_n_genes:

~~- Remove Unknown entirely by re-saving top_genes table. Also need to separate the top_n bit using rMT and SBP.~~

~~- Why is there a NULL printed right at the end? Issue with names~~

Questions:

- What day are these samples taken? d6?

- How many passages had 'SBP' been through?

- What are the differential counts of the samples taken? - Audrey's are on the Shared folder.

# Audrey's experimental description

24 Hours cycle P. chabaudi RNA seq.

Aim: Determine pir expression for each parasite stage in recently mosquito transmitted P. chabaudi parasites.

Principle: using a stabilate from recently mosquito transmitted P. chabaudi parasites infected C57BL/6 and extract RNA at each stage of parasite cycle (every 3 hours) during 24h.

Animals: 4 C57BL/6 J CD45.1 males for stabilates thawing (2 in RL, 2 in NL) and 27 C57BL/6 J CD45.1 females for experiments (12 in NL and 15 in RL) (issued 17/08/2016).

* AV16-35	2 C57BL/6 J CD45.1 males		RL.

* AV16-36	2 C57BL/6 J CD45.1 males		NL.

* AV16-37	15 C57BL/6 J CD45.1 females	RL.

* AV16-38	12 C57BL/6 J CD45.1 females	RL.

Experiment:

* d-14	11am	 Infected 2 mice (males) with recently mosquito transmitted parasite stabilate (MTC tower 6 box 6.1 position 31-65 (late troph)) in NL (AV16-36).

* d-14	5pm	Infecte 2 mice (males) with recently mosquito transmitted parasite stabilate (MTC tower 6 box 6.1 position 31-65 (late troph)) in RL (AV16-35).

* d-7	11am	pass AV16-35 into 15 mice (females) 105 iRBC/ mouse IP (AV16-37) RL.

* d0	 2pm	smear 3 mice and bleeding .

* d0	 2-5pm 	pellet parasites and trizol of 2pm samples.

* d0	 5pm	smear 3 mice and bleeding .

* d0	 5-8pm 	pellet parasites and trizol 5pm samples.

* d0	 8pm 	smear 3 mice and bleeding.

* d0	 8-11pm 	pellet parasites and trizol 8pm samples.

* d0	 11pm	smear 3 mice and bleeding .

* d0	11pm-2am 	pellet parasites and trizol 11pm samples.

* d1	 2am	smear 3 mice and bleeding .

* d1	 2-5am 	pellet parasites and trizol of 2am samples.

* d1	 5am	smear 3 mice and bleeding .

* d1	 5-8am 	pellet parasites and trizol 5pm samples.

* d1	 8am 	smear 3 mice in each group and bleeding .

* d1	 8-11am 	pellet parasites and trizol 8pm samples.

* d1	 11am	smear 3 mice in each group and bleeding .

* d1	 11am 	bleeding 3 mice (AV16-37) and parasite culture.

* d1	11am-2pm 	pellet parasites and trizol 11pm samples.

* d1	 4-6pm	schizonte pellet and trizol.

Dates:

d-14 : 25/08/2016.

d-7 : 01/09/2016.

d0 : 08/09/2016.

d1 : 09/09/2016.

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

# library('tidyverse')
library('readr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('DESeq2')
library("pheatmap")
library("RColorBrewer")
library('data.table')
library('gridExtra')
# library('apeglm')
# library('ashr')
library('tibble')
library('stringr')
# library('cowplot')
library('kableExtra')
library('viridis')
# library('VennDiagram')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('readxl') # for reading in Excel spreadsheets
library('plotly')
library('scales')
library('processx') # for downloading the plotly plots
library('cowplot')
library('writexl')

library('devtools')
#These packages were installed using github
library('ComplexHeatmap')
library('ggbiplot')
library('ggpattern')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select
```

```{r global_save_var}
#Make this TRUE if you want to run this script and save all the plots.
global_save_var <- TRUE
```

```{r tpm, results = 'hide', message = FALSE}
# # Calculating the transcript-per-million 
# 
# We need to write a function to calculate the TPM from the count data we have generated. This function uses the 'chabaudi_transcript_lengths' object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.

tpm = function(count_data, lengths = chabaudi_transcript_lengths) {

  #Print an error if the id columns do not match
  # first if statement to test that both have a GeneID column
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'Geneid', 
                                                  #This needs to be Geneid when some of the samples have 'id' in them
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      stop('Error: Count data and length gene ids DO NOT MATCH, please make sure they are ordered identically.')
    }
  } else {
    stop('Error: Count data and/or length table supplied do not have an id column.')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = T)) == T){
    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = T)]
  } else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
  }else{
    lengths.table <- lengths}
  lengths.table = matrix(lengths.table)
  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = T), drop = FALSE], 
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

```{r agree function}

agree = function(input_list){
  for(i in 1:(length(input_list)-1)){
    input_list[[i+1]] = input_list[[i]][ input_list[[i]] %in% input_list[[i+1]] ]
  }
  return(input_list[[length(input_list)]])
}

disagree = function(input_list){
  for(i in 1:(length(input_list)-1)){
    input_list[[i+1]] = input_list[[i]][ !input_list[[i]] %in% input_list[[i+1]] ]
  }
  return(input_list[[length(input_list)]])
}

mround <- function(x,base){ 
  base*round(x/base) 
} 
```

```{r function_ordering}
ordering <- function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}
```

```{r pca_function}
pca_function <- function(input_ct, 
                         genevec = NULL, 
                         scale.pca = TRUE, 
                         three.dimensional = FALSE,
                         legend_settings = list(),
                         colornames = NULL){

  twentyfour_tpm_pca <- input_ct$tpm
  if(!is.null(genevec)){
    twentyfour_tpm_pca <- input_ct$tpm %>% filter(Geneid %in% genevec)

    ### Re-scaling the TPM data
    twentyfour_tpm_pca_rescale <- lapply(colnames(twentyfour_tpm_pca), 
                                         function(column_name){
                                           column <- getElement(twentyfour_tpm_pca, 
                                                                paste(column_name))
                                           if(is.numeric(column)){
                                             #is.numeric to avoid trying to scale the Geneid column
                                             tpm_scale_factor <- 1e6 / sum(column)
                                             tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                           }else{
                                             tpm_scaled <- column
                                           }
                                           return(tpm_scaled)
                                         }) %>% 
      bind_cols

    colnames(twentyfour_tpm_pca_rescale) <- colnames(twentyfour_tpm_pca)
    twentyfour_tpm_pca <- twentyfour_tpm_pca_rescale
    ###

  }
  rownames(twentyfour_tpm_pca) <- twentyfour_tpm_pca$Geneid
  twentyfour_tpm_pca <- select(twentyfour_tpm_pca, -Geneid) %>% 
    prcomp(center = TRUE, scale. = scale.pca)
  #Need to look into what exactly scale and center are doing
  twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
  twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)

  twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                          PC1 = twentyfour_tpm_pca_coords$Dim.1,
                                          PC3 = twentyfour_tpm_pca_coords$Dim.3,
                                          PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
    dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
    dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
    dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])

  if(is.null(colornames)){
    # colornames <- c(cbPalette, 'black', 'red')
    colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
    colornames <- setNames(colornames, 
                           sort(unique(twentyfour_tpm_pca_coords$time)))
  }

  pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
  pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
  pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
  pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)

  if(three.dimensional){
    pl <- plot_ly(twentyfour_tpm_pca_coords, 
                  x  = ~PC1, y = ~PC2, z = ~PC3,  mode = 'markers',
                  type = 'scatter3d', color = ~time, text = ~sample,
                  colors = colornames,
                  # size = ~total_features,
                  marker = list(opacity = 1, size = 5),
                  symbol = ~transmission, 
                  symbols = c('circle','x')) %>% 
      layout(scene = list(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
                          yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
                          zaxis = list(title = paste0('PC3 ',pc3_importance,'%'))))
  }else{
    pl <- plot_ly(twentyfour_tpm_pca_coords, 
                  x  = ~PC1, y = ~PC2,
                  type = 'scatter', mode = 'markers',
                  color = ~time, text = ~sample,
                  colors = colornames,
                  # size = ~total_features,
                  marker = list(opacity = 1, size = 16),
                  symbol = ~transmission, symbols = c('circle','x')) %>% 
      layout(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
             yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
             legend = legend_settings)
  }
  pl
}
```

```{r function_draw_heatmap_chabaudi_24h}
# Function to draw a heatmap with a list of genes

order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")

# Note that there may seem to be discrepencies between the data and the heatmap because this function may be using the median and/or mean value of the TPM
draw_heatmap_chabaudi_24h <- function(genelist, 
                                scale = 'max', 
                                threshold = 0, 
                                datasets = twentyfour_ct, 
                                filter_tpm = NULL,
                                stage_diff = NULL, 
                                filter_stage = NULL, 
                                title_name = NULL, 
                                drop_stage = NULL, 
                                log_it = TRUE,
                                collate_samples_by_stage = TRUE,
                                order_by_med_tpm = TRUE,
                                mean_or_median = 'mean',
                                gene_names_for_title = NULL,
                                ...){
  #genelist <- cir_id
  #For de-bugging:

  # scale = 'max';
  # threshold = 0;
  # datasets = twentyfour_ct;
  # filter_tpm = NULL;
  # stage_diff = NULL;
  # filter_stage = NULL;
  # title_name = NULL;
  # drop_stage = NULL;
  # log_it = TRUE;
  # collate_samples_by_stage = TRUE;
  # order_by_med_tpm = TRUE;
  # mean_or_median = 'mean';

  # if(!is.null(datasets$tpm)){
  #   #Probably means only one dataset was input so lapply will malfunction
  #   datasets <- list(tpm = datasets)
  # }

  if(log_it == TRUE){
    threshold <- log1p(threshold)
    title <- 'log(TPM + 1)'
  }else{
    title <- 'TPM'
  }

  expt_tpms_top_genes = lapply(datasets, function(tpm_input){
    # tpm_input <- datasets[[1]]
    tpms = tpm_input$tpm %>% 
      #Melt the data frame for using dplyr
      filter(Geneid %in% genelist) %>% 
      melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm')

    if(collate_samples_by_stage){ 
      #Create a new variable 'stage' from the sample names without rep numbers
      tpms <- mutate(
        tpms, 
        stage = str_extract(
          sample,
          pattern = '.+(?=_[[:digit:]]{1,2})'
        )
      )
    }else{
      tpms <- mutate(tpms, 
                     stage = sample)
    }

    tpms <- mutate(tpms, time = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
      #Convert the times given in 'h' to robustly coded times in 24h clock
      mutate(time = ifelse(grepl(time, pattern = '[[:digit:]]h$'), 
                           strptime(time, format="%Hh") %>% format(., "%Hh"),
                           time)) %>% 
      mutate(time = ifelse(grepl(time, pattern = '[[:digit:]](a|p)m'), 
                           strptime(time, format="%I%p") %>% format(., "%Hh"),
                           time))

    colnam = sort(unique(tpms$stage))

    if(mean_or_median == 'median'){
      tpms = tpms %>% 
        group_by(Geneid, stage, time) %>%  
        dplyr::summarise(mean_tpm = median(tpm)) %>%  
        group_by(stage) %>% 
        recast(Geneid ~ stage + time + 'mean_tpm', median) #recast into short-form for complex heatmap
    } else if (mean_or_median == 'mean'){
      tpms = tpms %>% 
        group_by(Geneid, stage, time) %>%  
        dplyr::summarise(mean_tpm = mean(tpm)) %>%  
        group_by(stage) %>% 
        recast(Geneid ~ stage + time + 'median_tpm', mean) #recast into short-form for complex heatmap
    }

    colnames(tpms) = c('Geneid',as.character(colnam))

    if(!is.null(order_vector)){
      tpms <- tpms[,
                   ordering(colnames(tpms),
                            c('Geneid',
                              order_vector))]
      #Ordering using the ordering function and order_vector defined in the functions section.
    }

    #Apply drop_stage
    if(!is.null(drop_stage)){
      tpms <- select(tpms, -!!(paste(drop_stage)))
    }
    #Apply stage_diff to get genes which are expressed higher in one stage than another
    if(!is.null(stage_diff)){
      if(is.na(stage_diff[1])|is.na(stage_diff[2])){
        stop('stage_diff must be a vector of two named stages to select genes with TPM greater in the former than the latter')
      }
      tpms <- tpms[select(tpms, paste(stage_diff[1])) > select(tpms, paste(stage_diff[2])), ]
    }
    #Apply filter_tpm
    if(!is.null(filter_tpm)){
      tpms <- filter_at(tpms, vars(-Geneid), any_vars(. > filter_tpm))
    }
    #Apply filter_stage to remove genes expressed below filter_tpm in that stage
    if(!any(is.null(filter_stage), is.null(filter_tpm))){
      tpms <- filter(tpms, paste(filter_stage) >= filter_tpm)
    }
    #Remove PCHAS_ from the gene names for simplicity
    # rownames(tpms) = gsub(as.character(tpms$Geneid), pattern = 'PCHAS_', replacement = '')
    #Use the protein names for ease of understanding
    rownames(tpms) = chabaudi_gene_info$handy_descriptions[match(tpms$Geneid, 
                                                           chabaudi_gene_info$Geneid)]
    tpms = select(tpms, -Geneid)

    if(log_it == TRUE){
      tpms <- log1p(tpms)
    }

    if(tpm_input$experiment %>%  is.null){
      tpm_input$experiment = ''
    }
    list = list(tpms, tpm_input$experiment)
    names(list) = c('tpm','experiment')
    return(list)
  })

  if(
    !all(
      unlist(
        lapply(
          expt_tpms_top_genes, 
          function(x) all(
            x$tpm < threshold
          )
        )
      )
    )
  ){
    if(scale %in% 'max'){
      maxcol <- my_max(unlist(lapply(expt_tpms_top_genes, 
                                     function(tpms){my_max(tpms$tpm)})))
    }else{
      maxcol = scale
    }
  }

  if(is.na(maxcol)|maxcol<1){maxcol <- 1}

  if(order_by_med_tpm){
    #Work out the order by median TPM
    medians_across_expt_mat <- as.matrix(
      sapply(
        lapply(
          expt_tpms_top_genes, 
          function(x){
            rowMedians(as.matrix(x$tpm),
                       na.rm = TRUE) 
          }
        ), 
        rbind
      )
    )
    tpm_order <- order(
      rowMedians(
        medians_across_expt_mat, 
        na.rm = TRUE),
      decreasing = TRUE
    )
  }

  #Applying the threshold last because NAs mess things up
  expt_tpms_top_genes <- lapply(expt_tpms_top_genes, function(expt){
    tpms <- expt$tpm
    tpms = apply(tpms, 2, function(x){
      ifelse(
        x >= threshold, 
        x, 
        NA
      )
    }
    ) 
    #Get NA for values under threshold
    list = list(tpms, expt$experiment)
    names(list) = c('tpm','experiment')
    return(list)
  })

  heatmaps = lapply(expt_tpms_top_genes, function(tpms){
    if(tpms$experiment != last(expt_tpms_top_genes)$experiment){
      #If it is the final element don't remove the row names
      rownames(tpms$tpm) = c()
    }
    #Setting the name if one is not provided
    if(is.null(title_name)){
      title_name = tpms$experiment
    }

    #Re-order both data.frames by the median highest expressed pirs
    if(order_by_med_tpm){
      tpms$tpm <- tpms$tpm[tpm_order,, drop = FALSE]
    }

    # print(nrow(tpms$tpm))
    #Drawing each heatmap
    ht = Heatmap(tpms$tpm, 
                 name = 'TPM', 
                 col = colorRamp2(c(threshold, 
                                    maxcol/4, 
                                    maxcol/2, 
                                    3*maxcol/4, 
                                    maxcol), 
                                  viridis(5)),
                 cluster_rows = FALSE, 
                 cluster_columns = FALSE,
                 column_title = paste(title_name),
                 column_title_gp = gpar(fontsize = 7),
                 column_names_gp = gpar(fontsize = 7),
                 row_names_gp = gpar(fontsize = ifelse(
                   nrow(tpms$tpm) > 13, 
                   12*(13/nrow(tpms$tpm)), 
                   12)
                 ),
                 na_col = 'white', border = TRUE,
                 heatmap_legend_param = list(
                   title = paste(title), 
                   at = round(
                     c(threshold, 
                       maxcol/4, 
                       maxcol/2, 
                       3*maxcol/4, 
                       maxcol), 
                     digits = 1)
                 )
                 # ,...
    )
  })

  ht_list = Reduce('+', heatmaps) #To add the list elements together
  draw(ht_list, 
       row_title_side = 'left',
       column_title = bquote(~italic('P. chabaudi')~'24h asexual life cycle stage - '~italic(.(gene_names_for_title))~'genes'), 
       column_title_side = 'bottom',
       heatmap_legend_side = 'left')
}
```

```{r plotly_biplot_function}
plot_ly_biplot <- function(pca_object = twentyfour_tpm_pca, 
                           prin_components = c(1,2,3),
                           fontsize = 7,
                           scale.loads = 600,
                           separate_multigene = FALSE,
                           three_dimensions = FALSE,
                           col_var = 'time',
                           gene_info = chabaudi_gene_info,
                           textposition = "top right"){

  #Check that three components are provided for three dimensional biplot
  if(three_dimensions){
    if(length(prin_components) != 3){
      stop('For three dimensions, three principal components must be given for prin_components argument')
    }
  }

  pca_object_rotations <- rownames_to_column(data.frame(pca_object$rotation), 
                                             var = 'sample')  %>% 
    dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
    dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
    dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])

  #Setting the colours, which come from each life cycle experiment, called 'stage'.

  colornames <- setNames(
    viridis(
      length(
        unique(
          getElement(
            pca_object_rotations, 
            paste(col_var)
          )))),
    sort(
      unique(
        getElement(
          pca_object_rotations, 
          paste(col_var)
        ))))

  pca_object_loadings <- data.frame(Geneid = rownames(pca_object$x),
                                    pca_object$x)

  pca_object_loadings$Geneid <- paste0(
    str_replace(
          as.character(pca_object_loadings$Geneid),
          pattern = ' ',
          replacement = ''),
    ' ',
    gene_info$handy_descriptions
    [
      match(
        c(str_replace(
          as.character(pca_object_loadings$Geneid),
          pattern = ' ',
          replacement = '')
        ),
        gene_info$Geneid
      )
      ]
  )

  if(separate_multigene){
    #If the gene is a fam or a fam gene we want to know - haven't excluded pseudogenes here
    pca_object_loadings$multigene <- ifelse(
      #if it doesn't have fam or fam-something in its name then just say gene
    is.na(
      str_extract(
        pca_object_loadings$Geneid,
        pattern = 'fam-.|fam'
      )
    ),
    'Gene loadings',
    paste0(
      str_extract(
        pca_object_loadings$Geneid,
        pattern = 'fam-.|fam'
      ),
      ' gene loadings'
    )
  )
}else{
  pca_object_loadings$multigene <- rep('Gene loadings',
                                       stages = nrow(pca_object_loadings))
}

#Degrees of variation for the axes
pca_importance <- as.data.frame(summary(pca_object)$importance)
pc_importance_1 <- round(pca_importance[2,prin_components[1]]*100, digits = 1)
pc_importance_2 <- round(pca_importance[2,prin_components[2]]*100, digits = 1)
if(three_dimensions){
  pc_importance_3 <- round(pca_importance[2,prin_components[3]]*100, digits = 1)
}

#scale.loads controls how long the lines are, the longer they are the more prominent the loadings but the less
# obvious the genes.

pl <- plot_ly()

#For the font of the labels
t <- list(
  size = fontsize
  )

#For loop to add the PCA lines of the samples
for(k in 1:nrow(pca_object_rotations)) {
  #Have 0 so that the line comes from the origin (0,0)
  #PC first
  x <- c(0, pca_object_rotations[k,(prin_components[1]+1)])*scale.loads
  #PC second
  y <- c(0, pca_object_rotations[k,(prin_components[2]+1)])*scale.loads

  if(three_dimensions){
    #PC third
    z <- c(0, pca_object_rotations[k,(prin_components[3]+1)])*scale.loads
    df_plotly <- data.frame(x = x, 
                            y = y, 
                            z=z,
                            stage = getElement(pca_object_rotations, 
                                               paste(col_var))[k],
                            sample = pca_object_rotations$sample[k])
  }else{
    df_plotly <- data.frame(x = x, 
                            y = y,
                            stage = getElement(pca_object_rotations, 
                                               paste(col_var))[k],
                            sample = pca_object_rotations$sample[k])
  }

  #colour based on sample
  color <- colornames[
    match(
      getElement(
        pca_object_rotations, 
        paste(col_var)
      )[k], 
      rev(names(colornames)
      )
    )
    ]

  #Add these to the plot
  #To avoid duplication in the legend don't show it if the name is duplicated
  if(duplicated(getElement(pca_object_rotations, paste(col_var)))[k]){
    showlegend <- FALSE
  }else{
    showlegend <- TRUE
  }

  #Add the lines for the samples depending on their position in the PCA
  if(three_dimensions){
    pl <- pl %>%
      add_trace(data = df_plotly,
                x=~x,
                y=~y,
                z=~z,
                type='scatter3d',
                mode='line',
                color = ~stage,
                colors = colornames,
                line = list(width = 2),
                text = pca_object_rotations[k,1],
                name = getElement(pca_object_rotations, paste(col_var))[k],
                opacity = 1,
                showlegend = showlegend,
                legendgroup = getElement(pca_object_rotations, paste(col_var))[k],
                #Note that 3d plots only accept a few shapes e.g. hexagon is not accepted
                symbols = c('circle', 
                            'circle-open',
                            'square',
                            'square-open',
                            'x',
                            'diamond')) %>% 
            add_text(x = df_plotly$x[2],
               y = df_plotly$y[2],
               z = df_plotly$z[2],
               text = df_plotly$sample, 
               textfont = t,
               textposition = textposition,
               showlegend = ifelse(k == 1, TRUE, FALSE), #Only want this option to be TRUE once
               legendgroup = 'Toggle text',
               name = 'Toggle text')
  }else{
    pl <- pl %>% 
      add_trace(data = df_plotly,
                x=~x,
                y=~y,
                type='scatter',
                mode='line',
                color = ~stage,
                colors = colornames,
                line = list(width = 2),
                text = pca_object_rotations[k,1],
                name = getElement(pca_object_rotations, 
                                  paste(col_var))[k],
                opacity = 1,
                showlegend = showlegend,
                legendgroup = getElement(pca_object_rotations, 
                                         paste(col_var))[k],
                symbols = c('circle', 
                            'circle-open',
                            'square',
                            'square-open',
                            'x',
                            'diamond')) %>% 
      #I don't know why it only works if we put the symbols argument here and not in the add_trace later??
      add_text(x = df_plotly$x[2],
               y = df_plotly$y[2],
               text = df_plotly$sample,
               textfont = t,
               textposition = textposition,
               showlegend = ifelse(k == 1, TRUE, FALSE), #Only want this option to be TRUE once
               legendgroup = 'Toggle text',
               name = 'Toggle text')
  }
}

#Adding the gene loadings

if(three_dimensions){
  #Make data frame of the gene loadings
  pca_object_loadings_df <- select(data.frame(pca_object_loadings), c(Geneid, 
                                                                      multigene, 
                                                                      paste0('PC', prin_components[1]),
                                                                      paste0('PC', prin_components[2]),
                                                                      paste0('PC', prin_components[3])))
  colnames(pca_object_loadings_df) <- c('Geneid',
                                        'multigene',
                                        'prin_com_1',
                                        'prin_com_2',
                                        'prin_com_3')
  pl <- pl %>% 
    add_trace(data = pca_object_loadings_df, 
              inherit = FALSE,
              x = ~prin_com_1, 
              y = ~prin_com_2, 
              z = ~prin_com_3,
              type = 'scatter3d', 
              mode = 'markers',
              marker = list(color = 'black',
                            size = 4,
                            opacity = 0.7),
              text = ~Geneid,
              name = ~multigene,
              symbol = ~multigene,
              showlegend = TRUE,
              legendgroup = ~multigene) %>% 
    layout(
      scene = list(
        xaxis = list(
          title = paste0(
            'PC', 
            prin_components[1], 
            ' ', 
            pc_importance_1,
            '%'
          )
        ),
        yaxis = list(
          title = paste0(
            'PC', 
            prin_components[2], 
            ' ',
            pc_importance_2,
            '%'
          )
        ),
        zaxis = list(
          title = paste0(
            'PC', 
            prin_components[3], 
            ' ',
            pc_importance_3,
            '%'
          )
        )
      )
    )
    }else{
      #Make data frame of the gene loadings
      pca_object_loadings_df <- select(data.frame(pca_object_loadings), 
                                       c(Geneid, 
                                         multigene, 
                                         paste0('PC', prin_components[1]),
                                         paste0('PC', prin_components[2])))
      colnames(pca_object_loadings_df) <- c('Geneid',
                                            'multigene',
                                            'prin_com_1',
                                            'prin_com_2')
      pl <- pl %>% 
        add_trace(data = pca_object_loadings_df, 
                  inherit = FALSE,
                  x = ~prin_com_1, 
                  y = ~prin_com_2,
                  type = 'scatter', 
                  mode = 'markers',
                  marker = list(color = 'black',
                                size = 6,
                                opacity = 0.9),
                  text = ~Geneid,
                  name = ~multigene,
                  symbol = ~multigene,
                  showlegend = TRUE,
                  legendgroup = ~multigene) %>% 
        layout(
          xaxis = list(
            title = paste0(
              'PC', 
              prin_components[1], 
              ' ', 
              pc_importance_1,
              '%'
            )
          ),
          yaxis = list(
            title = paste0(
              'PC', 
              prin_components[2], 
              ' ',
              pc_importance_2,
              '%'
            )
          )
        )
    }
pl
}

#For debugging
# k <- 6

#Switching around the add_text and add_trace for the 'Toggle text' (so that it goes first) seems to bugger up the colours
```

```{r function_mround}
mround <- function(x,base){ 
  base*ceiling(x/base) 
} 
```

```{r function_my_max}
my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
```

```{r function_uniquify}
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
    name <- paste(vector_names[i],
                  df_num$rep[match(vector_names[i], df_num$names)],
                  sep = '_')
    #Increase rep counter
    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
    new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r function_firstup}
#Capitalize first letter
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
```

```{r function_density_samples}
density_samples = function(input, save = FALSE){
  #Get some colours for the lines
  # colours = rainbow_hcl(length(colnames(input$tpm)[-1]))
  colours = rep_len(cbPalette, length.out = length(colnames(input$tpm)[-1]))
  list_dens = lapply(colnames(input$tpm)[-1], function(sample){
    data = as.numeric(unlist(select(input$tpm, paste(sample))))
    dens = density(log1p(data), bw = bw_mean)
    #Use the index of the column to get the colour
    colouring = colours[grep(colnames(input$tpm)[-1], pattern = paste(sample))]
    list_return = list(dens, colouring)
    names(list_return) = c('dens','colouring')
    list_return
  })
  max_dens = max(unlist(lapply(list_dens, function(x){max(x$dens$y)})))
  max_x = max(unlist(lapply(list_dens, function(x){max(x$dens$x)})))
  line_width = 1
  par(new=FALSE)
  if(save){
    pdf(file = paste0(input$experiment, '_tl19-07_densityplot_fixedbw.pdf'), 
        width = 15/2.54, 
        height = 11.5/2.54)
  }
  lapply(seq_along(list_dens), function(index){
    x = list_dens[[index]]
    if(index <= 9){
      plot(x$dens, 
           xlim = c(0, max_x), 
           ylim = c(0,max_dens), 
           type = 'l', 
           col = x$colouring,
           main = paste(input$experiment),
           lwd = line_width)
    }else if(index <= 18){
      plot(x$dens, 
           xlim = c(0, max_x), 
           ylim = c(0,max_dens), 
           type = 'l', 
           col = x$colouring,
           main = paste(input$experiment),
           lwd = line_width,
           lty = 2)
    } else {
      plot(x$dens, 
           xlim = c(0, max_x), 
           ylim = c(0,max_dens), 
           type = 'l', 
           col = x$colouring,
           main = paste(input$experiment),
           lwd = line_width,
           lty = 3)
    }
    abline(v = log1p(min_median))
    legend(max_x-4, max_dens, 
           legend = paste(colnames(input$tpm)[-1]), 
           col = colours,
           lty = c(rep(1, times = 9), rep(2, times = 9), rep(3, times = 9)),
           cex = 0.5)
    par(new=TRUE)})
  if(save){
    dev.off()
  }
}
```

```{r function_total_tpm_expression}
total_tpm_expression <- function(list_genes, 
                                 save = FALSE,
                                 tpm_data = twentyfour_ct,
                                 stage_str_extract_regex_pattern = '(?<=_).+(?=_[[:digit:]]+)|sch.culture'){
  lapply(names(list_genes), 
         function(list_name){
           paste(list_name)
           gene_id <- getElement(list_genes, paste(list_name))
           list_tpms_line <- lapply(tpm_data, 
                                    function(input){
             tpm_data = input$tpm %>% 
               filter(Geneid %in% gene_id) %>% 
               melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
               mutate(stage = str_extract(sample, 
                                          pattern = paste(stage_str_extract_regex_pattern))) %>% 
               group_by(stage, sample) %>% 
               dplyr::summarise(total_tpm = sum(tpm))

             #All genes detected
             tpm_data_all_genes <- input$tpm %>% 
               melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
               mutate(stage = unlist(lapply(strsplit(as.character(sample), split = '_'),
                                            `[`,
                                            1))) %>% 
               # filter(tpm > threshold) %>% 
               group_by(stage, sample) %>% 
               dplyr::summarise(total_tpm = n()) %>% 
               ungroup %>%  
               group_by(stage) %>% 
               dplyr::summarise(total_tpm_median = median(total_tpm),
                                Number_of_replicates = n(),
                                total_tpm_max = max(total_tpm),
                                total_tpm_min = min(total_tpm)) %>% 
               arrange(stage)

             #Make stages with no cirs in the sample give a zero 
             if(!all(colnames(input$tpm) %in% tpm_data$sample)){
               missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
               missing_stages_data_zero = data.frame(lapply(missing_stages, function(missing_st){
                 data.frame(stage = unlist(str_extract(missing_st, 
                                                       pattern = paste(stage_str_extract_regex_pattern))),
                            sample = missing_st,
                            total_tpm = 0)
               }))
               tpm_data = bind_rows(tpm_data, missing_stages_data_zero)
             }

             #Add experiment column
             tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
             max_tpm = if(max(tpm_data$total_tpm) < 10){
               c(seq(0,mround(max(tpm_data$total_tpm),1),1))
             }else{
               c(seq(0,mround(max(tpm_data$total_tpm),5),5))
             }

             #Get the average data for the line graph
             tpm_data_line = tpm_data %>% 
               ungroup %>% 
               group_by(stage, experiment) %>% 
               dplyr::summarise(total_tpm_median = median(total_tpm),
                                Number_of_replicates = n(),
                                total_tpm_max = max(total_tpm),
                                total_tpm_min = min(total_tpm)) %>% 
               arrange(stage)

             tpm_data_line$total_tpm_median_allgenes <- tpm_data_all_genes$total_tpm_median

             return(tpm_data_line)
           })

           tpm_data_line = bind_rows(list_tpms_line)
           #Apply ordering if it is specified
           if(!is.null(order_vector)){
             tpm_data_line$stage = factor(tpm_data_line$stage, 
                                          levels = ordering(tpm_data_line$stage,
                                                            order_vector))
           }
           # tpm_data_line$threshold = rep(threshold, nrow(tpm_data_line))

           tpm_data_line

           #Calculate the range for the y value to make it look nicer
           if(max(
             tpm_data_line$total_tpm_max
           ) >= 500){
             max_tpm <- c(
               seq(
                 0,
                 mround(
                   max(
                     tpm_data_line$total_tpm_max
                   ),
                   500),
                 length.out = 11)
             )
           }else{
             max_tpm <- c(
               seq(
                 0,
                 mround(
                   max(
                     tpm_data_line$total_tpm_max
                   ),
                   10),
                 length.out = 11)
             )
           }
           # max_tpm <- seq(
           #       0,
           #       mround(
           #         max(
           #           tpm_data_line$total_tpm_max),
           #         500),
           #       by = 500)
           y = ggplot(data = tpm_data_line, aes(x = stage, y = total_tpm_median, 
                                                fill = experiment, group = experiment)) +
             geom_bar(position = position_dodge2(preserve = 'single'),
                      stat = 'identity') +
             scale_fill_manual(values = c(cbPalette, 'red', 'purple')[1:(length(unique(tpm_data_line$experiment)))]) +
             geom_errorbar(aes(x = stage,
                               ymax = total_tpm_max, 
                               ymin = total_tpm_min),
                           position = position_dodge2(preserve = 'single')) +
             scale_y_continuous(breaks = max_tpm) +
             theme_classic() +
             theme(axis.title.x=element_blank(),
                   axis.text.x=element_text(angle = 90, 
                                            size = 10, 
                                            vjust = 0.5),
                   axis.text.y = element_text(size = 5),
                   axis.title.y = element_text(size = 8),
                   legend.title = element_text(size = 7),
                   legend.text = element_text(size = 5),
                   legend.key.size = unit(3,'mm'),
                   legend.key.width = unit(2,'mm')) +
             scale_size(range = c(0,3)) +
             ylab(
               bquote('total '~italic(.(list_name))~' TPM')
             )
           if(save == TRUE){
             ggsave2(paste0('tl20-05_',
                            gsub(list_name, 
                                 pattern  = ' ', 
                                 replacement = '_'),
                            '_totalTPM_barchart.pdf'),
                     plot = y,
                     dpi = 300,
                     width = 15,
                     height = 10,
                     units = 'cm')
           }
           y
         })
}
```

```{r function_total_tpm_expression_stacked}
total_tpm_expression_stacked <- function(
  list_genes, 
  save = FALSE,
  tpm_data = twentyfour_ct,
  stage_str_extract_regex_pattern = '(?<=[[:digit:]]h_).+(?=_[[:digit:]]+)|sch.culture',
  stack_genes = cir_id,
  stack_names = c('Yes',
                  'No'),
  stack_label = 'Cir genes',
  stack_grouping = TRUE
){
  lapply(names(list_genes), 
         function(list_name){
           paste(list_name)
           gene_id <- getElement(list_genes, paste(list_name))
           list_tpms_line <- lapply(tpm_data, 
                                    function(input){
                                      tpm_data <- input$tpm %>% 
                                        filter(Geneid %in% gene_id) %>% 
                                        melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
                                        mutate(stage = str_extract(sample, 
                                                                   pattern = paste(stage_str_extract_regex_pattern)),
                                               # stack_group = ifelse(
                                               #   Geneid %in% stack_genes, #If the gene is in stack_genes
                                               #   ifelse(
                                               #     stack_grouping, #If stack_grouping is TRUE, use the group name from stack_names (the first element), otherwise use the gene name (useful is stack_genes is big e.g. cir_id)
                                               #     stack_names[1],
                                               #     Geneid
                                               #   ),
                                               #   stack_names[2]) #stack_names[2] is the negative name e.g. 'No'
                                               stack_group = case_when(
                                                 Geneid %in% stack_genes & stack_grouping ~ stack_names[1],
                                                 Geneid %in% stack_genes & !stack_grouping ~ Geneid,
                                                 TRUE ~ stack_names[2]
                                               )
                                        ) %>% 
                                        group_by(stage, sample, stack_group) %>% 
                                        dplyr::summarise(total_tpm = sum(tpm))

                                      #Make stages with no cirs in the sample give a zero 
                                      if(!all(colnames(input$tpm) %in% tpm_data$sample)){
                                        missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
                                        missing_stages_data_zero = data.frame(lapply(missing_stages, 
                                                                                     function(missing_st){
                                                              data.frame(stage = unlist(str_extract(missing_st, 
                                                                                                    pattern = paste(stage_str_extract_regex_pattern))),
                                                                         sample = missing_st,
                                                                         total_tpm = 0)
                                                            })
               )
               tpm_data = bind_rows(tpm_data, missing_stages_data_zero)
             }

             #Add experiment column
             tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
             max_tpm = if(max(tpm_data$total_tpm) < 10){
               c(seq(0,mround(max(tpm_data$total_tpm),1),1))
             }else{
               c(seq(0,mround(max(tpm_data$total_tpm),5),5))
             }

             #Get the average data for the line graph
             tpm_data_line = tpm_data %>% 
               ungroup %>% 
               group_by(stage, experiment, stack_group) %>% 
               dplyr::summarise(total_tpm_median = median(total_tpm),
                                Number_of_replicates = n(),
                                total_tpm_max = max(total_tpm),
                                total_tpm_min = min(total_tpm)) %>% 
               arrange(stage)

             return(tpm_data_line)
           })

           tpm_data_line <- bind_rows(list_tpms_line)
           #Apply ordering if it is specified
           if(!is.null(order_vector)){
             tpm_data_line$stage = factor(tpm_data_line$stage, 
                                          levels = ordering(tpm_data_line$stage,
                                                            order_vector))
           }

           tpm_data_line

           #Calculate the range for the y value to make it look nicer
           if(max(
             tpm_data_line$total_tpm_max
           ) >= 500){
             max_tpm <- c(
               seq(
                 0,
                 mround(
                   max(
                     tpm_data_line$total_tpm_max
                   ),
                   500),
                 length.out = 11)
             )
           }else{
             max_tpm <- c(
               seq(
                 0,
                 mround(
                   max(
                     tpm_data_line$total_tpm_max
                   ),
                   10),
                 length.out = 11)
             )
           }
           # max_tpm <- seq(
           #       0,
           #       mround(
           #         max(
           #           tpm_data_line$total_tpm_max),
           #         500),
           #       by = 500)
           y <- ggplot(data = tpm_data_line, 
                       aes(
                         x = experiment,
                         y = total_tpm_median, 
                         fill = stack_group
                       )
           ) +
             geom_bar(position = position_stack(),
                      stat = 'identity') +
             scale_fill_manual(
               limits = unique(tpm_data_line$stack_group),
               values = setNames(cbPalette[1:length(unique(tpm_data_line$stack_group))],
                                 unique(tpm_data_line$stack_group))
               #Need to have limits and values otherwise when there are only cir genes (i.e. only 'Yes') the colour 
               # palette will mess up and only use one, wrong colour. Limits tells it there are always two values.
             ) +
             labs(fill = stack_label) +
             # scale_y_continuous(breaks = max_tpm) +
             theme_classic() +
             theme(axis.title.x=element_blank(),
                   axis.text.x=element_text(angle = 90, 
                                            size = 10, 
                                            vjust = 0.5),
                   axis.text.y = element_text(size = 5),
                   axis.title.y = element_text(size = 8),
                   legend.title = element_text(size = 7),
                   legend.text = element_text(size = 5),
                   legend.key.size = unit(3,'mm'),
                   legend.key.width = unit(2,'mm')) +
             scale_size(range = c(0,3)) +
             ylab(
               bquote('total '~italic(.(list_name))~' TPM')
             ) + 
             facet_grid(.~stage) +
             theme(strip.text.x = element_text(size=4, face = 'bold', angle = 45),
                   strip.background = element_rect(colour="black", fill="lightgrey"))
           if(save == TRUE){
             ggsave2(paste0('tl20-08_',
                            gsub(list_name, 
                                 pattern  = ' ', 
                                 replacement = '_'),
                            '_totalTPM_barchart.pdf'),
                     plot = y,
                     dpi = 300,
                     width = 15,
                     height = 10,
                     units = 'cm')
           }
           y
         })
}
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info = read.csv('chabaudi_PccAS_PlasmoDBv44_geneinformation_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)
#.txt file is weirdly truncated, don't use it

chabaudi_transcript_lengths = select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID)

cir_info = chabaudi_gene_info %>% 
  select(Gene.ID, Product.Description, Transcript.Length) %>% 
  filter(Product.Description == 'PIR protein')
nrow(cir_info)

#Brugat 2017 cirs
list.files()
brugat_cir_names = read_xlsx('chabaudi_pir_cir_genes_brugat_2017_nmicrobiol2016276-s4.xlsx') %>% 
  .$id
length(brugat_cir_names)
#Two extra genes compared to from PlasmoDB

#Adam Reid Personal Comm 2019 cirs
reid_cir_clade = read.table('AdamReid_chabaudi_PccAS_v3_aa.pir.clades.simple.txt',
                            header = FALSE,
                            stringsAsFactors = FALSE,
                            sep = '\t',
                            col.names = c('Geneid','clade'))
reid_cir_names = reid_cir_clade$Geneid
length(reid_cir_names)
#An extra gene compared to Brugat

brugat_cir_names[!brugat_cir_names %in% cir_info$Gene.ID] #Not annotated as cirs by PlasmoDB v Brugat
cir_info$Gene.ID[!cir_info$Gene.ID %in% brugat_cir_names] #Not annotated as cirs by Brugat 2017? v PlasmoDB
cir_info$Gene.ID[!cir_info$Gene.ID %in% reid_cir_names] #Not annotated as cirs by Reid v PlasmoDB
reid_cir_names[!reid_cir_names %in% brugat_cir_names] #Not annotated as cirs by Brugat v Reid
reid_cir_names[!reid_cir_names %in% cir_info$Gene.ID] #Not annotated as cirs by PlasmoDB v Reid

#Adam Reid's list contains all the pirs we should assume, however I don't understand why it is missing the 'conserved rodent pir' - need to discuss this with him.
#After extensive discussion it has been decided that due to dissimilarity and lack of an InterPro pir domain the
# pan-rodent is not a pir gene

reid_missing_cir = cir_info$Gene.ID[!cir_info$Gene.ID %in% reid_cir_names]

#Add the missing cir to the list of cirs and their clades
cir_clade_info <- reid_cir_clade %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'))
write_delim(cir_clade_info,
            '../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')

cir_info = chabaudi_gene_info %>% 
  select(Gene.ID, 
         Product.Description, 
         Transcript.Length) %>% 
  filter(Gene.ID %in% c(reid_cir_names)) %>% 
  mutate(Geneid = Gene.ID)
cir_transcript_lengths = select(cir_info, -Product.Description, -Gene.ID)

cir_id <- cir_info$Gene.ID
length(cir_id)

cir_id_no_anc_or_rodcon <- cir_id[!cir_id %in% 'PCHAS_0101200']

#How many cirs are present on each chromosome?

cir_tbl = as_tibble(data.frame(Geneid = cir_id))
cir_chrom_tbl <- cir_tbl %>% 
  mutate(chrom = unlist(sapply(lapply(sapply(as.character(cir_tbl$Geneid), strsplit,
                                             split = ''), '[', 7:8), 
                               paste, collapse = ''))) 
cir_chrom_tbl %>% 
  group_by(chrom) %>% 
  dplyr::summarise(number_of_cirs_in_each_chromosome = n(),
                   percentage_of_all_cirs = round(n()/nrow(cir_tbl), 3)*100) %>% 
  dplyr::arrange(desc(percentage_of_all_cirs))

#How many in each clade?
cir_clade_tbl = reid_cir_clade %>% 
  as_tibble() %>% 
  group_by(clade) %>% 
  dplyr::summarise(n())
cir_clade_tbl

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)
```

```{r chapl_info}
cir_chapl_aapl_info <- read_xlsx('Brugat2017_pirexpressiondata_41564_2017_BFnmicrobiol2016276_MOESM25_ESM.xlsx',
                            sheet = 1) %>% 
  filter(id %in% cir_id) %>%
  mutate(
    Geneid = id,
    locus_type = `Locus type`,
    chapl_aapl = ifelse(is.na(`Locus type`),
                        NA,
                        paste0(Subtelomere, 
                               '_',
                               `Locus type`)
    )
  ) %>% 
  select(Geneid, locus_type, chapl_aapl)

cir_chapl_aapl_info %>% 
  mutate(chrom = str_extract(Geneid, pattern = '(?<=PCHAS_)[:digit:]{2}'),
         locus_type = ifelse(is.na(locus_type), 'None', locus_type)) %>% 
  group_by(chrom, 
           locus_type) %>% 
  summarise(num_subfam = n_distinct(Geneid))
```

```{r gene_info, message = FALSE, results='hide'}

#No redundancy
two.transcript <- chabaudi_gene_info %>% 
  dplyr::count(Geneid) %>% 
  filter(n > 1)

```

```{r handier_descriptions}
chabaudi_gene_info$is_putative <- ifelse(
  str_detect(
    chabaudi_gene_info$Description,
    pattern = 'putative'
  ),
  'Yes',
  'No'
)

chabaudi_gene_info$handy_descriptions <- case_when( 
  #Handy function for avoiding nested, messy ifelse statements
  str_detect(
    chabaudi_gene_info$Description, 
    pattern = 'early transcribed membrane protein'
  ) ~ 
    'ETRAMP',
  str_detect(
  chabaudi_gene_info$Description, 
    pattern = 'PIR protein'
  ) ~ 
    'PIR',
  str_detect(
    chabaudi_gene_info$Description, 
    pattern = 'hypothetical'
  ) ~ 
    'hypothetical',
  str_detect(
    chabaudi_gene_info$Description, 
    pattern = '(conserved Plasmodium)|(conserved protein, unknown function)'
  ) ~ 
    'cons.',
  TRUE ~ #This applies the case to all remaining
    str_extract(  
      str_remove_all( 
        #Remove all applies across all the patterns instead of just one.
        chabaudi_gene_info$Description, 
        pattern = '( protein)|( protein, )|(, pseudogene)|(, putative)' 
        #Remove the words protein and spaces for the word.
      ),
      pattern = '.{1,10}' 
      #Take the first 10 characters max.
    )
)

chabaudi_gene_info$handy_descriptions <- ifelse(
  grepl(chabaudi_gene_info$Description, pattern = 'pseudogene'),
  paste0(chabaudi_gene_info$handy_descriptions,
         '_ps'),
  paste0(chabaudi_gene_info$handy_descriptions)
)

chabaudi_gene_info$handy_descriptions <- uniquify(
  ifelse(
    chabaudi_gene_info$is_putative == 'Yes',
    paste0(chabaudi_gene_info$handy_descriptions, '_put'),
    paste0(chabaudi_gene_info$handy_descriptions)
  ),
  first_one = TRUE
)

#Annotations for individual pirs based on Frech and Chen 2013, PlasmoDB
chabaudi_gene_info$handy_descriptions[which(chabaudi_gene_info$Geneid == 'PCHAS_0101200')] <- 'PIR_highly_cons.'
# chabaudi_gene_info$handy_descriptions[which(chabaudi_gene_info$Geneid == 'PCHAS_0524800')] <- 'PIR_rodent_cons.'
chabaudi_gene_info$handy_descriptions[which(chabaudi_gene_info$Geneid == 'PCHAS_1100300')] <- 'PIR_SBP_high'
```

```{r function_featurecounts_to_table, results = 'hide', message = FALSE}

featurecounts_to_table = function(vector_of_files = NULL, 
                                  read_skip = 0){

  files_exp <- vector_of_files
  # print(files_exp)
  #Combine the data itself into one dataframe
  exp = lapply(files_exp, 
               read.delim, 
               skip = read_skip) %>% 
    purrr::reduce(left_join, 
                  by = 'Geneid')
  #Using purrr and left_join means that Geneid will have to match between the samples so I removed the test for this.

  return(exp)
}
```

```{r function_counttable_to_tpm}

counttable_to_tpm = function(count_table,
                             table_names,
                             transcript_lengths = chabaudi_transcript_lengths,
                             technical_reps = FALSE,
                             experiment_name = NULL){

  #Before assigning stage numbers we need to merge the technical replicates

  if(technical_reps == TRUE){

    if(all(!grepl(colnames(table_names), pattern = 'Technical_Replicates'))){
      stop('If technical_reps is TRUE then the table_names data frame must have a Technical_Replicates column')
    }

    geneid_col = count_table[,'Geneid', drop = FALSE]
    list_techreps = lapply(unique(table_names$Technical_Replicates), function(techrep){
      print(techrep)
      #For each technical replicate we need the mean of the counts
      # So get the count table for these replicates - Alias names haven't been assigned yet so names should be runs
      techrep_runs = filter(table_names, Technical_Replicates == paste(techrep)) %>% 
        `$`(Run)
      #Extract the counts and calculate the mean
      techrep_counts =  select(count_table, paste(techrep_runs)) %>% 
        apply(1, mean) %>% data.frame()
      colnames(techrep_counts) = paste(techrep)
      techrep_counts
    })

    exp = bind_cols(geneid_col, list_techreps)
    colnames(exp)[-1] = table_names$Alias[match(colnames(exp)[-1], table_names$Technical_Replicates)]

  }else{

    exp = count_table
    colnames(exp)[-1] = as.character(table_names$Alias[match(colnames(exp)[-1], table_names$Run)])

  }

  #Need a number for each sample based on the number of distinct samples for each stage
  #Also going to remove NA columns here as they are things we don't want.

  colnames(exp) <- ifelse(!is.na(colnames(exp)), 
                          uniquify(colnames(exp)), 
                          NA)
  exp <- exp[,!is.na(colnames(exp)),drop = FALSE]

  # print(colnames(exp))
  exp = exp %>% arrange(Geneid)

  #Calculate the tpm
  tpm_exp = tpm(exp, transcript_lengths)

  list_exp =  list(counts = exp, tpm = tpm_exp, experiment = experiment_name)
  return(list_exp)
}
```

# Downloading the count files

This part is long not because downloading the count files is tricky but because the names of the different samples ought to be standardised as much as possible to be helpful down the line.

Count data was generated on the Crick HPC from the fq.gz files between 
using Nextflow:

Using PlasmoDB v45 _P. chabaudi chabaudi_ genome.

Here I have written a function which can help with re-naming columns and removing those you don't need if the different stages expected and the vector of files/list of objects containing the counts are specified. The function returns a list containing the count and the TPM files.

The design table from online is missing a file ('21659_8#24') for some reason. This has been designated as 'MT_Unknown' until we can find out which one it is. I've actually decided to remove this.

```{r download_files}
#Use the rMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')

design_table <- read.delim('samples_lanes.list',
                           sep = '\t',
                           header = FALSE,
                           col.names = c('sample', 'transmission', 'sequence'),
                           stringsAsFactors = FALSE) %>% 
  #regex extract for the numbers (one or more) before am, pm or h, or just an 'sch' for schizont
  mutate(time = str_extract(sample, pattern = '[:digit:]+([ap]m|h)|sch')) %>% 
  #Convert the times given in 'h' to robustly coded times in 24h clock
  mutate(time = ifelse(grepl(time, pattern = '[[:digit:]]h$'), 
                       strptime(time, format="%Hh") %>% format(., "%Hh"),
                       time)) %>% 
  mutate(time = ifelse(grepl(time, pattern = '[[:digit:]](a|p)m'), 
                       strptime(time, format="%I%p") %>% format(., "%Hh"),
                       time)) %>% 
  mutate(Alias = paste0(transmission, '_', time)) %>% 
  #Paste on the stage from the diff count, or nothing if NA
  mutate(Alias = ifelse(is.na(match(time,
                                    times_stages_table$time)),
                        Alias,
                        paste0(Alias, '_', 
                               times_stages_table$stage_diff_count[match(time,
                                                                         times_stages_table$time)]))) %>% 
  mutate(Alias = gsub(Alias, pattern = 'MT', replacement = 'rMT')) %>% 
  mutate(Alias = ifelse(grepl(Alias, pattern = 'sch'), paste0(Alias, '.culture'), Alias)) %>% 
  #To make the names R compatible
  mutate(Alias = make.names(Alias)) %>% 
  mutate(sequence = gsub(sequence, 
                         pattern = '9478_[[:digit:]]', 
                         replacement = '9478_cat'))

#Get vector of the files of MT and merged SBP
# the merged SBP files are not separated into _3 and _4

unknown_sample <- data.frame(sample = 'PcUnknown', 
                             transmission = 'rMT',
                             time = 'Unknown',
                             sequence = '21659_8#24',
                             Alias = 'rMT_Unknown')
design_table <- bind_rows(design_table, unknown_sample)

# design_table %>% 
#   group_by(Alias) %>% 
#   dplyr::summarise(n())

# design_table %>% 
#   group_by(transmission) %>% 
#   dplyr::summarise(n())

#Old files
# vector_of_files <- list.files(pattern = 'txt$') %>% str_subset(pattern = '21659|9478_#')

#New files - methods needs updating
vector_of_files <- c(
  list.files('nextflow_working_dir_PlasmoDBv48_rna-seq-r1.4.2_dir/featureCounts/gene_counts',
             pattern = 'txt$',
             full.names = TRUE)
)

# length(vector_of_files)
# length(tmp)

# str_extract(tmp, pattern = '(21659|9478)_([[:digit:]]+|cat|)#[[:digit:]]+')
# str_extract(vector_of_files, pattern = '(21659|9478)_([[:digit:]]+|cat|)#[[:digit:]]+')

# tmp[tmp %in% vector_of_files]

twentyfour_counts <- featurecounts_to_table(vector_of_files = vector_of_files,
                                            read_skip = 1)

#The names of the files / Runs is now a bit longer thanks to my method of naming files
design_table$Run <- unlist(
  sapply(make.names(design_table$sequence), 
                            function(nam){str_subset(colnames(twentyfour_counts)[-1], 
                                                     pattern = paste0(paste(nam), '_'))})
  ) %>% 
  make.names() #Make names acts during read.delim in the next function

design_table$unique_alias <- uniquify(design_table$Alias, first_one = TRUE)
# disagree(list(colnames(twentyfour_counts), design_table$Run))
```

```{r}
#Getting a nice count table for the paper and Github
chab_count_table <- select(twentyfour_counts, 
       matches('Geneid|bam')) %>% 
  rename_with(function(colnam){
    design_table$unique_alias[match(colnam, design_table$Run)]
  },
  matches('bam')) %>% 
  select(-matches('sch|Unk|SBP')) %>% 
  .[sort(colnames(.))] %>% 
  arrange(Geneid)

write_xlsx(chab_count_table,
           'Manuscript_SuppInfo_chabaudi_24h_counts.xlsx')
write.csv(chab_count_table,
           'Manuscript_SuppInfo_chabaudi_24h_counts.csv')

# tpm(chab_count_table, 
#     lengths = chabaudi_transcript_lengths)

```

New list of sequence files has 9478_(3 or 4)_#(Number) while the old files just had 9478_#(Number). This appears to be the cause of discrepency. Why are there these extra files?? I wonder if they are paired files which weren't analysed together? - Issue was because I hadn't merged technical replicates.

```{r calculate_save_TPM}
# #Calculate and save TPM
# twentyfour_ct <- counttable_to_tpm(count_table = twentyfour_counts,
#                                   table_names = design_table,
#                                   experiment_name = NULL)
# # There's an issue with bind_cols in this function at the moment
# #
# # #saving this output
# write_csv(twentyfour_ct$tpm,
#           'tl19-07_twentyfour_tpm_samples.csv')
# write_csv(twentyfour_ct$counts,
#           'tl19-07_twentyfour_counts_samples.csv')
# 
# #Getting all the sample columns in order.
# twentyfour_tpm_sbp <- twentyfour_ct$tpm %>%
#   # select(-rMT_Unknown) %>%
#   select_at(vars(matches('Geneid|SBP_')))
# 
# twentyfour_tpm_sbp <- select(
#   twentyfour_tpm_sbp,
#   Geneid,
#   ordering(
#     colnames(twentyfour_tpm_sbp)[-1],
#     order_vector = order_vector
#   )
# )
# # Saving a tab-delimited file
# write_delim(twentyfour_tpm_sbp,
#             'tl19-07_twentyfour_tpm_samples_SBP.txt',
#             delim = '\t')
# 
# twentyfour_tpm_rMT <- twentyfour_ct$tpm %>%
#   select(-rMT_Unknown) %>%
#   select_at(vars(matches('Geneid|rMT_')))
# 
# twentyfour_tpm_rMT <- select(
#   twentyfour_tpm_rMT,
#   Geneid,
#   ordering(
#     colnames(twentyfour_tpm_rMT)[-1],
#     order_vector = order_vector
#   )
# )
# # Saving a tab-delimited file
# write_delim(twentyfour_tpm_rMT,
#             'tl19-07_twentyfour_tpm_samples_rMT.txt',
#             delim = '\t')

```

```{r read_tpm}

#Read TPM - no unknown - combined data frame
twentyfour_ct_comb <- list(
  tpm = read_csv('tl19-07_twentyfour_tpm_samples.csv') %>% 
    select(-rMT_Unknown),
  counts = read_csv('tl19-07_twentyfour_counts_samples.csv') %>% 
    select(-rMT_Unknown)
)

#Save counts

write_xlsx(twentyfour_ct_comb$counts[,sort(colnames(twentyfour_ct_comb$counts))] %>% 
             dplyr::select(!matches('sch')),
           'Manuscript_Frontiers_chabaudi_24h_counts.xlsx')

#Read TPM separately - no unknown
twentyfour_ct_rMT <- list(
  tpm = read_csv('tl19-07_twentyfour_tpm_samples.csv') %>% 
    select(-rMT_Unknown) %>%
    select_at(vars(matches('Geneid|rMT_'))),
  counts = read_csv('tl19-07_twentyfour_counts_samples.csv') %>% 
    select(-rMT_Unknown) %>%
    select_at(vars(matches('Geneid|rMT_'))),
  experiment = 'rMT')

twentyfour_ct_SBP <- list(
  tpm = read_csv('tl19-07_twentyfour_tpm_samples.csv') %>% 
                        select(-rMT_Unknown) %>%
                        select_at(vars(matches('Geneid|SBP_'))),
  counts = read_csv('tl19-07_twentyfour_counts_samples.csv') %>% 
                        select(-rMT_Unknown) %>%
                        select_at(vars(matches('Geneid|SBP_'))),
  experiment = 'SBP')

twentyfour_ct <- list(twentyfour_ct_rMT, 
                      twentyfour_ct_SBP)

#Removing the rMT and SBP prefix for the experiments now that they're separated by transmission
twentyfour_ct <- lapply(twentyfour_ct, 
                        function(list_ct){
                          lapply(list_ct, 
                                 function(ct){
                                   if(!is.null(colnames(ct))){
                                     colnames(ct) <- str_remove(colnames(ct),
                                                                pattern = '(rMT|SBP)_')
                                   }
                                   return(ct)
                                 })
                        })

```

```{r counts_v_features}
num_counts <- apply(twentyfour_ct_comb$counts[,-1], 2, sum)
num_features <- apply(twentyfour_ct_comb$counts[,-1], 2, function(x) sum(x > 0))
plot(num_counts, num_features 
     #, ylim = c(0, max(num_features))
     )
```

```{r sanity_test}

#sanity test
# twentyfour_ct$tpm %>% select(Geneid, 'SBP_17h_3') %>% .[450:455,]
# # read.delim('samples_lanes.list',
# #                            sep = '\t',
# #                            header = FALSE,
# #                            col.names = c('sample', 'transmission', 'sequence'),
# #                            stringsAsFactors = FALSE)
# read.delim(header = TRUE, sep = '\t', file = list.files(pattern = '9478_#6')[1], skip = 1) %>% 
#   arrange(Geneid) %>% 
#   select(Geneid, X9478_.6_hisat2alignment_v45chabaudigenome.sorted.merged.bam) %>% .[450:455,]
# 
# twentyfour_ct$counts %>% select(Geneid, 'MT_05h_3') %>% .[1284:1289,]
# read.delim(header = TRUE, sep = '\t', file = list.files(pattern = '21659_8#3')[1], skip = 1) %>% 
#   arrange(Geneid) %>% 
#   select(Geneid, X21659_8.3_hisat2alignment_v45chabaudigenome.sam) %>% .[1284:1289,]

```

# Hierarchical clustering of the samples

```{r hclusts}
hclusts = function(tpms){
  dists = dist(t(tpms$tpm[,-1]))
  plot(hclust(dists), 
       cex = 0.7, 
       main = tpms$experiment)
}

invisible(
  lapply(twentyfour_ct, hclusts)
)
# hclusts(twentyfour_ct_comb)
```

Using scale() to normalise the data variance and then hclust with 'average' method. The rMT samples and SBP samples still cluster separately.

```{r hclusts_scaled_data}
hclusts_scaled = function(tpm){
  dists = dist(t(scale(tpm$tpm[,-1])))
  plot(hclust(dists, method = 'average'), cex = 0.7)
}
hclusts_scaled(twentyfour_ct_comb)
```

Both here and in the correlation plots below most of the samples (except 11 and 14h) are distinct based not on time of day but be method of transmission. This isn't what one would expect. It of course could be due to batch effects since the SBP and MT experiments were done by different people in different years.

hclust without that one naughty pseudogene. This one pseudogene, PCHAS_0420800, is a major defining gene of the rMT samples, and removal of this gene makes the samples cluster as we'd expect; by asexual stage and not by transmission

```{r hclust_nopseudo}
tpm_nopseudo <- twentyfour_ct_comb$tpm %>% filter(!Geneid %in% 'PCHAS_0420800')

# sapply(tpm_nopseudo[,-1], sum)[rev(order(sapply(tpm_nopseudo[,-1], sum)))]
# Need to rescale the TPM

tpm_nopseudo_rescale <- lapply(colnames(tpm_nopseudo), 
                               function(column_name){
                                 column <- getElement(tpm_nopseudo, paste(column_name))
                                 if(is.numeric(column)){
                                   #is.numeric to avoid trying to scale the Geneid column
                                   tpm_scale_factor <- 1e6 / sum(column)
                                   tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                 }else{
                                   tpm_scaled <- column
                                 }
                                 return(tpm_scaled)
                               }) %>% 
  bind_cols

colnames(tpm_nopseudo_rescale) <- colnames(tpm_nopseudo)
# sapply(tpm_nopseudo_rescale[,-1], sum)[rev(order(sapply(tpm_nopseudo_rescale[,-1], sum)))]

#Note now the MT samples can be distinguished by having a lower total TPM
dists = dist(t(tpm_nopseudo_rescale[,-1]))
plot(hclust(dists), cex = 0.7)
```

# Density plots

These graphs show the distribution of TPM values for each gene in each sample. The shapes of the curves follow the generally expected pattern of one smaller peak at zero and another peak and higher values with a long tail. Individual samples however show a lot of variation from one another presumably because of differences in the amount of sequences detected by the experiments.

```{r density_plots, fig.show="hold", out.width="50%"}
#fig.show and out.width settings for having both plots on same output.
list_min = lapply(twentyfour_ct, function(input){
  input_tpm = input$tpm
  minima = sapply(colnames(select(input_tpm, -Geneid)), function(colnam){
    #Want to exclude the weird sporozoite densities
    if(grepl(colnam, pattern = 'sporozoite', ignore.case = TRUE)){
      NA
    }else{
      tpms = select(input_tpm, paste(colnam)) %>% unlist %>% as.numeric()
      den = density(log1p(tpms))
      #Get the difference between adjacent y values, find the positive change, first one coresponds to the first minimum
      first_min = den$x[which(diff(sign(diff(den$y)))==2)] %>% .[1]
    }
  }) %>% expm1()
  #expm1 is the reverse of log1p - almost forgot to take away the one!
  list(experiment = input$experiment,
       range = range(minima, na.rm = TRUE), 
       mean = mean(minima, na.rm = TRUE),
       median = median(minima, na.rm = TRUE))
})

min_median = median(unlist(lapply(list_min, function(x) x$median)))
min_min = min(unlist(lapply(list_min, function(x) x$range)))
min_max = max(unlist(lapply(list_min, function(x) x$range)))

threshold <- min_median

# print(min_median)

list_bw = lapply(twentyfour_ct, function(expt){ 
  lapply(colnames(expt$tpm)[-1], function(sample){
    data = as.numeric(unlist(select(expt$tpm, paste(sample))))
    dens = density(log1p(data))
    dens$bw
})
});
bw_mean = mean(unlist(list_bw))
bw_range = max(unlist(unlist(list_bw))) - min(unlist(unlist(list_bw)))
# print(paste('Bandwidth range = ', bw_range))

invisible(
  lapply(twentyfour_ct, 
       density_samples, 
       save = FALSE)
)
#Wrapping in Invisible to hide the console output of NULL (I don't know where it comes from...)
```

Based on these density plots we make the assumption of a threshold of expression as being `r round(min_median, digits = 2)` TPM.

# Top general and _cir_ genes in each stage

```{r top_genes}
top_n_genes = function(tpm_input,
                       threshold = 0,
                       stage_str_extract_regex_pattern = '(?<=_).+(?=_[[:digit:]]+)|sch.culture'){
  tpm = tpm_input$tpm %>% 
    #Melt the data frame for using dplyr
    melt(id.vars = 'Geneid', 
         variable.name = 'sample', 
         value.name = 'tpm') %>% 
    mutate(stage = str_extract(sample, 
                               pattern = paste(stage_str_extract_regex_pattern))) %>% 
    group_by(Geneid, stage) %>%  
    dplyr::summarise(median_tpm = median(tpm),
                     mean_tpm = mean(tpm), 
                     min_tpm = min(tpm), 
                     max_tpm = max(tpm), 
                     sd_tpm = sd(tpm)) %>% 
    group_by(stage) %>% 
    mutate(experiment = tpm_input$experiment)

  list_data = list('tpm' = tpm,
                   'experiment' = tpm_input$experiment)
  return(list_data)
}

### Generate and save the top gene data

# twentyfour_top <- lapply(twentyfour_ct, top_n_genes)
# 
# lapply(twentyfour_top, function(transmission_top){
#   tpm_data <- transmission_top$tpm
#   write_csv(tpm_data,
#             paste0('tl19-07_twentyfour_tpm_samplescollated_longform_',
#                    transmission_top$experiment,
#                    '.csv'))
# })
# 
# twentyfour_top_comb <- top_n_genes(twentyfour_ct_comb,
#                                    stage_str_extract_regex_pattern = '.+(?=_[[:digit:]]+)')
# 
# write_csv(twentyfour_top_comb$tpm,
#           'tl19-07_twentyfour_tpm_samplescollated_longform.csv')

###

### Read in the top gene data

twentyfour_top <- lapply(list.files(pattern = 'tl19-07_twentyfour_tpm_samplescollated_longform_', 
                  full.names = TRUE), 
       function(top_files){
         list(
           tpm = read_csv(top_files),
              experiment = str_extract(top_files, pattern = 'rMT|SBP')
         )
       })

twentyfour_top_comb <- list(tpm = read_csv('tl19-07_twentyfour_tpm_samplescollated_longform.csv'))

###
```

```{r kable_table_top_n_genes, results = "asis"}

top_n_genes_kable_table <- function(top_n_tpm,
                                    n_top = 5,
                                    subset_genes = NULL,
                                    subset_genes_title = 'all',
                                    gene_info = chabaudi_gene_info,
                                    highlight_genes_regex = 'PCHAS_0101200'){
  # top_n_tpm <- twentyfour_top[[1]]

  tpm <- top_n_tpm$tpm

  #Calculating the index used to make the kable table

  kable_index = rep(n_top, 
                    times = length(unique(tpm$stage)))
  names(kable_index) = unique(tpm$stage)

  if(is.null(subset_genes)){
    top_genes <- tpm
  }else{
    top_genes <- filter(tpm, Geneid %in% subset_genes)
  }

  top_genes = top_genes %>% 
    group_by(stage) %>% 
    top_n(wt = median_tpm, 
          n = n_top) %>% 
    dplyr::arrange(stage, 
                   desc(median_tpm)) %>% 
    mutate_if(is.numeric, 
              function(x){round(x, 
                                digits = 2)})

  top_genes$description = gene_info$handy_descriptions[match(top_genes$Geneid, 
                                                             gene_info$Geneid)]

  top_genes_kable <- top_genes %>% 
    mutate(median_tpm = cell_spec(median_tpm, "html", 
                                  color = ifelse(median_tpm > 10, "red", "blue"))) %>% 
    mutate(Geneid = as.character(Geneid)) %>% 
    mutate(Geneid = ifelse(grepl(Geneid, pattern = paste(highlight_genes_regex)),
                           cell_spec(Geneid, bold = TRUE),
                           Geneid)) %>% 
    kable(format = 'html', 
          caption = paste0(top_n_tpm$experiment, 
                           ' - Top ', 
                           paste(n_top), 
                           ' expressed ',
                           paste(subset_genes_title),
                           ' genes'), 
          escape = FALSE) %>%
    kable_styling() %>%
    pack_rows(index = kable_index) %>%
    scroll_box(width = "100%", height = "500px")

  return(top_genes_kable) 
}

lapply(twentyfour_top, 
       top_n_genes_kable_table)
lapply(twentyfour_top,
       top_n_genes_kable_table,
       subset_genes = cir_id, 
       subset_genes_title = 'cir')

```

~~At the moment these tables are not showing up in this document...~~

Wow! The SBP cir PCHAS_1100300 isn't expressed higher than ancestral or rodent-conserved pir-like in these SBP samples...

For all the other genes again in MT one of the highest expressed genes is a pseudogene while in SBP it is a series of ETRAMPs (although MT expresses these also it seems).

```{r most_variable_genes, eval = FALSE}
twentyfour_top$tpm %>%
  top_n(5, sd_tpm) %>%
  arrange(stage, desc(sd_tpm))
```

* _PCHAS_0101200_ = ancestral cir.

* _PCHAS_0524800_ = rodent conserved cir-like genes.

* _PCHAS_1100300_ = SBP-dominant cir gene (Spence et al, 2013).

```{r main-three-pir_heatmap}
# pdf('tl19-07_cirgenes_24h_heatmap.pdf', width = 10, height = 5)
draw_heatmap_chabaudi_24h(datasets =twentyfour_ct, 
                          genelist = c('PCHAS_0101200',
                                       # 'PCHAS_0524800',
                                       'PCHAS_1100300'),
                          scale = 'max',
                          threshold = 0,
                          log_it = TRUE)
# dev.off()
twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% c('PCHAS_0101200',
                       # 'PCHAS_0524800',
                       'PCHAS_1100300')) %>% 
  kable() %>% 
  kable_styling() %>% 
  scroll_box(width = "100%")
```

```{r Deirdres CHAPLs, eval = FALSE}
protein <- c('PCHAS_0400300',   
             'PCHAS_0600600',    
             'PCHAS_1200500',    
             'PCHAS_1300101',
             'PCHAS_0101200',   
             'PCHAS_1100300',
             'PCHAS_0626900')

draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
                          genelist = protein,
                          scale = 'max',
                          threshold = 0,
                          log_it = TRUE,
                          gene_names_for_title = 'Available protein')

chapl <- c('PCHAS_0626200',
           'PCHAS_0626700',
           'PCHAS_0627100')

draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
                          genelist = chapl,
                          scale = 'max',
                          threshold = 0,
                          log_it = FALSE,
                          gene_names_for_title = 'Deirdre\'s ChAPLs')

available_primers <- c('PCHAS_0600600',
                       'PCHAS_1100300',
                       'PCHAS_1301700',
                       'PCHAS_0400300',
                       'PCHAS_0301400',
                       'PCHAS_1200500',
                       'PCHAS_1300101')

draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
                          genelist = available_primers,
                          scale = 'max',
                          threshold = 0,
                          log_it = FALSE,
                          gene_names_for_title = 'Available primers')
```

Good _cir_ negative controls would be 0600600 and 1200500 for SBP and 1301700 for both (but especially MT).

# Correlation plots {.tabset}

## Correlation plot of collated stages (all genes)

```{r corelation_plots_collated}
# Collated across the stages for a more easily understood heatmap

big_top_twentyfour = twentyfour_ct_comb$tpm %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  group_by(Geneid, stage) %>% 
  dplyr::summarise(median_tpm = median(tpm)) %>% 
  dcast(Geneid ~ stage)

big_top_twentyfour_mat = big_top_twentyfour[,-1,drop=FALSE]

# pdf(file = 'tl19-07_corrplot_24h_rnaseq_chabaudi.pdf')
corrplot(cor(big_top_twentyfour_mat), method = 'color', order = 'hclust', cl.lim = c(-0,1), tl.cex = 0.7,
         col = c(viridis(100), magma(115)), 
         mar=c(0,0,1,0),
         title = 'Correlation plot')
# dev.off()

```

## Correlation plot of collated stages (only _cir_ genes)

```{r}

#SBP now all together, so the schizonts were pushing them elsewhere

#Now just the cir genes

big_top_twentyfour_cir_mat = big_top_twentyfour %>% 
  filter(Geneid %in% cir_id) %>% 
  select(-Geneid)
corrplot(cor(big_top_twentyfour_cir_mat), method = 'color', order = 'hclust', 
         cl.lim = c(0.0,1), tl.cex = 0.7,
         col = c(viridis(100), magma(100)), 
         mar=c(0,0,1,0),
         title = 'Correlation plot (only cirs)')
```

## Correlation plot of collated stages, but without schizonts (all genes)

```{r correlation_collated_nosch}

#How does this order with the schizonts removed

big_top_twentyfour_nosch = big_top_twentyfour %>% select(-rMT_sch.culture)

big_top_twentyfour_nosch_mat = big_top_twentyfour_nosch[,-1,drop=FALSE]

# pdf(file = 'tl19-07_corrplot_24h_rnaseq_chabaudi_noschz.pdf')
corrplot(cor(big_top_twentyfour_nosch_mat), method = 'color', order = 'hclust', cl.lim = c(0,1), tl.cex = 0.7,
         col = c(viridis(100), magma(115)), 
         mar=c(0,0,1,0),
         title = 'Correlation plot (no schizont samples)')
# dev.off()
```

Now all the SBP and MT groups cluster together

## Correlation plot of all samples (all genes)

```{r correlation_plot}
twentyfour_tpm_mat = twentyfour_ct_comb$tpm
rownames(twentyfour_tpm_mat) = twentyfour_tpm_mat$Geneid
twentyfour_tpm_mat = twentyfour_tpm_mat[,-1,drop=FALSE]

#Replicates
twentyfour_all_cor = cor(twentyfour_tpm_mat)
corrplot(twentyfour_all_cor, tl.cex = 0.5, mar=c(0,0,1,0),
         method = 'color', order = 'hclust', cl.lim = c(0,1), col = c(viridis(100), tl.cex = 0.7, magma(115)),
         title = 'Correlation plot - individual samples')
#MT and SBP separate, SBP 11/14/17h match MT schizonts
```

## Correlation plot of all samples (but only _cir_ genes)

```{r cir_correlation}

#Replicate cirs
twentyfour_tpm_mat_cir <- filter(twentyfour_ct_comb$tpm, Geneid %in% cir_id) #This drops the rownames for some reason
rownames(twentyfour_tpm_mat_cir) <- twentyfour_tpm_mat_cir$Geneid 
twentyfour_tpm_mat_cir <- twentyfour_tpm_mat_cir[,-1,drop=FALSE]
twentyfour_tpm_mat_cir_cor <- cor(twentyfour_tpm_mat_cir)
corrplot(twentyfour_tpm_mat_cir_cor, tl.cex = 0.5, method = 'color', order = 'hclust', 
         cl.lim = c(0,1), col = c(viridis(100), magma(115)),
         mar=c(0,0,1,0),
         title = 'Correlation plot - individual samples (only cirs)')
```

## Do the top 10 expressed genes in each sample give any extra clarity to the correlation plot?

Drawing a correlation plot with only the top 10 expressed genes in the samples.

Not really, I guess need to get the genes with the highest variance not total score.

```{r top10genes}
top10genes <- twentyfour_top_comb$tpm %>% 
  dplyr::arrange(desc(median_tpm)) %>% 
  group_by(stage) %>% 
  top_n(median_tpm, n = 10) %>% 
  .$Geneid %>% 
  unique

draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
                          genelist = top10genes,
                          scale = 'max',
                          threshold = 0,
                          log_it = TRUE)

big_top_twentyfour_10_mat = big_top_twentyfour %>% 
  filter(Geneid %in% cir_id) %>% 
  select(-Geneid)
corrplot(cor(big_top_twentyfour_10_mat), method = 'color', order = 'hclust', cl.lim = c(0.0,1), tl.cex = 0.7,
         col = c(viridis(100), magma(100)))
```

## Do the top 10 most variable genes across the samples give any extra clarity to the correlation plot?

The top ten most variable genes (highest standard deviation) between all the individual samples were identified.

Did with and without the pseodogene _PCHAS_0420800_ which dominates the expression profile of MT and not SBP.

```{r most_variable_genes_corr}

variable_genes <- twentyfour_top_comb$tpm %>% 
  ungroup %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(median_tpm),
                   std_tpm = sd(median_tpm)) %>% 
  mutate(product_description = 
           chabaudi_gene_info$Product.Description[match(Geneid, chabaudi_gene_info$Gene.ID)]) %>% 
  arrange(desc(std_tpm))

variable_gene_names <- variable_genes$Geneid[1:10]

draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
                          genelist = variable_gene_names,
                          scale = 'max',
                          threshold = 0,
                          log_it = TRUE)

big_top_twentyfour_mostvar_mat = big_top_twentyfour %>% 
  filter(Geneid %in% variable_gene_names) %>% 
  select(-Geneid)
corrplot(cor(big_top_twentyfour_mostvar_mat), method = 'color', order = 'hclust', tl.cex = 0.7,
         col = c(magma(200)), title = '10 most variable genes', mar = c(0,0,1,0))
#Note that the scale now extends to -1

#Again but without the pseudogene

variable_gene_names_notop <- variable_genes$Geneid[2:11]
big_top_twentyfour_mostvar_mat = big_top_twentyfour %>% 
  filter(Geneid %in% variable_gene_names_notop) %>% 
  select(-Geneid)
corrplot(cor(big_top_twentyfour_mostvar_mat), method = 'color', order = 'hclust', tl.cex = 0.7,
         col = c(magma(200)), title = '10 most variable genes - except pseudo', mar = c(0,0,1,0))

```

# Undetected _cir_ genes.

Which _cir_ genes are never expressed over the threshold of `r round(min_median, digits = 2)` TPM.

```{r never_expressed}
filter_at(twentyfour_ct_comb$tpm, 
          vars(-contains('Geneid')), 
          all_vars(. < min_median)) %>% 
  filter(Geneid %in% cir_id) %>% 
  kable() %>% 
  kable_styling() %>% 
  scroll_box(width = "100%", height = "500px")
```

# PCA of the life stages {.tabset}

## PCA with all genes {.tabset}

### 2D plot of PC1 v PC2

```{r pca_calculation, results = TRUE}
twentyfour_tpm <- twentyfour_ct_comb$tpm
rownames_geneid <- twentyfour_tpm$Geneid
twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
rownames(twentyfour_tpm_forpca) <- rownames_geneid
twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
#Need to look into what exactly scale and center are doing
```

```{r 2d_pca, results = TRUE}
twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)

twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])

# colornames <- c(cbPalette, 'black', 'red')
colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
colornames <- setNames(colornames, 
                       sort(unique(twentyfour_tpm_pca_coords$time)))

pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)

pl <- plot_ly(twentyfour_tpm_pca_coords, 
             x  = ~PC1, y = ~PC2,
             type = 'scatter', mode = 'markers',
             color = ~time, text = ~sample,
             colors = colornames,
             # size = ~total_features,
             marker = list(opacity = 1, size = 16),
             symbol = ~transmission, symbols = c('circle','x')) %>% 
  layout(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
         yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
         legend = list(font = list(size = 8)))
pl
orca(pl, "tl19-07_audreybrugat24hchabaudi_2D-PCA.pdf")

head(twentyfour_tpm_pca$x[rev(order(abs(twentyfour_tpm_pca$x[,1]))),,drop=FALSE])[,1:9]
```

```{r 2d_pca_allgenes_suppfig}
(
  plot <- ggplot(twentyfour_tpm_pca_coords,
                 aes(
                   x = PC1,
                   y = PC2,
                   col = time,
                   shape = transmission
                 )) +
    geom_point(size = 3,
               alpha = 0.8) +
    scale_color_manual(values = colornames) +
    scale_shape_manual(
      values = setNames(
        c(17,19),
        unique(twentyfour_tpm_pca_coords$transmission)
      )
    ) +
    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
    guides(shape=guide_legend(ncol=2),
           col = guide_legend(ncol=2)) +
    theme_classic() +
    theme(legend.key.size = unit(4, 'mm'),
          legend.text = element_text(size = 7))
)
ggsave2(filename = 'tl19-07_chabaudi24h_supp_figure_pca_allgenes.png',
        plot = plot,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')
```

Can do a plot of the gene loadings for PC1 and PC2 to see which genes are having the biggest impact on each.

```{r 2d_pca_loadings}
# plot_ly(data.frame(Geneid = rownames(twentyfour_tpm_pca$x),
#                    product = chabaudi_gene_info$Product.Description[match(rownames(twentyfour_tpm_pca$x), 
#                                                          chabaudi_gene_info$Gene.ID)],
#                    twentyfour_tpm_pca$x),
#         x = ~PC1, y = ~PC2,
#         type = 'scatter', mode = 'markers',
#         text = ~paste(Geneid, product))
plot_ly_biplot(twentyfour_tpm_pca,
               prin_components = c(1,2,3),
               scale.loads = 700,
               separate_multigene = FALSE,
               three_dimensions = FALSE,
               textposition = "top left",
               gene_info = chabaudi_gene_info)
orca(plot_ly_biplot(twentyfour_tpm_pca,
               prin_components = c(1,2,3),
               scale.loads = 700,
               separate_multigene = FALSE,
               three_dimensions = FALSE,
               textposition = "top left",
               gene_info = chabaudi_gene_info),
     "tl19-07_audreybrugat24hchabaudi_biplotPCA.pdf")

```

Negative loading for the the PCHAS_0420800 gene means that the samples which contain it are pushed into the positive of the PComponent. The absolute value is what's important loadings.

Most obvious separation is by method of transmission and not by time of day.

```{r pca_loading_genes}
top_loadings_pc1 <- twentyfour_tpm_pca$x[rev(order(abs(twentyfour_tpm_pca$x[,1]))),,drop=FALSE][1:5,1:8]
top_loadings_pc1_genes <- rownames(top_loadings_pc1)
chabaudi_gene_info %>% filter(Geneid %in% top_loadings_pc1_genes)

pdf('tl19-07_24hchabaudi_heatmap_topPCAloadingsgenes.pdf',width = 7, height = 5)
draw_heatmap_chabaudi_24h(tpm_input = twentyfour_ct_comb, 
                          genelist= top_loadings_pc1_genes,
                          scale = 'max',
                          log_it = TRUE)
dev.off()
```

### 3D plot of PC1 v PC2 v PC3

```{r 3d_pca, results = TRUE}

q <- plot_ly(twentyfour_tpm_pca_coords, 
             x  = ~PC1, y = ~PC2, z = ~PC3,  mode = 'markers',
             type = 'scatter3d', color = ~time, text = ~sample,
             colors = colornames,
             # size = ~total_features,
             marker = list(opacity = 1, size = 5),
             symbol = ~transmission, symbols = c('circle','x')) %>% 
  layout(scene = list(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
                      yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
                      zaxis = list(title = paste0('PC3 ',pc3_importance,'%'))))
q
```

PCA struggles to differentiate between the time stages. PC2 clearly separates by transmission method. While PC1 and PC3 vaguely clusters similar times together.

## PCA but only including the _cir_ sequences {.tabset}

### 2D:

```{r onlycir, fig.cap = 'PCA plot of the cir gene expression of 24h samples'}
pca_function(twentyfour_ct_comb, genevec = cir_id)
# orca(pca_function(twentyfour_ct_comb, 
#                   genevec = cir_id,
#                   scale.pca = TRUE, 
#                   legend_settings = list(font = list(size = 8))),
#      "tl19-07_audreybrugat24hchabaudi_2D-PCA-cir.pdf")
```

The rMT samples appear to contain much greater variation than the tightly clustered SBP samples.

### 3D:

```{r pca_onlycir_3d, fig.cap = 'Three dimensional PCA plot of the cir gene expression of 24h samples'}
pca_function(twentyfour_ct_comb, genevec = cir_id, three.dimensional = TRUE)
```

Looks like they both form their own separate clusters of cycling pir expression.

### Biplot:

```{r pca_onlycir_pc2-pc3, fig.cap = 'Biplot of cir gene expression of 24h samples'}
twentyfour_tpm_forpca <- data.frame(filter(twentyfour_tpm, Geneid %in% cir_id))
rownames(twentyfour_tpm_forpca) <- sapply(twentyfour_tpm_forpca$Geneid, 
                                          function(x){ 
                                            paste(x, 
                                                  chabaudi_gene_info$Product.Description[match(x,
                                                                              chabaudi_gene_info$Gene.ID)])
                                          })
twentyfour_tpm_pca_onlycir <- prcomp(select(twentyfour_tpm_forpca, -Geneid),
                                      center = TRUE, 
                                      scale. = TRUE)

plot_ly_biplot(pca_object = twentyfour_tpm_pca_onlycir,
               scale.loads = 300,
               gene_info = chabaudi_gene_info)

```

Heatmap of the five _pirs_ with the greatest impact on the cir-only PCA (greatest modulus):

```{r pca_onlycir_heatmap, fig.cap = 'Expression heatmap of the five cir genes with the highest magnitude variable loading values in the PCA'}
onlycir_loadings <- data.frame(Geneid = rownames(twentyfour_tpm_pca_onlycir$x),
                                twentyfour_tpm_pca_onlycir$x)

#Heatmap of the cir genes with greatest modulus from the origin
top_pca_cir <- data.frame(onlycir_loadings) %>% 
  mutate(length = sqrt((PC1^2)*(PC2^2))) %>% 
  dplyr::arrange(desc(length)) %>% 
  select(c(Geneid, PC1, PC2, PC3, length)) %>% 
  .$Geneid %>% 
  .[1:5] %>% 
  str_extract(pattern = 'PCHAS_[[:digit:]]+')
draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
                          genelist = top_pca_cir, 
                          log_it = TRUE)
# ggbiplot(twentyfour_tpm_pca_onlycir)
```

Both the major cirs are good for distinguishing the samples from each other, probably because they are the greatest signals. Some others clearly distinguish between rMT and SBP.

## PCA without the _cir_ sequences

```{r pca_no_cir, fig.cap='PCA plot of all gene expression, without the cir genes, of 24h samples'}
geneid_nocir <- twentyfour_ct_comb$counts$Geneid[which(!twentyfour_ct_comb$counts$Geneid %in% cir_id)]
pca_function(twentyfour_ct_comb, 
             genevec = geneid_nocir)
```

Looks much the same as the overall PCA.

## PCAs with the MT v SBP pseudogene removed {.tabset}

 ~~Should do this but with rescaling of the TPM.~~

### 2D:

```{r pca_no_pseudogene, fig.cap='PCA plot of all gene expression, without the rMT-high pseudogene, of 24h samples'}
geneid_nopseudo <- twentyfour_ct_comb$counts$Geneid[which(!twentyfour_ct_comb$counts$Geneid %in% 'PCHAS_0420800')] 
pca_function(twentyfour_ct_comb, 
             genevec = geneid_nopseudo, 
             scale.pca = TRUE, 
             legend_settings = list(font = list(size = 8)))
orca(pca_function(twentyfour_ct_comb, 
                  genevec = geneid_nopseudo, 
                  scale.pca = TRUE, 
                  legend_settings = list(font = list(size = 8))),
     "tl19-07_audreybrugat24hchabaudi_2D-PCA-nopseudogene.pdf")
# pca_function(twentyfour_ct_comb, genevec = geneid_nopseudo, scale.pca = FALSE)
```

### 3D:

```{r pca_3d_no_pseudogene, fig.cap = 'Three dimensional PCA plot of all gene expression, without the rMT-high pseudogene, of 24h samples'}
pca_function(twentyfour_ct_comb, 
             genevec = geneid_nopseudo, 
             scale.pca = TRUE, 
             three.dimensional = TRUE)
```

Removing cirs leads to no change in the PCA but removing the pseudogene PCHAS_0420800 has a massive impact!

Get a sort of ring shape (although there's quite a 'jump' between late trophs and earlier ring stages) showing transcriptional progression, although this is best described by PC2 and PC3 which only account for 15% of variation.

PC1 focusses on variation due to the schizont and SBP early-mid ring samples.

### Biplot:

```{r pca_no_pseudogene_loadings_plot}
twentyfour_tpm_forpca <- data.frame(
  filter(
    twentyfour_ct_comb$tpm, 
    Geneid %in% geneid_nopseudo)
)
rownames(twentyfour_tpm_forpca) <- sapply(twentyfour_tpm_forpca$Geneid, 
                                          function(x){ 
                                            paste(x, 
                                                  chabaudi_gene_info$Product.Description[match(x,
                                                                              chabaudi_gene_info$Gene.ID)])
                                          })
twentyfour_tpm_pca_nopseudo <- prcomp(select(twentyfour_tpm_forpca, -Geneid),
                                      center = TRUE, 
                                      scale. = TRUE)
plot_ly_biplot(twentyfour_tpm_pca_nopseudo,
               textposition = 'left')

```

## PCA without the schizont samples (and also without the pseudogene)

```{r}
twentyfour_ct_comb_noschiz <- list(
  tpm = select_at(
    twentyfour_ct_comb$tpm,
    vars(
      -matches('sch')
    )
  ),
  experiment = '24h-noschizont'
)
pca_function(twentyfour_ct_comb_noschiz, 
             genevec = geneid_nopseudo, 
             scale.pca = TRUE, 
             colornames = colornames,
             three.dimensional = TRUE)
```

## PCA without the schizont and Early.Mid.Rings samples which dominate PC1 (and also without the pseudogene)

```{r pca_no_pseudogene_no_schizont, messages = FALSE}
twentyfour_ct_comb_noschiz_noring <- list(
  tpm = select_at(
    twentyfour_ct_comb$tpm,
    vars(
      -matches('sch|Early.Mid.Rings')
    )
  ),
  experiment = '24h-noschizont-noSBPearlymidring'
)
pca_function(twentyfour_ct_comb_noschiz_noring, 
             genevec = geneid_nopseudo, 
             scale.pca = TRUE, 
             colornames = colornames,
             three.dimensional = TRUE)
# pca_function(twentyfour_ct_comb, genevec = geneid_nopseudo, scale.pca = FALSE)
orca(pca_function(twentyfour_ct_comb_noschiz, 
                  genevec = geneid_nopseudo, 
                  scale.pca = TRUE, 
                  colornames = colornames,
                  legend_settings = list(font = list(size = 8))),
     "tl19-07_audreybrugat24hchabaudi_2D-PCA-nopseudogene-noschizorearlymidrings.pdf")
```

Removing just the schizonts (not shown) doesn't make PC1 distinguish the samples much better, and the Early.Mid.Rings still are the most separated by the component. Remove both and a ring is produced just with PC1 and PC2.

## PC5 and PC6 of original PCA

Could also look at different principal component where the pseudogene is having less impact to remove the effect of the pseudogene.

```{r pseudogene_pca_component}
twentyfour_tpm_pca$x[rownames(twentyfour_tpm_pca$x) %in% 'PCHAS_0420800',,drop=FALSE][,1:9, drop=FALSE]
#PC5 and PC6 show much less impact from the pseudogene

twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                        twentyfour_tpm_pca_var$coord) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])

pc5_importance <- round(pca_importance$PC5[2]*100, digits = 1)
pc6_importance <- round(pca_importance$PC6[2]*100, digits = 1)

pl <- plot_ly(twentyfour_tpm_pca_coords, 
             x  = ~Dim.5, y = ~Dim.6,
             type = 'scatter', mode = 'markers',
             color = ~time, text = ~sample,
             colors = colornames,
             # size = ~total_features,
             marker = list(opacity = 1, size = 16),
             symbol = ~transmission, symbols = c('circle','x')) %>% 
  layout(xaxis = list(title = paste0('PC5 ',pc5_importance,'%')),
         yaxis = list(title = paste0('PC6 ',pc6_importance,'%')))
pl
```

I don't think this is particularly insightful because the variance explained by PC5 and PC6 is only 2.3% and the PCA doesn't really show much anyway.

Notes:

* Shouldn't make a big difference if we set scale. = FALSE, could try with scale. = TRUE on the counts?

* If you don't scale need to make sure that what's affecting the PCA is something that you're interested in.

* Sparse principal components could be used? Allows you to set a threshold for PCA.

* PC1 separates by mean effect, so basically the highest expressed genes. PC2 onwards separates by harder to understand, subtler differences.

# Removing the pseudogene and re-scaling the TPM data

```{r rescale_tpm}

twentyfour_tpm_nopseudo <- twentyfour_ct_comb$tpm %>% 
  filter(!Geneid %in% 'PCHAS_0420800')

twentyfour_tpm_rescale_nopseudo <- lapply(colnames(twentyfour_tpm_nopseudo), 
                                          function(column_name){
                                            column <- getElement(twentyfour_tpm_nopseudo, 
                                                                 paste(column_name))
                                            if(is.numeric(column)){
                                              #is.numeric to avoid trying to scale the Geneid column
                                              tpm_scale_factor <- 1e6 / sum(column)
                                              tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                            }else{
                                              tpm_scaled <- column
                                            }
                                            return(tpm_scaled)
                                          }) %>% 
  bind_cols

colnames(twentyfour_tpm_rescale_nopseudo) <- colnames(twentyfour_tpm_nopseudo)
twentyfour_tpm_rescale_nopseudo_t <- list(tpm = twentyfour_tpm_rescale_nopseudo)
# apply(twentyfour_tpm_rescale_nopseudo[,-1], 2, sum)

#Setting the re-scaled version as the default from now on
#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo_t

write_xlsx(select(twentyfour_ct_comb$tpm, 
                  -matches('sch')),
           'tl19-07_TPM_both_nopseudoRescaled_labelsStage_allSamples.xlsx')

twentyfour_ct_rescale <- lapply(twentyfour_ct, 
                                function(list_element){
                                  tpm_data <- list_element %>% 
                                    .$tpm %>% 
                                    filter(!Geneid %in% 'PCHAS_0420800')

                                  rescaled_tpm <- lapply(colnames(tpm_data), 
                                                         function(column_name){
                                                           column <- getElement(tpm_data, 
                                                                                paste(column_name))
                                                           if(is.numeric(column)){
                                                             #is.numeric to avoid trying to scale the Geneid column
                                                             tpm_scale_factor <- 1e6 / sum(column)
                                                             tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                                           }else{
                                                             tpm_scaled <- column
                                                           }
                                                           return(tpm_scaled)
                                                         }) %>% 
                                    bind_cols
                                  colnames(rescaled_tpm) <-  colnames(list_element$tpm)
                                  list_element$tpm <- rescaled_tpm
                                  return(
                                    list_element
                                  )
                                })
names(twentyfour_ct_rescale) <- unlist(lapply(twentyfour_ct_rescale, getElement, 'experiment'))
lapply(twentyfour_ct_rescale, function(list_element){
  apply(list_element$tpm[,-1], 2, sum) # == 1e6
})
lapply(twentyfour_ct_rescale, function(list_element){
  nrow(list_element$tpm) # == 5363
})

write_xlsx(twentyfour_ct_rescale$rMT$tpm,
           'tl19-07_TPM_rMT_nopseudoRescaled_labelsStage_allSamples.xlsx')
write_xlsx(twentyfour_ct_rescale$SBP$tpm,
           'tl19-07_TPM_SBP_nopseudoRescaled_labelsStage_allSamples.xlsx')

#Setting twentyfour_ct core dataset - removed pseudogene and re-scaled.
twentyfour_ct <- twentyfour_ct_rescale

```

Removing the pseuodgene then re-scaling so that all samples add up to a million TPM again.

```{r shortform_tpm_df}
## Setting core dataset twentyfour_avgtpm and twentyfour_avgtpm_noschz - both rescaled w/o the pseudogene.

#Generate short-form version of twentyfour_top  using the mean_tpm:
twentyfour_avgtpm <- lapply(twentyfour_top, 
       function(experiment_element){
         tpm_df <- dcast(experiment_element$tpm, 
                         Geneid ~ stage, 
                         value.var = 'mean_tpm') %>% 
           filter(!Geneid %in% 'PCHAS_0420800')

         #Remove pseudogene and rescale
         rescaled_tpm <- lapply(colnames(tpm_df), 
                                function(column_name){
                                  column <- getElement(tpm_df, 
                                                       paste(column_name))
                                  if(is.numeric(column)){
                                    #is.numeric to avoid trying to scale the Geneid column
                                    tpm_scale_factor <- 1e6 / sum(column)
                                    tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                  }else{
                                    tpm_scaled <- column
                                  }
                                  return(tpm_scaled)
                                }) %>% 
           bind_cols
         colnames(rescaled_tpm) <-  colnames(tpm_df)

         return(
           list(tpm = rescaled_tpm,
                experiment = experiment_element$experiment)
         )
       })

names(twentyfour_avgtpm) <- sapply(twentyfour_avgtpm, getElement, 'experiment')

write_xlsx(twentyfour_avgtpm$rMT$tpm,
           'tl19-07_TPM_rMT_nopseudoRescaled_labelsStage_meanTPM.xlsx')
write_xlsx(twentyfour_avgtpm$SBP$tpm,
           'tl19-07_TPM_SBP_nopseudoRescaled_labelsStage_meanTPM.xlsx')

twentyfour_avgtpm_labelTimepoints <- twentyfour_avgtpm

colnames(twentyfour_avgtpm_labelTimepoints$rMT$tpm) <- ifelse(
  !colnames(twentyfour_avgtpm_labelTimepoints$rMT$tpm) %in% c('Geneid', 'sch.culture'),
  times_stages_table$time[
    match(colnames(twentyfour_avgtpm_labelTimepoints$rMT$tpm),
          times_stages_table$r_stage_diff_count)
  ],
  colnames(twentyfour_avgtpm_labelTimepoints$rMT$tpm)
)
write_xlsx(twentyfour_avgtpm_labelTimepoints$rMT$tpm,
           'tl19-07_TPM_rMT_nopseudoRescaled_labelsTime_meanTPM.xlsx')

colnames(twentyfour_avgtpm_labelTimepoints$SBP$tpm) <- ifelse(
  !colnames(twentyfour_avgtpm_labelTimepoints$SBP$tpm) %in% c('Geneid', 'sch.culture'),
  times_stages_table$time[
    match(colnames(twentyfour_avgtpm_labelTimepoints$SBP$tpm),
          times_stages_table$r_stage_diff_count)
  ],
  colnames(twentyfour_avgtpm_labelTimepoints$SBP$tpm)
)
write_xlsx(twentyfour_avgtpm_labelTimepoints$SBP$tpm,
           'tl19-07_TPM_SBP_nopseudoRescaled_labelsTime_meanTPM.xlsx')

# lapply(twentyfour_avgtpm,
#        function(tmp){
#          apply(
#            tmp$tpm[-1],
#            2,
#            sum
#          )
#        })

#Above is a check that they all add up to 1e6

# Setting twentyfour_avgtpm_noschz core dataset

twentyfour_avgtpm_noschz <- lapply(twentyfour_avgtpm, 
                                   function(experiment_data){
                                     tpm_df <- experiment_data$tpm
                                     tpm_df <- tpm_df[!str_detect(colnames(tpm_df), pattern = 'sch')]
                                     # Remove the column with 'sch' pattern in it = schizonts
                                     list(tpm = tpm_df,
                                          experiment = experiment_data$experiment)
                                   })

# lapply(twentyfour_avgtpm_noschz,
#        function(tmp){
#          apply(
#            tmp$tpm[-1],
#            2,
#            sum
#          )
#        })

#Above is a check that they all add up to 1e6

```

# 3D PCA to save

```{r 3d_PCA_save_rgl}
library('rgl')
library('magick')

twentyfour_ct_rescale_allTPM <- lapply(names(twentyfour_ct_rescale), 
                                       function(list_name){
                                         getElement(twentyfour_ct_rescale, 
                                                    paste(list_name))$tpm %>% 
                                           select(!matches('sch')) %>% 
                                           rename_with(~ paste0(list_name,
                                                                '_',
                                                                .x),
                                                       matches('[[:digit:]]{2}h'))
                                       }) %>% 
  Reduce(merge, .) %>% 
  column_to_rownames(var = 'Geneid')
# apply(twentyfour_ct_rescale_allTPM, 2, function(x) 1e10*(sum(x) - 1e6))

twentyfour_ct_rescale_allTPM_pca <- prcomp(twentyfour_ct_rescale_allTPM,
       center = TRUE, 
       scale. = TRUE)

twentyfour_ct_rescale_allTPM_pca_var <- get_pca_var(twentyfour_ct_rescale_allTPM_pca)
twentyfour_ct_rescale_allTPM_pca_coords <- data.frame(twentyfour_ct_rescale_allTPM_pca_var$coord)

twentyfour_ct_rescale_allTPM_pca_coords <- data.frame(
  sample = rownames(twentyfour_ct_rescale_allTPM_pca_var$coord),
  PC1 = twentyfour_ct_rescale_allTPM_pca_coords$Dim.1, 
  PC3 = twentyfour_ct_rescale_allTPM_pca_coords$Dim.3, 
  PC2 = twentyfour_ct_rescale_allTPM_pca_coords$Dim.2
) %>% 
  dplyr::mutate(stage = gsub(sample,
                             pattern = "_[[:digit:]]$", 
                             replacement = "")) %>% 
  dplyr::mutate(time = str_split_fixed(stage, 
                                       pattern = "_", 
                                       n = 2)[, 2]) %>% 
  dplyr::mutate(transmission = str_split_fixed(sample, 
                                               pattern = "_", 
                                               n = 3)[, 1])

colornames <- viridis(length(unique(twentyfour_ct_rescale_allTPM_pca_coords$time)))
colornames <- setNames(colornames, sort(unique(twentyfour_ct_rescale_allTPM_pca_coords$time)))

twentyfour_ct_rescale_allTPM_pca_coords$cols <- colornames[
  match(twentyfour_ct_rescale_allTPM_pca_coords$time,
        names(colornames))
]

pl <- plot_ly(twentyfour_tpm_pca_coords, x = ~PC1, y = ~PC2, 
              z = ~PC3, mode = "markers", type = "scatter3d", color = ~time, 
              text = ~sample, colors = colornames, marker = list(opacity = 1, 
                                                                 size = 5), symbol = ~transmission, symbols = c("circle", 
                                                                                                                "x")) %>% layout(scene = list(xaxis = list(title = paste0("PC1 ", 
                                                                                                                                                                          pc1_importance, "%")), yaxis = list(title = paste0("PC2 ", 
                                                                                                                                                                                                                             pc2_importance, "%")), zaxis = list(title = paste0("PC3 ", 
                                                                                                                                                                                                                                                                                pc3_importance, "%"))))

plot3d(
  twentyfour_ct_rescale_allTPM_pca_coords$PC1,
  twentyfour_ct_rescale_allTPM_pca_coords$PC2,
  twentyfour_ct_rescale_allTPM_pca_coords$PC3,
  col = twentyfour_ct_rescale_allTPM_pca_coords$cols,
  type = "s", 
  radius = .025
)

# colors <- c("royalblue1", "darkcyan", "oldlace")
# iris$color <- colors[ as.numeric( as.factor(iris$Species) ) ]
# # Static chart
# plot3d( iris[,1], iris[,2], iris[,3], col = iris$color, type = "s", radius = .2 )
# 
# # We can indicate the axis and the rotation velocity
# play3d( spin3d( axis = c(0, 0, 1), rpm = 20), duration = 10 )
```

# Amount of _cir_ expression per stage/sample

```{r amount of cir expression}

twentyfour_ct_comb_cir <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
  mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
  mutate(diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3]) %>% 
  mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
  filter(!(time %in% c('sch', 'Unknown'))) %>% 
  group_by(sample, stage, transmission, time, diff_count) %>% 
  dplyr::summarise(sum_tpm = sum(tpm))

cir_plot <- twentyfour_ct_comb_cir

  #Want it to be in the right order, but the line function won't work with a discrete scale
  order_diff_counts <- c("Early.Rings",
                         "Early.Mid.Rings",
                         "Late.Mid.Rings",
                         "Late.Rings",
                         "Rings.Troph.conversion",
                         "Early.Trophs",
                         "Mid.Trophs",
                         "Late.Trophs")
  order_for_plot <- data.frame(order_diff_counts,
                               order = seq(1, length(order_diff_counts), 1))
  cir_plot <- mutate(cir_plot, 
                     order_plot = order_for_plot$order[match(diff_count, 
                                                             order_for_plot$order_diff_counts)])

# cir_plot <- filter(twentyfour_ct_comb_cir, !(time %in% c('schz', 'Unknown'))) %>% 
#   mutate(time = as.numeric(time))

p <- ggplot(data = cir_plot, 
            aes(x = order_plot, 
                y = sum_tpm,
                fill = transmission)) +
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
  geom_smooth(method = 'loess', se = FALSE, color = 'black', aes(linetype = transmission)) +
  scale_fill_manual(values = cbPalette) +
  scale_linetype_manual(values = c('dotted', 'dashed')) +
    scale_x_continuous(breaks = order_for_plot$order, 
                       labels = order_for_plot$order_diff_counts) +
  ylab(expression('Total'~italic(pir)~'TPM')) +
  theme_classic() +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
p
ggsave(p, filename = 'tl19-07_24hchabaudi_pirexpression_barchartLOESS.pdf')

total_tpm_expression_stacked(list(cir = cir_id),
                             stack_genes = c('PCHAS_0101200'),
                             stack_label = 'Ancestral pir?',
                             stack_grouping = FALSE)
total_tpm_expression_stacked(list_genes = list(cir = cir_id),
                             stack_genes = c('PCHAS_0101200', 
                                             # 'PCHAS_0524800',
                                             'PCHAS_1100300'),
                             stack_label = 'pir labels',
                             stack_grouping = FALSE,
                             save = TRUE)
```

##  Stacked bar chart

```{r function_total_tpm_expression_stacked_all}
total_tpm_expression_stacked_all <- function(
    list_genes, 
  save = FALSE,
  tpm_data = twentyfour_ct,
  stage_str_extract_regex_pattern = '(?<=[[:digit:]]h_).+(?=_[[:digit:]]+)|sch.culture',
  stack_genes = cir_id,
  stack_names = c('Yes',
                  'No'),
  stack_label = 'Cir genes',
  stack_grouping = TRUE
){
  lapply(names(list_genes), 
         function(list_name){
           #list_name <- names(list_genes)[1]
           paste(list_name)
           gene_id <- getElement(list_genes, paste(list_name))
           list_tpms_line <- lapply(tpm_data, 
                                    function(input){
                                      tpm_data <- input$tpm %>% 
                                        filter(Geneid %in% gene_id) %>% 
                                        melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
                                        mutate(stage = str_extract(sample, 
                                                                   pattern = paste(stage_str_extract_regex_pattern)),
                                               stack_group = Geneid
                                               ) %>% 
                                        group_by(stage, sample, stack_group) %>% 
                                        dplyr::summarise(total_tpm = sum(tpm))

                                      #Make stages with no cirs in the sample give a zero 
                                      if(!all(colnames(input$tpm) %in% tpm_data$sample)){
                                        missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
                                        missing_stages_data_zero = data.frame(
                                          lapply(missing_stages, 
                                                 function(missing_st){
                                                   data.frame(stage = unlist(
                                                     str_extract(
                                                       missing_st, 
                                                       pattern = paste(stage_str_extract_regex_pattern))),
                                                     sample = missing_st,
                                                     total_tpm = 0)
                                                 })
                                        )
                                        tpm_data = bind_rows(tpm_data, missing_stages_data_zero)
                                      }

             #Add experiment column
             tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
             max_tpm = if(max(tpm_data$total_tpm) < 10){
               c(seq(0,mround(max(tpm_data$total_tpm),1),1))
             }else{
               c(seq(0,mround(max(tpm_data$total_tpm),5),5))
             }

             #Get the average data for the line graph
             tpm_data_line = tpm_data %>% 
               ungroup %>% 
               group_by(stage, experiment, stack_group) %>% 
               dplyr::summarise(total_tpm_median = median(total_tpm),
                                Number_of_replicates = n()) %>% 
               arrange(stage)

             return(tpm_data_line)
           })

           tpm_data_line <- bind_rows(list_tpms_line)
           #Apply ordering if it is specified
           if(!is.null(order_vector)){
             tpm_data_line$stage = factor(tpm_data_line$stage, 
                                          levels = ordering(tpm_data_line$stage,
                                                            order_vector))
           }

           tpm_data_line

           #Calculate the range for the y value to make it look nicer
           stacked_max <- tpm_data_line %>% 
                                ungroup %>% 
                                group_by(stage, experiment) %>% 
                                dplyr::summarise(max_tpm = sum(total_tpm_median)) %>% 
                                .$max_tpm %>% 
             max

           max_tpm <- c(
             seq(
               0,
               mround(
                 stacked_max,
                 500),
               length.out = 11)
           )

           y <- ggplot(data = tpm_data_line, 
                       aes(
                         x = experiment,
                         y = total_tpm_median, 
                         fill = stack_group
                       )
           ) +
             geom_bar(position = position_stack(),
                      stat = 'identity',
                      colour = 'black',
                      size = 0.1) +
             scale_fill_manual(
               breaks = c(stack_genes, 'Other'),
               values = c( #Remove the stack_genes from the 'grey' cir group
                 setNames(cbPalette[1:length(c(stack_genes))],
                          c(stack_genes)),
                 setNames(rep_len('white', 
                                  length.out = length(
                                    str_subset(unique(tpm_data_line$stack_group),
                                               pattern = stack_genes, 
                                               negate = TRUE)
                                  )),
                          str_subset(unique(tpm_data_line$stack_group),
                                     pattern = stack_genes, 
                                     negate = TRUE)
                 )
               )
               #Need to have limits and values otherwise when there are only cir genes (i.e. only 'Yes') the colour 
               # palette will mess up and only use one, wrong colour. Limits tells it there are always two values.
             ) +
             scale_color_manual('black') %>% 
             labs(fill = stack_label) +
             scale_y_continuous(breaks = max_tpm) +
             theme_classic() +
             scale_size(range = c(0,3)) +
             ylab(
               bquote('total '~italic(.(list_name))~' TPM')
             ) + 
             facet_grid(.~stage, 
                        scales = "free_x",
                        space = "free_x") +
             theme(
               axis.title.x=element_blank(),
               axis.text.x=element_text(angle = 90, 
                                        size = 5, 
                                        vjust = 0.5),
               axis.text.y = element_text(size = 5),
               axis.title.y = element_text(size = 8),
               legend.title = element_text(size = 7),
               legend.text = element_text(size = 5),
               legend.key.size = unit(3,'mm'),
               legend.key.width = unit(2,'mm'),
               strip.text.x = element_text(size=4, 
                                           face = 'bold', 
                                           angle = 90),
               strip.background = element_rect(
                 color="black", 
                 size=0.1, 
                 linetype="solid"
               )
             )

           #This bit is to change the colour of the facet boxes

           g <- ggplot_gtable(ggplot_build(y))
           strip_t <- which(grepl('strip-t', g$layout$name))
           stage_vec <- unique(
             levels(tpm_data_line$stage)
           )
           fills <- case_when(str_detect(stage_vec, pattern = 'Asexual') ~ 'coral1' ,
                              str_detect(stage_vec, pattern = 'Liver') ~ 'plum2',
                              str_detect(stage_vec, pattern = 'Gametocyte') ~ 'lightskyblue1',
                              str_detect(stage_vec, pattern  = 'Blood-Meal|Sporozoite|Ookinete') ~ 'lightgreen')
           k <- 1
           for (i in strip_t) {
             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
             k <- k+1
           }

           if(save == TRUE){
             ggsave2(paste0('tl19-07_',
                            gsub(list_name, 
                                 pattern  = ' ', 
                                 replacement = '_'),
                            '_totalTPM_barchart.pdf'),
                     plot = g,
                     dpi = 300,
                     width = 15,
                     height = 10,
                     units = 'cm')
           }

           grid.newpage()
           grid.draw(g)

         })
}

total_tpm_expression_stacked_all(list_genes = list(cir = cir_id),
                             stack_genes = c('PCHAS_0101200', 
                                             # 'PCHAS_0524800',
                                             'PCHAS_1100300'),
                             stack_label = 'pir genes',
                             stack_names = c('-', 
                                             'Other'),
                             stack_grouping = FALSE,
                             save = TRUE)
```

Curves represent a local regression (LOESS) function.

The pattern of cir expression is the same between SBP and MT (more-or-less) and at the 17.00 and 20.00 time points it is possible that the amount of cir expression is actually greater in SBP v MT. Both transmission methods show an peak of expression around 11-14h, correlating with the onset of sequestration?

For completeness the schizont sample is included below:

```{r}
cir_plot_schz <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
  mutate(transmission = str_extract(stage, pattern = '^(rMT|SBP)')) %>%
  mutate(time = str_extract(stage, pattern = '(?<=[[:alpha:]]_)([[:alpha:]]+\\.[[:alpha:]]+)+$')) %>% 
  filter(grepl(time, pattern =  c('sch'))) %>% 
  group_by(sample, stage, transmission, time) %>% 
  dplyr::summarise(sum_tpm = sum(tpm))
p <- ggplot(data = cir_plot_schz, 
            aes(x = time, 
                y = sum_tpm,
                fill = transmission)) +
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = cbPalette) +
  ylab(expression('Total'~italic(cir)~'TPM')) +
  scale_y_continuous(limits = c(0, 10000)) +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  theme_classic()
p
```

# Heatmap of top _pir_ expression and orthology across _Plasmodium_

Need to verify that the strains removed don't have any 'extra' orthologs compared to the strain that I keep.

```{r download_ortholog_data}
chab_pir_ortho <- read_csv('PlasmoDBv48_PchabaudiPIRs_transformedbyorthology.csv') %>% 
  mutate(Geneid = `Gene ID`,
         Input_Geneid = `Input Ortholog(s)`) %>% 
  filter(!(Organism %in% 'Plasmodium chabaudi chabaudi')) %>% 
  tidyr::separate_rows(Input_Geneid, 
                       sep = ',') %>% 
  mutate(Organism = str_replace(Organism,
                                pattern = 'lasmodium',
                                replacement = '\\.'))

#Remove paralogs and split rows out by the Input sequences

chab_pir_ortho_synt <- read_csv('PlasmoDBv48_PchabaudiPIRs_transformedbyorthology_syntenic.csv')

chab_pir_ortho <- filter(
  chab_pir_ortho,
  !Organism %in% c(
    'P. vivax Sal-1',
    'P. yoelii yoelii 17X',
    'P. yoelii yoelii 17XML',
    'P. vinckei vinckei strain vinckei'
  )
) %>% 
  mutate(
    syntenic = ifelse(`Geneid` %in% chab_pir_ortho_synt$`Gene ID`,
                           'Yes',
                           'No'),
         Organism = str_extract(Organism,
                                pattern = 'P. [[:alpha:]]+')
  )

# unique(chab_pir_ortho$Organism)

chab_pir_ortho %>%
  mutate(species = str_extract(Geneid, pattern = '^.+(?>_)')) %>%
  group_by(Input_Geneid) %>%
  dplyr::summarise(orthologs = paste(unique(`Organism`), collapse = ','))  %>%
  tidyr::separate_rows(orthologs, sep = ',') %>%
  group_by(Input_Geneid) %>%
  dplyr::summarise(num = n()) %>%
  arrange(desc(num)) %>%
  mutate(product = chabaudi_gene_info$handy_descriptions[match(Input_Geneid,
                                                         chabaudi_gene_info$Geneid)])

chab_pir_ortho %>%
  mutate(species = str_extract(Geneid, 
                               pattern = '^.+(?>_)')) %>%
  group_by(Input_Geneid) %>%
  dplyr::summarise(orthologs = paste(unique(`Organism`), 
                                     collapse = ', '))

```

```{r ortholog_heatmap}

#   #pir_family_name <- names(pir_genes_list_ortho)[1]
#   pir_family <-  getElement(pir_genes_list_ortho, 
#                             paste(pir_family_name))
#   #Need to convert this into a character vector for the filter function
#   pir_family <- apply(pir_family, 1, as.character)

threshold <- 0

#Want Number for whether the orthologue is present in each species
pir_ortholog_species <- chab_pir_ortho %>% 
  filter(Input_Geneid %in% cir_id) %>% 
  purrr::modify_if(is.character, 
            as.factor) %>% 
  group_by(Organism, 
           Input_Geneid, 
           .drop = FALSE) %>%
  #.drop must be FALSE otherwise dplyr will drop zero values
  select(Input_Geneid, 
         Organism, syntenic) %>%
  dcast(Input_Geneid ~ Organism, 
        value.var = 'syntenic', 
        fun.aggregate = function(syntenic){
          if(any(syntenic == 'Yes')){
            2
          }else if(length(syntenic) > 0){
            1
          }else{
            0
          }
        })

#Add in the genes which have no orthologs
pir_ortholog_species <- pir_ortholog_species %>% 
  tidyr::complete(Input_Geneid = cir_id)

  #Alter here whether you want only the genes with orthologs, or all genes even if they don't have an ortholog

  pir_ortholog_species <-  dplyr::rename(pir_ortholog_species, 
                                  rowname = Input_Geneid) %>%
    mutate(rowname = str_replace(as.character(rowname), 
                                 pattern = 'PCHAS_', 
                                 replacement = '')) %>%
    arrange(
      match(
        rowname,
        str_sort(
          rowname,
          numeric = TRUE
        )
      )
    ) %>%
    #Use str_sort because the character format is sorted differently to how I want it normally
    column_to_rownames()
  pir_ortholog_species_rownam <- rownames(pir_ortholog_species)
  pir_ortholog_species <- apply(pir_ortholog_species, 2, function(x){
    case_when(
      x == 0 ~ NA_character_,
      x == 1 ~ 'No',
      x == 2 ~ "Yes")
  })
  rownames(pir_ortholog_species) <- pir_ortholog_species_rownam

  #Not going to worry about whether they are pseudogenes or not for now

  ortho_gene_vec <- rownames(pir_ortholog_species)

  heatmap_row_name_fonts <- 10
  heatmap_col_name_fonts <- 10
  heatmap_col_title_fonts <- 15

  #Get the TPM data for the pirs family being looked at
  expt_tpms_top_pirs <- lapply(twentyfour_avgtpm_noschz, 
                               function(tpm_input){
                                 # print(paste0('Now doing ', tpm_input$experiment))
                                 tpms <- tpm_input$tpm %>% 
                                   filter(Geneid %in% cir_id)
                                 #Remove PCHAS_ from the gene names for simplicity
                                 rownames(tpms) = str_replace(as.character(tpms$Geneid), 
                                                              pattern = 'PCHAS_', 
                                                              replacement = '')
                                 tpms <- tpms[str_sort(rownames(tpms), 
                                                       numeric = TRUE),]
                                 tpms <- select(tpms, -Geneid)

    colnames(tpms) <- times_stages_table$time[match(colnames(tpms),
                                                    times_stages_table$r_stage_diff_count)]

    tpms <- tpms[,
                 order(colnames(tpms)),
                 drop = FALSE]
    #Ordering using the ordering function and order_vector defined in the functions section.

    tpms <- apply(tpms, 2, function(x){ifelse(x >= threshold, 
                                              x, 
                                              NA)}) #Get NA for values under threshold
    tpms <- log1p(tpms)
    if(tpm_input$experiment %>%  is.null){
      tpm_input$experiment = ''
    }
    list = list(tpms, tpm_input$experiment)
    names(list) = c('tpm','experiment')
    return(list)
  })

  #if they are all NA then we don't want it to proceed further
  if(
    all(
      unlist(
        lapply(
          expt_tpms_top_pirs, 
          function(x) sapply(x$tpm, 
                             is.na)  
        )
      )
    )
  ){stop()}

    maxcol = max(
      unlist(
        lapply(
          expt_tpms_top_pirs, 
          function(x) max(
            x$tpm,
            na.rm = TRUE
          )
        )
      )
    )
    #Get the max TPM value

    if(is.na(maxcol)|maxcol<1){maxcol <- 1}

    #Re-order both data.frames by the median highest expressed pirs

    tpm_order <- order(
      rowMedians(
        sapply(
          lapply(
            expt_tpms_top_pirs,
            function(x) rowMedians(x$tpm)
          ),
          rbind),
        na.rm = TRUE),
      decreasing = TRUE
    )[1:10]

    heatmaps = lapply(expt_tpms_top_pirs, 
                      function(tpms){
                        if(tpms$experiment != last(expt_tpms_top_pirs)$experiment){
                          #If it is the final element don't remove the row names
                          rownames(tpms$tpm) = c()
                        }
                        #Setting the name if one is not provided
                        title_name = tpms$experiment
                        #Drawing each heatmap
                        ht = Heatmap(tpms$tpm[tpm_order,, drop = FALSE], 
                                     name = 'log(TPM + 1)', 
                                     col = colorRamp2(c(threshold, 
                                                        maxcol/4, 
                                        maxcol/2, 
                                        3*maxcol/4, 
                                        maxcol), 
                                      viridis(5)),
                     cluster_rows = FALSE, 
                     cluster_columns = FALSE,
                     column_title = paste(title_name),
                     column_title_gp = gpar(fontsize = heatmap_col_title_fonts),
                     row_names_gp = gpar(fontsize = heatmap_row_name_fonts),
                     column_title_rot = 0,
                     column_names_gp = gpar(fontsize = heatmap_col_name_fonts),
                     na_col = 'white', 
                     border = TRUE,
                     border_gp = gpar(col = "black", 
                                      lwd = 2),
                     heatmap_legend_param = list(
                       title = 'log(TPM + 1)', 
                       at = mround(
                         c(log1p(threshold), 
                           seq(0, 
                               maxcol, 
                               length.out = 4)[-1]
                         ), 
                         base = 0.05)
                     )
        )
    })

    ortholog_for_heatmap <- as.matrix(pir_ortholog_species)[tpm_order,
                                                            !apply(
                                                              pir_ortholog_species[tpm_order,,drop = FALSE],
                                                              2,
                                                              function(ortho) all(is.na(ortho))
                                                            ),
                                                            drop = FALSE]

    ha = Heatmap(ortholog_for_heatmap,
                 name = 'Syntenic Orthologs',
                 na_col = 'white',
                 col = structure(rev(plasma(2)), 
                                 names = c('Yes', 'No')),
                 cluster_rows = FALSE, 
                 cluster_columns = FALSE,
                 column_names_gp = gpar(col = 'darkslategray', 
                                    fontsize = 7*(50/(ncol(ortholog_for_heatmap)+26)), 
                                    fontface = 'italic'),
                 show_column_dend = FALSE,
                 rect_gp = gpar(col = 'white'), 
                 width = unit(4, "cm"),
                 heatmap_legend_param = list(

                     )
    )

    ht_list = Reduce('+', heatmaps) #To add the list elements together
    ht_list <- ha + ht_list

    if(global_save_var){
      pdf(
        paste0('tl19-07_pir_',
               'orthologs_heatmap.pdf')
        # , width = 17
        , height = 5
      )
      draw(ht_list,
           row_title_side = 'left',
           column_title = 'Malaria stage',
           column_title_side = 'bottom',
           heatmap_legend_side = 'left',
           ht_gap = unit(0.75, "cm"))
      dev.off()
    }

    draw(ht_list,
           row_title_side = 'left',
           column_title = 'Malaria stage',
           column_title_side = 'bottom',
           heatmap_legend_side = 'left',
           ht_gap = unit(0.6, "cm"))

```

```{r heatmap_ortho_most_conserved}

    #Re-order by most conserved genes

    tpm_order <- order(
      apply(pir_ortholog_species, 
            1, 
            function(ortho_row) sum(!is.na(ortho_row))),
      decreasing = TRUE
    )[1:10]

    heatmaps = lapply(expt_tpms_top_pirs, 
                      function(tpms){
                        if(tpms$experiment != last(expt_tpms_top_pirs)$experiment){
                          #If it is the final element don't remove the row names
                          rownames(tpms$tpm) = c()
                        }
                        #Setting the name if one is not provided
                        title_name = tpms$experiment
                        #Drawing each heatmap
                        ht = Heatmap(tpms$tpm[tpm_order,, drop = FALSE], 
                                     name = 'log(TPM + 1)', 
                                     col = colorRamp2(c(threshold, 
                                                        maxcol/4, 
                                        maxcol/2, 
                                        3*maxcol/4, 
                                        maxcol), 
                                      viridis(5)),
                     cluster_rows = FALSE, 
                     cluster_columns = FALSE,
                     column_title = paste(title_name),
                     column_title_gp = gpar(fontsize = heatmap_col_title_fonts),
                     row_names_gp = gpar(fontsize = heatmap_row_name_fonts),
                     column_title_rot = 90,
                     column_names_gp = gpar(fontsize = heatmap_col_name_fonts),
                     na_col = 'white', border = TRUE,
                     heatmap_legend_param = list(
                       title = 'log(TPM + 1)', 
                       at = mround(
                         c(log1p(threshold), 
                           seq(0, 
                               maxcol, 
                               length.out = 4)[-1]
                         ), 
                         base = 0.05)
                     )
        )
    })

    ha = Heatmap(as.matrix(pir_ortholog_species)[tpm_order,, drop = FALSE],
                 name = 'Syntenic Orthologs',
                 na_col = 'white',
                 col = structure(rev(plasma(2)), 
                                 names = c('Yes', 'No')),
                 cluster_rows = FALSE, 
                 cluster_columns = FALSE,
                 column_names_gp = gpar(col = 'darkslategray', 
                                        fontsize = 4*(50/(ncol(pir_ortholog_species)+26)), 
                                        fontface = 'italic'),
                 show_column_dend = FALSE,
                 rect_gp = gpar(col = 'white'), 
                 width = unit(4, "cm"),
                 heatmap_legend_param = list(

                     )
                 )

    ht_list = Reduce('+', heatmaps) #To add the list elements together
    ht_list <- ha + ht_list
    # pdf(
    #   paste0('tl19-07_pir_',
    #          'orthologs_heatmap.pdf')
    #   # , width = 17
    #   , height = 5
    # )
    draw(ht_list,
         row_title_side = 'left',
         column_title = 'Malaria stage',
         column_title_side = 'bottom',
         heatmap_legend_side = 'left')
    # dev.off()

```

# Figure 2b - Heatmap of top S and L gene expression.

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9',
        '#E69F00',
        '#999999'),
  sort(
    unique(cir_clade_info$SorL)
  )
)

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9',
    '#009E73',
    '#E69F00',
    '#0072B2',
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_clade_info$clade)
  )
)

```

```{r figure2b_heatmap_table}

#Get the TPM data for the pirs family being looked at
highest_SorL_24h <- lapply(twentyfour_avgtpm_noschz, 
                          function(tpm_input){

                            tpm_data_SorL_stackgenes_top_table <- tpm_input$tpm %>% 
                              filter(Geneid %in% cir_id) %>% 
                              melt(id.vars = 'Geneid', 
                                   variable.name = 'stage', 
                                   value.name = 'total_tpm') %>% 
                              mutate(total_tpm = as.numeric(total_tpm),
                                     SorL = cir_clade_info$SorL[match(Geneid, cir_clade_info$Geneid)],
                                     stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
                              filter(str_detect(SorL, 
                                                pattern = 'ancestral|unk', 
                                                negate = TRUE)) %>% 
                              group_by(stage) %>% 
                              mutate(sum_tpm = sum(total_tpm)) %>% 
                              group_by(stage, 
                                       SorL) %>% 
                              slice_max(order_by = total_tpm,
                                        n = 1) %>% 
                              mutate(total_tpm = round(total_tpm, digits = 2),
                                     prop_tpm = round(total_tpm/sum_tpm, digits = 2)) %>% 
                              select(stage, Geneid, SorL, total_tpm, prop_tpm)

                            write.csv(tpm_data_SorL_stackgenes_top_table,
                                      file = paste0('tl19-07_combined_SorL_topperstage_table_', 
                                                    tpm_input$experiment, 
                                                    '.csv'))

                            write_xlsx(tpm_data_SorL_stackgenes_top_table,
                                       paste0('tl19-07_combined_SorL_topperstage_table_', 
                                              tpm_input$experiment, 
                                              '.xlsx'))

                            SorL_kable_index <- rep.int(2, 
                                                        times = length(unique(tpm_data_SorL_stackgenes_top_table$stage)))
                            names(SorL_kable_index) <- unique(tpm_data_SorL_stackgenes_top_table$stage)

                            return(
                              list(dataframe = tpm_data_SorL_stackgenes_top_table, 
                                   table = 
                                     tpm_data_SorL_stackgenes_top_table %>% 
                                     kable(caption = paste(tpm_input$experiment) )%>% 
                                     kable_styling() %>% 
                                     pack_rows(index = SorL_kable_index) %>% 
                                     scroll_box(width = "100%", height = "500px"),
                                   experiment = tpm_input$experiment
                              )
                            )
                          }
)

lapply(highest_SorL_24h, function(expt) expt$table)

```

```{r figure2b_heatmap}

heatmap_row_name_fonts <- 20
heatmap_col_name_fonts <- 15

highest_SorL_pirs_bystage <- sort(
  unique(
    unlist(
      lapply(
        highest_SorL_24h, function(expt) expt$dataframe$Geneid)
    )))

highest_SorL_pirs_bystage_subfams <- data.frame(
  Geneid = str_replace(as.character(highest_SorL_pirs_bystage), 
                                    pattern = 'PCHAS_', 
                                    replacement = ''),
                       subfamily = cir_clade_info$SorL[match(highest_SorL_pirs_bystage
                                                             , cir_clade_info$Geneid)]
  )

ha = Heatmap(
  as.matrix(highest_SorL_pirs_bystage_subfams$subfamily),
  name = 'Subfamily',
  na_col = 'white',
  col = SorL_col,
  cluster_rows = FALSE, 
  cluster_columns = FALSE,
  column_names_gp = gpar(col = 'darkslategray', 
                         fontsize = 12, 
                         fontface = 'italic'),
  show_column_dend = FALSE,
  rect_gp = gpar(col = 'white'), 
  width = unit(1, "cm"),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10)
  ),
  show_heatmap_legend = FALSE
  # Going to hide the legend so I can make it manually later with the grid.points that I use in the heatmap body.
)

#Get the TPM data for the pirs family being looked at
expt_tpms_top_pirs <- lapply(twentyfour_avgtpm_noschz, 
                             function(tpm_input){
                               # print(paste0('Now doing ', tpm_input$experiment))
                               tpms <- tpm_input$tpm %>% 
                                 filter(Geneid %in% highest_SorL_pirs_bystage) %>% 
                                 as.data.frame

                               colnames(tpms) <- ifelse(
                                 !colnames(tpms) %in% 'Geneid',
                                                        times_stages_table$time[match(colnames(tpms),
                                                                                      times_stages_table$r_stage_diff_count)],
                                                        'Geneid'
                               )

                               tpms <- tpms[,
                                            order(colnames(tpms)),
                                            drop = FALSE]
                               #Ordering using the ordering function and order_vector defined in the functions section.

                               #Data for the grid.dots for heatmap to show max for each stage
                               max_SorL <- mutate(tpms, 
                                                  SorL = cir_clade_info$SorL[match(Geneid, 
                                                                                   cir_clade_info$Geneid)]) %>% 
                                 group_by(SorL) %>% 
                                 mutate(across(-matches('Geneid|SorL'), 
                                               ~ ifelse(.x == max(.x), 
                                                        SorL, 
                                                        NA))) %>% 
                                 ungroup() %>% 
                                 select(-Geneid, -SorL)

                               #Remove PCHAS_ from the gene names for simplicity
                               rownames(tpms) <- str_replace(as.character(tpms$Geneid), 
                                                             pattern = 'PCHAS_', 
                                                             replacement = '')
                               tpms <- tpms[str_sort(rownames(tpms), 
                                                     numeric = TRUE),]
                               tpms <- select(tpms, -Geneid)

                               tpms <- apply(tpms, 2, function(x){ifelse(x >= threshold, 
                                                                         x, 
                                                                         NA)}) #Get NA for values under threshold
                               tpms <- log1p(tpms)
                               if(tpm_input$experiment %>%  is.null){
                                 tpm_input$experiment = ''
                               }
                               list = list(tpm = tpms, 
                                           experiment = tpm_input$experiment,
                                           max_SorL = max_SorL)
                               return(list)
                             })

    maxcol <- max(
      unlist(
        lapply(
          expt_tpms_top_pirs, 
          function(x) max(
            x$tpm,
            na.rm = TRUE
          )
        )
      )
    )
    #Get the max TPM value

heatmaps = lapply(expt_tpms_top_pirs, 
                  function(tpms){

                    if(tpms$experiment != last(expt_tpms_top_pirs)$experiment){
                      #If it is the final element don't remove the row names
                      rownames(tpms$tpm) = c()
                    }
                    #Setting the name if one is not provided
                    title_name = tpms$experiment
                    #Drawing each heatmap
                    ht = Heatmap(tpms$tpm, 
                                 name = 'log(TPM + 1)', 
                                 col = colorRamp2(c(threshold, 
                                                    maxcol/4, 
                                                    maxcol/2, 
                                                    3*maxcol/4, 
                                                    maxcol), 
                                                  viridis(5)),
                                 cluster_rows = FALSE, 
                                 cluster_columns = FALSE,
                                 column_title = paste(title_name),
                                 column_title_gp = gpar(fontsize = heatmap_col_title_fonts),
                                 row_names_gp = gpar(fontsize = heatmap_row_name_fonts),
                                 column_title_rot = 0,
                                 column_names_rot = 0,
                                 column_names_centered = TRUE,
                                 column_names_gp = gpar(fontsize = heatmap_col_name_fonts),
                                 na_col = 'white', 
                                 border = TRUE,
                                 heatmap_legend_param = list(
                                   title = 'log(TPM + 1)', 
                                   at = mround(
                                     c(log1p(threshold), 
                                       seq(0, 
                                           maxcol, 
                                           length.out = 4)[-1]
                                     ), 
                                     base = 0.05),
                                   labels_gp = gpar(fontsize = 10),
                                   title_gp = gpar(fontsize = 10, fontface = "bold")
                                 ),
                                 cell_fun = function(j, i, x, y, width, height, fill) {
                                   if(!is.na(as.matrix(tpms$max_SorL)[i, j])){
                                     ifelse(
                                       as.matrix(tpms$max_SorL)[i, j] == 'S',
                                       grid.points(x, y, pch = 16, size = unit(3, "mm")),
                                       grid.points(x, y, pch = 18, size = unit(4, "mm"))
                                     )
                                   }
                                 }
                    )
                  })

ldg_subfam <- Legend(labels = c('S', 'L'),
                     type = 'points',
                     pch = c(16,18),
                     size = unit.c(unit(2, "mm"), 
                                   unit(3, "mm")),
                     background = SorL_col[c('S','L')],
                     title = 'Subfamily',
                     title_gp = gpar(fontsize = 10, fontface = "bold"))

ht_list = Reduce('+', heatmaps) #To add the list elements together
ht_list <- ha + ht_list
# pdf('tl19-07_pir_SorLhighest_figure2c_heatmap.pdf'
#     , width = 10
#     , height = 4
# )
draw(ht_list,
     row_title_side = 'left',
     column_title = 'Malaria stage',
     column_title_side = 'bottom',
     column_title_gp = gpar(fontsize = 16),
     heatmap_legend_side = 'left',
     ht_gap = unit(0.5, "cm"),
     heatmap_legend_list = list(ldg_subfam))
# dev.off()
```

# Ancestral and rodent-conserved gene expression curves

```{r ancestral_pir_curve}
gene_graph_24h <- function(input = twentyfour_ct_comb, 
                           genevec = cir_id,
                           genename,
                           use_times = FALSE) {
  cir_plot <- input$tpm %>% 
    filter(Geneid %in% genevec) %>% 
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
    mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
    mutate(diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3]) %>% 
    mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
    filter(!(time %in% c('sch', 'Unknown'))) %>% 
    group_by(sample, stage, transmission, time, diff_count) %>% 
    dplyr::summarise(sum_tpm = sum(tpm))

  #Want it to be in the right order, but the line function won't work with a discrete scale
  order_diff_counts <- c("Early.Rings",
                         "Early.Mid.Rings",
                         "Late.Mid.Rings",
                         "Late.Rings",
                         "Rings.Troph.conversion",
                         "Early.Trophs",
                         "Mid.Trophs",
                         "Late.Trophs")

  order_for_plot <- data.frame(order_diff_counts,
                               order = seq(1, 
                                           length(order_diff_counts), 
                                           1))
  cir_plot <- mutate(cir_plot, 
                     order_plot = order_for_plot$order[match(diff_count, 
                                                             order_for_plot$order_diff_counts)])
  if(use_times){
    order_for_plot$order_diff_counts <- times_stages_table$time[match(order_for_plot$order_diff_counts,
                                                                      times_stages_table$r_stage_diff_count)]
  }

  p <- ggplot(data = cir_plot, 
              aes(x = order_plot, 
                  y = sum_tpm,
                  fill = transmission)) +
    stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
    geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
    geom_smooth(method = 'loess', se = FALSE, color = 'black', aes(linetype = transmission)) +
    scale_fill_manual(values = cbPalette) +
    scale_linetype_manual(values = c('dotted', 'dashed')) +
    scale_x_continuous(breaks = order_for_plot$order, 
                       labels = order_for_plot$order_diff_counts) +
    ylab(as.expression(bquote('Total'~italic(.(genename))~'TPM'))) +
    xlab('Asexual blood stage development cycle') +
    theme_classic() +
    theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
  p
}

gene_graph_24h(genevec = 'PCHAS_0101200',
               genename = 'highly-conserved pir',
               use_times = TRUE)

gene_graph_24h(input = twentyfour_tpm_rescale_nopseudo_t,
               genevec = 'PCHAS_0101200',
               genename = 'highly-conserved pir')

ggsave(gene_graph_24h(genevec = 'PCHAS_0101200',
                      genename = 'highly-conserved pir'),
       filename = 'tl19-07_24hchabaudi_ancestral_expressionbarchartLOESS.pdf')
# gene_graph_24h(genevec = 'PCHAS_0524800',
#                genename = 'rodent-conserved pir')
# ggsave(gene_graph_24h(genevec = 'PCHAS_0524800',
#                       genename = 'rodent-conserved pir'),
#        filename = 'tl19-07_24hchabaudi_rodentpir_expressionbarchartLOESS.pdf')

cir_id_no_anc <- str_subset(cir_id, 
                            pattern = 'PCHAS_0101200', 
                            negate = TRUE)
gene_graph_24h(genevec = cir_id_no_anc,
               genename = 'pir without ancestral or rodent-conserved')

# gene_graph_24h(genevec = 'PCHAS_0420800', 
#                genename = 'rMT his-rich pseudogene')
```

## Expression of major expressed cirs rMT/SBP side-by-side.

_pir_ genes selected by being expressed above 1 TPM in all samples.

Adding in housekeeping genes for comparison:
HSP70 putative  PCHAS_0721000
MSP1 PCHAS_0831300
AMA1 PCHAS_0931000

```{r pirs_side-by-side_figure2c}

housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
                             'MSP1' = 'PCHAS_0831300',
                             'AMA1' = 'PCHAS_0931000')

cir_housekeeping_id <- c(cir_id, 
                         'PCHAS_0721000',
                         'PCHAS_0831300',
                         'PCHAS_0931000')

cir_plot <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown'))) %>% 
  group_by(Geneid) %>% 
  filter(any(tpm > 100)) %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a'))

cir_plot$Geneid <- ifelse(
  cir_plot$Geneid %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
  cir_plot$Geneid
)

cir_plot_order <- cir_plot %>% 
  filter(time == '14h',
         transmission == 'rMT') %>% 
  arrange(desc(tpm))

cir_plot <- mutate(cir_plot, 
                   Geneid = factor(Geneid, 
                                   levels = unique(cir_plot_order$Geneid)))
limits_cir <- c(0, mround(max(cir_plot$tpm), base = 100))

side_by_side_cir_plot <- function(time_vec = '11h',
                                  log_it = TRUE,
                                  save = TRUE,
                                  limits = limits_cir,
                                  point_size = 1){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

    # cir_plot %>% group_by(Geneid) %>% summarise(n())
    if(log_it){
      cir_plot$tpm <- log1p(cir_plot$tpm)
      limits = log1p(limits)
      breaks <- round(seq.int(0, 
                              max(limits), 
                              length.out = round(max(limits)) + 1, 
                              digits = 0), 
                      digits = 0)
    }else{
      breaks <- seq.int(0, max(limits), length.out = 11)
    }

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    cir_plot$subfam <- factor(
      cir_plot$subfam,
      levels = c(
        sort(unique(cir_plot$subfam)[!(unique(cir_plot$subfam) %in% 'ref')]),
        'ref'
      )
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      facet_grid(.~subfam, 
                 scales = "free_x",
                 space = "free_x") +
      theme_classic() +
      theme(
        axis.text.x = element_text(angle = 90)
      ) +
      ylab('log(TPM + 1)')+
      ggtitle(paste(unique(cir_plot$time)))
    if(save){
      ggsave2(
        paste0('tl20-05_side-by-side_barchart_facetbySubfam_',
               time_vec,
               '.png'),
        plot = y,
        dpi = 300,
        width = 28,
        height = 8,
        units = 'cm'
      )
    }
    return(y)
  }

  lapply(
    sort(unique(cir_plot$time)),
    side_by_side_cir_plot
  )

```

## Expression of the S7 pirs.

```{r pirs_side-by-side_different_criteria}

  cir_plot <- twentyfour_ct_comb$tpm %>% 
    filter(Geneid %in% cir_id) %>%
    melt(id.vars = 'Geneid', 
         variable.name = 'sample', 
         value.name = 'tpm') %>% 
    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
           subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]) %>% 
    filter(!(time %in% c('sch', 
                         'Unknown'))) %>% 
    group_by(Geneid)

  breaks = c(7,6,5,4, 3, 2, 1, 0)
  limits = c(0,7.2)
  point_size = 0.25

  lapply(sort(unique(cir_plot$subfam)),
  function(subfam_nam){
    cir_plot <- cir_plot %>% 
      filter(
        # all(tpm < 100),
        #      any(tpm > 50)
        subfam == subfam_nam
      ) %>% 
      arrange(desc(tpm))

    cir_plot_order <- cir_plot %>% 
      filter(time == '14h',
             transmission == 'rMT') %>% 
      arrange(desc(tpm))

    cir_plot <- mutate(cir_plot, 
                       Geneid = factor(Geneid, 
                                       levels = unique(cir_plot_order$Geneid)))

    lapply(
      sort(unique(cir_plot$time)),
      function(time_vec){

        cir_plot <- filter(cir_plot, 
                           time == time_vec)

      cir_plot$tpm <- log1p(cir_plot$tpm)

      cir_plot$Geneid <- factor(
        str_remove(cir_plot$Geneid, 
                   pattern = 'PCHAS_'),
        levels = str_remove(levels(cir_plot$Geneid), 
                            pattern = 'PCHAS_')
      )

      y <- ggplot(
        data = cir_plot,
        mapping = aes(
          x = Geneid,
          fill = transmission,
          y = tpm
        )
      ) +
        stat_summary(fun.y = median, 
                     fun.ymin = median, 
                     fun.ymax = median,
                     geom = "bar", 
                     width = 0.75, 
                     size = 0.25, 
                     position = 'dodge') +
        geom_point(stat = 'identity', 
                   position = position_dodge(width = 0.75),
                   size = point_size) +
        scale_y_continuous(breaks = breaks,
                           limits = limits) +
        ylab('ln(TPM + 1)')+
        facet_grid(.~subfam, 
                   scales = "free_x",
                   space = "free_x") +
        theme_classic() +
        theme(
          axis.text.x = element_text(angle = 90,
                                     size = 5)
        ) +
        ggtitle(paste(unique(cir_plot$time)))

      ggsave2(
        paste0('tl20-05_side-by-side_barchart_',
               subfam_nam,
               '_',
               time_vec,
               '.png'),
        plot = y,
        dpi = 300,
        width = 25,
        height = 10,
        units = 'cm'
      )
      return(y)
    })
})

```

## Total pir expression in each family across time.

```{r pir_family_expression}

  cir_plot <- twentyfour_ct_comb$tpm %>% 
    filter(Geneid %in% cir_id) %>%
    melt(id.vars = 'Geneid', 
         variable.name = 'sample', 
         value.name = 'tpm') %>% 
    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
           subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]) %>% 
    filter(!(time %in% c('sch', 
                         'Unknown'))) %>% 
    group_by(subfam, 
             time,
             transmission,
             sample) %>% 
    summarise(
      sum_tpm = sum(tpm)
    )

  breaks = seq.int(0, mround(max(cir_plot$sum_tpm), base = 100), length.out = 11)
    limits = c(0, mround(max(cir_plot$sum_tpm), base = 100))
    point_size = 0.25

  lapply(sort(unique(cir_plot$subfam)), 
         function(subfam_nam){

  # cir_plot_order <- cir_plot %>% 
  #   filter(time == '14h',
  #          transmission == 'rMT') %>% 
  #   arrange(desc(tpm))
  #   
  # cir_plot <- mutate(cir_plot, 
  #                    Geneid = factor(Geneid, 
  #                                    levels = unique(cir_plot_order$Geneid)))

    cir_plot <- filter(cir_plot, 
                       subfam == subfam_nam)

      # cir_plot$tpm <- log1p(cir_plot$sum_tpm)

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = time,
        fill = transmission,
        y = sum_tpm
      )
    ) +
    stat_summary(fun.y = median, 
                 fun.ymin = median, 
                 fun.ymax = median,
                 geom = "bar", 
                 width = 0.75, 
                 size = 0.25, 
                 position = 'dodge') +
    geom_point(stat = 'identity', 
               position = position_dodge(width = 0.75),
               size = point_size) +
    scale_y_continuous(breaks = breaks,
                       limits = limits) +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 90,
                                 size = 5)
    ) +
    ggtitle(subfam_nam)

    ggsave2(
      paste0('tl20-05_total-expression_by-subfam_',
             subfam_nam,
             '.png'),
      plot = y,
      dpi = 300,
      width = 25,
      height = 10,
      units = 'cm'
    )
  return(y)
  })
```

# Pir expression of AAPLs and ChAPLs.

```{r AapL-ChaAPl_expression}

  cir_plot <- twentyfour_ct_comb$tpm %>% 
    filter(Geneid %in% cir_id) %>%
    melt(id.vars = 'Geneid', 
         variable.name = 'sample', 
         value.name = 'tpm') %>% 
    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
           subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
           chapl_aapl = factor(cir_chapl_aapl_info$chapl_aapl[match(Geneid, 
                                                                    cir_chapl_aapl_info$Geneid)],
                               levels = sort(unique(cir_chapl_aapl_info$chapl_aapl[match(Geneid, 
                                                                                         cir_chapl_aapl_info$Geneid)])))) %>% 
    filter(!(time %in% c('sch', 
                         'Unknown')),
           !is.na(chapl_aapl) ) %>% 
    group_by(Geneid) %>% 
    # filter(
    #   # all(tpm < 100),
    #   #      any(tpm > 50)
    #   subfam == 'S7'
    #   ) %>% 
    arrange(desc(tpm))

  cir_plot_order <- cir_plot %>% 
    filter(time == '14h',
           transmission == 'SBP') %>% 
    arrange(desc(tpm))

  cir_plot <- mutate(cir_plot, 
                     Geneid = factor(Geneid, 
                                     levels = unique(cir_plot_order$Geneid)))

  breaks = c(7,6,5,4, 3, 2, 1, 0)
    limits = c(0,7)
    point_size = 0.25

  lapply(
    sort(unique(cir_plot$time)),
    function(time_vec){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

      cir_plot$tpm <- log1p(cir_plot$tpm)

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
    stat_summary(fun.y = median, 
                 fun.ymin = median, 
                 fun.ymax = median,
                 geom = "bar", 
                 width = 0.75, 
                 size = 0.25, 
                 position = 'dodge') +
    geom_point(stat = 'identity', 
               position = position_dodge(width = 0.75),
               size = point_size) +
    scale_y_continuous(breaks = breaks,
                       limits = limits) +
    facet_grid(.~chapl_aapl, 
             scales = "free_x",
             space = "free_x") +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 90,
                                 size = 5)
    ) +
    ggtitle(paste(unique(cir_plot$time)))

    ggsave2(
      paste0('tl20-05_side-by-side_barchart_chapl_aapl_',
             time_vec,
             '.png'),
      plot = y,
      dpi = 300,
      width = 25,
      height = 10,
      units = 'cm'
    )
  return(y)
})
```

# Pir expression across the genome

```{r pirs_across_genome}

  cir_plot <- twentyfour_ct_comb$tpm %>% 
    filter(Geneid %in% cir_id) %>%
    melt(id.vars = 'Geneid', 
         variable.name = 'sample', 
         value.name = 'tpm') %>% 
    mutate(
      stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
      transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
      diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
      time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
      subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
      chrom = paste0('chr',
                     str_extract(Geneid, 
                                 pattern = '(?<=PCHAS_)[[:digit:]]{2}'))
    ) %>% 
  filter(!(time %in% c('sch', 
                       'Unknown')),
         # transmission == 'rMT'
         ) %>% 
  group_by(Geneid)

breaks <- seq.int(0, mround(max(cir_plot$tpm), base = 100), length.out = 6)
limits <- c(0, mround(max(cir_plot$tpm), base = 100))

lapply(sort(unique(cir_plot$time)), 
       function(time_vec){
         cir_plot <- filter(cir_plot, 
                            time == time_vec)

           # cir_plot$tpm <- log1p(cir_plot$tpm)

         cir_plot$Geneid <- str_remove(cir_plot$Geneid, 
                                       pattern = 'PCHAS_')

         y <- ggplot(
           data = cir_plot,
           mapping = aes(
             x = Geneid,
             fill = transmission,
             y = tpm
           )
         ) +
           stat_summary(fun.y = median, 
                        fun.ymin = median, 
                        fun.ymax = median,
                        geom = "bar", 
                        width = 0.75, 
                        size = 0.25, 
                        position = 'dodge') +
           scale_y_continuous(breaks = breaks,
                              limits = limits) +
           scale_fill_manual(
             values = c('rMT' = 'blue', 
                        'SBP' = 'red')
           ) +
           facet_grid(.~chrom,
                    scales = "free_x",
                    space = "free_x") +
           theme_classic() +
           theme(
             axis.text.x = element_blank(),
             axis.ticks.x = element_blank(),
             panel.spacing = unit(0.1, "lines")
           ) +
           ggtitle(paste(unique(cir_plot$time)))

         # ggsave2(
         #   paste0('tl20-05_side-by-side_barchart_genome_facetChrom_',
         #          time_vec,
         #          '.png'),
         #   plot = y,
         #   dpi = 300,
         #   width = 25,
         #   height = 10,
         #   units = 'cm'
         # )

         return(y)
       })

```

# How does expression of the different clades change with time?

```{r clade_with-time_graph}
cir_plot <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id_no_anc_or_rodcon) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
  mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
  mutate(diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3]) %>% 
  mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
  filter(!(time %in% c('sch', 'Unknown'))) %>% 
  mutate(clade = reid_cir_clade$clade[match(Geneid, reid_cir_clade$Geneid)]) %>% 
  group_by(clade, transmission, time, diff_count) %>% 
  dplyr::summarise(median_tpm = median(tpm),
                   sd_tpm = sd(tpm),
                   n = n_distinct(Geneid))

# group_by(reid_cir_clade, clade) %>% 
#   dplyr::summarise(n())
# 
# cir_id[which(!(cir_id %in% reid_cir_clade$Geneid))]

order_diff_counts <- c("Early.Rings",
                       "Early.Mid.Rings",
                       "Late.Mid.Rings",
                       "Late.Rings",
                       "Rings.Troph.conversion",
                       "Early.Trophs",
                       "Mid.Trophs",
                       "Late.Trophs")
cir_plot$diff_count <- factor(cir_plot$diff_count, levels  = order_diff_counts)

p <- ggplot(data = cir_plot, 
            aes(x = diff_count, 
                y = log1p(median_tpm),
                fill = clade)) +
  geom_col(position = position_dodge()) +
  scale_fill_manual(values = cbPalette) +
  ylab(as.expression(bquote('Median'~italic('pir')~'log TPM + 1 by clade'))) +
  theme_classic() +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) + 
  facet_grid(. ~ transmission)
p

cir_plot <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
  mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
  mutate(diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3]) %>% 
  mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
  filter(!(time %in% c('sch', 'Unknown'))) %>% 
  mutate(clade = reid_cir_clade$clade[match(Geneid, reid_cir_clade$Geneid)]) %>% 
  mutate(clade = ifelse(grepl(clade, pattern = 'L'), 'Long', 
                        ifelse(grepl(clade, pattern = 'S'), 'Short',
                               'Unk'))) %>% 
  filter(clade != 'Unk')

order_diff_counts <- c("Early.Rings",
                       "Early.Mid.Rings",
                       "Late.Mid.Rings",
                       "Late.Rings",
                       "Rings.Troph.conversion",
                       "Early.Trophs",
                       "Mid.Trophs",
                       "Late.Trophs")
cir_plot$diff_count <- factor(cir_plot$diff_count, levels  = order_diff_counts)

p <- ggplot(data = cir_plot, 
            aes(x = diff_count,
                y = log1p(tpm),
                fill = clade,
                col = transmission)) +
  # geom_col(position = position_dodge()) +
  geom_boxplot() +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = c('dark blue','dark red')) +
  xlab('Stage') +
  ylab(as.expression(bquote('Median'~italic('pir')~'log TPM + 1 by clade'))) +
  theme_classic() +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
p

```

The only time that the SBP pir genes are being expressed higher than rMT are the burst of Short-forms in the Early Ring stage.

Splitting the _pir_ TPM graphs into S or L clades, and the sub-clades.

```{r total_tpm_pir_clades}
pir_clade_list <- lapply(unique(reid_cir_clade$clade), function(clade_name){
  reid_cir_clade$Geneid[reid_cir_clade$clade %in% clade_name]
})
names(pir_clade_list) <- as.character(unique(reid_cir_clade$clade))

#c can cat lists too!

total_tpm_expression(pir_clade_list, 
                     save = FALSE)
```

L4s and L1s appear to be particularly high in the rMT samples.

```{r total_tpm_pir_clades_SorL}
pir_SorL <- lapply(c('S','L'), function(clade_name){
  reid_cir_clade$Geneid[grepl(reid_cir_clade$clade, pattern = clade_name)]
})
names(pir_SorL) <- c('S','L')
total_tpm_expression(pir_SorL, save = FALSE)
```

#Expression of pirs from different chromosomes.

```{r total_tpm_pir_chrom}
pir_chrom_list <- lapply(unique(cir_chrom_tbl$chrom), 
                         function(chrom_name){
  cir_chrom_tbl$Geneid[cir_chrom_tbl$chrom %in% chrom_name]
})
names(pir_chrom_list) <- as.character(
  paste0(
    'chromosome ',
    unique(cir_chrom_tbl$chrom)
  )
)

total_tpm_expression(pir_chrom_list, save = FALSE)
```

Many big differences across the chromosomes. I wonder if this is consistent with Spence 2013 and Brugat 2017 data?

Early rings: Chr01/07/08/09/10 expressed higher in SBP. Chr05 expressed higher in SBP.

# Looking at the ETRAMP family which shows interesting expression patterns

```{r etramps}
etramp_info <- chabaudi_gene_info %>% 
  select(Geneid, Description, Transcript.Length) %>% 
  filter(Description == 'early transcribed membrane protein')
etramps <- unique(etramp_info$Geneid)

draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
                          genelist = etramps,
                          scale = 'max',
                          threshold = 0,
                          log_it = TRUE)
gene_graph_24h(genevec = etramps, genename = 'ETRAMPs')

twentyfour_ct_etramps <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% etramps) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
  mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
  mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
  #Convert the times given in 'h' to robustly coded times in 24h clock
  mutate(time = ifelse(grepl(time, pattern = '[[:digit:]]h$'), 
                       strptime(time, format="%Hh") %>% format(., "%Hh"),
                       time)) %>% 
  mutate(time = ifelse(grepl(time, pattern = '[[:digit:]](a|p)m'), 
                       strptime(time, format="%I%p") %>% format(., "%Hh"),
                       time)) %>% 
  group_by(sample, stage, transmission, time) %>% 
  dplyr::summarise(sum_tpm = sum(tpm))

p <- ggplot(data = twentyfour_ct_etramps, aes(x = time, 
                                          y = log1p(sum_tpm),
                                          fill = transmission)) +
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = cbPalette) +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
  theme_classic() + 
  ggtitle('24h chabaudi etramp expression')
p
```

When one ETRAMP goes down, does another replace it?

# Expression of HSP70

Often used as a housekeeping gene.

```{r hsp70}
total_tpm_expression(list_genes=list(hsp70 = 'PCHAS_0721000'),
                     save = TRUE)
```

# Calculate and cluster with z-scores {.tabset}

z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:

1. Convert the count/RPKM values of each gene into log values.
2. Calculate the mean and standard deviation of X gene log values in each sample
3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.

* Need to make sure that the z-score is calculated within each transmission method!

```{r zscores}
tpm_to_zscore = function(input){

  #Can use vars(matches(paste(input$experiment))) to get the columns with expression data
  # input_logged = input$tpm %>% 
  #   mutate_at(vars(matches(paste(input$experiment))), log1p)

  #If the input does not have a TPM element then assume that the input itself is the TPM object
  if(any(names(input) %in% 'tpm')){
    tpm_object <- input$tpm
  }else{
    tpm_object <- input
    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
  }

  #Want all of them so will just ignore Geneid
  input_logged = tpm_object %>%
    mutate_at(vars(-matches('Geneid')), 
              log1p)

  gene_mean_scores = input_logged %>% 
    melt(id.vars = 'Geneid') %>% 
    group_by(Geneid) %>% 
    dplyr::summarise(mean = mean(value),
                     std_dev = sd(value))

  gene_vector = as.character(input_logged$Geneid)

  input_z = lapply(gene_vector, function(gene){
    mean_gene = filter(gene_mean_scores, 
                       Geneid == paste(gene)) %>% 
      select(mean)
    std_gene = filter(gene_mean_scores, 
                      Geneid == paste(gene)) %>% 
      select(std_dev)
    gene_scores = input_logged %>% 
      filter(Geneid == paste(gene)) %>% 
      select(-Geneid)
    if(std_gene != 0){
      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
    }else{
      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
    }
    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
  })

  #Use reduce and rbind to turn the list into one df
  input_z_df = purrr::reduce(input_z, 
                             rbind)
  return(list(zscores = input_z_df,
              tpm = tpm_object,
              experiment = input$experiment))
}

#Make and save z-scores
# twentyfour_z <- lapply(twentyfour_ct, 
# tpm_to_zscore)

### Calculate and write zscores

# twentyfour_z <- lapply(twentyfour_avgtpm_noschz,
#                        tpm_to_zscore)
# 
# lapply(twentyfour_z, function(list_element){
#   z_df <- list_element$zscore
#   write_csv(x = z_df,
#             paste0('tl19-07_chabaudi24h_', list_element$experiment, '.csv'))
# })

###

### Read z-scores

twentyfour_z <- list(
  'rMT' = list(zscores = read_csv('tl19-07_chabaudi24h_rMT.csv'),
               tpm = twentyfour_avgtpm_noschz$rMT$tpm,
               experiment = 'rMT'),
  'SBP' = list(zscores = read_csv('tl19-07_chabaudi24h_SBP.csv'),
               tpm = twentyfour_avgtpm_noschz$SBP$tpm,
               experiment = 'SBP')
)

###
```

## Looking at all genes which have a z score over 2.4

```{r over2z}
over2_z_genes <- lapply(twentyfour_z, function(list_experiments){
  list_experiments$zscores %>%  
  filter_at(vars(-matches('Geneid')), 
            any_vars(. > 2.4)) %>% 
  .$Geneid
})

draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
                          genelist = unlist(over2_z_genes),
                          scale = 'max',
                          threshold = 0,
                          log_it = FALSE)
```

Hmm so the largest z scores aren't necessarily the most variable according to the correlation plot? Is this because they often only vary greatly in one sample?

# Phaseograms of the data. {.tabset}

```{r row_order_phaseograms}

#Getting the row order for each phaseogram for each experiment
# Want the data ordered by the highest zscore for each gene in each stage.
phaseogram_row_order <- lapply(twentyfour_z, 
                               function(list_elements){
                                 zscore_df <- list_elements$zscores
                                 zscore_ordered <- melt(zscore_df, 
                                                        id = 'Geneid',
                                                        variable.name = 'stage',
                                                        value.name = 'zscore') %>% 
                                   group_by(Geneid) %>% 
                                   top_n(zscore, n = 1) %>% 
                                   arrange(stage, 
                                           desc(zscore)) %>%
                                   distinct(Geneid, 
                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
                                   filter(zscore > 0) #For some reason some remain with zero zscore
                                 zscore_ordered <- zscore_ordered[order(
                                   match(
                                     zscore_ordered$stage, 
                                     order_vector
                                   )
                                 ),]
                                 return(c(list_elements, 
                                          list(zscore_order = zscore_ordered)))
                               })

phaseogram_time_order <- lapply(names(phaseogram_row_order), 
                                function(list_element){
                                  transmission_list <- getElement(phaseogram_row_order,
                                                                  list_element)
                                  order_by_time <- transmission_list$zscore_order %>% 
                                    mutate(stage = times_stages_table$time[match(stage,
                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
                                    arrange(stage)
                                  return(list(zscore_order = order_by_time))
                                }
)
names(phaseogram_time_order) <- names(phaseogram_row_order)

```

~~Change the lnTPM graph to a heatmap of max/min TPM a la Brugat 2017.~~

~~Add annotation for the clades,~~ or possibly separate the heatmap by clade.

Add threshold option.

```{r function_draw_heatmap_zscore}

draw_heatmap_zscore <- function(zscore_list,
                                title = '',
                                gene_ordering_vector = NULL,
                                filter_genes_vector = NULL,
                                extra_title = NULL,
                                annotation_clade = NULL,
                                annotation_col,
                                threshold_tpm = NULL,
                                use_times = FALSE,
                                use_order_vector = TRUE,
                                tpm_mat = TRUE,
                                tpm_breaks = NULL,
                                ...){
  if(is.null(tpm_breaks) & tpm_mat){
    #Need the generate the breaks in the data for the heatmap legend if not provided.
    max_tpm <- lapply(zscore_list, 
                      function(list_element){
                        #list_element <- zscore_list[[1]]
                        heatmap_data <- as.data.frame(list_element$zscores)
                        title <- list_element$experiment

                        if(!'Geneid' %in% colnames(heatmap_data)){
                          stop('No Geneid column in the heatmap data.frame')
                        }

                        if(!is.null(filter_genes_vector)){
                          heatmap_data <- as.data.frame(
                            filter(heatmap_data, 
                                   Geneid %in% filter_genes_vector)
                          )
                        }

                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                        if(!is.null(gene_ordering_vector)){
                          #Remove genes not in the main data frame but in the ordering vector, 
                          # and then order based on the vector
                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          heatmap_data <- heatmap_data[na.omit(
                            match(gene_ordering_vector, 
                                  rownames(heatmap_data))
                          ),,
                          drop=FALSE]
                        }

                        geneids <- rownames(heatmap_data)

                        #Get max/min TPM scores for the side annotation heatmap                       
                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                          list_element$tpm$Geneid),] %>% 
                          select(-Geneid) %>% 
                          as.matrix %>% 
                          log1p
                        rownames(tpm_mat) <- geneids

                        tpm_min <- rowMins(tpm_mat)
                        names(tpm_min) <- geneids
                        tpm_max <- rowMaxs(tpm_mat)
                        names(tpm_max) <- geneids
                        return(list(tpm_max = tpm_max,
                                    tpm_min = tpm_min))
                      })

    tpm_max_all <- mround(
      max(
        unlist(
          lapply(max_tpm, 
                 getElement,
                 'tpm_max')
        )
      ), 
      base = 1)

    tpm_breaks <- round(
      c(0,
        tpm_max_all/4,
        tpm_max_all/2,
        tpm_max_all*(3/4),
        tpm_max_all),
      digits = 2
    )
  }

  #Generate the colour function from the TPM legend breaks for the TPM annotation
  if(tpm_mat){
    tpm_col_fun <- colorRamp2(
      breaks = tpm_breaks,
      colors = viridis(length(tpm_breaks))
    )
  }

  if(is.numeric(threshold_tpm)){
    #Apply a threshold to the TPM in order to remove low expressed genes.
    # Remember to add log1p here! 
    # Threshold uses the max value, so it ony needs to be above thresh 
    #   in one stage avg in one transmission method.
    #Will alter the gene_filter vector in order to only select the genes above threshold

    gene_maxs <- bind_rows(
      lapply(max_tpm, 
             getElement,
             'tpm_max')
    )

    remove_genes <- names(gene_maxs)[apply(
      gene_maxs, 
      2, 
      function(maxs){
        all(maxs < log1p(threshold_tpm))
      }
    )]

    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]

  }

  if(is.character(annotation_clade)){
    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
    clade_info <- getElement(cir_clade_info,
                             annotation_clade)[match(order_cir, 
                                                     cir_clade_info$Geneid)]
    names(clade_info) <- str_split_fixed(order_cir, 
                                         pattern = '_', 
                                         n = 2)[,2]
    annotation_col <- annotation_col
  }

  heatmaps <- lapply(zscore_list, 
                     function(list_element){
                       #list_element <- zscore_list[[1]]
                       heatmap_data <- as.data.frame(list_element$zscores)
                       title <- list_element$experiment

                       if(!'Geneid' %in% colnames(heatmap_data)){
                         stop('No Geneid column in the heatmap data.frame')
                       }

                       if(!is.null(filter_genes_vector)){
                         heatmap_data <- as.data.frame(
                           filter(heatmap_data, 
                                  Geneid %in% filter_genes_vector)
                         )
                       }

                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                       if(!is.null(gene_ordering_vector)){
                         #Remove genes not in the main data frame but in the ordering vector,
                         #  and then order based on the vector
                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                         heatmap_data <- heatmap_data[na.omit(
                           match(gene_ordering_vector, 
                                 rownames(heatmap_data))
                         ),,
                         drop=FALSE]
                       }

                       if(tpm_mat){
                         #Get max/min TPM scores for the side annotation heatmap                       
                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                           list_element$tpm$Geneid),] %>%
                           select(-Geneid) %>% 
                           as.matrix %>% 
                           log1p

                         tpm_min <- rowMins(tpm_mat)
                         tpm_max <- rowMaxs(tpm_mat)

                         names(tpm_min) = rownames(heatmap_data)
                         names(tpm_max) = rownames(heatmap_data)

                         tpm_ha <- rowAnnotation(
                           min.max = anno_simple(cbind(tpm_min, 
                                                       tpm_max), 
                                                 col = tpm_col_fun,
                                                 gp = gpar(col = 'black',
                                                           lwd = 0.1),
                                                 border = TRUE,
                                                 simple_anno_size = unit(0.2, "cm")),
                           show_annotation_name = FALSE)
                       }

                       #Draw the left annotation is needed, only for the first element.

                       if(
                         identical(list_element,
                                   first(zscore_list)) & is.character(annotation_clade)
                       ){
                         clade_ha <- rowAnnotation(
                           'subfamily' = anno_simple(
                             clade_info,
                             col = annotation_col
                           )
                         )
                       }

                       #Change rownames to remove PCHAS_
                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
                                                                pattern = '_', 
                                                                n = 2)[,2]

                       if(use_times){
                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
                                                                                 times_stages_table$r_stage_diff_count)]
                       }

                       if(use_order_vector){
                         #Order the columns by ordering_vector
                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
                                                                   order_vector)),
                                                      drop = FALSE]
                       }else{
                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
                                                      drop = FALSE] 
                       }

                       h = Heatmap(heatmap_data,
                                   name = paste0('z-score',
                                                 extra_title),
                                   col = c(cbPalette[6],
                                           'white', 
                                           cbPalette[7]),
                                   row_names_gp = gpar(fontsize = ifelse(
                                     nrow(heatmap_data) > 13, 
                                     12*(13/nrow(heatmap_data)), 
                                     12)
                                   ),
                                   column_names_gp = gpar(fontsize = 10),
                                   row_dend_width = unit(3, "cm"),
                                   right_annotation = if(tpm_mat){
                                     tpm_ha
                                   }else{
                                     NULL
                                   },
                                   left_annotation = 
                                     if(identical(list_element,
                                                  first(zscore_list)) & is.character(annotation_clade)){
                                       clade_ha
                                     }else{
                                       NULL
                                     }, 
                                   #Only want the clade annotation when it is first heatmap
                                   width = unit(4.5, 'cm'),
                                   cluster_columns = FALSE,
                                   height = unit(7,'cm'),
                                   row_title = NULL,
                                   cluster_rows = FALSE,
                                   column_title = paste(title),
                                   heatmap_legend_param = list(
                                     at = c(3,2,1,0,-1,-2,-3),
                                     labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                   ),
                                   ...
                       )
                       return(h)
                     })

  ht_list <- Reduce('+', heatmaps) #To add the list elements together

  if(tpm_mat){
    #Draw the legend
    tpm_lgd <- Legend(
      title = "Min/Max log(TPM + 1)", 
      col_fun = tpm_col_fun, 
      at = tpm_breaks, 
      labels = tpm_breaks,
      title_gp = gpar(fontsize = 7)
    )
  }

  if(is.character(annotation_clade)){
    clade_lgd <- Legend(
      title = 'sub-family',
      at = names(annotation_col),
      legend_gp = gpar(fill = annotation_col)
    )
  }

  draw(ht_list, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       heatmap_legend_side = 'left',
       #Only include the annotation legends for the annotations we actually have.
       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
         list(tpm_lgd,
              clade_lgd)
       }else if(tpm_mat & !is.character(annotation_clade) ){
         list(tpm_lgd)
       } else if(!tpm_mat & is.character(annotation_clade) ){
         list(clade_lgd)
       })
}
```

## Phaseograms of the entire genome.

```{r phaseograms_whole_genome}
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)
```

## Phaseograms of the _cir_ genes with clade annotations.

```{r pir_gene_phaseogram_breaks}
  max_tpm <- lapply(twentyfour_z, 
                    function(list_element){
                      #list_element <- zscore_list[[1]]
                      heatmap_data <- as.data.frame(list_element$zscores)
                      title <- list_element$experiment

                      if(!'Geneid' %in% colnames(heatmap_data)){
                        stop('No Geneid column in the heatmap data.frame')
                      }

                            heatmap_data <- as.data.frame(
                              filter(heatmap_data, 
                                     Geneid %in% cir_id)
                            )

                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
                          # 
                          # if(!is.null(gene_ordering_vector)){
                          #   #Remove genes not in the main data frame but in the ordering vector, 
                          #   # and then order based on the vector
                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          #   heatmap_data <- heatmap_data[na.omit(
                          #     match(gene_ordering_vector, 
                          #           rownames(heatmap_data))
                          #   ),,
                          #   drop=FALSE]
                          # }

                          geneids <- rownames(heatmap_data)

                          #Get max/min TPM scores for the side annotation heatmap                       
                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                            list_element$tpm$Geneid),] %>% 
                            select(-Geneid) %>% 
                            as.matrix %>% 
                            log1p
                          rownames(tpm_mat) <- geneids

                          tpm_min <- rowMins(tpm_mat)
                          names(tpm_min) <- geneids
                          tpm_max <- rowMaxs(tpm_mat)
                          names(tpm_max) <- geneids
                          return(list(tpm_max = tpm_max,
                                      tpm_min = tpm_min))
                        })

  tpm_max_all <- mround(
    max(
      unlist(
        lapply(max_tpm, 
               getElement,
               'tpm_max')
      )
    ), 
    base = 1)

  tpm_breaks <- round(
    c(0,
      tpm_max_all/4,
      tpm_max_all/2,
      tpm_max_all*(3/4),
      tpm_max_all),
    digits = 2
  )
```

```{r phaseograms_cir_order_each}

#Generate the colour function from the TPM legend breaks for the TPM annotation

tpm_col_fun <- colorRamp2(
  breaks = tpm_breaks,
  colors = viridis(length(tpm_breaks))
)

#Apply a threshold to the TPM in order to remove low expressed genes.
# Remember to add log1p here! 
# Threshold uses the max value, so it ony needs to be above thresh 
#   in one stage avg in one transmission method.
#Will alter the gene_filter vector in order to only select the genes above threshold

gene_maxs <- bind_rows(
  lapply(max_tpm, 
         getElement,
         'tpm_max')
)

remove_genes <- names(gene_maxs)[apply(
  gene_maxs, 
  2, 
  function(maxs){
    all(maxs < log1p(1))
  }
)]

filter_genes_vector <- cir_id[!cir_id %in% remove_genes]

heatmaps <- lapply(names(twentyfour_z), 
                   function(list_name){
                     list_element <- getElement(twentyfour_z, 
                                                list_name)
                     #list_name <- names(twentyfour_z)[[1]]
                     heatmap_data <- as.data.frame(list_element$zscores)
                     title <- list_element$experiment

                     heatmap_data <- as.data.frame(
                       filter(heatmap_data, 
                              Geneid %in% filter_genes_vector)
                     )

                     rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                     heatmap_data <- heatmap_data[, 
                                                  !colnames(heatmap_data) %in% 'Geneid', 
                                                  drop = TRUE]
                     #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                     #Remove genes not in the main data frame but in the ordering vector,
                     #  and then order based on the vector
                     gene_ordering_vector <- getElement(phaseogram_time_order,
                                                        list_name)$zscore_order$Geneid

                     order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
                     clade_info <- getElement(cir_clade_info,
                                              'SorL')[match(order_cir, 
                                                            cir_clade_info$Geneid)]
                     names(clade_info) <- str_split_fixed(order_cir, 
                                                          pattern = '_', 
                                                          n = 2)[,2]
                     annotation_col <- SorL_col

                     remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                     heatmap_data <- heatmap_data[remove_index,
                                                  ,
                                                  drop = FALSE]
                     heatmap_data <- heatmap_data[na.omit(
                       match(gene_ordering_vector, 
                             rownames(heatmap_data))
                     ),
                     ,
                     drop=FALSE]

                     #Get max/min TPM scores for the side annotation heatmap                       
                     tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                       list_element$tpm$Geneid),] %>%
                       select(-Geneid) %>% 
                       as.matrix %>% 
                       log1p

                     tpm_min <- rowMins(tpm_mat)
                     tpm_max <- rowMaxs(tpm_mat)

                     names(tpm_min) = rownames(heatmap_data)
                     names(tpm_max) = rownames(heatmap_data)

                     tpm_ha <- rowAnnotation(
                       min.max = anno_simple(cbind(tpm_min, 
                                                   tpm_max), 
                                             col = tpm_col_fun,
                                             gp = gpar(col = 'black',
                                                       lwd = 0.1),
                                             border = TRUE,
                                             simple_anno_size = unit(0.2, "cm")),
                       show_annotation_name = FALSE)

                     #Clade annotation

                     clade_ha <- rowAnnotation(
                       'sub-family' = anno_simple(
                         clade_info,
                         col = annotation_col,
                       ),
                       annotation_name_gp = gpar(fontsize = 10)
                     )

                     #Change rownames to remove PCHAS_
                     rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
                                                              pattern = '_', 
                                                              n = 2)[,2]

                     colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
                                                                             times_stages_table$r_stage_diff_count)]

                     heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
                                                  drop = FALSE]

                     h = Heatmap(heatmap_data,
                                 name = paste0('z-score'),
                                 col = c(cbPalette[6],
                                         'white', 
                                         cbPalette[7]),
                                 row_names_gp = gpar(fontsize = ifelse(
                                   nrow(heatmap_data) > 13, 
                                   12*(13/nrow(heatmap_data)), 
                                   12)
                                 ),
                                 column_names_gp = gpar(fontsize = 10),
                                 row_dend_width = unit(3, "cm"),
                                 right_annotation = tpm_ha,
                                 left_annotation = clade_ha,
                                 width = unit(4.5, 'cm'),
                                 cluster_columns = FALSE,
                                 height = unit(7,'cm'),
                                 row_title = NULL,
                                 cluster_rows = FALSE,
                                 column_title = paste(title),
                                 heatmap_legend_param = list(
                                   at = c(3,2,1,0,-1,-2,-3),
                                   labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                 ),
                                 show_row_names = FALSE
                     )
                     return(h)
                   })

ht_list <- Reduce('+', heatmaps) #To add the list elements together

#Draw the legend
tpm_lgd <- Legend(
  title = "Min/Max log(TPM + 1)", 
  col_fun = tpm_col_fun,
  at = tpm_breaks, 
  labels = tpm_breaks,
  title_gp = gpar(fontsize = 7)
)

clade_lgd <- Legend(
  title = 'sub-family',
  at = names(SorL_col),
  legend_gp = gpar(fill = SorL_col)
)

png(filename = 'tl19-07_phaseogram_each-in-order_cir_clades.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 200)

  draw(ht_list, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       heatmap_legend_side = 'left',
       annotation_legend_list = 
         list(tpm_lgd,
              clade_lgd),
       ht_gap = unit(.5,'cm')
       )

  dev.off()

```

```{r phaseograms_cir}

phase_rMT <- draw_heatmap_zscore(zscore_list = list(twentyfour_z$rMT) ,
                    gene_ordering_vector = phaseogram_row_order$rMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_breaks = tpm_breaks)

phase_SBP <- draw_heatmap_zscore(zscore_list = list(twentyfour_z$SBP),
                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_breaks = tpm_breaks)

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_row_order$SBP$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'clade',
                    annotation_col = clade_col,
                    threshold_tpm = 200,
                    use_times = TRUE,
                    use_order_vector = FALSE)

png(filename = 'tl19-07_phaseogram_rMTorder_cir_clades.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 200)
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'clade',
                    annotation_col = clade_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    tpm_mat = FALSE)
dev.off()

png(filename = 'tl19-07_phaseogram_rMTorder_cir_SorL.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 200)
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)
dev.off()

png(filename = 'tl19-07_phaseogram_SBPorder_cir_SorL.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 200)
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_breaks = tpm_breaks,
                    tpm_mat = FALSE)
dev.off()

```

## Phaseograms by clade.

```{r phaseograms_by_clade}
cir_clades <- unique(cir_clade_info$clade) %>% sort
lapply(cir_clades, 
       function(cir_clade_name){
         cirs_in_clade <- cir_clade_info$Geneid[cir_clade_info$clade %in% cir_clade_name]
         if(length(cirs_in_clade) > 1){
           draw_heatmap_zscore(zscore_list = twentyfour_z,
                               gene_ordering_vector = phaseogram_row_order$rMT$zscore_order$Geneid,
                               filter_genes_vector = cirs_in_clade,
                               extra_title = paste0('_', cir_clade_name),
                               annotation_clade = TRUE)
           if(global_save_var){
             png(filename = paste0('tl19-07_phaseogram_rMTorder_cir_', 
                                   cir_clade_name,
                                   '.png'),
                 units = 'cm',
                 height = 20,
                 width = 22,
                 res = 200)
             draw_heatmap_zscore(zscore_list = twentyfour_z,
                                 gene_ordering_vector = phaseogram_row_order$rMT$zscore_order$Geneid,
                                 filter_genes_vector = cirs_in_clade,
                                 extra_title = paste0('_', cir_clade_name))
             dev.off()
           }

         }
       })
```

Putting them side by side you can see the delay with the whole transcriptome - genes upregulated in certain rMT stages appear to be upregulated in both the same stage and the next stage 3h later.

# Heatmap of all of the expressed pir genes in each stage, ordered by each transmission type. - Figure 3b

```{r highest_expressed_stage_heatmap_fig3b}

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm_noschz), 
       function(list_name){

         combined_data_pir <- getElement(twentyfour_avgtpm_noschz, 
                                     list_name)$tpm %>% 
           filter(Geneid %in% cir_id) %>% 
           column_to_rownames(var = 'Geneid')

         #Ordering that a more understandable order

         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                            order_vector)),
                                                drop = FALSE]

         # combined_data_ordered <- melt(
         #   (
         #     combined_data_pir %>% 
         #       rownames_to_column(var = 'Geneid')
         #   ),
         #   id = 'Geneid',
         #   variable.name = 'stage',
         #   value.name = 'tpm') %>% 
         #   group_by(Geneid) %>% 
         #   top_n(tpm, 
         #         n = 1) %>% 
         #   arrange(stage,
         #           desc(tpm)) %>%
         #   distinct(Geneid, 
         #            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
         #   filter(tpm > 0)

         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                         means = apply(combined_data_pir, 1, mean),
                                         subfam = cir_clade_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
                                                                            cir_clade_info$Geneid)]) %>% 
           filter(means > 1) %>% 
           arrange(subfam, desc(means)) %>% 
           .$Geneid

         combined_data_pir_ht <- combined_data_pir[
           order(
             match(rownames(combined_data_pir), 
                   combined_data_ordered),
             na.last = NA #Remove NAs
           ),
         ]

         return(
           list(order = combined_data_pir_ht)
         )
       })
names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm_noschz)

lapply(names(combined_data_pir_ht_list),
       function(list_name_to_order){

         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
                                            list_name_to_order)$order

         pir_highest_tpm_order <- rownames(combined_data_pir_ht)

         maxcol <- ceiling(
           max(log1p(combined_data_pir_ht)
           )
         )

         #Sub-family annotation
         clade_info <- getElement(cir_clade_info,
                                  'SorL')[match(pir_highest_tpm_order, 
                                                cir_clade_info$Geneid)]
         annotation_col <- SorL_col

         #Subfamily annotation

         clade_ha <- rowAnnotation(
           'sub-family' = anno_simple(
             clade_info,
             col = annotation_col,
           ),
           annotation_name_gp = gpar(fontsize = 10)
         )

         anc_ha <- rowAnnotation(foo = anno_mark(at = 1, 
                                                 labels = 'ancestral',
                                                 labels_gp = gpar(fontsize = 7),
                                                 link_width = unit(3, "mm")))

         #lapply across the datasets now, for each order
         ht_list <- lapply(names(combined_data_pir_ht_list),
                           function(list_name){

                             combined_data_pir_ht <- getElement(twentyfour_avgtpm_noschz, 
                                                                list_name)$tpm %>% 
                               column_to_rownames(var = 'Geneid')

                             #Get the times instead of the stage names.
                             colnames(combined_data_pir_ht) <- times_stages_table$time[match(colnames(combined_data_pir_ht),

                                                                                             times_stages_table$r_stage_diff_count)]

                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                                          drop = FALSE]

                             Heatmap(
                               log1p(combined_data_pir_ht)[pir_highest_tpm_order,],
                               cluster_rows = FALSE,
                               cluster_columns = FALSE,
                               show_row_names = FALSE,
                               name = 'log(TPM+1)', 
                               col = colorRamp2(c(0, 
                                                  maxcol/4, 
                                                  maxcol/2, 
                                                  3*maxcol/4, 
                                                  maxcol), 
                                                viridis(5)),
                               na_col = 'white', 
                               border = TRUE,
                               border_gp = gpar(lwd = 0.1),
                               column_names_gp = gpar(fontsize = 10),
                               row_dend_width = unit(3, "cm"),
                               width = unit(4.5, 'cm'),
                               height = unit(7,'cm'),
                               row_title = NULL,
                               left_annotation = if(
                                 list_name %in% first(names(combined_data_pir_ht_list))
                               ){
                                 clade_ha
                               }else{
                                 NULL
                               },
                               right_annotation = if(
                                 list_name %in% last(names(combined_data_pir_ht_list))
                               ){
                                 anc_ha
                               }else{
                                 NULL
                               },
                               # row_split =  cir_clade_info$SorL[match(pir_highest_tpm_order, 
                               #                                        cir_clade_info$Geneid)],
                               column_title = paste(list_name),
                               heatmap_legend_param = list(
                                 labels_gp = gpar(fontsize = 10),
                                 title_gp = gpar(fontsize = 10,
                                                 fontface = 'bold')
                               )
                             )
                           }
         )
         ht_list <- Reduce(`+`, ht_list)
         clade_lgd <- Legend(
           title = 'sub-family',
           at = names(SorL_col),
           legend_gp = gpar(fill = SorL_col)
         )

         png(filename = paste0('tl19-07_combined_pirs_ordered_highest_stage_expression_heatmap_TPM_',
                               list_name_to_order, 'order',
                               '.png'),
             width = 20,
             height = 10,
             units = 'cm',
             res = 330)

         draw(ht_list, 
              row_title_side = 'left',
              column_title_side = 'bottom',
              heatmap_legend_side = 'right',
              annotation_legend_side = 'right',
              annotation_legend_list = list(clade_lgd),
              ht_gap = unit(.35,'cm'),
              merge_legend = TRUE
         )

         dev.off()
       })

```

```{r highest_genes_stage}
highest_pirs_stage_transmission <- lapply(names(twentyfour_avgtpm_noschz), 
       function(list_name){

         combined_data_pir <- getElement(twentyfour_avgtpm_noschz, 
                                     list_name)$tpm %>% 
           filter(Geneid %in% cir_id) %>% 
           column_to_rownames(var = 'Geneid')

         #Ordering that a more understandable order

         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                            order_vector)),
                                                drop = FALSE]

         combined_data_ordered <- melt(
           (
             combined_data_pir %>%
               rownames_to_column(var = 'Geneid')
           ),
           id = 'Geneid',
           variable.name = 'stage',
           value.name = 'tpm') %>%
           group_by(Geneid) %>%
           top_n(tpm,
                 n = 1) %>%
           arrange(stage,
                   desc(tpm)) %>%
           distinct(Geneid,
                    .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
           filter(tpm > 0)

         highest_pir_stage_list <- lapply(unique(combined_data_ordered$stage), 
                                          function(stage_name){
                                            filter(combined_data_ordered, 
                                                   stage %in% stage_name)$Geneid
                                          }
         )
         names(highest_pir_stage_list) <- unique(combined_data_ordered$stage)
         return(highest_pir_stage_list)
       })
names(highest_pirs_stage_transmission) <- names(twentyfour_avgtpm_noschz)
```

```{r different_highest_genes_between_transmission}
stage_names <- ordering(colnames(twentyfour_avgtpm_noschz$rMT$tpm)[-1],
                        order_vector)
agree_highest_pirs <- lapply(stage_names, 
       function(stage_name){
  agree(
    lapply(highest_pirs_stage_transmission, 
           getElement, 
           stage_name)
  )
})
names(agree_highest_pirs) <- stage_names

disagree_highest_pirs_rMThigher <- lapply(stage_names, 
       function(stage_name){
  disagree(
    lapply(highest_pirs_stage_transmission, 
           getElement, 
           stage_name)
  )
})
names(disagree_highest_pirs_rMThigher) <- stage_names

disagree_highest_pirs_SBPhigher <- lapply(stage_names, 
       function(stage_name){
  disagree(
    lapply(rev(highest_pirs_stage_transmission), 
           getElement, 
           stage_name)
  )
})
names(disagree_highest_pirs_SBPhigher) <- stage_names
```

# pir expression by clade/sub-family

```{r highest_genes_clade, results = "asis"}
lapply(names(highest_pirs_stage_transmission), 
       function(transmission_name){
         transmission_highest_pirs <- getElement(highest_pirs_stage_transmission,
                                                 transmission_name)
         lapply(names(transmission_highest_pirs), 
                function(top_pirs_name){
                  top_pirs <- getElement(transmission_highest_pirs, 
                                         top_pirs_name)
                  mutate(cir_clade_info, 
                         in_top = str_detect(Geneid, 
                                             pattern = paste(top_pirs, collapse = '|')) ) %>% 
                    group_by(clade) %>% 
                    dplyr::summarise(num = sum(in_top),
                                     prop_of_stage_top_pirs = round(sum(in_top)/length(top_pirs),
                                                                    digits = 3)*100,
                                     prop_of_clade = round(sum(in_top)/length(Geneid),
                                                           digits = 3)*100) %>% 
                    kable(format = 'html', 
                          caption = paste0(transmission_name, 
                                           '_', 
                                           top_pirs_name), 
                          escape = FALSE) %>% 
                    kable_styling()
                })
       })
```

Unknowns always highest in the ring stages.

## Proportion table of S/L clade expression by stage.

```{r proportion_data}

tpm_data_twentyfour <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_clade_info$clade[match(Geneid, 
                                            cir_clade_info$Geneid)],
         chrom = str_extract(Geneid, 
                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
         ) %>% 
  mutate(
    SorL = str_extract(clade, 
                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
    # stage =  factor(stage, 
    #                 levels = ordering(stage, order_vector))
  ) %>% 
  group_by(transmission,
           stage, 
           Geneid,
           clade,
           chrom,
           SorL) %>% 
  dplyr::summarise(tpm = mean(tpm))

```

```{r clade_proportion}
tpm_data_twentyfour %>% 
    group_by(transmission,
           stage, 
           clade) %>% 
  dplyr::summarise(max_gene = max(tpm),
                   total_clade = sum(tpm)) %>%
  group_by(stage) %>% 
  dplyr::mutate(total_stage = sum(total_clade)) %>% 
  dplyr::mutate(
    prop = round(total_clade/total_stage, digits = 2),
    prop_from_max_gene = ifelse(total_clade > 0,
                                round(max_gene/total_clade, digits = 2),
                                0)
  ) %>% 
  # Prop from max gene to show how much of the clade expression is from the highest expressed gene of the clade.
  filter(prop > 0.0001) %>%
  dplyr::arrange(transmission, desc(prop)) %>% 
  select(-max_gene) %>% 
  kable(format = 'html', 
        caption = 'Proportion of pir clade expression for each stage', 
        escape = FALSE) %>% 
  kable_styling() %>% 
  scroll_box(width = "100%", height = "500px")

```

## Stacked bar chart of clade expression.

```{r clades_barchart}
tpm_data_clade <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_clade_info$clade[match(Geneid, 
                                            cir_clade_info$Geneid)]) %>% 
  # mutate(stage =  factor(stage, 
  #                        levels = ordering(stage, 
  #                                          order_vector))) %>% 
  group_by(transmission,
           stage, 
           clade) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  arrange(stage)

y <- ggplot(data = tpm_data_clade, 
              aes(
                x = transmission,
                y = total_tpm, 
                fill = clade,
                group = transmission
              )) +
    geom_bar(position = position_stack(),
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
    scale_fill_manual(
      breaks = unique(tpm_data_clade$clade),
      values = setNames(
        c(cbPalette, 
          'red', 
          'white')[1:length(unique(tpm_data_clade$clade))],
        unique(tpm_data_clade$clade)
      )
    ) +
    scale_color_manual('black') + 
    # labs(fill = stack_label) +
    # scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~stage, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_subclades_totalTPM_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

## Clades proportion bar chart

```{r clades_prop_barchart}
y <- ggplot(data = tpm_data_clade, 
              aes(
                x = stage,
                y = total_tpm, 
                fill = clade,
                group = transmission
              )) +
    geom_bar(position = 'fill',
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
    scale_fill_manual(
      breaks = sort(unique(tpm_data_clade$clade)),
      values = clade_col,
      guide = guide_legend(reverse = TRUE) 
    ) +
    scale_color_manual('black') + 
    # labs(fill = stack_label) +
    # scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('proportion of total '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~transmission, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_text(size = 7, 
                                 angle = 90),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_subclades_propTPM_barchart.png',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

## S/L clade member proportions among the pirs expressed highest at each stage.

```{r highest_pirs_SorL, results = "asis"}
lapply(names(highest_pirs_stage_transmission), 
       function(transmission_name){
         transmission_highest_pirs <- getElement(highest_pirs_stage_transmission,
                                                 transmission_name)
         lapply(names(transmission_highest_pirs), 
                function(top_pirs_name){
                  top_pirs <- getElement(transmission_highest_pirs, 
                                         top_pirs_name)
                  mutate(cir_clade_info, 
                         clade = str_extract(clade, pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
                         in_top = str_detect(Geneid, 
                                             pattern = paste(top_pirs, collapse = '|')) ) %>% 
                    group_by(clade) %>% 
                    dplyr::summarise(num = sum(in_top),
                                     prop_of_stage_top_pirs = round(sum(in_top)/length(top_pirs),
                                                                    digits = 3)*100,
                                     prop_of_clade = round(sum(in_top)/length(Geneid),
                                                           digits = 3)*100) %>% 
                    kable(format = 'html', 
                          caption = paste0(transmission_name, 
                                           '_', 
                                           top_pirs_name), 
                          escape = FALSE) %>% 
                    kable_styling()
                })
       })
```

SBP Late Trophs have highest expression of more Ls than the rMT late trophs. This is probably because the expression of the pirs is generally more spread out in the rMT, does this culminate as them having more genes in different phases in the Fourier results?

All the unknown clades are expressed in the earlier ring stages, however I don't think they themselves are expressed much at all.

## Bar chart of S and L clade expression across time.

```{r SorL_barchart}
tpm_data_SorL <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_clade_info$clade[match(Geneid, 
                                            cir_clade_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm))

y <- ggplot(data = tpm_data_SorL, 
              aes(
                x = transmission,
                y = total_tpm, 
                fill = SorL,
                group = transmission
              )) +
    geom_bar(position = position_stack(),
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
    scale_fill_manual(
      values = SorL_col,
      guide = guide_legend(reverse = TRUE)
    ) +
    scale_color_manual('black') + 
    # labs(fill = stack_label) +
    # scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~stage, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_clades_totalTPM_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

```{r SorL_barchart_transmission-facet}
tpm_data_SorL <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_clade_info$clade[match(Geneid, 
                                            cir_clade_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  mutate(SorL = factor(SorL, 
                       levels = sort(unique(SorL))))

y <- ggplot(data = tpm_data_SorL, 
              aes(
                x = stage,
                y = total_tpm, 
                fill = SorL,
                # group = transmission
              )) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) +
  scale_fill_manual(
    values = SorL_col,
    guide = guide_legend(reverse = FALSE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~transmission, 
             scales = "free_x",
             space = "free_x") +
  labs(fill = 'sub-family') +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_SorL_totalTPM_facet-transmission_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

## S or L proportion of expression bar chart

```{r add_berghei_SorL_prop}
#Add berghei data
# berghei_SorL_file <- list.files('..', 
#                                 recursive = TRUE, 
#                                 pattern = 'tl20-05_SorL-subfamily_pir.csv', 
#                                 full.names = TRUE)
berghei_SorL <- read_csv('../tl20-05_bergheisystematicanalysis_nextflowoutput/tl20-05_SorL-subfamily_pir.csv') %>% 
  filter(stage %in% c('Asexual.SchizontA',
                      'Asexual.SchizontB',
                      'Asexual.Ring',
                      'Asexual.Trophozoite',
                      'Asexual.Mixed')) 
berghei_SorL_avg <- berghei_SorL %>% 
  group_by(stage,
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
  mutate(SorL = str_extract(SorL,
                            pattern = 'S|L|ancestral'),
         transmission = 'berghei')

```

```{r}
y <- ggplot(data = rbind(
  tpm_data_SorL,
  berghei_SorL_avg
), 
aes(
  x = stage,
                y = total_tpm, 
                fill = SorL,
                # group = transmission
              )) +
  geom_bar(position = 'fill',
           stat = 'identity',
           colour = 'black',
           size = 0.1) +
  scale_fill_manual(
    values = SorL_col,
    guide = guide_legend(reverse = FALSE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('proportion of total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~transmission, 
             scales = "free_x",
             space = "free_x") +
  labs(fill = 'sub-family') +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_text(angle = 90),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_SorL_totalTPM_facet-transmission_prop_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')
```

## S or L proportion of expression.

```{r SorL_proportion_table}
tpm_data_twentyfour %>% 
    group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(max_gene = max(tpm),
                   total_clade = sum(tpm)) %>%
  group_by(stage) %>% 
  dplyr::mutate(total_stage = sum(total_clade)) %>% 
  dplyr::mutate(
    prop = round(total_clade/total_stage, digits = 2),
    prop_from_max_gene = ifelse(total_clade > 0,
                                round(max_gene/total_clade, digits = 2),
                                0)
  ) %>% 
  # Prop from max gene to show how much of the clade expression is from the highest expressed gene of the clade.
  filter(prop > 0.0001) %>%
  dplyr::arrange(transmission, stage, desc(prop)) %>% 
  select(-max_gene) %>% 
  kable(format = 'html', 
        caption = 'Proportion of pir clade expression for each stage', 
        escape = FALSE) %>% 
  kable_styling() %>% 
  scroll_box(width = "100%", height = "500px")

```

# Expression by chromosome

```{r chrom_proportion}
tpm_data_twentyfour %>% 
      group_by(transmission,
           stage, 
           chrom) %>% 
  dplyr::summarise(max_gene = max(tpm),
                   total_chrom = sum(tpm)) %>%
  group_by(stage) %>% 
  dplyr::mutate(total_stage = sum(total_chrom)) %>% 
  dplyr::mutate(
    prop = round(total_chrom/total_stage, digits = 2),
    prop_from_max_gene = ifelse(total_chrom > 0,
                                round(max_gene/total_chrom, digits = 2),
                                0)
  ) %>% 
  # Prop from max gene to show how much of the chrom expression is from the highest expressed gene of the chrom.
  filter(prop > 0.0001) %>%
  dplyr::arrange(transmission, desc(prop)) %>% 
  select(-max_gene) %>% 
  kable(format = 'html', 
        caption = 'Proportion of pir chrom expression for each stage', 
        escape = FALSE) %>% 
  kable_styling() %>% 
  scroll_box(width = "100%", height = "500px")
```

```{r chrom_total_bar_chart}
tpm_data_chrom <- tpm_data_twentyfour %>% 
  mutate(stage = times_stages_table$time[match(stage, 
                                               times_stages_table$r_stage_diff_count)]) %>% 
  group_by(transmission,
           stage, 
           chrom) %>% 
  dplyr::summarise(total_tpm = sum(tpm))

y <- ggplot(data = tpm_data_chrom, 
              aes(
                x = transmission,
                y = total_tpm, 
                fill = chrom,
                group = transmission
              )) +
    geom_bar(position = position_stack(),
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
  scale_fill_manual(
    breaks = unique(tpm_data_chrom$chrom),
    values = setNames(
      c(
        plasma(length(unique(tpm_data_chrom$chrom)))[c(TRUE, NA)]
      ),
      unique(tpm_data_chrom$chrom)
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

# ggsave2('tl20-05_chroms_TPM_barchart.pdf',
#         plot = y,
#         dpi = 300,
#         width = 15,
#         height = 10,
#         units = 'cm')
#   
```

```{r chrom_prop_bar_chart}

y <- ggplot(data = tpm_data_chrom) +
  geom_bar_pattern(
    aes(
      x = stage,
      y = total_tpm, 
      fill = chrom,
      group = transmission,
      pattern = chrom
    ),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1,
    pattern_density = 0.05,
    pattern_spacing = 0.02,
    pattern_key_scale_factor = 0.5
  ) +
  scale_pattern_manual(
    values = rep(c('none', 
                   'circle'), 
                 each = length(c(cbPalette, 
                                 'red', 
                                 'white')), 
                 length.out = length(unique(tpm_data_chrom$chrom))),
    labels = unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_fill_manual(
    breaks = unique(tpm_data_chrom$chrom),
    values = rep_len(c(cbPalette, 
                       'red', 
                       'white'),
                     length.out = length(unique(tpm_data_chrom$chrom))),
    labels =  unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~transmission, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_text(size = 10, 
                                 angle = 0),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=10, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_chroms_propTPM_barchart.png',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

Clearly a massive upregulation of chr6 in rMT.

```{r chrom_6_L4s}
cir_clade_info %>% 
  mutate(
    chrom = str_extract(Geneid, 
                        '(?<=PCHAS_)[[:digit:]]{2}')
  ) %>% 
  dplyr::group_by(clade, 
                  chrom) %>% 
  dplyr::summarise(num = n()) %>% 
  dplyr::mutate(freq = num/sum(num)*100) %>% 
  arrange(clade,
          desc(freq))

cir_clade_info %>% 
  mutate(
    chrom = str_extract(Geneid, 
                        '(?<=PCHAS_)[[:digit:]]{2}')
  ) %>% 
  dplyr::group_by(chrom, 
                  clade) %>% 
  dplyr::summarise(num = n()) %>% 
  dplyr::mutate(freq = num/sum(num)*100) %>% 
  arrange(chrom,
          desc(freq))
```

L4 _pir_\s mostly on chr3 and chr6 (20.6% of the L4s each). 17% of L1s are on chr6.

L1 pirs make up 58.3% of chr6, and L4 pirs make up another 29.2%, only three other pirs are left, one unknown and two S7s.

## Bar chart of chromosome expression

```{r chrom_expression_bar_charts}
lapply(
    unique(tpm_data_twentyfour$chrom), 
  function(chrom_cirs){

         tpm_df <- filter(tpm_data_twentyfour,
                          chrom %in% chrom_cirs)

         #Get the order

         tpm_df <- mutate(tpm_df, 
                          stage = factor(stage,
                                         levels = ordering(stage,
                                                           order_vector)))

         maximum_tpm_value <- group_by(tpm_df, 
                                       stage, 
                                       transmission) %>% 
           dplyr::summarise(sum_tpm = sum(tpm)) %>% 
           .$sum_tpm %>% 
           max

         max_tpm <- as.integer(
           mround(
             seq(0, 
                 maximum_tpm_value, 
                 length.out = 5),
             base = 5
           )
         )

         y <- ggplot(data = tpm_df, 
                     aes(
                       x = transmission,
                       y = tpm, 
                       fill = Geneid
                     )) +
           geom_bar(position = position_stack(),
                    stat = 'identity',
                    colour = 'black',
                    size = 0.1) +
           scale_fill_manual(
             breaks = unique(tpm_df$Geneid),
             values = setNames(
               rep_len(
                 'white', 
                 length.out = length(unique(tpm_df$Geneid))
               ),
               unique(tpm_df$Geneid)
             )
           )+
           scale_color_manual('black') + 
           # labs(fill = stack_label) +
           scale_y_continuous(breaks = max_tpm) +
           theme_classic() +
           scale_size(range = c(0,3)) +
           ylab(
             bquote('total chr'~italic(.(chrom_cirs))~italic(.('pir'))~' TPM')
           ) + 
           facet_grid(.~stage, 
                      scales = "free_x",
                      space = "free_x") +
           theme(
             axis.title.x=element_blank(),
             axis.text.x=element_text(size = 8),
             axis.text.y = element_text(size = 5),
             axis.title.y = element_text(size = 8),
             legend.title = element_text(size = 7),
             legend.text = element_text(size = 5),
             legend.key.size = unit(3,'mm'),
             legend.key.width = unit(2,'mm'),
             strip.text.x = element_text(size=4, 
                                         face = 'bold', 
                                         angle = 90),
             strip.background = element_rect(
               color="black", 
               size=0.1, 
               linetype="solid"
             )
           )
         if(global_save_var){
             ggsave2(paste0('tl19-07_chr',
                            gsub(chrom_cirs, 
                                 pattern  = ' ', 
                                 replacement = '_'),
                            '_totalTPM_barchart.pdf'),
                     plot = y,
                     dpi = 300,
                     width = 15,
                     height = 10,
                     units = 'cm')
         }
         y
       })
```

# Expression by locus

```{r defining_loci}
cir_loci <- data.frame(
  Geneid = cir_id,
  chrom = str_extract(cir_id, 
                      pattern = '(?<=_)[[:digit:]]{2}'),
  number = as.numeric(
    str_extract(cir_id, 
                pattern = '(?<=_[[:digit:]]{2})[[:digit:]]+')
  )
) %>% 
  filter(!chrom %in% '00') %>% 
  #Remove chrom00 (contigs)
  group_by(chrom) %>% 
  mutate(jump = ifelse(lag(number, 
                           default = 1001) < 3000 & number - lag(number, 
                                                                 default = 0) > 1000,
                       1,
                       0)) %>% 
  mutate(
    jump = ifelse(cumsum(jump) == 1 | jump == 1, 1, 0)
  ) %>% 
  mutate(
    locus = paste0(
      chrom,
      ifelse(jump == 0 & number < 3000, 
             'Left',
             'Right')
    )
  )

# Detect 'jumps' of the gene number over 1000 within each chromosome, setting the lag default as 1001 in the
#   first part of the ifelse statement so that chroms don't have an erroneous jump if they start at high numbers
#   >1000
# Later cumsum is do that every gene after the 'jump' is included allowing us to define telomere ends.

cir_loci_list <- lapply(
  unique(cir_loci$locus),
  function(locus_name){
    cir_loci$Geneid[cir_loci$locus == locus_name]
  })

names(cir_loci_list) <- unique(cir_loci$locus)
```

```{r loci_numbers}
cir_loci %>% 
  group_by(locus) %>% 
  dplyr::summarise(num = n()) %>% 
  arrange(desc(num))
```

## Loci expression stacked bar graph

```{r loci_total_bar_chart}

tpm_data_loci <- filter(tpm_data_twentyfour,
                                       Geneid %in% cir_id) %>%
  mutate(locus = cir_loci$locus[match(Geneid, 
                                      cir_loci$Geneid)],
         stage = times_stages_table$time[match(stage, 
                                               times_stages_table$r_stage_diff_count)]) %>% 
  ungroup() %>% 
  group_by(stage, 
           transmission,
           locus) %>% 
  dplyr::summarise(max_gene = max(tpm),
                   total_tpm = sum(tpm))

y <- ggplot(data = tpm_data_loci, 
              aes(
                x = transmission,
                y = total_tpm, 
                fill = locus,
                group = transmission
              )) +
    geom_bar(position = position_stack(),
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
  scale_fill_manual(
    breaks = unique(tpm_data_loci$locus),
    values = setNames(
      c(
        plasma(length(unique(tpm_data_loci$locus)))[
          ifelse(str_detect(unique(tpm_data_loci$locus),
                            pattern = 'Left'),
                 TRUE,
                 NA)
        ]
      ),
      unique(tpm_data_loci$locus)
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

# ggsave2('tl20-05_loci_TPM_barchart.pdf',
#         plot = y,
#         dpi = 300,
#         width = 15,
#         height = 10,
#         units = 'cm')
#   
```

## Expression by locus proportions

```{r locus_prop}

tpm_data_loci %>%
  group_by(stage, 
           transmission) %>% 
  dplyr::mutate(total_stage = sum(total_tpm)) %>% 
  dplyr::mutate(
    prop = round(total_tpm/total_stage, digits = 2),
    prop_from_max_gene = ifelse(total_tpm > 0,
                                round(max_gene/total_tpm, digits = 2),
                                0)
  ) %>% 
  # Prop from max gene to show how much of the locus expression is from the highest expressed gene of the locus.
  filter(prop > 0.0001) %>%
  dplyr::arrange(desc(prop)) %>% 
  select(-max_gene) %>% 
  kable(format = 'html', 
        caption = 'Proportion of pir loci expression for each stage', 
        escape = FALSE) %>% 
  kable_styling() %>% 
  scroll_box(width = "100%", height = "500px")
```

* 34% of early and mid rMT troph comes from 06Left - 27% and 33% contribution from one gene respectively. Also 28% of late rings (17% from one gene). So spread out over multiple pir genes. Actually it is a majorly expressed locus it seems in rMT. 06Left has 14 genes, the most of any locus bar 03Left (18 genes).

```{r}
y <- ggplot(data = tpm_data_loci, 
              aes(
                x = transmission,
                y = total_tpm, 
                fill = locus,
                group = transmission
              )) +
    geom_bar(position = 'fill',
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
  scale_fill_manual(
    breaks = unique(tpm_data_loci$locus),
    values = setNames(
      c(
        rep(cbPalette, 
            length.out = length(unique(tpm_data_loci$locus)),
            each = 2)[
          ifelse(str_detect(unique(tpm_data_loci$locus),
                            pattern = 'Left'),
                 TRUE,
                 NA)
        ]
      ),
      unique(tpm_data_loci$locus)
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

# ggsave2('tl20-05_loci_propTPM_barchart.pdf',
#         plot = y,
#         dpi = 300,
#         width = 15,
#         height = 10,
#         units = 'cm')
#   
```

## Expression by locus over time

```{r loci_expression}

lapply(names(cir_loci_list), 
       function(loci_name){

         cir_loci_genes <- getElement(cir_loci_list, 
                                      loci_name)

         tpm_df <- filter(tpm_data_twentyfour,
                          Geneid %in% cir_loci_genes)

         #Get the order

         tpm_df <- mutate(tpm_df, 
                          stage = factor(stage,
                                         levels = ordering(stage,
                                                           order_vector)))

         maximum_tpm_value <- group_by(tpm_df, 
                                       stage, 
                                       transmission) %>% 
           dplyr::summarise(sum_tpm = sum(tpm)) %>% 
           .$sum_tpm %>% 
           max

         max_tpm <- as.integer(
           mround(
             seq(0, 
                 maximum_tpm_value, 
                 length.out = 5),
             base = 5
           )
         )

         y <- ggplot(data = tpm_df, 
                     aes(
                       x = transmission,
                       y = tpm, 
                       fill = Geneid
                     )) +
           geom_bar(position = position_stack(),
                    stat = 'identity',
                    colour = 'black',
                    size = 0.1) +
           scale_fill_manual(
             breaks = unique(tpm_df$Geneid),
             values = setNames(
               rep_len(
                 'white', 
                 length.out = length(unique(tpm_df$Geneid))
               ),
               unique(tpm_df$Geneid)
             )
           )+
           scale_color_manual('black') + 
           # labs(fill = stack_label) +
           scale_y_continuous(breaks = max_tpm) +
           theme_classic() +
           scale_size(range = c(0,3)) +
           ylab(
             bquote('total '~italic(.(loci_name))~italic(.('pir'))~' TPM')
           ) + 
           facet_grid(.~stage, 
                      scales = "free_x",
                      space = "free_x") +
           theme(
             axis.title.x=element_blank(),
             axis.text.x=element_text(size = 8),
             axis.text.y = element_text(size = 5),
             axis.title.y = element_text(size = 8),
             legend.title = element_text(size = 7),
             legend.text = element_text(size = 5),
             legend.key.size = unit(3,'mm'),
             legend.key.width = unit(2,'mm'),
             strip.text.x = element_text(size=4, 
                                         face = 'bold', 
                                         angle = 90),
             strip.background = element_rect(
               color="black", 
               size=0.1, 
               linetype="solid"
             )
           )
         if(global_save_var){
             ggsave2(paste0('tl19-07_',
                            gsub(loci_name, 
                                 pattern  = ' ', 
                                 replacement = '_'),
                            '_totalTPM_barchart.pdf'),
                     plot = y,
                     dpi = 300,
                     width = 15,
                     height = 10,
                     units = 'cm')
         }
         y
       })
```

* Cyclical expression generally similar between different chromosomes.
* 01Left, 05Right dominated by the ancestral and pan-rodent pirs.
* 01Right, 03Right, 07Right, 08Left, 08Right, 09Left, 10Left, 10Right, 11Left, 11Right, 12Right, 13Right, 14Left, 14Right show association with SBP expression.
* 04Right in rMT peaks in Early Trophs, and is spread across multiple cir genes.
* 10Left peaks in Early-Mid Rings mostly due to one pir PCHAS_1001351. 12Right also peaks then in SBP due to one or two pirs.
* In some loci the higher expression in rMT is massive; 03Left, 06Left, 06Right, 12Left.

# Lower expression of _pir_ genes versus the rest of the genome

```{r low_pir_expression}
set.seed(4534)

gene_expression_df <- twentyfour_ct_comb$tpm %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(pir_gene = ifelse(Geneid %in% cir_id,
                           'pir',
                           'non-pir')) %>% 
  group_by(stage, 
           pir_gene,
           transmission) %>% 
  slice_sample(n = length(cir_id)) %>%
  summarise(mean_tpm = mean(tpm),
            med_tpm = round(median(tpm),
                            digits = 6))

gene_expression_df %>% 
  group_by(transmission, 
           pir_gene) %>% 
  slice_min(mean_tpm, n =1)

gene_expression_df %>% 
  group_by(transmission, 
           pir_gene) %>% 
  slice_max(mean_tpm, n =1)

y <- ggplot(gene_expression_df,
       aes(x = pir_gene,
           y = log1p(mean_tpm))) +
  facet_grid(~transmission) +
  ggtitle(expression(paste(italic('P. chabaudi'), ' 24h cycle'))) +
  xlab('Gene group')+
  ylab('log(Median TPM + 1)')+
  # geom_violin()+
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               # binwidth = 1,
               # dotsize = 2,
               method = 'dotdensity') +
  geom_path(aes(group = stage,
                col = stage)) +
  theme_classic()
y

if(global_save_var){
ggsave2('tl19-07_pir-v-nonpir_mean_TPM_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')
}
```

# Fourier analysis of the expression data.

Used a script written by Chris Illingham and shared by Adam Reid. This is written in C++ so I don't quite understand all of it, but from what I can tell it calculates the phase order for each gene given the data points using the sine/cosine Fourier transform (as opposed to using complex numbers).

```{r different_phase_genes}
phase_df <- read_delim('./Fourier_code/Fourier_data.out',
                       delim = " ",
                       trim_ws = TRUE)
# This definitely isn't reading in correctly past Phase. There's spaces in the column names for 'Normalised input' which is missing stuff up.

# total_tpm_expression(list_genes = list(PCHAS_0100651 = 'PCHAS_0100651',
#                                        PCHAS_0100600 = 'PCHAS_0100600',
#                                        PCHAS_0100700 = 'PCHAS_0100700'))

phase_ordered_df <- dplyr::arrange(phase_df,
                                   Phase)

select_gene_index <- ceiling(
  sapply(
    c(1/nrow(phase_ordered_df), 
      1/4, 
      1/2, 
      3/4, 
      1),
    `*`,
    nrow(phase_ordered_df)
  )
)
#Oddly I think the 1/nrow multiplication gives a double/float (i.e. a decimel) which subsetting then converts to an integer giving zero!

# (phaseiest_genes <- phase_ordered_df[select_gene_index,])
# phaseiest_genes <- phaseiest_genes$ID
# phaseiest_genes_list <- as.list(phaseiest_genes)
# names(phaseiest_genes_list) <- phaseiest_genes
# total_tpm_expression(
#   phaseiest_genes_list
#   )

#Note that the genes at certain phase extremities may have very low expression

#The proportionally top genes of each phase

phaseiest_genes <- list(
  top_phase = slice_max(phase_df,
                        order_by = Phase,
                        n = 100)$ID, #Highest
  upp_quart_phase = slice_min(
    slice_max(phase_df,
              order_by = Phase,
              prop = 0.25),
    order_by = Phase,
    n = 100
  )$ID,
  upp_third_phase = slice_min(
    slice_max(phase_df,
              order_by = Phase,
              prop = 0.33),
    order_by = Phase,
    n = 100
  )$ID,
  mid_phase = slice_min(
    slice_max(phase_df,
              order_by = Phase,
              prop = 0.55),
    order_by = Phase,
    n = 100
  )$ID, #Middle (lowest of the top 50%)
    low_quart_phase = slice_max(
    slice_min(phase_df,
              order_by = Phase,
              prop = 0.25),
    order_by = Phase,
    n = 100
  )$ID,
  min_phase = slice_min(phase_df,
                        order_by = Phase,
                        n = 100)$ID #Lowest
)
total_tpm_expression(phaseiest_genes)
```

Expression profiles of the pir genes with Fourier transforms predicting a frequency over 1Hz (all 2Hz)

```{r pirs_over_1Hz}
two_freq_pirs <- as.list(
  filter(phase_df,
       ID %in% cir_id,
       Max.freq > 1)$ID
  )
names(two_freq_pirs) <- unlist(two_freq_pirs)
total_tpm_expression(list_genes = two_freq_pirs)
```

```{r}
filter(phase_df,
       Max.freq > 1) %>% 
  arrange(desc(Significance))

filter(phase_df,
       Max.freq > 1) %>% 
  group_by(Max.freq) %>% 
  top_n(Significance, 
        n = 3) %>% 
  arrange(desc(Max.freq), 
          desc(Significance))
total_tpm_expression(list_genes = list(
  four_Hz = 'PCHAS_1460200',
  three_Hz = 'PCHAS_1317100',
  two_Hz = 'PCHAS_1411500'
))

phase_df$Max.freq %>% unique

arrange(phase_df, 
        Significance)

filter(phase_df, 
        abs(Phase) < 0.5)
```

# Sessioninfo

```{r}
sessionInfo()
```

### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_chabaudi24h_PCA_3D_rgl.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

---
title: "tl19-07_chabaudi24h_PCA_3D_rgl"
author: "Timothy Little"
date: "12/01/2021"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/tl19-07_audreybrugat24hchabaudi")
#Note that the above doesn't give an error if the dir doesn't exist
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r}
library('rgl')
library('magick')
library('dplyr')
library('readxl')
library('tibble')
library('factoextra')
library('stringr')
library('viridis')
```

```{r}
rmt <- read_xlsx('tl19-07_TPM_rMT_nopseudoRescaled_labelsStage_allSamples.xlsx') %>% 
  select(!matches('sch')) %>% 
  rename_with(~ paste0('rMT_',
                       .x),
              matches('[[:digit:]]{2}h'))
sbp <- read_xlsx('tl19-07_TPM_SBP_nopseudoRescaled_labelsStage_allSamples.xlsx') %>% 
  select(!matches('sch')) %>% 
  rename_with(~ paste0('SBP_',
                       .x),
              matches('[[:digit:]]{2}h'))
tpm_data <- merge(rmt, 
                  sbp) %>% 
  column_to_rownames(var = 'Geneid')
tpm_data <- tpm_data[,sort(colnames(tpm_data))]

# apply(tpm_data, 2, function(x) 1e10*(sum(x) - 1e6))
```

# 3D PCA to save

```{r 3d_PCA_save_rgl}

twentyfour_TPM_pca <- prcomp(tpm_data,
                             center = TRUE, 
                             scale. = TRUE)

twentyfour_TPM_pca_var <- get_pca_var(twentyfour_TPM_pca)
twentyfour_TPM_pca_coords <- data.frame(twentyfour_TPM_pca_var$coord)

pca_importance <- as.data.frame(summary(twentyfour_TPM_pca)$importance)
pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)

twentyfour_TPM_pca_coords <- data.frame(
  sample = rownames(twentyfour_TPM_pca_var$coord),
  PC1 = twentyfour_TPM_pca_coords$Dim.1, 
  PC3 = twentyfour_TPM_pca_coords$Dim.3, 
  PC2 = twentyfour_TPM_pca_coords$Dim.2
) %>% 
  dplyr::mutate(stage = gsub(sample,
                             pattern = "_[[:digit:]]$", 
                             replacement = "")) %>% 
  dplyr::mutate(time = str_extract(sample, 
                                   pattern = '[[:digit:]]{2}h')) %>% 
  dplyr::mutate(transmission = str_split_fixed(sample, 
                                               pattern = "_", 
                                               n = 3)[, 1])

colornames <- viridis(length(unique(twentyfour_TPM_pca_coords$time)))
colornames <- setNames(colornames, sort(unique(twentyfour_TPM_pca_coords$time)))

twentyfour_TPM_pca_coords$cols <- colornames[
  match(twentyfour_TPM_pca_coords$time,
        names(colornames))
]

twentyfour_TPM_pca_coords_rmt <- filter(twentyfour_TPM_pca_coords, 
                                        transmission == 'rMT')
twentyfour_TPM_pca_coords_sbp <-  filter(twentyfour_TPM_pca_coords, 
                                        transmission == 'SBP')

```

```{r movie3d_with_loop}
movie3d_loop <- function(f, duration, dev = cur3d(), ..., fps = 10, 
                         movie = "movie", frames = movie, dir = tempdir(), 
                         convert = NULL, clean = TRUE, verbose = TRUE,
                         top = TRUE, type = "gif", startTime = 0){
  olddir <- setwd(dir)
  on.exit(setwd(olddir))
  for (i in round(startTime * fps):(duration * fps)) {
    time <- i/fps
    if (cur3d() != dev) 
      set3d(dev)
    stopifnot(cur3d() != 0)
    args <- f(time, ...)
    subs <- args$subscene
    if (is.null(subs)) 
      subs <- currentSubscene3d(dev)
    else args$subscene <- NULL
    for (s in subs) par3d(args, subscene = s)
    filename <- sprintf("%s%03d.png", frames, i)
    if (verbose) {
      cat(gettextf("Writing '%s'\r", filename))
      flush.console()
    }
    rgl.snapshot(filename = filename, fmt = "png", top = top)
  }
  cat("\n")
  if (.Platform$OS.type == "windows") 
    system <- shell
  if (is.null(convert) && requireNamespace("magick")) {
    m <- NULL
    for (i in round(startTime * fps):(duration * fps)) {
      filename <- sprintf("%s%03d.png", frames, i)
      frame <- magick::image_read(filename)
      if (is.null(m)) 
        m <- frame
      else m <- c(m, frame)
      if (clean) 
        unlink(filename)
    }
    m <- magick::image_animate(m, fps = fps, loop = 0, dispose = "previous") 
    #Only change is here; 'loop = 1' -> 'loop = 0'
    magick::image_write(m, paste0(movie, ".", type))
    return(invisible(m))
  }
  else if (is.null(convert)) {
    warning("R package 'magick' is not installed; trying external package.")
    convert <- TRUE
  }
  if (is.logical(convert) && convert) {
    progname <- "magick"
    version <- try(system2(progname, "--version", stdout = TRUE, 
                           stderr = TRUE), silent = TRUE)
    if (inherits(version, "try-error") || !length(grep("ImageMagick", 
                                                       version))) {
      progname <- "convert"
      version <- try(system2(progname, "--version", stdout = TRUE, 
                             stderr = TRUE), silent = TRUE)
    }
    if (inherits(version, "try-error") || !length(grep("ImageMagick", 
                                                       version))) 
      stop("'ImageMagick' not found")
    filename <- paste0(movie, ".", type)
    if (verbose) 
      cat(gettextf("Will create: %s\n", file.path(dir, 
                                                  filename)))
    convert <- paste(progname, "-delay 1x%d %s*.png %s.%s")
  }
  if (is.character(convert)) {
    convert <- sprintf(convert, fps, frames, movie, type, 
                       duration, dir)
    if (verbose) {
      cat(gettextf("Executing: '%s'\n", convert))
      flush.console()
    }
    system(convert)
    if (clean) {
      if (verbose) 
        cat(gettext("Deleting frames\n"))
      for (i in 0:(duration * fps)) {
        filename <- sprintf("%s%03d.png", frames, i)
        unlink(filename)
      }
    }
  }
  invisible(convert)
}
```

```{r}
r3dDefaults$windowRect <- c(100, 100, 750, 750)
rgl.init()
plot3d(
  twentyfour_TPM_pca_coords_rmt$PC1,
  twentyfour_TPM_pca_coords_rmt$PC2,
  twentyfour_TPM_pca_coords_rmt$PC3,
  col = twentyfour_TPM_pca_coords_rmt$cols,
  type = "s", 
  radius = .025,
  alpha = 0.65,
  xlab = '',
  ylab = '',
  zlab = ''
)
cubes <- shapelist3d(
  cube3d(),
  twentyfour_TPM_pca_coords_sbp$PC1,
  twentyfour_TPM_pca_coords_sbp$PC2,
  twentyfour_TPM_pca_coords_sbp$PC3,
  size = 0.017,
  alpha = 0.65,
  # ambient = 'black',
  emission = twentyfour_TPM_pca_coords_sbp$cols,
  col = twentyfour_TPM_pca_coords_sbp$cols,
  # shininess = 25,
  # specular = 'black'
  lit = TRUE
)
#Add outlines (copy pasted from 
# https://stackoverflow.com/questions/39730889/rgl-drawing-a-cube-with-colored-faces-vertex-points-and-lines )
for (i in 1:length(cubes)){
  c3d <- cubes[[i]]
  for (i in 1:6){
    lines3d(t(c3d$vb)[c3d$ib[,i],])
  }
}
# add titles
title3d('', 
        '', 
        paste0('PC1 ', pc1_importance, '%'), 
        paste0('PC2 ', pc2_importance, '%'), 
        paste0('PC3 ', pc3_importance, '%'))

empty_length <- floor(length(unique(twentyfour_TPM_pca_coords$time))/2)
legend3d("topright",
         legend = c(
           unique(twentyfour_TPM_pca_coords$time),
           rep_len('', 
                   length.out = empty_length),
           'RMT',
           'SBP',
           rep_len('', 
                   length.out = empty_length)
         ), 
         pch = c(
           rep_len(16, 
                   length.out = empty_length*2),
           rep_len(NA, 
                   length.out = empty_length) ,
           16,
           15,
           rep_len(NA, 
                   length.out = empty_length)
           ), 
         col = c(
           unique(twentyfour_TPM_pca_coords$cols),
           rep_len(NA, 
                   length.out = empty_length),
           'blue',
           'blue',
           rep_len(NA, 
                   length.out = empty_length)
         ), 
         cex=1, 
         inset=c(0.02),
         ncol = 2)
aspect3d("iso")

# spin3d rpm defaults at 5 rpm - so 60/5 = 12s carries one full rotation.
movie3d_loop(spin3d(axis = c(0,0,1)),
        duration = 12,
        movie = 'tl19-07_pca3d_test',
        type = 'gif',
        dir = getwd())

# pca3dplot <- scene3d()
```

This works alright at the moment, although the additional shape I add for the 'SBP' samples seems stretched, probably because of the different lengths of the xyz axes. "iso" setting in aspect3d fixes this, so it is indeed to do with that.

### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_chabaudi24h_pir_GeneSetEnrichment.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

---
title: "tl19-07_chabaudi24h_pir_GeneSetEnrichment"
author: "Timothy Little"
date: "21/01/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)

# Set the working directory for the document where the data is located.
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/tl19-07_audreybrugat_24h_chabaudi")
# Change options so that numbers are displayed with all the zeros (e,g, as 1000000 not 1e6)
options(scipen = 999) 
```

```{r load_packages}
library('dplyr')
library('readr')
library('readxl')
library('data.table')
library('tibble')
library('qusage')
library('ggplot2')
library('stringr')
library('RColorBrewer')
```

```{r read_count_data}
# chab_counts <- read_xlsx('Manuscript_Frontiers_chabaudi_24h_counts.xlsx') %>% 
#   filter(!Geneid %in% 'PCHAS_0420800') #Removing the meddlesome pseudogene

chab_tpm <- read_xlsx('Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
          sheet = 'chab_TPM_bioreps')

# chab_tpm %>% filter(Geneid %in% 'PCHAS_0420800')

#Filter low expressed genes (greater than 10 tpm in at least one sample), convert to a matrix, add 1 for the log transformation, then transform to log2
chab_tpm <- column_to_rownames(chab_tpm, var = 'Geneid')

chab_tpm <- chab_tpm[apply(chab_tpm, 1, function(x) {any(x >= 10)}),]

chab_tpm_log2plus1.mat <- chab_tpm %>% 
  as.matrix %>% 
  `+`(.+1) %>% 
  log(base = 2)

chab_tpm_log2plus1.mat[1:5,1:5]
```

```{r generate_labels}
#Use make.names to remove the '-' from the labels
colnames(chab_tpm_log2plus1.mat) <- make.names(colnames(chab_tpm_log2plus1.mat))

labels <- colnames(chab_tpm_log2plus1.mat)

#We want the life cycle stages to be the different groups for now.

lifeCycle_labels <- str_extract(labels, 
                                pattern = '[[:digit:]]{2}h')
unique(lifeCycle_labels) %>% sort

transmission_labels <- str_extract(labels, 
                                pattern = 'rMT|SBP')
unique(transmission_labels) %>% sort

```

```{r generate_genesets}

pir_info <- read_xlsx('Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
                      sheet = 1)

pir_geneset <- list(pirs = pir_info$Geneid)

chrom_list <- lapply(unique(pir_info$chrom), 
                     function(chrom_nam) filter(pir_info, 
                                                chrom == chrom_nam)$Geneid)

names(chrom_list) <- unique(pir_info$chrom)

#Remove groups with only three members or fewer.
chrom_list <- chrom_list[unlist(lapply(chrom_list, function(x) length(x) > 3))]

subfam_list <- lapply(unique(pir_info$subfam), 
                     function(clade_nam) filter(pir_info, 
                                                subfam == clade_nam)$Geneid)
names(subfam_list) <- unique(pir_info$subfam)
subfam_list <- subfam_list[unlist(lapply(subfam_list, function(x) length(x) > 3))]

ShortLong_list <- 
  lapply(unique(pir_info$SorL), 
                     function(SorL_nam) filter(pir_info, 
                                                SorL == SorL_nam)$Geneid)
names(ShortLong_list) <- unique(pir_info$SorL)
ShortLong_list <- ShortLong_list[unlist(lapply(ShortLong_list, function(x) length(x) > 3))]

```

```{r running_qusage_pirs}
#For each life cycle stage:
qs.results.list <- lapply(
  unique(transmission_labels), 
  function(transmission){

    index_trans <- str_detect(colnames(chab_tpm_log2plus1.mat), 
                              pattern = transmission)

    #Perform qusage
    qs.results <- qusage(chab_tpm_log2plus1.mat[,index_trans],
                         lifeCycle_labels[index_trans],
                         '14h-02h',
                         pir_geneset,
                         n.points = 2^18)
  }
)

names(qs.results.list) <- unique(transmission_labels)

qsTable(qs.results.list$SBP)
qsTable(qs.results.list$rMT)
```

```{r}
 qs.results.trans.chrom <- qusage(chab_tpm_log2plus1.mat,
                              transmission_labels,
                              'rMT-SBP',
                              chrom_list,
                              n.points = 2^18)
```

```{r}
qsTable(qs.results.trans.chrom)
```

```{r}
 qs.results.trans.pirs <- qusage(chab_tpm_log2plus1.mat,
                              transmission_labels,
                              'rMT-SBP',
                              pir_geneset,
                              n.points = 2^18)
```

```{r}
qsTable(qs.results.trans.pirs)
```

WTF

```{r}

index_14h <- str_detect(colnames(chab_tpm_log2plus1.mat), pattern = '14h')

qs.results.trans.pirs.14h <- qusage(chab_tpm_log2plus1.mat[,index_14h],
                                transmission_labels[index_14h],
                                'rMT-SBP',
                                pir_geneset,
                                n.points = 2^18)

qsTable(qs.results.trans.pirs.14h)

plotGeneSetDistributions(qs.results.trans.pirs.14h)
```

```{r qusage_pirs-expressed-higher-RMT}

rmt_14h_hi <- read_xlsx('Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
                      sheet = 'chab_TPM_bioreps') %>% 
  filter(Geneid %in% cir_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  filter(time == '14h') %>% 
  group_by(transmission, time, Geneid) %>% 
  summarise(tpm_mean = mean(tpm)) %>% 
  dcast(Geneid ~ transmission, 
        margins = c('tpm_mean')) %>% 
  filter((rMT - SBP) > 150) %>% 
  .$Geneid

qs.results.trans.pirs.14h.rmthi <- qusage(chab_tpm_log2plus1.mat,
                                          transmission_labels,
                                          'rMT-SBP',
                                          list(rmt_hi = rmt_14h_hi),
                                          n.points = 2^18)

qsTable(qs.results.trans.pirs.14h.rmthi)

```

### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_manuscript_figures.Rmd
Rename and create chabaudi pir analysis files and update code for SBP and improved all pir heatmap figure

--- 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_audreybrugat24hchabaudi.Rmd
@@
 ---
-title: "TL19-07_Audrey_24h_chabaudi"
-date: "17/06/2020"
-output: 
-  html_document:
-    theme: cosmo
-    code_download: true
-    toc: true
-    number_sections: true
-    toc_float:
-      collapsed: false
-      smooth_scroll: true
+title: "tl19-07_manuscript_figures"
+author: "Timothy Little"
+date: "06/11/2020"
+output: html_document
 ---
@@
 ```
-
-To take into account:
-
-- SBP is more synchronous than MT, do we have counts?
-
-- MT and SBP may seem very different due to differences in the amount of material for sequencing.
-
-Suggestions:
-
-- PCA of SBP on its own, and MT on its own, UMAP maybe too?
-
-- Use DESeq2 to get the diff expr genes between time points and then use this to plot a PCA.
-
-- z-score plots: Have a max/min expression heatmap instead of showing the median logTPM.
-
-- Just remove the Unknown sample since it contributes nothing without an identity.
-
-- The value of the expression threshold (variable: min_median) is very high (over 4 TPM) - is this reasonable?
-
-Top_n_genes:
-
-~~- Remove Unknown entirely by re-saving top_genes table. Also need to separate the top_n bit using rMT and SBP.~~
-
-~~- Why is there a NULL printed right at the end? Issue with names~~
-
-Questions:
-
-- What day are these samples taken? d6?
-
-- How many passages had 'SBP' been through?
-
-- What are the differential counts of the samples taken? - Audrey's are on the Shared folder.
-
-# Audrey's experimental description
-
-24 Hours cycle P. chabaudi RNA seq.
-
-Aim: Determine pir expression for each parasite stage in recently mosquito transmitted P. chabaudi parasites.
-
-Principle: using a stabilate from recently mosquito transmitted P. chabaudi parasites infected C57BL/6 and extract RNA at each stage of parasite cycle (every 3 hours) during 24h.
-
-Animals: 4 C57BL/6 J CD45.1 males for stabilates thawing (2 in RL, 2 in NL) and 27 C57BL/6 J CD45.1 females for experiments (12 in NL and 15 in RL) (issued 17/08/2016).
-
-* AV16-35	2 C57BL/6 J CD45.1 males		RL.
-
-* AV16-36	2 C57BL/6 J CD45.1 males		NL.
-
-* AV16-37	15 C57BL/6 J CD45.1 females	RL.
-
-* AV16-38	12 C57BL/6 J CD45.1 females	RL.
-
-Experiment:
-
-* d-14	11am	 Infected 2 mice (males) with recently mosquito transmitted parasite stabilate (MTC tower 6 box 6.1 position 31-65 (late troph)) in NL (AV16-36).
-
-* d-14	5pm	Infecte 2 mice (males) with recently mosquito transmitted parasite stabilate (MTC tower 6 box 6.1 position 31-65 (late troph)) in RL (AV16-35).
-
-* d-7	11am	pass AV16-35 into 15 mice (females) 105 iRBC/ mouse IP (AV16-37) RL.
-
-* d0	 2pm	smear 3 mice and bleeding .
-
-* d0	 2-5pm 	pellet parasites and trizol of 2pm samples.
-
-* d0	 5pm	smear 3 mice and bleeding .
-
-* d0	 5-8pm 	pellet parasites and trizol 5pm samples.
-
-* d0	 8pm 	smear 3 mice and bleeding.
-
-* d0	 8-11pm 	pellet parasites and trizol 8pm samples.
-
-* d0	 11pm	smear 3 mice and bleeding .
-
-* d0	11pm-2am 	pellet parasites and trizol 11pm samples.
-
-* d1	 2am	smear 3 mice and bleeding .
-
-* d1	 2-5am 	pellet parasites and trizol of 2am samples.
-
-* d1	 5am	smear 3 mice and bleeding .
-
-* d1	 5-8am 	pellet parasites and trizol 5pm samples.
-
-* d1	 8am 	smear 3 mice in each group and bleeding .
-
-* d1	 8-11am 	pellet parasites and trizol 8pm samples.
-
-* d1	 11am	smear 3 mice in each group and bleeding .
-
-* d1	 11am 	bleeding 3 mice (AV16-37) and parasite culture.
-
-* d1	11am-2pm 	pellet parasites and trizol 11pm samples.
-
-* d1	 4-6pm	schizonte pellet and trizol.
-
-Dates:
-
-d-14 : 25/08/2016.
-
-d-7 : 01/09/2016.
-
-d0 : 08/09/2016.
-
-d1 : 09/09/2016.
 
@@
 
-```{r global_save_var}
-#Make this TRUE if you want to run this script and save all the plots.
-global_save_var <- TRUE
-```
-
-```{r tpm, results = 'hide', message = FALSE}
-# # Calculating the transcript-per-million 
-# 
-# We need to write a function to calculate the TPM from the count data we have generated.
-This function uses the 'chabaudi_transcript_lengths' object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.
-
-tpm = function(count_data, lengths = chabaudi_transcript_lengths) {
-
-  #Print an error if the id columns do not match
-  # first if statement to test that both have a GeneID column
-  if(any(grepl(colnames(lengths), 
-               pattern = 'id', 
-               ignore.case = T)) == T & any(grepl(colnames(count_data), 
-                                                  pattern = 'Geneid', 
-                                                  #This needs to be Geneid when some of the samples have 'id' in them
-                                                  ignore.case = T)) == T){
-    if(all(count_data[grep(colnames(count_data), 
-                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
-                                                                             pattern = 'id', ignore.case = T)])) {
-      print('Count data and length gene ids both match')
-    } else {
-      stop('Error: Count data and length gene ids DO NOT MATCH, please make sure they are ordered identically.')
-    }
-  } else {
-    stop('Error: Count data and/or length table supplied do not have an id column.')
-  }
-
-  #Remove the id column from supplied count_data
-  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = T)) == T){
-    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = T)]
-  } else{
-    counts_tpm <- count_data}
-
-  #Remove the id column from supplied lengths
-  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
-    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
-  }else{
-    lengths.table <- lengths}
-  lengths.table = matrix(lengths.table)
-  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop
-
-  #Now for the actual calculation
-  rpm <- counts_tpm
-  for (i in 1:dim(counts_tpm)[2]) {
-    for (j in 1:dim(counts_tpm)[1]) {
-      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
-    }
-    scale.factor <- sum(rpm[,i])/(1E6)
-    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
-  }
-  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
-    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = T), drop = FALSE], 
-                           counts_tpm)
-  } 
-  return(counts_tpm)
-}
-```
-
-```{r agree function}
-
-agree = function(input_list){
-  for(i in 1:(length(input_list)-1)){
-    input_list[[i+1]] = input_list[[i]][ input_list[[i]] %in% input_list[[i+1]] ]
-  }
-  return(input_list[[length(input_list)]])
-}
-
-disagree = function(input_list){
-  for(i in 1:(length(input_list)-1)){
-    input_list[[i+1]] = input_list[[i]][ !input_list[[i]] %in% input_list[[i+1]] ]
-  }
-  return(input_list[[length(input_list)]])
-}
-
+```{r function_mround}
 mround <- function(x,base){ 
-  base*round(x/base) 
-} 
+  base*ceiling(x/base) 
+} 
+```
+
+```{r function_my_max}
+my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
 ```
@@
 }
-```
-
-```{r pca_function}
-pca_function <- function(input_ct, 
-                         genevec = NULL, 
-                         scale.pca = TRUE, 
-                         three.dimensional = FALSE,
-                         legend_settings = list(),
-                         colornames = NULL){
-
-  twentyfour_tpm_pca <- input_ct$tpm
-  if(!is.null(genevec)){
-    twentyfour_tpm_pca <- input_ct$tpm %>% filter(Geneid %in% genevec)
-
-    ### Re-scaling the TPM data
-    twentyfour_tpm_pca_rescale <- lapply(colnames(twentyfour_tpm_pca), 
-                                         function(column_name){
-                                           column <- getElement(twentyfour_tpm_pca, 
-                                                                paste(column_name))
-                                           if(is.numeric(column)){
-                                             #is.numeric to avoid trying to scale the Geneid column
-                                             tpm_scale_factor <- 1e6 / sum(column)
-                                             tpm_scaled <- as.numeric(column*tpm_scale_factor)
-                                           }else{
-                                             tpm_scaled <- column
-                                           }
-                                           return(tpm_scaled)
-                                         }) %>% 
-      bind_cols
-
-    colnames(twentyfour_tpm_pca_rescale) <- colnames(twentyfour_tpm_pca)
-    twentyfour_tpm_pca <- twentyfour_tpm_pca_rescale
-    ###
-
-  }
-  rownames(twentyfour_tpm_pca) <- twentyfour_tpm_pca$Geneid
-  twentyfour_tpm_pca <- select(twentyfour_tpm_pca, -Geneid) %>% 
-    prcomp(center = TRUE, scale. = scale.pca)
-  #Need to look into what exactly scale and center are doing
-  twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
-  twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
-
-  twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
-                                          PC1 = twentyfour_tpm_pca_coords$Dim.1,
-                                          PC3 = twentyfour_tpm_pca_coords$Dim.3,
-                                          PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
-    dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-    dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
-    dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
-
-  if(is.null(colornames)){
-    # colornames <- c(cbPalette, 'black', 'red')
-    colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
-    colornames <- setNames(colornames, 
-                           sort(unique(twentyfour_tpm_pca_coords$time)))
-  }
-
-  pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
-  pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
-  pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
-  pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
-
-  if(three.dimensional){
-    pl <- plot_ly(twentyfour_tpm_pca_coords, 
-                  x  = ~PC1, y = ~PC2, z = ~PC3,  mode = 'markers',
-                  type = 'scatter3d', color = ~time, text = ~sample,
-                  colors = colornames,
-                  # size = ~total_features,
-                  marker = list(opacity = 1, size = 5),
-                  symbol = ~transmission, 
-                  symbols = c('circle','x')) %>% 
-      layout(scene = list(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
-                          yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
-                          zaxis = list(title = paste0('PC3 ',pc3_importance,'%'))))
-  }else{
-    pl <- plot_ly(twentyfour_tpm_pca_coords, 
-                  x  = ~PC1, y = ~PC2,
-                  type = 'scatter', mode = 'markers',
-                  color = ~time, text = ~sample,
-                  colors = colornames,
-                  # size = ~total_features,
-                  marker = list(opacity = 1, size = 16),
-                  symbol = ~transmission, symbols = c('circle','x')) %>% 
-      layout(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
-             yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
-             legend = legend_settings)
-  }
-  pl
-}
-```
-
-```{r function_draw_heatmap_chabaudi_24h}
-# Function to draw a heatmap with a list of genes
 
@@
                   "Late.Trophs")
-
-# Note that there may seem to be discrepencies between the data and the heatmap because this function may be using the median and/or mean value of the TPM
-draw_heatmap_chabaudi_24h <- function(genelist, 
-                                scale = 'max', 
-                                threshold = 0, 
-                                datasets = twentyfour_ct, 
-                                filter_tpm = NULL,
-                                stage_diff = NULL, 
-                                filter_stage = NULL, 
-                                title_name = NULL, 
-                                drop_stage = NULL, 
-                                log_it = TRUE,
-                                collate_samples_by_stage = TRUE,
-                                order_by_med_tpm = TRUE,
-                                mean_or_median = 'mean',
-                                gene_names_for_title = NULL,
-                                ...){
-  #genelist <- cir_id
-  #For de-bugging:
-
-  # scale = 'max';
-  # threshold = 0;
-  # datasets = twentyfour_ct;
-  # filter_tpm = NULL;
-  # stage_diff = NULL;
-  # filter_stage = NULL;
-  # title_name = NULL;
-  # drop_stage = NULL;
-  # log_it = TRUE;
-  # collate_samples_by_stage = TRUE;
-  # order_by_med_tpm = TRUE;
-  # mean_or_median = 'mean';
-
-  # if(!is.null(datasets$tpm)){
-  #   #Probably means only one dataset was input so lapply will malfunction
-  #   datasets <- list(tpm = datasets)
-  # }
-
-  if(log_it == TRUE){
-    threshold <- log1p(threshold)
-    title <- 'log(TPM + 1)'
-  }else{
-    title <- 'TPM'
-  }
-
-  expt_tpms_top_genes = lapply(datasets, function(tpm_input){
-    # tpm_input <- datasets[[1]]
-    tpms = tpm_input$tpm %>% 
-      #Melt the data frame for using dplyr
-      filter(Geneid %in% genelist) %>% 
-      melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm')
-
-    if(collate_samples_by_stage){ 
-      #Create a new variable 'stage' from the sample names without rep numbers
-      tpms <- mutate(
-        tpms, 
-        stage = str_extract(
-          sample,
-          pattern = '.+(?=_[[:digit:]]{1,2})'
-        )
-      )
-    }else{
-      tpms <- mutate(tpms, 
-                     stage = sample)
-    }
-
-    tpms <- mutate(tpms, time = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
-      #Convert the times given in 'h' to robustly coded times in 24h clock
-      mutate(time = ifelse(grepl(time, pattern = '[[:digit:]]h$'), 
-                           strptime(time, format="%Hh") %>% format(., "%Hh"),
-                           time)) %>% 
-      mutate(time = ifelse(grepl(time, pattern = '[[:digit:]](a|p)m'), 
-                           strptime(time, format="%I%p") %>% format(., "%Hh"),
-                           time))
-
-    colnam = sort(unique(tpms$stage))
-
-    if(mean_or_median == 'median'){
-      tpms = tpms %>% 
-        group_by(Geneid, stage, time) %>%  
-        dplyr::summarise(mean_tpm = median(tpm)) %>%  
-        group_by(stage) %>% 
-        recast(Geneid ~ stage + time + 'mean_tpm', median) #recast into short-form for complex heatmap
-    } else if (mean_or_median == 'mean'){
-      tpms = tpms %>% 
-        group_by(Geneid, stage, time) %>%  
-        dplyr::summarise(mean_tpm = mean(tpm)) %>%  
-        group_by(stage) %>% 
-        recast(Geneid ~ stage + time + 'median_tpm', mean) #recast into short-form for complex heatmap
-    }
-
-    colnames(tpms) = c('Geneid',as.character(colnam))
-
-    if(!is.null(order_vector)){
-      tpms <- tpms[,
-                   ordering(colnames(tpms),
-                            c('Geneid',
-                              order_vector))]
-      #Ordering using the ordering function and order_vector defined in the functions section.
-    }
-
-    #Apply drop_stage
-    if(!is.null(drop_stage)){
-      tpms <- select(tpms, -!!(paste(drop_stage)))
-    }
-    #Apply stage_diff to get genes which are expressed higher in one stage than another
-    if(!is.null(stage_diff)){
-      if(is.na(stage_diff[1])|is.na(stage_diff[2])){
-        stop('stage_diff must be a vector of two named stages to select genes with TPM greater in the former than the latter')
-      }
-      tpms <- tpms[select(tpms, paste(stage_diff[1])) > select(tpms, paste(stage_diff[2])), ]
-    }
-    #Apply filter_tpm
-    if(!is.null(filter_tpm)){
-      tpms <- filter_at(tpms, vars(-Geneid), any_vars(. > filter_tpm))
-    }
-    #Apply filter_stage to remove genes expressed below filter_tpm in that stage
-    if(!any(is.null(filter_stage), is.null(filter_tpm))){
-      tpms <- filter(tpms, paste(filter_stage) >= filter_tpm)
-    }
-    #Remove PCHAS_ from the gene names for simplicity
-    # rownames(tpms) = gsub(as.character(tpms$Geneid), pattern = 'PCHAS_', replacement = '')
-    #Use the protein names for ease of understanding
-    rownames(tpms) = chabaudi_gene_info$handy_descriptions[match(tpms$Geneid, 
-                                                           chabaudi_gene_info$Geneid)]
-    tpms = select(tpms, -Geneid)
-
-    if(log_it == TRUE){
-      tpms <- log1p(tpms)
-    }
-
-    if(tpm_input$experiment %>%  is.null){
-      tpm_input$experiment = ''
-    }
-    list = list(tpms, tpm_input$experiment)
-    names(list) = c('tpm','experiment')
-    return(list)
-  })
-
-  if(
-    !all(
-      unlist(
-        lapply(
-          expt_tpms_top_genes, 
-          function(x) all(
-            x$tpm < threshold
-          )
-        )
-      )
-    )
-  ){
-    if(scale %in% 'max'){
-      maxcol <- my_max(unlist(lapply(expt_tpms_top_genes, 
-                                     function(tpms){my_max(tpms$tpm)})))
-    }else{
-      maxcol = scale
-    }
-  }
-
-  if(is.na(maxcol)|maxcol<1){maxcol <- 1}
-
-  if(order_by_med_tpm){
-    #Work out the order by median TPM
-    medians_across_expt_mat <- as.matrix(
-      sapply(
-        lapply(
-          expt_tpms_top_genes, 
-          function(x){
-            rowMedians(as.matrix(x$tpm),
-                       na.rm = TRUE) 
-          }
-        ), 
-        rbind
-      )
-    )
-    tpm_order <- order(
-      rowMedians(
-        medians_across_expt_mat, 
-        na.rm = TRUE),
-      decreasing = TRUE
-    )
-  }
-
-  #Applying the threshold last because NAs mess things up
-  expt_tpms_top_genes <- lapply(expt_tpms_top_genes, function(expt){
-    tpms <- expt$tpm
-    tpms = apply(tpms, 2, function(x){
-      ifelse(
-        x >= threshold, 
-        x, 
-        NA
-      )
-    }
-    ) 
-    #Get NA for values under threshold
-    list = list(tpms, expt$experiment)
-    names(list) = c('tpm','experiment')
-    return(list)
-  })
-
-  heatmaps = lapply(expt_tpms_top_genes, function(tpms){
-    if(tpms$experiment != last(expt_tpms_top_genes)$experiment){
-      #If it is the final element don't remove the row names
-      rownames(tpms$tpm) = c()
-    }
-    #Setting the name if one is not provided
-    if(is.null(title_name)){
-      title_name = tpms$experiment
-    }
-
-    #Re-order both data.frames by the median highest expressed pirs
-    if(order_by_med_tpm){
-      tpms$tpm <- tpms$tpm[tpm_order,, drop = FALSE]
-    }
-
-    # print(nrow(tpms$tpm))
-    #Drawing each heatmap
-    ht = Heatmap(tpms$tpm, 
-                 name = 'TPM', 
-                 col = colorRamp2(c(threshold, 
-                                    maxcol/4, 
-                                    maxcol/2, 
-                                    3*maxcol/4, 
-                                    maxcol), 
-                                  viridis(5)),
-                 cluster_rows = FALSE, 
-                 cluster_columns = FALSE,
-                 column_title = paste(title_name),
-                 column_title_gp = gpar(fontsize = 7),
-                 column_names_gp = gpar(fontsize = 7),
-                 row_names_gp = gpar(fontsize = ifelse(
-                   nrow(tpms$tpm) > 13, 
-                   12*(13/nrow(tpms$tpm)), 
-                   12)
-                 ),
-                 na_col = 'white', border = TRUE,
-                 heatmap_legend_param = list(
-                   title = paste(title), 
-                   at = round(
-                     c(threshold, 
-                       maxcol/4, 
-                       maxcol/2, 
-                       3*maxcol/4, 
-                       maxcol), 
-                     digits = 1)
-                 )
-                 # ,...
-    )
-  })
-
-  ht_list = Reduce('+', heatmaps) #To add the list elements together
-  draw(ht_list, 
-       row_title_side = 'left',
-       column_title = bquote(~italic('P. chabaudi')~'24h asexual life cycle stage - '~italic(.(gene_names_for_title))~'genes'), 
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'left')
-}
-```
-
-```{r plotly_biplot_function}
-plot_ly_biplot <- function(pca_object = twentyfour_tpm_pca, 
-                           prin_components = c(1,2,3),
-                           fontsize = 7,
-                           scale.loads = 600,
-                           separate_multigene = FALSE,
-                           three_dimensions = FALSE,
-                           col_var = 'time',
-                           gene_info = chabaudi_gene_info,
-                           textposition = "top right"){
-
-  #Check that three components are provided for three dimensional biplot
-  if(three_dimensions){
-    if(length(prin_components) != 3){
-      stop('For three dimensions, three principal components must be given for prin_components argument')
-    }
-  }
-
-  pca_object_rotations <- rownames_to_column(data.frame(pca_object$rotation), 
-                                             var = 'sample')  %>% 
-    dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-    dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
-    dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
-
-  #Setting the colours, which come from each life cycle experiment, called 'stage'.
-
-  colornames <- setNames(
-    viridis(
-      length(
-        unique(
-          getElement(
-            pca_object_rotations, 
-            paste(col_var)
-          )))),
-    sort(
-      unique(
-        getElement(
-          pca_object_rotations, 
-          paste(col_var)
-        ))))
-
-  pca_object_loadings <- data.frame(Geneid = rownames(pca_object$x),
-                                    pca_object$x)
-
-  pca_object_loadings$Geneid <- paste0(
-    str_replace(
-          as.character(pca_object_loadings$Geneid),
-          pattern = ' ',
-          replacement = ''),
-    ' ',
-    gene_info$handy_descriptions
-    [
-      match(
-        c(str_replace(
-          as.character(pca_object_loadings$Geneid),
-          pattern = ' ',
-          replacement = '')
-        ),
-        gene_info$Geneid
-      )
-      ]
-  )
-
-  if(separate_multigene){
-    #If the gene is a fam or a fam gene we want to know - haven't excluded pseudogenes here
-    pca_object_loadings$multigene <- ifelse(
-      #if it doesn't have fam or fam-something in its name then just say gene
-    is.na(
-      str_extract(
-        pca_object_loadings$Geneid,
-        pattern = 'fam-.|fam'
-      )
-    ),
-    'Gene loadings',
-    paste0(
-      str_extract(
-        pca_object_loadings$Geneid,
-        pattern = 'fam-.|fam'
-      ),
-      ' gene loadings'
-    )
-  )
-}else{
-  pca_object_loadings$multigene <- rep('Gene loadings',
-                                       stages = nrow(pca_object_loadings))
-}
-
-#Degrees of variation for the axes
-pca_importance <- as.data.frame(summary(pca_object)$importance)
-pc_importance_1 <- round(pca_importance[2,prin_components[1]]*100, digits = 1)
-pc_importance_2 <- round(pca_importance[2,prin_components[2]]*100, digits = 1)
-if(three_dimensions){
-  pc_importance_3 <- round(pca_importance[2,prin_components[3]]*100, digits = 1)
-}
-
-#scale.loads controls how long the lines are, the longer they are the more prominent the loadings but the less
-# obvious the genes.
-
-pl <- plot_ly()
-
-#For the font of the labels
-t <- list(
-  size = fontsize
-  )
-
-#For loop to add the PCA lines of the samples
-for(k in 1:nrow(pca_object_rotations)) {
-  #Have 0 so that the line comes from the origin (0,0)
-  #PC first
-  x <- c(0, pca_object_rotations[k,(prin_components[1]+1)])*scale.loads
-  #PC second
-  y <- c(0, pca_object_rotations[k,(prin_components[2]+1)])*scale.loads
-
-  if(three_dimensions){
-    #PC third
-    z <- c(0, pca_object_rotations[k,(prin_components[3]+1)])*scale.loads
-    df_plotly <- data.frame(x = x, 
-                            y = y, 
-                            z=z,
-                            stage = getElement(pca_object_rotations, 
-                                               paste(col_var))[k],
-                            sample = pca_object_rotations$sample[k])
-  }else{
-    df_plotly <- data.frame(x = x, 
-                            y = y,
-                            stage = getElement(pca_object_rotations, 
-                                               paste(col_var))[k],
-                            sample = pca_object_rotations$sample[k])
-  }
-
-  #colour based on sample
-  color <- colornames[
-    match(
-      getElement(
-        pca_object_rotations, 
-        paste(col_var)
-      )[k], 
-      rev(names(colornames)
-      )
-    )
-    ]
-
-  #Add these to the plot
-  #To avoid duplication in the legend don't show it if the name is duplicated
-  if(duplicated(getElement(pca_object_rotations, paste(col_var)))[k]){
-    showlegend <- FALSE
-  }else{
-    showlegend <- TRUE
-  }
-
-  #Add the lines for the samples depending on their position in the PCA
-  if(three_dimensions){
-    pl <- pl %>%
-      add_trace(data = df_plotly,
-                x=~x,
-                y=~y,
-                z=~z,
-                type='scatter3d',
-                mode='line',
-                color = ~stage,
-                colors = colornames,
-                line = list(width = 2),
-                text = pca_object_rotations[k,1],
-                name = getElement(pca_object_rotations, paste(col_var))[k],
-                opacity = 1,
-                showlegend = showlegend,
-                legendgroup = getElement(pca_object_rotations, paste(col_var))[k],
-                #Note that 3d plots only accept a few shapes e.g. hexagon is not accepted
-                symbols = c('circle', 
-                            'circle-open',
-                            'square',
-                            'square-open',
-                            'x',
-                            'diamond')) %>% 
-            add_text(x = df_plotly$x[2],
-               y = df_plotly$y[2],
-               z = df_plotly$z[2],
-               text = df_plotly$sample, 
-               textfont = t,
-               textposition = textposition,
-               showlegend = ifelse(k == 1, TRUE, FALSE), #Only want this option to be TRUE once
-               legendgroup = 'Toggle text',
-               name = 'Toggle text')
-  }else{
-    pl <- pl %>% 
-      add_trace(data = df_plotly,
-                x=~x,
-                y=~y,
-                type='scatter',
-                mode='line',
-                color = ~stage,
-                colors = colornames,
-                line = list(width = 2),
-                text = pca_object_rotations[k,1],
-                name = getElement(pca_object_rotations, 
-                                  paste(col_var))[k],
-                opacity = 1,
-                showlegend = showlegend,
-                legendgroup = getElement(pca_object_rotations, 
-                                         paste(col_var))[k],
-                symbols = c('circle', 
-                            'circle-open',
-                            'square',
-                            'square-open',
-                            'x',
-                            'diamond')) %>% 
-      #I don't know why it only works if we put the symbols argument here and not in the add_trace later??
-      add_text(x = df_plotly$x[2],
-               y = df_plotly$y[2],
-               text = df_plotly$sample,
-               textfont = t,
-               textposition = textposition,
-               showlegend = ifelse(k == 1, TRUE, FALSE), #Only want this option to be TRUE once
-               legendgroup = 'Toggle text',
-               name = 'Toggle text')
-  }
-}
-
-#Adding the gene loadings
-
-if(three_dimensions){
-  #Make data frame of the gene loadings
-  pca_object_loadings_df <- select(data.frame(pca_object_loadings), c(Geneid, 
-                                                                      multigene, 
-                                                                      paste0('PC', prin_components[1]),
-                                                                      paste0('PC', prin_components[2]),
-                                                                      paste0('PC', prin_components[3])))
-  colnames(pca_object_loadings_df) <- c('Geneid',
-                                        'multigene',
-                                        'prin_com_1',
-                                        'prin_com_2',
-                                        'prin_com_3')
-  pl <- pl %>% 
-    add_trace(data = pca_object_loadings_df, 
-              inherit = FALSE,
-              x = ~prin_com_1, 
-              y = ~prin_com_2, 
-              z = ~prin_com_3,
-              type = 'scatter3d', 
-              mode = 'markers',
-              marker = list(color = 'black',
-                            size = 4,
-                            opacity = 0.7),
-              text = ~Geneid,
-              name = ~multigene,
-              symbol = ~multigene,
-              showlegend = TRUE,
-              legendgroup = ~multigene) %>% 
-    layout(
-      scene = list(
-        xaxis = list(
-          title = paste0(
-            'PC', 
-            prin_components[1], 
-            ' ', 
-            pc_importance_1,
-            '%'
-          )
-        ),
-        yaxis = list(
-          title = paste0(
-            'PC', 
-            prin_components[2], 
-            ' ',
-            pc_importance_2,
-            '%'
-          )
-        ),
-        zaxis = list(
-          title = paste0(
-            'PC', 
-            prin_components[3], 
-            ' ',
-            pc_importance_3,
-            '%'
-          )
-        )
-      )
-    )
-    }else{
-      #Make data frame of the gene loadings
-      pca_object_loadings_df <- select(data.frame(pca_object_loadings), 
-                                       c(Geneid, 
-                                         multigene, 
-                                         paste0('PC', prin_components[1]),
-                                         paste0('PC', prin_components[2])))
-      colnames(pca_object_loadings_df) <- c('Geneid',
-                                            'multigene',
-                                            'prin_com_1',
-                                            'prin_com_2')
-      pl <- pl %>% 
-        add_trace(data = pca_object_loadings_df, 
-                  inherit = FALSE,
-                  x = ~prin_com_1, 
-                  y = ~prin_com_2,
-                  type = 'scatter', 
-                  mode = 'markers',
-                  marker = list(color = 'black',
-                                size = 6,
-                                opacity = 0.9),
-                  text = ~Geneid,
-                  name = ~multigene,
-                  symbol = ~multigene,
-                  showlegend = TRUE,
-                  legendgroup = ~multigene) %>% 
-        layout(
-          xaxis = list(
-            title = paste0(
-              'PC', 
-              prin_components[1], 
-              ' ', 
-              pc_importance_1,
-              '%'
-            )
-          ),
-          yaxis = list(
-            title = paste0(
-              'PC', 
-              prin_components[2], 
-              ' ',
-              pc_importance_2,
-              '%'
-            )
-          )
-        )
-    }
-pl
-}
-
-#For debugging
-# k <- 6
-
-#Switching around the add_text and add_trace for the 'Toggle text' (so that it goes first) seems to bugger up the colours
-```
-
-```{r function_mround}
-mround <- function(x,base){ 
-  base*ceiling(x/base) 
-} 
-```
-
-```{r function_my_max}
-my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
-```
-
-```{r function_uniquify}
-uniquify <- function(vector_names, first_one = TRUE){
-  new_vector <- c()
-  df_num <- data.frame(names = unique(vector_names), rep = 1)
-  if(first_one){
-    duplicated_names <- unique(vector_names[duplicated(vector_names)])
-  }else{
-    duplicated_names <- c()
-  }
-  for(i in 1:length(vector_names)){
-    if(vector_names[i] %in% duplicated_names|!first_one){
-    name <- paste(vector_names[i],
-                  df_num$rep[match(vector_names[i], df_num$names)],
-                  sep = '_')
-    #Increase rep counter
-    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
-    new_vector[i] <- name
-    }else{
-      new_vector[i] <- vector_names[i]
-    }
-  }
-  new_vector
-}
-```
-
-```{r function_firstup}
-#Capitalize first letter
-firstup <- function(x) {
-  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
-  x
-}
-```
-
-```{r function_density_samples}
-density_samples = function(input, save = FALSE){
-  #Get some colours for the lines
-  # colours = rainbow_hcl(length(colnames(input$tpm)[-1]))
-  colours = rep_len(cbPalette, length.out = length(colnames(input$tpm)[-1]))
-  list_dens = lapply(colnames(input$tpm)[-1], function(sample){
-    data = as.numeric(unlist(select(input$tpm, paste(sample))))
-    dens = density(log1p(data), bw = bw_mean)
-    #Use the index of the column to get the colour
-    colouring = colours[grep(colnames(input$tpm)[-1], pattern = paste(sample))]
-    list_return = list(dens, colouring)
-    names(list_return) = c('dens','colouring')
-    list_return
-  })
-  max_dens = max(unlist(lapply(list_dens, function(x){max(x$dens$y)})))
-  max_x = max(unlist(lapply(list_dens, function(x){max(x$dens$x)})))
-  line_width = 1
-  par(new=FALSE)
-  if(save){
-    pdf(file = paste0(input$experiment, '_tl19-07_densityplot_fixedbw.pdf'), 
-        width = 15/2.54, 
-        height = 11.5/2.54)
-  }
-  lapply(seq_along(list_dens), function(index){
-    x = list_dens[[index]]
-    if(index <= 9){
-      plot(x$dens, 
-           xlim = c(0, max_x), 
-           ylim = c(0,max_dens), 
-           type = 'l', 
-           col = x$colouring,
-           main = paste(input$experiment),
-           lwd = line_width)
-    }else if(index <= 18){
-      plot(x$dens, 
-           xlim = c(0, max_x), 
-           ylim = c(0,max_dens), 
-           type = 'l', 
-           col = x$colouring,
-           main = paste(input$experiment),
-           lwd = line_width,
-           lty = 2)
-    } else {
-      plot(x$dens, 
-           xlim = c(0, max_x), 
-           ylim = c(0,max_dens), 
-           type = 'l', 
-           col = x$colouring,
-           main = paste(input$experiment),
-           lwd = line_width,
-           lty = 3)
-    }
-    abline(v = log1p(min_median))
-    legend(max_x-4, max_dens, 
-           legend = paste(colnames(input$tpm)[-1]), 
-           col = colours,
-           lty = c(rep(1, times = 9), rep(2, times = 9), rep(3, times = 9)),
-           cex = 0.5)
-    par(new=TRUE)})
-  if(save){
-    dev.off()
-  }
-}
-```
-
-```{r function_total_tpm_expression}
-total_tpm_expression <- function(list_genes, 
-                                 save = FALSE,
-                                 tpm_data = twentyfour_ct,
-                                 stage_str_extract_regex_pattern = '(?<=_).+(?=_[[:digit:]]+)|sch.culture'){
-  lapply(names(list_genes), 
-         function(list_name){
-           paste(list_name)
-           gene_id <- getElement(list_genes, paste(list_name))
-           list_tpms_line <- lapply(tpm_data, 
-                                    function(input){
-             tpm_data = input$tpm %>% 
-               filter(Geneid %in% gene_id) %>% 
-               melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-               mutate(stage = str_extract(sample, 
-                                          pattern = paste(stage_str_extract_regex_pattern))) %>% 
-               group_by(stage, sample) %>% 
-               dplyr::summarise(total_tpm = sum(tpm))
-
-             #All genes detected
-             tpm_data_all_genes <- input$tpm %>% 
-               melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-               mutate(stage = unlist(lapply(strsplit(as.character(sample), split = '_'),
-                                            `[`,
-                                            1))) %>% 
-               # filter(tpm > threshold) %>% 
-               group_by(stage, sample) %>% 
-               dplyr::summarise(total_tpm = n()) %>% 
-               ungroup %>%  
-               group_by(stage) %>% 
-               dplyr::summarise(total_tpm_median = median(total_tpm),
-                                Number_of_replicates = n(),
-                                total_tpm_max = max(total_tpm),
-                                total_tpm_min = min(total_tpm)) %>% 
-               arrange(stage)
-
-             #Make stages with no cirs in the sample give a zero 
-             if(!all(colnames(input$tpm) %in% tpm_data$sample)){
-               missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
-               missing_stages_data_zero = data.frame(lapply(missing_stages, function(missing_st){
-                 data.frame(stage = unlist(str_extract(missing_st, 
-                                                       pattern = paste(stage_str_extract_regex_pattern))),
-                            sample = missing_st,
-                            total_tpm = 0)
-               }))
-               tpm_data = bind_rows(tpm_data, missing_stages_data_zero)
-             }
-
-             #Add experiment column
-             tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
-             max_tpm = if(max(tpm_data$total_tpm) < 10){
-               c(seq(0,mround(max(tpm_data$total_tpm),1),1))
-             }else{
-               c(seq(0,mround(max(tpm_data$total_tpm),5),5))
-             }
-
-             #Get the average data for the line graph
-             tpm_data_line = tpm_data %>% 
-               ungroup %>% 
-               group_by(stage, experiment) %>% 
-               dplyr::summarise(total_tpm_median = median(total_tpm),
-                                Number_of_replicates = n(),
-                                total_tpm_max = max(total_tpm),
-                                total_tpm_min = min(total_tpm)) %>% 
-               arrange(stage)
-
-             tpm_data_line$total_tpm_median_allgenes <- tpm_data_all_genes$total_tpm_median
-
-             return(tpm_data_line)
-           })
-
-           tpm_data_line = bind_rows(list_tpms_line)
-           #Apply ordering if it is specified
-           if(!is.null(order_vector)){
-             tpm_data_line$stage = factor(tpm_data_line$stage, 
-                                          levels = ordering(tpm_data_line$stage,
-                                                            order_vector))
-           }
-           # tpm_data_line$threshold = rep(threshold, nrow(tpm_data_line))
-
-           tpm_data_line
-
-           #Calculate the range for the y value to make it look nicer
-           if(max(
-             tpm_data_line$total_tpm_max
-           ) >= 500){
-             max_tpm <- c(
-               seq(
-                 0,
-                 mround(
-                   max(
-                     tpm_data_line$total_tpm_max
-                   ),
-                   500),
-                 length.out = 11)
-             )
-           }else{
-             max_tpm <- c(
-               seq(
-                 0,
-                 mround(
-                   max(
-                     tpm_data_line$total_tpm_max
-                   ),
-                   10),
-                 length.out = 11)
-             )
-           }
-           # max_tpm <- seq(
-           #       0,
-           #       mround(
-           #         max(
-           #           tpm_data_line$total_tpm_max),
-           #         500),
-           #       by = 500)
-           y = ggplot(data = tpm_data_line, aes(x = stage, y = total_tpm_median, 
-                                                fill = experiment, group = experiment)) +
-             geom_bar(position = position_dodge2(preserve = 'single'),
-                      stat = 'identity') +
-             scale_fill_manual(values = c(cbPalette, 'red', 'purple')[1:(length(unique(tpm_data_line$experiment)))]) +
-             geom_errorbar(aes(x = stage,
-                               ymax = total_tpm_max, 
-                               ymin = total_tpm_min),
-                           position = position_dodge2(preserve = 'single')) +
-             scale_y_continuous(breaks = max_tpm) +
-             theme_classic() +
-             theme(axis.title.x=element_blank(),
-                   axis.text.x=element_text(angle = 90, 
-                                            size = 10, 
-                                            vjust = 0.5),
-                   axis.text.y = element_text(size = 5),
-                   axis.title.y = element_text(size = 8),
-                   legend.title = element_text(size = 7),
-                   legend.text = element_text(size = 5),
-                   legend.key.size = unit(3,'mm'),
-                   legend.key.width = unit(2,'mm')) +
-             scale_size(range = c(0,3)) +
-             ylab(
-               bquote('total '~italic(.(list_name))~' TPM')
-             )
-           if(save == TRUE){
-             ggsave2(paste0('tl20-05_',
-                            gsub(list_name, 
-                                 pattern  = ' ', 
-                                 replacement = '_'),
-                            '_totalTPM_barchart.pdf'),
-                     plot = y,
-                     dpi = 300,
-                     width = 15,
-                     height = 10,
-                     units = 'cm')
-           }
-           y
-         })
-}
-```
-
-```{r function_total_tpm_expression_stacked}
-total_tpm_expression_stacked <- function(
-  list_genes, 
-  save = FALSE,
-  tpm_data = twentyfour_ct,
-  stage_str_extract_regex_pattern = '(?<=[[:digit:]]h_).+(?=_[[:digit:]]+)|sch.culture',
-  stack_genes = cir_id,
-  stack_names = c('Yes',
-                  'No'),
-  stack_label = 'Cir genes',
-  stack_grouping = TRUE
-){
-  lapply(names(list_genes), 
-         function(list_name){
-           paste(list_name)
-           gene_id <- getElement(list_genes, paste(list_name))
-           list_tpms_line <- lapply(tpm_data, 
-                                    function(input){
-                                      tpm_data <- input$tpm %>% 
-                                        filter(Geneid %in% gene_id) %>% 
-                                        melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-                                        mutate(stage = str_extract(sample, 
-                                                                   pattern = paste(stage_str_extract_regex_pattern)),
-                                               # stack_group = ifelse(
-                                               #   Geneid %in% stack_genes, #If the gene is in stack_genes
-                                               #   ifelse(
-                                               #     stack_grouping, #If stack_grouping is TRUE, use the group name from stack_names (the first element), otherwise use the gene name (useful is stack_genes is big e.g.
-cir_id)
-                                               #     stack_names[1],
-                                               #     Geneid
-                                               #   ),
-                                               #   stack_names[2]) #stack_names[2] is the negative name e.g. 'No'
-                                               stack_group = case_when(
-                                                 Geneid %in% stack_genes & stack_grouping ~ stack_names[1],
-                                                 Geneid %in% stack_genes & !stack_grouping ~ Geneid,
-                                                 TRUE ~ stack_names[2]
-                                               )
-                                        ) %>% 
-                                        group_by(stage, sample, stack_group) %>% 
-                                        dplyr::summarise(total_tpm = sum(tpm))
-
-                                      #Make stages with no cirs in the sample give a zero 
-                                      if(!all(colnames(input$tpm) %in% tpm_data$sample)){
-                                        missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
-                                        missing_stages_data_zero = data.frame(lapply(missing_stages, 
-                                                                                     function(missing_st){
-                                                              data.frame(stage = unlist(str_extract(missing_st, 
-                                                                                                    pattern = paste(stage_str_extract_regex_pattern))),
-                                                                         sample = missing_st,
-                                                                         total_tpm = 0)
-                                                            })
-               )
-               tpm_data = bind_rows(tpm_data, missing_stages_data_zero)
-             }
-
-             #Add experiment column
-             tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
-             max_tpm = if(max(tpm_data$total_tpm) < 10){
-               c(seq(0,mround(max(tpm_data$total_tpm),1),1))
-             }else{
-               c(seq(0,mround(max(tpm_data$total_tpm),5),5))
-             }
-
-             #Get the average data for the line graph
-             tpm_data_line = tpm_data %>% 
-               ungroup %>% 
-               group_by(stage, experiment, stack_group) %>% 
-               dplyr::summarise(total_tpm_median = median(total_tpm),
-                                Number_of_replicates = n(),
-                                total_tpm_max = max(total_tpm),
-                                total_tpm_min = min(total_tpm)) %>% 
-               arrange(stage)
-
-             return(tpm_data_line)
-           })
-
-           tpm_data_line <- bind_rows(list_tpms_line)
-           #Apply ordering if it is specified
-           if(!is.null(order_vector)){
-             tpm_data_line$stage = factor(tpm_data_line$stage, 
-                                          levels = ordering(tpm_data_line$stage,
-                                                            order_vector))
-           }
-
-           tpm_data_line
-
-           #Calculate the range for the y value to make it look nicer
-           if(max(
-             tpm_data_line$total_tpm_max
-           ) >= 500){
-             max_tpm <- c(
-               seq(
-                 0,
-                 mround(
-                   max(
-                     tpm_data_line$total_tpm_max
-                   ),
-                   500),
-                 length.out = 11)
-             )
-           }else{
-             max_tpm <- c(
-               seq(
-                 0,
-                 mround(
-                   max(
-                     tpm_data_line$total_tpm_max
-                   ),
-                   10),
-                 length.out = 11)
-             )
-           }
-           # max_tpm <- seq(
-           #       0,
-           #       mround(
-           #         max(
-           #           tpm_data_line$total_tpm_max),
-           #         500),
-           #       by = 500)
-           y <- ggplot(data = tpm_data_line, 
-                       aes(
-                         x = experiment,
-                         y = total_tpm_median, 
-                         fill = stack_group
-                       )
-           ) +
-             geom_bar(position = position_stack(),
-                      stat = 'identity') +
-             scale_fill_manual(
-               limits = unique(tpm_data_line$stack_group),
-               values = setNames(cbPalette[1:length(unique(tpm_data_line$stack_group))],
-                                 unique(tpm_data_line$stack_group))
-               #Need to have limits and values otherwise when there are only cir genes (i.e. only 'Yes') the colour 
-               # palette will mess up and only use one, wrong colour. Limits tells it there are always two values.
-             ) +
-             labs(fill = stack_label) +
-             # scale_y_continuous(breaks = max_tpm) +
-             theme_classic() +
-             theme(axis.title.x=element_blank(),
-                   axis.text.x=element_text(angle = 90, 
-                                            size = 10, 
-                                            vjust = 0.5),
-                   axis.text.y = element_text(size = 5),
-                   axis.title.y = element_text(size = 8),
-                   legend.title = element_text(size = 7),
-                   legend.text = element_text(size = 5),
-                   legend.key.size = unit(3,'mm'),
-                   legend.key.width = unit(2,'mm')) +
-             scale_size(range = c(0,3)) +
-             ylab(
-               bquote('total '~italic(.(list_name))~' TPM')
-             ) + 
-             facet_grid(.~stage) +
-             theme(strip.text.x = element_text(size=4, face = 'bold', angle = 45),
-                   strip.background = element_rect(colour="black", fill="lightgrey"))
-           if(save == TRUE){
-             ggsave2(paste0('tl20-08_',
-                            gsub(list_name, 
-                                 pattern  = ' ', 
-                                 replacement = '_'),
-                            '_totalTPM_barchart.pdf'),
-                     plot = y,
-                     dpi = 300,
-                     width = 15,
-                     height = 10,
-                     units = 'cm')
-           }
-           y
-         })
-}
 ```
@@
 
-cir_info = chabaudi_gene_info %>% 
+cir_info <- chabaudi_gene_info %>% 
   select(Gene.ID, 
@@
   filter(Gene.ID %in% c(reid_cir_names)) %>% 
-  mutate(Geneid = Gene.ID)
+  mutate(
+    Geneid = Gene.ID,
+    chrom = str_extract(Gene.ID, pattern = '(?<=PCHAS_)[[:digit:]]{2}'),
+    subfam = reid_cir_clade$clade[match(Geneid, reid_cir_clade$Geneid)]
+  ) %>% 
+  mutate(SorL = str_extract(subfam, pattern = 'L|S|ancestral|unk'))
 cir_transcript_lengths = select(cir_info, -Product.Description, -Gene.ID)
@@
 
+pir_clade <- reid_cir_clade %>% 
+  mutate(SorL = str_extract(clade, pattern = 'S|L|anc|unk'))
+
 #How many in each clade?
@@
 ```{r chapl_info}
-cir_chapl_aapl_info <- read_xlsx('Brugat2017_pirexpressiondata_41564_2017_BFnmicrobiol2016276_MOESM25_ESM.xlsx',
+cir_loci <- read_xlsx('Brugat2017_pirexpressiondata_41564_2017_BFnmicrobiol2016276_MOESM25_ESM.xlsx',
                             sheet = 1) %>% 
-  filter(id %in% cir_id) %>%
+  filter(id %in% cir_id)
+
+cir_chapl_aapl_info <- cir_loci %>%
   mutate(
@@
 
-cir_chapl_aapl_info %>% 
-  mutate(chrom = str_extract(Geneid, pattern = '(?<=PCHAS_)[:digit:]{2}'),
-         locus_type = ifelse(is.na(locus_type), 'None', locus_type)) %>% 
-  group_by(chrom, 
-           locus_type) %>% 
-  summarise(num_subfam = n_distinct(Geneid))
-```
-
-```{r gene_info, message = FALSE, results='hide'}
-
-#No redundancy
-two.transcript <- chabaudi_gene_info %>% 
-  dplyr::count(Geneid) %>% 
-  filter(n > 1)
-
-```
-
-```{r handier_descriptions}
-chabaudi_gene_info$is_putative <- ifelse(
-  str_detect(
-    chabaudi_gene_info$Description,
-    pattern = 'putative'
-  ),
-  'Yes',
-  'No'
-)
-
-chabaudi_gene_info$handy_descriptions <- case_when( 
-  #Handy function for avoiding nested, messy ifelse statements
-  str_detect(
-    chabaudi_gene_info$Description, 
-    pattern = 'early transcribed membrane protein'
-  ) ~ 
-    'ETRAMP',
-  str_detect(
-  chabaudi_gene_info$Description, 
-    pattern = 'PIR protein'
-  ) ~ 
-    'PIR',
-  str_detect(
-    chabaudi_gene_info$Description, 
-    pattern = 'hypothetical'
-  ) ~ 
-    'hypothetical',
-  str_detect(
-    chabaudi_gene_info$Description, 
-    pattern = '(conserved Plasmodium)|(conserved protein, unknown function)'
-  ) ~ 
-    'cons.',
-  TRUE ~ #This applies the case to all remaining
-    str_extract(  
-      str_remove_all( 
-        #Remove all applies across all the patterns instead of just one.
-        chabaudi_gene_info$Description, 
-        pattern = '( protein)|( protein, )|(, pseudogene)|(, putative)' 
-        #Remove the words protein and spaces for the word.
-      ),
-      pattern = '.{1,10}' 
-      #Take the first 10 characters max.
-    )
-)
-
-chabaudi_gene_info$handy_descriptions <- ifelse(
-  grepl(chabaudi_gene_info$Description, pattern = 'pseudogene'),
-  paste0(chabaudi_gene_info$handy_descriptions,
-         '_ps'),
-  paste0(chabaudi_gene_info$handy_descriptions)
-)
-
-chabaudi_gene_info$handy_descriptions <- uniquify(
-  ifelse(
-    chabaudi_gene_info$is_putative == 'Yes',
-    paste0(chabaudi_gene_info$handy_descriptions, '_put'),
-    paste0(chabaudi_gene_info$handy_descriptions)
-  ),
-  first_one = TRUE
-)
-
-#Annotations for individual pirs based on Frech and Chen 2013, PlasmoDB
-chabaudi_gene_info$handy_descriptions[which(chabaudi_gene_info$Geneid == 'PCHAS_0101200')] <- 'PIR_highly_cons.'
-# chabaudi_gene_info$handy_descriptions[which(chabaudi_gene_info$Geneid == 'PCHAS_0524800')] <- 'PIR_rodent_cons.'
-chabaudi_gene_info$handy_descriptions[which(chabaudi_gene_info$Geneid == 'PCHAS_1100300')] <- 'PIR_SBP_high'
-```
-
-```{r function_featurecounts_to_table, results = 'hide', message = FALSE}
-
-featurecounts_to_table = function(vector_of_files = NULL, 
-                                  read_skip = 0){
-
-  files_exp <- vector_of_files
-  # print(files_exp)
-  #Combine the data itself into one dataframe
-  exp = lapply(files_exp, 
-               read.delim, 
-               skip = read_skip) %>% 
-    purrr::reduce(left_join, 
-                  by = 'Geneid')
-  #Using purrr and left_join means that Geneid will have to match between the samples so I removed the test for this.
-
-  return(exp)
-}
-```
-
-```{r function_counttable_to_tpm}
-
-counttable_to_tpm = function(count_table,
-                             table_names,
-                             transcript_lengths = chabaudi_transcript_lengths,
-                             technical_reps = FALSE,
-                             experiment_name = NULL){
-
-  #Before assigning stage numbers we need to merge the technical replicates
-
-  if(technical_reps == TRUE){
-
-    if(all(!grepl(colnames(table_names), pattern = 'Technical_Replicates'))){
-      stop('If technical_reps is TRUE then the table_names data frame must have a Technical_Replicates column')
-    }
-
-    geneid_col = count_table[,'Geneid', drop = FALSE]
-    list_techreps = lapply(unique(table_names$Technical_Replicates), function(techrep){
-      print(techrep)
-      #For each technical replicate we need the mean of the counts
-      # So get the count table for these replicates - Alias names haven't been assigned yet so names should be runs
-      techrep_runs = filter(table_names, Technical_Replicates == paste(techrep)) %>% 
-        `$`(Run)
-      #Extract the counts and calculate the mean
-      techrep_counts =  select(count_table, paste(techrep_runs)) %>% 
-        apply(1, mean) %>% data.frame()
-      colnames(techrep_counts) = paste(techrep)
-      techrep_counts
-    })
-
-    exp = bind_cols(geneid_col, list_techreps)
-    colnames(exp)[-1] = table_names$Alias[match(colnames(exp)[-1], table_names$Technical_Replicates)]
-
-  }else{
-
-    exp = count_table
-    colnames(exp)[-1] = as.character(table_names$Alias[match(colnames(exp)[-1], table_names$Run)])
-
-  }
-
-  #Need a number for each sample based on the number of distinct samples for each stage
-  #Also going to remove NA columns here as they are things we don't want.
-
-  colnames(exp) <- ifelse(!is.na(colnames(exp)), 
-                          uniquify(colnames(exp)), 
-                          NA)
-  exp <- exp[,!is.na(colnames(exp)),drop = FALSE]
-
-  # print(colnames(exp))
-  exp = exp %>% arrange(Geneid)
-
-  #Calculate the tpm
-  tpm_exp = tpm(exp, transcript_lengths)
-
-  list_exp =  list(counts = exp, tpm = tpm_exp, experiment = experiment_name)
-  return(list_exp)
-}
-```
-
-# Downloading the count files
-
-This part is long not because downloading the count files is tricky but because the names of the different samples ought to be standardised as much as possible to be helpful down the line.
-
-Count data was generated on the Crick HPC from the fq.gz files between 
-using Nextflow:
-
-Using PlasmoDB v45 _P. chabaudi chabaudi_ genome.
-
-Here I have written a function which can help with re-naming columns and removing those you don't need if the different stages expected and the vector of files/list of objects containing the counts are specified.
-The function returns a list containing the count and the TPM files.
-
-The design table from online is missing a file ('21659_8#24') for some reason. This has been designated as 'MT_Unknown' until we can find out which one it is. I've actually decided to remove this.
-
-```{r download_files}
+cir_info <- mutate(cir_info, 
+                   subtelomere = cir_loci$Subtelomere[match(Geneid, cir_loci$id)],
+                   chapl_aapl = cir_loci$`Locus type`[match(Geneid, cir_loci$id)],
+                   chapl_aapl_name = cir_chapl_aapl_info$chapl_aapl[match(Geneid, cir_chapl_aapl_info$Geneid)])
+
+```
+
+```{r writing_cir_info_file}
+
+write_delim(cir_info,
+            path = 'tl19-07_dataset_pirs_chabaudi_info.txt',
+            delim = '\t')
+
+```
+
+```{r times_stages_labels}
 #Use the rMT stage differential count in order to name the samples, but keep time to retain order
@@
                                                          replacement = '\\.')
-
-design_table <- read.delim('samples_lanes.list',
-                           sep = '\t',
-                           header = FALSE,
-                           col.names = c('sample', 'transmission', 'sequence'),
-                           stringsAsFactors = FALSE) %>% 
-  #regex extract for the numbers (one or more) before am, pm or h, or just an 'sch' for schizont
-  mutate(time = str_extract(sample, pattern = '[:digit:]+([ap]m|h)|sch')) %>% 
-  #Convert the times given in 'h' to robustly coded times in 24h clock
-  mutate(time = ifelse(grepl(time, pattern = '[[:digit:]]h$'), 
-                       strptime(time, format="%Hh") %>% format(., "%Hh"),
-                       time)) %>% 
-  mutate(time = ifelse(grepl(time, pattern = '[[:digit:]](a|p)m'), 
-                       strptime(time, format="%I%p") %>% format(., "%Hh"),
-                       time)) %>% 
-  mutate(Alias = paste0(transmission, '_', time)) %>% 
-  #Paste on the stage from the diff count, or nothing if NA
-  mutate(Alias = ifelse(is.na(match(time,
-                                    times_stages_table$time)),
-                        Alias,
-                        paste0(Alias, '_', 
-                               times_stages_table$stage_diff_count[match(time,
-                                                                         times_stages_table$time)]))) %>% 
-  mutate(Alias = gsub(Alias, pattern = 'MT', replacement = 'rMT')) %>% 
-  mutate(Alias = ifelse(grepl(Alias, pattern = 'sch'), paste0(Alias, '.culture'), Alias)) %>% 
-  #To make the names R compatible
-  mutate(Alias = make.names(Alias)) %>% 
-  mutate(sequence = gsub(sequence, 
-                         pattern = '9478_[[:digit:]]', 
-                         replacement = '9478_cat'))
-
-#Get vector of the files of MT and merged SBP
-# the merged SBP files are not separated into _3 and _4
-
-unknown_sample <- data.frame(sample = 'PcUnknown', 
-                             transmission = 'rMT',
-                             time = 'Unknown',
-                             sequence = '21659_8#24',
-                             Alias = 'rMT_Unknown')
-design_table <- bind_rows(design_table, unknown_sample)
-
-# design_table %>% 
-#   group_by(Alias) %>% 
-#   dplyr::summarise(n())
-
-# design_table %>% 
-#   group_by(transmission) %>% 
-#   dplyr::summarise(n())
-
-#Old files
-# vector_of_files <- list.files(pattern = 'txt$') %>% str_subset(pattern = '21659|9478_#')
-
-#New files - methods needs updating
-vector_of_files <- c(
-  list.files('nextflow_working_dir_PlasmoDBv48_rna-seq-r1.4.2_dir/featureCounts/gene_counts',
-             pattern = 'txt$',
-             full.names = TRUE)
-)
-
-# length(vector_of_files)
-# length(tmp)
-
-# str_extract(tmp, pattern = '(21659|9478)_([[:digit:]]+|cat|)#[[:digit:]]+')
-# str_extract(vector_of_files, pattern = '(21659|9478)_([[:digit:]]+|cat|)#[[:digit:]]+')
-
-# tmp[tmp %in% vector_of_files]
-
-twentyfour_counts <- featurecounts_to_table(vector_of_files = vector_of_files,
-                                            read_skip = 1)
-
-#The names of the files / Runs is now a bit longer thanks to my method of naming files
-design_table$Run <- unlist(
-  sapply(make.names(design_table$sequence), 
-                            function(nam){str_subset(colnames(twentyfour_counts)[-1], 
-                                                     pattern = paste0(paste(nam), '_'))})
-  ) %>% 
-  make.names() #Make names acts during read.delim in the next function
-
-design_table$unique_alias <- uniquify(design_table$Alias, first_one = TRUE)
-# disagree(list(colnames(twentyfour_counts), design_table$Run))
-```
-
-```{r}
-#Getting a nice count table for the paper and Github
-chab_count_table <- select(twentyfour_counts, 
-       matches('Geneid|bam')) %>% 
-  rename_with(function(colnam){
-    design_table$unique_alias[match(colnam, design_table$Run)]
-  },
-  matches('bam')) %>% 
-  select(-matches('sch|Unk|SBP')) %>% 
-  .[sort(colnames(.))] %>% 
-  arrange(Geneid)
-
-write_xlsx(chab_count_table,
-           'Manuscript_SuppInfo_chabaudi_24h_counts.xlsx')
-write.csv(chab_count_table,
-           'Manuscript_SuppInfo_chabaudi_24h_counts.csv')
-
-# tpm(chab_count_table, 
-#     lengths = chabaudi_transcript_lengths)
-
-```
-
-New list of sequence files has 9478_(3 or 4)_#(Number) while the old files just had 9478_#(Number).
-This appears to be the cause of discrepency.
-Why are there these extra files??
-I wonder if they are paired files which weren't analysed together?
-- Issue was because I hadn't merged technical replicates.
-
-```{r calculate_save_TPM}
-# #Calculate and save TPM
-# twentyfour_ct <- counttable_to_tpm(count_table = twentyfour_counts,
-#                                   table_names = design_table,
-#                                   experiment_name = NULL)
-# # There's an issue with bind_cols in this function at the moment
-# #
-# # #saving this output
-# write_csv(twentyfour_ct$tpm,
-#           'tl19-07_twentyfour_tpm_samples.csv')
-# write_csv(twentyfour_ct$counts,
-#           'tl19-07_twentyfour_counts_samples.csv')
-# 
-# #Getting all the sample columns in order.
-# twentyfour_tpm_sbp <- twentyfour_ct$tpm %>%
-#   # select(-rMT_Unknown) %>%
-#   select_at(vars(matches('Geneid|SBP_')))
-# 
-# twentyfour_tpm_sbp <- select(
-#   twentyfour_tpm_sbp,
-#   Geneid,
-#   ordering(
-#     colnames(twentyfour_tpm_sbp)[-1],
-#     order_vector = order_vector
-#   )
-# )
-# # Saving a tab-delimited file
-# write_delim(twentyfour_tpm_sbp,
-#             'tl19-07_twentyfour_tpm_samples_SBP.txt',
-#             delim = '\t')
-# 
-# twentyfour_tpm_rMT <- twentyfour_ct$tpm %>%
-#   select(-rMT_Unknown) %>%
-#   select_at(vars(matches('Geneid|rMT_')))
-# 
-# twentyfour_tpm_rMT <- select(
-#   twentyfour_tpm_rMT,
-#   Geneid,
-#   ordering(
-#     colnames(twentyfour_tpm_rMT)[-1],
-#     order_vector = order_vector
-#   )
-# )
-# # Saving a tab-delimited file
-# write_delim(twentyfour_tpm_rMT,
-#             'tl19-07_twentyfour_tpm_samples_rMT.txt',
-#             delim = '\t')
-
-```
-
-```{r read_tpm}
-
-#Read TPM - no unknown - combined data frame
-twentyfour_ct_comb <- list(
-  tpm = read_csv('tl19-07_twentyfour_tpm_samples.csv') %>% 
-    select(-rMT_Unknown),
-  counts = read_csv('tl19-07_twentyfour_counts_samples.csv') %>% 
-    select(-rMT_Unknown)
-)
-
-#Save counts
-
-write_xlsx(twentyfour_ct_comb$counts[,sort(colnames(twentyfour_ct_comb$counts))] %>% 
-             dplyr::select(!matches('sch')),
-           'Manuscript_Frontiers_chabaudi_24h_counts.xlsx')
-
-#Read TPM separately - no unknown
-twentyfour_ct_rMT <- list(
-  tpm = read_csv('tl19-07_twentyfour_tpm_samples.csv') %>% 
-    select(-rMT_Unknown) %>%
-    select_at(vars(matches('Geneid|rMT_'))),
-  counts = read_csv('tl19-07_twentyfour_counts_samples.csv') %>% 
-    select(-rMT_Unknown) %>%
-    select_at(vars(matches('Geneid|rMT_'))),
-  experiment = 'rMT')
-
-twentyfour_ct_SBP <- list(
-  tpm = read_csv('tl19-07_twentyfour_tpm_samples.csv') %>% 
-                        select(-rMT_Unknown) %>%
-                        select_at(vars(matches('Geneid|SBP_'))),
-  counts = read_csv('tl19-07_twentyfour_counts_samples.csv') %>% 
-                        select(-rMT_Unknown) %>%
-                        select_at(vars(matches('Geneid|SBP_'))),
-  experiment = 'SBP')
-
-twentyfour_ct <- list(twentyfour_ct_rMT, 
-                      twentyfour_ct_SBP)
-
-#Removing the rMT and SBP prefix for the experiments now that they're separated by transmission
-twentyfour_ct <- lapply(twentyfour_ct, 
-                        function(list_ct){
-                          lapply(list_ct, 
-                                 function(ct){
-                                   if(!is.null(colnames(ct))){
-                                     colnames(ct) <- str_remove(colnames(ct),
-                                                                pattern = '(rMT|SBP)_')
-                                   }
-                                   return(ct)
-                                 })
-                        })
-
-```
-
-```{r counts_v_features}
-num_counts <- apply(twentyfour_ct_comb$counts[,-1], 2, sum)
-num_features <- apply(twentyfour_ct_comb$counts[,-1], 2, function(x) sum(x > 0))
-plot(num_counts, num_features 
-     #, ylim = c(0, max(num_features))
-     )
-```
-
-```{r sanity_test}
-
-#sanity test
-# twentyfour_ct$tpm %>% select(Geneid, 'SBP_17h_3') %>% .[450:455,]
-# # read.delim('samples_lanes.list',
-# #                            sep = '\t',
-# #                            header = FALSE,
-# #                            col.names = c('sample', 'transmission', 'sequence'),
-# #                            stringsAsFactors = FALSE)
-# read.delim(header = TRUE, sep = '\t', file = list.files(pattern = '9478_#6')[1], skip = 1) %>% 
-#   arrange(Geneid) %>% 
-#   select(Geneid, X9478_.6_hisat2alignment_v45chabaudigenome.sorted.merged.bam) %>% .[450:455,]
-# 
-# twentyfour_ct$counts %>% select(Geneid, 'MT_05h_3') %>% .[1284:1289,]
-# read.delim(header = TRUE, sep = '\t', file = list.files(pattern = '21659_8#3')[1], skip = 1) %>% 
-#   arrange(Geneid) %>% 
-#   select(Geneid, X21659_8.3_hisat2alignment_v45chabaudigenome.sam) %>% .[1284:1289,]
-
-```
-
-# Hierarchical clustering of the samples
-
-```{r hclusts}
-hclusts = function(tpms){
-  dists = dist(t(tpms$tpm[,-1]))
-  plot(hclust(dists), 
-       cex = 0.7, 
-       main = tpms$experiment)
-}
-
-invisible(
-  lapply(twentyfour_ct, hclusts)
-)
-# hclusts(twentyfour_ct_comb)
-```
-
-Using scale() to normalise the data variance and then hclust with 'average' method. The rMT samples and SBP samples still cluster separately.
-
-```{r hclusts_scaled_data}
-hclusts_scaled = function(tpm){
-  dists = dist(t(scale(tpm$tpm[,-1])))
-  plot(hclust(dists, method = 'average'), cex = 0.7)
-}
-hclusts_scaled(twentyfour_ct_comb)
-```
-
-Both here and in the correlation plots below most of the samples (except 11 and 14h) are distinct based not on time of day but be method of transmission.
-This isn't what one would expect.
-It of course could be due to batch effects since the SBP and MT experiments were done by different people in different years.
-
-hclust without that one naughty pseudogene.
-This one pseudogene, PCHAS_0420800, is a major defining gene of the rMT samples, and removal of this gene makes the samples cluster as we'd expect; by asexual stage and not by transmission
-
-```{r hclust_nopseudo}
-tpm_nopseudo <- twentyfour_ct_comb$tpm %>% filter(!Geneid %in% 'PCHAS_0420800')
-
-# sapply(tpm_nopseudo[,-1], sum)[rev(order(sapply(tpm_nopseudo[,-1], sum)))]
-# Need to rescale the TPM
-
-tpm_nopseudo_rescale <- lapply(colnames(tpm_nopseudo), 
-                               function(column_name){
-                                 column <- getElement(tpm_nopseudo, paste(column_name))
-                                 if(is.numeric(column)){
-                                   #is.numeric to avoid trying to scale the Geneid column
-                                   tpm_scale_factor <- 1e6 / sum(column)
-                                   tpm_scaled <- as.numeric(column*tpm_scale_factor)
-                                 }else{
-                                   tpm_scaled <- column
-                                 }
-                                 return(tpm_scaled)
-                               }) %>% 
-  bind_cols
-
-colnames(tpm_nopseudo_rescale) <- colnames(tpm_nopseudo)
-# sapply(tpm_nopseudo_rescale[,-1], sum)[rev(order(sapply(tpm_nopseudo_rescale[,-1], sum)))]
-
-#Note now the MT samples can be distinguished by having a lower total TPM
-dists = dist(t(tpm_nopseudo_rescale[,-1]))
-plot(hclust(dists), cex = 0.7)
-```
-
-# Density plots
-
-These graphs show the distribution of TPM values for each gene in each sample.
-The shapes of the curves follow the generally expected pattern of one smaller peak at zero and another peak and higher values with a long tail.
-Individual samples however show a lot of variation from one another presumably because of differences in the amount of sequences detected by the experiments.
-
-```{r density_plots, fig.show="hold", out.width="50%"}
-#fig.show and out.width settings for having both plots on same output.
-list_min = lapply(twentyfour_ct, function(input){
-  input_tpm = input$tpm
-  minima = sapply(colnames(select(input_tpm, -Geneid)), function(colnam){
-    #Want to exclude the weird sporozoite densities
-    if(grepl(colnam, pattern = 'sporozoite', ignore.case = TRUE)){
-      NA
-    }else{
-      tpms = select(input_tpm, paste(colnam)) %>% unlist %>% as.numeric()
-      den = density(log1p(tpms))
-      #Get the difference between adjacent y values, find the positive change, first one coresponds to the first minimum
-      first_min = den$x[which(diff(sign(diff(den$y)))==2)] %>% .[1]
-    }
-  }) %>% expm1()
-  #expm1 is the reverse of log1p - almost forgot to take away the one!
-  list(experiment = input$experiment,
-       range = range(minima, na.rm = TRUE), 
-       mean = mean(minima, na.rm = TRUE),
-       median = median(minima, na.rm = TRUE))
-})
-
-min_median = median(unlist(lapply(list_min, function(x) x$median)))
-min_min = min(unlist(lapply(list_min, function(x) x$range)))
-min_max = max(unlist(lapply(list_min, function(x) x$range)))
-
-threshold <- min_median
-
-# print(min_median)
-
-list_bw = lapply(twentyfour_ct, function(expt){ 
-  lapply(colnames(expt$tpm)[-1], function(sample){
-    data = as.numeric(unlist(select(expt$tpm, paste(sample))))
-    dens = density(log1p(data))
-    dens$bw
-})
-});
-bw_mean = mean(unlist(list_bw))
-bw_range = max(unlist(unlist(list_bw))) - min(unlist(unlist(list_bw)))
-# print(paste('Bandwidth range = ', bw_range))
-
-invisible(
-  lapply(twentyfour_ct, 
-       density_samples, 
-       save = FALSE)
-)
-#Wrapping in Invisible to hide the console output of NULL (I don't know where it comes from...)
-```
-
-Based on these density plots we make the assumption of a threshold of expression as being `r round(min_median, digits = 2)` TPM.
-
-# Top general and _cir_ genes in each stage
-
-```{r top_genes}
-top_n_genes = function(tpm_input,
-                       threshold = 0,
-                       stage_str_extract_regex_pattern = '(?<=_).+(?=_[[:digit:]]+)|sch.culture'){
-  tpm = tpm_input$tpm %>% 
-    #Melt the data frame for using dplyr
-    melt(id.vars = 'Geneid', 
-         variable.name = 'sample', 
-         value.name = 'tpm') %>% 
-    mutate(stage = str_extract(sample, 
-                               pattern = paste(stage_str_extract_regex_pattern))) %>% 
-    group_by(Geneid, stage) %>%  
-    dplyr::summarise(median_tpm = median(tpm),
-                     mean_tpm = mean(tpm), 
-                     min_tpm = min(tpm), 
-                     max_tpm = max(tpm), 
-                     sd_tpm = sd(tpm)) %>% 
-    group_by(stage) %>% 
-    mutate(experiment = tpm_input$experiment)
-
-  list_data = list('tpm' = tpm,
-                   'experiment' = tpm_input$experiment)
-  return(list_data)
-}
-
-### Generate and save the top gene data
-
-# twentyfour_top <- lapply(twentyfour_ct, top_n_genes)
-# 
-# lapply(twentyfour_top, function(transmission_top){
-#   tpm_data <- transmission_top$tpm
-#   write_csv(tpm_data,
-#             paste0('tl19-07_twentyfour_tpm_samplescollated_longform_',
-#                    transmission_top$experiment,
-#                    '.csv'))
-# })
-# 
-# twentyfour_top_comb <- top_n_genes(twentyfour_ct_comb,
-#                                    stage_str_extract_regex_pattern = '.+(?=_[[:digit:]]+)')
-# 
-# write_csv(twentyfour_top_comb$tpm,
-#           'tl19-07_twentyfour_tpm_samplescollated_longform.csv')
-
-###
-
-### Read in the top gene data
-
-twentyfour_top <- lapply(list.files(pattern = 'tl19-07_twentyfour_tpm_samplescollated_longform_', 
-                  full.names = TRUE), 
-       function(top_files){
-         list(
-           tpm = read_csv(top_files),
-              experiment = str_extract(top_files, pattern = 'rMT|SBP')
-         )
-       })
-
-twentyfour_top_comb <- list(tpm = read_csv('tl19-07_twentyfour_tpm_samplescollated_longform.csv'))
-
-###
-```
-
-```{r kable_table_top_n_genes, results = "asis"}
-
-top_n_genes_kable_table <- function(top_n_tpm,
-                                    n_top = 5,
-                                    subset_genes = NULL,
-                                    subset_genes_title = 'all',
-                                    gene_info = chabaudi_gene_info,
-                                    highlight_genes_regex = 'PCHAS_0101200'){
-  # top_n_tpm <- twentyfour_top[[1]]
-
-  tpm <- top_n_tpm$tpm
-
-  #Calculating the index used to make the kable table
-
-  kable_index = rep(n_top, 
-                    times = length(unique(tpm$stage)))
-  names(kable_index) = unique(tpm$stage)
-
-  if(is.null(subset_genes)){
-    top_genes <- tpm
-  }else{
-    top_genes <- filter(tpm, Geneid %in% subset_genes)
-  }
-
-  top_genes = top_genes %>% 
-    group_by(stage) %>% 
-    top_n(wt = median_tpm, 
-          n = n_top) %>% 
-    dplyr::arrange(stage, 
-                   desc(median_tpm)) %>% 
-    mutate_if(is.numeric, 
-              function(x){round(x, 
-                                digits = 2)})
-
-  top_genes$description = gene_info$handy_descriptions[match(top_genes$Geneid, 
-                                                             gene_info$Geneid)]
-
-  top_genes_kable <- top_genes %>% 
-    mutate(median_tpm = cell_spec(median_tpm, "html", 
-                                  color = ifelse(median_tpm > 10, "red", "blue"))) %>% 
-    mutate(Geneid = as.character(Geneid)) %>% 
-    mutate(Geneid = ifelse(grepl(Geneid, pattern = paste(highlight_genes_regex)),
-                           cell_spec(Geneid, bold = TRUE),
-                           Geneid)) %>% 
-    kable(format = 'html', 
-          caption = paste0(top_n_tpm$experiment, 
-                           ' - Top ', 
-                           paste(n_top), 
-                           ' expressed ',
-                           paste(subset_genes_title),
-                           ' genes'), 
-          escape = FALSE) %>%
-    kable_styling() %>%
-    pack_rows(index = kable_index) %>%
-    scroll_box(width = "100%", height = "500px")
-
-  return(top_genes_kable) 
-}
-
-lapply(twentyfour_top, 
-       top_n_genes_kable_table)
-lapply(twentyfour_top,
-       top_n_genes_kable_table,
-       subset_genes = cir_id, 
-       subset_genes_title = 'cir')
-
-```
-
-~~At the moment these tables are not showing up in this document...~~
-
-Wow! The SBP cir PCHAS_1100300 isn't expressed higher than ancestral or rodent-conserved pir-like in these SBP samples...
-
-For all the other genes again in MT one of the highest expressed genes is a pseudogene while in SBP it is a series of ETRAMPs (although MT expresses these also it seems).
-
-```{r most_variable_genes, eval = FALSE}
-twentyfour_top$tpm %>%
-  top_n(5, sd_tpm) %>%
-  arrange(stage, desc(sd_tpm))
-```
-
-* _PCHAS_0101200_ = ancestral cir.
-
-* _PCHAS_0524800_ = rodent conserved cir-like genes.
-
-* _PCHAS_1100300_ = SBP-dominant cir gene (Spence et al, 2013).
-
-```{r main-three-pir_heatmap}
-# pdf('tl19-07_cirgenes_24h_heatmap.pdf', width = 10, height = 5)
-draw_heatmap_chabaudi_24h(datasets =twentyfour_ct, 
-                          genelist = c('PCHAS_0101200',
-                                       # 'PCHAS_0524800',
-                                       'PCHAS_1100300'),
-                          scale = 'max',
-                          threshold = 0,
-                          log_it = TRUE)
-# dev.off()
-twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% c('PCHAS_0101200',
-                       # 'PCHAS_0524800',
-                       'PCHAS_1100300')) %>% 
-  kable() %>% 
-  kable_styling() %>% 
-  scroll_box(width = "100%")
-```
-
-```{r Deirdres CHAPLs, eval = FALSE}
-protein <- c('PCHAS_0400300',   
-             'PCHAS_0600600',    
-             'PCHAS_1200500',    
-             'PCHAS_1300101',
-             'PCHAS_0101200',   
-             'PCHAS_1100300',
-             'PCHAS_0626900')
-
-draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
-                          genelist = protein,
-                          scale = 'max',
-                          threshold = 0,
-                          log_it = TRUE,
-                          gene_names_for_title = 'Available protein')
-
-chapl <- c('PCHAS_0626200',
-           'PCHAS_0626700',
-           'PCHAS_0627100')
-
-draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
-                          genelist = chapl,
-                          scale = 'max',
-                          threshold = 0,
-                          log_it = FALSE,
-                          gene_names_for_title = 'Deirdre\'s ChAPLs')
-
-available_primers <- c('PCHAS_0600600',
-                       'PCHAS_1100300',
-                       'PCHAS_1301700',
-                       'PCHAS_0400300',
-                       'PCHAS_0301400',
-                       'PCHAS_1200500',
-                       'PCHAS_1300101')
-
-draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
-                          genelist = available_primers,
-                          scale = 'max',
-                          threshold = 0,
-                          log_it = FALSE,
-                          gene_names_for_title = 'Available primers')
-```
-
-Good _cir_ negative controls would be 0600600 and 1200500 for SBP and 1301700 for both (but especially MT).
-
-# Correlation plots {.tabset}
-
-## Correlation plot of collated stages (all genes)
-
-```{r corelation_plots_collated}
-# Collated across the stages for a more easily understood heatmap
-
-big_top_twentyfour = twentyfour_ct_comb$tpm %>% 
-  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  group_by(Geneid, stage) %>% 
-  dplyr::summarise(median_tpm = median(tpm)) %>% 
-  dcast(Geneid ~ stage)
-
-big_top_twentyfour_mat = big_top_twentyfour[,-1,drop=FALSE]
-
-# pdf(file = 'tl19-07_corrplot_24h_rnaseq_chabaudi.pdf')
-corrplot(cor(big_top_twentyfour_mat), method = 'color', order = 'hclust', cl.lim = c(-0,1), tl.cex = 0.7,
-         col = c(viridis(100), magma(115)), 
-         mar=c(0,0,1,0),
-         title = 'Correlation plot')
-# dev.off()
-
-```
-
-## Correlation plot of collated stages (only _cir_ genes)
-
-```{r}
-
-#SBP now all together, so the schizonts were pushing them elsewhere
-
-#Now just the cir genes
-
-big_top_twentyfour_cir_mat = big_top_twentyfour %>% 
-  filter(Geneid %in% cir_id) %>% 
-  select(-Geneid)
-corrplot(cor(big_top_twentyfour_cir_mat), method = 'color', order = 'hclust', 
-         cl.lim = c(0.0,1), tl.cex = 0.7,
-         col = c(viridis(100), magma(100)), 
-         mar=c(0,0,1,0),
-         title = 'Correlation plot (only cirs)')
-```
-
-## Correlation plot of collated stages, but without schizonts (all genes)
-
-```{r correlation_collated_nosch}
-
-#How does this order with the schizonts removed
-
-big_top_twentyfour_nosch = big_top_twentyfour %>% select(-rMT_sch.culture)
-
-big_top_twentyfour_nosch_mat = big_top_twentyfour_nosch[,-1,drop=FALSE]
-
-# pdf(file = 'tl19-07_corrplot_24h_rnaseq_chabaudi_noschz.pdf')
-corrplot(cor(big_top_twentyfour_nosch_mat), method = 'color', order = 'hclust', cl.lim = c(0,1), tl.cex = 0.7,
-         col = c(viridis(100), magma(115)), 
-         mar=c(0,0,1,0),
-         title = 'Correlation plot (no schizont samples)')
-# dev.off()
-```
-
-Now all the SBP and MT groups cluster together
-
-## Correlation plot of all samples (all genes)
-
-```{r correlation_plot}
-twentyfour_tpm_mat = twentyfour_ct_comb$tpm
-rownames(twentyfour_tpm_mat) = twentyfour_tpm_mat$Geneid
-twentyfour_tpm_mat = twentyfour_tpm_mat[,-1,drop=FALSE]
-
-#Replicates
-twentyfour_all_cor = cor(twentyfour_tpm_mat)
-corrplot(twentyfour_all_cor, tl.cex = 0.5, mar=c(0,0,1,0),
-         method = 'color', order = 'hclust', cl.lim = c(0,1), col = c(viridis(100), tl.cex = 0.7, magma(115)),
-         title = 'Correlation plot - individual samples')
-#MT and SBP separate, SBP 11/14/17h match MT schizonts
-```
-
-## Correlation plot of all samples (but only _cir_ genes)
-
-```{r cir_correlation}
-
-#Replicate cirs
-twentyfour_tpm_mat_cir <- filter(twentyfour_ct_comb$tpm, Geneid %in% cir_id) #This drops the rownames for some reason
-rownames(twentyfour_tpm_mat_cir) <- twentyfour_tpm_mat_cir$Geneid 
-twentyfour_tpm_mat_cir <- twentyfour_tpm_mat_cir[,-1,drop=FALSE]
-twentyfour_tpm_mat_cir_cor <- cor(twentyfour_tpm_mat_cir)
-corrplot(twentyfour_tpm_mat_cir_cor, tl.cex = 0.5, method = 'color', order = 'hclust', 
-         cl.lim = c(0,1), col = c(viridis(100), magma(115)),
-         mar=c(0,0,1,0),
-         title = 'Correlation plot - individual samples (only cirs)')
-```
-
-## Do the top 10 expressed genes in each sample give any extra clarity to the correlation plot?
-
-Drawing a correlation plot with only the top 10 expressed genes in the samples.
-
-Not really, I guess need to get the genes with the highest variance not total score.
-
-```{r top10genes}
-top10genes <- twentyfour_top_comb$tpm %>% 
-  dplyr::arrange(desc(median_tpm)) %>% 
-  group_by(stage) %>% 
-  top_n(median_tpm, n = 10) %>% 
-  .$Geneid %>% 
-  unique
-
-draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
-                          genelist = top10genes,
-                          scale = 'max',
-                          threshold = 0,
-                          log_it = TRUE)
-
-big_top_twentyfour_10_mat = big_top_twentyfour %>% 
-  filter(Geneid %in% cir_id) %>% 
-  select(-Geneid)
-corrplot(cor(big_top_twentyfour_10_mat), method = 'color', order = 'hclust', cl.lim = c(0.0,1), tl.cex = 0.7,
-         col = c(viridis(100), magma(100)))
-```
-
-## Do the top 10 most variable genes across the samples give any extra clarity to the correlation plot?
-
-The top ten most variable genes (highest standard deviation) between all the individual samples were identified.
-
-Did with and without the pseodogene _PCHAS_0420800_ which dominates the expression profile of MT and not SBP.
-
-```{r most_variable_genes_corr}
-
-variable_genes <- twentyfour_top_comb$tpm %>% 
-  ungroup %>% 
-  group_by(Geneid) %>% 
-  dplyr::summarise(med_tpm = median(median_tpm),
-                   std_tpm = sd(median_tpm)) %>% 
-  mutate(product_description = 
-           chabaudi_gene_info$Product.Description[match(Geneid, chabaudi_gene_info$Gene.ID)]) %>% 
-  arrange(desc(std_tpm))
-
-variable_gene_names <- variable_genes$Geneid[1:10]
-
-draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
-                          genelist = variable_gene_names,
-                          scale = 'max',
-                          threshold = 0,
-                          log_it = TRUE)
-
-big_top_twentyfour_mostvar_mat = big_top_twentyfour %>% 
-  filter(Geneid %in% variable_gene_names) %>% 
-  select(-Geneid)
-corrplot(cor(big_top_twentyfour_mostvar_mat), method = 'color', order = 'hclust', tl.cex = 0.7,
-         col = c(magma(200)), title = '10 most variable genes', mar = c(0,0,1,0))
-#Note that the scale now extends to -1
-
-#Again but without the pseudogene
-
-variable_gene_names_notop <- variable_genes$Geneid[2:11]
-big_top_twentyfour_mostvar_mat = big_top_twentyfour %>% 
-  filter(Geneid %in% variable_gene_names_notop) %>% 
-  select(-Geneid)
-corrplot(cor(big_top_twentyfour_mostvar_mat), method = 'color', order = 'hclust', tl.cex = 0.7,
-         col = c(magma(200)), title = '10 most variable genes - except pseudo', mar = c(0,0,1,0))
-
-```
-
-# Undetected _cir_ genes.
-
-Which _cir_ genes are never expressed over the threshold of `r round(min_median, digits = 2)` TPM.
-
-```{r never_expressed}
-filter_at(twentyfour_ct_comb$tpm, 
-          vars(-contains('Geneid')), 
-          all_vars(. < min_median)) %>% 
-  filter(Geneid %in% cir_id) %>% 
-  kable() %>% 
-  kable_styling() %>% 
-  scroll_box(width = "100%", height = "500px")
-```
-
-# PCA of the life stages {.tabset}
-
-## PCA with all genes {.tabset}
-
-### 2D plot of PC1 v PC2
-
-```{r pca_calculation, results = TRUE}
-twentyfour_tpm <- twentyfour_ct_comb$tpm
-rownames_geneid <- twentyfour_tpm$Geneid
-twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
-rownames(twentyfour_tpm_forpca) <- rownames_geneid
-twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
-#Need to look into what exactly scale and center are doing
-```
-
-```{r 2d_pca, results = TRUE}
-twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
-twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
-
-twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
-                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
-                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
-                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
-  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
-  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
-
-# colornames <- c(cbPalette, 'black', 'red')
-colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(twentyfour_tpm_pca_coords$time)))
-
-pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
-pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
-pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
-pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
-
-pl <- plot_ly(twentyfour_tpm_pca_coords, 
-             x  = ~PC1, y = ~PC2,
-             type = 'scatter', mode = 'markers',
-             color = ~time, text = ~sample,
-             colors = colornames,
-             # size = ~total_features,
-             marker = list(opacity = 1, size = 16),
-             symbol = ~transmission, symbols = c('circle','x')) %>% 
-  layout(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
-         yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
-         legend = list(font = list(size = 8)))
-pl
-orca(pl, "tl19-07_audreybrugat24hchabaudi_2D-PCA.pdf")
-
-head(twentyfour_tpm_pca$x[rev(order(abs(twentyfour_tpm_pca$x[,1]))),,drop=FALSE])[,1:9]
-```
-
-```{r 2d_pca_allgenes_suppfig}
-(
-  plot <- ggplot(twentyfour_tpm_pca_coords,
-                 aes(
-                   x = PC1,
-                   y = PC2,
-                   col = time,
-                   shape = transmission
-                 )) +
-    geom_point(size = 3,
-               alpha = 0.8) +
-    scale_color_manual(values = colornames) +
-    scale_shape_manual(
-      values = setNames(
-        c(17,19),
-        unique(twentyfour_tpm_pca_coords$transmission)
-      )
-    ) +
-    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-    guides(shape=guide_legend(ncol=2),
-           col = guide_legend(ncol=2)) +
-    theme_classic() +
-    theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7))
-)
-ggsave2(filename = 'tl19-07_chabaudi24h_supp_figure_pca_allgenes.png',
-        plot = plot,
-        dpi = 300,
-        width = 15,
-        height = 10,
-        units = 'cm')
-```
-
-Can do a plot of the gene loadings for PC1 and PC2 to see which genes are having the biggest impact on each.
-
-```{r 2d_pca_loadings}
-# plot_ly(data.frame(Geneid = rownames(twentyfour_tpm_pca$x),
-#                    product = chabaudi_gene_info$Product.Description[match(rownames(twentyfour_tpm_pca$x), 
-#                                                          chabaudi_gene_info$Gene.ID)],
-#                    twentyfour_tpm_pca$x),
-#         x = ~PC1, y = ~PC2,
-#         type = 'scatter', mode = 'markers',
-#         text = ~paste(Geneid, product))
-plot_ly_biplot(twentyfour_tpm_pca,
-               prin_components = c(1,2,3),
-               scale.loads = 700,
-               separate_multigene = FALSE,
-               three_dimensions = FALSE,
-               textposition = "top left",
-               gene_info = chabaudi_gene_info)
-orca(plot_ly_biplot(twentyfour_tpm_pca,
-               prin_components = c(1,2,3),
-               scale.loads = 700,
-               separate_multigene = FALSE,
-               three_dimensions = FALSE,
-               textposition = "top left",
-               gene_info = chabaudi_gene_info),
-     "tl19-07_audreybrugat24hchabaudi_biplotPCA.pdf")
-
-```
-
-Negative loading for the the PCHAS_0420800 gene means that the samples which contain it are pushed into the positive of the PComponent. The absolute value is what's important loadings.
-
-Most obvious separation is by method of transmission and not by time of day.
-
-```{r pca_loading_genes}
-top_loadings_pc1 <- twentyfour_tpm_pca$x[rev(order(abs(twentyfour_tpm_pca$x[,1]))),,drop=FALSE][1:5,1:8]
-top_loadings_pc1_genes <- rownames(top_loadings_pc1)
-chabaudi_gene_info %>% filter(Geneid %in% top_loadings_pc1_genes)
-
-pdf('tl19-07_24hchabaudi_heatmap_topPCAloadingsgenes.pdf',width = 7, height = 5)
-draw_heatmap_chabaudi_24h(tpm_input = twentyfour_ct_comb, 
-                          genelist= top_loadings_pc1_genes,
-                          scale = 'max',
-                          log_it = TRUE)
-dev.off()
-```
-
-### 3D plot of PC1 v PC2 v PC3
-
-```{r 3d_pca, results = TRUE}
-
-q <- plot_ly(twentyfour_tpm_pca_coords, 
-             x  = ~PC1, y = ~PC2, z = ~PC3,  mode = 'markers',
-             type = 'scatter3d', color = ~time, text = ~sample,
-             colors = colornames,
-             # size = ~total_features,
-             marker = list(opacity = 1, size = 5),
-             symbol = ~transmission, symbols = c('circle','x')) %>% 
-  layout(scene = list(xaxis = list(title = paste0('PC1 ',pc1_importance,'%')),
-                      yaxis = list(title = paste0('PC2 ',pc2_importance,'%')),
-                      zaxis = list(title = paste0('PC3 ',pc3_importance,'%'))))
-q
-```
-
-PCA struggles to differentiate between the time stages. PC2 clearly separates by transmission method. While PC1 and PC3 vaguely clusters similar times together.
-
-## PCA but only including the _cir_ sequences {.tabset}
-
-### 2D:
-
-```{r onlycir, fig.cap = 'PCA plot of the cir gene expression of 24h samples'}
-pca_function(twentyfour_ct_comb, genevec = cir_id)
-# orca(pca_function(twentyfour_ct_comb, 
-#                   genevec = cir_id,
-#                   scale.pca = TRUE, 
-#                   legend_settings = list(font = list(size = 8))),
-#      "tl19-07_audreybrugat24hchabaudi_2D-PCA-cir.pdf")
-```
-
-The rMT samples appear to contain much greater variation than the tightly clustered SBP samples.
-
-### 3D:
-
-```{r pca_onlycir_3d, fig.cap = 'Three dimensional PCA plot of the cir gene expression of 24h samples'}
-pca_function(twentyfour_ct_comb, genevec = cir_id, three.dimensional = TRUE)
-```
-
-Looks like they both form their own separate clusters of cycling pir expression.
-
-### Biplot:
-
-```{r pca_onlycir_pc2-pc3, fig.cap = 'Biplot of cir gene expression of 24h samples'}
-twentyfour_tpm_forpca <- data.frame(filter(twentyfour_tpm, Geneid %in% cir_id))
-rownames(twentyfour_tpm_forpca) <- sapply(twentyfour_tpm_forpca$Geneid, 
-                                          function(x){ 
-                                            paste(x, 
-                                                  chabaudi_gene_info$Product.Description[match(x,
-                                                                              chabaudi_gene_info$Gene.ID)])
-                                          })
-twentyfour_tpm_pca_onlycir <- prcomp(select(twentyfour_tpm_forpca, -Geneid),
-                                      center = TRUE, 
-                                      scale. = TRUE)
-
-plot_ly_biplot(pca_object = twentyfour_tpm_pca_onlycir,
-               scale.loads = 300,
-               gene_info = chabaudi_gene_info)
-
-```
-
-Heatmap of the five _pirs_ with the greatest impact on the cir-only PCA (greatest modulus):
-
-```{r pca_onlycir_heatmap, fig.cap = 'Expression heatmap of the five cir genes with the highest magnitude variable loading values in the PCA'}
-onlycir_loadings <- data.frame(Geneid = rownames(twentyfour_tpm_pca_onlycir$x),
-                                twentyfour_tpm_pca_onlycir$x)
-
-#Heatmap of the cir genes with greatest modulus from the origin
-top_pca_cir <- data.frame(onlycir_loadings) %>% 
-  mutate(length = sqrt((PC1^2)*(PC2^2))) %>% 
-  dplyr::arrange(desc(length)) %>% 
-  select(c(Geneid, PC1, PC2, PC3, length)) %>% 
-  .$Geneid %>% 
-  .[1:5] %>% 
-  str_extract(pattern = 'PCHAS_[[:digit:]]+')
-draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
-                          genelist = top_pca_cir, 
-                          log_it = TRUE)
-# ggbiplot(twentyfour_tpm_pca_onlycir)
-```
-
-Both the major cirs are good for distinguishing the samples from each other, probably because they are the greatest signals. Some others clearly distinguish between rMT and SBP.
-
-## PCA without the _cir_ sequences
-
-```{r pca_no_cir, fig.cap='PCA plot of all gene expression, without the cir genes, of 24h samples'}
-geneid_nocir <- twentyfour_ct_comb$counts$Geneid[which(!twentyfour_ct_comb$counts$Geneid %in% cir_id)]
-pca_function(twentyfour_ct_comb, 
-             genevec = geneid_nocir)
-```
-
-Looks much the same as the overall PCA.
-
-## PCAs with the MT v SBP pseudogene removed {.tabset}
-
- ~~Should do this but with rescaling of the TPM.~~
-
-### 2D:
-
-```{r pca_no_pseudogene, fig.cap='PCA plot of all gene expression, without the rMT-high pseudogene, of 24h samples'}
-geneid_nopseudo <- twentyfour_ct_comb$counts$Geneid[which(!twentyfour_ct_comb$counts$Geneid %in% 'PCHAS_0420800')] 
-pca_function(twentyfour_ct_comb, 
-             genevec = geneid_nopseudo, 
-             scale.pca = TRUE, 
-             legend_settings = list(font = list(size = 8)))
-orca(pca_function(twentyfour_ct_comb, 
-                  genevec = geneid_nopseudo, 
-                  scale.pca = TRUE, 
-                  legend_settings = list(font = list(size = 8))),
-     "tl19-07_audreybrugat24hchabaudi_2D-PCA-nopseudogene.pdf")
-# pca_function(twentyfour_ct_comb, genevec = geneid_nopseudo, scale.pca = FALSE)
-```
-
-### 3D:
-
-```{r pca_3d_no_pseudogene, fig.cap = 'Three dimensional PCA plot of all gene expression, without the rMT-high pseudogene, of 24h samples'}
-pca_function(twentyfour_ct_comb, 
-             genevec = geneid_nopseudo, 
-             scale.pca = TRUE, 
-             three.dimensional = TRUE)
-```
-
-Removing cirs leads to no change in the PCA but removing the pseudogene PCHAS_0420800 has a massive impact!
-
-Get a sort of ring shape (although there's quite a 'jump' between late trophs and earlier ring stages) showing transcriptional progression, although this is best described by PC2 and PC3 which only account for 15% of variation.
-
-PC1 focusses on variation due to the schizont and SBP early-mid ring samples.
-
-### Biplot:
-
-```{r pca_no_pseudogene_loadings_plot}
-twentyfour_tpm_forpca <- data.frame(
-  filter(
-    twentyfour_ct_comb$tpm, 
-    Geneid %in% geneid_nopseudo)
-)
-rownames(twentyfour_tpm_forpca) <- sapply(twentyfour_tpm_forpca$Geneid, 
-                                          function(x){ 
-                                            paste(x, 
-                                                  chabaudi_gene_info$Product.Description[match(x,
-                                                                              chabaudi_gene_info$Gene.ID)])
-                                          })
-twentyfour_tpm_pca_nopseudo <- prcomp(select(twentyfour_tpm_forpca, -Geneid),
-                                      center = TRUE, 
-                                      scale. = TRUE)
-plot_ly_biplot(twentyfour_tpm_pca_nopseudo,
-               textposition = 'left')
-
-```
-
-## PCA without the schizont samples (and also without the pseudogene)
-
-```{r}
-twentyfour_ct_comb_noschiz <- list(
-  tpm = select_at(
-    twentyfour_ct_comb$tpm,
-    vars(
-      -matches('sch')
-    )
+```
+
+```{r read_tpm_rescaled_data}
+#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
+
+twentyfour_ct_comb <- list( tpm = read_xlsx('tl19-07_TPM_both_nopseudoRescaled_labelsStage_allSamples.xlsx') )
+
+twentyfour_ct_comb_noschz <- list( tpm = read_xlsx('tl19-07_TPM_both_nopseudoRescaled_labelsStage_allSamples.xlsx') %>% 
+                                    dplyr::select(!matches('sch'))  )
+
+twentyfour_avgtpm <- list(
+  'rMT' = list(
+    tpm = read_xlsx(
+                    'tl19-07_TPM_rMT_nopseudoRescaled_labelsStage_meanTPM.xlsx'),
+    experiment = 'rMT'
   ),
-  experiment = '24h-noschizont'
-)
-pca_function(twentyfour_ct_comb_noschiz, 
-             genevec = geneid_nopseudo, 
-             scale.pca = TRUE, 
-             colornames = colornames,
-             three.dimensional = TRUE)
-```
-
-## PCA without the schizont and Early.Mid.Rings samples which dominate PC1 (and also without the pseudogene)
-
-```{r pca_no_pseudogene_no_schizont, messages = FALSE}
-twentyfour_ct_comb_noschiz_noring <- list(
-  tpm = select_at(
-    twentyfour_ct_comb$tpm,
-    vars(
-      -matches('sch|Early.Mid.Rings')
-    )
-  ),
-  experiment = '24h-noschizont-noSBPearlymidring'
-)
-pca_function(twentyfour_ct_comb_noschiz_noring, 
-             genevec = geneid_nopseudo, 
-             scale.pca = TRUE, 
-             colornames = colornames,
-             three.dimensional = TRUE)
-# pca_function(twentyfour_ct_comb, genevec = geneid_nopseudo, scale.pca = FALSE)
-orca(pca_function(twentyfour_ct_comb_noschiz, 
-                  genevec = geneid_nopseudo, 
-                  scale.pca = TRUE, 
-                  colornames = colornames,
-                  legend_settings = list(font = list(size = 8))),
-     "tl19-07_audreybrugat24hchabaudi_2D-PCA-nopseudogene-noschizorearlymidrings.pdf")
-```
-
-Removing just the schizonts (not shown) doesn't make PC1 distinguish the samples much better, and the Early.Mid.Rings still are the most separated by the component.
-Remove both and a ring is produced just with PC1 and PC2.
-
-## PC5 and PC6 of original PCA
-
-Could also look at different principal component where the pseudogene is having less impact to remove the effect of the pseudogene.
-
-```{r pseudogene_pca_component}
-twentyfour_tpm_pca$x[rownames(twentyfour_tpm_pca$x) %in% 'PCHAS_0420800',,drop=FALSE][,1:9, drop=FALSE]
-#PC5 and PC6 show much less impact from the pseudogene
-
-twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
-                                        twentyfour_tpm_pca_var$coord) %>% 
-  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
-  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
-
-pc5_importance <- round(pca_importance$PC5[2]*100, digits = 1)
-pc6_importance <- round(pca_importance$PC6[2]*100, digits = 1)
-
-pl <- plot_ly(twentyfour_tpm_pca_coords, 
-             x  = ~Dim.5, y = ~Dim.6,
-             type = 'scatter', mode = 'markers',
-             color = ~time, text = ~sample,
-             colors = colornames,
-             # size = ~total_features,
-             marker = list(opacity = 1, size = 16),
-             symbol = ~transmission, symbols = c('circle','x')) %>% 
-  layout(xaxis = list(title = paste0('PC5 ',pc5_importance,'%')),
-         yaxis = list(title = paste0('PC6 ',pc6_importance,'%')))
-pl
-```
-
-I don't think this is particularly insightful because the variance explained by PC5 and PC6 is only 2.3% and the PCA doesn't really show much anyway.
-
-Notes:
-
-* Shouldn't make a big difference if we set scale. = FALSE, could try with scale. = TRUE on the counts?
-
-* If you don't scale need to make sure that what's affecting the PCA is something that you're interested in.
-
-* Sparse principal components could be used? Allows you to set a threshold for PCA.
-
-* PC1 separates by mean effect, so basically the highest expressed genes. PC2 onwards separates by harder to understand, subtler differences.
-
-# Removing the pseudogene and re-scaling the TPM data
-
-```{r rescale_tpm}
-
-twentyfour_tpm_nopseudo <- twentyfour_ct_comb$tpm %>% 
-  filter(!Geneid %in% 'PCHAS_0420800')
-
-twentyfour_tpm_rescale_nopseudo <- lapply(colnames(twentyfour_tpm_nopseudo), 
-                                          function(column_name){
-                                            column <- getElement(twentyfour_tpm_nopseudo, 
-                                                                 paste(column_name))
-                                            if(is.numeric(column)){
-                                              #is.numeric to avoid trying to scale the Geneid column
-                                              tpm_scale_factor <- 1e6 / sum(column)
-                                              tpm_scaled <- as.numeric(column*tpm_scale_factor)
-                                            }else{
-                                              tpm_scaled <- column
-                                            }
-                                            return(tpm_scaled)
-                                          }) %>% 
-  bind_cols
-
-colnames(twentyfour_tpm_rescale_nopseudo) <- colnames(twentyfour_tpm_nopseudo)
-twentyfour_tpm_rescale_nopseudo_t <- list(tpm = twentyfour_tpm_rescale_nopseudo)
-# apply(twentyfour_tpm_rescale_nopseudo[,-1], 2, sum)
-
-#Setting the re-scaled version as the default from now on
-#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
-twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo_t
-
-write_xlsx(select(twentyfour_ct_comb$tpm, 
-                  -matches('sch')),
-           'tl19-07_TPM_both_nopseudoRescaled_labelsStage_allSamples.xlsx')
-
-twentyfour_ct_rescale <- lapply(twentyfour_ct, 
-                                function(list_element){
-                                  tpm_data <- list_element %>% 
-                                    .$tpm %>% 
-                                    filter(!Geneid %in% 'PCHAS_0420800')
-
-                                  rescaled_tpm <- lapply(colnames(tpm_data), 
-                                                         function(column_name){
-                                                           column <- getElement(tpm_data, 
-                                                                                paste(column_name))
-                                                           if(is.numeric(column)){
-                                                             #is.numeric to avoid trying to scale the Geneid column
-                                                             tpm_scale_factor <- 1e6 / sum(column)
-                                                             tpm_scaled <- as.numeric(column*tpm_scale_factor)
-                                                           }else{
-                                                             tpm_scaled <- column
-                                                           }
-                                                           return(tpm_scaled)
-                                                         }) %>% 
-                                    bind_cols
-                                  colnames(rescaled_tpm) <-  colnames(list_element$tpm)
-                                  list_element$tpm <- rescaled_tpm
-                                  return(
-                                    list_element
-                                  )
-                                })
-names(twentyfour_ct_rescale) <- unlist(lapply(twentyfour_ct_rescale, getElement, 'experiment'))
-lapply(twentyfour_ct_rescale, function(list_element){
-  apply(list_element$tpm[,-1], 2, sum) # == 1e6
-})
-lapply(twentyfour_ct_rescale, function(list_element){
-  nrow(list_element$tpm) # == 5363
-})
-
-write_xlsx(twentyfour_ct_rescale$rMT$tpm,
-           'tl19-07_TPM_rMT_nopseudoRescaled_labelsStage_allSamples.xlsx')
-write_xlsx(twentyfour_ct_rescale$SBP$tpm,
-           'tl19-07_TPM_SBP_nopseudoRescaled_labelsStage_allSamples.xlsx')
-
-#Setting twentyfour_ct core dataset - removed pseudogene and re-scaled.
-twentyfour_ct <- twentyfour_ct_rescale
-
-```
-
-Removing the pseuodgene then re-scaling so that all samples add up to a million TPM again.
-
-```{r shortform_tpm_df}
-## Setting core dataset twentyfour_avgtpm and twentyfour_avgtpm_noschz - both rescaled w/o the pseudogene.
-
-#Generate short-form version of twentyfour_top  using the mean_tpm:
-twentyfour_avgtpm <- lapply(twentyfour_top, 
-       function(experiment_element){
-         tpm_df <- dcast(experiment_element$tpm, 
-                         Geneid ~ stage, 
-                         value.var = 'mean_tpm') %>% 
-           filter(!Geneid %in% 'PCHAS_0420800')
-
-         #Remove pseudogene and rescale
-         rescaled_tpm <- lapply(colnames(tpm_df), 
-                                function(column_name){
-                                  column <- getElement(tpm_df, 
-                                                       paste(column_name))
-                                  if(is.numeric(column)){
-                                    #is.numeric to avoid trying to scale the Geneid column
-                                    tpm_scale_factor <- 1e6 / sum(column)
-                                    tpm_scaled <- as.numeric(column*tpm_scale_factor)
-                                  }else{
-                                    tpm_scaled <- column
-                                  }
-                                  return(tpm_scaled)
-                                }) %>% 
-           bind_cols
-         colnames(rescaled_tpm) <-  colnames(tpm_df)
-
-         return(
-           list(tpm = rescaled_tpm,
-                experiment = experiment_element$experiment)
-         )
-       })
-
-names(twentyfour_avgtpm) <- sapply(twentyfour_avgtpm, getElement, 'experiment')
-
-write_xlsx(twentyfour_avgtpm$rMT$tpm,
-           'tl19-07_TPM_rMT_nopseudoRescaled_labelsStage_meanTPM.xlsx')
-write_xlsx(twentyfour_avgtpm$SBP$tpm,
-           'tl19-07_TPM_SBP_nopseudoRescaled_labelsStage_meanTPM.xlsx')
-
-twentyfour_avgtpm_labelTimepoints <- twentyfour_avgtpm
-
-colnames(twentyfour_avgtpm_labelTimepoints$rMT$tpm) <- ifelse(
-  !colnames(twentyfour_avgtpm_labelTimepoints$rMT$tpm) %in% c('Geneid', 'sch.culture'),
-  times_stages_table$time[
-    match(colnames(twentyfour_avgtpm_labelTimepoints$rMT$tpm),
-          times_stages_table$r_stage_diff_count)
-  ],
-  colnames(twentyfour_avgtpm_labelTimepoints$rMT$tpm)
-)
-write_xlsx(twentyfour_avgtpm_labelTimepoints$rMT$tpm,
-           'tl19-07_TPM_rMT_nopseudoRescaled_labelsTime_meanTPM.xlsx')
-
-colnames(twentyfour_avgtpm_labelTimepoints$SBP$tpm) <- ifelse(
-  !colnames(twentyfour_avgtpm_labelTimepoints$SBP$tpm) %in% c('Geneid', 'sch.culture'),
-  times_stages_table$time[
-    match(colnames(twentyfour_avgtpm_labelTimepoints$SBP$tpm),
-          times_stages_table$r_stage_diff_count)
-  ],
-  colnames(twentyfour_avgtpm_labelTimepoints$SBP$tpm)
-)
-write_xlsx(twentyfour_avgtpm_labelTimepoints$SBP$tpm,
-           'tl19-07_TPM_SBP_nopseudoRescaled_labelsTime_meanTPM.xlsx')
-
-# lapply(twentyfour_avgtpm,
-#        function(tmp){
-#          apply(
-#            tmp$tpm[-1],
-#            2,
-#            sum
-#          )
-#        })
-
-#Above is a check that they all add up to 1e6
-
-# Setting twentyfour_avgtpm_noschz core dataset
+  'SBP' = list(
+    tpm = read_xlsx(
+                     'tl19-07_TPM_SBP_nopseudoRescaled_labelsStage_meanTPM.xlsx'),
+    experiment = 'SBP'
+  )
+)
 
@@
 
-# lapply(twentyfour_avgtpm_noschz,
-#        function(tmp){
-#          apply(
-#            tmp$tpm[-1],
-#            2,
-#            sum
-#          )
-#        })
-
-#Above is a check that they all add up to 1e6
-
-```
-
-# 3D PCA to save
-
-```{r 3d_PCA_save_rgl}
-library('rgl')
-library('magick')
-
-twentyfour_ct_rescale_allTPM <- lapply(names(twentyfour_ct_rescale), 
-                                       function(list_name){
-                                         getElement(twentyfour_ct_rescale, 
-                                                    paste(list_name))$tpm %>% 
-                                           select(!matches('sch')) %>% 
-                                           rename_with(~ paste0(list_name,
-                                                                '_',
-                                                                .x),
-                                                       matches('[[:digit:]]{2}h'))
-                                       }) %>% 
-  Reduce(merge, .) %>% 
-  column_to_rownames(var = 'Geneid')
-# apply(twentyfour_ct_rescale_allTPM, 2, function(x) 1e10*(sum(x) - 1e6))
-
-twentyfour_ct_rescale_allTPM_pca <- prcomp(twentyfour_ct_rescale_allTPM,
-       center = TRUE, 
-       scale. = TRUE)
-
-twentyfour_ct_rescale_allTPM_pca_var <- get_pca_var(twentyfour_ct_rescale_allTPM_pca)
-twentyfour_ct_rescale_allTPM_pca_coords <- data.frame(twentyfour_ct_rescale_allTPM_pca_var$coord)
-
-twentyfour_ct_rescale_allTPM_pca_coords <- data.frame(
-  sample = rownames(twentyfour_ct_rescale_allTPM_pca_var$coord),
-  PC1 = twentyfour_ct_rescale_allTPM_pca_coords$Dim.1, 
-  PC3 = twentyfour_ct_rescale_allTPM_pca_coords$Dim.3, 
-  PC2 = twentyfour_ct_rescale_allTPM_pca_coords$Dim.2
-) %>% 
-  dplyr::mutate(stage = gsub(sample,
-                             pattern = "_[[:digit:]]$", 
-                             replacement = "")) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, 
-                                       pattern = "_", 
-                                       n = 2)[, 2]) %>% 
-  dplyr::mutate(transmission = str_split_fixed(sample, 
-                                               pattern = "_", 
-                                               n = 3)[, 1])
-
-colornames <- viridis(length(unique(twentyfour_ct_rescale_allTPM_pca_coords$time)))
-colornames <- setNames(colornames, sort(unique(twentyfour_ct_rescale_allTPM_pca_coords$time)))
-
-twentyfour_ct_rescale_allTPM_pca_coords$cols <- colornames[
-  match(twentyfour_ct_rescale_allTPM_pca_coords$time,
-        names(colornames))
-]
-
-pl <- plot_ly(twentyfour_tpm_pca_coords, x = ~PC1, y = ~PC2, 
-              z = ~PC3, mode = "markers", type = "scatter3d", color = ~time, 
-              text = ~sample, colors = colornames, marker = list(opacity = 1, 
-                                                                 size = 5), symbol = ~transmission, symbols = c("circle", 
-                                                                                                                "x")) %>% layout(scene = list(xaxis = list(title = paste0("PC1 ", 
-                                                                                                                                                                          pc1_importance, "%")), yaxis = list(title = paste0("PC2 ",
-                                                                                                                                                                                                                             pc2_importance, "%")), zaxis = list(title = paste0("PC3 ",
-                                                                                                                                                                                                                                                                                pc3_importance, "%"))))
-
-plot3d(
-  twentyfour_ct_rescale_allTPM_pca_coords$PC1,
-  twentyfour_ct_rescale_allTPM_pca_coords$PC2,
-  twentyfour_ct_rescale_allTPM_pca_coords$PC3,
-  col = twentyfour_ct_rescale_allTPM_pca_coords$cols,
-  type = "s", 
-  radius = .025
-)
-
-# colors <- c("royalblue1", "darkcyan", "oldlace")
-# iris$color <- colors[ as.numeric( as.factor(iris$Species) ) ]
-# # Static chart
-# plot3d( iris[,1], iris[,2], iris[,3], col = iris$color, type = "s", radius = .2 )
-# 
-# # We can indicate the axis and the rotation velocity
-# play3d( spin3d( axis = c(0, 0, 1), rpm = 20), duration = 10 )
-```
-
-# Amount of _cir_ expression per stage/sample
-
-```{r amount of cir expression}
-
-twentyfour_ct_comb_cir <- twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
-  mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
-  mutate(diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3]) %>% 
-  mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
-  filter(!(time %in% c('sch', 'Unknown'))) %>% 
-  group_by(sample, stage, transmission, time, diff_count) %>% 
-  dplyr::summarise(sum_tpm = sum(tpm))
-
-cir_plot <- twentyfour_ct_comb_cir
-
-  #Want it to be in the right order, but the line function won't work with a discrete scale
-  order_diff_counts <- c("Early.Rings",
-                         "Early.Mid.Rings",
-                         "Late.Mid.Rings",
-                         "Late.Rings",
-                         "Rings.Troph.conversion",
-                         "Early.Trophs",
-                         "Mid.Trophs",
-                         "Late.Trophs")
-  order_for_plot <- data.frame(order_diff_counts,
-                               order = seq(1, length(order_diff_counts), 1))
-  cir_plot <- mutate(cir_plot, 
-                     order_plot = order_for_plot$order[match(diff_count, 
-                                                             order_for_plot$order_diff_counts)])
-
-# cir_plot <- filter(twentyfour_ct_comb_cir, !(time %in% c('schz', 'Unknown'))) %>% 
-#   mutate(time = as.numeric(time))
-
-p <- ggplot(data = cir_plot, 
-            aes(x = order_plot, 
-                y = sum_tpm,
-                fill = transmission)) +
-  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
-               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
-  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
-  geom_smooth(method = 'loess', se = FALSE, color = 'black', aes(linetype = transmission)) +
-  scale_fill_manual(values = cbPalette) +
-  scale_linetype_manual(values = c('dotted', 'dashed')) +
-    scale_x_continuous(breaks = order_for_plot$order, 
-                       labels = order_for_plot$order_diff_counts) +
-  ylab(expression('Total'~italic(pir)~'TPM')) +
-  theme_classic() +
-  theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
-p
-ggsave(p, filename = 'tl19-07_24hchabaudi_pirexpression_barchartLOESS.pdf')
-
-total_tpm_expression_stacked(list(cir = cir_id),
-                             stack_genes = c('PCHAS_0101200'),
-                             stack_label = 'Ancestral pir?',
-                             stack_grouping = FALSE)
-total_tpm_expression_stacked(list_genes = list(cir = cir_id),
-                             stack_genes = c('PCHAS_0101200', 
-                                             # 'PCHAS_0524800',
-                                             'PCHAS_1100300'),
-                             stack_label = 'pir labels',
-                             stack_grouping = FALSE,
-                             save = TRUE)
-```
-
-##  Stacked bar chart
-
-```{r function_total_tpm_expression_stacked_all}
-total_tpm_expression_stacked_all <- function(
-    list_genes, 
-  save = FALSE,
-  tpm_data = twentyfour_ct,
-  stage_str_extract_regex_pattern = '(?<=[[:digit:]]h_).+(?=_[[:digit:]]+)|sch.culture',
-  stack_genes = cir_id,
-  stack_names = c('Yes',
-                  'No'),
-  stack_label = 'Cir genes',
-  stack_grouping = TRUE
-){
-  lapply(names(list_genes), 
-         function(list_name){
-           #list_name <- names(list_genes)[1]
-           paste(list_name)
-           gene_id <- getElement(list_genes, paste(list_name))
-           list_tpms_line <- lapply(tpm_data, 
-                                    function(input){
-                                      tpm_data <- input$tpm %>% 
-                                        filter(Geneid %in% gene_id) %>% 
-                                        melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-                                        mutate(stage = str_extract(sample, 
-                                                                   pattern = paste(stage_str_extract_regex_pattern)),
-                                               stack_group = Geneid
-                                               ) %>% 
-                                        group_by(stage, sample, stack_group) %>% 
-                                        dplyr::summarise(total_tpm = sum(tpm))
-
-                                      #Make stages with no cirs in the sample give a zero 
-                                      if(!all(colnames(input$tpm) %in% tpm_data$sample)){
-                                        missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
-                                        missing_stages_data_zero = data.frame(
-                                          lapply(missing_stages, 
-                                                 function(missing_st){
-                                                   data.frame(stage = unlist(
-                                                     str_extract(
-                                                       missing_st, 
-                                                       pattern = paste(stage_str_extract_regex_pattern))),
-                                                     sample = missing_st,
-                                                     total_tpm = 0)
-                                                 })
-                                        )
-                                        tpm_data = bind_rows(tpm_data, missing_stages_data_zero)
-                                      }
-
-             #Add experiment column
-             tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
-             max_tpm = if(max(tpm_data$total_tpm) < 10){
-               c(seq(0,mround(max(tpm_data$total_tpm),1),1))
-             }else{
-               c(seq(0,mround(max(tpm_data$total_tpm),5),5))
-             }
-
-             #Get the average data for the line graph
-             tpm_data_line = tpm_data %>% 
-               ungroup %>% 
-               group_by(stage, experiment, stack_group) %>% 
-               dplyr::summarise(total_tpm_median = median(total_tpm),
-                                Number_of_replicates = n()) %>% 
-               arrange(stage)
-
-             return(tpm_data_line)
-           })
-
-           tpm_data_line <- bind_rows(list_tpms_line)
-           #Apply ordering if it is specified
-           if(!is.null(order_vector)){
-             tpm_data_line$stage = factor(tpm_data_line$stage, 
-                                          levels = ordering(tpm_data_line$stage,
-                                                            order_vector))
-           }
-
-           tpm_data_line
-
-           #Calculate the range for the y value to make it look nicer
-           stacked_max <- tpm_data_line %>% 
-                                ungroup %>% 
-                                group_by(stage, experiment) %>% 
-                                dplyr::summarise(max_tpm = sum(total_tpm_median)) %>% 
-                                .$max_tpm %>% 
-             max
-
-           max_tpm <- c(
-             seq(
-               0,
-               mround(
-                 stacked_max,
-                 500),
-               length.out = 11)
-           )
-
-           y <- ggplot(data = tpm_data_line, 
-                       aes(
-                         x = experiment,
-                         y = total_tpm_median, 
-                         fill = stack_group
-                       )
-           ) +
-             geom_bar(position = position_stack(),
-                      stat = 'identity',
-                      colour = 'black',
-                      size = 0.1) +
-             scale_fill_manual(
-               breaks = c(stack_genes, 'Other'),
-               values = c( #Remove the stack_genes from the 'grey' cir group
-                 setNames(cbPalette[1:length(c(stack_genes))],
-                          c(stack_genes)),
-                 setNames(rep_len('white', 
-                                  length.out = length(
-                                    str_subset(unique(tpm_data_line$stack_group),
-                                               pattern = stack_genes, 
-                                               negate = TRUE)
-                                  )),
-                          str_subset(unique(tpm_data_line$stack_group),
-                                     pattern = stack_genes, 
-                                     negate = TRUE)
-                 )
-               )
-               #Need to have limits and values otherwise when there are only cir genes (i.e. only 'Yes') the colour 
-               # palette will mess up and only use one, wrong colour. Limits tells it there are always two values.
-             ) +
-             scale_color_manual('black') %>% 
-             labs(fill = stack_label) +
-             scale_y_continuous(breaks = max_tpm) +
-             theme_classic() +
-             scale_size(range = c(0,3)) +
-             ylab(
-               bquote('total '~italic(.(list_name))~' TPM')
-             ) + 
-             facet_grid(.~stage, 
-                        scales = "free_x",
-                        space = "free_x") +
-             theme(
-               axis.title.x=element_blank(),
-               axis.text.x=element_text(angle = 90, 
-                                        size = 5, 
-                                        vjust = 0.5),
-               axis.text.y = element_text(size = 5),
-               axis.title.y = element_text(size = 8),
-               legend.title = element_text(size = 7),
-               legend.text = element_text(size = 5),
-               legend.key.size = unit(3,'mm'),
-               legend.key.width = unit(2,'mm'),
-               strip.text.x = element_text(size=4, 
-                                           face = 'bold', 
-                                           angle = 90),
-               strip.background = element_rect(
-                 color="black", 
-                 size=0.1, 
-                 linetype="solid"
-               )
-             )
-
-           #This bit is to change the colour of the facet boxes
-
-           g <- ggplot_gtable(ggplot_build(y))
-           strip_t <- which(grepl('strip-t', g$layout$name))
-           stage_vec <- unique(
-             levels(tpm_data_line$stage)
-           )
-           fills <- case_when(str_detect(stage_vec, pattern = 'Asexual') ~ 'coral1' ,
-                              str_detect(stage_vec, pattern = 'Liver') ~ 'plum2',
-                              str_detect(stage_vec, pattern = 'Gametocyte') ~ 'lightskyblue1',
-                              str_detect(stage_vec, pattern  = 'Blood-Meal|Sporozoite|Ookinete') ~ 'lightgreen')
-           k <- 1
-           for (i in strip_t) {
-             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-             k <- k+1
-           }
-
-           if(save == TRUE){
-             ggsave2(paste0('tl19-07_',
-                            gsub(list_name, 
-                                 pattern  = ' ', 
-                                 replacement = '_'),
-                            '_totalTPM_barchart.pdf'),
-                     plot = g,
-                     dpi = 300,
-                     width = 15,
-                     height = 10,
-                     units = 'cm')
-           }
-
-           grid.newpage()
-           grid.draw(g)
-
-         })
-}
-
-total_tpm_expression_stacked_all(list_genes = list(cir = cir_id),
-                             stack_genes = c('PCHAS_0101200', 
-                                             # 'PCHAS_0524800',
-                                             'PCHAS_1100300'),
-                             stack_label = 'pir genes',
-                             stack_names = c('-', 
-                                             'Other'),
-                             stack_grouping = FALSE,
-                             save = TRUE)
-```
-
-Curves represent a local regression (LOESS) function.
-
-The pattern of cir expression is the same between SBP and MT (more-or-less) and at the 17.00 and 20.00 time points it is possible that the amount of cir expression is actually greater in SBP v MT.
-Both transmission methods show an peak of expression around 11-14h, correlating with the onset of sequestration?
-
-For completeness the schizont sample is included below:
-
-```{r}
-cir_plot_schz <- twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
-  mutate(transmission = str_extract(stage, pattern = '^(rMT|SBP)')) %>%
-  mutate(time = str_extract(stage, pattern = '(?<=[[:alpha:]]_)([[:alpha:]]+\\.[[:alpha:]]+)+$')) %>% 
-  filter(grepl(time, pattern =  c('sch'))) %>% 
-  group_by(sample, stage, transmission, time) %>% 
-  dplyr::summarise(sum_tpm = sum(tpm))
-p <- ggplot(data = cir_plot_schz, 
-            aes(x = time, 
-                y = sum_tpm,
-                fill = transmission)) +
-  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
-               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
-  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
-  scale_fill_manual(values = cbPalette) +
-  ylab(expression('Total'~italic(cir)~'TPM')) +
-  scale_y_continuous(limits = c(0, 10000)) +
-  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
-  theme_classic()
-p
-```
-
-# Heatmap of top _pir_ expression and orthology across _Plasmodium_
-
-Need to verify that the strains removed don't have any 'extra' orthologs compared to the strain that I keep.
-
-```{r download_ortholog_data}
-chab_pir_ortho <- read_csv('PlasmoDBv48_PchabaudiPIRs_transformedbyorthology.csv') %>% 
-  mutate(Geneid = `Gene ID`,
-         Input_Geneid = `Input Ortholog(s)`) %>% 
-  filter(!(Organism %in% 'Plasmodium chabaudi chabaudi')) %>% 
-  tidyr::separate_rows(Input_Geneid, 
-                       sep = ',') %>% 
-  mutate(Organism = str_replace(Organism,
-                                pattern = 'lasmodium',
-                                replacement = '\\.'))
-
-#Remove paralogs and split rows out by the Input sequences
-
-chab_pir_ortho_synt <- read_csv('PlasmoDBv48_PchabaudiPIRs_transformedbyorthology_syntenic.csv')
-
-chab_pir_ortho <- filter(
-  chab_pir_ortho,
-  !Organism %in% c(
-    'P. vivax Sal-1',
-    'P. yoelii yoelii 17X',
-    'P. yoelii yoelii 17XML',
-    'P. vinckei vinckei strain vinckei'
-  )
-) %>% 
-  mutate(
-    syntenic = ifelse(`Geneid` %in% chab_pir_ortho_synt$`Gene ID`,
-                           'Yes',
-                           'No'),
-         Organism = str_extract(Organism,
-                                pattern = 'P. [[:alpha:]]+')
-  )
-
-# unique(chab_pir_ortho$Organism)
-
-chab_pir_ortho %>%
-  mutate(species = str_extract(Geneid, pattern = '^.+(?>_)')) %>%
-  group_by(Input_Geneid) %>%
-  dplyr::summarise(orthologs = paste(unique(`Organism`), collapse = ','))  %>%
-  tidyr::separate_rows(orthologs, sep = ',') %>%
-  group_by(Input_Geneid) %>%
-  dplyr::summarise(num = n()) %>%
-  arrange(desc(num)) %>%
-  mutate(product = chabaudi_gene_info$handy_descriptions[match(Input_Geneid,
-                                                         chabaudi_gene_info$Geneid)])
-
-chab_pir_ortho %>%
-  mutate(species = str_extract(Geneid, 
-                               pattern = '^.+(?>_)')) %>%
-  group_by(Input_Geneid) %>%
-  dplyr::summarise(orthologs = paste(unique(`Organism`), 
-                                     collapse = ', '))
-
-```
-
-```{r ortholog_heatmap}
-
-#   #pir_family_name <- names(pir_genes_list_ortho)[1]
-#   pir_family <-  getElement(pir_genes_list_ortho, 
-#                             paste(pir_family_name))
-#   #Need to convert this into a character vector for the filter function
-#   pir_family <- apply(pir_family, 1, as.character)
-
-threshold <- 0
-
-#Want Number for whether the orthologue is present in each species
-pir_ortholog_species <- chab_pir_ortho %>% 
-  filter(Input_Geneid %in% cir_id) %>% 
-  purrr::modify_if(is.character, 
-            as.factor) %>% 
-  group_by(Organism, 
-           Input_Geneid, 
-           .drop = FALSE) %>%
-  #.drop must be FALSE otherwise dplyr will drop zero values
-  select(Input_Geneid, 
-         Organism, syntenic) %>%
-  dcast(Input_Geneid ~ Organism, 
-        value.var = 'syntenic', 
-        fun.aggregate = function(syntenic){
-          if(any(syntenic == 'Yes')){
-            2
-          }else if(length(syntenic) > 0){
-            1
-          }else{
-            0
-          }
-        })
-
-#Add in the genes which have no orthologs
-pir_ortholog_species <- pir_ortholog_species %>% 
-  tidyr::complete(Input_Geneid = cir_id)
-
-  #Alter here whether you want only the genes with orthologs, or all genes even if they don't have an ortholog
-
-  pir_ortholog_species <-  dplyr::rename(pir_ortholog_species, 
-                                  rowname = Input_Geneid) %>%
-    mutate(rowname = str_replace(as.character(rowname), 
-                                 pattern = 'PCHAS_', 
-                                 replacement = '')) %>%
-    arrange(
-      match(
-        rowname,
-        str_sort(
-          rowname,
-          numeric = TRUE
-        )
-      )
-    ) %>%
-    #Use str_sort because the character format is sorted differently to how I want it normally
-    column_to_rownames()
-  pir_ortholog_species_rownam <- rownames(pir_ortholog_species)
-  pir_ortholog_species <- apply(pir_ortholog_species, 2, function(x){
-    case_when(
-      x == 0 ~ NA_character_,
-      x == 1 ~ 'No',
-      x == 2 ~ "Yes")
-  })
-  rownames(pir_ortholog_species) <- pir_ortholog_species_rownam
-
-  #Not going to worry about whether they are pseudogenes or not for now
-
-  ortho_gene_vec <- rownames(pir_ortholog_species)
-
-  heatmap_row_name_fonts <- 10
-  heatmap_col_name_fonts <- 10
-  heatmap_col_title_fonts <- 15
-
-  #Get the TPM data for the pirs family being looked at
-  expt_tpms_top_pirs <- lapply(twentyfour_avgtpm_noschz, 
-                               function(tpm_input){
-                                 # print(paste0('Now doing ', tpm_input$experiment))
-                                 tpms <- tpm_input$tpm %>% 
-                                   filter(Geneid %in% cir_id)
-                                 #Remove PCHAS_ from the gene names for simplicity
-                                 rownames(tpms) = str_replace(as.character(tpms$Geneid), 
-                                                              pattern = 'PCHAS_', 
-                                                              replacement = '')
-                                 tpms <- tpms[str_sort(rownames(tpms), 
-                                                       numeric = TRUE),]
-                                 tpms <- select(tpms, -Geneid)
-
-    colnames(tpms) <- times_stages_table$time[match(colnames(tpms),
-                                                    times_stages_table$r_stage_diff_count)]
-
-    tpms <- tpms[,
-                 order(colnames(tpms)),
-                 drop = FALSE]
-    #Ordering using the ordering function and order_vector defined in the functions section.
-
-    tpms <- apply(tpms, 2, function(x){ifelse(x >= threshold, 
-                                              x, 
-                                              NA)}) #Get NA for values under threshold
-    tpms <- log1p(tpms)
-    if(tpm_input$experiment %>%  is.null){
-      tpm_input$experiment = ''
-    }
-    list = list(tpms, tpm_input$experiment)
-    names(list) = c('tpm','experiment')
-    return(list)
-  })
-
-  #if they are all NA then we don't want it to proceed further
-  if(
-    all(
-      unlist(
-        lapply(
-          expt_tpms_top_pirs, 
-          function(x) sapply(x$tpm, 
-                             is.na)  
-        )
-      )
-    )
-  ){stop()}
-
-    maxcol = max(
-      unlist(
-        lapply(
-          expt_tpms_top_pirs, 
-          function(x) max(
-            x$tpm,
-            na.rm = TRUE
-          )
-        )
-      )
-    )
-    #Get the max TPM value
-
-    if(is.na(maxcol)|maxcol<1){maxcol <- 1}
-
-    #Re-order both data.frames by the median highest expressed pirs
-
-    tpm_order <- order(
-      rowMedians(
-        sapply(
-          lapply(
-            expt_tpms_top_pirs,
-            function(x) rowMedians(x$tpm)
-          ),
-          rbind),
-        na.rm = TRUE),
-      decreasing = TRUE
-    )[1:10]
-
-    heatmaps = lapply(expt_tpms_top_pirs, 
-                      function(tpms){
-                        if(tpms$experiment != last(expt_tpms_top_pirs)$experiment){
-                          #If it is the final element don't remove the row names
-                          rownames(tpms$tpm) = c()
-                        }
-                        #Setting the name if one is not provided
-                        title_name = tpms$experiment
-                        #Drawing each heatmap
-                        ht = Heatmap(tpms$tpm[tpm_order,, drop = FALSE], 
-                                     name = 'log(TPM + 1)', 
-                                     col = colorRamp2(c(threshold, 
-                                                        maxcol/4, 
-                                        maxcol/2, 
-                                        3*maxcol/4, 
-                                        maxcol), 
-                                      viridis(5)),
-                     cluster_rows = FALSE, 
-                     cluster_columns = FALSE,
-                     column_title = paste(title_name),
-                     column_title_gp = gpar(fontsize = heatmap_col_title_fonts),
-                     row_names_gp = gpar(fontsize = heatmap_row_name_fonts),
-                     column_title_rot = 0,
-                     column_names_gp = gpar(fontsize = heatmap_col_name_fonts),
-                     na_col = 'white', 
-                     border = TRUE,
-                     border_gp = gpar(col = "black", 
-                                      lwd = 2),
-                     heatmap_legend_param = list(
-                       title = 'log(TPM + 1)', 
-                       at = mround(
-                         c(log1p(threshold), 
-                           seq(0, 
-                               maxcol, 
-                               length.out = 4)[-1]
-                         ), 
-                         base = 0.05)
-                     )
-        )
-    })
-
-    ortholog_for_heatmap <- as.matrix(pir_ortholog_species)[tpm_order,
-                                                            !apply(
-                                                              pir_ortholog_species[tpm_order,,drop = FALSE],
-                                                              2,
-                                                              function(ortho) all(is.na(ortho))
-                                                            ),
-                                                            drop = FALSE]
-
-    ha = Heatmap(ortholog_for_heatmap,
-                 name = 'Syntenic Orthologs',
-                 na_col = 'white',
-                 col = structure(rev(plasma(2)), 
-                                 names = c('Yes', 'No')),
-                 cluster_rows = FALSE, 
-                 cluster_columns = FALSE,
-                 column_names_gp = gpar(col = 'darkslategray', 
-                                    fontsize = 7*(50/(ncol(ortholog_for_heatmap)+26)), 
-                                    fontface = 'italic'),
-                 show_column_dend = FALSE,
-                 rect_gp = gpar(col = 'white'), 
-                 width = unit(4, "cm"),
-                 heatmap_legend_param = list(
-
-                     )
-    )
-
-    ht_list = Reduce('+', heatmaps) #To add the list elements together
-    ht_list <- ha + ht_list
-
-    if(global_save_var){
-      pdf(
-        paste0('tl19-07_pir_',
-               'orthologs_heatmap.pdf')
-        # , width = 17
-        , height = 5
-      )
-      draw(ht_list,
-           row_title_side = 'left',
-           column_title = 'Malaria stage',
-           column_title_side = 'bottom',
-           heatmap_legend_side = 'left',
-           ht_gap = unit(0.75, "cm"))
-      dev.off()
-    }
-
-    draw(ht_list,
-           row_title_side = 'left',
-           column_title = 'Malaria stage',
-           column_title_side = 'bottom',
-           heatmap_legend_side = 'left',
-           ht_gap = unit(0.6, "cm"))
-
-```
-
-```{r heatmap_ortho_most_conserved}
-
-    #Re-order by most conserved genes
-
-    tpm_order <- order(
-      apply(pir_ortholog_species, 
-            1, 
-            function(ortho_row) sum(!is.na(ortho_row))),
-      decreasing = TRUE
-    )[1:10]
-
-    heatmaps = lapply(expt_tpms_top_pirs, 
-                      function(tpms){
-                        if(tpms$experiment != last(expt_tpms_top_pirs)$experiment){
-                          #If it is the final element don't remove the row names
-                          rownames(tpms$tpm) = c()
-                        }
-                        #Setting the name if one is not provided
-                        title_name = tpms$experiment
-                        #Drawing each heatmap
-                        ht = Heatmap(tpms$tpm[tpm_order,, drop = FALSE], 
-                                     name = 'log(TPM + 1)', 
-                                     col = colorRamp2(c(threshold, 
-                                                        maxcol/4, 
-                                        maxcol/2, 
-                                        3*maxcol/4, 
-                                        maxcol), 
-                                      viridis(5)),
-                     cluster_rows = FALSE, 
-                     cluster_columns = FALSE,
-                     column_title = paste(title_name),
-                     column_title_gp = gpar(fontsize = heatmap_col_title_fonts),
-                     row_names_gp = gpar(fontsize = heatmap_row_name_fonts),
-                     column_title_rot = 90,
-                     column_names_gp = gpar(fontsize = heatmap_col_name_fonts),
-                     na_col = 'white', border = TRUE,
-                     heatmap_legend_param = list(
-                       title = 'log(TPM + 1)', 
-                       at = mround(
-                         c(log1p(threshold), 
-                           seq(0, 
-                               maxcol, 
-                               length.out = 4)[-1]
-                         ), 
-                         base = 0.05)
-                     )
-        )
-    })
-
-    ha = Heatmap(as.matrix(pir_ortholog_species)[tpm_order,, drop = FALSE],
-                 name = 'Syntenic Orthologs',
-                 na_col = 'white',
-                 col = structure(rev(plasma(2)), 
-                                 names = c('Yes', 'No')),
-                 cluster_rows = FALSE, 
-                 cluster_columns = FALSE,
-                 column_names_gp = gpar(col = 'darkslategray', 
-                                        fontsize = 4*(50/(ncol(pir_ortholog_species)+26)), 
-                                        fontface = 'italic'),
-                 show_column_dend = FALSE,
-                 rect_gp = gpar(col = 'white'), 
-                 width = unit(4, "cm"),
-                 heatmap_legend_param = list(
-
-                     )
-                 )
-
-    ht_list = Reduce('+', heatmaps) #To add the list elements together
-    ht_list <- ha + ht_list
-    # pdf(
-    #   paste0('tl19-07_pir_',
-    #          'orthologs_heatmap.pdf')
-    #   # , width = 17
-    #   , height = 5
-    # )
-    draw(ht_list,
-         row_title_side = 'left',
-         column_title = 'Malaria stage',
-         column_title_side = 'bottom',
-         heatmap_legend_side = 'left')
-    # dev.off()
-
-```
-
-# Figure 2b - Heatmap of top S and L gene expression.
+```
 
@@
   c(    '#F0E442',
-        '#56B4E9',
-        '#E69F00',
+        '#56B4E9', #L
+        '#E69F00', #S
         '#999999'),
@@
 
+SorL_col_num <- SorL_col
+names(SorL_col_num) <- sapply(names(SorL_col), 
+                              function(clade_name) {
+                                paste0(
+                                  clade_name, 
+                                  ' (n = ',
+                                  sum(pir_clade$SorL == clade_name),
+                                  ')'
+                                )
+                              })
+
 clade_col <- setNames(
   c('#F0E442',
-    '#56B4E9',
+    '#56B4E9', #L1
     '#009E73',
-    '#E69F00',
     '#0072B2',
+    '#E69F00', #S1
     '#D55E00',
@@
 
-```
-
-```{r figure2b_heatmap_table}
+clade_col_num <- clade_col
+names(clade_col_num) <- sapply(names(clade_col), 
+                               function(clade_name) {
+                                 paste0(
+                                   clade_name, 
+                                   ' (n = ',
+                                   sum(pir_clade$clade == clade_name),
+                                   ')'
+                                 )
+                               })
+
+#Be careful to ensure that the above is consistent with other images in the same paper.
+
+```
+
+# Figure  - Heatmap of top S and L gene expression.
+
+```{r figure_heatmap_table}
 
@@
 
-                            write.csv(tpm_data_SorL_stackgenes_top_table,
-                                      file = paste0('tl19-07_combined_SorL_topperstage_table_', 
-                                                    tpm_input$experiment, 
-                                                    '.csv'))
-
-                            write_xlsx(tpm_data_SorL_stackgenes_top_table,
-                                       paste0('tl19-07_combined_SorL_topperstage_table_', 
-                                              tpm_input$experiment, 
-                                              '.xlsx'))
+                            # write.csv(tpm_data_SorL_stackgenes_top_table,
+                            #           file = paste0('tl19-07_combined_SorL_topperstage_table_', 
+                            #                         tpm_input$experiment, 
+                            #                         '.csv'))
+
+                            # write_xlsx(tpm_data_SorL_stackgenes_top_table,
+                            #            paste0('tl19-07_combined_SorL_topperstage_table_', 
+                            #                   tpm_input$experiment, 
+                            #                   '.xlsx'))
 
@@
 
-```{r figure2b_heatmap}
+```{r figure_heatmap}
 
@@
 heatmap_col_name_fonts <- 15
+heatmap_col_title_fonts <- 15 
+threshold <- 0
 
@@
                                                                          NA)}) #Get NA for values under threshold
-                               tpms <- log1p(tpms)
+                               tpms <- log2(1+tpms)
                                if(tpm_input$experiment %>%  is.null){
@@
 
-                    if(tpms$experiment != last(expt_tpms_top_pirs)$experiment){
+                    if(tpms$experiment != dplyr::last(expt_tpms_top_pirs)$experiment){
                       #If it is the final element don't remove the row names
@@
                                    at = mround(
-                                     c(log1p(threshold), 
+                                     c(log2(1+threshold), 
                                        seq(0, 
@@
 
-# Ancestral and rodent-conserved gene expression curves
-
-```{r ancestral_pir_curve}
-gene_graph_24h <- function(input = twentyfour_ct_comb, 
-                           genevec = cir_id,
-                           genename,
-                           use_times = FALSE) {
-  cir_plot <- input$tpm %>% 
-    filter(Geneid %in% genevec) %>% 
-    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
-    mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
-    mutate(diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3]) %>% 
-    mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
-    filter(!(time %in% c('sch', 'Unknown'))) %>% 
-    group_by(sample, stage, transmission, time, diff_count) %>% 
-    dplyr::summarise(sum_tpm = sum(tpm))
-
-  #Want it to be in the right order, but the line function won't work with a discrete scale
-  order_diff_counts <- c("Early.Rings",
-                         "Early.Mid.Rings",
-                         "Late.Mid.Rings",
-                         "Late.Rings",
-                         "Rings.Troph.conversion",
-                         "Early.Trophs",
-                         "Mid.Trophs",
-                         "Late.Trophs")
-
-  order_for_plot <- data.frame(order_diff_counts,
-                               order = seq(1, 
-                                           length(order_diff_counts), 
-                                           1))
-  cir_plot <- mutate(cir_plot, 
-                     order_plot = order_for_plot$order[match(diff_count, 
-                                                             order_for_plot$order_diff_counts)])
-  if(use_times){
-    order_for_plot$order_diff_counts <- times_stages_table$time[match(order_for_plot$order_diff_counts,
-                                                                      times_stages_table$r_stage_diff_count)]
-  }
-
-  p <- ggplot(data = cir_plot, 
-              aes(x = order_plot, 
-                  y = sum_tpm,
-                  fill = transmission)) +
-    stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
-                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
-    geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
-    geom_smooth(method = 'loess', se = FALSE, color = 'black', aes(linetype = transmission)) +
-    scale_fill_manual(values = cbPalette) +
-    scale_linetype_manual(values = c('dotted', 'dashed')) +
-    scale_x_continuous(breaks = order_for_plot$order, 
-                       labels = order_for_plot$order_diff_counts) +
-    ylab(as.expression(bquote('Total'~italic(.(genename))~'TPM'))) +
-    xlab('Asexual blood stage development cycle') +
-    theme_classic() +
-    theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
-  p
-}
-
-gene_graph_24h(genevec = 'PCHAS_0101200',
-               genename = 'highly-conserved pir',
-               use_times = TRUE)
-
-gene_graph_24h(input = twentyfour_tpm_rescale_nopseudo_t,
-               genevec = 'PCHAS_0101200',
-               genename = 'highly-conserved pir')
-
-ggsave(gene_graph_24h(genevec = 'PCHAS_0101200',
-                      genename = 'highly-conserved pir'),
-       filename = 'tl19-07_24hchabaudi_ancestral_expressionbarchartLOESS.pdf')
-# gene_graph_24h(genevec = 'PCHAS_0524800',
-#                genename = 'rodent-conserved pir')
-# ggsave(gene_graph_24h(genevec = 'PCHAS_0524800',
-#                       genename = 'rodent-conserved pir'),
-#        filename = 'tl19-07_24hchabaudi_rodentpir_expressionbarchartLOESS.pdf')
-
-cir_id_no_anc <- str_subset(cir_id, 
-                            pattern = 'PCHAS_0101200', 
-                            negate = TRUE)
-gene_graph_24h(genevec = cir_id_no_anc,
-               genename = 'pir without ancestral or rodent-conserved')
-
-# gene_graph_24h(genevec = 'PCHAS_0420800', 
-#                genename = 'rMT his-rich pseudogene')
-```
-
-## Expression of major expressed cirs rMT/SBP side-by-side.
+## Figure 4 - Expression of major expressed cirs rMT/SBP side-by-side.
 
@@
 
-```{r pirs_side-by-side_figure2c}
+```{r pirs_side-by-side_figure4}
 
@@
                        'Unknown'))) %>% 
-  group_by(Geneid) %>% 
-  filter(any(tpm > 100)) %>% 
   mutate(subfam = str_replace(subfam, 
@@
 
+limits_cir <- c(0, 
+                mround(max(cir_plot$tpm), 
+                       base = 100))
+
 cir_plot_order <- cir_plot %>% 
@@
                                    levels = unique(cir_plot_order$Geneid)))
-limits_cir <- c(0, mround(max(cir_plot$tpm), base = 100))
-
-side_by_side_cir_plot <- function(time_vec = '11h',
+
+side_by_side_cir_plot <- function(time_vec = '14h',
                                   log_it = TRUE,
@@
                                   limits = limits_cir,
-                                  point_size = 1){
-
-    cir_plot <- filter(cir_plot, 
-                       time == time_vec)
+                                  point_size = 1,
+                                  text_size = 10,
+                                  file_name_end = '',
+                                  threshold_min = 100,
+                                  threshold_max = Inf){
+
+  cir_plot <- group_by(cir_plot, 
+                       Geneid) %>% 
+    filter(
+      time == time_vec,
+      !Geneid %in% cir_id | any(tpm >= threshold_min),
+      !Geneid %in% cir_id | all(tpm <= threshold_max)
+    )
 
@@
   if(log_it){
-      cir_plot$tpm <- log1p(cir_plot$tpm)
-      limits = log1p(limits)
+    cir_plot$tpm <- log2(1+cir_plot$tpm)
+      limits = log2(1+limits)
       breaks <- round(seq.int(0, 
@@
                          limits = limits) +
+      scale_fill_manual(values = c('skyblue','red')) +
       facet_grid(.~subfam, 
@@
       theme(
-        axis.text.x = element_text(angle = 90)
+        axis.text.x = element_text(size = text_size,
+                                   angle = 90)
       ) +
@@
       ggtitle(paste(unique(cir_plot$time)))
+
+    #This bit is to change the colour, and text colour, of the facet boxes
+
+    g <- ggplot_gtable(ggplot_build(y))
+    strip_t <- which(grepl('strip-t', g$layout$name))
+    subfam_vec <- sort(unique(cir_plot$subfam))
+    fills <- case_when(str_detect(subfam_vec, 
+                                  pattern = 'a') ~ clade_col['ancestral'],
+                       subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
+                       str_detect(subfam_vec, 
+                                  pattern = 'ref') ~ 'white')
+    k <- 1
+    for (i in strip_t) {
+      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
+      # if(getGrob(g$grobs[[i]]$grobs[[1]], 'GRID', grep = TRUE)$label == 'a'){
+      #   #If grob is for the 'a' label we want to change the font
+      #   m <- which(grepl('text', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+      #   n <- which(grepl('GRID.text', g$grobs[[i]]$grobs[[1]]$children[[m]]$childrenOrder))
+      #   g$grobs[[i]]$grobs[[1]]$children[[m]]$children[[n]]$gp$col <- 'white'
+      # }
+      k <- k+1
+    }
+
     if(save){
@@
                time_vec,
+               file_name_end,
                '.png'),
-        plot = y,
+        plot = g,
         dpi = 300,
@@
     }
-    return(y)
+
+    grid.newpage()
+    grid.draw(g)
+
 }
@@
     sort(unique(cir_plot$time)),
-    side_by_side_cir_plot
-  )
-
-```
-
-## Expression of the S7 pirs.
-
-```{r pirs_side-by-side_different_criteria}
+    side_by_side_cir_plot,
+    threshold_min = 50,
+    threshold_max = Inf,
+    point_size = 1,
+    text_size = 10,
+    file_name_end = '_Over100TPM',
+    save = TRUE
+  )
+```
+
+```{r}
+  lapply(
+    sort(unique(cir_plot$time)),
+    side_by_side_cir_plot,
+    threshold_min = 0,
+    threshold_max = 50,
+    point_size = 0.1,
+    text_size = 3,
+    file_name_end = '_Under50TPM',
+    save = TRUE
+  )
+#Note that the file name above may not be accurate - this is just so that I don't have to make the supp figure from scratch.
+```
+
+```{r pirs_side-by-side_figure4}
+
+housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
+                             'MSP1' = 'PCHAS_0831300',
+                             'AMA1' = 'PCHAS_0931000')
+
+cir_housekeeping_id <- c(cir_id, 
+                         'PCHAS_0721000',
+                         'PCHAS_0831300',
+                         'PCHAS_0931000')
 
 cir_plot <- twentyfour_ct_comb$tpm %>% 
-    filter(Geneid %in% cir_id) %>%
+  filter(Geneid %in% cir_housekeeping_id) %>%
   melt(id.vars = 'Geneid', 
@@
        value.name = 'tpm') %>% 
-    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
+  mutate(
+    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
     transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
@@
     time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
-           subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]) %>% 
-    filter(!(time %in% c('sch', 
+    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
+  ) %>% 
+  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
+  filter(!(time %in% c('sch.culture', 
                        'Unknown'))) %>% 
-    group_by(Geneid)
-
-  breaks = c(7,6,5,4, 3, 2, 1, 0)
-  limits = c(0,7.2)
-  point_size = 0.25
-
-  lapply(sort(unique(cir_plot$subfam)),
-  function(subfam_nam){
-    cir_plot <- cir_plot %>% 
-      filter(
-        # all(tpm < 100),
-        #      any(tpm > 50)
-        subfam == subfam_nam
-      ) %>% 
-      arrange(desc(tpm))
+  mutate(subfam = str_replace(subfam, 
+                              'ancestral',
+                              'a')) %>% 
+  mutate(
+    subfam = interaction(
+      transmission,
+      forcats::fct_shift(
+        relevel(
+          as.factor(subfam),
+          ref = 'ref')
+      ),
+      sep = '_',
+      lex.order = TRUE
+    )
+  )
+#Need to add the transmission name to the subfam so that they are ordered correctly.
+# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
+# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.
+
+cir_plot$Geneid <- ifelse(
+  cir_plot$Geneid %in% housekeeping_genes[1,],
+  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
+  cir_plot$Geneid
+)
+
+limits_cir <- c(0, 
+                mround(max(cir_plot$tpm), 
+                       base = 100))
 
@@
 
-    lapply(
-      sort(unique(cir_plot$time)),
-      function(time_vec){
-
-        cir_plot <- filter(cir_plot, 
-                           time == time_vec)
-
-      cir_plot$tpm <- log1p(cir_plot$tpm)
+transmissionSplit_cir_plot <- function(time_vec = '14h',
+                                  log_it = TRUE,
+                                  save = TRUE,
+                                  limits = limits_cir,
+                                  point_size = 1,
+                                  text_size = 10,
+                                  file_name_end = '',
+                                  threshold_min = 100,
+                                  threshold_max = Inf){
+
+  cir_plot <- group_by(cir_plot, 
+                       Geneid) %>% 
+    filter(
+      time == time_vec,
+      !Geneid %in% cir_id | any(tpm >= threshold_min),
+      !Geneid %in% cir_id | all(tpm <= threshold_max)
+    )
+
+  # cir_plot %>% group_by(Geneid) %>% summarise(n())
+  if(log_it){
+    cir_plot$tpm <- log2(1+cir_plot$tpm)
+      limits = log2(1+limits)
+      breaks <- round(seq.int(0, 
+                              max(limits), 
+                              length.out = round(max(limits)) + 1, 
+                              digits = 0), 
+                      digits = 0)
+    }else{
+      breaks <- seq.int(0, max(limits), length.out = 11)
+    }
 
@@
                          limits = limits) +
-        ylab('ln(TPM + 1)')+
+      scale_fill_manual(values = c('skyblue','red')) +
       facet_grid(.~subfam, 
                  scales = "free_x",
-                   space = "free_x") +
+                 space = "free_x",
+                 labeller = labeller(subfam = function(string) {
+                   str_remove_all(string, 
+                                  pattern = '(rMT|SBP)_')
+                 })) +
       theme_classic() +
       theme(
-          axis.text.x = element_text(angle = 90,
-                                     size = 5)
+        axis.text.x = element_text(size = text_size,
+                                   angle = 90,
+                                   # hjust = 2,
+                                   vjust = 0.5)
       ) +
+      ylab('log(TPM + 1)')+
       ggtitle(paste(unique(cir_plot$time)))
 
+    #This bit is to change the colour, and text colour, of the facet boxes
+
+    g <- ggplot_gtable(ggplot_build(y))
+
+    # g$widths[5] = 2*g$widths[5]
+
+    strip_t <- which(grepl('strip-t', g$layout$name))
+    subfam_vec <- sort(unique(cir_plot$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
+    fills <- case_when(str_detect(subfam_vec, 
+                                  pattern = 'a') ~ clade_col['ancestral'],
+                       subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
+                       str_detect(subfam_vec, 
+                                  pattern = 'ref') ~ 'white')
+    k <- 1
+    for (i in strip_t) {
+      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
+      k <- k+1
+    }
+
+    #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
+    # manually make them a bit wider.
+
+    short_panels <- which(str_detect(subfam_vec, pattern = 'a|ref'))
+
+    for(i in short_panels){
+      num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
+      g$widths[num] = 1.75*g$widths[num]
+      #1.75 times the width of the right panels.
+      # print(num)
+    }
+
+    if(save){
       ggsave2(
-        paste0('tl20-05_side-by-side_barchart_',
-               subfam_nam,
-               '_',
+        paste0('tl19-07_barchart_facetTrans_facetSubfam_',
                time_vec,
+               file_name_end,
                '.png'),
-        plot = y,
+        plot = g,
         dpi = 300,
-        width = 25,
-        height = 10,
+        width = 28,
+        height = 8,
         units = 'cm'
       )
-      return(y)
-    })
-})
-
-```
-
-## Total pir expression in each family across time.
-
-```{r pir_family_expression}
+    }
+
+    grid.newpage()
+    grid.draw(g)
+
+}
+
+# transmissionSplit_cir_plot(
+#   sort(unique(cir_plot$time))[1],
+#     threshold_min = 100,
+#     threshold_max = Inf,
+#     point_size = 1,
+#     text_size = 4,
+#     file_name_end = '_Over100TPM',
+#     save = FALSE)
+
+ lapply(
+    sort(unique(cir_plot$time)),
+    transmissionSplit_cir_plot,
+    threshold_min = 100,
+    threshold_max = Inf,
+    point_size = 1,
+    text_size = 5,
+    file_name_end = '_Over100TPM',
+    save = TRUE
+  )
+```
+
+```{r pirs_side-by-side_figure4_RMT-SBP-completelySeparate}
 
 cir_plot <- twentyfour_ct_comb$tpm %>% 
-    filter(Geneid %in% cir_id) %>%
+  filter(Geneid %in% cir_housekeeping_id) %>%
   melt(id.vars = 'Geneid', 
@@
     time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
-           subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]) %>% 
-    filter(!(time %in% c('sch', 
+    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
+  ) %>% 
+  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
+  filter(!(time %in% c('sch.culture', 
                        'Unknown'))) %>% 
-    group_by(subfam, 
-             time,
+  mutate(subfam = str_replace(subfam, 
+                              'ancestral',
+                              'a')) %>% 
+  mutate(
+    subfam = interaction(
       transmission,
-             sample) %>% 
-    summarise(
-      sum_tpm = sum(tpm)
-    )
-
-  breaks = seq.int(0, mround(max(cir_plot$sum_tpm), base = 100), length.out = 11)
-    limits = c(0, mround(max(cir_plot$sum_tpm), base = 100))
-    point_size = 0.25
-
-  lapply(sort(unique(cir_plot$subfam)), 
-         function(subfam_nam){
-
-  # cir_plot_order <- cir_plot %>% 
-  #   filter(time == '14h',
-  #          transmission == 'rMT') %>% 
-  #   arrange(desc(tpm))
-  #   
-  # cir_plot <- mutate(cir_plot, 
-  #                    Geneid = factor(Geneid, 
-  #                                    levels = unique(cir_plot_order$Geneid)))
-
-    cir_plot <- filter(cir_plot, 
-                       subfam == subfam_nam)
-
-      # cir_plot$tpm <- log1p(cir_plot$sum_tpm)
+      forcats::fct_shift(
+        relevel(
+          as.factor(subfam),
+          ref = 'ref')
+      ),
+      sep = '_',
+      lex.order = TRUE
+    )
+  )
+#Need to add the transmission name to the subfam so that they are ordered correctly.
+# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
+# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.
+
+cir_plot$Geneid <- ifelse(
+  cir_plot$Geneid %in% housekeeping_genes[1,],
+  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
+  cir_plot$Geneid
+)
+
+limits_cir <- c(0, 
+                mround(max(cir_plot$tpm), 
+                       base = 100))
+
+cir_plot_order <- cir_plot %>% 
+  filter(time == '14h',
+         transmission == 'rMT') %>% 
+  arrange(desc(tpm))
+
+cir_plot <- mutate(cir_plot, 
+                   Geneid = factor(Geneid, 
+                                   levels = unique(cir_plot_order$Geneid)))
+
+transmissionSep_cir_plot <- function(time_vec = '14h',
+                                     log_it = TRUE,
+                                     save = TRUE,
+                                     limits = limits_cir,
+                                     point_size = 1,
+                                     text_size = 10,
+                                     file_name_end = '',
+                                     threshold_min = 100,
+                                     threshold_max = Inf){
+
+  cir_plot_filter <- group_by(cir_plot, 
+                       Geneid) %>% 
+    filter(
+      time == time_vec,
+      !Geneid %in% cir_id | any(tpm >= threshold_min),
+      !Geneid %in% cir_id | all(tpm <= threshold_max)
+    )
+  #Filter for pirs expressed above or below the thresholds, have the OR so that the ref genes aren't removed.
+  # e.g. Either 'Not a cir' OR 'Any TPM is above the minimum'.
+
+  # cir_plot_filter %>% group_by(Geneid) %>% summarise(n())
+  if(log_it){
+    cir_plot_filter$tpm <- log2(1+cir_plot_filter$tpm)
+    limits = log2(1+limits)
+    breaks <- round(seq.int(0, 
+                            max(limits), 
+                            length.out = round(max(limits)) + 1, 
+                            digits = 0), 
+                    digits = 0)
+  }else{
+    breaks <- seq.int(0, max(limits), length.out = 11)
+  }
+
+  cir_plot_filter$Geneid <- factor(
+    str_remove(cir_plot_filter$Geneid, 
+               pattern = 'PCHAS_'),
+    levels = str_remove(levels(cir_plot_filter$Geneid), 
+                        pattern = 'PCHAS_')
+  )
+
+  lapply(unique(cir_plot_filter$transmission), 
+         function(transmission_method){
+
+           cir_plot_filter <- filter(cir_plot_filter, 
+                              transmission %in% transmission_method)
 
            y <- ggplot(
-      data = cir_plot,
+             data = cir_plot_filter,
              mapping = aes(
-        x = time,
-        fill = transmission,
-        y = sum_tpm
+               x = Geneid,
+               y = tpm
              )
            ) +
+             stat_summary(fun.y = median, 
+                          fun.ymin = median, 
+                          fun.ymax = median,
+                          geom = "bar",
+                          fill = 'lightblue',
+                          width = 0.75, 
+                          size = 0.25, 
+                          position = 'dodge') +
+             geom_point(stat = 'identity', 
+                        position = position_dodge(width = 0.75),
+                        size = point_size) +
+             scale_y_continuous(breaks = breaks,
+                                limits = limits) +
+             facet_grid(.~subfam, 
+                        scales = "free_x",
+                        space = "free_x",
+                        labeller = labeller(subfam = function(string) {
+                          str_remove_all(string, 
+                                         pattern = '(rMT|SBP)_')
+                        })) +
+             theme_classic() +
+             ylab('log(TPM + 1)')+
+             ggtitle(paste(unique(cir_plot_filter$time)))
+
+           if(grepl(time_vec, pattern = '20|23')) {
+             y <- y + theme(
+               axis.text.x = element_text(size = text_size,
+                                          angle = 90,
+                                          vjust = 0.5)
+             )
+           } else {
+             y <- y + theme(
+               axis.text.x = element_blank(),
+               axis.ticks.x = element_blank()
+             ) 
+           }
+
+           #This bit is to change the colour, and text colour, of the facet boxes
+
+           g <- ggplot_gtable(ggplot_build(y))
+
+           # g$widths[5] = 2*g$widths[5]
+
+           strip_t <- which(grepl('strip-t', g$layout$name))
+           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
+           fills <- case_when(str_detect(subfam_vec, 
+                                         pattern = 'a') ~ clade_col['ancestral'],
+                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
+                              str_detect(subfam_vec, 
+                                         pattern = 'ref') ~ 'white')
+           k <- 1
+           for (i in strip_t) {
+             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
+             k <- k+1
+           }
+
+           #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
+           # manually make them a bit wider.
+
+           short_panels <- which(str_detect(subfam_vec, pattern = 'a|S1'))
+
+           for(i in short_panels){
+             num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
+             g$widths[num] = 1.75*g$widths[num]
+             #1.75 times the width of the right panels.
+             # print(num)
+           }
+
+           if(save){
+             ggsave2(
+               paste0('tl19-07_barchart_',
+                      transmission_method,
+                      '_',
+                      'facetSubfam_',
+                      time_vec,
+                      file_name_end,
+                      '.png'),
+               plot = g,
+               dpi = 300,
+               width = 28,
+               height = 8,
+               units = 'cm'
+             )
+           }
+
+           grid.newpage()
+           grid.draw(g)
+         })
+
+}
+
+# transmissionSep_cir_plot(
+#   sort(unique(cir_plot$time))[1],
+#   threshold_min = 50,
+#   threshold_max = Inf,
+#   point_size = 1,
+#   text_size = 6,
+#   file_name_end = '_Over50TPM',
+#   save = FALSE)
+
+lapply(
+  sort(unique(cir_plot$time)),
+  transmissionSep_cir_plot,
+  threshold_min = 50,
+  threshold_max = Inf,
+  point_size = 1,
+  text_size = 7.5,
+  file_name_end = '_Over50TPM',
+  save = TRUE
+)
+```
+
+```{r ancestral_expression_barchart}
+
+# housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
+#                              'MSP1' = 'PCHAS_0831300',
+#                              'AMA1' = 'PCHAS_0931000')
+
+cir_plot <- twentyfour_ct_comb$tpm %>% 
+  filter(Geneid %in% c('PCHAS_0101200', 
+                       'PCHAS_0931000')) %>%
+  melt(id.vars = 'Geneid', 
+       variable.name = 'sample', 
+       value.name = 'tpm') %>% 
+  mutate(
+    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
+    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
+    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
+    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
+    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
+  ) %>% 
+  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
+  filter(!(time %in% c('sch.culture', 
+                       'Unknown')),
+         transmission == 'rMT') %>% 
+  mutate(subfam = str_replace(subfam, 
+                              'ancestral',
+                              'a')) %>% 
+  mutate(
+    subfam = interaction(
+      transmission,
+      forcats::fct_shift(
+        relevel(
+          as.factor(subfam),
+          ref = 'ref')
+      ),
+      sep = '_',
+      lex.order = TRUE
+    )
+  )
+
+anc <- ggplot(data = cir_plot,
+       aes(x = time,
+           y = tpm,
+           fill = Geneid)) +
   stat_summary(fun.y = median, 
@@
   geom_point(stat = 'identity', 
-               position = position_dodge(width = 0.75),
-               size = point_size) +
-    scale_y_continuous(breaks = breaks,
-                       limits = limits) +
-    theme_classic() +
-    theme(
-      axis.text.x = element_text(angle = 90,
-                                 size = 5)
-    ) +
-    ggtitle(subfam_nam)
-
-    ggsave2(
-      paste0('tl20-05_total-expression_by-subfam_',
-             subfam_nam,
-             '.png'),
-      plot = y,
+             position = position_dodge(width = 0.75)) +
+  scale_y_continuous(n.breaks = 11) +
+   scale_fill_manual(values = cbPalette[c(5,8)],
+                     labels = c('ancestral pir', 
+                                'AMA1')) +
+  xlab('Time point') + ylab('TPM') +
+  theme_classic()
+
+ggsave(filename = 'tl19-07_barchart_tpm_RMT_ancestralpirAMA1.png',
+       plot = anc,
        dpi = 300,
-      width = 25,
+       width = 20,
        height = 10,
-      units = 'cm'
-    )
-  return(y)
-  })
-```
-
-# Pir expression of AAPLs and ChAPLs.
-
-```{r AapL-ChaAPl_expression}
+       units = 'cm')
+
+```
+
+```{r rmt_pirs_higher_than_sbp_are_expressed_more}
+cir_rmt_sbp_hi <- twentyfour_ct_comb$tpm %>% 
+  filter(Geneid %in% cir_id) %>%
+  melt(id.vars = 'Geneid', 
+       variable.name = 'sample', 
+       value.name = 'tpm') %>% 
+  mutate(
+    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
+    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
+    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
+    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
+    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
+  ) %>% 
+  filter(time == '14h') %>% 
+  group_by(transmission, time, Geneid) %>% 
+  summarise(tpm_mean = mean(tpm)) %>% 
+  dcast(Geneid ~ transmission, margins = c('tpm_mean')) %>% 
+  mutate(diff = ifelse(rMT - SBP > 0, 'RMT', 'SBP')) %>% 
+  melt(id.vars = c('Geneid', 'diff'), 
+       variable.name = 'transmission',
+       value.name = 'tpm_mean')
+
+cir_rmt_sbp_hi %>% 
+  group_by(diff) %>% 
+  summarise(tpm_mean = mean(tpm_mean))
+
+ggplot(cir_rmt_sbp_hi,
+       aes(x = diff,
+           y = log2(tpm_mean+1),
+           col = transmission)) +
+  geom_point(position = 'jitter') + geom_boxplot()
+```
+
+# Supp figure - Pir expression of AAPLs and ChAPLs.
+
+```{r Supp_fig_AapL-ChaAPl_expression}
 
@@
                                      levels = unique(cir_plot_order$Geneid)))
+  cir_plot_chapl_aapl_supp_info <- cir_plot
+
+  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))
+
+  breaks = seq(0, max_aapl_chapl, by = 1)
+    limits = c(0, max_aapl_chapl)
+    point_size = 0.25
+
+  lapply(
+    sort(unique(cir_plot$time)),
+    function(time_vec){
+
+    cir_plot <- filter(cir_plot, 
+                       time == time_vec)
+
+      cir_plot$tpm <- log2(1+cir_plot$tpm)
+
+    cir_plot$Geneid <- factor(
+      str_remove(cir_plot$Geneid, 
+                 pattern = 'PCHAS_'),
+      levels = str_remove(levels(cir_plot$Geneid), 
+                          pattern = 'PCHAS_')
+    )
+
+    y <- ggplot(
+      data = cir_plot,
+      mapping = aes(
+        x = Geneid,
+        fill = transmission,
+        y = tpm
+      )
+    ) +
+    stat_summary(fun.y = median, 
+                 fun.ymin = median, 
+                 fun.ymax = median,
+                 geom = "bar", 
+                 width = 0.75, 
+                 size = 0.25, 
+                 position = 'dodge') +
+    geom_point(stat = 'identity', 
+               position = position_dodge(width = 0.75),
+               size = point_size) +
+    scale_y_continuous(breaks = breaks,
+                       limits = limits) +
+      scale_fill_manual(values = c('skyblue','red')) +
+      ylab('log(TPM +1)') +
+    facet_grid(.~chapl_aapl, 
+             scales = "free_x",
+             space = "free_x") +
+    theme_classic() +
+    theme(
+      axis.text.x = element_text(angle = 90,
+                                 size = 5)
+    ) +
+    ggtitle(paste(unique(cir_plot$time)))
+
+    ggsave2(
+      paste0('tl20-05_side-by-side_barchart_chapl_aapl_',
+             time_vec,
+             '.png'),
+      plot = y,
+      dpi = 300,
+      width = 25,
+      height = 10,
+      units = 'cm'
+    )
+  return(y)
+})
+```
+
+# Chromosome 6 expression, split by ChAPL designation (mostly a ChAPL)
+
+```{r Chr6-ChaAPl_expression}
+
+  cir_plot <- twentyfour_ct_comb$tpm %>% 
+    filter(Geneid %in% cir_id) %>%
+    melt(id.vars = 'Geneid', 
+         variable.name = 'sample', 
+         value.name = 'tpm') %>% 
+    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
+           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
+           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
+           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
+           subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
+           chrom = str_extract(Geneid, pattern = '(?<=PCHAS_)[:digit:]{2}'),
+           chapl_aapl = factor(cir_chapl_aapl_info$chapl_aapl[match(Geneid, 
+                                                                    cir_chapl_aapl_info$Geneid)],
+                               levels = sort(unique(cir_chapl_aapl_info$chapl_aapl[match(Geneid, 
+                                                                                         cir_chapl_aapl_info$Geneid)])))) %>% 
+    filter(!(time %in% c('sch', 
+                         'Unknown')),
+             chrom == '06') %>% 
+    group_by(Geneid) %>% 
+    # filter(
+    #   # all(tpm < 100),
+    #   #      any(tpm > 50)
+    #   subfam == 'S7'
+    #   ) %>% 
+    arrange(desc(tpm))
+
+  cir_plot_order <- cir_plot %>% 
+    filter(time == '14h',
+           transmission == 'SBP') %>% 
+    arrange(desc(tpm))
+
+  cir_plot <- mutate(cir_plot, 
+                     Geneid = factor(Geneid, 
+                                     levels = unique(cir_plot_order$Geneid)))
 
@@
 
-      cir_plot$tpm <- log1p(cir_plot$tpm)
+      cir_plot$tpm <- log2(1+cir_plot$tpm)
 
@@
     ggsave2(
-      paste0('tl20-05_side-by-side_barchart_chapl_aapl_',
+      paste0('tl20-05_side-by-side_barchart_chr6_chapl_',
              time_vec,
@@
 
-# Pir expression across the genome
-
-```{r pirs_across_genome}
-
-  cir_plot <- twentyfour_ct_comb$tpm %>% 
-    filter(Geneid %in% cir_id) %>%
-    melt(id.vars = 'Geneid', 
-         variable.name = 'sample', 
-         value.name = 'tpm') %>% 
-    mutate(
-      stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-      transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-      diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
-      time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
-      subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
-      chrom = paste0('chr',
-                     str_extract(Geneid, 
-                                 pattern = '(?<=PCHAS_)[[:digit:]]{2}'))
-    ) %>% 
-  filter(!(time %in% c('sch', 
-                       'Unknown')),
-         # transmission == 'rMT'
-         ) %>% 
-  group_by(Geneid)
-
-breaks <- seq.int(0, mround(max(cir_plot$tpm), base = 100), length.out = 6)
-limits <- c(0, mround(max(cir_plot$tpm), base = 100))
-
-lapply(sort(unique(cir_plot$time)), 
-       function(time_vec){
-         cir_plot <- filter(cir_plot, 
-                            time == time_vec)
-
-           # cir_plot$tpm <- log1p(cir_plot$tpm)
-
-         cir_plot$Geneid <- str_remove(cir_plot$Geneid, 
-                                       pattern = 'PCHAS_')
-
-         y <- ggplot(
-           data = cir_plot,
-           mapping = aes(
-             x = Geneid,
-             fill = transmission,
-             y = tpm
-           )
-         ) +
-           stat_summary(fun.y = median, 
-                        fun.ymin = median, 
-                        fun.ymax = median,
-                        geom = "bar", 
-                        width = 0.75, 
-                        size = 0.25, 
-                        position = 'dodge') +
-           scale_y_continuous(breaks = breaks,
-                              limits = limits) +
-           scale_fill_manual(
-             values = c('rMT' = 'blue', 
-                        'SBP' = 'red')
-           ) +
-           facet_grid(.~chrom,
-                    scales = "free_x",
-                    space = "free_x") +
-           theme_classic() +
-           theme(
-             axis.text.x = element_blank(),
-             axis.ticks.x = element_blank(),
-             panel.spacing = unit(0.1, "lines")
-           ) +
-           ggtitle(paste(unique(cir_plot$time)))
-
-         # ggsave2(
-         #   paste0('tl20-05_side-by-side_barchart_genome_facetChrom_',
-         #          time_vec,
-         #          '.png'),
-         #   plot = y,
-         #   dpi = 300,
-         #   width = 25,
-         #   height = 10,
-         #   units = 'cm'
-         # )
-
-         return(y)
-       })
-
-```
-
-# How does expression of the different clades change with time?
-
-```{r clade_with-time_graph}
-cir_plot <- twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% cir_id_no_anc_or_rodcon) %>% 
-  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
-  mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
-  mutate(diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3]) %>% 
-  mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
-  filter(!(time %in% c('sch', 'Unknown'))) %>% 
-  mutate(clade = reid_cir_clade$clade[match(Geneid, reid_cir_clade$Geneid)]) %>% 
-  group_by(clade, transmission, time, diff_count) %>% 
-  dplyr::summarise(median_tpm = median(tpm),
-                   sd_tpm = sd(tpm),
-                   n = n_distinct(Geneid))
-
-# group_by(reid_cir_clade, clade) %>% 
-#   dplyr::summarise(n())
-# 
-# cir_id[which(!(cir_id %in% reid_cir_clade$Geneid))]
-
-order_diff_counts <- c("Early.Rings",
-                       "Early.Mid.Rings",
-                       "Late.Mid.Rings",
-                       "Late.Rings",
-                       "Rings.Troph.conversion",
-                       "Early.Trophs",
-                       "Mid.Trophs",
-                       "Late.Trophs")
-cir_plot$diff_count <- factor(cir_plot$diff_count, levels  = order_diff_counts)
-
-p <- ggplot(data = cir_plot, 
-            aes(x = diff_count, 
-                y = log1p(median_tpm),
-                fill = clade)) +
-  geom_col(position = position_dodge()) +
-  scale_fill_manual(values = cbPalette) +
-  ylab(as.expression(bquote('Median'~italic('pir')~'log TPM + 1 by clade'))) +
-  theme_classic() +
-  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) + 
-  facet_grid(. ~ transmission)
-p
-
-cir_plot <- twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
-  mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
-  mutate(diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3]) %>% 
-  mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
-  filter(!(time %in% c('sch', 'Unknown'))) %>% 
-  mutate(clade = reid_cir_clade$clade[match(Geneid, reid_cir_clade$Geneid)]) %>% 
-  mutate(clade = ifelse(grepl(clade, pattern = 'L'), 'Long', 
-                        ifelse(grepl(clade, pattern = 'S'), 'Short',
-                               'Unk'))) %>% 
-  filter(clade != 'Unk')
-
-order_diff_counts <- c("Early.Rings",
-                       "Early.Mid.Rings",
-                       "Late.Mid.Rings",
-                       "Late.Rings",
-                       "Rings.Troph.conversion",
-                       "Early.Trophs",
-                       "Mid.Trophs",
-                       "Late.Trophs")
-cir_plot$diff_count <- factor(cir_plot$diff_count, levels  = order_diff_counts)
-
-p <- ggplot(data = cir_plot, 
-            aes(x = diff_count,
-                y = log1p(tpm),
-                fill = clade,
-                col = transmission)) +
-  # geom_col(position = position_dodge()) +
-  geom_boxplot() +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = c('dark blue','dark red')) +
-  xlab('Stage') +
-  ylab(as.expression(bquote('Median'~italic('pir')~'log TPM + 1 by clade'))) +
-  theme_classic() +
-  theme(axis.text.x=element_text(angle = 90, vjust = 0.5))
-p
-
-```
-
-The only time that the SBP pir genes are being expressed higher than rMT are the burst of Short-forms in the Early Ring stage.
-
-Splitting the _pir_ TPM graphs into S or L clades, and the sub-clades.
-
-```{r total_tpm_pir_clades}
-pir_clade_list <- lapply(unique(reid_cir_clade$clade), function(clade_name){
-  reid_cir_clade$Geneid[reid_cir_clade$clade %in% clade_name]
-})
-names(pir_clade_list) <- as.character(unique(reid_cir_clade$clade))
-
-#c can cat lists too!
-
-total_tpm_expression(pir_clade_list, 
-                     save = FALSE)
-```
-
-L4s and L1s appear to be particularly high in the rMT samples.
-
-```{r total_tpm_pir_clades_SorL}
-pir_SorL <- lapply(c('S','L'), function(clade_name){
-  reid_cir_clade$Geneid[grepl(reid_cir_clade$clade, pattern = clade_name)]
-})
-names(pir_SorL) <- c('S','L')
-total_tpm_expression(pir_SorL, save = FALSE)
-```
-
-#Expression of pirs from different chromosomes.
-
-```{r total_tpm_pir_chrom}
-pir_chrom_list <- lapply(unique(cir_chrom_tbl$chrom), 
-                         function(chrom_name){
-  cir_chrom_tbl$Geneid[cir_chrom_tbl$chrom %in% chrom_name]
-})
-names(pir_chrom_list) <- as.character(
-  paste0(
-    'chromosome ',
-    unique(cir_chrom_tbl$chrom)
-  )
-)
-
-total_tpm_expression(pir_chrom_list, save = FALSE)
-```
-
-Many big differences across the chromosomes. I wonder if this is consistent with Spence 2013 and Brugat 2017 data?
-
-Early rings: Chr01/07/08/09/10 expressed higher in SBP. Chr05 expressed higher in SBP.
-
-# Looking at the ETRAMP family which shows interesting expression patterns
-
-```{r etramps}
-etramp_info <- chabaudi_gene_info %>% 
-  select(Geneid, Description, Transcript.Length) %>% 
-  filter(Description == 'early transcribed membrane protein')
-etramps <- unique(etramp_info$Geneid)
-
-draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
-                          genelist = etramps,
-                          scale = 'max',
-                          threshold = 0,
-                          log_it = TRUE)
-gene_graph_24h(genevec = etramps, genename = 'ETRAMPs')
-
-twentyfour_ct_etramps <- twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% etramps) %>% 
-  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-  mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
-  mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1]) %>%
-  mutate(time = str_split_fixed(sample, pattern = '_', n = 3)[,2]) %>% 
-  #Convert the times given in 'h' to robustly coded times in 24h clock
-  mutate(time = ifelse(grepl(time, pattern = '[[:digit:]]h$'), 
-                       strptime(time, format="%Hh") %>% format(., "%Hh"),
-                       time)) %>% 
-  mutate(time = ifelse(grepl(time, pattern = '[[:digit:]](a|p)m'), 
-                       strptime(time, format="%I%p") %>% format(., "%Hh"),
-                       time)) %>% 
-  group_by(sample, stage, transmission, time) %>% 
-  dplyr::summarise(sum_tpm = sum(tpm))
-
-p <- ggplot(data = twentyfour_ct_etramps, aes(x = time, 
-                                          y = log1p(sum_tpm),
-                                          fill = transmission)) +
-  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
-               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
-  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
-  scale_fill_manual(values = cbPalette) +
-  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) +
-  theme_classic() + 
-  ggtitle('24h chabaudi etramp expression')
-p
-```
-
-When one ETRAMP goes down, does another replace it?
-
-# Expression of HSP70
-
-Often used as a housekeeping gene.
-
-```{r hsp70}
-total_tpm_expression(list_genes=list(hsp70 = 'PCHAS_0721000'),
-                     save = TRUE)
-```
-
-# Calculate and cluster with z-scores {.tabset}
-
-z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:
-
-1. Convert the count/RPKM values of each gene into log values.
-2. Calculate the mean and standard deviation of X gene log values in each sample
-3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
-4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.
-
-* Need to make sure that the z-score is calculated within each transmission method!
-
-```{r zscores}
-tpm_to_zscore = function(input){
-
-  #Can use vars(matches(paste(input$experiment))) to get the columns with expression data
-  # input_logged = input$tpm %>% 
-  #   mutate_at(vars(matches(paste(input$experiment))), log1p)
-
-  #If the input does not have a TPM element then assume that the input itself is the TPM object
-  if(any(names(input) %in% 'tpm')){
-    tpm_object <- input$tpm
-  }else{
-    tpm_object <- input
-    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
-  }
-
-  #Want all of them so will just ignore Geneid
-  input_logged = tpm_object %>%
-    mutate_at(vars(-matches('Geneid')), 
-              log1p)
-
-  gene_mean_scores = input_logged %>% 
-    melt(id.vars = 'Geneid') %>% 
-    group_by(Geneid) %>% 
-    dplyr::summarise(mean = mean(value),
-                     std_dev = sd(value))
-
-  gene_vector = as.character(input_logged$Geneid)
-
-  input_z = lapply(gene_vector, function(gene){
-    mean_gene = filter(gene_mean_scores, 
-                       Geneid == paste(gene)) %>% 
-      select(mean)
-    std_gene = filter(gene_mean_scores, 
-                      Geneid == paste(gene)) %>% 
-      select(std_dev)
-    gene_scores = input_logged %>% 
-      filter(Geneid == paste(gene)) %>% 
-      select(-Geneid)
-    if(std_gene != 0){
-      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
-    }else{
-      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
-    }
-    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
-    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
-  })
-
-  #Use reduce and rbind to turn the list into one df
-  input_z_df = purrr::reduce(input_z, 
-                             rbind)
-  return(list(zscores = input_z_df,
-              tpm = tpm_object,
-              experiment = input$experiment))
-}
-
-#Make and save z-scores
-# twentyfour_z <- lapply(twentyfour_ct, 
-# tpm_to_zscore)
-
-### Calculate and write zscores
-
-# twentyfour_z <- lapply(twentyfour_avgtpm_noschz,
-#                        tpm_to_zscore)
-# 
-# lapply(twentyfour_z, function(list_element){
-#   z_df <- list_element$zscore
-#   write_csv(x = z_df,
-#             paste0('tl19-07_chabaudi24h_', list_element$experiment, '.csv'))
-# })
-
-###
-
-### Read z-scores
-
-twentyfour_z <- list(
-  'rMT' = list(zscores = read_csv('tl19-07_chabaudi24h_rMT.csv'),
-               tpm = twentyfour_avgtpm_noschz$rMT$tpm,
-               experiment = 'rMT'),
-  'SBP' = list(zscores = read_csv('tl19-07_chabaudi24h_SBP.csv'),
-               tpm = twentyfour_avgtpm_noschz$SBP$tpm,
-               experiment = 'SBP')
-)
-
-###
-```
-
-## Looking at all genes which have a z score over 2.4
-
-```{r over2z}
-over2_z_genes <- lapply(twentyfour_z, function(list_experiments){
-  list_experiments$zscores %>%  
-  filter_at(vars(-matches('Geneid')), 
-            any_vars(. > 2.4)) %>% 
-  .$Geneid
-})
-
-draw_heatmap_chabaudi_24h(datasets = twentyfour_ct, 
-                          genelist = unlist(over2_z_genes),
-                          scale = 'max',
-                          threshold = 0,
-                          log_it = FALSE)
-```
-
-Hmm so the largest z scores aren't necessarily the most variable according to the correlation plot? Is this because they often only vary greatly in one sample?
-
-# Phaseograms of the data. {.tabset}
-
-```{r row_order_phaseograms}
-
-#Getting the row order for each phaseogram for each experiment
-# Want the data ordered by the highest zscore for each gene in each stage.
-phaseogram_row_order <- lapply(twentyfour_z, 
-                               function(list_elements){
-                                 zscore_df <- list_elements$zscores
-                                 zscore_ordered <- melt(zscore_df, 
-                                                        id = 'Geneid',
-                                                        variable.name = 'stage',
-                                                        value.name = 'zscore') %>% 
-                                   group_by(Geneid) %>% 
-                                   top_n(zscore, n = 1) %>% 
-                                   arrange(stage, 
-                                           desc(zscore)) %>%
-                                   distinct(Geneid, 
-                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-                                   filter(zscore > 0) #For some reason some remain with zero zscore
-                                 zscore_ordered <- zscore_ordered[order(
-                                   match(
-                                     zscore_ordered$stage, 
-                                     order_vector
-                                   )
-                                 ),]
-                                 return(c(list_elements, 
-                                          list(zscore_order = zscore_ordered)))
-                               })
-
-phaseogram_time_order <- lapply(names(phaseogram_row_order), 
-                                function(list_element){
-                                  transmission_list <- getElement(phaseogram_row_order,
-                                                                  list_element)
-                                  order_by_time <- transmission_list$zscore_order %>% 
-                                    mutate(stage = times_stages_table$time[match(stage,
-                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
-                                    arrange(stage)
-                                  return(list(zscore_order = order_by_time))
-                                }
-)
-names(phaseogram_time_order) <- names(phaseogram_row_order)
-
-```
-
-~~Change the lnTPM graph to a heatmap of max/min TPM a la Brugat 2017.~~
-
-~~Add annotation for the clades,~~ or possibly separate the heatmap by clade.
-
-Add threshold option.
-
-```{r function_draw_heatmap_zscore}
-
-draw_heatmap_zscore <- function(zscore_list,
-                                title = '',
-                                gene_ordering_vector = NULL,
-                                filter_genes_vector = NULL,
-                                extra_title = NULL,
-                                annotation_clade = NULL,
-                                annotation_col,
-                                threshold_tpm = NULL,
-                                use_times = FALSE,
-                                use_order_vector = TRUE,
-                                tpm_mat = TRUE,
-                                tpm_breaks = NULL,
-                                ...){
-  if(is.null(tpm_breaks) & tpm_mat){
-    #Need the generate the breaks in the data for the heatmap legend if not provided.
-    max_tpm <- lapply(zscore_list, 
-                      function(list_element){
-                        #list_element <- zscore_list[[1]]
-                        heatmap_data <- as.data.frame(list_element$zscores)
-                        title <- list_element$experiment
-
-                        if(!'Geneid' %in% colnames(heatmap_data)){
-                          stop('No Geneid column in the heatmap data.frame')
-                        }
-
-                        if(!is.null(filter_genes_vector)){
-                          heatmap_data <- as.data.frame(
-                            filter(heatmap_data, 
-                                   Geneid %in% filter_genes_vector)
-                          )
-                        }
-
-                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                        if(!is.null(gene_ordering_vector)){
-                          #Remove genes not in the main data frame but in the ordering vector, 
-                          # and then order based on the vector
-                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          heatmap_data <- heatmap_data[na.omit(
-                            match(gene_ordering_vector, 
-                                  rownames(heatmap_data))
-                          ),,
-                          drop=FALSE]
-                        }
-
-                        geneids <- rownames(heatmap_data)
-
-                        #Get max/min TPM scores for the side annotation heatmap                       
-                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                          list_element$tpm$Geneid),] %>% 
-                          select(-Geneid) %>% 
-                          as.matrix %>% 
-                          log1p
-                        rownames(tpm_mat) <- geneids
-
-                        tpm_min <- rowMins(tpm_mat)
-                        names(tpm_min) <- geneids
-                        tpm_max <- rowMaxs(tpm_mat)
-                        names(tpm_max) <- geneids
-                        return(list(tpm_max = tpm_max,
-                                    tpm_min = tpm_min))
-                      })
-
-    tpm_max_all <- mround(
-      max(
-        unlist(
-          lapply(max_tpm, 
-                 getElement,
-                 'tpm_max')
-        )
-      ), 
-      base = 1)
-
-    tpm_breaks <- round(
-      c(0,
-        tpm_max_all/4,
-        tpm_max_all/2,
-        tpm_max_all*(3/4),
-        tpm_max_all),
-      digits = 2
-    )
-  }
-
-  #Generate the colour function from the TPM legend breaks for the TPM annotation
-  if(tpm_mat){
-    tpm_col_fun <- colorRamp2(
-      breaks = tpm_breaks,
-      colors = viridis(length(tpm_breaks))
-    )
-  }
-
-  if(is.numeric(threshold_tpm)){
-    #Apply a threshold to the TPM in order to remove low expressed genes.
-    # Remember to add log1p here! 
-    # Threshold uses the max value, so it ony needs to be above thresh 
-    #   in one stage avg in one transmission method.
-    #Will alter the gene_filter vector in order to only select the genes above threshold
-
-    gene_maxs <- bind_rows(
-      lapply(max_tpm, 
-             getElement,
-             'tpm_max')
-    )
-
-    remove_genes <- names(gene_maxs)[apply(
-      gene_maxs, 
-      2, 
-      function(maxs){
-        all(maxs < log1p(threshold_tpm))
-      }
-    )]
-
-    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
-
-  }
-
-  if(is.character(annotation_clade)){
-    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
-    clade_info <- getElement(cir_clade_info,
-                             annotation_clade)[match(order_cir, 
-                                                     cir_clade_info$Geneid)]
-    names(clade_info) <- str_split_fixed(order_cir, 
-                                         pattern = '_', 
-                                         n = 2)[,2]
-    annotation_col <- annotation_col
-  }
-
-  heatmaps <- lapply(zscore_list, 
-                     function(list_element){
-                       #list_element <- zscore_list[[1]]
-                       heatmap_data <- as.data.frame(list_element$zscores)
-                       title <- list_element$experiment
-
-                       if(!'Geneid' %in% colnames(heatmap_data)){
-                         stop('No Geneid column in the heatmap data.frame')
-                       }
-
-                       if(!is.null(filter_genes_vector)){
-                         heatmap_data <- as.data.frame(
-                           filter(heatmap_data, 
-                                  Geneid %in% filter_genes_vector)
-                         )
-                       }
-
-                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                       if(!is.null(gene_ordering_vector)){
-                         #Remove genes not in the main data frame but in the ordering vector,
-                         #  and then order based on the vector
-                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                         heatmap_data <- heatmap_data[na.omit(
-                           match(gene_ordering_vector, 
-                                 rownames(heatmap_data))
-                         ),,
-                         drop=FALSE]
-                       }
-
-                       if(tpm_mat){
-                         #Get max/min TPM scores for the side annotation heatmap                       
-                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                           list_element$tpm$Geneid),] %>%
-                           select(-Geneid) %>% 
-                           as.matrix %>% 
-                           log1p
-
-                         tpm_min <- rowMins(tpm_mat)
-                         tpm_max <- rowMaxs(tpm_mat)
-
-                         names(tpm_min) = rownames(heatmap_data)
-                         names(tpm_max) = rownames(heatmap_data)
-
-                         tpm_ha <- rowAnnotation(
-                           min.max = anno_simple(cbind(tpm_min, 
-                                                       tpm_max), 
-                                                 col = tpm_col_fun,
-                                                 gp = gpar(col = 'black',
-                                                           lwd = 0.1),
-                                                 border = TRUE,
-                                                 simple_anno_size = unit(0.2, "cm")),
-                           show_annotation_name = FALSE)
-                       }
-
-                       #Draw the left annotation is needed, only for the first element.
-
-                       if(
-                         identical(list_element,
-                                   first(zscore_list)) & is.character(annotation_clade)
-                       ){
-                         clade_ha <- rowAnnotation(
-                           'subfamily' = anno_simple(
-                             clade_info,
-                             col = annotation_col
-                           )
-                         )
-                       }
-
-                       #Change rownames to remove PCHAS_
-                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
-                                                                pattern = '_', 
-                                                                n = 2)[,2]
-
-                       if(use_times){
-                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
-                                                                                 times_stages_table$r_stage_diff_count)]
-                       }
-
-                       if(use_order_vector){
-                         #Order the columns by ordering_vector
-                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
-                                                                   order_vector)),
-                                                      drop = FALSE]
-                       }else{
-                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
-                                                      drop = FALSE] 
-                       }
-
-                       h = Heatmap(heatmap_data,
-                                   name = paste0('z-score',
-                                                 extra_title),
-                                   col = c(cbPalette[6],
-                                           'white', 
-                                           cbPalette[7]),
-                                   row_names_gp = gpar(fontsize = ifelse(
-                                     nrow(heatmap_data) > 13, 
-                                     12*(13/nrow(heatmap_data)), 
-                                     12)
-                                   ),
-                                   column_names_gp = gpar(fontsize = 10),
-                                   row_dend_width = unit(3, "cm"),
-                                   right_annotation = if(tpm_mat){
-                                     tpm_ha
-                                   }else{
-                                     NULL
-                                   },
-                                   left_annotation = 
-                                     if(identical(list_element,
-                                                  first(zscore_list)) & is.character(annotation_clade)){
-                                       clade_ha
-                                     }else{
-                                       NULL
-                                     }, 
-                                   #Only want the clade annotation when it is first heatmap
-                                   width = unit(4.5, 'cm'),
-                                   cluster_columns = FALSE,
-                                   height = unit(7,'cm'),
-                                   row_title = NULL,
-                                   cluster_rows = FALSE,
-                                   column_title = paste(title),
-                                   heatmap_legend_param = list(
-                                     at = c(3,2,1,0,-1,-2,-3),
-                                     labels_gp = gpar(fontsize = 10),
-                                     title_gp = gpar(fontsize = 10)
-                                   ),
-                                   ...
-                       )
-                       return(h)
-                     })
-
-  ht_list <- Reduce('+', heatmaps) #To add the list elements together
-
-  if(tpm_mat){
-    #Draw the legend
-    tpm_lgd <- Legend(
-      title = "Min/Max log(TPM + 1)", 
-      col_fun = tpm_col_fun, 
-      at = tpm_breaks, 
-      labels = tpm_breaks,
-      title_gp = gpar(fontsize = 7)
-    )
-  }
-
-  if(is.character(annotation_clade)){
-    clade_lgd <- Legend(
-      title = 'sub-family',
-      at = names(annotation_col),
-      legend_gp = gpar(fill = annotation_col)
-    )
-  }
-
-  draw(ht_list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'left',
-       #Only include the annotation legends for the annotations we actually have.
-       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
-         list(tpm_lgd,
-              clade_lgd)
-       }else if(tpm_mat & !is.character(annotation_clade) ){
-         list(tpm_lgd)
-       } else if(!tpm_mat & is.character(annotation_clade) ){
-         list(clade_lgd)
-       })
-}
-```
-
-## Phaseograms of the entire genome.
-
-```{r phaseograms_whole_genome}
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-```
-
-## Phaseograms of the _cir_ genes with clade annotations.
-
-```{r pir_gene_phaseogram_breaks}
-  max_tpm <- lapply(twentyfour_z, 
-                    function(list_element){
-                      #list_element <- zscore_list[[1]]
-                      heatmap_data <- as.data.frame(list_element$zscores)
-                      title <- list_element$experiment
-
-                      if(!'Geneid' %in% colnames(heatmap_data)){
-                        stop('No Geneid column in the heatmap data.frame')
-                      }
-
-                            heatmap_data <- as.data.frame(
-                              filter(heatmap_data, 
-                                     Geneid %in% cir_id)
-                            )
-
-                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-                          # 
-                          # if(!is.null(gene_ordering_vector)){
-                          #   #Remove genes not in the main data frame but in the ordering vector, 
-                          #   # and then order based on the vector
-                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          #   heatmap_data <- heatmap_data[na.omit(
-                          #     match(gene_ordering_vector, 
-                          #           rownames(heatmap_data))
-                          #   ),,
-                          #   drop=FALSE]
-                          # }
-
-                          geneids <- rownames(heatmap_data)
-
-                          #Get max/min TPM scores for the side annotation heatmap                       
-                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                            list_element$tpm$Geneid),] %>% 
-                            select(-Geneid) %>% 
-                            as.matrix %>% 
-                            log1p
-                          rownames(tpm_mat) <- geneids
-
-                          tpm_min <- rowMins(tpm_mat)
-                          names(tpm_min) <- geneids
-                          tpm_max <- rowMaxs(tpm_mat)
-                          names(tpm_max) <- geneids
-                          return(list(tpm_max = tpm_max,
-                                      tpm_min = tpm_min))
-                        })
-
-  tpm_max_all <- mround(
-    max(
-      unlist(
-        lapply(max_tpm, 
-               getElement,
-               'tpm_max')
-      )
-    ), 
-    base = 1)
-
-  tpm_breaks <- round(
-    c(0,
-      tpm_max_all/4,
-      tpm_max_all/2,
-      tpm_max_all*(3/4),
-      tpm_max_all),
-    digits = 2
-  )
-```
-
-```{r phaseograms_cir_order_each}
-
-#Generate the colour function from the TPM legend breaks for the TPM annotation
-
-tpm_col_fun <- colorRamp2(
-  breaks = tpm_breaks,
-  colors = viridis(length(tpm_breaks))
-)
-
-#Apply a threshold to the TPM in order to remove low expressed genes.
-# Remember to add log1p here! 
-# Threshold uses the max value, so it ony needs to be above thresh 
-#   in one stage avg in one transmission method.
-#Will alter the gene_filter vector in order to only select the genes above threshold
-
-gene_maxs <- bind_rows(
-  lapply(max_tpm, 
-         getElement,
-         'tpm_max')
-)
-
-remove_genes <- names(gene_maxs)[apply(
-  gene_maxs, 
-  2, 
-  function(maxs){
-    all(maxs < log1p(1))
-  }
-)]
-
-filter_genes_vector <- cir_id[!cir_id %in% remove_genes]
-
-heatmaps <- lapply(names(twentyfour_z), 
-                   function(list_name){
-                     list_element <- getElement(twentyfour_z, 
-                                                list_name)
-                     #list_name <- names(twentyfour_z)[[1]]
-                     heatmap_data <- as.data.frame(list_element$zscores)
-                     title <- list_element$experiment
-
-                     heatmap_data <- as.data.frame(
-                       filter(heatmap_data, 
-                              Geneid %in% filter_genes_vector)
-                     )
-
-                     rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                     heatmap_data <- heatmap_data[, 
-                                                  !colnames(heatmap_data) %in% 'Geneid', 
-                                                  drop = TRUE]
-                     #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                     #Remove genes not in the main data frame but in the ordering vector,
-                     #  and then order based on the vector
-                     gene_ordering_vector <- getElement(phaseogram_time_order,
-                                                        list_name)$zscore_order$Geneid
-
-                     order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
-                     clade_info <- getElement(cir_clade_info,
-                                              'SorL')[match(order_cir, 
-                                                            cir_clade_info$Geneid)]
-                     names(clade_info) <- str_split_fixed(order_cir, 
-                                                          pattern = '_', 
-                                                          n = 2)[,2]
-                     annotation_col <- SorL_col
-
-                     remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                     heatmap_data <- heatmap_data[remove_index,
-                                                  ,
-                                                  drop = FALSE]
-                     heatmap_data <- heatmap_data[na.omit(
-                       match(gene_ordering_vector, 
-                             rownames(heatmap_data))
-                     ),
-                     ,
-                     drop=FALSE]
-
-                     #Get max/min TPM scores for the side annotation heatmap                       
-                     tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                       list_element$tpm$Geneid),] %>%
-                       select(-Geneid) %>% 
-                       as.matrix %>% 
-                       log1p
-
-                     tpm_min <- rowMins(tpm_mat)
-                     tpm_max <- rowMaxs(tpm_mat)
-
-                     names(tpm_min) = rownames(heatmap_data)
-                     names(tpm_max) = rownames(heatmap_data)
-
-                     tpm_ha <- rowAnnotation(
-                       min.max = anno_simple(cbind(tpm_min, 
-                                                   tpm_max), 
-                                             col = tpm_col_fun,
-                                             gp = gpar(col = 'black',
-                                                       lwd = 0.1),
-                                             border = TRUE,
-                                             simple_anno_size = unit(0.2, "cm")),
-                       show_annotation_name = FALSE)
-
-                     #Clade annotation
-
-                     clade_ha <- rowAnnotation(
-                       'sub-family' = anno_simple(
-                         clade_info,
-                         col = annotation_col,
-                       ),
-                       annotation_name_gp = gpar(fontsize = 10)
-                     )
-
-                     #Change rownames to remove PCHAS_
-                     rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
-                                                              pattern = '_', 
-                                                              n = 2)[,2]
-
-                     colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
-                                                                             times_stages_table$r_stage_diff_count)]
-
-                     heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
-                                                  drop = FALSE]
-
-                     h = Heatmap(heatmap_data,
-                                 name = paste0('z-score'),
-                                 col = c(cbPalette[6],
-                                         'white', 
-                                         cbPalette[7]),
-                                 row_names_gp = gpar(fontsize = ifelse(
-                                   nrow(heatmap_data) > 13, 
-                                   12*(13/nrow(heatmap_data)), 
-                                   12)
-                                 ),
-                                 column_names_gp = gpar(fontsize = 10),
-                                 row_dend_width = unit(3, "cm"),
-                                 right_annotation = tpm_ha,
-                                 left_annotation = clade_ha,
-                                 width = unit(4.5, 'cm'),
-                                 cluster_columns = FALSE,
-                                 height = unit(7,'cm'),
-                                 row_title = NULL,
-                                 cluster_rows = FALSE,
-                                 column_title = paste(title),
-                                 heatmap_legend_param = list(
-                                   at = c(3,2,1,0,-1,-2,-3),
-                                   labels_gp = gpar(fontsize = 10),
-                                     title_gp = gpar(fontsize = 10)
-                                 ),
-                                 show_row_names = FALSE
-                     )
-                     return(h)
-                   })
-
-ht_list <- Reduce('+', heatmaps) #To add the list elements together
-
-#Draw the legend
-tpm_lgd <- Legend(
-  title = "Min/Max log(TPM + 1)", 
-  col_fun = tpm_col_fun,
-  at = tpm_breaks, 
-  labels = tpm_breaks,
-  title_gp = gpar(fontsize = 7)
-)
-
-clade_lgd <- Legend(
-  title = 'sub-family',
-  at = names(SorL_col),
-  legend_gp = gpar(fill = SorL_col)
-)
-
-png(filename = 'tl19-07_phaseogram_each-in-order_cir_clades.png',
-    units = 'cm',
-    height = 20,
-    width = 22,
-    res = 200)
-
-  draw(ht_list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'left',
-       annotation_legend_list = 
-         list(tpm_lgd,
-              clade_lgd),
-       ht_gap = unit(.5,'cm')
-       )
-
-  dev.off()
-
-```
-
-```{r phaseograms_cir}
-
-phase_rMT <- draw_heatmap_zscore(zscore_list = list(twentyfour_z$rMT) ,
-                    gene_ordering_vector = phaseogram_row_order$rMT$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_breaks = tpm_breaks)
-
-phase_SBP <- draw_heatmap_zscore(zscore_list = list(twentyfour_z$SBP),
-                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_breaks = tpm_breaks)
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_row_order$SBP$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'clade',
-                    annotation_col = clade_col,
-                    threshold_tpm = 200,
-                    use_times = TRUE,
-                    use_order_vector = FALSE)
-
-png(filename = 'tl19-07_phaseogram_rMTorder_cir_clades.png',
-    units = 'cm',
-    height = 20,
-    width = 22,
-    res = 200)
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'clade',
-                    annotation_col = clade_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    tpm_mat = FALSE)
-dev.off()
-
-png(filename = 'tl19-07_phaseogram_rMTorder_cir_SorL.png',
-    units = 'cm',
-    height = 20,
-    width = 22,
-    res = 200)
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-dev.off()
-
-png(filename = 'tl19-07_phaseogram_SBPorder_cir_SorL.png',
-    units = 'cm',
-    height = 20,
-    width = 22,
-    res = 200)
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_breaks = tpm_breaks,
-                    tpm_mat = FALSE)
-dev.off()
-
-```
-
-## Phaseograms by clade.
-
-```{r phaseograms_by_clade}
-cir_clades <- unique(cir_clade_info$clade) %>% sort
-lapply(cir_clades, 
-       function(cir_clade_name){
-         cirs_in_clade <- cir_clade_info$Geneid[cir_clade_info$clade %in% cir_clade_name]
-         if(length(cirs_in_clade) > 1){
-           draw_heatmap_zscore(zscore_list = twentyfour_z,
-                               gene_ordering_vector = phaseogram_row_order$rMT$zscore_order$Geneid,
-                               filter_genes_vector = cirs_in_clade,
-                               extra_title = paste0('_', cir_clade_name),
-                               annotation_clade = TRUE)
-           if(global_save_var){
-             png(filename = paste0('tl19-07_phaseogram_rMTorder_cir_', 
-                                   cir_clade_name,
-                                   '.png'),
-                 units = 'cm',
-                 height = 20,
-                 width = 22,
-                 res = 200)
-             draw_heatmap_zscore(zscore_list = twentyfour_z,
-                                 gene_ordering_vector = phaseogram_row_order$rMT$zscore_order$Geneid,
-                                 filter_genes_vector = cirs_in_clade,
-                                 extra_title = paste0('_', cir_clade_name))
-             dev.off()
-           }
-
-         }
-       })
-```
-
-Putting them side by side you can see the delay with the whole transcriptome - genes upregulated in certain rMT stages appear to be upregulated in both the same stage and the next stage 3h later.
-
-# Heatmap of all of the expressed pir genes in each stage, ordered by each transmission type. - Figure 3b
-
-```{r highest_expressed_stage_heatmap_fig3b}
-
-# Get order of pir genes by highest expressed genes in each sample.
-
-combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm_noschz), 
-       function(list_name){
-
-         combined_data_pir <- getElement(twentyfour_avgtpm_noschz, 
-                                     list_name)$tpm %>% 
-           filter(Geneid %in% cir_id) %>% 
-           column_to_rownames(var = 'Geneid')
-
-         #Ordering that a more understandable order
-
-         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
-                                                            order_vector)),
-                                                drop = FALSE]
-
-         # combined_data_ordered <- melt(
-         #   (
-         #     combined_data_pir %>% 
-         #       rownames_to_column(var = 'Geneid')
-         #   ),
-         #   id = 'Geneid',
-         #   variable.name = 'stage',
-         #   value.name = 'tpm') %>% 
-         #   group_by(Geneid) %>% 
-         #   top_n(tpm, 
-         #         n = 1) %>% 
-         #   arrange(stage,
-         #           desc(tpm)) %>%
-         #   distinct(Geneid, 
-         #            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-         #   filter(tpm > 0)
-
-         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
-                                         means = apply(combined_data_pir, 1, mean),
-                                         subfam = cir_clade_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
-                                                                            cir_clade_info$Geneid)]) %>% 
-           filter(means > 1) %>% 
-           arrange(subfam, desc(means)) %>% 
-           .$Geneid
-
-         combined_data_pir_ht <- combined_data_pir[
-           order(
-             match(rownames(combined_data_pir), 
-                   combined_data_ordered),
-             na.last = NA #Remove NAs
-           ),
-         ]
-
-         return(
-           list(order = combined_data_pir_ht)
-         )
-       })
-names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm_noschz)
-
-lapply(names(combined_data_pir_ht_list),
-       function(list_name_to_order){
-
-         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
-                                            list_name_to_order)$order
-
-         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
-         maxcol <- ceiling(
-           max(log1p(combined_data_pir_ht)
-           )
-         )
-
-         #Sub-family annotation
-         clade_info <- getElement(cir_clade_info,
-                                  'SorL')[match(pir_highest_tpm_order, 
-                                                cir_clade_info$Geneid)]
-         annotation_col <- SorL_col
-
-         #Subfamily annotation
-
-         clade_ha <- rowAnnotation(
-           'sub-family' = anno_simple(
-             clade_info,
-             col = annotation_col,
-           ),
-           annotation_name_gp = gpar(fontsize = 10)
-         )
-
-         anc_ha <- rowAnnotation(foo = anno_mark(at = 1, 
-                                                 labels = 'ancestral',
-                                                 labels_gp = gpar(fontsize = 7),
-                                                 link_width = unit(3, "mm")))
-
-         #lapply across the datasets now, for each order
-         ht_list <- lapply(names(combined_data_pir_ht_list),
-                           function(list_name){
-
-                             combined_data_pir_ht <- getElement(twentyfour_avgtpm_noschz, 
-                                                                list_name)$tpm %>% 
-                               column_to_rownames(var = 'Geneid')
-
-                             #Get the times instead of the stage names.
-                             colnames(combined_data_pir_ht) <- times_stages_table$time[match(colnames(combined_data_pir_ht),
-
-                                                                                             times_stages_table$r_stage_diff_count)]
-
-                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                                          drop = FALSE]
-
-                             Heatmap(
-                               log1p(combined_data_pir_ht)[pir_highest_tpm_order,],
-                               cluster_rows = FALSE,
-                               cluster_columns = FALSE,
-                               show_row_names = FALSE,
-                               name = 'log(TPM+1)', 
-                               col = colorRamp2(c(0, 
-                                                  maxcol/4, 
-                                                  maxcol/2, 
-                                                  3*maxcol/4, 
-                                                  maxcol), 
-                                                viridis(5)),
-                               na_col = 'white', 
-                               border = TRUE,
-                               border_gp = gpar(lwd = 0.1),
-                               column_names_gp = gpar(fontsize = 10),
-                               row_dend_width = unit(3, "cm"),
-                               width = unit(4.5, 'cm'),
-                               height = unit(7,'cm'),
-                               row_title = NULL,
-                               left_annotation = if(
-                                 list_name %in% first(names(combined_data_pir_ht_list))
-                               ){
-                                 clade_ha
-                               }else{
-                                 NULL
-                               },
-                               right_annotation = if(
-                                 list_name %in% last(names(combined_data_pir_ht_list))
-                               ){
-                                 anc_ha
-                               }else{
-                                 NULL
-                               },
-                               # row_split =  cir_clade_info$SorL[match(pir_highest_tpm_order, 
-                               #                                        cir_clade_info$Geneid)],
-                               column_title = paste(list_name),
-                               heatmap_legend_param = list(
-                                 labels_gp = gpar(fontsize = 10),
-                                 title_gp = gpar(fontsize = 10,
-                                                 fontface = 'bold')
-                               )
-                             )
-                           }
-         )
-         ht_list <- Reduce(`+`, ht_list)
-         clade_lgd <- Legend(
-           title = 'sub-family',
-           at = names(SorL_col),
-           legend_gp = gpar(fill = SorL_col)
-         )
-
-         png(filename = paste0('tl19-07_combined_pirs_ordered_highest_stage_expression_heatmap_TPM_',
-                               list_name_to_order, 'order',
-                               '.png'),
-             width = 20,
-             height = 10,
-             units = 'cm',
-             res = 330)
-
-         draw(ht_list, 
-              row_title_side = 'left',
-              column_title_side = 'bottom',
-              heatmap_legend_side = 'right',
-              annotation_legend_side = 'right',
-              annotation_legend_list = list(clade_lgd),
-              ht_gap = unit(.35,'cm'),
-              merge_legend = TRUE
-         )
-
-         dev.off()
-       })
-
-```
-
-```{r highest_genes_stage}
-highest_pirs_stage_transmission <- lapply(names(twentyfour_avgtpm_noschz), 
-       function(list_name){
-
-         combined_data_pir <- getElement(twentyfour_avgtpm_noschz, 
-                                     list_name)$tpm %>% 
-           filter(Geneid %in% cir_id) %>% 
-           column_to_rownames(var = 'Geneid')
-
-         #Ordering that a more understandable order
-
-         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
-                                                            order_vector)),
-                                                drop = FALSE]
-
-         combined_data_ordered <- melt(
-           (
-             combined_data_pir %>%
-               rownames_to_column(var = 'Geneid')
-           ),
-           id = 'Geneid',
-           variable.name = 'stage',
-           value.name = 'tpm') %>%
-           group_by(Geneid) %>%
-           top_n(tpm,
-                 n = 1) %>%
-           arrange(stage,
-                   desc(tpm)) %>%
-           distinct(Geneid,
-                    .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-           filter(tpm > 0)
-
-         highest_pir_stage_list <- lapply(unique(combined_data_ordered$stage), 
-                                          function(stage_name){
-                                            filter(combined_data_ordered, 
-                                                   stage %in% stage_name)$Geneid
-                                          }
-         )
-         names(highest_pir_stage_list) <- unique(combined_data_ordered$stage)
-         return(highest_pir_stage_list)
-       })
-names(highest_pirs_stage_transmission) <- names(twentyfour_avgtpm_noschz)
-```
-
-```{r different_highest_genes_between_transmission}
-stage_names <- ordering(colnames(twentyfour_avgtpm_noschz$rMT$tpm)[-1],
-                        order_vector)
-agree_highest_pirs <- lapply(stage_names, 
-       function(stage_name){
-  agree(
-    lapply(highest_pirs_stage_transmission, 
-           getElement, 
-           stage_name)
-  )
-})
-names(agree_highest_pirs) <- stage_names
-
-disagree_highest_pirs_rMThigher <- lapply(stage_names, 
-       function(stage_name){
-  disagree(
-    lapply(highest_pirs_stage_transmission, 
-           getElement, 
-           stage_name)
-  )
-})
-names(disagree_highest_pirs_rMThigher) <- stage_names
-
-disagree_highest_pirs_SBPhigher <- lapply(stage_names, 
-       function(stage_name){
-  disagree(
-    lapply(rev(highest_pirs_stage_transmission), 
-           getElement, 
-           stage_name)
-  )
-})
-names(disagree_highest_pirs_SBPhigher) <- stage_names
-```
-
-# pir expression by clade/sub-family
-
-```{r highest_genes_clade, results = "asis"}
-lapply(names(highest_pirs_stage_transmission), 
-       function(transmission_name){
-         transmission_highest_pirs <- getElement(highest_pirs_stage_transmission,
-                                                 transmission_name)
-         lapply(names(transmission_highest_pirs), 
-                function(top_pirs_name){
-                  top_pirs <- getElement(transmission_highest_pirs, 
-                                         top_pirs_name)
-                  mutate(cir_clade_info, 
-                         in_top = str_detect(Geneid, 
-                                             pattern = paste(top_pirs, collapse = '|')) ) %>% 
-                    group_by(clade) %>% 
-                    dplyr::summarise(num = sum(in_top),
-                                     prop_of_stage_top_pirs = round(sum(in_top)/length(top_pirs),
-                                                                    digits = 3)*100,
-                                     prop_of_clade = round(sum(in_top)/length(Geneid),
-                                                           digits = 3)*100) %>% 
-                    kable(format = 'html', 
-                          caption = paste0(transmission_name, 
-                                           '_', 
-                                           top_pirs_name), 
-                          escape = FALSE) %>% 
-                    kable_styling()
-                })
-       })
-```
-
-Unknowns always highest in the ring stages.
-
-## Proportion table of S/L clade expression by stage.
-
-```{r proportion_data}
-
-tpm_data_twentyfour <- twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_clade_info$clade[match(Geneid, 
-                                            cir_clade_info$Geneid)],
-         chrom = str_extract(Geneid, 
-                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
-         ) %>% 
-  mutate(
-    SorL = str_extract(clade, 
-                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-    # stage =  factor(stage, 
-    #                 levels = ordering(stage, order_vector))
-  ) %>% 
-  group_by(transmission,
-           stage, 
-           Geneid,
-           clade,
-           chrom,
-           SorL) %>% 
-  dplyr::summarise(tpm = mean(tpm))
-
-```
-
-```{r clade_proportion}
-tpm_data_twentyfour %>% 
-    group_by(transmission,
-           stage, 
-           clade) %>% 
-  dplyr::summarise(max_gene = max(tpm),
-                   total_clade = sum(tpm)) %>%
-  group_by(stage) %>% 
-  dplyr::mutate(total_stage = sum(total_clade)) %>% 
-  dplyr::mutate(
-    prop = round(total_clade/total_stage, digits = 2),
-    prop_from_max_gene = ifelse(total_clade > 0,
-                                round(max_gene/total_clade, digits = 2),
-                                0)
-  ) %>% 
-  # Prop from max gene to show how much of the clade expression is from the highest expressed gene of the clade.
-  filter(prop > 0.0001) %>%
-  dplyr::arrange(transmission, desc(prop)) %>% 
-  select(-max_gene) %>% 
-  kable(format = 'html', 
-        caption = 'Proportion of pir clade expression for each stage', 
-        escape = FALSE) %>% 
-  kable_styling() %>% 
-  scroll_box(width = "100%", height = "500px")
-
-```
-
-## Stacked bar chart of clade expression.
-
-```{r clades_barchart}
-tpm_data_clade <- twentyfour_ct_comb$tpm %>% 
+## Bar chart of S and L clade expression across time.
+
+```{r SorL_barchart}
+tpm_data_SorL <- twentyfour_ct_comb$tpm %>% 
   filter(Geneid %in% cir_id) %>% 
@@
                                             cir_clade_info$Geneid)]) %>% 
-  # mutate(stage =  factor(stage, 
-  #                        levels = ordering(stage, 
-  #                                          order_vector))) %>% 
+  mutate(SorL = str_extract(clade, 
+                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
+         # stage =  factor(stage, levels = ordering(stage, order_vector))
+         ) %>% 
   group_by(transmission,
            stage, 
-           clade) %>% 
-  dplyr::summarise(total_tpm = sum(tpm)) %>% 
-  arrange(stage)
-
-y <- ggplot(data = tpm_data_clade, 
+           SorL) %>% 
+  dplyr::summarise(total_tpm = sum(tpm))
+
+y <- ggplot(data = tpm_data_SorL, 
               aes(
@@
                 y = total_tpm, 
-                fill = clade,
+                fill = SorL,
                 group = transmission
@@
     scale_fill_manual(
-      breaks = unique(tpm_data_clade$clade),
-      values = setNames(
-        c(cbPalette, 
-          'red', 
-          'white')[1:length(unique(tpm_data_clade$clade))],
-        unique(tpm_data_clade$clade)
-      )
+      values = SorL_col,
+      guide = guide_legend(reverse = TRUE)
     ) +
@@
 
-ggsave2('tl20-05_subclades_totalTPM_barchart.pdf',
+ggsave2('tl20-05_clades_totalTPM_barchart.pdf',
         plot = y,
@@
 ```{r clades_prop_barchart}
+tpm_data_clade <- twentyfour_ct_comb$tpm %>% 
+  filter(Geneid %in% cir_id) %>% 
+  melt(id.vars = 'Geneid', 
+       variable.name = 'sample', 
+       value.name = 'tpm') %>% 
+  mutate(stage = str_extract(sample, 
+                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
+  mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
+  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
+  filter(stage != 'sch.culture') %>% 
+  mutate(clade = cir_clade_info$clade[match(Geneid, 
+                                            cir_clade_info$Geneid)]) %>% 
+  # mutate(stage =  factor(stage, 
+  #                        levels = ordering(stage, 
+  #                                          order_vector))) %>% 
+  group_by(transmission,
+           stage, 
+           clade) %>% 
+  dplyr::summarise(total_tpm = sum(tpm)) %>% 
+  arrange(stage)
+
 y <- ggplot(data = tpm_data_clade, 
@@
 
-## S/L clade member proportions among the pirs expressed highest at each stage.
-
-```{r highest_pirs_SorL, results = "asis"}
-lapply(names(highest_pirs_stage_transmission), 
-       function(transmission_name){
-         transmission_highest_pirs <- getElement(highest_pirs_stage_transmission,
-                                                 transmission_name)
-         lapply(names(transmission_highest_pirs), 
-                function(top_pirs_name){
-                  top_pirs <- getElement(transmission_highest_pirs, 
-                                         top_pirs_name)
-                  mutate(cir_clade_info, 
-                         clade = str_extract(clade, pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-                         in_top = str_detect(Geneid, 
-                                             pattern = paste(top_pirs, collapse = '|')) ) %>% 
-                    group_by(clade) %>% 
-                    dplyr::summarise(num = sum(in_top),
-                                     prop_of_stage_top_pirs = round(sum(in_top)/length(top_pirs),
-                                                                    digits = 3)*100,
-                                     prop_of_clade = round(sum(in_top)/length(Geneid),
-                                                           digits = 3)*100) %>% 
-                    kable(format = 'html', 
-                          caption = paste0(transmission_name, 
-                                           '_', 
-                                           top_pirs_name), 
-                          escape = FALSE) %>% 
-                    kable_styling()
-                })
-       })
-```
-
-SBP Late Trophs have highest expression of more Ls than the rMT late trophs.
-This is probably because the expression of the pirs is generally more spread out in the rMT, does this culminate as them having more genes in different phases in the Fourier results?
-
-All the unknown clades are expressed in the earlier ring stages, however I don't think they themselves are expressed much at all.
-
-## Bar chart of S and L clade expression across time.
-
-```{r SorL_barchart}
-tpm_data_SorL <- twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_clade_info$clade[match(Geneid, 
-                                            cir_clade_info$Geneid)]) %>% 
-  mutate(SorL = str_extract(clade, 
-                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-         # stage =  factor(stage, levels = ordering(stage, order_vector))
-         ) %>% 
-  group_by(transmission,
-           stage, 
-           SorL) %>% 
-  dplyr::summarise(total_tpm = sum(tpm))
-
-y <- ggplot(data = tpm_data_SorL, 
-              aes(
-                x = transmission,
-                y = total_tpm, 
-                fill = SorL,
-                group = transmission
-              )) +
-    geom_bar(position = position_stack(),
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-    scale_fill_manual(
-      values = SorL_col,
-      guide = guide_legend(reverse = TRUE)
-    ) +
-    scale_color_manual('black') + 
-    # labs(fill = stack_label) +
-    # scale_y_continuous(breaks = max_tpm) +
-    theme_classic() +
-    scale_size(range = c(0,3)) +
-    ylab(
-      bquote('total '~italic(.('pir'))~' TPM')
-    ) + 
-    facet_grid(.~stage, 
-               scales = "free_x",
-               space = "free_x") +
-    theme(
-      axis.title.x=element_blank(),
-      # axis.text.x=element_blank(),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-ggsave2('tl20-05_clades_totalTPM_barchart.pdf',
-        plot = y,
-        dpi = 300,
-        width = 15,
-        height = 10,
-        units = 'cm')
-
-```
-
-```{r SorL_barchart_transmission-facet}
-tpm_data_SorL <- twentyfour_ct_comb$tpm %>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_clade_info$clade[match(Geneid, 
-                                            cir_clade_info$Geneid)]) %>% 
-  mutate(SorL = str_extract(clade, 
-                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-         # stage =  factor(stage, levels = ordering(stage, order_vector))
-         ) %>% 
-  group_by(transmission,
-           stage, 
-           SorL) %>% 
-  dplyr::summarise(total_tpm = sum(tpm)) %>% 
-  mutate(SorL = factor(SorL, 
-                       levels = sort(unique(SorL))))
-
-y <- ggplot(data = tpm_data_SorL, 
-              aes(
-                x = stage,
-                y = total_tpm, 
-                fill = SorL,
-                # group = transmission
-              )) +
-  geom_bar(position = position_stack(),
-           stat = 'identity',
-           colour = 'black',
-           size = 0.1) +
-  scale_fill_manual(
-    values = SorL_col,
-    guide = guide_legend(reverse = FALSE)
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  # scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~transmission, 
-             scales = "free_x",
-             space = "free_x") +
-  labs(fill = 'sub-family') +
-    theme(
-      axis.title.x=element_blank(),
-      # axis.text.x=element_blank(),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-ggsave2('tl20-05_SorL_totalTPM_facet-transmission_barchart.pdf',
-        plot = y,
-        dpi = 300,
-        width = 15,
-        height = 10,
-        units = 'cm')
-
-```
-
-## S or L proportion of expression bar chart
-
 ```{r add_berghei_SorL_prop}
 #Add berghei data
-# berghei_SorL_file <- list.files('..', 
-#                                 recursive = TRUE, 
-#                                 pattern = 'tl20-05_SorL-subfamily_pir.csv', 
-#                                 full.names = TRUE)
-berghei_SorL <- read_csv('../tl20-05_bergheisystematicanalysis_nextflowoutput/tl20-05_SorL-subfamily_pir.csv') %>% 
-  filter(stage %in% c('Asexual.SchizontA',
-                      'Asexual.SchizontB',
-                      'Asexual.Ring',
-                      'Asexual.Trophozoite',
-                      'Asexual.Mixed')) 
+berghei_SorL_file <- list.files('../Manuscript_pirAnalysis_Github/Pberghei/', 
+                                recursive = TRUE, 
+                                pattern = 'tl20-05_SorL-subfamily_pir.csv', 
+                                full.names = TRUE)
+berghei_SorL <- read_csv(berghei_SorL_file) %>% 
+  filter(stage %in% c('Asex.SchizA',
+                      'Asex.SchizB',
+                      'Asex.Ring',
+                      'Asex.Troph',
+                      'Asex.Mixed')) 
 berghei_SorL_avg <- berghei_SorL %>% 
@@
 
-## S or L proportion of expression.
-
-```{r SorL_proportion_table}
-tpm_data_twentyfour %>% 
+```{r SorL_barchart_transmission-facet}
+tpm_data_SorL <- twentyfour_ct_comb$tpm %>% 
+  filter(Geneid %in% cir_id) %>% 
+  melt(id.vars = 'Geneid', 
+       variable.name = 'sample', 
+       value.name = 'tpm') %>% 
+  mutate(stage = str_extract(sample, 
+                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
+    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
+  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
+  filter(stage != 'sch.culture') %>% 
+  mutate(clade = cir_clade_info$clade[match(Geneid, 
+                                            cir_clade_info$Geneid)]) %>% 
+  mutate(SorL = str_extract(clade, 
+                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
+         # stage =  factor(stage, levels = ordering(stage, order_vector))
+         ) %>% 
   group_by(transmission,
@@
            SorL) %>% 
-  dplyr::summarise(max_gene = max(tpm),
-                   total_clade = sum(tpm)) %>%
-  group_by(stage) %>% 
-  dplyr::mutate(total_stage = sum(total_clade)) %>% 
-  dplyr::mutate(
-    prop = round(total_clade/total_stage, digits = 2),
-    prop_from_max_gene = ifelse(total_clade > 0,
-                                round(max_gene/total_clade, digits = 2),
-                                0)
-  ) %>% 
-  # Prop from max gene to show how much of the clade expression is from the highest expressed gene of the clade.
-  filter(prop > 0.0001) %>%
-  dplyr::arrange(transmission, stage, desc(prop)) %>% 
-  select(-max_gene) %>% 
-  kable(format = 'html', 
-        caption = 'Proportion of pir clade expression for each stage', 
-        escape = FALSE) %>% 
-  kable_styling() %>% 
-  scroll_box(width = "100%", height = "500px")
-
-```
-
-# Expression by chromosome
-
-```{r chrom_proportion}
-tpm_data_twentyfour %>% 
-      group_by(transmission,
-           stage, 
-           chrom) %>% 
-  dplyr::summarise(max_gene = max(tpm),
-                   total_chrom = sum(tpm)) %>%
-  group_by(stage) %>% 
-  dplyr::mutate(total_stage = sum(total_chrom)) %>% 
-  dplyr::mutate(
-    prop = round(total_chrom/total_stage, digits = 2),
-    prop_from_max_gene = ifelse(total_chrom > 0,
-                                round(max_gene/total_chrom, digits = 2),
-                                0)
-  ) %>% 
-  # Prop from max gene to show how much of the chrom expression is from the highest expressed gene of the chrom.
-  filter(prop > 0.0001) %>%
-  dplyr::arrange(transmission, desc(prop)) %>% 
-  select(-max_gene) %>% 
-  kable(format = 'html', 
-        caption = 'Proportion of pir chrom expression for each stage', 
-        escape = FALSE) %>% 
-  kable_styling() %>% 
-  scroll_box(width = "100%", height = "500px")
-```
-
-```{r chrom_total_bar_chart}
-tpm_data_chrom <- tpm_data_twentyfour %>% 
-  mutate(stage = times_stages_table$time[match(stage, 
-                                               times_stages_table$r_stage_diff_count)]) %>% 
-  group_by(transmission,
-           stage, 
-           chrom) %>% 
-  dplyr::summarise(total_tpm = sum(tpm))
-
-y <- ggplot(data = tpm_data_chrom, 
+  dplyr::summarise(total_tpm = sum(tpm)) %>% 
+  mutate(SorL = factor(SorL, 
+                       levels = sort(unique(SorL))))
+
+y <- ggplot(data = tpm_data_SorL, 
               aes(
-                x = transmission,
+                x = stage,
                 y = total_tpm, 
-                fill = chrom,
-                group = transmission
+                fill = SorL,
+                # group = transmission
               )) +
@@
   scale_fill_manual(
-    breaks = unique(tpm_data_chrom$chrom),
-    values = setNames(
-      c(
-        plasma(length(unique(tpm_data_chrom$chrom)))[c(TRUE, NA)]
-      ),
-      unique(tpm_data_chrom$chrom)
-    ),
-    guide = guide_legend(reverse = TRUE)
+    values = SorL_col,
+    guide = guide_legend(reverse = FALSE)
   ) +
@@
   ) + 
-  facet_grid(.~stage, 
+  facet_grid(.~transmission, 
              scales = "free_x",
              space = "free_x") +
+  labs(fill = 'sub-family') +
     theme(
@@
 
-# ggsave2('tl20-05_chroms_TPM_barchart.pdf',
-#         plot = y,
-#         dpi = 300,
-#         width = 15,
-#         height = 10,
-#         units = 'cm')
-#   
+ggsave2('tl20-05_SorL_totalTPM_facet-transmission_barchart.pdf',
+        plot = y,
+        dpi = 300,
+        width = 15,
+        height = 10,
+        units = 'cm')
+
+```
+
+```{r proportion_data}
+
+tpm_data_twentyfour <- twentyfour_ct_comb$tpm %>% 
+  filter(Geneid %in% cir_id) %>% 
+  melt(id.vars = 'Geneid', 
+       variable.name = 'sample', 
+       value.name = 'tpm') %>% 
+  mutate(stage = str_extract(sample, 
+                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
+  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
+  filter(stage != 'sch.culture') %>% 
+  mutate(clade = cir_clade_info$clade[match(Geneid, 
+                                            cir_clade_info$Geneid)],
+         chrom = str_extract(Geneid, 
+                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
+         ) %>% 
+  mutate(
+    SorL = str_extract(clade, 
+                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
+    # stage =  factor(stage, 
+    #                 levels = ordering(stage, order_vector))
+  ) %>% 
+  group_by(transmission,
+           stage, 
+           Geneid,
+           clade,
+           chrom,
+           SorL) %>% 
+  dplyr::summarise(tpm = mean(tpm))
+
 ```
@@
 ```{r chrom_prop_bar_chart}
+tpm_data_chrom <- tpm_data_twentyfour %>%
+  mutate(stage = times_stages_table$time[match(stage,
+                                               times_stages_table$r_stage_diff_count)]) %>%
+  group_by(transmission,
+           stage,
+           chrom) %>%
+  dplyr::summarise(total_tpm = sum(tpm))
 
@@
     ylab(
-      bquote('total '~italic(.('pir'))~' TPM')
+      bquote('proportion of '~italic(.('pir'))~' TPM')
     ) + 
@@
 
-ggsave2('tl20-05_chroms_propTPM_barchart.png',
+ggsave2('tl19-05_chroms_propTPM_barchart.png',
         plot = y,
+        dpi = 330,
+        width = 15,
+        height = 10,
+        units = 'cm')
+
+```
+
+### 2D plot of PC1 v PC2 - Supp figure
+
+```{r pca_calculation, results = TRUE}
+twentyfour_tpm <- twentyfour_ct_comb_noschz$tpm
+rownames_geneid <- twentyfour_tpm$Geneid
+twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
+rownames(twentyfour_tpm_forpca) <- rownames_geneid
+twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
+#Need to look into what exactly scale and center are doing
+```
+
+```{r 2d_pca, results = TRUE}
+twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
+twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
+
+twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
+                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
+                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
+                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
+  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
+  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
+  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
+
+colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
+colornames <- setNames(colornames, 
+                       sort(unique(twentyfour_tpm_pca_coords$time)))
+
+pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
+pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
+pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
+pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
+
+(
+  plot <- ggplot(twentyfour_tpm_pca_coords,
+                 aes(
+                   x = PC1,
+                   y = PC2,
+                   col = time,
+                   shape = transmission
+                 )) +
+    geom_point(size = 3,
+               alpha = 0.8) +
+    scale_color_manual(values = colornames) +
+    scale_shape_manual(
+      values = setNames(
+        c(17,19),
+        unique(twentyfour_tpm_pca_coords$transmission)
+      )
+    ) +
+    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
+    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
+    guides(shape=guide_legend(ncol=2),
+           col = guide_legend(ncol=2)) +
+    theme_classic() +
+    theme(legend.key.size = unit(4, 'mm'),
+          legend.text = element_text(size = 7))
+)
+ggsave2(filename = 'tl19-07_chabaudi24h_supp_figure_pca_nopseudo.png',
+        plot = plot,
         dpi = 300,
@@
 
-Clearly a massive upregulation of chr6 in rMT.
-
-```{r chrom_6_L4s}
-cir_clade_info %>% 
-  mutate(
-    chrom = str_extract(Geneid, 
-                        '(?<=PCHAS_)[[:digit:]]{2}')
-  ) %>% 
-  dplyr::group_by(clade, 
-                  chrom) %>% 
-  dplyr::summarise(num = n()) %>% 
-  dplyr::mutate(freq = num/sum(num)*100) %>% 
-  arrange(clade,
-          desc(freq))
-
-cir_clade_info %>% 
-  mutate(
-    chrom = str_extract(Geneid, 
-                        '(?<=PCHAS_)[[:digit:]]{2}')
-  ) %>% 
-  dplyr::group_by(chrom, 
-                  clade) %>% 
-  dplyr::summarise(num = n()) %>% 
-  dplyr::mutate(freq = num/sum(num)*100) %>% 
-  arrange(chrom,
-          desc(freq))
-```
-
-L4 _pir_\s mostly on chr3 and chr6 (20.6% of the L4s each). 17% of L1s are on chr6.
-
-L1 pirs make up 58.3% of chr6, and L4 pirs make up another 29.2%, only three other pirs are left, one unknown and two S7s.
-
-## Bar chart of chromosome expression
-
-```{r chrom_expression_bar_charts}
-lapply(
-    unique(tpm_data_twentyfour$chrom), 
-  function(chrom_cirs){
-
-         tpm_df <- filter(tpm_data_twentyfour,
-                          chrom %in% chrom_cirs)
-
-         #Get the order
-
-         tpm_df <- mutate(tpm_df, 
-                          stage = factor(stage,
-                                         levels = ordering(stage,
-                                                           order_vector)))
-
-         maximum_tpm_value <- group_by(tpm_df, 
-                                       stage, 
-                                       transmission) %>% 
-           dplyr::summarise(sum_tpm = sum(tpm)) %>% 
-           .$sum_tpm %>% 
-           max
-
-         max_tpm <- as.integer(
-           mround(
-             seq(0, 
-                 maximum_tpm_value, 
-                 length.out = 5),
-             base = 5
-           )
-         )
-
-         y <- ggplot(data = tpm_df, 
-                     aes(
-                       x = transmission,
-                       y = tpm, 
-                       fill = Geneid
-                     )) +
-           geom_bar(position = position_stack(),
-                    stat = 'identity',
-                    colour = 'black',
-                    size = 0.1) +
-           scale_fill_manual(
-             breaks = unique(tpm_df$Geneid),
-             values = setNames(
-               rep_len(
+# z scores
+
+```{r read_z}
+### Read z-scores
+
+twentyfour_z <- list(
+  'rMT' = list(zscores = read_csv('tl19-07_chabaudi24h_rMT.csv'),
+               tpm = twentyfour_avgtpm_noschz$rMT$tpm,
+               experiment = 'rMT'),
+  'SBP' = list(zscores = read_csv('tl19-07_chabaudi24h_SBP.csv'),
+               tpm = twentyfour_avgtpm_noschz$SBP$tpm,
+               experiment = 'SBP')
+)
+
+###
+```
+
+# Phaseograms of the data. {.tabset}
+
+```{r row_order_phaseograms}
+
+#Getting the row order for each phaseogram for each experiment
+# Want the data ordered by the highest zscore for each gene in each stage.
+phaseogram_row_order <- lapply(twentyfour_z, 
+                               function(list_elements){
+                                 zscore_df <- list_elements$zscores
+                                 zscore_ordered <- melt(zscore_df, 
+                                                        id = 'Geneid',
+                                                        variable.name = 'stage',
+                                                        value.name = 'zscore') %>% 
+                                   group_by(Geneid) %>% 
+                                   top_n(zscore, n = 1) %>% 
+                                   arrange(stage, 
+                                           desc(zscore)) %>%
+                                   distinct(Geneid, 
+                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
+                                   filter(zscore > 0) #For some reason some remain with zero zscore
+                                 # zscore_ordered <- zscore_ordered[order(
+                                 #   match(
+                                 #     zscore_ordered$stage, 
+                                 #     order_vector
+                                 #   )
+                                 # ),]
+                                 return(c(list_elements, 
+                                          list(zscore_order = zscore_ordered)))
+                               })
+
+phaseogram_time_order <- lapply(names(phaseogram_row_order), 
+                                function(list_element){
+                                  transmission_list <- getElement(phaseogram_row_order,
+                                                                  list_element)
+                                  order_by_time <- transmission_list$zscore_order %>% 
+                                    mutate(stage = times_stages_table$time[match(stage,
+                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
+                                    arrange(stage)
+                                  return(list(zscore_order = order_by_time))
+                                }
+)
+names(phaseogram_time_order) <- names(phaseogram_row_order)
+
+```
+
+~~Change the lnTPM graph to a heatmap of max/min TPM a la Brugat 2017.~~
+
+~~Add annotation for the clades,~~ or possibly separate the heatmap by clade.
+
+Add threshold option.
+
+```{r function_draw_heatmap_zscore}
+
+draw_heatmap_zscore <- function(zscore_list,
+                                title = '',
+                                gene_ordering_vector = NULL,
+                                filter_genes_vector = NULL,
+                                extra_title = NULL,
+                                annotation_clade = NULL,
+                                annotation_col,
+                                threshold_tpm = NULL,
+                                use_times = FALSE,
+                                use_order_vector = TRUE,
+                                tpm_mat = TRUE,
+                                tpm_breaks = NULL,
+                                ...){
+  if(is.null(tpm_breaks) & tpm_mat | is.numeric(threshold_tpm)){
+    #Need the generate the breaks in the data for the heatmap legend if not provided.
+    # Threshold_tpm also needs this in order to filter lowly expressed genes.
+    max_tpm <- lapply(zscore_list, 
+                      function(list_element){
+                        #list_element <- zscore_list[[1]]
+                        heatmap_data <- as.data.frame(list_element$zscores)
+                        title <- list_element$experiment
+
+                        if(!'Geneid' %in% colnames(heatmap_data)){
+                          stop('No Geneid column in the heatmap data.frame')
+                        }
+
+                        if(!is.null(filter_genes_vector)){
+                          heatmap_data <- as.data.frame(
+                            filter(heatmap_data, 
+                                   Geneid %in% filter_genes_vector)
+                          )
+                        }
+
+                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+
+                        if(!is.null(gene_ordering_vector)){
+                          #Remove genes not in the main data frame but in the ordering vector, 
+                          # and then order based on the vector
+                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                          heatmap_data <- heatmap_data[na.omit(
+                            match(gene_ordering_vector, 
+                                  rownames(heatmap_data))
+                          ),,
+                          drop=FALSE]
+                        }
+
+                        geneids <- rownames(heatmap_data)
+
+                        #Get max/min TPM scores for the side annotation heatmap                       
+                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                          list_element$tpm$Geneid),] %>% 
+                          select(-Geneid) %>% 
+                          as.matrix %>% 
+                          log1p
+                        rownames(tpm_mat) <- geneids
+
+                        tpm_min <- rowMins(tpm_mat)
+                        names(tpm_min) <- geneids
+                        tpm_max <- rowMaxs(tpm_mat)
+                        names(tpm_max) <- geneids
+                        return(list(tpm_max = tpm_max,
+                                    tpm_min = tpm_min))
+                      })
+
+    tpm_max_all <- mround(
+      max(
+        unlist(
+          lapply(max_tpm, 
+                 getElement,
+                 'tpm_max')
+        )
+      ), 
+      base = 1)
+
+    tpm_breaks <- round(
+      c(0,
+        tpm_max_all/4,
+        tpm_max_all/2,
+        tpm_max_all*(3/4),
+        tpm_max_all),
+      digits = 2
+    )
+  }
+
+  #Generate the colour function from the TPM legend breaks for the TPM annotation
+  if(tpm_mat){
+    tpm_col_fun <- colorRamp2(
+      breaks = tpm_breaks,
+      colors = viridis(length(tpm_breaks))
+    )
+  }
+
+  if(is.numeric(threshold_tpm)){
+    #Apply a threshold to the TPM in order to remove low expressed genes.
+    # Remember to add log1p here! 
+    # Threshold uses the max value, so it ony needs to be above thresh 
+    #   in one stage avg in one transmission method.
+    #Will alter the gene_filter vector in order to only select the genes above threshold
+
+    gene_maxs <- bind_rows(
+      lapply(max_tpm, 
+             getElement,
+             'tpm_max')
+    )
+
+    remove_genes <- names(gene_maxs)[apply(
+      gene_maxs, 
+      2, 
+      function(maxs){
+        all(maxs < log2(1+threshold_tpm))
+      }
+    )]
+
+    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
+
+  }
+
+  if(is.character(annotation_clade)){
+    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
+    clade_info <- getElement(cir_clade_info,
+                             annotation_clade)[match(order_cir, 
+                                                     cir_clade_info$Geneid)]
+    names(clade_info) <- str_split_fixed(order_cir, 
+                                         pattern = '_', 
+                                         n = 2)[,2]
+    annotation_col <- annotation_col
+  }
+
+  heatmaps <- lapply(zscore_list, 
+                     function(list_element){
+                       #list_element <- zscore_list[[1]]
+                       heatmap_data <- as.data.frame(list_element$zscores)
+                       title <- list_element$experiment
+
+                       if(!'Geneid' %in% colnames(heatmap_data)){
+                         stop('No Geneid column in the heatmap data.frame')
+                       }
+
+                       if(!is.null(filter_genes_vector)){
+                         heatmap_data <- as.data.frame(
+                           filter(heatmap_data, 
+                                  Geneid %in% filter_genes_vector)
+                         )
+                       }
+
+                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+
+                       if(!is.null(gene_ordering_vector)){
+                         #Remove genes not in the main data frame but in the ordering vector,
+                         #  and then order based on the vector
+                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                         heatmap_data <- heatmap_data[na.omit(
+                           match(gene_ordering_vector, 
+                                 rownames(heatmap_data))
+                         ),,
+                         drop=FALSE]
+                       }
+
+                       if(tpm_mat){
+                         #Get max/min TPM scores for the side annotation heatmap                       
+                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                           list_element$tpm$Geneid),] %>%
+                           select(-Geneid) %>% 
+                           as.matrix %>% 
+                           log1p
+
+                         tpm_min <- rowMins(tpm_mat)
+                         tpm_max <- rowMaxs(tpm_mat)
+
+                         names(tpm_min) = rownames(heatmap_data)
+                         names(tpm_max) = rownames(heatmap_data)
+
+                         tpm_ha <- rowAnnotation(
+                           min.max = anno_simple(cbind(tpm_min, 
+                                                       tpm_max), 
+                                                 col = tpm_col_fun,
+                                                 gp = gpar(col = 'black',
+                                                           lwd = 0.1),
+                                                 border = TRUE,
+                                                 simple_anno_size = unit(0.2, "cm")),
+                           show_annotation_name = FALSE)
+                       }
+
+                       #Draw the left annotation is needed, only for the first element.
+
+                       if(
+                         identical(list_element,
+                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
+                       ){
+                         clade_ha <- rowAnnotation(
+                           'subfamily' = anno_simple(
+                             clade_info,
+                             col = annotation_col
+                           )
+                         )
+                       }
+
+                       #Change rownames to remove PCHAS_
+                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
+                                                                pattern = '_', 
+                                                                n = 2)[,2]
+
+                       if(use_times){
+                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
+                                                                                 times_stages_table$r_stage_diff_count)]
+                       }
+
+                       if(use_order_vector){
+                         #Order the columns by ordering_vector
+                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
+                                                                   order_vector)),
+                                                      drop = FALSE]
+                       }else{
+                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
+                                                      drop = FALSE] 
+                       }
+
+                       h = Heatmap(heatmap_data,
+                                   name = paste0('z-score',
+                                                 extra_title),
+                                   col = c(cbPalette[6],
                                            'white', 
-                 length.out = length(unique(tpm_df$Geneid))
+                                           cbPalette[7]),
+                                   row_names_gp = gpar(fontsize = ifelse(
+                                     nrow(heatmap_data) > 13, 
+                                     12*(13/nrow(heatmap_data)), 
+                                     12)
                                    ),
-               unique(tpm_df$Geneid)
-             )
-           )+
-           scale_color_manual('black') + 
-           # labs(fill = stack_label) +
-           scale_y_continuous(breaks = max_tpm) +
-           theme_classic() +
-           scale_size(range = c(0,3)) +
-           ylab(
-             bquote('total chr'~italic(.(chrom_cirs))~italic(.('pir'))~' TPM')
-           ) + 
-           facet_grid(.~stage, 
-                      scales = "free_x",
-                      space = "free_x") +
-           theme(
-             axis.title.x=element_blank(),
-             axis.text.x=element_text(size = 8),
-             axis.text.y = element_text(size = 5),
-             axis.title.y = element_text(size = 8),
-             legend.title = element_text(size = 7),
-             legend.text = element_text(size = 5),
-             legend.key.size = unit(3,'mm'),
-             legend.key.width = unit(2,'mm'),
-             strip.text.x = element_text(size=4, 
-                                         face = 'bold', 
-                                         angle = 90),
-             strip.background = element_rect(
-               color="black", 
-               size=0.1, 
-               linetype="solid"
-             )
-           )
-         if(global_save_var){
-             ggsave2(paste0('tl19-07_chr',
-                            gsub(chrom_cirs, 
-                                 pattern  = ' ', 
-                                 replacement = '_'),
-                            '_totalTPM_barchart.pdf'),
-                     plot = y,
-                     dpi = 300,
-                     width = 15,
+                                   column_names_gp = gpar(fontsize = 10),
+                                   row_dend_width = unit(3, "cm"),
+                                   right_annotation = if(tpm_mat){
+                                     tpm_ha
+                                   }else{
+                                     NULL
+                                   },
+                                   left_annotation = 
+                                     if(identical(list_element,
+                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
+                                       clade_ha
+                                     }else{
+                                       NULL
+                                     }, 
+                                   #Only want the clade annotation when it is first heatmap
+                                   width = unit(4.5, 'cm'),
+                                   cluster_columns = FALSE,
+                                   height = unit(7,'cm'),
+                                   row_title = NULL,
+                                   cluster_rows = FALSE,
+                                   column_title = paste(title),
+                                   heatmap_legend_param = list(
+                                     at = c(3,2,1,0,-1,-2,-3),
+                                     labels_gp = gpar(fontsize = 10),
+                                     title_gp = gpar(fontsize = 10)
+                                   ),
+                                   ...
+                       )
+                       return(h)
+                     })
+
+  ht_list <- Reduce('+', heatmaps) #To add the list elements together
+
+  if(tpm_mat){
+    #Draw the legend
+    tpm_lgd <- Legend(
+      title = "Min/Max log(TPM + 1)", 
+      col = tpm_col_fun, 
+      at = tpm_breaks, 
+      labels = tpm_breaks,
+      title_gp = gpar(fontsize = 7)
+    )
+  }
+
+  if(is.character(annotation_clade)){
+    clade_lgd <- Legend(
+      title = 'sub-family',
+      at = names(annotation_col),
+      legend_gp = gpar(fill = annotation_col)
+    )
+  }
+
+  draw(ht_list, 
+       row_title_side = 'left',
+       column_title_side = 'bottom',
+       heatmap_legend_side = 'left',
+       #Only include the annotation legends for the annotations we actually have.
+       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
+         list(tpm_lgd,
+              clade_lgd)
+       }else if(tpm_mat & !is.character(annotation_clade) ){
+         list(tpm_lgd)
+       } else if(!tpm_mat & is.character(annotation_clade) ){
+         list(clade_lgd)
+       })
+}
+```
+
+## Phaseograms of the entire genome.
+
+```{r phaseograms_whole_genome}
+png(filename = 'tl19-07_phaseogram_rMTorder_allgenes_filter1TPM.png',
+    units = 'cm',
+    height = 20,
+    width = 22,
+    res = 330)
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE,
+                    threshold_tpm = 1)
+dev.off()
+```
+
+## Phaseograms of the _cir_ genes with clade annotations.
+
+```{r pir_gene_phaseogram_breaks}
+  max_tpm <- lapply(twentyfour_z, 
+                    function(list_element){
+                      #list_element <- zscore_list[[1]]
+                      heatmap_data <- as.data.frame(list_element$zscores)
+                      title <- list_element$experiment
+
+                      if(!'Geneid' %in% colnames(heatmap_data)){
+                        stop('No Geneid column in the heatmap data.frame')
+                      }
+
+                            heatmap_data <- as.data.frame(
+                              filter(heatmap_data, 
+                                     Geneid %in% cir_id)
+                            )
+
+                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+                          # 
+                          # if(!is.null(gene_ordering_vector)){
+                          #   #Remove genes not in the main data frame but in the ordering vector, 
+                          #   # and then order based on the vector
+                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                          #   heatmap_data <- heatmap_data[na.omit(
+                          #     match(gene_ordering_vector, 
+                          #           rownames(heatmap_data))
+                          #   ),,
+                          #   drop=FALSE]
+                          # }
+
+                          geneids <- rownames(heatmap_data)
+
+                          #Get max/min TPM scores for the side annotation heatmap                       
+                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                            list_element$tpm$Geneid),] %>% 
+                            select(-Geneid) %>% 
+                            as.matrix %>% 
+                            log1p
+                          rownames(tpm_mat) <- geneids
+
+                          tpm_min <- rowMins(tpm_mat)
+                          names(tpm_min) <- geneids
+                          tpm_max <- rowMaxs(tpm_mat)
+                          names(tpm_max) <- geneids
+                          return(list(tpm_max = tpm_max,
+                                      tpm_min = tpm_min))
+                        })
+
+  tpm_max_all <- mround(
+    max(
+      unlist(
+        lapply(max_tpm, 
+               getElement,
+               'tpm_max')
+      )
+    ), 
+    base = 1)
+
+  tpm_breaks <- round(
+    c(0,
+      tpm_max_all/4,
+      tpm_max_all/2,
+      tpm_max_all*(3/4),
+      tpm_max_all),
+    digits = 2
+  )
+```
+
+```{r phaseograms_cir_order_each}
+
+#Generate the colour function from the TPM legend breaks for the TPM annotation
+
+tpm_col_fun <- colorRamp2(
+  breaks = tpm_breaks,
+  colors = viridis(length(tpm_breaks))
+)
+
+#Apply a threshold to the TPM in order to remove low expressed genes.
+# Remember to add log1p here! 
+# Threshold uses the max value, so it ony needs to be above thresh 
+#   in one stage avg in one transmission method.
+#Will alter the gene_filter vector in order to only select the genes above threshold
+
+gene_maxs <- bind_rows(
+  lapply(max_tpm, 
+         getElement,
+         'tpm_max')
+)
+
+remove_genes <- names(gene_maxs)[apply(
+  gene_maxs, 
+  2, 
+  function(maxs){
+    all(maxs < log2(1+1))
+  }
+)]
+
+filter_genes_vector <- cir_id[!cir_id %in% remove_genes]
+
+heatmaps <- lapply(names(twentyfour_z), 
+                   function(list_name){
+                     list_element <- getElement(twentyfour_z, 
+                                                list_name)
+                     #list_name <- names(twentyfour_z)[[1]]
+                     heatmap_data <- as.data.frame(list_element$zscores)
+                     title <- list_element$experiment
+
+                     heatmap_data <- as.data.frame(
+                       filter(heatmap_data, 
+                              Geneid %in% filter_genes_vector)
+                     )
+
+                     rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                     heatmap_data <- heatmap_data[, 
+                                                  !colnames(heatmap_data) %in% 'Geneid', 
+                                                  drop = TRUE]
+                     #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+
+                     #Remove genes not in the main data frame but in the ordering vector,
+                     #  and then order based on the vector
+                     gene_ordering_vector <- getElement(phaseogram_time_order,
+                                                        list_name)$zscore_order$Geneid
+
+                     order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
+                     clade_info <- getElement(cir_clade_info,
+                                              'SorL')[match(order_cir, 
+                                                            cir_clade_info$Geneid)]
+                     names(clade_info) <- str_split_fixed(order_cir, 
+                                                          pattern = '_', 
+                                                          n = 2)[,2]
+                     annotation_col <- SorL_col
+
+                     remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                     heatmap_data <- heatmap_data[remove_index,
+                                                  ,
+                                                  drop = FALSE]
+                     heatmap_data <- heatmap_data[na.omit(
+                       match(gene_ordering_vector, 
+                             rownames(heatmap_data))
+                     ),
+                     ,
+                     drop=FALSE]
+
+                     #Get max/min TPM scores for the side annotation heatmap                       
+                     tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                       list_element$tpm$Geneid),] %>%
+                       select(-Geneid) %>% 
+                       as.matrix %>% 
+                       log1p
+
+                     tpm_min <- rowMins(tpm_mat)
+                     tpm_max <- rowMaxs(tpm_mat)
+
+                     names(tpm_min) = rownames(heatmap_data)
+                     names(tpm_max) = rownames(heatmap_data)
+
+                     tpm_ha <- rowAnnotation(
+                       min.max = anno_simple(cbind(tpm_min, 
+                                                   tpm_max), 
+                                             col = tpm_col_fun,
+                                             gp = gpar(col = 'black',
+                                                       lwd = 0.1),
+                                             border = TRUE,
+                                             simple_anno_size = unit(0.2, "cm")),
+                       show_annotation_name = FALSE)
+
+                     #Clade annotation
+
+                     clade_ha <- rowAnnotation(
+                       'sub-family' = anno_simple(
+                         clade_info,
+                         col = annotation_col,
+                       ),
+                       annotation_name_gp = gpar(fontsize = 10)
+                     )
+
+                     #Change rownames to remove PCHAS_
+                     rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
+                                                              pattern = '_', 
+                                                              n = 2)[,2]
+
+                     colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
+                                                                             times_stages_table$r_stage_diff_count)]
+
+                     heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
+                                                  drop = FALSE]
+
+                     h = Heatmap(heatmap_data,
+                                 name = paste0('z-score'),
+                                 col = c(cbPalette[6],
+                                         'white', 
+                                         cbPalette[7]),
+                                 row_names_gp = gpar(fontsize = ifelse(
+                                   nrow(heatmap_data) > 13, 
+                                   12*(13/nrow(heatmap_data)), 
+                                   12)
+                                 ),
+                                 column_names_gp = gpar(fontsize = 10),
+                                 row_dend_width = unit(3, "cm"),
+                                 right_annotation = tpm_ha,
+                                 left_annotation = clade_ha,
+                                 width = unit(4.5, 'cm'),
+                                 cluster_columns = FALSE,
+                                 height = unit(7,'cm'),
+                                 row_title = NULL,
+                                 cluster_rows = FALSE,
+                                 column_title = paste(title),
+                                 heatmap_legend_param = list(
+                                   at = c(3,2,1,0,-1,-2,-3),
+                                   labels_gp = gpar(fontsize = 10),
+                                     title_gp = gpar(fontsize = 10)
+                                 ),
+                                 show_row_names = FALSE
+                     )
+                     return(h)
+                   })
+
+ht_list <- Reduce('+', heatmaps) #To add the list elements together
+
+#Draw the legend
+tpm_lgd <- Legend(
+  title = "Min/Max log(TPM + 1)", 
+  col = tpm_col_fun, 
+  at = tpm_breaks, 
+  labels = tpm_breaks,
+  title_gp = gpar(fontsize = 7)
+)
+
+clade_lgd <- Legend(
+  title = 'sub-family',
+  at = names(SorL_col),
+  legend_gp = gpar(fill = SorL_col)
+)
+
+png(filename = 'tl19-07_phaseogram_each-in-order_cir_clades.png',
+    units = 'cm',
+    height = 20,
+    width = 22,
+    res = 200)
+
+  draw(ht_list, 
+       row_title_side = 'left',
+       column_title_side = 'bottom',
+       heatmap_legend_side = 'left',
+       annotation_legend_list = 
+         list(tpm_lgd,
+              clade_lgd),
+       ht_gap = unit(.5,'cm')
+       )
+
+  dev.off()
+
+```
+
+```{r phaseograms_cir}
+png(filename = 'tl19-07_phaseogram_rMTorder_cir_clades_filter1TPM.png',
+    units = 'cm',
+    height = 20,
+    width = 22,
+    res = 330)
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'clade',
+                    annotation_col = clade_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    tpm_mat = FALSE)
+dev.off()
+
+png(filename = 'tl19-07_phaseogram_rMTorder_cir_SorL_filter1TPM.png',
+    units = 'cm',
+    height = 20,
+    width = 22,
+    res = 330)
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE)
+dev.off()
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE)
+
+png(filename = 'tl19-07_phaseogram_SBPorder_cir_SorL_filter1TPM.png',
+    units = 'cm',
+    height = 20,
+    width = 22,
+    res = 330)
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_breaks = tpm_breaks,
+                    tpm_mat = FALSE)
+dev.off()
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_breaks = tpm_breaks,
+                    tpm_mat = FALSE)
+
+```
+## Figure 3b - Highest expressed all _pir_ heatmap
+
+```{r highest_expressed_stage_heatmap_fig3b}
+
+num_SorL <- cir_info %>% 
+  group_by(SorL) %>% 
+  summarise(num = n_distinct(Gene.ID))
+
+labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
+                        num_SorL$SorL)
+
+# Get order of pir genes by highest expressed genes in each sample.
+
+combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm_noschz), 
+       function(list_name){
+
+         combined_data_pir <- getElement(twentyfour_avgtpm_noschz, 
+                                     list_name)$tpm %>% 
+           filter(Geneid %in% cir_id) %>% 
+           column_to_rownames(var = 'Geneid')
+
+         #Ordering that a more understandable order
+
+         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
+                                                            order_vector)),
+                                                drop = FALSE]
+
+         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
+                                         means = apply(combined_data_pir, 1, mean),
+                                         subfam = cir_clade_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
+                                                                            cir_clade_info$Geneid)]) %>% 
+           filter(means > 1) %>% 
+           arrange(subfam, 
+                   desc(means)) %>% 
+           .$Geneid
+
+         #Order the heatmap data by the order defined above (SorL, then mean expression).
+         combined_data_pir_ht <- combined_data_pir[
+           order(
+             match(rownames(combined_data_pir),
+                   combined_data_ordered),
+             na.last = NA #Remove NAs
+           ),
+         ]
+
+         #Change the colnames to the times
+         colnames(combined_data_pir_ht) <- times_stages_table$time[match(colnames(combined_data_pir_ht),
+
+                                                                         times_stages_table$r_stage_diff_count)]
+
+         #Order by time point
+         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+                                                      drop = FALSE] 
+         #Add a column for the SorL
+         combined_data_pir_ht$SorL <- cir_clade_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
+                                                                cir_clade_info$Geneid)]
+
+         return(combined_data_pir_ht)
+       })
+names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm_noschz)
+
+maxcol <- ceiling(
+  max(
+    unlist(
+      lapply(combined_data_pir_ht_list,
+             function(list_element){
+               max(
+                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
+               )
+             }
+      )
+    )
+  )
+)
+
+lapply(names(combined_data_pir_ht_list),
+       function(list_name_to_order){
+
+         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
+                                            list_name_to_order)
+
+         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
+
+         #Sub-family annotation
+         clade_info <- combined_data_pir_ht$SorL
+         annotation_col <- SorL_col
+
+         #Subfamily annotation
+
+         clade_ha <- rowAnnotation(
+           'sub-family' = anno_simple(
+             clade_info,
+             col = annotation_col,
+           ),
+           annotation_name_gp = gpar(fontsize = 10)
+         )
+
+         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
+                                                 labels = 'ancestral',
+                                                 labels_gp = gpar(fontsize = 7),
+                                                 link_width = unit(3, "mm")))
+
+         #lapply across the datasets now, for each order
+         ht_list <- lapply(names(combined_data_pir_ht_list),
+                           function(list_name){
+
+                             combined_data_pir_ht <- getElement(twentyfour_avgtpm_noschz, 
+                                                                list_name)$tpm %>% 
+                               column_to_rownames(var = 'Geneid')
+
+                             #Get the times instead of the stage names.
+                             colnames(combined_data_pir_ht) <- times_stages_table$time[match(colnames(combined_data_pir_ht),
+
+                                                                                             times_stages_table$r_stage_diff_count)]
+
+                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+                                                                          drop = FALSE]
+
+                             Heatmap(
+                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
+                               cluster_rows = FALSE,
+                               cluster_columns = FALSE,
+                               show_row_names = FALSE,
+                               name = 'log(TPM+1)', 
+                               col = colorRamp2(c(0,
+                                                  maxcol/4,
+                                                  maxcol/2,
+                                                  3*maxcol/4,
+                                                  maxcol),
+                                                viridis(5)),
+                               na_col = 'white', 
+                               border = TRUE,
+                               border_gp = gpar(lwd = 0.1),
+                               column_names_gp = gpar(fontsize = 10),
+                               row_dend_width = unit(3, "cm"),
+                               width = unit(4.5, 'cm'),
+                               height = unit(7,'cm'),
+                               row_title = NULL,
+                               left_annotation = if(
+                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
+                               ){
+                                 clade_ha
+                               }else{
+                                 NULL
+                               },
+                               right_annotation = if(
+                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
+                               ){
+                                 anc_ha
+                               }else{
+                                 NULL
+                               },
+                               # row_split =  cir_clade_info$SorL[match(pir_highest_tpm_order, 
+                               #                                        cir_clade_info$Geneid)],
+                               column_title = paste(list_name),
+                               heatmap_legend_param = list(
+                                 labels_gp = gpar(fontsize = 10),
+                                 title_gp = gpar(fontsize = 10,
+                                                 fontface = 'bold'),
+                                 at = c(0,
+                                        maxcol/4,
+                                        maxcol/2,
+                                        3*maxcol/4,
+                                        maxcol)
+                               )
+                             )
+                           }
+         )
+         ht_list <- Reduce(`+`, ht_list)
+         clade_lgd <- Legend(
+           title = 'sub-family',
+           at = labels_SorL,
+           legend_gp = gpar(fill = SorL_col)
+         )
+
+         png(filename = paste0('tl19-07_combined_pirs_ordered_highest_stage_expression_heatmap_TPM_',
+                               list_name_to_order, 'order',
+                               '.png'),
+             width = 20,
              height = 10,
-                     units = 'cm')
-         }
-         y
+             units = 'cm',
+             res = 330)
+
+         draw(ht_list, 
+              row_title_side = 'left',
+              column_title_side = 'bottom',
+              heatmap_legend_side = 'right',
+              annotation_legend_side = 'right',
+              annotation_legend_list = list(clade_lgd),
+              ht_gap = unit(.35,'cm'),
+              merge_legend = TRUE
+         )
+
+         dev.off()
+         return(
+           draw(ht_list, 
+                row_title_side = 'left',
+                column_title_side = 'bottom',
+                heatmap_legend_side = 'right',
+                annotation_legend_side = 'right',
+                annotation_legend_list = list(clade_lgd),
+                ht_gap = unit(.35,'cm'),
+                merge_legend = TRUE
+           ))
        })
-```
-
-# Expression by locus
-
-```{r defining_loci}
-cir_loci <- data.frame(
-  Geneid = cir_id,
-  chrom = str_extract(cir_id, 
-                      pattern = '(?<=_)[[:digit:]]{2}'),
-  number = as.numeric(
-    str_extract(cir_id, 
-                pattern = '(?<=_[[:digit:]]{2})[[:digit:]]+')
-  )
-) %>% 
-  filter(!chrom %in% '00') %>% 
-  #Remove chrom00 (contigs)
-  group_by(chrom) %>% 
-  mutate(jump = ifelse(lag(number, 
-                           default = 1001) < 3000 & number - lag(number, 
-                                                                 default = 0) > 1000,
-                       1,
-                       0)) %>% 
-  mutate(
-    jump = ifelse(cumsum(jump) == 1 | jump == 1, 1, 0)
-  ) %>% 
-  mutate(
-    locus = paste0(
-      chrom,
-      ifelse(jump == 0 & number < 3000, 
-             'Left',
-             'Right')
-    )
-  )
-
-# Detect 'jumps' of the gene number over 1000 within each chromosome, setting the lag default as 1001 in the
-#   first part of the ifelse statement so that chroms don't have an erroneous jump if they start at high numbers
-#   >1000
-# Later cumsum is do that every gene after the 'jump' is included allowing us to define telomere ends.
-
-cir_loci_list <- lapply(
-  unique(cir_loci$locus),
-  function(locus_name){
-    cir_loci$Geneid[cir_loci$locus == locus_name]
-  })
-
-names(cir_loci_list) <- unique(cir_loci$locus)
-```
-
-```{r loci_numbers}
-cir_loci %>% 
-  group_by(locus) %>% 
-  dplyr::summarise(num = n()) %>% 
-  arrange(desc(num))
-```
-
-## Loci expression stacked bar graph
-
-```{r loci_total_bar_chart}
-
-tpm_data_loci <- filter(tpm_data_twentyfour,
-                                       Geneid %in% cir_id) %>%
-  mutate(locus = cir_loci$locus[match(Geneid, 
-                                      cir_loci$Geneid)],
-         stage = times_stages_table$time[match(stage, 
-                                               times_stages_table$r_stage_diff_count)]) %>% 
-  ungroup() %>% 
-  group_by(stage, 
-           transmission,
-           locus) %>% 
-  dplyr::summarise(max_gene = max(tpm),
-                   total_tpm = sum(tpm))
-
-y <- ggplot(data = tpm_data_loci, 
-              aes(
-                x = transmission,
-                y = total_tpm, 
-                fill = locus,
-                group = transmission
-              )) +
-    geom_bar(position = position_stack(),
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-  scale_fill_manual(
-    breaks = unique(tpm_data_loci$locus),
-    values = setNames(
-      c(
-        plasma(length(unique(tpm_data_loci$locus)))[
-          ifelse(str_detect(unique(tpm_data_loci$locus),
-                            pattern = 'Left'),
-                 TRUE,
-                 NA)
-        ]
-      ),
-      unique(tpm_data_loci$locus)
-    ),
-    guide = guide_legend(reverse = TRUE)
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  # scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~stage, 
-             scales = "free_x",
-               space = "free_x") +
-    theme(
-      axis.title.x=element_blank(),
-      # axis.text.x=element_blank(),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-# ggsave2('tl20-05_loci_TPM_barchart.pdf',
-#         plot = y,
-#         dpi = 300,
-#         width = 15,
-#         height = 10,
-#         units = 'cm')
+
+```
+
+# ```{r figure_3b_suppinfo}
 # 
-```
-
-## Expression by locus proportions
-
-```{r locus_prop}
-
-tpm_data_loci %>%
-  group_by(stage, 
-           transmission) %>% 
-  dplyr::mutate(total_stage = sum(total_tpm)) %>% 
-  dplyr::mutate(
-    prop = round(total_tpm/total_stage, digits = 2),
-    prop_from_max_gene = ifelse(total_tpm > 0,
-                                round(max_gene/total_tpm, digits = 2),
-                                0)
-  ) %>% 
-  # Prop from max gene to show how much of the locus expression is from the highest expressed gene of the locus.
-  filter(prop > 0.0001) %>%
-  dplyr::arrange(desc(prop)) %>% 
-  select(-max_gene) %>% 
-  kable(format = 'html', 
-        caption = 'Proportion of pir loci expression for each stage', 
-        escape = FALSE) %>% 
-  kable_styling() %>% 
-  scroll_box(width = "100%", height = "500px")
-```
-
-* 34% of early and mid rMT troph comes from 06Left - 27% and 33% contribution from one gene respectively.
-Also 28% of late rings (17% from one gene).
-So spread out over multiple pir genes.
-Actually it is a majorly expressed locus it seems in rMT.
-06Left has 14 genes, the most of any locus bar 03Left (18 genes).
-
-```{r}
-y <- ggplot(data = tpm_data_loci, 
-              aes(
-                x = transmission,
-                y = total_tpm, 
-                fill = locus,
-                group = transmission
-              )) +
-    geom_bar(position = 'fill',
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-  scale_fill_manual(
-    breaks = unique(tpm_data_loci$locus),
-    values = setNames(
-      c(
-        rep(cbPalette, 
-            length.out = length(unique(tpm_data_loci$locus)),
-            each = 2)[
-          ifelse(str_detect(unique(tpm_data_loci$locus),
-                            pattern = 'Left'),
-                 TRUE,
-                 NA)
-        ]
-      ),
-      unique(tpm_data_loci$locus)
-    ),
-    guide = guide_legend(reverse = TRUE)
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  # scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~stage, 
-             scales = "free_x",
-               space = "free_x") +
-    theme(
-      axis.title.x=element_blank(),
-      # axis.text.x=element_blank(),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-# ggsave2('tl20-05_loci_propTPM_barchart.pdf',
-#         plot = y,
-#         dpi = 300,
-#         width = 15,
-#         height = 10,
-#         units = 'cm')
+# supp_info_fig_3b <- lapply(names(combined_data_pir_ht_list),
+#        function(list_name){
 #          
-```
-
-## Expression by locus over time
-
-```{r loci_expression}
-
-lapply(names(cir_loci_list), 
-       function(loci_name){
-
-         cir_loci_genes <- getElement(cir_loci_list, 
-                                      loci_name)
-
-         tpm_df <- filter(tpm_data_twentyfour,
-                          Geneid %in% cir_loci_genes)
-
-         #Get the order
-
-         tpm_df <- mutate(tpm_df, 
-                          stage = factor(stage,
-                                         levels = ordering(stage,
-                                                           order_vector)))
-
-         maximum_tpm_value <- group_by(tpm_df, 
-                                       stage, 
-                                       transmission) %>% 
-           dplyr::summarise(sum_tpm = sum(tpm)) %>% 
-           .$sum_tpm %>% 
-           max
-
-         max_tpm <- as.integer(
-           mround(
-             seq(0, 
-                 maximum_tpm_value, 
-                 length.out = 5),
-             base = 5
-           )
-         )
-
-         y <- ggplot(data = tpm_df, 
-                     aes(
-                       x = transmission,
-                       y = tpm, 
-                       fill = Geneid
-                     )) +
-           geom_bar(position = position_stack(),
-                    stat = 'identity',
-                    colour = 'black',
-                    size = 0.1) +
-           scale_fill_manual(
-             breaks = unique(tpm_df$Geneid),
-             values = setNames(
-               rep_len(
-                 'white', 
-                 length.out = length(unique(tpm_df$Geneid))
-               ),
-               unique(tpm_df$Geneid)
-             )
-           )+
-           scale_color_manual('black') + 
-           # labs(fill = stack_label) +
-           scale_y_continuous(breaks = max_tpm) +
-           theme_classic() +
-           scale_size(range = c(0,3)) +
-           ylab(
-             bquote('total '~italic(.(loci_name))~italic(.('pir'))~' TPM')
-           ) + 
-           facet_grid(.~stage, 
-                      scales = "free_x",
-                      space = "free_x") +
-           theme(
-             axis.title.x=element_blank(),
-             axis.text.x=element_text(size = 8),
-             axis.text.y = element_text(size = 5),
-             axis.title.y = element_text(size = 8),
-             legend.title = element_text(size = 7),
-             legend.text = element_text(size = 5),
-             legend.key.size = unit(3,'mm'),
-             legend.key.width = unit(2,'mm'),
-             strip.text.x = element_text(size=4, 
-                                         face = 'bold', 
-                                         angle = 90),
-             strip.background = element_rect(
-               color="black", 
-               size=0.1, 
-               linetype="solid"
-             )
-           )
-         if(global_save_var){
-             ggsave2(paste0('tl19-07_',
-                            gsub(loci_name, 
-                                 pattern  = ' ', 
-                                 replacement = '_'),
-                            '_totalTPM_barchart.pdf'),
-                     plot = y,
-                     dpi = 300,
-                     width = 15,
-                     height = 10,
-                     units = 'cm')
-         }
-         y
-       })
-```
-
-* Cyclical expression generally similar between different chromosomes.
-* 01Left, 05Right dominated by the ancestral and pan-rodent pirs.
-* 01Right, 03Right, 07Right, 08Left, 08Right, 09Left, 10Left, 10Right, 11Left, 11Right, 12Right, 13Right, 14Left, 14Right show association with SBP expression.
-* 04Right in rMT peaks in Early Trophs, and is spread across multiple cir genes.
-* 10Left peaks in Early-Mid Rings mostly due to one pir PCHAS_1001351. 12Right also peaks then in SBP due to one or two pirs.
-* In some loci the higher expression in rMT is massive; 03Left, 06Left, 06Right, 12Left.
-
-# Lower expression of _pir_ genes versus the rest of the genome
+#          combined_data_pir_ht <- getElement(twentyfour_avgtpm_noschz, 
+#                                             list_name)$tpm %>% 
+#            column_to_rownames(var = 'Geneid')
+#          
+#          #Get the times instead of the stage names.
+#          colnames(combined_data_pir_ht) <- times_stages_table$time[match(colnames(combined_data_pir_ht),
+#                                                                          
+#                                                                          times_stages_table$r_stage_diff_count)]
+#          
+#          combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+#                                                       drop = FALSE] 
+#          
+#          colnames(combined_data_pir_ht) <- paste0(list_name, 
+#                                                   '_', 
+#                                                   colnames(combined_data_pir_ht))
+#          
+#          log2(1+combined_data_pir_ht)[pir_highest_tpm_order,]
+#        }) %>% bind_cols
+# 
+# supp_info_fig_3b$SorL <- cir_clade_info$SorL[match(names(apply(supp_info_fig_3b, 1, mean)), 
+#                                                                 cir_clade_info$Geneid)]
+# 
+# ```
+
+# Lower pir expression vs rest of genome
 
@@
        aes(x = pir_gene,
-           y = log1p(mean_tpm))) +
+           y = log2(1+mean_tpm))) +
   facet_grid(~transmission) +
@@
 
-if(global_save_var){
 ggsave2('tl19-07_pir-v-nonpir_mean_TPM_barchart.pdf',
@@
         units = 'cm')
-}
-```
-
-# Fourier analysis of the expression data.
-
-Used a script written by Chris Illingham and shared by Adam Reid.
-This is written in C++ so I don't quite understand all of it, but from what I can tell it calculates the phase order for each gene given the data points using the sine/cosine Fourier transform (as opposed to using complex numbers).
-
-```{r different_phase_genes}
-phase_df <- read_delim('./Fourier_code/Fourier_data.out',
-                       delim = " ",
-                       trim_ws = TRUE)
-# This definitely isn't reading in correctly past Phase. There's spaces in the column names for 'Normalised input' which is missing stuff up.
-
-# total_tpm_expression(list_genes = list(PCHAS_0100651 = 'PCHAS_0100651',
-#                                        PCHAS_0100600 = 'PCHAS_0100600',
-#                                        PCHAS_0100700 = 'PCHAS_0100700'))
-
-phase_ordered_df <- dplyr::arrange(phase_df,
-                                   Phase)
-
-select_gene_index <- ceiling(
-  sapply(
-    c(1/nrow(phase_ordered_df), 
-      1/4, 
-      1/2, 
-      3/4, 
-      1),
-    `*`,
-    nrow(phase_ordered_df)
-  )
-)
-#Oddly I think the 1/nrow multiplication gives a double/float (i.e. a decimel) which subsetting then converts to an integer giving zero!
-
-# (phaseiest_genes <- phase_ordered_df[select_gene_index,])
-# phaseiest_genes <- phaseiest_genes$ID
-# phaseiest_genes_list <- as.list(phaseiest_genes)
-# names(phaseiest_genes_list) <- phaseiest_genes
-# total_tpm_expression(
-#   phaseiest_genes_list
-#   )
-
-#Note that the genes at certain phase extremities may have very low expression
-
-#The proportionally top genes of each phase
-
-phaseiest_genes <- list(
-  top_phase = slice_max(phase_df,
-                        order_by = Phase,
-                        n = 100)$ID, #Highest
-  upp_quart_phase = slice_min(
-    slice_max(phase_df,
-              order_by = Phase,
-              prop = 0.25),
-    order_by = Phase,
-    n = 100
-  )$ID,
-  upp_third_phase = slice_min(
-    slice_max(phase_df,
-              order_by = Phase,
-              prop = 0.33),
-    order_by = Phase,
-    n = 100
-  )$ID,
-  mid_phase = slice_min(
-    slice_max(phase_df,
-              order_by = Phase,
-              prop = 0.55),
-    order_by = Phase,
-    n = 100
-  )$ID, #Middle (lowest of the top 50%)
-    low_quart_phase = slice_max(
-    slice_min(phase_df,
-              order_by = Phase,
-              prop = 0.25),
-    order_by = Phase,
-    n = 100
-  )$ID,
-  min_phase = slice_min(phase_df,
-                        order_by = Phase,
-                        n = 100)$ID #Lowest
-)
-total_tpm_expression(phaseiest_genes)
-```
-
-Expression profiles of the pir genes with Fourier transforms predicting a frequency over 1Hz (all 2Hz)
-
-```{r pirs_over_1Hz}
-two_freq_pirs <- as.list(
-  filter(phase_df,
-       ID %in% cir_id,
-       Max.freq > 1)$ID
-  )
-names(two_freq_pirs) <- unlist(two_freq_pirs)
-total_tpm_expression(list_genes = two_freq_pirs)
-```
-
-```{r}
-filter(phase_df,
-       Max.freq > 1) %>% 
-  arrange(desc(Significance))
-
-filter(phase_df,
-       Max.freq > 1) %>% 
-  group_by(Max.freq) %>% 
-  top_n(Significance, 
-        n = 3) %>% 
-  arrange(desc(Max.freq), 
-          desc(Significance))
-total_tpm_expression(list_genes = list(
-  four_Hz = 'PCHAS_1460200',
-  three_Hz = 'PCHAS_1317100',
-  two_Hz = 'PCHAS_1411500'
-))
-
-phase_df$Max.freq %>% unique
-
-arrange(phase_df, 
-        Significance)
-
-filter(phase_df, 
-        abs(Phase) < 0.5)
-```
-
-# Sessioninfo
-
-```{r}
-sessionInfo()
-```
+```
+
+```{r write_excel_data_sheets}
+
+twentyfour_avgtpm_noschz_timecolnames <- lapply(names(twentyfour_avgtpm_noschz), 
+                                                function(list_name){
+                                                  list_element <- getElement(twentyfour_avgtpm_noschz, list_name)
+                                                  colnames(list_element$tpm)[-1] <- times_stages_table$time[match(colnames(list_element$tpm)[-1],
+                                                                                                              times_stages_table$r_stage_diff_count)]
+                                                  list_element$tpm <- list_element$tpm[c('Geneid',sort(colnames(list_element$tpm)[-1]))]
+                                                  return(list_element)
+                                                })
+names(twentyfour_avgtpm_noschz_timecolnames) <- names(twentyfour_avgtpm_noschz)
+
+twentyfour_z_timecolnames <- lapply(names(twentyfour_z), 
+                                    function(list_name){
+                                      list_element <- getElement(twentyfour_z, list_name)
+                                      colnames(list_element$zscores)[-1] <- times_stages_table$time[match(colnames(list_element$zscores)[-1],
+                                                                                                          times_stages_table$r_stage_diff_count)]
+                                      list_element$zscores <- list_element$zscores[c('Geneid',sort(colnames(list_element$zscores)[-1]))]
+                                      return(list_element)
+                                    })
+names(twentyfour_z_timecolnames) <- names(twentyfour_z)
+
+excel_data <- list(
+  pir_gene_info = cir_info,
+  chab_TPM_bioreps = twentyfour_ct_comb_noschz$tpm[,c('Geneid',sort(colnames(twentyfour_ct_comb_noschz$tpm)[-1]))] %>% 
+    select(matches('Geneid|rMT')) %>% 
+    dplyr::rename_with(~ str_remove_all(., pattern = 'rMT_'), everything()),
+  chab_TPM_avg = twentyfour_avgtpm_noschz_timecolnames$rMT$tpm,
+  # chab_TPM_avg_SBP = twentyfour_avgtpm_noschz_timecolnames$SBP$tpm,
+  # chab_logTPM_SorL_rMTorder = supp_info_fig_3b,
+  chab_TPM_ChaplAapl = select(cir_plot_chapl_aapl_supp_info, c(Geneid, time, transmission, chapl_aapl, sample, tpm)) %>% 
+    filter(transmission == 'rMT') %>% 
+    select(-transmission) %>% 
+    filter(time != 'sch.culture') %>% 
+    mutate(sample = str_remove_all(sample, pattern = 'rMT_')) %>% 
+    arrange(Geneid, time),
+  # chab_TPM_SorLSums = tpm_data_SorL,
+  # chab_TPM_SubfamSums = tpm_data_clade,
+  # chab_TPM_ChromSums = tpm_data_chrom,
+  chab_zscore = twentyfour_z_timecolnames$rMT$zscores
+  # chab_zscore_SBP = twentyfour_z_timecolnames$SBP$zscores
+)
+
+write_xlsx(excel_data,
+           path = 'Manuscript_SuppInfo_chabaudi24h.xlsx')
+
+```

### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Make more chabaudi figures

@@
 library('circlize') # for circular plots and colour
+library("factoextra")
 
@@
          colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h')
+                                                                           pattern = '[[:digit:]]{2}h|sch.culture')
 
@@
          clade_ha <- rowAnnotation(
-           'sub-family' = anno_simple(
+           'sub-fam' = anno_simple(
              clade_info,
@@
 
-         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
-                                                 labels = 'ancestral',
-                                                 labels_gp = gpar(fontsize = 7),
-                                                 link_width = unit(3, "mm")))
-
          #lapply across the datasets now, for each order
@@
                              colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h')
+                                                                           pattern = '[[:digit:]]{2}h|sch.culture')
 
@@
                                },
-                               right_annotation = if(
-                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
-                               ){
-                                 anc_ha
-                               }else{
-                                 NULL
-                               },
-                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
-                               #                                        cir_info$Geneid)],
                                column_title = paste(list_name),
@@
          clade_lgd <- Legend(
-           title = 'sub-family',
+           title = 'sub-fam',
            at = labels_SorL,
@@
          if(global_save_var){        
-           pdf('Manuscript_figure3c_heatmap.pdf'
+           pdf(paste0('plots/Heatmap_pir_RMT-SBP_',
+                      list_name_to_order,
+                      'order.pdf')
                , width = 10
@@
                 ht_gap = unit(.35,'cm'),
-                merge_legend = TRUE
-           ))
+                merge_legend = TRUE)
+         )
        })
@@
 ```{r pca_calculation, results = TRUE}
-twentyfour_tpm <- twentyfour_rMT
-rownames_geneid <- twentyfour_tpm$Geneid
-twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
-rownames(twentyfour_tpm_forpca) <- rownames_geneid
-twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
+rownames_geneid <- chab_tpm.df$Geneid
+chab_tpm.df_forpca <- data.frame(select(chab_tpm.df, -Geneid))
+rownames(chab_tpm.df_forpca) <- rownames_geneid
+chab_tpm.df_pca <- prcomp(chab_tpm.df_forpca, 
+                              center = TRUE, 
+                              scale. = TRUE)
 #Need to look into what exactly scale and center are doing
@@
 ```{r 2d_pca, results = TRUE}
-twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
-twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
-
-twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
-                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
-                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
-                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
+chab_tpm.df_pca_var <- get_pca_var(chab_tpm.df_pca)
+chab_tpm.df_pca_coords <- data.frame(chab_tpm.df_pca_var$coord)
+
+chab_tpm.df_pca_coords <- data.frame(sample = rownames(chab_tpm.df_pca_var$coord),
+                                        PC1 = chab_tpm.df_pca_coords$Dim.1,
+                                        PC3 = chab_tpm.df_pca_coords$Dim.3,
+                                        PC2 = chab_tpm.df_pca_coords$Dim.2) %>% 
   dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,1])
-
-colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
+  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
+          time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture'))
+
+colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
 colornames <- setNames(colornames, 
-                       sort(unique(twentyfour_tpm_pca_coords$time)))
-
-pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
+                       sort(unique(chab_tpm.df_pca_coords$time)))
+
+pca_importance <- as.data.frame(summary(chab_tpm.df_pca)$importance)
 pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
@@
 (
-  plot <- ggplot(twentyfour_tpm_pca_coords,
+  plot <- ggplot(chab_tpm.df_pca_coords,
                  aes(
@@
                    y = PC2,
-                   col = time
+                   col = time,
+                   shape = transmission
                  )) +
@@
 ```{r}
-  cir_plot <- twentyfour_rMT %>% 
+  cir_plot <- chab_tpm.df %>% 
     filter(Geneid %in% 'PCHAS_0101200') %>% 
@@
     # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
-    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h')) %>% 
-    group_by(sample, time) %>% 
+    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
+           transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
+    group_by(sample, time, transmission) %>% 
     dplyr::summarise(sum_tpm = sum(tpm))
@@
               aes(x = time, 
-                  y = sum_tpm)) +
+                  y = sum_tpm,
+                  fill = transmission)) +
     stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
-                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge',
-                 fill = 'lightblue') +
+                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
     scale_y_continuous(breaks = y_scale) +

### 2022-07-22 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Commented out now but inserted code for PC1/PC3 figure

@@
 
+# (
+#   plot <- ggplot(chab_tpm.df_pca_coords,
+#                  aes(
+#                    x = PC3,
+#                    y = PC1,
+#                    col = time,
+#                    shape = transmission
+#                  )) +
+#     geom_point(size = 3,
+#                alpha = 0.8) +
+#     scale_color_manual(values = colornames) +
+#     xlab(paste0('PC3 ', round(pca_importance$PC3[2]*100, digits = 1),'%')) +
+#     ylab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
+#     guides(shape=guide_legend(ncol=2),
+#            col = guide_legend(ncol=2)) +
+#     theme_classic() +
+#     theme(legend.key.size = unit(4, 'mm'),
+#           legend.text = element_text(size = 7))
+# )
+
 if(global_save_var){

### 2022-07-22 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Fix and clean-up the all pir heatmap with RMT/SBP order

@@
 
-         combined_data_pir <- getElement(twentyfour_avgtpm, 
+         combined_data_pir <- getElement(chab_tpm_avg.list, 
                                      list_name)$tpm %>% 
@@
 
-         #Ordering that a more understandable order
-
-         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
-                                                            order_vector)),
-                                                drop = FALSE]
-
+         #Calculate the means, then arrange by mean tpm and sub-family, 
+         # then get the Geneid list from this
          combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                          means = apply(combined_data_pir, 1, mean),
-                                         subfam = cir_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
-                                                                            cir_info$Geneid)]) %>% 
-           filter(means > 1) %>% 
+                                         subfam = cir_info$SorL[
+                                           match(names(apply(combined_data_pir, 1, mean)), 
+                                                 cir_info$Geneid)
+                                         ]) %>% 
+           # filter(means > 1) %>% 
            arrange(subfam, 
@@
                                                       drop = FALSE]
-         #Add a column for the SorL
-         combined_data_pir_ht$SorL <- cir_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
-                                                                cir_info$Geneid)]
 
@@
        })
-names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)
-
+names(combined_data_pir_ht_list) <- names(chab_tpm_avg.list)
+```
+
+```{r heatmap}
 #Need to reach into the data to find the maximum value of the log(TPM+1) for 
@@
                max(
-                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
+                 log2(1+list_element)
                )
@@
          #Sub-family annotation
-         clade_info <- combined_data_pir_ht$SorL
+         clade_info <- combined_data_pir_ht
+                  #Add a column for the SorL
+         combined_data_pir_ht$SorL <- cir_info$SorL[match(rownames(combined_data_pir_ht), 
+                                                                cir_info$Geneid)]
          annotation_col <- SorL_col
@@
 
-                             combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
-                                                                list_name)$tpm %>% 
-                               column_to_rownames(var = 'Geneid')
+                             combined_data_pir_ht <- getElement(combined_data_pir_ht_list, 
+                                                                list_name)
 
@@
 
-```{r pca_calculation, results = TRUE}
+```{r pca_plot, results = TRUE}
 rownames_geneid <- chab_tpm.df$Geneid
@@
                               scale. = TRUE)
-#Need to look into what exactly scale and center are doing
-```
-
-```{r 2d_pca, results = TRUE}
+
 chab_tpm.df_pca_var <- get_pca_var(chab_tpm.df_pca)
@@
 
+```
+
+```{r pca_plot_withPseudogene}
+chab_tpm_wPseudo.df <- read_csv('data/Pchabaudi_TPM_all_withPseudo.csv')
+
+rownames_geneid <- chab_tpm_wPseudo.df$Geneid
+chab_tpm_wPseudo.df_forpca <- data.frame(select(chab_tpm_wPseudo.df, -Geneid))
+rownames(chab_tpm_wPseudo.df_forpca) <- rownames_geneid
+chab_tpm_wPseudo.df_pca <- prcomp(chab_tpm_wPseudo.df_forpca, 
+                              center = TRUE, 
+                              scale. = TRUE)
+
+chab_tpm_wPseudo.df_pca_var <- get_pca_var(chab_tpm_wPseudo.df_pca)
+chab_tpm_wPseudo.df_pca_coords <- data.frame(chab_tpm_wPseudo.df_pca_var$coord)
+
+chab_tpm_wPseudo.df_pca_coords <- data.frame(sample = rownames(chab_tpm_wPseudo.df_pca_var$coord),
+                                        PC1 = chab_tpm_wPseudo.df_pca_coords$Dim.1,
+                                        PC3 = chab_tpm_wPseudo.df_pca_coords$Dim.3,
+                                        PC2 = chab_tpm_wPseudo.df_pca_coords$Dim.2) %>% 
+  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
+  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
+          time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture'))
+
+colornames <- viridis(length(unique(chab_tpm_wPseudo.df_pca_coords$time)))
+colornames <- setNames(colornames, 
+                       sort(unique(chab_tpm_wPseudo.df_pca_coords$time)))
+
+pca_importance <- as.data.frame(summary(chab_tpm_wPseudo.df_pca)$importance)
+pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
+pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
+pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
+
+(
+  plot <- ggplot(chab_tpm_wPseudo.df_pca_coords,
+                 aes(
+                   x = PC1,
+                   y = PC2,
+                   col = time,
+                   shape = transmission
+                 )) +
+    geom_point(size = 3,
+               alpha = 0.8) +
+    scale_color_manual(values = colornames) +
+    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
+    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
+    guides(shape=guide_legend(ncol=2),
+           col = guide_legend(ncol=2)) +
+    theme_classic() +
+    theme(legend.key.size = unit(4, 'mm'),
+          legend.text = element_text(size = 7))
+)
+
+if(global_save_var){
+  ggsave2(filename = 'Manuscript_SuppFigure3_chabaudi24h_pca_wPseudo.png',
+          plot = plot,
+          dpi = 300,
+          width = 15,
+          height = 10,
+          units = 'cm')
+}
 ```

### 2022-07-22 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Fix ggsave2 to ggsave

@@
          #Sub-family annotation
-         clade_info <- combined_data_pir_ht
-                  #Add a column for the SorL
-         combined_data_pir_ht$SorL <- cir_info$SorL[match(rownames(combined_data_pir_ht), 
+         clade_info <- cir_info$SorL[match(rownames(combined_data_pir_ht), 
                                                           cir_info$Geneid)]
@@
 
-                             #Get the times instead of the stage names.
-                             colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h|sch.culture')
-
-                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                                          drop = FALSE]
+                             combined_data_pir_ht <- combined_data_pir_ht[
+                               ,order(colnames(combined_data_pir_ht)),
+                               drop = FALSE
+                             ]
 
@@
 if(global_save_var){
-  ggsave2(filename = 'Manuscript_SuppFigure3_chabaudi24h_pca_nopseudo.png',
+  ggsave2(filename = 'plots/pca_nopseudo.png',
           plot = plot,
@@
 if(global_save_var){
-  ggsave2(filename = 'Manuscript_SuppFigure3_chabaudi24h_pca_wPseudo.png',
+  ggsave(filename = 'plots/pca_wPseudo.png',
           plot = plot,

### 2022-07-22 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Add titles for PCA plots

@@
     theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7))
+          legend.text = element_text(size = 7)) +
+    ggtitle('PCA without RMT-high pseudogene')
 )
@@
 if(global_save_var){
-  ggsave2(filename = 'plots/pca_nopseudo.png',
+  ggsave(filename = 'plots/pca_nopseudo.png',
           plot = plot,
@@
     theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7))
+          legend.text = element_text(size = 7)) +
+    ggtitle('PCA with all genes')
 )

### 2022-07-22 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Use the times of the life cycle stages for the pseudogene TPM graph

@@
   mutate(time = str_extract(sample, 
-                            pattern = '([[:digit:]]{2}h_.+|sch.culture)(?=_[[:digit:]])'),
+                            pattern = '([[:digit:]]{2}h|sch.culture)'),
          transmission = str_extract(sample,
@@
        aes(x = time, 
-           y = log1p(tpm),
+           y = tpm,
            fill = transmission)) +

### 2022-07-24 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Save ancestral pir expression plot

@@
     scale_y_continuous(breaks = y_scale) +
+    scale_fill_manual(values = cbPalette) +
     geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
-    ylab(as.expression(bquote('Total'~italic(.('PCHAS_0101200'))~'TPM'))) +
+    ylab(as.expression(bquote('Total ancestral '~italic(.('pir'))~'TPM'))) +
     xlab('Asexual blood stage development cycle') +
@@
                                    vjust = 0.5,
-                                   size = 13))
+                                   size = 10))
 )
@@
 if(global_save_var){
-  ggsave2(filename = 'Manuscript_ancestralPir_BarChart.pdf',
+  ggsave(filename = 'plots/ancestralPir_BarChart.pdf',
           plot = p,

### 2022-07-24 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Start calculating z-scores

@@
 
+#Remove schizont culture from this as we are interested in z-scores
+# across the 24h cycle and schizont will distort this and don't fit into comparison.
+chab_tpm_avg_forZ.df <- select(chab_tpm_avg.df, 
+                               -rMT_sch.culture)
+chab_tpm_avg_forZ.list <- lapply(c('rMT', 'SBP'), 
+                                 function(transmission){
+                                   #Subset for the Geneid column and either rMT or SBP
+                                   tpm.df <- chab_tpm_avg_forZ.df[
+                                     ,
+                                     grepl(colnames(chab_tpm_avg_forZ.df), 
+                                           pattern = paste0('Geneid|', transmission))
+                                   ]
+                                   list(tpm = tpm.df,
+                                        experiment = transmission)
+                                 })
+
 #Make and save z-scores
-twentyfour_z <- lapply(twentyfour_avgtpm,
+chab_z.list <- lapply(chab_tpm_avg_forZ.list,
                       tpm_to_zscore)
@@
 # Want the data ordered by the highest zscore for each gene in each stage.
-phaseogram_row_order <- lapply(twentyfour_z, 
+phaseogram_row_order <- lapply(chab_z.list, 
                                function(list_elements){
@@
       width = 5)
-  draw_heatmap_zscore(zscore_list = twentyfour_z,
+  draw_heatmap_zscore(zscore_list = chab_z.list,
                       gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
@@
 
-draw_heatmap_zscore(zscore_list = twentyfour_z,
+draw_heatmap_zscore(zscore_list = chab_z.list,
                     gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
@@
 ```{r pir_gene_phaseogram_breaks}
-  max_tpm <- lapply(twentyfour_z, 
+  max_tpm <- lapply(chab_z.list, 
                     function(list_element){

### 2022-07-25 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Phaseogram heatmap now has the max/min tpm annotation working again

@@
 library("factoextra")
+library("MatrixGenerics") # for rowMins and rowMax
 
@@
 
+names(phaseogram_row_order) <- c('rMT', 'SBP')
+
 phaseogram_time_order <- lapply(names(phaseogram_row_order), 
@@
                                   order_by_time <- transmission_list$zscore_order %>% 
-                                    mutate(stage = times_stages_table$time[match(stage,
-                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
+                                    mutate(stage = str_extract(stage,
+                                                               pattern = '[[:digit:]]{2}h')) %>% 
                                     arrange(stage)
@@
 
+    #Set the filter_genes_vector as the genes to remove if it doesn't already exist.
+    if(is.null(filter_genes_vector)){
+      filter_genes_vector <- remove_genes
+    }else{
       filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
+    }
 
@@
                          #Get max/min TPM scores for the side annotation heatmap                       
-                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                         tpm_mat.mat <- list_element$tpm[match(rownames(heatmap_data),
                                                            list_element$tpm$Geneid),] %>%
@@
 
-                         tpm_min <- rowMins(tpm_mat)
-                         tpm_max <- rowMaxs(tpm_mat)
+                         tpm_min <- rowMins(tpm_mat.mat)
+                         tpm_max <- rowMaxs(tpm_mat.mat)
 
@@
                                                  col = tpm_col_fun,
-                                                 gp = gpar(col = 'black',
-                                                           lwd = 0.1),
-                                                 border = TRUE,
+                                                 # gp = gpar(col = 'black',
+                                                 #           lwd = 0.1),
+                                                 # border = TRUE,
                                                  simple_anno_size = unit(0.2, "cm")),
@@
 
-                       #Draw the left annotation is needed, only for the first element.
+                       #Draw the left annotation only for the first element of the heatmap list.
 
@@
       title = "Min/Max log(TPM + 1)", 
-      col = tpm_col_fun, 
+      col_fun = tpm_col_fun, 
       at = tpm_breaks, 
@@
   draw_heatmap_zscore(zscore_list = chab_z.list,
-                      gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
+                      gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                       use_times = TRUE,
@@
 draw_heatmap_zscore(zscore_list = chab_z.list,
-                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
+                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                     use_times = TRUE,
@@
                     show_row_names = FALSE,
-                    tpm_mat = FALSE)
+                    tpm_mat = TRUE)
 

### 2022-07-25 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Also fixed pir subfam annotation

@@
     }
-
   }
@@
 if(global_save_var){
-  pdf(file = 'Manuscript_SuppFig5_phaseogram_allgenes.pdf',
+  pdf(file = 'plots/phaseogram_allgenes_rMTorder.pdf',
       height = 5,
@@
                       show_row_names = FALSE,
-                      tpm_mat = FALSE)
+                      tpm_mat = TRUE)
   dev.off()
@@
   draw_heatmap_zscore(zscore_list = chab_z.list,
-                      gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
+                      gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                       filter_genes_vector = cir_id,
@@
 draw_heatmap_zscore(zscore_list = chab_z.list,
-                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
+                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                     filter_genes_vector = cir_id,

### 2022-07-25 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Save phaseogram plots in each order for all and pir genes

@@
 
+if(global_save_var){
+  pdf(file = 'plots/phaseogram_allgenes_SBPorder.pdf',
+      height = 5,
+      width = 5)
 draw_heatmap_zscore(zscore_list = chab_z.list,
@@
                     tpm_mat = TRUE)
+  dev.off()
+}
 
@@
 if(global_save_var){
- pdf(file = 'Manuscript_SuppFig6_phaseogram_cirgenes.pdf',
+ pdf(file = 'plots/phaseogram_pirgenes_rMTorder.pdf',
       height = 5,
@@
 
+if(global_save_var){
+ pdf(file = 'plots/phaseogram_pirgenes_SBPorder.pdf',
+      height = 5,
+      width = 5)
 draw_heatmap_zscore(zscore_list = chab_z.list,

### 2022-07-25 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Save zscore information and tables

@@
                                    heatmap_legend_param = list(
-                                     at = c(3,2,1,0,-1,-2,-3),
+                                     at = rev(c(3,2,1,0,-1,-2,-3)),
                                      labels_gp = gpar(fontsize = 10),
@@
       height = 5,
-      width = 5)
+      width = 7)
   draw_heatmap_zscore(zscore_list = chab_z.list,
@@
       height = 5,
-      width = 5)
+      width = 7)
 draw_heatmap_zscore(zscore_list = chab_z.list,
@@
 
+```
+
+```{r zscore_save}
+chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
+                        chab_z.list[[2]]$zscores, 
+                        by = 'Geneid')
+write.csv(chab_z.df,
+          file = 'data/Pchabaudi_z.csv',
+          row.names = FALSE)
+write.table(x = phaseogram_time_order$rMT$zscore_order$Geneid,
+            file = 'data/Pchabaudi_z_phaseOrder_rMT.list',
+            row.names = FALSE, col.names = FALSE)
+write.table(x = phaseogram_time_order$SBP$zscore_order$Geneid,
+            file = 'data/Pchabaudi_z_phaseOrder_SBP.list',
+            row.names = FALSE, col.names = FALSE)
 ```

### 2022-07-25 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Plot and analysis of phase differences of genes between rMT and SBP

@@
 
+```{r gene_sync}
+
+#Which genes are out of sync between rMT and SBP?
+head(chab_z.list[[1]]$zscores[,-1])
+
+#Melt the z-score dataframe into long form, then slice out the maximum z-score value.
+# Filter out any gene which has multiple max values for simplicity.
+# Summarise this with the numbers that each stage (the max stage) corresponds to.
+
+z_max <- lapply(chab_z.list, function(x) {
+  x$zscores %>% 
+  reshape2::melt(id.vars = 'Geneid', 
+       variable.name = 'sample', 
+       value.name = 'zscore') %>% 
+  group_by(Geneid) %>% 
+  slice_max(order_by = zscore, n = 1) %>% 
+  filter(n() == 1) %>% 
+  summarise(sample_max = times_stages_table$num[
+    str_detect(sample, 
+               times_stages_table$time)
+    ])
+})
+
+#Bind the two together with inner_join so that it only includes genes in both.
+# Then calculate the peak difference by subtracting the difference between the sample numbers.
+# Since this is a cycle, no point of the cycle is theoretically 
+#  more than 4 time points away from the next one. 
+#  e.g. 23h and 02h (8 and 1) are only 1 time point apart.
+#Hence we subtract values above four from eight
+z_max_diff <- inner_join(z_max[[1]], z_max[[2]], by = 'Geneid') %>% 
+  mutate(diff = ifelse(abs(sample_max.x - sample_max.y) > 4,
+                       8 - abs(sample_max.x - sample_max.y),
+                       abs(sample_max.x - sample_max.y)))
+
+#Probably unnecessarily complicated code to make a barplot of how many genes 
+# are peaking at different timepoints between rMT and SBP
+barplot(group_by(z_max_diff, diff) %>% summarise(total = n()) %>% .$total,
+        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))
+
+```
+
 ## Figure 4 and supplementary figure 7

### 2022-07-26 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Plot of cir genes out of phase too
Add comments and start GO enrichment analysis

@@
 #Which genes are out of sync between rMT and SBP?
-head(chab_z.list[[1]]$zscores[,-1])
 
@@
 #  more than 4 time points away from the next one. 
-#  e.g. 23h and 02h (8 and 1) are only 1 time point apart.
-#Hence we subtract values above four from eight
+#  e.g. 23h and 02h (8 and 1) are only 1 time point apart, not seven.
+#  Hence we subtract values above four from eight
 z_max_diff <- inner_join(z_max[[1]], z_max[[2]], by = 'Geneid') %>% 
@@
 
+#And for the pir genes
+barplot(group_by(z_max_diff, diff) %>% 
+          filter(Geneid %in% cir_id) %>% 
+          summarise(total = n()) %>% 
+          .$total,
+        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))
+
+```
+
+```{r phase_go_enrich}
+#Get the gene list for each group of genes in and out of phase.
+z_max_diff.list <- lapply(seq(0,4), 
+                          function(diff_num) filter(z_max_diff, diff == diff_num) %>% 
+                            .$Geneid)
+names(z_max_diff.list) <- seq(0,4)
 ```

### 2022-07-26 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Move through GO enrichment analysis

@@
 library("MatrixGenerics") # for rowMins and rowMax
+library("topGO")
+library("FoldGO")
 
@@
 #Get the gene list for each group of genes in and out of phase.
+gene.list <- chabaudi_gene_info$Geneid
 z_max_diff.list <- lapply(seq(0,4), 
-                          function(diff_num) filter(z_max_diff, diff == diff_num) %>% 
-                            .$Geneid)
+                          function(diff_num) {
+                            diff.list <- filter(z_max_diff, diff == diff_num) %>% 
+                              .$Geneid
+                            #Get 1 for on the list and 0 for not.
+                            gene_diff.list <- as.integer(gene.list %in% diff.list)
+                            names(gene_diff.list) <- gene.list
+                            return(gene_diff.list)
+                          })
 names(z_max_diff.list) <- seq(0,4)
+
+#Read in the GO-Annotation file using GAFReader from FoldGO.
+# getAnnotation to obtain a list of the GO terms and associated genes.
+# inverseList (from topGO) to turn this into list of geneID with associated GO terms.
+gaf <- GAFReader("../standard_files/PlasmoDB-48_Pchabaudichabaudi_GO.gaf")
+go_annotation <- inverseList(getAnnotation(gaf))
+
+#Make a topGOdata object 
+GOdata <- new("topGOdata", 
+              ontology = "CC", 
+              geneSel = function(x){x == 1},
+              allGenes = z_max_diff.list$`0`,
+              annot = annFUN.gene2GO, 
+              gene2GO = go_annotation)
+
 ```

### 2022-07-26 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Summary table of top GO terms for Cellular Components

@@
               gene2GO = go_annotation)
-
+resultFis <- runTest(GOdata, algorithm = "classic", statistic = "fisher")
+hist(score(resultFis), 50, xlab = "p-values")
+GenTable(GOdata, 
+         classic = resultFis, 
+         orderBy = "classic", 
+         ranksOf = "classic",
+         topNodes = 25)
 ```

### 2022-07-26 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
GO enrichment of terms across the phases and looking at how many genes don't have any GO terms

--- 2022-07-26 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
@@
 
+# How many genes for each group do not have any GO annotations?
+
+lapply(z_max_diff.list,
+       function(gene_diff.list){
+         total_noGO <- sum(!names(gene_diff.list[gene_diff.list == 1]) %in% 
+                             names(go_annotation))
+         prop_noGO <- total_noGO/sum(gene_diff.list == 1)
+         return(paste0(total_noGO, ' genes without annotation = ',
+                       round(prop_noGO*100, digits = 2), '%'))
+       })
+
+```
+
+```{r}
 #Make a topGOdata object 
@@
               allGenes = z_max_diff.list$`0`,
+              nodeSize = 5,
               annot = annFUN.gene2GO, 
               gene2GO = go_annotation)
-
-```
+resultFis <- runTest(GOdata, 
+                     algorithm = "classic", 
+                     statistic = "fisher")
+hist(score(resultFis), 50, xlab = "p-values")
+GenTable(GOdata, 
+         classic = resultFis, 
+         orderBy = "classic", 
+         ranksOf = "classic",
+         topNodes = 25) %>% 
+  filter(classic < 0.05)
+
+lapply(seq(0,4), 
+       function(phase){
+         GOdata <- new("topGOdata", 
+                       ontology = "BP", 
+                       geneSel = function(x){x == 1},
+                       allGenes = getElement(z_max_diff.list, paste(phase)),
+                       nodeSize = 5,
+                       annot = annFUN.gene2GO, 
+                       gene2GO = go_annotation)
+         resultFis <- runTest(GOdata, 
+                              algorithm = "classic", 
+                              statistic = "fisher")
+         hist(score(resultFis), 50, xlab = "p-values")
+         GenTable(GOdata, 
+                  classic = resultFis, 
+                  orderBy = "classic", 
+                  ranksOf = "classic",
+                  topNodes = 25) %>% 
+           filter(classic < 0.05)
+       })
+
+# showSigOfNodes(GOdata, 
+#                score(resultFis), 
+#                firstSigNodes = 5, 
+#                useInfo = 'all')
+```
+
+Noticeable enrichments of genes that are:
+* In-phase: DNA and nucleotide processes, and amino acid processing, such as DNA replication, tRNA aminoacylation and amino acid activation.
+* 1\*3h Out-of-phase: Translation and peptide biosynthesis, gene expression and ribosome biogenesis.
+* 2\*3h Out-of-phase: (Based on much fewer genes) Metabolism of carbohydrates and respiration.
+* 3\*3h Out-of-phase: (Based on even fewerer genes) More metabolism and lipid biosynthesis.
+* 4\*3h Out-of-phase: (Based on a few more genes actually) mRNA processing.
 

### 2022-07-27 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Further investigations into the phase differences

@@
 ```{r}
+#Load for the showSigOfNodes network function.
+library('Rgraphviz')
+
 #Make a topGOdata object 
 GOdata <- new("topGOdata", 
-              ontology = "CC", 
+              ontology = "BP", 
               geneSel = function(x){x == 1},
-              allGenes = z_max_diff.list$`0`,
+              allGenes = z_max_diff.list$`1`,
               nodeSize = 5,
@@
                      statistic = "fisher")
+plot.new()
 hist(score(resultFis), 50, xlab = "p-values")
-GenTable(GOdata, 
+dev.off()
+showSigOfNodes(GOdata,
+               score(resultFis),
+               firstSigNodes = 5,
+               useInfo = 'all')
+goRes <- GenTable(GOdata, 
          classic = resultFis, 
@@
   filter(classic < 0.05)
+
+#Get first (most significant) enriched GO id 
+goID <- goRes[4, "GO.ID"]
+#Show distribution of GO term genes compared to null distribution (all other genes)
+print(showGroupDensity(GOdata, goID, ranks = TRUE))
+
+#Unlist the lapply below which goes through each element of the list and 
+# returns TRUE is the go_id is listed for that gene.
+#This makes a TRUE/FALSE index for the list/names of the list which can get 
+# the genes associated with that GO id.
+goID_genes <- names(go_annotation)[unlist(
+  lapply(go_annotation, 
+         function(go_ids){
+           any(goID %in% go_ids)
+         })
+)]
+
+draw_heatmap_zscore(zscore_list = chab_z.list,
+                    filter_genes_vector = names(z_max_diff.list$`1`[z_max_diff.list$`2` == 1]),
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = TRUE)
+```
+
+```{r}
 
@@
                               statistic = "fisher")
-         hist(score(resultFis), 50, xlab = "p-values")
+         draw_heatmap_zscore(zscore_list = chab_z.list,
+                             filter_genes_vector = names(
+                               getElement(z_max_diff.list, paste(phase))[
+                                 getElement(z_max_diff.list, paste(phase)) == 1
+                               ]
+                             ),
+                             gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                             threshold_tpm = 1,
+                             use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = TRUE)
          GenTable(GOdata, 
@@
 
-# showSigOfNodes(GOdata, 
-#                score(resultFis), 
-#                firstSigNodes = 5, 
-#                useInfo = 'all')
 ```
@@
 * 4\*3h Out-of-phase: (Based on a few more genes actually) mRNA processing.
+
+Fore node visualisation:
+"Rectangles indicate the 5 most signicant terms. Rectangle color represents the relative signicance,
+ranging from dark red (most signicant) to bright yellow (least signicant). For each node, some basic information
+is displayed"
 

### 2022-07-27 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Move draw_heatmap_zscore to a separate function to source

@@
 source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
+source('../rfunctions/draw_heatmap_zscore.R')
 ```
@@
 
-```{r function_draw_heatmap_zscore}
-
-draw_heatmap_zscore <- function(zscore_list,
-                                title = '',
-                                gene_ordering_vector = NULL,
-                                filter_genes_vector = NULL,
-                                extra_title = NULL,
-                                annotation_clade = NULL,
-                                annotation_col,
-                                threshold_tpm = NULL,
-                                use_times = FALSE,
-                                use_order_vector = TRUE,
-                                tpm_mat = TRUE,
-                                tpm_breaks = NULL,
-                                ...){
-  if(is.null(tpm_breaks) & tpm_mat){
-    #Need the generate the breaks in the data for the heatmap legend if not provided.
-    max_tpm <- lapply(zscore_list, 
-                      function(list_element){
-                        #list_element <- zscore_list[[1]]
-                        heatmap_data <- as.data.frame(list_element$zscores)
-                        title <- list_element$experiment
-
-                        if(!'Geneid' %in% colnames(heatmap_data)){
-                          stop('No Geneid column in the heatmap data.frame')
-                        }
-
-                        if(!is.null(filter_genes_vector)){
-                          heatmap_data <- as.data.frame(
-                            filter(heatmap_data, 
-                                   Geneid %in% filter_genes_vector)
-                          )
-                        }
-
-                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                        if(!is.null(gene_ordering_vector)){
-                          #Remove genes not in the main data frame but in the ordering vector, 
-                          # and then order based on the vector
-                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          heatmap_data <- heatmap_data[na.omit(
-                            match(gene_ordering_vector, 
-                                  rownames(heatmap_data))
-                          ),,
-                          drop=FALSE]
-                        }
-
-                        geneids <- rownames(heatmap_data)
-
-                        #Get max/min TPM scores for the side annotation heatmap                       
-                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                          list_element$tpm$Geneid),] %>% 
-                          select(-Geneid) %>% 
-                          as.matrix %>% 
-                          log1p
-                        rownames(tpm_mat) <- geneids
-
-                        tpm_min <- rowMins(tpm_mat)
-                        names(tpm_min) <- geneids
-                        tpm_max <- rowMaxs(tpm_mat)
-                        names(tpm_max) <- geneids
-                        return(list(tpm_max = tpm_max,
-                                    tpm_min = tpm_min))
-                      })
-
-    tpm_max_all <- mround(
-      max(
-        unlist(
-          lapply(max_tpm, 
-                 getElement,
-                 'tpm_max')
-        )
-      ), 
-      base = 1)
-
-    tpm_breaks <- round(
-      c(0,
-        tpm_max_all/4,
-        tpm_max_all/2,
-        tpm_max_all*(3/4),
-        tpm_max_all),
-      digits = 2
-    )
-  }
-
-  #Generate the colour function from the TPM legend breaks for the TPM annotation
-  if(tpm_mat){
-    tpm_col_fun <- colorRamp2(
-      breaks = tpm_breaks,
-      colors = viridis(length(tpm_breaks))
-    )
-  }
-
-  if(is.numeric(threshold_tpm)){
-    #Apply a threshold to the TPM in order to remove low expressed genes.
-    # Remember to add log1p here! 
-    # Threshold uses the max value, so it ony needs to be above thresh 
-    #   in one stage avg in one transmission method.
-    #Will alter the gene_filter vector in order to only select the genes above threshold
-
-    gene_maxs <- bind_rows(
-      lapply(max_tpm, 
-             getElement,
-             'tpm_max')
-    )
-
-    remove_genes <- names(gene_maxs)[apply(
-      gene_maxs, 
-      2, 
-      function(maxs){
-        all(maxs < log2(1+threshold_tpm))
-      }
-    )]
-
-    #Set the filter_genes_vector as the genes to remove if it doesn't already exist.
-    if(is.null(filter_genes_vector)){
-      filter_genes_vector <- remove_genes
-    }else{
-      filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
-    }
-  }
-
-  if(is.character(annotation_clade)){
-    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
-    clade_info <- getElement(cir_info,
-                             annotation_clade)[match(order_cir, 
-                                                     cir_info$Geneid)]
-    names(clade_info) <- str_split_fixed(order_cir, 
-                                         pattern = '_', 
-                                         n = 2)[,2]
-    annotation_col <- annotation_col
-  }
-
-  heatmaps <- lapply(zscore_list, 
-                     function(list_element){
-                       #list_element <- zscore_list[[1]]
-                       heatmap_data <- as.data.frame(list_element$zscores)
-                       title <- list_element$experiment
-
-                       if(!'Geneid' %in% colnames(heatmap_data)){
-                         stop('No Geneid column in the heatmap data.frame')
-                       }
-
-                       if(!is.null(filter_genes_vector)){
-                         heatmap_data <- as.data.frame(
-                           filter(heatmap_data, 
-                                  Geneid %in% filter_genes_vector)
-                         )
-                       }
-
-                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                       if(!is.null(gene_ordering_vector)){
-                         #Remove genes not in the main data frame but in the ordering vector,
-                         #  and then order based on the vector
-                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                         heatmap_data <- heatmap_data[na.omit(
-                           match(gene_ordering_vector, 
-                                 rownames(heatmap_data))
-                         ),,
-                         drop=FALSE]
-                       }
-
-                       if(tpm_mat){
-                         #Get max/min TPM scores for the side annotation heatmap                       
-                         tpm_mat.mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                           list_element$tpm$Geneid),] %>%
-                           select(-Geneid) %>% 
-                           as.matrix %>% 
-                           log1p
-
-                         tpm_min <- rowMins(tpm_mat.mat)
-                         tpm_max <- rowMaxs(tpm_mat.mat)
-
-                         names(tpm_min) = rownames(heatmap_data)
-                         names(tpm_max) = rownames(heatmap_data)
-
-                         tpm_ha <- rowAnnotation(
-                           min.max = anno_simple(cbind(tpm_min, 
-                                                       tpm_max), 
-                                                 col = tpm_col_fun,
-                                                 # gp = gpar(col = 'black',
-                                                 #           lwd = 0.1),
-                                                 # border = TRUE,
-                                                 simple_anno_size = unit(0.2, "cm")),
-                           show_annotation_name = FALSE)
-                       }
-
-                       #Draw the left annotation only for the first element of the heatmap list.
-
-                       if(
-                         identical(list_element,
-                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
-                       ){
-                         clade_ha <- rowAnnotation(
-                           'subfamily' = anno_simple(
-                             clade_info,
-                             col = annotation_col
-                           )
-                         )
-                       }
-
-                       #Change rownames to remove PCHAS_
-                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
-                                                                pattern = '_', 
-                                                                n = 2)[,2]
-
-                       if(use_times){
-                         colnames(heatmap_data) <- str_extract(colnames(heatmap_data), 
-                                                               pattern = '[[:digit:]]{2}h')
-                       }
-
-                       if(use_order_vector){
-                         #Order the columns by ordering_vector
-                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
-                                                                   order_vector)),
-                                                      drop = FALSE]
-                       }else{
-                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
-                                                      drop = FALSE] 
-                       }
-
-                       h = Heatmap(heatmap_data,
-                                   name = paste0('z-score',
-                                                 extra_title),
-                                   col = c(cbPalette[6],
-                                           'white', 
-                                           cbPalette[7]),
-                                   row_names_gp = gpar(fontsize = ifelse(
-                                     nrow(heatmap_data) > 13, 
-                                     12*(13/nrow(heatmap_data)), 
-                                     12)
-                                   ),
-                                   column_names_gp = gpar(fontsize = 10),
-                                   row_dend_width = unit(3, "cm"),
-                                   right_annotation = if(tpm_mat){
-                                     tpm_ha
-                                   }else{
-                                     NULL
-                                   },
-                                   left_annotation = 
-                                     if(identical(list_element,
-                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
-                                       clade_ha
-                                     }else{
-                                       NULL
-                                     }, 
-                                   #Only want the clade annotation when it is first heatmap
-                                   width = unit(4.5, 'cm'),
-                                   cluster_columns = FALSE,
-                                   height = unit(7,'cm'),
-                                   row_title = NULL,
-                                   cluster_rows = FALSE,
-                                   column_title = paste(title),
-                                   heatmap_legend_param = list(
-                                     at = rev(c(3,2,1,0,-1,-2,-3)),
-                                     labels_gp = gpar(fontsize = 10),
-                                     title_gp = gpar(fontsize = 10)
-                                   ),
-                                   ...
-                       )
-                       return(h)
-                     })
-
-  ht_list <- Reduce('+', heatmaps) #To add the list elements together
-
-  if(tpm_mat){
-    #Draw the legend
-    tpm_lgd <- Legend(
-      title = "Min/Max log(TPM + 1)", 
-      col_fun = tpm_col_fun, 
-      at = tpm_breaks, 
-      labels = tpm_breaks,
-      title_gp = gpar(fontsize = 7)
-    )
-  }
-
-  if(is.character(annotation_clade)){
-    clade_lgd <- Legend(
-      title = 'sub-family',
-      at = names(annotation_col),
-      legend_gp = gpar(fill = annotation_col)
-    )
-  }
-
-  draw(ht_list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'left',
-       #Only include the annotation legends for the annotations we actually have.
-       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
-         list(tpm_lgd,
-              clade_lgd)
-       }else if(tpm_mat & !is.character(annotation_clade) ){
-         list(tpm_lgd)
-       } else if(!tpm_mat & is.character(annotation_clade) ){
-         list(clade_lgd)
-       })
-}
-```
-
 ## Supplementary figure 5 - Phaseograms of the entire genome.

### 2022-07-27 rfunctions/draw_heatmap_zscore.R

@@ -1,0 +1,305 @@
+draw_heatmap_zscore <- function(zscore_list,
+                                title = '',
+                                gene_ordering_vector = NULL,
+                                filter_genes_vector = NULL,
+                                extra_title = NULL,
+                                annotation_clade = NULL,
+                                annotation_col,
+                                threshold_tpm = NULL,
+                                use_times = FALSE,
+                                use_order_vector = TRUE,
+                                tpm_mat = TRUE,
+                                tpm_breaks = NULL,
+                                ...){
+  if(is.null(tpm_breaks) & tpm_mat){
+    #Need the generate the breaks in the data for the heatmap legend if not provided.
+    max_tpm <- lapply(zscore_list, 
+                      function(list_element){
+                        #list_element <- zscore_list[[1]]
+                        heatmap_data <- as.data.frame(list_element$zscores)
+                        title <- list_element$experiment
+                        
+                        if(!'Geneid' %in% colnames(heatmap_data)){
+                          stop('No Geneid column in the heatmap data.frame')
+                        }
+                        
+                        if(!is.null(filter_genes_vector)){
+                          heatmap_data <- as.data.frame(
+                            filter(heatmap_data, 
+                                   Geneid %in% filter_genes_vector)
+                          )
+                        }
+                        
+                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+                        
+                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+                        
+                        if(!is.null(gene_ordering_vector)){
+                          #Remove genes not in the main data frame but in the ordering vector, 
+                          # and then order based on the vector
+                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                          heatmap_data <- heatmap_data[na.omit(
+                            match(gene_ordering_vector, 
+                                  rownames(heatmap_data))
+                          ),,
+                          drop=FALSE]
+                        }
+                        
+                        geneids <- rownames(heatmap_data)
+                        
+                        #Get max/min TPM scores for the side annotation heatmap                       
+                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
[…]

### 2022-07-27 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_phaseGenes.Rmd
Start moving phase/GO enrichment code to its own Rmd file

--- 2022-07-27 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
@@
 ---
-title: "Manuscript_Github_Pchabaudi_pir"
+title: "Manuscript_Github_Pchabaudi_phaseGene"
 author: "Timothy Little"
-date: "22/03/2021"
+date: "27/07/2022"
 output: html_document
@@
 library("FoldGO")
-
-# library('DESeq2')
-# library("pheatmap")
-# 
-# library('gridExtra')
-# 
-# library('corrplot')
-# 
-# library('cluster')    # clustering algorithms
-# library('factoextra') # clustering algorithms & visualization
-# library('ggfortify') # so ggplot understands prcomp objects
-# library('ape') # for phylogeny plotting
-# library('dendextend') # for dendrogram plotting
-# library('readxl') # for reading in Excel spreadsheets
-# library('plotly')
-# library('scales')
-# library('processx') # for downloading the plotly plots
-# library('cowplot')
-# 
-# library('ggbiplot')
-# library('ggpattern')
 
@@
 
+```{r import_z}
+chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
+                        chab_z.list[[2]]$zscores, 
+                        by = 'Geneid')
+chab_z.df <- read.csv(file = 'data/Pchabaudi_z.csv')
+
+chab_z.list <- list('rMT' = list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
+                                                                             pattern = 'Geneid|rMT')]),
+                    'SBP' = list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
+                                                                             pattern = 'Geneid|SBP')]))
+
+order_z_rMT <- read.table(file = 'data/Pchabaudi_z_phaseOrder_rMT.list')[,1]
+order_z_SBP <- read.table(file = 'data/Pchabaudi_z_phaseOrder_SBP.list')[,1]
+
+```
+
 ```{r colours_clades_phase}
@@
 
-```
-
-## Figure 3c - Highest expressed all _pir_ heatmap
-
-```{r highest_expressed_stage_heatmap_fig3c}
-
-num_SorL <- cir_info %>% 
-  group_by(SorL) %>% 
-  summarise(num = n_distinct(Gene.ID))
-
-labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
-                        num_SorL$SorL)
-
-# Get order of pir genes by highest expressed genes in each sample.
-
-combined_data_pir_ht_list <- lapply(names(chab_tpm_avg.list), 
-       function(list_name){
-
-         combined_data_pir <- getElement(chab_tpm_avg.list, 
-                                     list_name)$tpm %>% 
-           filter(Geneid %in% cir_id) %>% 
-           column_to_rownames(var = 'Geneid')
-
-         #Calculate the means, then arrange by mean tpm and sub-family, 
-         # then get the Geneid list from this
-         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
-                                         means = apply(combined_data_pir, 1, mean),
-                                         subfam = cir_info$SorL[
-                                           match(names(apply(combined_data_pir, 1, mean)), 
-                                                 cir_info$Geneid)
-                                         ]) %>% 
-           # filter(means > 1) %>% 
-           arrange(subfam, 
-                   desc(means)) %>% 
-           .$Geneid
-
-         #Order the heatmap data by the order defined above (SorL, then mean expression).
-         combined_data_pir_ht <- combined_data_pir[
-           order(
-             match(rownames(combined_data_pir),
-                   combined_data_ordered),
-             na.last = NA #Remove NAs
-           ),
-         ]
-
-         #Change the colnames to the times
-         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                       pattern = '[[:digit:]]{2}h|sch.culture')
-
-         #Order by time point
-         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                      drop = FALSE]
-
-         return(combined_data_pir_ht)
-       })
-names(combined_data_pir_ht_list) <- names(chab_tpm_avg.list)
-```
-
-```{r heatmap}
-#Need to reach into the data to find the maximum value of the log(TPM+1) for 
-# the colouring in order to maintain consistency.
-
-maxcol <- ceiling(
-  max(
-    unlist(
-      lapply(combined_data_pir_ht_list,
-             function(list_element){
-               max(
-                 log2(1+list_element)
-               )
-             }
-      )
-    )
-  )
-)
-
-lapply(names(combined_data_pir_ht_list),
-       function(list_name_to_order){
-
-         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
-                                            list_name_to_order)
-
-         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
-         #Sub-family annotation
-         clade_info <- cir_info$SorL[match(rownames(combined_data_pir_ht), 
-                                                          cir_info$Geneid)]
-         annotation_col <- SorL_col
-
-         #Subfamily annotation
-
-         clade_ha <- rowAnnotation(
-           'sub-fam' = anno_simple(
-             clade_info,
-             col = annotation_col,
-           ),
-           annotation_name_gp = gpar(fontsize = 10)
-         )
-
-         #lapply across the datasets now, for each order
-         ht_list <- lapply(names(combined_data_pir_ht_list),
-                           function(list_name){
-
-                             combined_data_pir_ht <- getElement(combined_data_pir_ht_list, 
-                                                                list_name)
-
-                             combined_data_pir_ht <- combined_data_pir_ht[
-                               ,order(colnames(combined_data_pir_ht)),
-                               drop = FALSE
-                             ]
-
-                             Heatmap(
-                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
-                               cluster_rows = FALSE,
-                               cluster_columns = FALSE,
-                               show_row_names = FALSE,
-                               name = 'log(TPM+1)', 
-                               col = colorRamp2(c(0,
-                                                  maxcol/4,
-                                                  maxcol/2,
-                                                  3*maxcol/4,
-                                                  maxcol),
-                                                viridis(5)),
-                               na_col = 'white', 
-                               border = TRUE,
-                               border_gp = gpar(lwd = 0.1),
-                               column_names_gp = gpar(fontsize = 10),
-                               row_dend_width = unit(3, "cm"),
-                               width = unit(4.5, 'cm'),
-                               height = unit(7,'cm'),
-                               row_title = NULL,
-                               split = clade_info,
-                               left_annotation = if(
-                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
-                               ){
-                                 clade_ha
-                               }else{
-                                 NULL
-                               },
-                               column_title = paste(list_name),
-                               heatmap_legend_param = list(
-                                 labels_gp = gpar(fontsize = 10),
-                                 title_gp = gpar(fontsize = 10,
-                                                 fontface = 'bold'),
-                                 at = c(0,
-                                        maxcol/4,
-                                        maxcol/2,
-                                        3*maxcol/4,
-                                        maxcol)
-                               )
-                             )
-                           }
-         )
-         ht_list <- Reduce(`+`, ht_list)
-         clade_lgd <- Legend(
-           title = 'sub-fam',
-           at = labels_SorL,
-           legend_gp = gpar(fill = SorL_col)
-         )
-
-         if(global_save_var){        
-           pdf(paste0('plots/Heatmap_pir_RMT-SBP_',
-                      list_name_to_order,
-                      'order.pdf')
-               , width = 10
-               , height = 4
-           )
-           draw(ht_list, 
-                row_title_side = 'left',
-                column_title_side = 'bottom',
-                heatmap_legend_side = 'right',
-                annotation_legend_side = 'right',
-                annotation_legend_list = list(clade_lgd),
-                ht_gap = unit(.35,'cm'),
-                merge_legend = TRUE)
-           dev.off()
-         }
-
-         return(
-           draw(ht_list, 
-                row_title_side = 'left',
-                column_title_side = 'bottom',
-                heatmap_legend_side = 'right',
-                annotation_legend_side = 'right',
-                annotation_legend_list = list(clade_lgd),
-                ht_gap = unit(.35,'cm'),
-                merge_legend = TRUE)
-         )
-       })
-
-```
-
-## 2D plot of PC1 v PC2 - Supp figure 4A
-
-```{r pca_plot, results = TRUE}
-rownames_geneid <- chab_tpm.df$Geneid
-chab_tpm.df_forpca <- data.frame(select(chab_tpm.df, -Geneid))
-rownames(chab_tpm.df_forpca) <- rownames_geneid
-chab_tpm.df_pca <- prcomp(chab_tpm.df_forpca, 
-                              center = TRUE, 
-                              scale. = TRUE)
-
-chab_tpm.df_pca_var <- get_pca_var(chab_tpm.df_pca)
-chab_tpm.df_pca_coords <- data.frame(chab_tpm.df_pca_var$coord)
-
-chab_tpm.df_pca_coords <- data.frame(sample = rownames(chab_tpm.df_pca_var$coord),
-                                        PC1 = chab_tpm.df_pca_coords$Dim.1,
-                                        PC3 = chab_tpm.df_pca_coords$Dim.3,
-                                        PC2 = chab_tpm.df_pca_coords$Dim.2) %>% 
-  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
-          time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture'))
-
-colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(chab_tpm.df_pca_coords$time)))
-
-pca_importance <- as.data.frame(summary(chab_tpm.df_pca)$importance)
-pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
-pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
-pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
-
-(
-  plot <- ggplot(chab_tpm.df_pca_coords,
-                 aes(
-                   x = PC1,
-                   y = PC2,
-                   col = time,
-                   shape = transmission
-                 )) +
-    geom_point(size = 3,
-               alpha = 0.8) +
-    scale_color_manual(values = colornames) +
-    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-    guides(shape=guide_legend(ncol=2),
-           col = guide_legend(ncol=2)) +
-    theme_classic() +
-    theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7)) +
-    ggtitle('PCA without RMT-high pseudogene')
-)
-
-# (
-#   plot <- ggplot(chab_tpm.df_pca_coords,
-#                  aes(
-#                    x = PC3,
-#                    y = PC1,
-#                    col = time,
-#                    shape = transmission
-#                  )) +
-#     geom_point(size = 3,
-#                alpha = 0.8) +
-#     scale_color_manual(values = colornames) +
-#     xlab(paste0('PC3 ', round(pca_importance$PC3[2]*100, digits = 1),'%')) +
-#     ylab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-#     guides(shape=guide_legend(ncol=2),
-#            col = guide_legend(ncol=2)) +
-#     theme_classic() +
-#     theme(legend.key.size = unit(4, 'mm'),
-#           legend.text = element_text(size = 7))
-# )
-
-if(global_save_var){
-  ggsave(filename = 'plots/pca_nopseudo.png',
-          plot = plot,
-          dpi = 300,
-          width = 15,
-          height = 10,
-          units = 'cm')
-}
-
-```
-
-```{r pca_plot_withPseudogene}
-chab_tpm_wPseudo.df <- read_csv('data/Pchabaudi_TPM_all_withPseudo.csv')
-
-rownames_geneid <- chab_tpm_wPseudo.df$Geneid
-chab_tpm_wPseudo.df_forpca <- data.frame(select(chab_tpm_wPseudo.df, -Geneid))
-rownames(chab_tpm_wPseudo.df_forpca) <- rownames_geneid
-chab_tpm_wPseudo.df_pca <- prcomp(chab_tpm_wPseudo.df_forpca, 
-                              center = TRUE, 
-                              scale. = TRUE)
-
-chab_tpm_wPseudo.df_pca_var <- get_pca_var(chab_tpm_wPseudo.df_pca)
-chab_tpm_wPseudo.df_pca_coords <- data.frame(chab_tpm_wPseudo.df_pca_var$coord)
-
-chab_tpm_wPseudo.df_pca_coords <- data.frame(sample = rownames(chab_tpm_wPseudo.df_pca_var$coord),
-                                        PC1 = chab_tpm_wPseudo.df_pca_coords$Dim.1,
-                                        PC3 = chab_tpm_wPseudo.df_pca_coords$Dim.3,
-                                        PC2 = chab_tpm_wPseudo.df_pca_coords$Dim.2) %>% 
-  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
-          time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture'))
-
-colornames <- viridis(length(unique(chab_tpm_wPseudo.df_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(chab_tpm_wPseudo.df_pca_coords$time)))
-
-pca_importance <- as.data.frame(summary(chab_tpm_wPseudo.df_pca)$importance)
-pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
-pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
-pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
-
-(
-  plot <- ggplot(chab_tpm_wPseudo.df_pca_coords,
-                 aes(
-                   x = PC1,
-                   y = PC2,
-                   col = time,
-                   shape = transmission
-                 )) +
-    geom_point(size = 3,
-               alpha = 0.8) +
-    scale_color_manual(values = colornames) +
-    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-    guides(shape=guide_legend(ncol=2),
-           col = guide_legend(ncol=2)) +
-    theme_classic() +
-    theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7)) +
-    ggtitle('PCA with all genes')
-)
-
-if(global_save_var){
-  ggsave(filename = 'plots/pca_wPseudo.png',
-          plot = plot,
-          dpi = 300,
-          width = 15,
-          height = 10,
-          units = 'cm')
-}
-```
-
-# Bar chart of ancestral _pir_ expression.
-
-```{r}
-  cir_plot <- chab_tpm.df %>% 
-    filter(Geneid %in% 'PCHAS_0101200') %>% 
-    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
-    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
-           transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
-    group_by(sample, time, transmission) %>% 
-    dplyr::summarise(sum_tpm = sum(tpm))
-
-y_scale <- seq(0, 
-               signif(ceiling(max(cir_plot$sum_tpm)), 
-                      digits = 2), 
-               length.out = 5)
-
-(
-  p <- ggplot(data = cir_plot, 
-              aes(x = time, 
-                  y = sum_tpm,
-                  fill = transmission)) +
-    stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
-                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
-    scale_y_continuous(breaks = y_scale) +
-    scale_fill_manual(values = cbPalette) +
-    geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
-    ylab(as.expression(bquote('Total ancestral '~italic(.('pir'))~'TPM'))) +
-    xlab('Asexual blood stage development cycle') +
-    theme_classic() +
-    theme(axis.text.x=element_text(angle = 0, 
-                                   vjust = 0.5,
-                                   size = 10))
-)
-
-if(global_save_var){
-  ggsave(filename = 'plots/ancestralPir_BarChart.pdf',
-          plot = p,
-          dpi = 300,
-          width = 15,
-          height = 10,
-          units = 'cm')
-}
-
-```
-
-# Calculate and cluster with z-scores {.tabset}
-
-z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:
-
-1. Convert the count/RPKM values of each gene into log values.
-2. Calculate the mean and standard deviation of X gene log values in each sample
-3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
-4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.
-
-* Need to make sure that the z-score is calculated within each transmission method!
-
-```{r zscores}
-tpm_to_zscore = function(input){
-
-  #If the input does not have a TPM element then assume that the input itself is the TPM object
-  if(any(names(input) %in% 'tpm')){
-    tpm_object <- input$tpm
-  }else{
-    tpm_object <- input
-    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
-  }
-
-  #Want all of them so will just ignore Geneid
-  input_logged <- tpm_object %>%
-    mutate_at(vars(-matches('Geneid')), 
-              log1p)
-
-  gene_mean_scores = input_logged %>% 
-    melt(id.vars = 'Geneid') %>% 
-    group_by(Geneid) %>% 
-    dplyr::summarise(mean = mean(value),
-                     std_dev = sd(value))
-
-  gene_vector = as.character(input_logged$Geneid)
-
-  input_z = lapply(gene_vector, function(gene){
-    mean_gene = filter(gene_mean_scores, 
-                       Geneid == paste(gene)) %>% 
-      select(mean)
-    std_gene = filter(gene_mean_scores, 
-                      Geneid == paste(gene)) %>% 
-      select(std_dev)
-    gene_scores = input_logged %>% 
-      filter(Geneid == paste(gene)) %>% 
-      select(-Geneid)
-    if(std_gene != 0){
-      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
-    }else{
-      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
-    }
-    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
-    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
-  })
-
-  #Use reduce and rbind to turn the list into one df
-  input_z_df = purrr::reduce(input_z, 
-                             rbind)
-  return(list(zscores = input_z_df,
-              tpm = tpm_object,
-              experiment = input$experiment))
-}
-
-#Remove schizont culture from this as we are interested in z-scores
-# across the 24h cycle and schizont will distort this and don't fit into comparison.
-chab_tpm_avg_forZ.df <- select(chab_tpm_avg.df, 
-                               -rMT_sch.culture)
-chab_tpm_avg_forZ.list <- lapply(c('rMT', 'SBP'), 
-                                 function(transmission){
-                                   #Subset for the Geneid column and either rMT or SBP
-                                   tpm.df <- chab_tpm_avg_forZ.df[
-                                     ,
-                                     grepl(colnames(chab_tpm_avg_forZ.df), 
-                                           pattern = paste0('Geneid|', transmission))
-                                   ]
-                                   list(tpm = tpm.df,
-                                        experiment = transmission)
-                                 })
-
-#Make and save z-scores
-chab_z.list <- lapply(chab_tpm_avg_forZ.list,
-                      tpm_to_zscore)
-
-###
-```
-
-# Phaseograms of all the genes.
-
-```{r row_order_phaseograms}
-
-#Getting the row order for each phaseogram for each experiment
-# Want the data ordered by the highest zscore for each gene in each stage.
-phaseogram_row_order <- lapply(chab_z.list, 
-                               function(list_elements){
-                                 zscore_df <- list_elements$zscores
-                                 zscore_ordered <- melt(zscore_df, 
-                                                        id = 'Geneid',
-                                                        variable.name = 'stage',
-                                                        value.name = 'zscore') %>% 
-                                   group_by(Geneid) %>% 
-                                   top_n(zscore, n = 1) %>% 
-                                   arrange(stage, 
-                                           desc(zscore)) %>%
-                                   distinct(Geneid, 
-                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-                                   filter(zscore > 0) #For some reason some remain with zero zscore
-                                 # zscore_ordered <- zscore_ordered[order(
-                                 #   match(
-                                 #     zscore_ordered$stage, 
-                                 #     order_vector
-                                 #   )
-                                 # ),]
-                                 return(c(list_elements, 
-                                          list(zscore_order = zscore_ordered)))
-                               })
-
-names(phaseogram_row_order) <- c('rMT', 'SBP')
-
-phaseogram_time_order <- lapply(names(phaseogram_row_order), 
-                                function(list_element){
-                                  transmission_list <- getElement(phaseogram_row_order,
-                                                                  list_element)
-                                  order_by_time <- transmission_list$zscore_order %>% 
-                                    mutate(stage = str_extract(stage,
-                                                               pattern = '[[:digit:]]{2}h')) %>% 
-                                    arrange(stage)
-                                  return(list(zscore_order = order_by_time))
-                                }
-)
-names(phaseogram_time_order) <- names(phaseogram_row_order)
-
-```
-
-## Supplementary figure 5 - Phaseograms of the entire genome.
-
-```{r phaseograms_whole_genome}
-if(global_save_var){
-  pdf(file = 'plots/phaseogram_allgenes_rMTorder.pdf',
-      height = 5,
-      width = 7)
-  draw_heatmap_zscore(zscore_list = chab_z.list,
-                      gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                      use_times = TRUE,
-                      use_order_vector = FALSE,
-                      show_row_names = FALSE,
-                      tpm_mat = TRUE)
-  dev.off()
-}
-
-if(global_save_var){
-  pdf(file = 'plots/phaseogram_allgenes_SBPorder.pdf',
-      height = 5,
-      width = 7)
-draw_heatmap_zscore(zscore_list = chab_z.list,
-                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = TRUE)
-  dev.off()
-}
-
-```
-
-##  Supplementary figure 6 - Phaseograms of the _cir_ genes with clade annotations
-
-```{r pir_gene_phaseogram_breaks}
-  max_tpm <- lapply(chab_z.list, 
-                    function(list_element){
-                      #list_element <- zscore_list[[1]]
-                      heatmap_data <- as.data.frame(list_element$zscores)
-                      title <- list_element$experiment
-
-                      if(!'Geneid' %in% colnames(heatmap_data)){
-                        stop('No Geneid column in the heatmap data.frame')
-                      }
-
-                            heatmap_data <- as.data.frame(
-                              filter(heatmap_data, 
-                                     Geneid %in% cir_id)
-                            )
-
-                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-                          # 
-                          # if(!is.null(gene_ordering_vector)){
-                          #   #Remove genes not in the main data frame but in the ordering vector, 
-                          #   # and then order based on the vector
-                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          #   heatmap_data <- heatmap_data[na.omit(
-                          #     match(gene_ordering_vector, 
-                          #           rownames(heatmap_data))
-                          #   ),,
-                          #   drop=FALSE]
-                          # }
-
-                          geneids <- rownames(heatmap_data)
-
-                          #Get max/min TPM scores for the side annotation heatmap                       
-                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                            list_element$tpm$Geneid),] %>% 
-                            select(-Geneid) %>% 
-                            as.matrix %>% 
-                            log1p
-                          rownames(tpm_mat) <- geneids
-
-                          tpm_min <- rowMins(tpm_mat)
-                          names(tpm_min) <- geneids
-                          tpm_max <- rowMaxs(tpm_mat)
-                          names(tpm_max) <- geneids
-                          return(list(tpm_max = tpm_max,
-                                      tpm_min = tpm_min))
-                        })
-
-  tpm_max_all <- mround(
-    max(
-      unlist(
-        lapply(max_tpm, 
-               getElement,
-               'tpm_max')
-      )
-    ), 
-    base = 1)
-
-  tpm_breaks <- round(
-    c(0,
-      tpm_max_all/4,
-      tpm_max_all/2,
-      tpm_max_all*(3/4),
-      tpm_max_all),
-    digits = 2
-  )
-```
-
-```{r phaseograms_cir}
-
-if(global_save_var){
- pdf(file = 'plots/phaseogram_pirgenes_rMTorder.pdf',
-      height = 5,
-      width = 5)
-  draw_heatmap_zscore(zscore_list = chab_z.list,
-                      gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                      filter_genes_vector = cir_id,
-                      annotation_clade = 'SorL',
-                      annotation_col = SorL_col,
-                      threshold_tpm = 1,
-                      use_times = TRUE,
-                      use_order_vector = FALSE,
-                      show_row_names = FALSE,
-                      tpm_mat = FALSE)
-  dev.off()
-}
-
-if(global_save_var){
-  pdf(file = 'plots/phaseogram_pirgenes_SBPorder.pdf',
-      height = 5,
-      width = 5)
-  draw_heatmap_zscore(zscore_list = chab_z.list,
-                      gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
-                      filter_genes_vector = cir_id,
-                      annotation_clade = 'SorL',
-                      annotation_col = SorL_col,
-                      threshold_tpm = 1,
-                      use_times = TRUE,
-                      use_order_vector = FALSE,
-                      show_row_names = FALSE,
-                      tpm_mat = TRUE)
-  dev.off()
-}
-
-```
-
-```{r zscore_save}
-chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
-                        chab_z.list[[2]]$zscores, 
-                        by = 'Geneid')
-write.csv(chab_z.df,
-          file = 'data/Pchabaudi_z.csv',
-          row.names = FALSE)
-write.table(x = phaseogram_time_order$rMT$zscore_order$Geneid,
-            file = 'data/Pchabaudi_z_phaseOrder_rMT.list',
-            row.names = FALSE, col.names = FALSE)
-write.table(x = phaseogram_time_order$SBP$zscore_order$Geneid,
-            file = 'data/Pchabaudi_z_phaseOrder_SBP.list',
-            row.names = FALSE, col.names = FALSE)
 ```
@@
 ```
-
-Noticeable enrichments of genes that are:
-* In-phase: DNA and nucleotide processes, and amino acid processing, such as DNA replication, tRNA aminoacylation and amino acid activation.
-* 1\*3h Out-of-phase: Translation and peptide biosynthesis, gene expression and ribosome biogenesis.
-* 2\*3h Out-of-phase: (Based on much fewer genes) Metabolism of carbohydrates and respiration.
-* 3\*3h Out-of-phase: (Based on even fewerer genes) More metabolism and lipid biosynthesis.
-* 4\*3h Out-of-phase: (Based on a few more genes actually) mRNA processing.
-
-Fore node visualisation:
-"Rectangles indicate the 5 most signicant terms. Rectangle color represents the relative signicance,
-ranging from dark red (most signicant) to bright yellow (least signicant). For each node, some basic information
-is displayed"
-
-## Figure 4 and supplementary figure 7
-
-_pir_ genes selected by being expressed above 1 TPM in all samples.
-
-Adding in housekeeping genes for comparison:
-HSP70 putative  PCHAS_0721000
-MSP1 PCHAS_0831300
-AMA1 PCHAS_0931000
-
-```{r pirs_side-by-side_figure4}
-
-housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
-                                 'MSP1' = 'PCHAS_0831300',
-                                 'AMA1' = 'PCHAS_0931000')
-
-cir_housekeeping_id <- c(cir_id, 
-                         'PCHAS_0721000',
-                         'PCHAS_0831300',
-                         'PCHAS_0931000')
-
-cir_plot <- twentyfour_rMT %>% 
-  filter(Geneid %in% cir_housekeeping_id) %>%
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(
-    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
-    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
-  ) %>% 
-  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
-  filter(!(time %in% c('sch.culture', 
-                       'Unknown'))) %>% 
-  mutate(subfam = str_replace(subfam, 
-                              'ancestral',
-                              'a')) %>% 
-  mutate(subfam =
-           forcats::fct_shift(
-             relevel(
-               as.factor(subfam),
-               ref = 'ref')
-           )
-  )
-
-#Need to add the transmission name to the subfam so that they are ordered correctly.
-# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
-# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.
-
-cir_plot$Geneid <- ifelse(
-  cir_plot$Geneid %in% housekeeping_genes[1,],
-  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
-  cir_plot$Geneid
-)
-
-limits_cir <- c(0, 
-                mround(max(cir_plot$tpm), 
-                       base = 100))
-
-cir_plot_order <- cir_plot %>% 
-  filter(time == '14h') %>% 
-  arrange(desc(tpm))
-
-cir_plot <- mutate(cir_plot, 
-                   Geneid = factor(Geneid, 
-                                   levels = unique(cir_plot_order$Geneid)))
-
-transmissionSep_cir_plot <- function(time_vec = '14h',
-                                     log_it = TRUE,
-                                     limits = limits_cir,
-                                     point_size = 1,
-                                     text_size = 10,
-                                     threshold_min = 100,
-                                     threshold_max = Inf){
-
-  cir_plot_filter <- group_by(cir_plot, 
-                              Geneid) %>% 
-    filter(
-      time == time_vec,
-      !Geneid %in% cir_id | any(tpm >= threshold_min),
-      !Geneid %in% cir_id | all(tpm <= threshold_max)
-    )
-  #Filter for pirs expressed above or below the thresholds, have the OR so that the ref genes aren't removed.
-  # e.g. Either 'Not a cir' OR 'Any TPM is above the minimum'.
-
-  # cir_plot_filter %>% group_by(Geneid) %>% summarise(n())
-  if(log_it){
-    cir_plot_filter$tpm <- log2(1+cir_plot_filter$tpm)
-    limits = log2(1+limits)
-    breaks <- round(seq.int(0, 
-                            max(limits), 
-                            length.out = round(max(limits)) + 1, 
-                            digits = 0), 
-                    digits = 0)
-  }else{
-    breaks <- seq.int(0, max(limits), length.out = 11)
-  }
-
-  cir_plot_filter$Geneid <- factor(
-    str_remove(cir_plot_filter$Geneid, 
-               pattern = 'PCHAS_'),
-    levels = str_remove(levels(cir_plot_filter$Geneid), 
-                        pattern = 'PCHAS_')
-  )
-
-  y <- ggplot(
-    data = cir_plot_filter,
-    mapping = aes(
-      x = Geneid,
-      y = tpm
-    )
-  ) +
-    stat_summary(fun.y = median, 
-                 fun.ymin = median, 
-                 fun.ymax = median,
-                 geom = "bar",
-                 fill = 'lightblue',
-                 width = 0.75, 
-                 size = 0.25, 
-                 position = 'dodge') +
-    geom_point(stat = 'identity', 
-               position = position_dodge(width = 0.75),
-               size = point_size) +
-    scale_y_continuous(breaks = breaks,
-                       limits = limits) +
-    facet_grid(.~subfam, 
-               scales = "free_x",
-               space = "free_x",
-               labeller = labeller(subfam = function(string) {
-                 str_remove_all(string, 
-                                pattern = '(RMT|SBP)_')
-               })) +
-    theme_classic() +
-    ylab('log(TPM + 1)')+
-    ggtitle(paste(unique(cir_plot_filter$time)))
-
-  if(grepl(time_vec, pattern = '20|23')) {
-    y <- y + theme(
-      axis.text.x = element_text(size = text_size,
-                                 angle = 90,
-                                 vjust = 0.5)
-    )
-  } else {
-    y <- y + theme(
-      axis.text.x = element_blank(),
-      axis.ticks.x = element_blank()
-    ) 
-  }
-
-           #This bit is to change the colour, and text colour, of the facet boxes
-
-           g <- ggplot_gtable(ggplot_build(y))
-
-           # g$widths[5] = 2*g$widths[5]
-
-           strip_t <- which(grepl('strip-t', g$layout$name))
-           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(RMT|SBP)_')
-           fills <- case_when(str_detect(subfam_vec, 
-                                         pattern = 'a') ~ clade_col['ancestral'],
-                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
-                              str_detect(subfam_vec, 
-                                         pattern = 'ref') ~ 'white')
-           k <- 1
-           for (i in strip_t) {
-             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-             k <- k+1
-           }
-
-           #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
-           # manually make them a bit wider.
-
-           short_panels <- which(str_detect(subfam_vec, pattern = 'a|S1'))
-
-           for(i in short_panels){
-             num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
-             g$widths[num] = 1.75*g$widths[num]
-             #1.75 times the width of the right panels.
-             # print(num)
-           }
-
-           if(global_save_var){
-             ggsave2(
-               paste0('Pchabaudi_side-by-side_barchart_facetbySubfam_',
-                      time_vec,
-                      '.png'),
-               plot = g,
-               dpi = 300,
-               width = 28,
-               height = 8,
-               units = 'cm'
-             )
-           }
-
-           grid.newpage()
-           grid.draw(g)
-
-}
-
-lapply(
-  sort(unique(cir_plot$time)),
-  transmissionSep_cir_plot,
-  threshold_min = 50,
-  threshold_max = Inf,
-  point_size = 1,
-  text_size = 7.5
-)
-```
-
-# Supp figure 9 - Pir expression of AAPLs and ChAPLs.
-
-```{r Supp_fig_AapL-ChaAPl_expression}
-
-cir_plot <- twentyfour_rMT %>% 
-  filter(Geneid %in% cir_id) %>%
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(
-    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
-    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
-    chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
-                                                       cir_info$Geneid)],
-                        levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
-                                                                            cir_info$Geneid)])))) %>% 
-  filter(!(time %in% c('sch', 
-                       'Unknown')),
-         !is.na(chapl_aapl) ) %>% 
-  group_by(Geneid) %>% 
-    arrange(desc(tpm))
-
-  cir_plot_order <- cir_plot %>% 
-    filter(time == '14h') %>% 
-    arrange(desc(tpm))
-
-  cir_plot <- mutate(cir_plot, 
-                     Geneid = factor(Geneid, 
-                                     levels = unique(cir_plot_order$Geneid)))
-  cir_plot_chapl_aapl_supp_info <- cir_plot
-
-  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))
-
-  breaks = seq(0, max_aapl_chapl, by = 1)
-    limits = c(0, max_aapl_chapl)
-    point_size = 0.25
-
-  lapply(
-    sort(unique(cir_plot$time)),
-    function(time_vec){
-
-    cir_plot <- filter(cir_plot, 
-                       time == time_vec)
-
-      cir_plot$tpm <- log2(1+cir_plot$tpm)
-
-    cir_plot$Geneid <- factor(
-      str_remove(cir_plot$Geneid, 
-                 pattern = 'PCHAS_'),
-      levels = str_remove(levels(cir_plot$Geneid), 
-                          pattern = 'PCHAS_')
-    )
-
-    y <- ggplot(
-      data = cir_plot,
-      mapping = aes(
-        x = Geneid,
-        y = tpm
-      )
-    ) +
-      stat_summary(fun.y = median, 
-                   fun.ymin = median, 
-                   fun.ymax = median,
-                   geom = "bar", 
-                   width = 0.75, 
-                   size = 0.25, 
-                   position = 'dodge',
-                   fill = 'lightblue') +
-      geom_point(stat = 'identity', 
-                 position = position_dodge(width = 0.75),
-                 size = point_size) +
-      scale_y_continuous(breaks = breaks,
-                         limits = limits) +
-      scale_fill_manual(values = c('skyblue','red')) +
-      ylab('log(TPM +1)') +
-      facet_grid(.~chapl_aapl, 
-                 scales = "free_x",
-                 space = "free_x") +
-      theme_classic() +
-      theme(
-        axis.text.x = element_text(angle = 90,
-                                   size = 5)
-      ) +
-      ggtitle(paste(unique(cir_plot$time)))
-
-    if(global_save_var){
-      ggsave2(
-        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
-               time_vec,
-               '.png'),
-        plot = y,
-        dpi = 300,
-        width = 25,
-        height = 10,
-        units = 'cm'
-      )
-    }
-
-    return(y)
-    })
-```

### 2022-07-27 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Start moving phase/GO enrichment code to its own Rmd file

@@
 
-```{r gene_sync}
-
-#Which genes are out of sync between rMT and SBP?
-
-#Melt the z-score dataframe into long form, then slice out the maximum z-score value.
-# Filter out any gene which has multiple max values for simplicity.
-# Summarise this with the numbers that each stage (the max stage) corresponds to.
-
-z_max <- lapply(chab_z.list, function(x) {
-  x$zscores %>% 
-  reshape2::melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'zscore') %>% 
-  group_by(Geneid) %>% 
-  slice_max(order_by = zscore, n = 1) %>% 
-  filter(n() == 1) %>% 
-  summarise(sample_max = times_stages_table$num[
-    str_detect(sample, 
-               times_stages_table$time)
-    ])
-})
-
-#Bind the two together with inner_join so that it only includes genes in both.
-# Then calculate the peak difference by subtracting the difference between the sample numbers.
-# Since this is a cycle, no point of the cycle is theoretically 
-#  more than 4 time points away from the next one. 
-#  e.g. 23h and 02h (8 and 1) are only 1 time point apart, not seven.
-#  Hence we subtract values above four from eight
-z_max_diff <- inner_join(z_max[[1]], z_max[[2]], by = 'Geneid') %>% 
-  mutate(diff = ifelse(abs(sample_max.x - sample_max.y) > 4,
-                       8 - abs(sample_max.x - sample_max.y),
-                       abs(sample_max.x - sample_max.y)))
-
-#Probably unnecessarily complicated code to make a barplot of how many genes 
-# are peaking at different timepoints between rMT and SBP
-barplot(group_by(z_max_diff, diff) %>% summarise(total = n()) %>% .$total,
-        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))
-
-#And for the pir genes
-barplot(group_by(z_max_diff, diff) %>% 
-          filter(Geneid %in% cir_id) %>% 
-          summarise(total = n()) %>% 
-          .$total,
-        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))
-
-```
-
-```{r phase_go_enrich}
-#Get the gene list for each group of genes in and out of phase.
-gene.list <- chabaudi_gene_info$Geneid
-z_max_diff.list <- lapply(seq(0,4), 
-                          function(diff_num) {
-                            diff.list <- filter(z_max_diff, diff == diff_num) %>% 
-                              .$Geneid
-                            #Get 1 for on the list and 0 for not.
-                            gene_diff.list <- as.integer(gene.list %in% diff.list)
-                            names(gene_diff.list) <- gene.list
-                            return(gene_diff.list)
-                          })
-names(z_max_diff.list) <- seq(0,4)
-
-#Read in the GO-Annotation file using GAFReader from FoldGO.
-# getAnnotation to obtain a list of the GO terms and associated genes.
-# inverseList (from topGO) to turn this into list of geneID with associated GO terms.
-gaf <- GAFReader("../standard_files/PlasmoDB-48_Pchabaudichabaudi_GO.gaf")
-go_annotation <- inverseList(getAnnotation(gaf))
-
-# How many genes for each group do not have any GO annotations?
-
-lapply(z_max_diff.list,
-       function(gene_diff.list){
-         total_noGO <- sum(!names(gene_diff.list[gene_diff.list == 1]) %in% 
-                             names(go_annotation))
-         prop_noGO <- total_noGO/sum(gene_diff.list == 1)
-         return(paste0(total_noGO, ' genes without annotation = ',
-                       round(prop_noGO*100, digits = 2), '%'))
-       })
-
-```
-
-```{r}
-#Load for the showSigOfNodes network function.
-library('Rgraphviz')
-
-#Make a topGOdata object 
-GOdata <- new("topGOdata", 
-              ontology = "BP", 
-              geneSel = function(x){x == 1},
-              allGenes = z_max_diff.list$`1`,
-              nodeSize = 5,
-              annot = annFUN.gene2GO, 
-              gene2GO = go_annotation)
-resultFis <- runTest(GOdata, 
-                     algorithm = "classic", 
-                     statistic = "fisher")
-plot.new()
-hist(score(resultFis), 50, xlab = "p-values")
-dev.off()
-showSigOfNodes(GOdata,
-               score(resultFis),
-               firstSigNodes = 5,
-               useInfo = 'all')
-goRes <- GenTable(GOdata, 
-         classic = resultFis, 
-         orderBy = "classic", 
-         ranksOf = "classic",
-         topNodes = 25) %>% 
-  filter(classic < 0.05)
-
-#Get first (most significant) enriched GO id 
-goID <- goRes[4, "GO.ID"]
-#Show distribution of GO term genes compared to null distribution (all other genes)
-print(showGroupDensity(GOdata, goID, ranks = TRUE))
-
-#Unlist the lapply below which goes through each element of the list and 
-# returns TRUE is the go_id is listed for that gene.
-#This makes a TRUE/FALSE index for the list/names of the list which can get 
-# the genes associated with that GO id.
-goID_genes <- names(go_annotation)[unlist(
-  lapply(go_annotation, 
-         function(go_ids){
-           any(goID %in% go_ids)
-         })
-)]
-
-draw_heatmap_zscore(zscore_list = chab_z.list,
-                    filter_genes_vector = names(z_max_diff.list$`1`[z_max_diff.list$`2` == 1]),
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = TRUE)
-```
-
-```{r}
-
-lapply(seq(0,4), 
-       function(phase){
-         GOdata <- new("topGOdata", 
-                       ontology = "BP", 
-                       geneSel = function(x){x == 1},
-                       allGenes = getElement(z_max_diff.list, paste(phase)),
-                       nodeSize = 5,
-                       annot = annFUN.gene2GO, 
-                       gene2GO = go_annotation)
-         resultFis <- runTest(GOdata, 
-                              algorithm = "classic", 
-                              statistic = "fisher")
-         draw_heatmap_zscore(zscore_list = chab_z.list,
-                             filter_genes_vector = names(
-                               getElement(z_max_diff.list, paste(phase))[
-                                 getElement(z_max_diff.list, paste(phase)) == 1
-                               ]
-                             ),
-                             gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                             threshold_tpm = 1,
-                             use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = TRUE)
-         GenTable(GOdata, 
-                  classic = resultFis, 
-                  orderBy = "classic", 
-                  ranksOf = "classic",
-                  topNodes = 25) %>% 
-           filter(classic < 0.05)
-       })
-
-```
-
 Noticeable enrichments of genes that are:
@@
 
-Fore node visualisation:
-"Rectangles indicate the 5 most signicant terms. Rectangle color represents the relative signicance,
-ranging from dark red (most signicant) to bright yellow (least signicant). For each node, some basic information
+For node visualisation:
+"Rectangles indicate the 5 most significant terms. Rectangle color represents the relative significance,
+ranging from dark red (most significant) to bright yellow (least significant). For each node, some basic information
 is displayed"

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_phaseGenes.Rmd
Make phaseograms of the most signif GO annotation genes and make graphs nicer

@@
 library("FoldGO")
+library('Rgraphviz')
 
@@
                                                          replacement = '\\.')
+#Add a number from 1 to 8 for each stage, useful for phase calculations
+times_stages_table$num <- 1:8
 ```
@@
 ```{r import_z}
-chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
-                        chab_z.list[[2]]$zscores, 
-                        by = 'Geneid')
 chab_z.df <- read.csv(file = 'data/Pchabaudi_z.csv')
 
-chab_z.list <- list('rMT' = list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
-                                                                             pattern = 'Geneid|rMT')]),
-                    'SBP' = list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
-                                                                             pattern = 'Geneid|SBP')]))
+#Make list object of the TPM values and the z-scores value by concatenating the 
+# lists of the zscore dataframes and the TPM lists from above.
+chab_z.list <- list('RMT' = c(chab_tpm_avg.list[['RMT']],
+                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
+                                                           pattern = 'Geneid|rMT')])),
+                    'SBP' = c(chab_tpm_avg.list[['SBP']],
+                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
+                                                           pattern = 'Geneid|SBP')])))
 
@@
 
-z_max <- lapply(chab_z.list, function(x) {
+z_max <- lapply(chab_z.list, 
+                function(x) {
   x$zscores %>% 
@@
 
-```{r phase_go_enrich}
+```{r z_genelist}
 #Get the gene list for each group of genes in and out of phase.
@@
                           })
-names(z_max_diff.list) <- seq(0,4)
+names(z_max_diff.list) <- paste0(seq(0,4)*3, 'h')
+
+```
+
+```{r phaseograms}
+lapply(paste0(seq(0,4)*3, 'h'), 
+       function(phase){
+         draw_heatmap_zscore(zscore_list = chab_z.list,
+                             filter_genes_vector = names(
+                               getElement(z_max_diff.list, paste(phase))[
+                                 getElement(z_max_diff.list, paste(phase)) == 1
+                               ]
+                             ),
+                             gene_ordering_vector = order_z_rMT,
+                             threshold_tpm = 1,
+                             use_times = TRUE,
+                             use_order_vector = FALSE,
+                             show_row_names = FALSE,
+                             tpm_mat = TRUE,
+                             column_title_name = paste0('Genes ',phase, ' out-of-phase'))
+         return()
+       })
+```
+
+How many of these genes do not have GO annotations?
+
+```{r phase_go_enrich}
 
@@
 #Load for the showSigOfNodes network function.
-library('Rgraphviz')
 
@@
 #Get first (most significant) enriched GO id 
-goID <- goRes[4, "GO.ID"]
+goID <- goRes[1, "GO.ID"]
 #Show distribution of GO term genes compared to null distribution (all other genes)
@@
 
-draw_heatmap_zscore(zscore_list = chab_z.list,
-                    filter_genes_vector = names(z_max_diff.list$`1`[z_max_diff.list$`2` == 1]),
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = TRUE)
 ```
@@
 
-lapply(seq(0,4), 
+GOres.list <- lapply(seq(0,4), 
                      function(phase){
@@
                                             statistic = "fisher")
-         draw_heatmap_zscore(zscore_list = chab_z.list,
-                             filter_genes_vector = names(
-                               getElement(z_max_diff.list, paste(phase))[
-                                 getElement(z_max_diff.list, paste(phase)) == 1
-                               ]
-                             ),
-                             gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                             threshold_tpm = 1,
-                             use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = TRUE)
-         GenTable(GOdata, 
+                       gentable <- GenTable(GOdata, 
                                               classic = resultFis, 
@@
                            filter(classic < 0.05)
+                       return(list(GOdata = GOdata,
+                                   resultFis = resultFis,
+                                   gentable = gentable))
+                     })
+names(GOres.list) <- paste0(seq(0,4)*3, 'h')
+lapply(GOres.list, getElement, 'gentable')
+
+```
+
+```{r topGO_phase}
+lapply(names(GOres.list), 
+       function(GOres.num){
+         GOres <- getElement(GOres.list, GOres.num)
+         #Get first (most significant) enriched GO id 
+         goID <- GOres$gentable[1, "GO.ID"]
+         goTerm <- GOres$gentable[1, "Term"]
+
+         #Show distribution of GO term genes compared to null distribution (all other genes)
+         print(showGroupDensity(GOres$GOdata, goID, ranks = TRUE))
+
+         #Can use genesInTerm function to extract the genes related to that GO term
+         genesinterm <- genesInTerm(GOres$GOdata, goID)
+         goID_genes <- unlist(genesinterm)
+         names(goID_genes) <- genesinterm
+
+         draw_heatmap_zscore(zscore_list = chab_z.list,
+                             filter_genes_vector = goID_genes,
+                             gene_ordering_vector = order_z_rMT,
+                             threshold_tpm = 1,
+                             use_times = TRUE,
+                             use_order_vector = FALSE,
+                             show_row_names = FALSE,
+                             tpm_mat = TRUE,
+                             column_title_name = paste0(GOres.num, ' - ', goTerm)
+         )
        })
@@
 ```
+
+Note that some enriched GO terms may not appear in the annotation file (.gaf) directly as they are found to be enriched through the network of related/ancestral terms from the GO network of relations.
+
+Also these phaseograms show every gene in entire dataset associated with the top GO term, so they may not all be as out-of-phase as one another.
+This is a good test to see if this GO enrichment seems correct - if most of the genes look like they're not in the phase expected then the GO enrichment is probably erroneous.

### 2022-07-28 rfunctions/draw_heatmap_zscore.R

@@ -10,6 +10,7 @@ draw_heatmap_zscore <- function(zscore_list,
                                 use_order_vector = TRUE,
                                 tpm_mat = TRUE,
                                 tpm_breaks = NULL,
+                                column_title_name = NULL,
                                 ...){
   if(is.null(tpm_breaks) & tpm_mat){
     #Need the generate the breaks in the data for the heatmap legend if not provided.
@@ -291,8 +292,9 @@ draw_heatmap_zscore <- function(zscore_list,
   
   draw(ht_list, 
        row_title_side = 'left',
-       column_title_side = 'bottom',
+       column_title_side = 'top',
        heatmap_legend_side = 'left',
+       column_title = column_title_name,
        #Only include the annotation legends for the annotations we actually have.
        annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
          list(tpm_lgd,

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_phaseGenes.Rmd
Clean-up to remove stuff from practising wiht topGO

@@
 barplot(group_by(z_max_diff, diff) %>% summarise(total = n()) %>% .$total,
-        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))
+        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'),
+        main = 'all genes out-of-phase')
 
@@
           .$total,
-        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))
+        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'),
+        main = 'pir genes out-of-phase')
 
@@
 
-```{r}
-#Load for the showSigOfNodes network function.
-
-#Make a topGOdata object 
-GOdata <- new("topGOdata", 
-              ontology = "BP", 
-              geneSel = function(x){x == 1},
-              allGenes = z_max_diff.list$`1`,
-              nodeSize = 5,
-              annot = annFUN.gene2GO, 
-              gene2GO = go_annotation)
-resultFis <- runTest(GOdata, 
-                     algorithm = "classic", 
-                     statistic = "fisher")
-plot.new()
-hist(score(resultFis), 50, xlab = "p-values")
-dev.off()
-showSigOfNodes(GOdata,
-               score(resultFis),
-               firstSigNodes = 5,
-               useInfo = 'all')
-goRes <- GenTable(GOdata, 
-         classic = resultFis, 
-         orderBy = "classic", 
-         ranksOf = "classic",
-         topNodes = 25) %>% 
-  filter(classic < 0.05)
-
-#Get first (most significant) enriched GO id 
-goID <- goRes[1, "GO.ID"]
-#Show distribution of GO term genes compared to null distribution (all other genes)
-print(showGroupDensity(GOdata, goID, ranks = TRUE))
-
-#Unlist the lapply below which goes through each element of the list and 
-# returns TRUE is the go_id is listed for that gene.
-#This makes a TRUE/FALSE index for the list/names of the list which can get 
-# the genes associated with that GO id.
-goID_genes <- names(go_annotation)[unlist(
-  lapply(go_annotation, 
-         function(go_ids){
-           any(goID %in% go_ids)
-         })
-)]
-
-```
-
-```{r}
-
-GOres.list <- lapply(seq(0,4), 
+```{r go-list}
+# Use the topGO functions to make a topGOdata object, 
+# conduct a Fisher test for statistical significance, 
+# then create a table of these signifcant GO terms.
+GOres.list <- lapply(paste0(seq(0,4)*3, 'h'), 
                      function(phase){

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_phaseGenes.Rmd
Use which_go variable to assign number to which highest enriched GO termn we want for phaseograms

@@
                        GOdata <- new("topGOdata", 
-                                     ontology = "BP", 
+                                     ontology = "CC", 
                                      geneSel = function(x){x == 1},
@@
          #Get first (most significant) enriched GO id 
-         goID <- GOres$gentable[1, "GO.ID"]
-         goTerm <- GOres$gentable[1, "Term"]
+         which_go <- 2
+         goID <- GOres$gentable[which_go, "GO.ID"]
+         goTerm <- GOres$gentable[which_go, "Term"]
 
          #Show distribution of GO term genes compared to null distribution (all other genes)
-         print(showGroupDensity(GOres$GOdata, goID, ranks = TRUE))
+         # print(showGroupDensity(GOres$GOdata, goID, ranks = TRUE))
 
@@
          )
+         return()
        })

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_phaseGenes.Rmd
Add notes, BP and CC topGO sections, and move notes from pirAnalysis to phaseGenes

@@
 ---
+
+Looking at GO enrichment of genes in- and out-of-phase between the 24 asexual life cycles of rMT and SBP transmitted P. chabaudi.
 
@@
 
-The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.
-
 ```{r cir_id, message=FALSE, results='hide'}
-
+# The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.
 # Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million
@@
 
+Working out whether a given gene is out-of-sync based on when it peaks in transcription.
+If it peaks at different timepoints in rMT compared to SBP, then the time difference between these points is considered the gene's synchronicity.
+0h refers to a gene that's transcriped maximally at the same time in rMT and SBP infections.
+
+Note that this isn't directional; 6h out-of-sync could be 6h before or after.
+
+# Number of genes out-of-phase.
+
 ```{r gene_sync}
@@
 
-How many of these genes do not have GO annotations?
+The out-of-phase values represent the data reasonably well, although it is notable that for some genes only 3 or 6h out-of-sync the peak of transcription is quite broad and overlaps.
+
+# How many of these genes do not have GO annotations?
 
@@
 ```
+
+# GO term enrichment.
+
+Perform a Fishers test using values of 1 for genes of out-of-sync group being tested, and 0 for all others.
+
+## Cellular Component.
 
@@
 
+Phaseograms of the genes represented by the most significant GO term returned for each out-of-sync group.
+
+These phaseograms show every gene in the entire dataset associated with the 1st or 2nd most statistically signif enriched GO term, so they may not all be as out-of-phase as one another.
+It is likely that only a subset of GO term genes will be out-of-sync by the same number of timepoints but this can be sufficient still for the GO term to be enriched.
+See the tables above for the number of genes in that GO term group ('Annotated') and the number that are in that out-of-sync group ('Significant').
+
+This is a good test to see if this GO enrichment seems correct - if most of the genes look like they're not in the phase expected then the GO enrichment is probably erroneous.
+
 ```{r topGO_phase}
+
 lapply(names(GOres.list), 
@@
          GOres <- getElement(GOres.list, GOres.num)
-         #Get first (most significant) enriched GO id 
-         which_go <- 2
+         lapply(c(1,2), 
+                function(which_go){
+                  #Get first or second (most significant) enriched GO id.
                   goID <- GOres$gentable[which_go, "GO.ID"]
@@
                              tpm_mat = TRUE,
-                             column_title_name = paste0(GOres.num, ' - ', goTerm)
+                             column_title_name = paste0(GOres.num, ' - ',
+                                                        which_go, ' most sig',
+                                                        ' - ', goTerm)
          )
+         return()
+                })
          return()
@@
 
-Also these phaseograms show every gene in entire dataset associated with the top GO term, so they may not all be as out-of-phase as one another.
-This is a good test to see if this GO enrichment seems correct - if most of the genes look like they're not in the phase expected then the GO enrichment is probably erroneous.
+## Biological Process
+
+```{r go-list}
+# Use the topGO functions to make a topGOdata object, 
+# conduct a Fisher test for statistical significance, 
+# then create a table of these signifcant GO terms.
+GOres.bp.list <- lapply(paste0(seq(0,4)*3, 'h'), 
+                     function(phase){
+                       GOdata <- new("topGOdata", 
+                                     ontology = "BP", 
+                                     geneSel = function(x){x == 1},
+                                     allGenes = getElement(z_max_diff.list, paste(phase)),
+                                     nodeSize = 5,
+                                     annot = annFUN.gene2GO, 
+                                     gene2GO = go_annotation)
+                       resultFis <- runTest(GOdata, 
+                                            algorithm = "classic", 
+                                            statistic = "fisher")
+                       gentable <- GenTable(GOdata, 
+                                              classic = resultFis, 
+                                              orderBy = "classic", 
+                                              ranksOf = "classic",
+                                              topNodes = 25) %>% 
+                           filter(classic < 0.05)
+                       return(list(GOdata = GOdata,
+                                   resultFis = resultFis,
+                                   gentable = gentable))
+                     })
+names(GOres.bp.list) <- paste0(seq(0,4)*3, 'h')
+lapply(GOres.bp.list, getElement, 'gentable')
+
+```
+
+```{r topGO_phase}
+
+lapply(names(GOres.bp.list), 
+       function(GOres.num){
+         GOres <- getElement(GOres.bp.list, GOres.num)
+         lapply(c(1,2), 
+                function(which_go){
+                  #Get first or second (most significant) enriched GO id.
+                  goID <- GOres$gentable[which_go, "GO.ID"]
+                  goTerm <- GOres$gentable[which_go, "Term"]
+
+                  #Show distribution of GO term genes compared to null distribution (all other genes)
+                  # print(showGroupDensity(GOres$GOdata, goID, ranks = TRUE))
+
+                  #Can use genesInTerm function to extract the genes related to that GO term
+                  genesinterm <- genesInTerm(GOres$GOdata, goID)
+                  goID_genes <- unlist(genesinterm)
+                  names(goID_genes) <- genesinterm
+
+                  draw_heatmap_zscore(zscore_list = chab_z.list,
+                                      filter_genes_vector = goID_genes,
+                             gene_ordering_vector = order_z_rMT,
+                             threshold_tpm = 1,
+                             use_times = TRUE,
+                             use_order_vector = FALSE,
+                             show_row_names = FALSE,
+                             tpm_mat = TRUE,
+                             column_title_name = paste0(GOres.num, ' - ',
+                                                        which_go, ' most sig',
+                                                        ' - ', goTerm)
+         )
+         return()
+                })
+         return()
+       })
+
+```
+
+Noticeable enrichments of genes that are:
+* In-phase: DNA and nucleotide processes, and amino acid processing, such as DNA replication, tRNA aminoacylation and amino acid activation.
+* 1\*3h Out-of-phase: Translation and peptide biosynthesis, gene expression and ribosome biogenesis.
+* 2\*3h Out-of-phase: (Based on much fewer genes) Metabolism of carbohydrates and respiration.
+* 3\*3h Out-of-phase: (Based on even fewerer genes) More metabolism and lipid biosynthesis.
+* 4\*3h Out-of-phase: (Based on a few more genes actually) mRNA processing.

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Add notes, BP and CC topGO sections, and move notes from pirAnalysis to phaseGenes

@@
 
-Noticeable enrichments of genes that are:
-* In-phase: DNA and nucleotide processes, and amino acid processing, such as DNA replication, tRNA aminoacylation and amino acid activation.
-* 1\*3h Out-of-phase: Translation and peptide biosynthesis, gene expression and ribosome biogenesis.
-* 2\*3h Out-of-phase: (Based on much fewer genes) Metabolism of carbohydrates and respiration.
-* 3\*3h Out-of-phase: (Based on even fewerer genes) More metabolism and lipid biosynthesis.
-* 4\*3h Out-of-phase: (Based on a few more genes actually) mRNA processing.
-
-For node visualisation:
-"Rectangles indicate the 5 most significant terms. Rectangle color represents the relative significance,
-ranging from dark red (most significant) to bright yellow (least significant). For each node, some basic information
-is displayed"
-
 ## Figure 4 and supplementary figure 7

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_phaseGenes.Rmd
Add all pir bar plot

@@
 
-```{r go-list}
+```{r go-list-bp}
 # Use the topGO functions to make a topGOdata object, 
@@
 
-```{r topGO_phase}
+```{r topGO_phase-bp}
 
@@
 Noticeable enrichments of genes that are:
+
 * In-phase: DNA and nucleotide processes, and amino acid processing, such as DNA replication, tRNA aminoacylation and amino acid activation.
-* 1\*3h Out-of-phase: Translation and peptide biosynthesis, gene expression and ribosome biogenesis.
-* 2\*3h Out-of-phase: (Based on much fewer genes) Metabolism of carbohydrates and respiration.
-* 3\*3h Out-of-phase: (Based on even fewerer genes) More metabolism and lipid biosynthesis.
-* 4\*3h Out-of-phase: (Based on a few more genes actually) mRNA processing.
+* 3h Out-of-phase: Translation and peptide biosynthesis, gene expression and ribosome biogenesis.
+* 6h Out-of-phase: (Based on much fewer genes) Metabolism of carbohydrates and respiration.
+* 9h Out-of-phase: (Based on even fewerer genes - phaseogram not quite convincing) More metabolism and lipid biosynthesis.
+* 12h Out-of-phase: (Based on a few more genes actually - phaseogram not quite convincing) mRNA processing.

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Add all pir bar plot

@@
 
+```{r all_pir_barchart}
+  cir_plot <- chab_tpm.df %>% 
+    filter(Geneid %in% cir_id) %>% 
+    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
+    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
+    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
+           transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
+    group_by(sample, time, transmission) %>% 
+    dplyr::summarise(sum_tpm = sum(tpm))
+
+y_scale <- seq(0, 
+               signif(ceiling(max(cir_plot$sum_tpm)), 
+                      digits = 2), 
+               length.out = 5)
+
+(
+  p <- ggplot(data = cir_plot, 
+              aes(x = time, 
+                  y = sum_tpm,
+                  fill = transmission)) +
+    stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
+                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
+    scale_y_continuous(breaks = y_scale) +
+    scale_fill_manual(values = cbPalette) +
+    geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
+    ylab(as.expression(bquote('Total '~italic(.('pir'))~'TPM'))) +
+    xlab('Asexual blood stage development cycle') +
+    theme_classic() +
+    theme(axis.text.x=element_text(angle = 0, 
+                                   vjust = 0.5,
+                                   size = 10))
+)
+
+if(global_save_var){
+  ggsave(filename = 'plots/allPir_BarChart.pdf',
+          plot = p,
+          dpi = 300,
+          width = 15,
+          height = 10,
+          units = 'cm')
+}
+
+```
+
 # Calculate and cluster with z-scores {.tabset}

### 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Changed total TPM of pir bar chart to a stacked chart with sub-families shown

--- 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
@@
 
+```{r all_pir_barchart}
+cir_plot <- chab_tpm.df %>% 
+  filter(Geneid %in% cir_id) %>% 
+  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
+  mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
+         transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1],
+         SorL = cir_clade_info$SorL[match(Geneid, cir_clade_info$Geneid)]) %>% 
+  group_by(Geneid, time, transmission, SorL) %>% 
+  dplyr::summarise(tpm = median(tpm)) %>% 
+  arrange(SorL, tpm)
+
+y_scale <- seq(0, 
+               signif(ceiling(
+                 max(
+                   group_by(cir_plot, time, transmission) %>%
+                     dplyr::summarise(sum_tpm = sum(tpm)) %>%  .$sum_tpm
+                   )
+               ), 
+               digits = 2), 
+               length.out = 5)
+
+(
+  p <- ggplot(data = cir_plot, 
+              aes(x = factor(time), 
+                  y = tpm,
+                  fill = factor(SorL))) +
+    geom_col(colour = 'black',
+             size = 0.1,
+             position = 'stack') +
+    scale_y_continuous(breaks = y_scale) +
+    scale_fill_manual(values = SorL_col) +
+    ylab(as.expression(bquote('Total '~italic(.('pir'))~'TPM'))) +
+    xlab('Asexual blood stage development cycle') +
+    theme_classic() +
+    theme(axis.text.x=element_text(angle = 90, 
+                                   vjust = 0.5,
+                                   size = 7)) +
+    facet_grid(~transmission, scales = 'free')
+)
+
+if(global_save_var){
+  ggsave(filename = 'plots/allPir_BarChart.pdf',
+          plot = p,
+          dpi = 300,
+          width = 20,
+          height = 10,
+          units = 'cm')
+}
+
+```
+
 # Calculate and cluster with z-scores {.tabset}
@@
 #Make and save z-scores
-chab_z.list <- lapply(chab_tpm_avg_forZ.list,
-                      tpm_to_zscore)
+# chab_z.list <- lapply(chab_tpm_avg_forZ.list,
+#                       tpm_to_zscore)
 
@@
 ```{r row_order_phaseograms}
-
-#Getting the row order for each phaseogram for each experiment
-# Want the data ordered by the highest zscore for each gene in each stage.
-phaseogram_row_order <- lapply(chab_z.list, 
-                               function(list_elements){
-                                 zscore_df <- list_elements$zscores
-                                 zscore_ordered <- melt(zscore_df, 
-                                                        id = 'Geneid',
-                                                        variable.name = 'stage',
-                                                        value.name = 'zscore') %>% 
-                                   group_by(Geneid) %>% 
-                                   top_n(zscore, n = 1) %>% 
-                                   arrange(stage, 
-                                           desc(zscore)) %>%
-                                   distinct(Geneid, 
-                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-                                   filter(zscore > 0) #For some reason some remain with zero zscore
-                                 # zscore_ordered <- zscore_ordered[order(
-                                 #   match(
-                                 #     zscore_ordered$stage, 
-                                 #     order_vector
+# 
+# #Getting the row order for each phaseogram for each experiment
+# # Want the data ordered by the highest zscore for each gene in each stage.
+# phaseogram_row_order <- lapply(chab_z.list, 
+#                                function(list_elements){
+#                                  zscore_df <- list_elements$zscores
+#                                  zscore_ordered <- melt(zscore_df, 
+#                                                         id = 'Geneid',
+#                                                         variable.name = 'stage',
+#                                                         value.name = 'zscore') %>% 
+#                                    group_by(Geneid) %>% 
+#                                    top_n(zscore, n = 1) %>% 
+#                                    arrange(stage, 
+#                                            desc(zscore)) %>%
+#                                    distinct(Geneid, 
+#                                             .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
+#                                    filter(zscore > 0) #For some reason some remain with zero zscore
+#                                  # zscore_ordered <- zscore_ordered[order(
+#                                  #   match(
+#                                  #     zscore_ordered$stage, 
+#                                  #     order_vector
+#                                  #   )
+#                                  # ),]
+#                                  return(c(list_elements, 
+#                                           list(zscore_order = zscore_ordered)))
+#                                })
+# 
+# names(phaseogram_row_order) <- c('rMT', 'SBP')
+# 
+# phaseogram_time_order <- lapply(names(phaseogram_row_order), 
+#                                 function(list_element){
+#                                   transmission_list <- getElement(phaseogram_row_order,
+#                                                                   list_element)
+#                                   order_by_time <- transmission_list$zscore_order %>% 
+#                                     mutate(stage = str_extract(stage,
+#                                                                pattern = '[[:digit:]]{2}h')) %>% 
+#                                     arrange(stage)
+#                                   return(list(zscore_order = order_by_time))
+#                                 }
 # )
-                                 # ),]
-                                 return(c(list_elements, 
-                                          list(zscore_order = zscore_ordered)))
-                               })
-
-names(phaseogram_row_order) <- c('rMT', 'SBP')
-
-phaseogram_time_order <- lapply(names(phaseogram_row_order), 
-                                function(list_element){
-                                  transmission_list <- getElement(phaseogram_row_order,
-                                                                  list_element)
-                                  order_by_time <- transmission_list$zscore_order %>% 
-                                    mutate(stage = str_extract(stage,
-                                                               pattern = '[[:digit:]]{2}h')) %>% 
-                                    arrange(stage)
-                                  return(list(zscore_order = order_by_time))
-                                }
-)
-names(phaseogram_time_order) <- names(phaseogram_row_order)
+# names(phaseogram_time_order) <- names(phaseogram_row_order)
+```
+
+```{r read_z}
+#Can read in the z scores to avoid re-doing all the above.
+chab_z.df <- read.csv(file = 'data/Pchabaudi_z.csv')
+#Make list object of the TPM values and the z-scores value by concatenating the 
+# lists of the zscore dataframes and the TPM lists from above.
+chab_z.list <- list('RMT' = c(chab_tpm_avg.list[['RMT']],
+                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
+                                                           pattern = 'Geneid|rMT')])),
+                    'SBP' = c(chab_tpm_avg.list[['SBP']],
+                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
+                                                           pattern = 'Geneid|SBP')])))
+phaseogram_time_order <- list(rMT = list(zscore_order = list()),
+                              SBP = list(zscore_order = list()))
+
+phaseogram_time_order$rMT$zscore_order$Geneid <- read.table(file = 'data/Pchabaudi_z_phaseOrder_rMT.list')[,1]
+phaseogram_time_order$SBP$zscore_order$Geneid <- read.table(file = 'data/Pchabaudi_z_phaseOrder_SBP.list')[,1]
 
@@
       height = 5,
-      width = 5)
+      width = 6)
   draw_heatmap_zscore(zscore_list = chab_z.list,
@@
       height = 5,
-      width = 5)
+      width = 6)
   draw_heatmap_zscore(zscore_list = chab_z.list,
                       gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
+                      # gene_ordering_vector = arrange(cir_clade_info, SorL)$Geneid,
                       filter_genes_vector = cir_id,
@@
                       annotation_col = SorL_col,
-                      threshold_tpm = 1,
+                      threshold_tpm = 0,
                       use_times = TRUE,
@@
 ```{r zscore_save}
-chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
-                        chab_z.list[[2]]$zscores, 
-                        by = 'Geneid')
-write.csv(chab_z.df,
-          file = 'data/Pchabaudi_z.csv',
-          row.names = FALSE)
-write.table(x = phaseogram_time_order$rMT$zscore_order$Geneid,
-            file = 'data/Pchabaudi_z_phaseOrder_rMT.list',
-            row.names = FALSE, col.names = FALSE)
-write.table(x = phaseogram_time_order$SBP$zscore_order$Geneid,
-            file = 'data/Pchabaudi_z_phaseOrder_SBP.list',
-            row.names = FALSE, col.names = FALSE)
+# chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
+#                         chab_z.list[[2]]$zscores, 
+#                         by = 'Geneid')
+# write.csv(chab_z.df,
+#           file = 'data/Pchabaudi_z.csv',
+#           row.names = FALSE)
+# write.table(x = phaseogram_time_order$rMT$zscore_order$Geneid,
+#             file = 'data/Pchabaudi_z_phaseOrder_rMT.list',
+#             row.names = FALSE, col.names = FALSE)
+# write.table(x = phaseogram_time_order$SBP$zscore_order$Geneid,
+#             file = 'data/Pchabaudi_z_phaseOrder_SBP.list',
+#             row.names = FALSE, col.names = FALSE)
 ```

### 2022-07-29 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Improved all pir stacked bar chart plot, accidentally may have altered the ancestral bar chart

@@
                                    vjust = 0.5,
-                                   size = 10))
+                                   size = 12))
 )
@@
     theme_classic() +
+    guides(fill = guide_legend(title = 'Subfamily')) +
     theme(axis.text.x=element_text(angle = 90, 
                                    vjust = 0.5,
-                                   size = 7)) +
+                                   size = 10),
+          axis.text.y = element_text(size = 12),
+          axis.title.y = element_text(size = 15),
+          axis.title.x = element_text(size = 15),
+          strip.text = element_text(size = 15)) +
     facet_grid(~transmission, scales = 'free')
@@
 if(global_save_var){
-  ggsave(filename = 'plots/allPir_BarChart.pdf',
+  ggsave(filename = 'plots/allPir_stackedBarChart.pdf',
           plot = p,
@@
           width = 20,
-          height = 10,
+          height = 20,
           units = 'cm')

### 2022-07-29 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Add pca plot of just pirs

@@
 
+```{r pca_pir}
+rownames_geneid <- chab_tpm.df$Geneid[chab_tpm.df$Geneid %in% cir_id]
+chab_tpm.df_forpca <- data.frame(select(filter(chab_tpm.df, Geneid %in% cir_id), -Geneid))
+rownames(chab_tpm.df_forpca) <- rownames_geneid
+chab_tpm.df_pca <- prcomp(chab_tpm.df_forpca, 
+                              center = TRUE, 
+                              scale. = TRUE)
+
+chab_tpm.df_pca_var <- get_pca_var(chab_tpm.df_pca)
+chab_tpm.df_pca_coords <- data.frame(chab_tpm.df_pca_var$coord)
+
+chab_tpm.df_pca_coords <- data.frame(sample = rownames(chab_tpm.df_pca_var$coord),
+                                        PC1 = chab_tpm.df_pca_coords$Dim.1,
+                                        PC3 = chab_tpm.df_pca_coords$Dim.3,
+                                        PC2 = chab_tpm.df_pca_coords$Dim.2) %>% 
+  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
+  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
+          time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture'))
+
+colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
+colornames <- setNames(colornames, 
+                       sort(unique(chab_tpm.df_pca_coords$time)))
+
+pca_importance <- as.data.frame(summary(chab_tpm.df_pca)$importance)
+pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
+pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
+pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
+
+(
+  plot <- ggplot(chab_tpm.df_pca_coords,
+                 aes(
+                   x = PC1,
+                   y = PC2,
+                   col = time,
+                   shape = transmission
+                 )) +
+    geom_point(size = 3,
+               alpha = 0.8) +
+    scale_color_manual(values = colornames) +
+    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
+    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
+    guides(shape=guide_legend(ncol=2),
+           col = guide_legend(ncol=2)) +
+    theme_classic() +
+    theme(legend.key.size = unit(4, 'mm'),
+          legend.text = element_text(size = 7)) +
+    ggtitle('PCA of pir genes')
+)
+
+```
+
 # Bar chart of ancestral _pir_ expression.
@@
     theme_classic() +
-    theme(axis.text.x=element_text(angle = 0, 
+    theme(axis.text.x=element_text(angle = 90, 
                                    vjust = 0.5,

### 2022-07-29 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_boilerplate.Rmd
Create boilerplate Rmd for easier time making future Rmd scripts

--- 2022-07-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_phaseGenes.Rmd
@@
 ---
-title: "Manuscript_Github_Pchabaudi_phaseGene"
+title: "Manuscript_Github_Pchabaudi_boilerplate"
 author: "Timothy Little"
-date: "27/07/2022"
+date: "29/07/2022"
 output: html_document
 ---
+
+Looking at GO enrichment of genes in- and out-of-phase between the 24 asexual life cycles of rMT and SBP transmitted P. chabaudi.
 
@@
 library("MatrixGenerics") # for rowMins and rowMax
-library("topGO")
-library("FoldGO")
-library('Rgraphviz')
 
@@
 ```
-
-```{r import_z}
-chab_z.df <- read.csv(file = 'data/Pchabaudi_z.csv')
-
-#Make list object of the TPM values and the z-scores value by concatenating the 
-# lists of the zscore dataframes and the TPM lists from above.
-chab_z.list <- list('RMT' = c(chab_tpm_avg.list[['RMT']],
-                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
-                                                           pattern = 'Geneid|rMT')])),
-                    'SBP' = c(chab_tpm_avg.list[['SBP']],
-                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
-                                                           pattern = 'Geneid|SBP')])))
-
-order_z_rMT <- read.table(file = 'data/Pchabaudi_z_phaseOrder_rMT.list')[,1]
-order_z_SBP <- read.table(file = 'data/Pchabaudi_z_phaseOrder_SBP.list')[,1]
-
-```
-
-```{r colours_clades_phase}
-SorL_col <- setNames(
-  c(    '#F0E442',
-        '#56B4E9', #L
-        '#E69F00', #S
-        '#999999'),
-  sort(
-    unique(cir_info$SorL)
-  )
-)
-
-SorL_col_num <- SorL_col
-names(SorL_col_num) <- sapply(names(SorL_col), 
-                              function(clade_name) {
-                                paste0(
-                                  clade_name, 
-                                  ' (n = ',
-                                  sum(cir_info$SorL == clade_name),
-                                  ')'
-                                )
-                              })
-
-clade_col <- setNames(
-  c('#F0E442',
-    '#56B4E9', #L1
-    '#009E73',
-    '#0072B2',
-    '#E69F00', #S1
-    '#D55E00',
-    '#CC79A7',
-    '#999999'),
-  sort(
-    unique(cir_info$subfam)
-  )
-)
-
-clade_col_num <- clade_col
-names(clade_col_num) <- sapply(names(clade_col), 
-                               function(clade_name) {
-                                 paste0(
-                                   clade_name, 
-                                   ' (n = ',
-                                   sum(cir_info$subfam == clade_name),
-                                   ')'
-                                 )
-                               })
-
-```
-
-```{r gene_sync}
-
-#Which genes are out of sync between rMT and SBP?
-
-#Melt the z-score dataframe into long form, then slice out the maximum z-score value.
-# Filter out any gene which has multiple max values for simplicity.
-# Summarise this with the numbers that each stage (the max stage) corresponds to.
-
-z_max <- lapply(chab_z.list, 
-                function(x) {
-  x$zscores %>% 
-  reshape2::melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'zscore') %>% 
-  group_by(Geneid) %>% 
-  slice_max(order_by = zscore, n = 1) %>% 
-  filter(n() == 1) %>% 
-  summarise(sample_max = times_stages_table$num[
-    str_detect(sample, 
-               times_stages_table$time)
-    ])
-})
-
-#Bind the two together with inner_join so that it only includes genes in both.
-# Then calculate the peak difference by subtracting the difference between the sample numbers.
-# Since this is a cycle, no point of the cycle is theoretically 
-#  more than 4 time points away from the next one. 
-#  e.g. 23h and 02h (8 and 1) are only 1 time point apart, not seven.
-#  Hence we subtract values above four from eight
-z_max_diff <- inner_join(z_max[[1]], z_max[[2]], by = 'Geneid') %>% 
-  mutate(diff = ifelse(abs(sample_max.x - sample_max.y) > 4,
-                       8 - abs(sample_max.x - sample_max.y),
-                       abs(sample_max.x - sample_max.y)))
-
-#Probably unnecessarily complicated code to make a barplot of how many genes 
-# are peaking at different timepoints between rMT and SBP
-barplot(group_by(z_max_diff, diff) %>% summarise(total = n()) %>% .$total,
-        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'),
-        main = 'all genes out-of-phase')
-
-#And for the pir genes
-barplot(group_by(z_max_diff, diff) %>% 
-          filter(Geneid %in% cir_id) %>% 
-          summarise(total = n()) %>% 
-          .$total,
-        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'),
-        main = 'pir genes out-of-phase')
-
-```
-
-```{r z_genelist}
-#Get the gene list for each group of genes in and out of phase.
-gene.list <- chabaudi_gene_info$Geneid
-z_max_diff.list <- lapply(seq(0,4), 
-                          function(diff_num) {
-                            diff.list <- filter(z_max_diff, diff == diff_num) %>% 
-                              .$Geneid
-                            #Get 1 for on the list and 0 for not.
-                            gene_diff.list <- as.integer(gene.list %in% diff.list)
-                            names(gene_diff.list) <- gene.list
-                            return(gene_diff.list)
-                          })
-names(z_max_diff.list) <- paste0(seq(0,4)*3, 'h')
-
-```
-
-```{r phaseograms}
-lapply(paste0(seq(0,4)*3, 'h'), 
-       function(phase){
-         draw_heatmap_zscore(zscore_list = chab_z.list,
-                             filter_genes_vector = names(
-                               getElement(z_max_diff.list, paste(phase))[
-                                 getElement(z_max_diff.list, paste(phase)) == 1
-                               ]
-                             ),
-                             gene_ordering_vector = order_z_rMT,
-                             threshold_tpm = 1,
-                             use_times = TRUE,
-                             use_order_vector = FALSE,
-                             show_row_names = FALSE,
-                             tpm_mat = TRUE,
-                             column_title_name = paste0('Genes ',phase, ' out-of-phase'))
-         return()
-       })
-```
-
-How many of these genes do not have GO annotations?
-
-```{r phase_go_enrich}
-
-#Read in the GO-Annotation file using GAFReader from FoldGO.
-# getAnnotation to obtain a list of the GO terms and associated genes.
-# inverseList (from topGO) to turn this into list of geneID with associated GO terms.
-gaf <- GAFReader("../standard_files/PlasmoDB-48_Pchabaudichabaudi_GO.gaf")
-go_annotation <- inverseList(getAnnotation(gaf))
-
-# How many genes for each group do not have any GO annotations?
-
-lapply(z_max_diff.list,
-       function(gene_diff.list){
-         total_noGO <- sum(!names(gene_diff.list[gene_diff.list == 1]) %in% 
-                             names(go_annotation))
-         prop_noGO <- total_noGO/sum(gene_diff.list == 1)
-         return(paste0(total_noGO, ' genes without annotation = ',
-                       round(prop_noGO*100, digits = 2), '%'))
-       })
-
-```
-
-```{r go-list}
-# Use the topGO functions to make a topGOdata object, 
-# conduct a Fisher test for statistical significance, 
-# then create a table of these signifcant GO terms.
-GOres.list <- lapply(paste0(seq(0,4)*3, 'h'), 
-                     function(phase){
-                       GOdata <- new("topGOdata", 
-                                     ontology = "CC", 
-                                     geneSel = function(x){x == 1},
-                                     allGenes = getElement(z_max_diff.list, paste(phase)),
-                                     nodeSize = 5,
-                                     annot = annFUN.gene2GO, 
-                                     gene2GO = go_annotation)
-                       resultFis <- runTest(GOdata, 
-                                            algorithm = "classic", 
-                                            statistic = "fisher")
-                       gentable <- GenTable(GOdata, 
-                                              classic = resultFis, 
-                                              orderBy = "classic", 
-                                              ranksOf = "classic",
-                                              topNodes = 25) %>% 
-                           filter(classic < 0.05)
-                       return(list(GOdata = GOdata,
-                                   resultFis = resultFis,
-                                   gentable = gentable))
-                     })
-names(GOres.list) <- paste0(seq(0,4)*3, 'h')
-lapply(GOres.list, getElement, 'gentable')
-
-```
-
-```{r topGO_phase}
-lapply(names(GOres.list), 
-       function(GOres.num){
-         GOres <- getElement(GOres.list, GOres.num)
-         #Get first (most significant) enriched GO id 
-         which_go <- 2
-         goID <- GOres$gentable[which_go, "GO.ID"]
-         goTerm <- GOres$gentable[which_go, "Term"]
-
-         #Show distribution of GO term genes compared to null distribution (all other genes)
-         # print(showGroupDensity(GOres$GOdata, goID, ranks = TRUE))
-
-         #Can use genesInTerm function to extract the genes related to that GO term
-         genesinterm <- genesInTerm(GOres$GOdata, goID)
-         goID_genes <- unlist(genesinterm)
-         names(goID_genes) <- genesinterm
-
-         draw_heatmap_zscore(zscore_list = chab_z.list,
-                             filter_genes_vector = goID_genes,
-                             gene_ordering_vector = order_z_rMT,
-                             threshold_tpm = 1,
-                             use_times = TRUE,
-                             use_order_vector = FALSE,
-                             show_row_names = FALSE,
-                             tpm_mat = TRUE,
-                             column_title_name = paste0(GOres.num, ' - ', goTerm)
-         )
-         return()
-       })
-
-```
-
-Note that some enriched GO terms may not appear in the annotation file (.gaf) directly as they are found to be enriched through the network of related/ancestral terms from the GO network of relations.
-
-Also these phaseograms show every gene in entire dataset associated with the top GO term, so they may not all be as out-of-phase as one another.
-This is a good test to see if this GO enrichment seems correct - if most of the genes look like they're not in the phase expected then the GO enrichment is probably erroneous.

### 2022-07-29 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Create boilerplate Rmd for easier time making future Rmd scripts

@@
 rownames_geneid <- chab_tpm.df$Geneid[chab_tpm.df$Geneid %in% cir_id]
-chab_tpm.df_forpca <- data.frame(select(filter(chab_tpm.df, Geneid %in% cir_id), -Geneid))
+chab_tpm.df_forpca <- data.frame(select(filter(chab_tpm.df, 
+                                               Geneid %in% cir_id), 
+                                        -Geneid))
+
+# To prove scale and centre function work:
+# filter(chab_tpm.df, Geneid %in% cir_id) %>% mutate(across(matches('rMT|SBP'), ~ 1e6*.x/sum(.x), .names = '{.col}_{.fn}')) %>% select(-Geneid) %>% apply(.,2, sum) -> chab_tpm.df_forpca
+
 rownames(chab_tpm.df_forpca) <- rownames_geneid
@@
 
+if(global_save_var){
+  ggsave(filename = 'plots/pca_pir.png',
+          plot = plot,
+          dpi = 300,
+          width = 15,
+          height = 10,
+          units = 'cm')
+}
+
 ```
@@
 
+```
+
+```{r sbp-hi_barchart}
+  cir_plot <- chab_tpm.df %>% 
+    filter(Geneid %in% "PCHAS_1100300") %>% 
+    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
+    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
+    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
+           transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
+    group_by(sample, time, transmission) %>% 
+    dplyr::summarise(sum_tpm = sum(tpm))
+
+y_scale <- seq(0, 
+               signif(ceiling(max(cir_plot$sum_tpm)), 
+                      digits = 2), 
+               length.out = 5)
+
+(
+  p <- ggplot(data = cir_plot, 
+              aes(x = time, 
+                  y = sum_tpm,
+                  fill = transmission)) +
+    stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
+                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
+    scale_y_continuous(breaks = y_scale) +
+    scale_fill_manual(values = cbPalette) +
+    geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
+    ylab(as.expression(bquote('Total SBP-high '~italic(.('pir'))~'TPM'))) +
+    xlab('Asexual blood stage development cycle') +
+    theme_classic() +
+    theme(axis.text.x=element_text(angle = 90, 
+                                   vjust = 0.5,
+                                   size = 12))
+)
+
+if(global_save_var){
+  ggsave(filename = 'plots/SBP-hi_Pir_BarChart.pdf',
+          plot = p,
+          dpi = 300,
+          width = 15,
+          height = 10,
+          units = 'cm')
+}
 ```

### 2022-07-29 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Start Pchabaudi TPM heatmap Rmd

--- 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
@@
 ---
-title: "Manuscript_Github_Pchabaudi_pir_TPMHeatmaps"
+title: "Manuscript_Github_Pchabaudi_boilerplate"
 author: "Timothy Little"
-date: "20/04/2021"
+date: "29/07/2022"
 output: html_document
 ---
+
+Looking at GO enrichment of genes in- and out-of-phase between the 24 asexual life cycles of rMT and SBP transmitted P. chabaudi.
 
@@
 library('viridis')
-library('corrplot')
 library('reshape2')
-library('circlize') # for circular plots and colours
-library('cluster')    # clustering algorithms
-library('factoextra') # clustering algorithms & visualization
-library('ggfortify') # so ggplot understands prcomp objects
-library('ape') # for phylogeny plotting
-library('dendextend') # for dendrogram plotting
-library('readxl') # for reading in Excel spreadsheets
-library('plotly')
-library('scales')
-library('processx') # for downloading the plotly plots
-library('cowplot')
 library('ComplexHeatmap')
-library('ggbiplot')
-library('ggpattern')
+library('circlize') # for circular plots and colour
+library("factoextra")
+library("MatrixGenerics") # for rowMins and rowMax
 
@@
 
-```{r source_fun}
+```{r sourcing_functions}
+source('../rfunctions/uniquify_fun.R')
+source('../rfunctions/ordering_fun.R')
+source('../rfunctions/tpm_fun.R')
 source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
+source('../rfunctions/draw_heatmap_zscore.R')
 ```
 
-```{r order_vec}
+```{r ordering_vec}
 order_vector <- c("Early.Rings",
@@
             delim = '\t')
-
-#Add chapl_aapl/subfam numbers for easier heatmap drawing later.
-cir_info <- cir_info %>% 
-  group_by(chapl_aapl_name) %>% 
-  mutate(chapl_aapl_name_num = n_distinct(Geneid)) %>% 
-  mutate(
-    chapl_aapl_name_num = paste0(
-      chapl_aapl_name, 
-      ' (n = ',
-      chapl_aapl_name_num,
-      ')'
-    )) %>% 
-  group_by(subfam) %>% 
-  mutate(subfam_num = n_distinct(Geneid)) %>% 
-  mutate(
-    subfam_num = paste0(
-      subfam, 
-      ' (n = ',
-      subfam_num,
-      ')'
-    )
-  )
 ```
@@
                                                          replacement = '\\.')
-```
-
-```{r read_TPM}
-#Need to use the calculateTPM script first.
-
-twentyfour_rMT <- read_csv('Pchabaudi_TPM_allBioReps.csv')
-
-twentyfour_avgtpm <- list(
-  'RMT' = list(
-    tpm = read_csv('Pchabaudi_TPM_average.csv'),
-    experiment = 'RMT'
-  )
-)
-
+#Add a number from 1 to 8 for each stage, useful for phase calculations
+times_stages_table$num <- 1:8
 ```
@@
 
-_pir_ genes selected by being expressed above 1 TPM in at least one sample.
+```{r read_TPM}
+#Need to use the calculateTPM script first.
+
+chab_tpm.df <- read_csv('data/Pchabaudi_TPM_all.csv')
+
+chab_tpm_avg.df <- read_csv('data/Pchabaudi_TPM_average.csv')
+
+chab_tpm_avg.list <- list(
+  'RMT' = list(
+    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
+                                pattern = 'rMT|Geneid')],
+    experiment = 'RMT'
+  ),
+  'SBP' = list(
+    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
+                                pattern = 'SBP|Geneid')],
+    experiment = 'SBP'
+  )
+)
+
+```
 
@@
 
-combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
-  filter(Geneid %in% cir_housekeeping_id) %>% 
+combined_data_pir <-  filter(chab_tpm_avg.df,
+                             Geneid %in% cir_housekeeping_id) %>% 
   column_to_rownames(var = 'Geneid') %>% 
@@
 
+```
+
+```{r}
 #Ordering that a more understandable order
@@
                                              drop = FALSE] 
+```
 
+```{r}
 #Add a column for the subfam (ref for the reference genes)
@@
 ```
-
-Add chapl/aapl annotation to above
-
-```{r chapl_aapl_ann}
-
-chapl_aapl_name_vec <- ifelse(grepl(cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), 
-                                                                                     cir_info$Geneid)], 
-                                                  pattern = 'NA'),
-                                            NA,
-                                            cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), cir_info$Geneid)])
-
-#chapl_aapl_name annotation
-
-annotation_col <- chapl_aapl_name_col_num[ordering(chapl_aapl_name_vec,
-                                                   order_vector = 'ChAPL')]
-
-chapl_aapl_name_ha <- rowAnnotation(
-  'CHAPL/AAPL' = anno_simple(
-    chapl_aapl_name_vec,
-    col = annotation_col,
-    na_col = 'white'
-  ),
-  annotation_name_gp = gpar(fontsize = 10)
-)
-
-chapl_aapl_name_lgd <- Legend(
-  title = 'CHAPL/AAPL',
-  at = names(annotation_col)
-,
-  legend_gp = gpar(fill = na.omit(annotation_col))
-)
-
-ht <- Heatmap(
-  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
-  cluster_rows = FALSE,
-  cluster_columns = FALSE,
-  show_row_names = FALSE,
-  name = 'log(TPM+1)', 
-  col = colorRamp2(c(0,
-                     maxcol/4,
-                     maxcol/2,
-                     3*maxcol/4,
-                     maxcol),
-                   viridis(5)),
-  na_col = 'white', 
-  border = TRUE,
-  border_gp = gpar(lwd = 0.1),
-  column_names_gp = gpar(fontsize = 10),
-  row_dend_width = unit(3, "cm"),
-  width = unit(4.5, 'cm'),
-  height = unit(7,'cm'),
-  # row_title = NULL,
-  left_annotation = chapl_aapl_name_ha,
-  row_title_rot = 0,
-  row_title_gp = gpar(fontsize = 10),
-  row_split = forcats::fct_shift(
-    relevel(
-      as.factor(combined_data_pir_ht$subfam),
-      ref = 'ref')
-  ),
-  heatmap_legend_param = list(
-    labels_gp = gpar(fontsize = 10),
-    title_gp = gpar(fontsize = 10,
-                    fontface = 'bold'),
-    at = c(0,
-           maxcol/4,
-           maxcol/2,
-           3*maxcol/4,
-           maxcol)
-  )
-)
-
-draw(ht, 
-     row_title_side = 'left',
-     column_title_side = 'bottom',
-     heatmap_legend_side = 'right',
-     annotation_legend_side = 'right',
-     annotation_legend_list = list(chapl_aapl_name_lgd),
-     merge_legend = TRUE)
-
-if(global_save_var){
-  pdf(file = 'Interest_SubfamSplit_ChAPL-AAPLAnnotate_heatmap.pdf',
-      height = 5,
-      width = 5)
-draw(ht, 
-     row_title_side = 'left',
-     column_title_side = 'bottom',
-     heatmap_legend_side = 'right',
-     annotation_legend_side = 'right',
-     annotation_legend_list = list(chapl_aapl_name_lgd),
-     merge_legend = TRUE)
-  dev.off()
-}
-
-```
-
-# Supp figure 9 - Pir expression of AAPLs and ChAPLs.
-
-_pir_ genes selected by being expressed above 1 TPM in at least one sample.
-
-```{r chapl_appl_heatmap_suppfig9}
-
-# Get order of pir genes by highest expressed genes in each sample.
-
-combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
-  filter(Geneid %in% cir_id) %>% 
-  mutate(chapl_aapl_name = cir_info$chapl_aapl_name_num[match(Geneid, cir_info$Geneid)]) %>% 
-  filter(!str_detect(chapl_aapl_name, pattern = 'NA')) %>% 
-  column_to_rownames(var = 'Geneid') %>% 
-  filter_all(
-      any_vars(. > 1)
-  )
-
-combined_data_ordered <- tibble(Geneid = rownames(combined_data_pir),
-                                means = apply(combined_data_pir[,
-                                                                str_detect(colnames(combined_data_pir), 
-                                                                           'chapl_aapl',
-                                                                           negate = TRUE)],
-                                              1,
-                                              mean),
-                                chapl_aapl_name = factor(
-                                  combined_data_pir$chapl_aapl_name
-                                )) %>% 
-  arrange(chapl_aapl_name, 
-          desc(means)) %>% 
-  .$Geneid
-
-#Order the heatmap data by the order defined above (chapl_aapl_name, then mean expression).
-combined_data_pir_ht <- combined_data_pir[
-  order(
-    match(rownames(combined_data_pir),
-          combined_data_ordered),
-    na.last = NA #Remove NAs
-  ),
-]
-
-#Change the colnames to the times
-colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                              pattern = '[[:digit:]]{2}h|chapl_aapl_name')
-
-#Order by time point
-combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                             drop = FALSE]
-
-#Get the ref gene actual names instead of the Geneid
-rownames(combined_data_pir_ht) <- ifelse(
-  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
-  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
-  rownames(combined_data_pir_ht)
-)
-
-#Get the colour range
-maxcol <- ceiling(
-  max(
-    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name'])
-  ))
-
-#chapl_aapl_name name labels
-num_chapl_aapl_name <- cir_info %>% 
-  group_by(chapl_aapl_name) %>% 
-  summarise(num = n_distinct(Gene.ID))
-
-labels_chapl_aapl_name <- setNames(paste0(num_chapl_aapl_name$chapl_aapl_name, 
-                                          ' (n = ', num_chapl_aapl_name$num, ')'),
-                                   num_chapl_aapl_name$chapl_aapl_name)
-
-#loci annotation
-chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
-annotation_col <- chapl_aapl_name_col_num[ordering(labels_chapl_aapl_name,
-                                                    order_vector = 'ChAPL')]
-
-ht <- Heatmap(
-  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
-  cluster_rows = FALSE,
-  cluster_columns = FALSE,
-  show_row_names = FALSE,
-  name = 'log(TPM+1)', 
-  col = colorRamp2(c(0,
-                     maxcol/4,
-                     maxcol/2,
-                     3*maxcol/4,
-                     maxcol),
-                   viridis(5)),
-  na_col = 'white', 
-  border = TRUE,
-  border_gp = gpar(lwd = 0.1),
-  column_names_gp = gpar(fontsize = 10),
-  row_dend_width = unit(3, "cm"),
-  width = unit(4.5, 'cm'),
-  height = unit(7,'cm'),
-  # row_title = NULL,
-  row_title_rot = 0,
-  row_title_gp = gpar(fontsize = 7),
-  row_split = 
-    factor(combined_data_pir_ht$chapl_aapl_name,
-              levels = ordering(combined_data_pir_ht$chapl_aapl_name,
-                                order_vector = 'ChAPL')),
-  heatmap_legend_param = list(
-    labels_gp = gpar(fontsize = 10),
-    title_gp = gpar(fontsize = 10,
-                    fontface = 'bold'),
-    at = c(0,
-           maxcol/4,
-           maxcol/2,
-           3*maxcol/4,
-           maxcol)
-  ),
-  row_gap = unit(c(rep.int(1, times = 2),
-                   3,
-                   rep.int(1, times = 4)), "mm")
-)
-
-if(global_save_var){
-  pdf(file = 'Manuscript_SuppFig_ChAPL-AAPL_heatmap.pdf',
-      height = 5,
-      width = 5)
-  draw(ht, 
-       row_title_side = 'left',
-       column_title_side = 'bottom'
-  )
-  dev.off()
-}
-
-  draw(ht, 
-       row_title_side = 'left',
-       column_title_side = 'bottom'
-  )
-```
-
-```{r}
-chapl_aapl_name_ha <- rowAnnotation(
-  'CHAPL/AAPL' = anno_simple(
-    chapl_aapl_name_info,
-    col = annotation_col,
-  ),
-  annotation_name_gp = gpar(fontsize = 10)
-)
-
-chapl_aapl_name_lgd <- Legend(
-  title = 'CHAPL/AAPL',
-  at = str_subset(
-    sort(
-      factor(labels_chapl_aapl_name,
-             ordering(labels_chapl_aapl_name,
-                      order_vector = 'ChAPL'))
-    ),
-    pattern = 'NA',
-    negate = TRUE
-  ),
-  legend_gp = gpar(fill = na.omit(annotation_col))
-)
-
-ht <- Heatmap(
-  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
-  cluster_rows = TRUE,
-  cluster_columns = FALSE,
-  show_row_names = FALSE,
-  name = 'log(TPM+1)', 
-  col = colorRamp2(c(0,
-                     maxcol/4,
-                     maxcol/2,
-                     3*maxcol/4,
-                     maxcol),
-                   viridis(5)),
-  na_col = 'white', 
-  border = TRUE,
-  border_gp = gpar(lwd = 0.1),
-  column_names_gp = gpar(fontsize = 10),
-  row_dend_width = unit(3, "cm"),
-  width = unit(4.5, 'cm'),
-  height = unit(7,'cm'),
-  row_title = NULL,
-  left_annotation = chapl_aapl_name_ha,
-  row_km = 9,
-  heatmap_legend_param = list(
-    labels_gp = gpar(fontsize = 10),
-    title_gp = gpar(fontsize = 10,
-                    fontface = 'bold'),
-    at = c(0,
-           maxcol/4,
-           maxcol/2,
-           3*maxcol/4,
-           maxcol)
-  )
-)
-
-draw(ht, 
-     row_title_side = 'left',
-     column_title_side = 'bottom',
-     heatmap_legend_side = 'right',
-     annotation_legend_side = 'right',
-     annotation_legend_list = list(chapl_aapl_name_lgd),
-     merge_legend = TRUE)
-
-if(global_save_var){
-  pdf(file = 'Interest_chapl-aapl_clustered_heatmap.pdf',
-      height = 5,
-      width = 5)
-draw(ht, 
-     row_title_side = 'left',
-     column_title_side = 'bottom',
-     heatmap_legend_side = 'right',
-     annotation_legend_side = 'right',
-     annotation_legend_list = list(chapl_aapl_name_lgd),
-     merge_legend = TRUE)
-  dev.off()
-}
-```

### 2022-07-30 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Fix heatmap for list of rMT and SBP heatmaps

@@
             delim = '\t')
+
+#Add chapl_aapl/subfam numbers for easier heatmap drawing later.
+cir_info <- cir_info %>% 
+  group_by(chapl_aapl_name) %>% 
+  mutate(chapl_aapl_name_num = n_distinct(Geneid)) %>% 
+  mutate(
+    chapl_aapl_name_num = paste0(
+      chapl_aapl_name, 
+      ' (n = ',
+      chapl_aapl_name_num,
+      ')'
+    )) %>% 
+  group_by(subfam) %>% 
+  mutate(subfam_num = n_distinct(Geneid)) %>% 
+  mutate(
+    subfam_num = paste0(
+      subfam, 
+      ' (n = ',
+      subfam_num,
+      ')'
+    )
+  )
 ```
@@
 colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                              pattern = '[[:digit:]]{2}h')
+                                              pattern = '.{3}_([[:digit:]]{2}h|sch.culture)')
 
@@
                    'ref' = 'ref')
+```
+
+```{r draw_heatmap}
+ht.list <- lapply(c('rMT', 'SBP'),
+                  function(transmission){
+
+                    ht.df <- combined_data_pir_ht[
+                      ,
+                      grepl(colnames(combined_data_pir_ht), 
+                            pattern = paste0(transmission))
+                    ]
+
+                    colnames(ht.df) <- str_remove(colnames(ht.df),
+                                                  pattern = paste0(transmission, 
+                                                                   '_'))
 
                     ht <- Heatmap(
-  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
+                      log2(
+                        1+ht.df
+                      ),
                       cluster_rows = FALSE,
@@
                     )
+                    return(ht)
+                  })
+
+ht.list <- Reduce('+', ht.list)
 
@@
 
-draw(ht, 
-     row_title_side = 'left',
-     column_title_side = 'bottom',
-     # heatmap_legend_side = 'right',
-     # annotation_legend_side = 'right',
-     # annotation_legend_list = list(clade_lgd),
-     merge_legend = TRUE)
-
 if(global_save_var){
@@
       width = 5)
-draw(ht, 
+  draw(ht.list, 
        row_title_side = 'left',
@@
        heatmap_legend_side = 'right',
-     # annotation_legend_side = 'right',
-     # annotation_legend_list = list(clade_lgd),
        merge_legend = TRUE)

### 2022-07-30 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Save tpm subfam heatmap

@@
 if(global_save_var){
-  pdf(file = 'Manuscript_Fig_subfam_heatmap.pdf',
+  pdf(file = 'plots/subfam_TPMheatmap.pdf',
       height = 5,
-      width = 5)
+      width = 8)
   draw(ht.list, 

### 2022-07-30 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Heatmap column titles

@@
                       height = unit(7,'cm'),
+                      column_title = paste(transmission),
                       row_title_rot = 0,

### 2022-07-30 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Alter dimensions of saved heatmap

@@
   pdf(file = 'plots/subfam_TPMheatmap.pdf',
-      height = 5,
-      width = 8)
+      height = 8,
+      width = 6)
   draw(ht.list, 

### 2022-07-30 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Order heatmap df by mean TPM because Complex Heatmaps will split by sub-fams itself

@@
 
-```{r}
-#Ordering that a more understandable order
-
-combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
-                                means = apply(combined_data_pir, 1, mean),
-                                subfam = factor(
-                                  ifelse(rownames(combined_data_pir) %in% cir_id,
-                                         cir_info$subfam[match(rownames(combined_data_pir), 
-                                                               cir_info$Geneid)],
-                                         'ref')
-                                )) %>% 
-  mutate(subfam =
-           forcats::fct_shift(
-             relevel(
-               as.factor(subfam),
-               ref = 'ref')
-           )
-  ) %>% 
-  # filter(means > 1) %>% 
-  arrange(subfam, 
-          desc(means)) %>% 
-  .$Geneid
-
-#Order the heatmap data by the order defined above (Subfam, then mean expression).
-combined_data_pir_ht <- combined_data_pir[
-  order(
-    match(rownames(combined_data_pir),
-          combined_data_ordered),
-    na.last = NA #Remove NAs
-  ),
-]
-
+```{r heatmap_df}
+combined_data_pir_ht <- combined_data_pir
 #Change the colnames to the times
@@
                                              drop = FALSE]
-```
 
-```{r}
+#Add mean column for ordering
+combined_data_pir_ht$mean_tpm <- apply(combined_data_pir_ht, 1, mean)
+
+combined_data_pir_ht <- arrange(combined_data_pir_ht, desc(mean_tpm))
+
 #Add a column for the subfam (ref for the reference genes)

### 2022-08-02 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Add subfam annotations to the chapl aapl heatmap

@@
 ```
+
+```{r chapl_aapl_heatmap}
+
+combined_data_pir_ht$locus <- cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), 
+                                    cir_info$Gene.ID)]
+
+#Remove the pirs in the 'NA' locus (i.e. not in a chapl or aapl)
+# don't remove actual NAs because these are the ref genes.
+combined_data_pir_locus_ht <- filter(combined_data_pir_ht,
+                                     !grepl(locus, pattern = "NA \\(n" ))
+
+combined_data_pir_locus_ht$locus <- ifelse(is.na(combined_data_pir_locus_ht$locus),
+                                     'ref',
+                                     combined_data_pir_locus_ht$locus)
+
+#Change locus to factor and change levels so that AAPLs are before ChAPLs and are together.
+combined_data_pir_locus_ht$locus <- factor(combined_data_pir_locus_ht$locus,
+                                           levels = ordering(combined_data_pir_locus_ht$locus,
+                                                             order_vector = c('AAPL',
+                                                                              'ChAPL',
+                                                                              'ref')))
+
+clade_info <- getElement(cir_info,
+                         'subfam')[match(rownames(combined_data_pir_locus_ht), 
+                                                 cir_info$Geneid)]
+clade_info[is.na(clade_info)] <- 'ref'
+names(clade_info) <- rownames(combined_data_pir_locus_ht)
+clade_ha <- rowAnnotation(
+  'subfamily' = anno_simple(
+    clade_info,
+    col = annotation_col
+  )
+)
+#Draw the clade legend for the subfamily
+clade_lgd <- Legend(
+  title = 'sub-family',
+  at = names(annotation_col),
+  legend_gp = gpar(fill = annotation_col)
+)
+
+#Get the colour range
+maxcol <- ceiling(
+  max(
+    log2(1+combined_data_pir_locus_ht[
+      !colnames(combined_data_pir_locus_ht) %in% c('subfam', 
+                                                   'locus')
+    ])
+  ))
+```
+
+```{r draw_locus_heatmap}
+
+ht.list <- lapply(c('rMT', 'SBP'),
+                  function(transmission){
+
+                    #Get either the rMT or the SBP data
+                    ht.df <- combined_data_pir_locus_ht[
+                      ,
+                      grepl(colnames(combined_data_pir_locus_ht), 
+                            pattern = paste0(transmission))
+                    ]
+
+                    #Remove the transmission name from the column names
+                    colnames(ht.df) <- str_remove(colnames(ht.df),
+                                                  pattern = paste0(transmission, 
+                                                                   '_'))
+
+                    ht <- Heatmap(
+                      log2(
+                        1+ht.df
+                      ),
+                      cluster_rows = FALSE,
+                      cluster_columns = FALSE,
+                      show_row_names = FALSE,
+                      name = 'log(TPM+1)', 
+                      col = colorRamp2(c(0,
+                                         maxcol/4,
+                                         maxcol/2,
+                                         3*maxcol/4,
+                                         maxcol),
+                                       viridis(5)),
+                      na_col = 'white', 
+                      border = TRUE,
+                      border_gp = gpar(lwd = 0.1),
+                      column_names_gp = gpar(fontsize = 10),
+                      row_dend_width = unit(3, "cm"),
+                      width = unit(4.5, 'cm'),
+                      height = unit(12,'cm'),
+                      column_title = paste(transmission),
+                      row_title_rot = 0,
+                      row_title_gp = gpar(fontsize = 10),
+                      row_split = forcats::fct_shift(
+                        relevel(
+                          as.factor(combined_data_pir_locus_ht$locus),
+                          ref = 'ref')
+                      ),
+                      heatmap_legend_param = list(
+                        labels_gp = gpar(fontsize = 10),
+                        title_gp = gpar(fontsize = 10,
+                                        fontface = 'bold'),
+                        at = c(0,
+                               maxcol/4,
+                               maxcol/2,
+                               3*maxcol/4,
+                               maxcol)
+                      ),
+                      #Only want the clade annotation on one heatmap not both.
+                      left_annotation = if(transmission == 'rMT'){clade_ha}else{NULL}
+                    )
+                    return(ht)
+                  })
+
+ht.list <- Reduce('+', ht.list)
+
+if(global_save_var){
+  pdf(file = 'plots/ChAPL-AAPL_TPMheatmap.pdf',
+      height = 9,
+      width = 6)
+  draw(ht.list, 
+       row_title_side = 'left',
+       column_title_side = 'bottom',
+       heatmap_legend_side = 'right',
+       merge_legend = TRUE,
+       annotation_legend_list = list(clade_lgd))
+  dev.off()
+}
+
+```

### 2022-08-02 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Removed variable combined_data_pir as it was redundant

@@
 
-combined_data_pir <-  filter(chab_tpm_avg.df,
+combined_data_pir_ht <-  filter(chab_tpm_avg.df,
                              Geneid %in% cir_housekeeping_id) %>% 
@@
 ```{r heatmap_df}
-combined_data_pir_ht <- combined_data_pir
 #Change the colnames to the times

### 2022-08-03 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
Make labels for qusage comparison of chabaudi 24h

@@
                       width = unit(4.5, 'cm'),
-                      height = unit(7,'cm'),
+                      height = unit(12,'cm'),
                       column_title = paste(transmission),

### 2022-08-03 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
Make heatmap of qusage results for stages between transmissions, also do qusage analysis for said comparisons

--- 2022-08-02 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
@@
 library("MatrixGenerics") # for rowMins and rowMax
+library('qusage')
 
@@
 
-cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
+cir_info <- read_delim('../standard_files/Pchabaudi_cir_info.txt',
            delim = '\t')
@@
 
-chab_tpm_avg.list <- list(
-  'RMT' = list(
-    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
-                                pattern = 'rMT|Geneid')],
-    experiment = 'RMT'
-  ),
-  'SBP' = list(
-    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
-                                pattern = 'SBP|Geneid')],
-    experiment = 'SBP'
-  )
-)
+# chab_tpm_avg.list <- list(
+#   'RMT' = list(
+#     tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
+#                                 pattern = 'rMT|Geneid')],
+#     experiment = 'RMT'
+#   ),
+#   'SBP' = list(
+#     tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
+#                                 pattern = 'SBP|Geneid')],
+#     experiment = 'SBP'
+#   )
+# )
 
@@
 
-Adding in housekeeping genes for comparison:
-HSP70 putative  PCHAS_0721000
-MSP1 PCHAS_0831300
-AMA1 PCHAS_0931000
+```{r df}
+#Filter low expressed genes (greater than 10 tpm in at least one sample), convert to a matrix, add 1 for the log transformation, then transform to log2
+chab_tpm.df <-column_to_rownames(chab_tpm.df, var = 'Geneid')
+chab_tpm.df <- chab_tpm.df[apply(chab_tpm.df, 1, function(x) {any(x >= 10)}),]
+chab_tpm_log2plus1.mat <- chab_tpm.df %>% 
+  as.matrix %>% 
+  `+`(.+1) %>% 
+  log(base = 2)
+chab_tpm_log2plus1.mat[1:5,1:5]
+```
 
-```{r highest_expressed_stage_heatmap_fig3b}
+```{r generate_labels}
+#Use make.names to remove the '-' from the labels
+colnames(chab_tpm_log2plus1.mat) <- make.names(colnames(chab_tpm_log2plus1.mat))
+```
 
-housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
-                                 'MSP1' = 'PCHAS_0831300',
-                                 'AMA1' = 'PCHAS_0931000')
+Removing genesets with < 3 members.
 
-cir_housekeeping_id <- c(cir_id, 
-                         'PCHAS_0721000',
-                         'PCHAS_0831300',
-                         'PCHAS_0931000')
+```{r generate_genesets}
+pir_geneset <- list(pirs = cir_info$Geneid)
 
-# Get order of pir genes by highest expressed genes in each sample.
+subfam_list <- lapply(unique(cir_info$subfam), 
+                      function(clade_nam) filter(cir_info, 
+                                                 subfam == clade_nam)$Geneid)
+names(subfam_list) <- unique(cir_info$subfam)
+subfam_list <- subfam_list[unlist(lapply(subfam_list, 
+                                         function(x) length(x) > 3))]
 
-combined_data_pir_ht <-  filter(chab_tpm_avg.df,
-                             Geneid %in% cir_housekeeping_id) %>% 
-  column_to_rownames(var = 'Geneid') %>% 
-  filter_all(
-    any_vars(. > 10)
-  )
+ShortLong_list <- 
+  lapply(unique(cir_info$SorL), 
+         function(SorL_nam) filter(cir_info, 
+                                   SorL == SorL_nam)$Geneid)
+names(ShortLong_list) <- unique(cir_info$SorL)
+ShortLong_list <- ShortLong_list[unlist(lapply(ShortLong_list, 
+                                               function(x) length(x) > 3))]
+order_path.names <- ordering(c('pirs', unique(cir_info$SorL), sort(unique(cir_info$subfam))), 
+                             order_vector = c('pir','L','S'))
+
+locus_list <- 
+  lapply(unique(cir_info$chapl_aapl_name), 
+         function(chapl_aapl_name_nam) filter(cir_info, 
+                                   chapl_aapl_name == chapl_aapl_name_nam)$Geneid)
+names(locus_list) <- unique(cir_info$chapl_aapl_name)
+locus_list <- locus_list[unlist(lapply(locus_list, 
+                                               function(x) length(x) > 3))]
 
@@
 
-```{r heatmap_df}
-#Change the colnames to the times
-colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                              pattern = '.{3}_([[:digit:]]{2}h|sch.culture)')
+Which comparisons do I want to make?
 
-#Order by time point
-combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                             drop = FALSE]
-
-#Add mean column for ordering
-combined_data_pir_ht$mean_tpm <- apply(combined_data_pir_ht, 1, mean)
-
-combined_data_pir_ht <- arrange(combined_data_pir_ht, desc(mean_tpm))
-
-#Add a column for the subfam (ref for the reference genes)
-combined_data_pir_ht$subfam <- ifelse(rownames(combined_data_pir_ht) %in% cir_id,
-                                      cir_info$subfam_num[match(rownames(combined_data_pir_ht), 
-                                                            cir_info$Geneid)],
-                                      'ref')
-
-#Get the ref gene actual names instead of the Geneid
-rownames(combined_data_pir_ht) <- ifelse(
-  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
-  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
-  rownames(combined_data_pir_ht)
+```{r running_qusage}
+ordered_stages <- ordering(unique(lifeCycle_labels),
+         order_vector = times_stages_table$time)
+#Remove sch.culture from the comparisons for now.
+ordered_stages <- ordered_stages[!ordered_stages == 'sch.culture']
+comparisons <- c(
+  unlist(
+    lapply(seq_along(ordered_stages),
+           function(i){ 
+             if(i < length(ordered_stages)){ 
+               paste0(ordered_stages[i+1],'-',ordered_stages[i]) 
+             } else {
+               #When at the end, want it to cycle back round
+               paste0(ordered_stages[1],'-',ordered_stages[i]) 
+             } 
+           }
     )
-
-#Get the colour range
-maxcol <- ceiling(
-  max(
-    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'subfam'])
-  ))
-
-#Sub-family annotation
-clade_info <- combined_data_pir_ht$subfam
-annotation_col <- c(clade_col, 'ref' = 'black')
-
-#Subfamily annotation
-
-clade_ha <- rowAnnotation(
-  'sub-family' = anno_simple(
-    clade_info,
-    col = annotation_col,
-  ),
-  annotation_name_gp = gpar(fontsize = 10)
   )
-
-#Subfamily name labels
-num_subfam <- cir_info %>% 
-  group_by(subfam) %>% 
-  summarise(num = n_distinct(Gene.ID))
-
-labels_subfam <- c(setNames(paste0(num_subfam$subfam, ' (n = ', num_subfam$num, ')'),
-                            num_subfam$subfam),
-                   'ref' = 'ref')
+)
 ```
 
-```{r draw_heatmap}
-ht.list <- lapply(c('rMT', 'SBP'),
-                  function(transmission){
+Compare the gene lists between the same stages in different transmission
 
-                    ht.df <- combined_data_pir_ht[
-                      ,
-                      grepl(colnames(combined_data_pir_ht), 
-                            pattern = paste0(transmission))
+```{r qusage_stagesBetweenTransmission}
+
+comparisons_stagesTrans <- paste0(paste('rMT', 
+                                        times_stages_table$time, 
+                                        sep = '_'),
+                                  '-',
+                                  paste('SBP', 
+                                        times_stages_table$time, 
+                                        sep = '_'))
+
+comp_stagesTrans.mat <- chab_tpm_log2plus1.mat[
+  ,!grepl(colnames(chab_tpm_log2plus1.mat), pattern = 'sch.culture')
 ]
+# For the life cycle labels then extract the transmission + timepoint label
+lifeCycle_labels <- str_extract(colnames(comp_stagesTrans.mat),
+                                pattern = '^(rMT|SBP)_([[:digit:]]{2}h|sch.culture)')
 
-                    colnames(ht.df) <- str_remove(colnames(ht.df),
-                                                  pattern = paste0(transmission, 
-                                                                   '_'))
+all_comparisons <- lapply(comparisons_stagesTrans, 
+                          function(comparison){
+                            print(comparison)
+                            list_comp <- lapply(
+                              list(pir_geneset,
+                                   ShortLong_list,
+                                   subfam_list,
+                                   locus_list), 
+                              function(gene_list){
+                                print(deparse(substitute(gene_list)))
+                                qs.results.subfam <- qusage(comp_stagesTrans.mat,
+                                                            lifeCycle_labels,
+                                                            comparison,
+                                                            gene_list,
+                                                            n.points = 2^18)
+                                list(results = qs.results.subfam, 
+                                     qstable = qsTable(qs.results.subfam) %>% 
+                                       mutate(contrast = qs.results.subfam$contrast))
+                              })
+                            names(list_comp) <- c('pir', 'shortlong', 'subfam', 'locus')
+                            list_comp$combined <- bind_rows(lapply(list_comp, 
+                                                                   getElement, 
+                                                                   'qstable'))
+                            return(list_comp)
+                          })
+names(all_comparisons) <- comparisons
 
-                    ht <- Heatmap(
-                      log2(
-                        1+ht.df
-                      ),
+all_comparisons.df <- bind_rows(lapply(all_comparisons, 
+                                       function(x) bind_rows(lapply(x, 
+                                                                    function(y) y$qstable)))) %>% 
+  arrange(pathway.name)
+
+stagesTrans_ht.mat <- filter(all_comparisons.df, FDR < 0.05) %>% 
+  mutate(pathway.name = factor(pathway.name, 
+                           levels = ordering(pathway.name, 
+                                             order_vector = c('pirs', 'L', 'S', 'AAPL', 'ChAPL')))) %>% 
+  arrange(pathway.name) %>% 
+  melt(id.vars = c('contrast', 'pathway.name'),
+       measure.vars  = c('log.fold.change')) %>% 
+  dcast(contrast ~ pathway.name, 
+        value.var = 'value') %>% 
+  column_to_rownames('contrast') %>% 
+  as.matrix
+
+maxcol <- my_max(stagesTrans_ht.mat)
+mincol <- my_max(stagesTrans_ht.mat*-1)*-1
+
+ht <- Heatmap(stagesTrans_ht.mat,
         cluster_rows = FALSE,
         cluster_columns = FALSE,
-                      show_row_names = FALSE,
-                      name = 'log(TPM+1)', 
-                      col = colorRamp2(c(0,
+        cluster_column_slices = FALSE,
+        na_col = 'white',
+        column_split = factor(str_extract(colnames(stagesTrans_ht.mat),
+                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'),
+                              levels = unique(str_extract(colnames(stagesTrans_ht.mat),
+                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'))),
+        column_order = colnames(stagesTrans_ht.mat),
+        column_gap = unit(.5,"cm"),
+        border = TRUE,
+        col = colorRamp2(c(mincol,
+                           0,
                            maxcol/4,
                            maxcol/2,
-                                         3*maxcol/4,
                            maxcol),
-                                       viridis(5)),
-                      na_col = 'white', 
-                      border = TRUE,
-                      border_gp = gpar(lwd = 0.1),
-                      column_names_gp = gpar(fontsize = 10),
-                      row_dend_width = unit(3, "cm"),
-                      width = unit(4.5, 'cm'),
-                      height = unit(7,'cm'),
-                      column_title = paste(transmission),
-                      row_title_rot = 0,
-                      row_title_gp = gpar(fontsize = 10),
-                      row_split = forcats::fct_shift(
-                        relevel(
-                          as.factor(combined_data_pir_ht$subfam),
-                          ref = 'ref')
-                      ),
-                      heatmap_legend_param = list(
-                        labels_gp = gpar(fontsize = 10),
-                        title_gp = gpar(fontsize = 10,
-                                        fontface = 'bold'),
-                        at = c(0,
-                               maxcol/4,
-                               maxcol/2,
-                               3*maxcol/4,
-                               maxcol)
-                      )
-                    )
-                    return(ht)
-                  })
-
-ht.list <- Reduce('+', ht.list)
-
-clade_lgd <- Legend(
-  title = 'sub-family',
-  at = c(labels_subfam),
-  legend_gp = gpar(fill = annotation_col)
-)
+                         c('dodgerblue1', 'white', 'yellow', 'orange', 'red')),
+        heatmap_legend_param = list(title = 'LogFoldChange'))
 
 if(global_save_var){
-  pdf(file = 'plots/subfam_TPMheatmap.pdf',
+  pdf(file = 'plots/stageTrans_logFoldheatmap.pdf',
       height = 8,
       width = 6)
-  draw(ht.list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'right',
-       merge_legend = TRUE)
-  dev.off()
-}
-```
-
-```{r chapl_aapl_heatmap}
-
-combined_data_pir_ht$locus <- cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), 
-                                    cir_info$Gene.ID)]
-
-#Remove the pirs in the 'NA' locus (i.e. not in a chapl or aapl)
-# don't remove actual NAs because these are the ref genes.
-combined_data_pir_locus_ht <- filter(combined_data_pir_ht,
-                                     !grepl(locus, pattern = "NA \\(n" ))
-
-combined_data_pir_locus_ht$locus <- ifelse(is.na(combined_data_pir_locus_ht$locus),
-                                     'ref',
-                                     combined_data_pir_locus_ht$locus)
-
-#Change locus to factor and change levels so that AAPLs are before ChAPLs and are together.
-combined_data_pir_locus_ht$locus <- factor(combined_data_pir_locus_ht$locus,
-                                           levels = ordering(combined_data_pir_locus_ht$locus,
-                                                             order_vector = c('AAPL',
-                                                                              'ChAPL',
-                                                                              'ref')))
-
-clade_info <- getElement(cir_info,
-                         'subfam')[match(rownames(combined_data_pir_locus_ht), 
-                                                 cir_info$Geneid)]
-clade_info[is.na(clade_info)] <- 'ref'
-names(clade_info) <- rownames(combined_data_pir_locus_ht)
-clade_ha <- rowAnnotation(
-  'subfamily' = anno_simple(
-    clade_info,
-    col = annotation_col
-  )
-)
-#Draw the clade legend for the subfamily
-clade_lgd <- Legend(
-  title = 'sub-family',
-  at = names(annotation_col),
-  legend_gp = gpar(fill = annotation_col)
-)
-
-#Get the colour range
-maxcol <- ceiling(
-  max(
-    log2(1+combined_data_pir_locus_ht[
-      !colnames(combined_data_pir_locus_ht) %in% c('subfam', 
-                                                   'locus')
-    ])
-  ))
-```
-
-```{r draw_locus_heatmap}
-
-ht.list <- lapply(c('rMT', 'SBP'),
-                  function(transmission){
-
-                    #Get either the rMT or the SBP data
-                    ht.df <- combined_data_pir_locus_ht[
-                      ,
-                      grepl(colnames(combined_data_pir_locus_ht), 
-                            pattern = paste0(transmission))
-                    ]
-
-                    #Remove the transmission name from the column names
-                    colnames(ht.df) <- str_remove(colnames(ht.df),
-                                                  pattern = paste0(transmission, 
-                                                                   '_'))
-
-                    ht <- Heatmap(
-                      log2(
-                        1+ht.df
-                      ),
-                      cluster_rows = FALSE,
-                      cluster_columns = FALSE,
-                      show_row_names = FALSE,
-                      name = 'log(TPM+1)', 
-                      col = colorRamp2(c(0,
-                                         maxcol/4,
-                                         maxcol/2,
-                                         3*maxcol/4,
-                                         maxcol),
-                                       viridis(5)),
-                      na_col = 'white', 
-                      border = TRUE,
-                      border_gp = gpar(lwd = 0.1),
-                      column_names_gp = gpar(fontsize = 10),
-                      row_dend_width = unit(3, "cm"),
-                      width = unit(4.5, 'cm'),
-                      height = unit(12,'cm'),
-                      column_title = paste(transmission),
-                      row_title_rot = 0,
-                      row_title_gp = gpar(fontsize = 10),
-                      row_split = forcats::fct_shift(
-                        relevel(
-                          as.factor(combined_data_pir_locus_ht$locus),
-                          ref = 'ref')
-                      ),
-                      heatmap_legend_param = list(
-                        labels_gp = gpar(fontsize = 10),
-                        title_gp = gpar(fontsize = 10,
-                                        fontface = 'bold'),
-                        at = c(0,
-                               maxcol/4,
-                               maxcol/2,
-                               3*maxcol/4,
-                               maxcol)
-                      ),
-                      #Only want the clade annotation on one heatmap not both.
-                      left_annotation = if(transmission == 'rMT'){clade_ha}else{NULL}
-                    )
-                    return(ht)
-                  })
-
-ht.list <- Reduce('+', ht.list)
-
-if(global_save_var){
-  pdf(file = 'plots/ChAPL-AAPL_TPMheatmap.pdf',
-      height = 9,
-      width = 6)
-  draw(ht.list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'right',
-       merge_legend = TRUE,
-       annotation_legend_list = list(clade_lgd))
+  draw(ht)
   dev.off()

### 2022-08-04 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
Transpose the stages comp between transmission heatmap qusage analysis

@@
 options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
+
+#Set random seed for reproducible randomisation
+set.seed(34576)
 ```
@@
 
-ht <- Heatmap(stagesTrans_ht.mat,
+ht <- Heatmap(t(stagesTrans_ht.mat),
         cluster_rows = FALSE,
@@
         na_col = 'white',
-        column_split = factor(str_extract(colnames(stagesTrans_ht.mat),
+        row_split = factor(str_extract(colnames(stagesTrans_ht.mat),
                                    pattern = 'pirs|^L|^S|AAPL|ChAPL'),
@@
                                    pattern = 'pirs|^L|^S|AAPL|ChAPL'))),
-        column_order = colnames(stagesTrans_ht.mat),
-        column_gap = unit(.5,"cm"),
+        row_order = colnames(stagesTrans_ht.mat),
+        row_gap = unit(.5,"cm"),
+        row_names_gp = gpar(fontsize = 9),
         border = TRUE,

### 2022-08-04 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
Add some comments

@@
 maxcol <- my_max(stagesTrans_ht.mat)
+#Multiply by minus one to use my_max to get the most negative value
 mincol <- my_max(stagesTrans_ht.mat*-1)*-1
 
+#Transpose the matrix so that the stages (comparison) are on the x axis like other figures.
 ht <- Heatmap(t(stagesTrans_ht.mat),
@@
         na_col = 'white',
+        #Use factors to keep the ordering that we want.
         row_split = factor(str_extract(colnames(stagesTrans_ht.mat),
@@
         border = TRUE,
+        #Using this colour scheme because mincol is a lot less negative 
+        # than maxcol is positive
         col = colorRamp2(c(mincol,

### 2022-08-04 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
Analyse the stages comp within transmission heatmap

@@
 
-Which comparisons do I want to make?
-
-```{r running_qusage}
-ordered_stages <- ordering(unique(lifeCycle_labels),
-         order_vector = times_stages_table$time)
-#Remove sch.culture from the comparisons for now.
-ordered_stages <- ordered_stages[!ordered_stages == 'sch.culture']
-comparisons <- c(
-  unlist(
-    lapply(seq_along(ordered_stages),
-           function(i){ 
-             if(i < length(ordered_stages)){ 
-               paste0(ordered_stages[i+1],'-',ordered_stages[i]) 
-             } else {
-               #When at the end, want it to cycle back round
-               paste0(ordered_stages[1],'-',ordered_stages[i]) 
-             } 
-           }
-    )
-  )
-)
-```
-
 Compare the gene lists between the same stages in different transmission
@@
                                levels = ordering(pathway.name, 
-                                             order_vector = c('pirs', 'L', 'S', 'AAPL', 'ChAPL')))) %>% 
+                                                 order_vector = c('pirs', 
+                                                                  'L', 
+                                                                  'S', 
+                                                                  'AAPL', 
+                                                                  'ChAPL')))) %>% 
   arrange(pathway.name) %>% 
@@
 }
+```
+
+Compare between stages within the same transmission method.
+
+```{r qusage_transStage}
+ordered_stages <- ordering(unique(lifeCycle_labels),
+         order_vector = times_stages_table$time)
+#Remove sch.culture from the comparisons for now.
+ordered_stages <- ordered_stages[!ordered_stages == 'sch.culture']
+comparisons_transStage.list <- lapply(c('rMT','SBP'), function(transmission){
+  #Getting the comparisons across the time points for each transmission
+  ordered_stages_trans <- str_subset(ordered_stages,
+                               pattern = transmission)
+  comparisons <- c(
+    unlist(
+      lapply(seq_along(ordered_stages_trans),
+             function(i){ 
+               if(i < length(ordered_stages_trans)){ 
+                 paste0(ordered_stages_trans[i+1],'-',ordered_stages_trans[i]) 
+               } else {
+                 #When at the end, want it to cycle back round
+                 paste0(ordered_stages_trans[1],'-',ordered_stages_trans[i]) 
+               } 
+             }
+      )
+    )
+  )
+})
+names(comparisons_transStage.list) <- c('rMT', 'SBP')
+
+# For the life cycle labels then extract the transmission + timepoint label
+lifeCycle_labels <- str_extract(colnames(comp_stagesTrans.mat),
+                                pattern = '^(rMT|SBP)_([[:digit:]]{2}h|sch.culture)')
+
+all_comparisons_transStage <- 
+  lapply(c('rMT', 'SBP'), 
+         function(transmission){
+           #Extract the comparisons for this transmission method
+           comparisons_transStage <- comparisons_transStage.list[[transmission]]
+           #Subset the dataset based on the transmission and removing sch.cultures
+           comp_transStage.mat <- chab_tpm_log2plus1.mat[
+             ,str_detect(colnames(chab_tpm_log2plus1.mat), pattern = paste0(transmission,'_(?!sch.culture)'))
+           ]
+           #Get appropriate life cycle labels from this.
+           lifeCycle_labels <- str_extract(colnames(comp_transStage.mat),
+                                pattern = '^(rMT|SBP)_([[:digit:]]{2}h|sch.culture)')
+           results_transStage.list <- lapply(comparisons_transStage, 
+                  function(comparison){
+                    print(comparison)
+                    list_comp <- lapply(
+                      list(pir_geneset,
+                           ShortLong_list,
+                           subfam_list,
+                           locus_list), 
+                      function(gene_list){
+                        print(deparse(substitute(gene_list)))
+                        qs.results.subfam <- qusage(comp_transStage.mat,
+                                                    lifeCycle_labels,
+                                                    comparison,
+                                                    gene_list,
+                                                    n.points = 2^18)
+                        list(results = qs.results.subfam, 
+                             qstable = qsTable(qs.results.subfam) %>% 
+                               mutate(contrast = qs.results.subfam$contrast))
+                      })
+                    names(list_comp) <- c('pir', 'shortlong', 'subfam', 'locus')
+                    list_comp$combined <- bind_rows(lapply(list_comp, 
+                                                           getElement, 
+                                                           'qstable'))
+                    return(list_comp)
+                  })
+           names(results_transStage.list) <- comparisons_transStage
+           return(results_transStage.list)
+         })
+names(all_comparisons_transStage) <- c('rMT', 'SBP')
+
+all_comparisons_transStage.df <- bind_rows(lapply(all_comparisons_transStage, 
+                                           function(z) bind_rows(
+                                             lapply(z, 
+                                                    function(x) bind_rows(
+                                                      lapply(x, 
+                                                             function(y) y$qstable)))))) %>% 
+  arrange(pathway.name)
 
 ```
+
+```{r heatmap_df_transStage}
+transStages_ht.mat <- filter(all_comparisons_transStage.df, FDR < 0.05) %>% 
+  mutate(pathway.name = factor(pathway.name, 
+                               levels = ordering(pathway.name, 
+                                                 order_vector = c('pirs', 
+                                                                  'L', 
+                                                                  'S', 
+                                                                  'AAPL', 
+                                                                  'ChAPL')))) %>% 
+  arrange(pathway.name) %>% 
+  melt(id.vars = c('contrast', 'pathway.name'),
+       measure.vars  = c('log.fold.change')) %>% 
+  dcast(contrast ~ pathway.name, 
+        value.var = 'value') %>% 
+  column_to_rownames('contrast') %>% 
+  as.matrix
+
+maxcol <- my_max(transStages_ht.mat)
+#Multiply by minus one to use my_max to get the most negative value
+mincol <- my_max(transStages_ht.mat*-1)*-1
+
+#Transpose the matrix so that the stages (comparison) are on the x axis like other figures.
+# Change the colnames of the transposed matrix so that they're just the times.
+# Will split the heatmap by rMT and SBP when drawing the heatmap.
+transStages_ht.mat.t <- t(transStages_ht.mat)
+colnames(transStages_ht.mat.t) <- str_remove_all(colnames(transStages_ht.mat.t),
+                                                 pattern = '(rMT|SBP)_')
+
+ht <- Heatmap(transStages_ht.mat.t,
+        cluster_rows = FALSE,
+        cluster_columns = FALSE,
+        cluster_column_slices = FALSE,
+        na_col = 'white',
+        #Use factors to keep the ordering that we want.
+        row_split = factor(str_extract(colnames(transStages_ht.mat),
+                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'),
+                              levels = unique(str_extract(colnames(transStages_ht.mat),
+                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'))),
+        row_order = colnames(transStages_ht.mat),
+        row_gap = unit(.5,"cm"),
+        row_names_gp = gpar(fontsize = 9),
+        column_split = factor(str_extract(rownames(transStages_ht.mat),
+                                   pattern = 'rMT|SBP'),
+                              levels = unique(str_extract(rownames(transStages_ht.mat),
+                                   pattern = 'rMT|SBP'))),
+        column_gap = unit(.3, "cm"),
+        border = TRUE,
+        #Using this colour scheme because mincol is a lot less negative 
+        # than maxcol is positive
+        col = colorRamp2(c(mincol,
+                           0,
+                           maxcol/4,
+                           maxcol/2,
+                           maxcol),
+                         c('blue', 'white', 'yellow', 'orange', 'red')),
+        heatmap_legend_param = list(title = 'LogFoldChange'))
+draw(ht)
+```

### 2022-08-05 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
qusage rmd title and intro

@@
 ---
-title: "Manuscript_Github_Pchabaudi_boilerplate"
+title: "Manuscript_Github_Pchabaudi_qusage"
 author: "Timothy Little"
@@
 
-Looking at GO enrichment of genes in- and out-of-phase between the 24 asexual life cycles of rMT and SBP transmitted P. chabaudi.
+Qusage analysis of gene set enrichment (e.g. pirs, L-forms, ChAPL loci) between the transmission methods and timepoints of the asexual life cycle.
 
@@
         heatmap_legend_param = list(title = 'LogFoldChange'))
+if(global_save_var){
+  pdf(file = 'plots/stagesWithinTrans_logFoldheatmap.pdf',
+      height = 8,
+      width = 6)
   draw(ht)
+  dev.off()
+}
 ```

### 2022-08-05 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Create compBerghei rmd file

--- 2022-08-02 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_TPMheatmaps.Rmd
@@
 ```
-
-Adding in housekeeping genes for comparison:
-HSP70 putative  PCHAS_0721000
-MSP1 PCHAS_0831300
-AMA1 PCHAS_0931000
-
-```{r highest_expressed_stage_heatmap_fig3b}
-
-housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
-                                 'MSP1' = 'PCHAS_0831300',
-                                 'AMA1' = 'PCHAS_0931000')
-
-cir_housekeeping_id <- c(cir_id, 
-                         'PCHAS_0721000',
-                         'PCHAS_0831300',
-                         'PCHAS_0931000')
-
-# Get order of pir genes by highest expressed genes in each sample.
-
-combined_data_pir_ht <-  filter(chab_tpm_avg.df,
-                             Geneid %in% cir_housekeeping_id) %>% 
-  column_to_rownames(var = 'Geneid') %>% 
-  filter_all(
-    any_vars(. > 10)
-  )
-
-```
-
-```{r heatmap_df}
-#Change the colnames to the times
-colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                              pattern = '.{3}_([[:digit:]]{2}h|sch.culture)')
-
-#Order by time point
-combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                             drop = FALSE]
-
-#Add mean column for ordering
-combined_data_pir_ht$mean_tpm <- apply(combined_data_pir_ht, 1, mean)
-
-combined_data_pir_ht <- arrange(combined_data_pir_ht, desc(mean_tpm))
-
-#Add a column for the subfam (ref for the reference genes)
-combined_data_pir_ht$subfam <- ifelse(rownames(combined_data_pir_ht) %in% cir_id,
-                                      cir_info$subfam_num[match(rownames(combined_data_pir_ht), 
-                                                            cir_info$Geneid)],
-                                      'ref')
-
-#Get the ref gene actual names instead of the Geneid
-rownames(combined_data_pir_ht) <- ifelse(
-  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
-  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
-  rownames(combined_data_pir_ht)
-)
-
-#Get the colour range
-maxcol <- ceiling(
-  max(
-    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'subfam'])
-  ))
-
-#Sub-family annotation
-clade_info <- combined_data_pir_ht$subfam
-annotation_col <- c(clade_col, 'ref' = 'black')
-
-#Subfamily annotation
-
-clade_ha <- rowAnnotation(
-  'sub-family' = anno_simple(
-    clade_info,
-    col = annotation_col,
-  ),
-  annotation_name_gp = gpar(fontsize = 10)
-)
-
-#Subfamily name labels
-num_subfam <- cir_info %>% 
-  group_by(subfam) %>% 
-  summarise(num = n_distinct(Gene.ID))
-
-labels_subfam <- c(setNames(paste0(num_subfam$subfam, ' (n = ', num_subfam$num, ')'),
-                            num_subfam$subfam),
-                   'ref' = 'ref')
-```
-
-```{r draw_heatmap}
-ht.list <- lapply(c('rMT', 'SBP'),
-                  function(transmission){
-
-                    ht.df <- combined_data_pir_ht[
-                      ,
-                      grepl(colnames(combined_data_pir_ht), 
-                            pattern = paste0(transmission))
-                    ]
-
-                    colnames(ht.df) <- str_remove(colnames(ht.df),
-                                                  pattern = paste0(transmission, 
-                                                                   '_'))
-
-                    ht <- Heatmap(
-                      log2(
-                        1+ht.df
-                      ),
-                      cluster_rows = FALSE,
-                      cluster_columns = FALSE,
-                      show_row_names = FALSE,
-                      name = 'log(TPM+1)', 
-                      col = colorRamp2(c(0,
-                                         maxcol/4,
-                                         maxcol/2,
-                                         3*maxcol/4,
-                                         maxcol),
-                                       viridis(5)),
-                      na_col = 'white', 
-                      border = TRUE,
-                      border_gp = gpar(lwd = 0.1),
-                      column_names_gp = gpar(fontsize = 10),
-                      row_dend_width = unit(3, "cm"),
-                      width = unit(4.5, 'cm'),
-                      height = unit(7,'cm'),
-                      column_title = paste(transmission),
-                      row_title_rot = 0,
-                      row_title_gp = gpar(fontsize = 10),
-                      row_split = forcats::fct_shift(
-                        relevel(
-                          as.factor(combined_data_pir_ht$subfam),
-                          ref = 'ref')
-                      ),
-                      heatmap_legend_param = list(
-                        labels_gp = gpar(fontsize = 10),
-                        title_gp = gpar(fontsize = 10,
-                                        fontface = 'bold'),
-                        at = c(0,
-                               maxcol/4,
-                               maxcol/2,
-                               3*maxcol/4,
-                               maxcol)
-                      )
-                    )
-                    return(ht)
-                  })
-
-ht.list <- Reduce('+', ht.list)
-
-clade_lgd <- Legend(
-  title = 'sub-family',
-  at = c(labels_subfam),
-  legend_gp = gpar(fill = annotation_col)
-)
-
-if(global_save_var){
-  pdf(file = 'plots/subfam_TPMheatmap.pdf',
-      height = 8,
-      width = 6)
-  draw(ht.list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'right',
-       merge_legend = TRUE)
-  dev.off()
-}
-```
-
-```{r chapl_aapl_heatmap}
-
-combined_data_pir_ht$locus <- cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), 
-                                    cir_info$Gene.ID)]
-
-#Remove the pirs in the 'NA' locus (i.e. not in a chapl or aapl)
-# don't remove actual NAs because these are the ref genes.
-combined_data_pir_locus_ht <- filter(combined_data_pir_ht,
-                                     !grepl(locus, pattern = "NA \\(n" ))
-
-combined_data_pir_locus_ht$locus <- ifelse(is.na(combined_data_pir_locus_ht$locus),
-                                     'ref',
-                                     combined_data_pir_locus_ht$locus)
-
-#Change locus to factor and change levels so that AAPLs are before ChAPLs and are together.
-combined_data_pir_locus_ht$locus <- factor(combined_data_pir_locus_ht$locus,
-                                           levels = ordering(combined_data_pir_locus_ht$locus,
-                                                             order_vector = c('AAPL',
-                                                                              'ChAPL',
-                                                                              'ref')))
-
-clade_info <- getElement(cir_info,
-                         'subfam')[match(rownames(combined_data_pir_locus_ht), 
-                                                 cir_info$Geneid)]
-clade_info[is.na(clade_info)] <- 'ref'
-names(clade_info) <- rownames(combined_data_pir_locus_ht)
-clade_ha <- rowAnnotation(
-  'subfamily' = anno_simple(
-    clade_info,
-    col = annotation_col
-  )
-)
-#Draw the clade legend for the subfamily
-clade_lgd <- Legend(
-  title = 'sub-family',
-  at = names(annotation_col),
-  legend_gp = gpar(fill = annotation_col)
-)
-
-#Get the colour range
-maxcol <- ceiling(
-  max(
-    log2(1+combined_data_pir_locus_ht[
-      !colnames(combined_data_pir_locus_ht) %in% c('subfam', 
-                                                   'locus')
-    ])
-  ))
-```
-
-```{r draw_locus_heatmap}
-
-ht.list <- lapply(c('rMT', 'SBP'),
-                  function(transmission){
-
-                    #Get either the rMT or the SBP data
-                    ht.df <- combined_data_pir_locus_ht[
-                      ,
-                      grepl(colnames(combined_data_pir_locus_ht), 
-                            pattern = paste0(transmission))
-                    ]
-
-                    #Remove the transmission name from the column names
-                    colnames(ht.df) <- str_remove(colnames(ht.df),
-                                                  pattern = paste0(transmission, 
-                                                                   '_'))
-
-                    ht <- Heatmap(
-                      log2(
-                        1+ht.df
-                      ),
-                      cluster_rows = FALSE,
-                      cluster_columns = FALSE,
-                      show_row_names = FALSE,
-                      name = 'log(TPM+1)', 
-                      col = colorRamp2(c(0,
-                                         maxcol/4,
-                                         maxcol/2,
-                                         3*maxcol/4,
-                                         maxcol),
-                                       viridis(5)),
-                      na_col = 'white', 
-                      border = TRUE,
-                      border_gp = gpar(lwd = 0.1),
-                      column_names_gp = gpar(fontsize = 10),
-                      row_dend_width = unit(3, "cm"),
-                      width = unit(4.5, 'cm'),
-                      height = unit(12,'cm'),
-                      column_title = paste(transmission),
-                      row_title_rot = 0,
-                      row_title_gp = gpar(fontsize = 10),
-                      row_split = forcats::fct_shift(
-                        relevel(
-                          as.factor(combined_data_pir_locus_ht$locus),
-                          ref = 'ref')
-                      ),
-                      heatmap_legend_param = list(
-                        labels_gp = gpar(fontsize = 10),
-                        title_gp = gpar(fontsize = 10,
-                                        fontface = 'bold'),
-                        at = c(0,
-                               maxcol/4,
-                               maxcol/2,
-                               3*maxcol/4,
-                               maxcol)
-                      ),
-                      #Only want the clade annotation on one heatmap not both.
-                      left_annotation = if(transmission == 'rMT'){clade_ha}else{NULL}
-                    )
-                    return(ht)
-                  })
-
-ht.list <- Reduce('+', ht.list)
-
-if(global_save_var){
-  pdf(file = 'plots/ChAPL-AAPL_TPMheatmap.pdf',
-      height = 9,
-      width = 6)
-  draw(ht.list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'right',
-       merge_legend = TRUE,
-       annotation_legend_list = list(clade_lgd))
-  dev.off()
-}
-
-```

### 2022-08-06 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Analyse berghei subfam transcription as well as chab

@@
 ---
-title: "Manuscript_Github_Pchabaudi_boilerplate"
+title: "Manuscript_Github_Pchabaudi_compBerghei"
 author: "Timothy Little"
@@
 
-Looking at GO enrichment of genes in- and out-of-phase between the 24 asexual life cycles of rMT and SBP transmitted P. chabaudi.
+Comparing the P. chabaudi and P. berghei pir expression patterns for any similarities.
 
@@
 cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
-           delim = '\t')
+           delim = '\t') %>% 
+  group_by(subfam) %>% 
+  mutate(subfam_total_members = n_distinct(Geneid)) %>% 
+  mutate(
+    subfam_num = paste0(
+      subfam, 
+      ' (n = ',
+      subfam_total_members,
+      ')'
+    )
+  )
 cir_id <- cir_info$Geneid
@@
     '#CC79A7',
-    '#999999'),
+    'black'),
   sort(
@@
 
-chab_tpm_avg.list <- list(
-  'RMT' = list(
-    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
-                                pattern = 'rMT|Geneid')],
-    experiment = 'RMT'
-  ),
-  'SBP' = list(
-    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
-                                pattern = 'SBP|Geneid')],
-    experiment = 'SBP'
+# chab_tpm_avg.list <- list(
+#   'RMT' = list(
+#     tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
+#                                 pattern = 'rMT|Geneid')],
+#     experiment = 'RMT'
+#   ),
+#   'SBP' = list(
+#     tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
+#                                 pattern = 'SBP|Geneid')],
+#     experiment = 'SBP'
+#   )
+# )
+
+```
+
+```{r berghei_info}
+bir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
+                        delim = '\t') %>% 
+  group_by(clade) %>% 
+  mutate(subfam_total_members = n_distinct(Geneid),
+         subfam_num = paste0(clade, ' (n = ', n_distinct(Geneid), ')'))
+bir_id <- bir_clade$Geneid
+
+bir_clade_col <- setNames(
+  c('#F0E442',
+    '#56B4E9',
+    '#009E73',
+    '#0072B2',
+    '#E69F00',
+    '#D55E00',
+    '#CC79A7',
+    '#999999',
+    'black',
+    'red'),
+  sort(
+   unique(bir_clade$clade)
   )
@@
 
+bir_clade_col_num <- bir_clade_col
+names(bir_clade_col_num) <- sapply(names(bir_clade_col), 
+                               function(clade_name) {
+                                 paste0(
+                                   clade_name, 
+                                   ' (n = ',
+                                   sum(bir_clade$clade == clade_name),
+                                   ')'
+                                 )
+                               })
 ```
+
+```{r berg_tpm}
+berghei_tpm <- read.csv('../chapter3_results_bergheiPir/berghei_tpm.csv', 
+                        row.names = 1)
+```
+
+```{r chab_subfam_normByGeneNum}
+chab_tpm_subfam.df.long <- chab_tpm.df %>% 
+  rownames_to_column('Geneid') %>% 
+  filter(Geneid %in% cir_id) %>% 
+  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
+  mutate(subfam = cir_info$subfam_num[match(Geneid, cir_info$Geneid)],
+         transmission = str_extract(sample, pattern = 'rMT|SBP'),
+         time = str_extract(sample, '[[:digit:]]{2}h|sch.culture'),
+         name = paste0(time, str_extract(sample, 
+                                pattern = '(_[[:digit:]]$)')))
+
+ggplot(chab_tpm_subfam.df.long, 
+       aes(x = factor(name,
+                      levels = ordering(name, 
+                                        order_vector = c(times_stages_table$time, 'sch.culture'))),
+           y = tpm,
+           fill = subfam)) +
+  geom_col(position = "fill") +
+  scale_fill_manual(values = clade_col_num) +
+  facet_wrap(~transmission, scales = "free_x")+
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+Dividing the TPM by the number of genes in each sub-family to see TPM/gene.
+
+```{r normalised_subfam}
+
+subfam_number.table <- group_by(cir_info, subfam_num) %>% 
+  summarise(subfam_total_members = n_distinct(Geneid))
+
+#Don't use n_distinct(Geneid) to nornalise by number of genes in a subfam group
+# There has been some filtering so not all members are in the dataset.
+chab_tpm_subfam_norm.df.long <- chab_tpm_subfam.df.long %>% 
+  group_by(transmission, time, subfam) %>% 
+         mutate(norm_tpm = tpm / cir_info$subfam_total_members[match(Geneid, cir_info$Geneid)])
+
+ggplot(filter(chab_tpm_subfam_norm.df.long, 
+              !subfam == 'ancestral (n = 1)'), 
+       aes(x = factor(name,
+                      levels = ordering(name, 
+                                        order_vector = c(times_stages_table$time, 'sch.culture'))),
+           y = norm_tpm,
+           fill = subfam)) +
+  geom_col(position = "stack") +
+  facet_wrap(~transmission, scales = "free_x") +
+  scale_fill_manual(values = clade_col_num[!names(clade_col_num) == 'ancestral (n = 1)']) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90))
+
+```
+
+```{r berghei_subfam}
+berg_tpm_subfam.df.long <- berghei_tpm[,
+                                       c('Geneid',
+                                         str_subset(colnames(berghei_tpm), 
+                                                    pattern = 'Asex'))] %>% 
+  filter(Geneid %in% bir_id) %>% 
+  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
+  mutate(subfam = bir_clade$subfam_num[match(Geneid, bir_clade$Geneid)],
+         experiment = str_extract(sample, pattern = '^E[[:digit:]]{2}'),
+         stage = str_extract(sample, 'Asex.+(?=_[[:digit:]])'),
+         name = str_extract(sample, 'Asex.+'))
+
+ggplot(berg_tpm_subfam.df.long, 
+       aes(x = factor(name,
+                      levels = ordering(name,
+                                        order_vector = c('Ring','Troph','Schiz','Mixed'))),
+           y = tpm,
+           fill = subfam)) +
+  geom_col(position = "fill") +
+  scale_fill_manual(values = bir_clade_col_num) +
+  facet_wrap(~experiment, scales = "free_x")+
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+```{r berghei_subfam_normByGeneNum}
+
+#Don't use n_distinct(Geneid) to nornalise by number of genes in a subfam group
+# There has been some filtering so not all members are in the dataset.
+berg_tpm_subfam_norm.df.long <- berg_tpm_subfam.df.long %>% 
+  group_by(experiment, stage, subfam) %>% 
+  mutate(norm_tpm = tpm / bir_clade$subfam_total_members[match(Geneid, bir_clade$Geneid)])
+
+ggplot(filter(berg_tpm_subfam_norm.df.long, 
+              !subfam == 'ancestral (n = 1)'), 
+       aes(x = factor(name,
+                      levels = ordering(name, 
+                                        order_vector = c('Ring','Troph','Schiz','Mixed'))),
+           y = norm_tpm,
+           fill = subfam)) +
+  geom_col(position = "stack") +
+  facet_wrap(~experiment, scales = "free_x") +
+  scale_fill_manual(values = bir_clade_col_num[!names(bir_clade_col_num) %in% c('ancestral (n = 1)',
+                                                                                'S6 (n = 1)')]) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90))
+```

### 2022-08-08 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Improve graph labels

@@
            fill = subfam)) +
-  geom_col(position = "stack") +
+  geom_col(position = "fill") +
+    xlab('stages') +
+  ylab('TPM per gene') +
   facet_wrap(~transmission, scales = "free_x") +
@@
          stage = str_extract(sample, 'Asex.+(?=_[[:digit:]])'),
-         name = str_extract(sample, 'Asex.+'))
+         name = str_extract(sample, '(?<=Asex\\.).+'))
 
@@
   geom_col(position = "fill") +
+  xlab('stages') +
   scale_fill_manual(values = bir_clade_col_num) +
@@
            fill = subfam)) +
-  geom_col(position = "stack") +
+  geom_col(position = "fill") +
+  xlab('stages') +
+  ylab('TPM per gene') +
   facet_wrap(~experiment, scales = "free_x") +

### 2022-08-08 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Perform prop.tests on chab samples

@@
 ```
+
+Hypothesis: There is a difference in the transcription level of L-form pirs as a proportion of all S/L-form pirs between Asexual Schizont samples and non-schizont samples
+
+```{r}
+chab_tpm_subfam.L.df.long <- chab_tpm_subfam.df.long %>% 
+  filter(!subfam == 'ancestral (n = 1)') %>% 
+  mutate(subfam_L = grepl(subfam, pattern = '^L')) %>% 
+  group_by(transmission, time, name, sample, subfam_L) %>% 
+  summarise(sum_tpm = sum(tpm)) %>% 
+  group_by(sample) %>% 
+  mutate(sum_tpm_all = sum(sum_tpm)) %>% 
+  filter(subfam_L)
+
+chab_tpm_subfam.L.df.long %>% 
+  group_by(sample) %>% 
+  mutate(prop_L = sum_tpm/sum_tpm_all) %>% 
+  arrange(desc(prop_L))
+
+bind_rows(
+  lapply(str_subset(unique(chab_tpm_subfam.L.df.long$sample), 
+                    pattern = 'rMT(?!_sch.culture)'),
+         function(sample.name){
+           bind_rows(
+             lapply(c('rMT_sch.culture_1', 
+                      'rMT_sch.culture_2', 
+                      'rMT_sch.culture_3'), 
+                    function(sch.cult){
+                      L_forms <- c(chab_tpm_subfam.L.df.long$sum_tpm[
+                        chab_tpm_subfam.L.df.long$sample == sample.name
+                      ],
+                      chab_tpm_subfam.L.df.long$sum_tpm[
+                        chab_tpm_subfam.L.df.long$sample == sch.cult
+                      ])
+                      SorL_forms <- c(chab_tpm_subfam.L.df.long$sum_tpm_all[
+                        chab_tpm_subfam.L.df.long$sample == sample.name
+                      ],
+                      chab_tpm_subfam.L.df.long$sum_tpm_all[
+                        chab_tpm_subfam.L.df.long$sample == sch.cult
+                      ])
+                      test.stat <- prop.test(L_forms, SorL_forms,
+                                             alternative = 'less')
+                      data.frame(comp = paste0(sample.name, '-', sch.cult),
+                                 pval = test.stat$p.value)
+                    })
+           )
+         })
+)%>% 
+  mutate(padj = round(p.adjust(pval), digits = 5),
+         pval = round(pval, digits = 5)) %>% 
+  arrange(desc(padj))
+```
+
+```{r berg_prop_test}
+
+```

### 2022-08-09 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Changed figures to not show ancestral as comparing S vs L proportions

--- 2022-08-08 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
@@
   mutate(subfam = cir_info$subfam_num[match(Geneid, cir_info$Geneid)],
+         SorL = cir_info$SorL[match(Geneid, cir_info$Geneid)],
          transmission = str_extract(sample, pattern = 'rMT|SBP'),
@@
 
-ggplot(chab_tpm_subfam.df.long, 
+ggplot(filter(chab_tpm_subfam.df.long, SorL != 'ancestral'), 
        aes(x = factor(name,
@@
   scale_fill_manual(values = clade_col_num) +
-  facet_wrap(~transmission, scales = "free_x")+
+  facet_grid(~transmission, scales = "free_x") +
+  # facet_grid(transmission ~ SorL, scales = "free")+
   theme_classic()+
-  theme(axis.text.x = element_text(angle = 90))
+  theme(axis.text.x = element_text(angle = 90, size = 3))
 ```
@@
                                          str_subset(colnames(berghei_tpm), 
-                                                    pattern = 'Asex'))] %>% 
+                                                    pattern = 'Asex\\.(?!Mixed)'))] %>% 
   filter(Geneid %in% bir_id) %>% 
@@
 
-ggplot(berg_tpm_subfam.df.long, 
+ggplot(filter(berg_tpm_subfam.df.long, subfam != 'ancestral (n = 1)'), 
        aes(x = factor(name,
@@
 ```
+
+Hypothesis: There is a difference in the transcription level of L-form pirs as a proportion of all S/L-form pirs between Asexual Schizont samples and non-schizont samples
+
+```{r chab_prop_test}
+chab_tpm_subfam.L.df.long <- chab_tpm_subfam.df.long %>% 
+  filter(!subfam == 'ancestral (n = 1)') %>% 
+  mutate(subfam_L = grepl(subfam, pattern = '^L')) %>% 
+  group_by(transmission, time, name, sample, subfam_L) %>% 
+  summarise(sum_tpm = sum(tpm)) %>% 
+  group_by(sample) %>% 
+  mutate(sum_tpm_all = sum(sum_tpm)) %>% 
+  filter(subfam_L)
+
+chab_tpm_subfam.L.df.long %>% 
+  group_by(sample) %>% 
+  mutate(prop_L = sum_tpm/sum_tpm_all) %>% 
+  arrange(desc(prop_L))
+
+bind_rows(
+  lapply(str_subset(unique(chab_tpm_subfam.L.df.long$sample), 
+                    pattern = 'sch.culture',
+                    negate = TRUE),
+         function(sample.name){
+           bind_rows(
+             lapply(c('rMT_sch.culture_1', 
+                      'rMT_sch.culture_2', 
+                      'rMT_sch.culture_3'), 
+                    function(sch.cult){
+                      L_forms <- c(chab_tpm_subfam.L.df.long$sum_tpm[
+                        chab_tpm_subfam.L.df.long$sample == sample.name
+                      ],
+                      chab_tpm_subfam.L.df.long$sum_tpm[
+                        chab_tpm_subfam.L.df.long$sample == sch.cult
+                      ])
+                      SorL_forms <- c(chab_tpm_subfam.L.df.long$sum_tpm_all[
+                        chab_tpm_subfam.L.df.long$sample == sample.name
+                      ],
+                      chab_tpm_subfam.L.df.long$sum_tpm_all[
+                        chab_tpm_subfam.L.df.long$sample == sch.cult
+                      ])
+                      test.stat <- prop.test(L_forms, SorL_forms,
+                                             alternative = 'less')
+                      data.frame(comp = paste0(sample.name, '-', sch.cult),
+                                 pval = test.stat$p.value)
+                    })
+           )
+         })
+)%>% 
+  mutate(padj = round(p.adjust(pval), digits = 5),
+         pval = round(pval, digits = 5)) %>% 
+  arrange(desc(padj))
+```
+
+```{r berg_prop_test}
+berg_tpm_subfam.L.df.long <- berg_tpm_subfam.df.long %>% 
+  filter(!subfam == 'ancestral (n = 1)') %>% 
+  mutate(subfam_L = grepl(subfam, pattern = '^L')) %>% 
+  group_by(experiment, name, sample, subfam_L) %>% 
+  summarise(sum_tpm = sum(tpm)) %>% 
+  group_by(sample) %>% 
+  mutate(sum_tpm_all = sum(sum_tpm)) %>% 
+  filter(subfam_L)
+
+berg_tpm_subfam.L.df.long %>% 
+  group_by(sample) %>% 
+  mutate(prop_L = sum_tpm/sum_tpm_all) %>% 
+  arrange(desc(prop_L))
+
+bind_rows(
+  lapply(str_subset(unique(berg_tpm_subfam.L.df.long$sample), 
+                    pattern = 'Schiz|Mixed',
+                    negate = TRUE),
+         function(sample.name){
+           bind_rows(
+             lapply(str_subset(unique(berg_tpm_subfam.L.df.long$sample), 
+                               pattern = 'Schiz'), 
+                    function(sch.cult){
+                      L_forms <- c(berg_tpm_subfam.L.df.long$sum_tpm[
+                        berg_tpm_subfam.L.df.long$sample == sample.name
+                      ],
+                      berg_tpm_subfam.L.df.long$sum_tpm[
+                        berg_tpm_subfam.L.df.long$sample == sch.cult
+                      ])
+                      SorL_forms <- c(berg_tpm_subfam.L.df.long$sum_tpm_all[
+                        berg_tpm_subfam.L.df.long$sample == sample.name
+                      ],
+                      berg_tpm_subfam.L.df.long$sum_tpm_all[
+                        berg_tpm_subfam.L.df.long$sample == sch.cult
+                      ])
+                      test.stat <- prop.test(L_forms, SorL_forms,
+                                             alternative = 'less')
+                      data.frame(comp = paste0(sample.name, '-', sch.cult),
+                                 pval = test.stat$p.value)
+                    })
+           )
+         })
+)%>% 
+  mutate(padj = round(p.adjust(pval), digits = 5),
+         pval = round(pval, digits = 5)) %>% 
+  arrange(desc(padj))
+```

### 2022-08-10 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Alter figure sizes and swap around schizont and other samples and change prop.test() to greater

@@
     '#009E73',
-    '#0072B2',
+    '#00517f',
     '#E69F00',
-    '#D55E00',
-    '#CC79A7',
+    '#893c00',
+    '#c879cc',
     '#999999',
@@
 
-ggplot(filter(chab_tpm_subfam.df.long, SorL != 'ancestral'), 
+( plot <- ggplot(filter(chab_tpm_subfam.df.long, 
+                        SorL != 'ancestral'), 
                  aes(x = factor(name,
@@
     geom_col(position = "fill") +
-  scale_fill_manual(values = clade_col_num) +
+    xlab('Stages') + ylab('TPM') +
+    scale_fill_manual(values = clade_col_num[!grepl(names(clade_col_num), pattern =  'ancestral')]) +
     facet_grid(~transmission, scales = "free_x") +
@@
     theme_classic()+
-  theme(axis.text.x = element_text(angle = 90, size = 3))
+    theme(axis.text.x = element_text(angle = 90,
+                                     hjust = 1),
+          strip.text = element_text(size = 12),
+          axis.text.y = element_text(size = 10),
+          axis.title = element_text(size = 12))  +
+  guides(fill = guide_legend(title = 'Sub-family')) )
+
+if(global_save_var){
+  ggsave(plot,
+         file = 'plots/chab_subfamProp.png',
+         width = 7,
+         height = 8,
+         units = "in")
+}
 ```
@@
 
-ggplot(filter(berg_tpm_subfam.df.long, subfam != 'ancestral (n = 1)'), 
+( plot <- ggplot(filter(berg_tpm_subfam.df.long, subfam != 'ancestral (n = 1)'), 
                  aes(x = factor(name,
@@
     geom_col(position = "fill") +
+    ylab("TPM") +
     xlab('stages') +
-  scale_fill_manual(values = bir_clade_col_num) +
+    scale_fill_manual(values = bir_clade_col_num[!grepl(names(bir_clade_col_num), 
+                                                        pattern =  'ancestral')]) +
     facet_wrap(~experiment, scales = "free_x")+
     theme_classic()+
-  theme(axis.text.x = element_text(angle = 90))
+    theme(axis.text.x = element_text(angle = 90)) )
+
+if(global_save_var){
+  ggsave(plot,
+         file = 'plots/berghei_subfamProp_AsexExact.png',
+         width = 7,
+         height = 9,
+         units = "in")
+}
 ```
@@
                       L_forms <- c(chab_tpm_subfam.L.df.long$sum_tpm[
-                        chab_tpm_subfam.L.df.long$sample == sample.name
+                        chab_tpm_subfam.L.df.long$sample == sch.cult 
                       ],
                       chab_tpm_subfam.L.df.long$sum_tpm[
-                        chab_tpm_subfam.L.df.long$sample == sch.cult
+                        chab_tpm_subfam.L.df.long$sample == sample.name
                       ])
                       SorL_forms <- c(chab_tpm_subfam.L.df.long$sum_tpm_all[
-                        chab_tpm_subfam.L.df.long$sample == sample.name
+                        chab_tpm_subfam.L.df.long$sample == sch.cult 
                       ],
                       chab_tpm_subfam.L.df.long$sum_tpm_all[
-                        chab_tpm_subfam.L.df.long$sample == sch.cult
+                        chab_tpm_subfam.L.df.long$sample == sample.name
                       ])
                       test.stat <- prop.test(L_forms, SorL_forms,
-                                             alternative = 'less')
+                                             alternative = 'greater')
                       data.frame(comp = paste0(sample.name, '-', sch.cult),
@@
                       L_forms <- c(berg_tpm_subfam.L.df.long$sum_tpm[
-                        berg_tpm_subfam.L.df.long$sample == sample.name
+                        berg_tpm_subfam.L.df.long$sample == sch.cult 
                       ],
                       berg_tpm_subfam.L.df.long$sum_tpm[
-                        berg_tpm_subfam.L.df.long$sample == sch.cult
+                        berg_tpm_subfam.L.df.long$sample == sample.name
                       ])
                       SorL_forms <- c(berg_tpm_subfam.L.df.long$sum_tpm_all[
-                        berg_tpm_subfam.L.df.long$sample == sample.name
+                        berg_tpm_subfam.L.df.long$sample == sch.cult 
                       ],
                       berg_tpm_subfam.L.df.long$sum_tpm_all[
-                        berg_tpm_subfam.L.df.long$sample == sch.cult
+                        berg_tpm_subfam.L.df.long$sample == sample.name
                       ])
                       test.stat <- prop.test(L_forms, SorL_forms,
-                                             alternative = 'less')
+                                             alternative = 'greater')
                       data.frame(comp = paste0(sample.name, '-', sch.cult),

### 2022-08-26 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Added a TPM filter to the stacked bar chart

@@
 ```{r all_pir_barchart}
+#Filter for pirs above 1 TPM in at least one sample
 cir_plot <- chab_tpm.df %>% 
-  filter(Geneid %in% cir_id) %>% 
+  filter(Geneid %in% cir_id,
+         if_any(!matches('Geneid'), ~ . > 1)) %>% 
   melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 

### 2022-08-26 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Alter plots of berg comparison subfam proportions so that the graphs have same font size between each and don't show gene stacks

@@
 chab_tpm_subfam.df.long <- chab_tpm.df %>% 
-  rownames_to_column('Geneid') %>% 
+  # rownames_to_column('Geneid') %>% 
   filter(Geneid %in% cir_id) %>% 
@@
          name = paste0(time, str_extract(sample, 
-                                         pattern = '(_[[:digit:]]$)')))
+                                         pattern = '(_[[:digit:]]$)'))) %>% 
+  group_by(subfam, SorL, transmission, time, name) %>% summarise(tpm = sum(tpm))
 
@@
          stage = str_extract(sample, 'Asex.+(?=_[[:digit:]])'),
-         name = str_extract(sample, '(?<=Asex\\.).+'))
+         name = str_extract(sample, '(?<=Asex\\.).+')) %>% 
+  group_by(subfam, experiment, stage, name) %>% summarise(tpm = sum(tpm))
 
@@
     ylab("TPM") +
-    xlab('stages') +
+    xlab('Stages') +
     scale_fill_manual(values = bir_clade_col_num[!grepl(names(bir_clade_col_num), 
@@
     theme_classic()+
-    theme(axis.text.x = element_text(angle = 90)) )
+    theme(axis.text.x = element_text(angle = 90,
+                                     hjust = 1),
+          strip.text = element_text(size = 12),
+          axis.text.y = element_text(size = 10),
+          axis.title = element_text(size = 12))
+)
 
@@
          width = 7,
-         height = 9,
+         height = 8,
          units = "in")

### 2022-08-26 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Looking at sr10 transcription

@@
   ggsave(filename = 'plots/SBP-hi_Pir_BarChart.pdf',
+          plot = p,
+          dpi = 300,
+          width = 15,
+          height = 10,
+          units = 'cm')
+}
+```
+
+```{r}
+ cir_plot <- chab_tpm.df %>% 
+    filter(Geneid %in% "PCHAS_1433600") %>% 
+    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
+    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
+    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
+           transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
+    group_by(sample, time, transmission) %>% 
+    dplyr::summarise(sum_tpm = sum(tpm))
+
+y_scale <- seq(0, 
+               signif(ceiling(max(cir_plot$sum_tpm)), 
+                      digits = 2), 
+               length.out = 5)
+
+(
+  p <- ggplot(data = cir_plot, 
+              aes(x = time, 
+                  y = sum_tpm,
+                  fill = transmission)) +
+    stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
+                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
+    scale_y_continuous(breaks = y_scale) +
+    scale_fill_manual(values = cbPalette) +
+    geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
+    ylab(as.expression(bquote('Total '~italic(.('SR10'))~'TPM'))) +
+    xlab('Asexual blood stage development cycle') +
+    theme_classic() +
+    theme(axis.text.x=element_text(angle = 90, 
+                                   vjust = 0.5,
+                                   size = 12))
+)
+
+if(global_save_var){
+  ggsave(filename = 'plots/SR10_BarChart.pdf',
           plot = p,

### 2022-08-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
Clean-up

@@
 
-cir_info <- read_delim('../standard_files/Pchabaudi_cir_info.txt',
+cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
            delim = '\t')
@@
                               function(gene_list){
-                                print(deparse(substitute(gene_list)))
                                 qs.results.subfam <- qusage(comp_stagesTrans.mat,
@@
                           })
-names(all_comparisons) <- comparisons
+names(all_comparisons) <- comparisons_stagesTrans
 

### 2022-08-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
Look at L4s ChAPL and AAPLs separate

@@
 
+cir_info <- mutate(cir_info, 
+                   subfam_loci = paste(subfam,
+                                       chapl_aapl, 
+                                       sep = '_'))
+L4_loci_list <- lapply(c('L4_ChAPL', 'L4_AAPL'), 
+                  function(clade_nam) filter(cir_info, 
+                                             subfam_loci == clade_nam)$Geneid)
+names(L4_loci_list) <- c('L4_ChAPL', 'L4_AAPL')
+L4_loci_list <- L4_loci_list[unlist(lapply(L4_loci_list, 
+                                         function(x) length(x) > 3))]
+
 ShortLong_list <- 
@@
                                    subfam_list,
+                                   L4_loci_list,
                                    locus_list), 
@@
                            subfam_list,
+                           L4_loci_list,
                            locus_list), 
                       function(gene_list){
-                        print(deparse(substitute(gene_list)))
                         qs.results.subfam <- qusage(comp_transStage.mat,

### 2022-09-27 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Change PCA plots to have capitalised RMT and colour blind colour scheme

@@
   dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
+  mutate( transmission = toupper(str_split_fixed(stage, pattern = '_', n = 2)[,1]),
           time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture'))
 
-colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
+# colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
+colornames <- c(cbPalette, 'black')
+# colornames <- c('#FEFE62', '#D35FB7','#FFC20A', '#0C7BDC',
+#                 '#1AFF1A', '#4B0092', '#E66100', '#5D3A9B',
+#                 '#D41159')
 colornames <- setNames(colornames, 
@@
 
-# (
-#   plot <- ggplot(chab_tpm.df_pca_coords,
-#                  aes(
-#                    x = PC3,
-#                    y = PC1,
-#                    col = time,
-#                    shape = transmission
-#                  )) +
-#     geom_point(size = 3,
-#                alpha = 0.8) +
-#     scale_color_manual(values = colornames) +
-#     xlab(paste0('PC3 ', round(pca_importance$PC3[2]*100, digits = 1),'%')) +
-#     ylab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-#     guides(shape=guide_legend(ncol=2),
-#            col = guide_legend(ncol=2)) +
-#     theme_classic() +
-#     theme(legend.key.size = unit(4, 'mm'),
-#           legend.text = element_text(size = 7))
-# )
-
 if(global_save_var){
@@
   dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
+  mutate( transmission = toupper(str_split_fixed(stage, pattern = '_', n = 2)[,1]),
           time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture'))
 
-colornames <- viridis(length(unique(chab_tpm_wPseudo.df_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(chab_tpm_wPseudo.df_pca_coords$time)))
+# colornames <- viridis(length(unique(chab_tpm_wPseudo.df_pca_coords$time)))
+# colornames <- setNames(colornames, 
+#                        sort(unique(chab_tpm_wPseudo.df_pca_coords$time)))
 

### 2022-10-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Add Chapter 5 vivax de novo results folder

@@
   dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
+  mutate( transmission = toupper(str_split_fixed(stage, pattern = '_', n = 2)[,1]),
           time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture')) 
-
-colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(chab_tpm.df_pca_coords$time)))
+# 
+# colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
+# colornames <- setNames(colornames, 
+#                        sort(unique(chab_tpm.df_pca_coords$time)))
 
@@
     mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
-           transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
+           transmission = toupper(str_split_fixed(sample, pattern = '_', n = 2)[,1])) %>% 
     group_by(sample, time, transmission) %>% 
@@
   mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
-         transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1],
+         transmission = toupper(str_split_fixed(sample, pattern = '_', n = 2)[,1]),
          SorL = cir_clade_info$SorL[match(Geneid, cir_clade_info$Geneid)]) %>% 
@@
 # lists of the zscore dataframes and the TPM lists from above.
-chab_z.list <- list('RMT' = c(chab_tpm_avg.list[['RMT']],
-                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
-                                                           pattern = 'Geneid|rMT')])),
+# Also rename the RMT dataframe columns to change 'rMT' to 'RMT'
+chab_z.list <- list('RMT' = list('experiment' = chab_tpm_avg.list[['RMT']]$experiment,
+                                 'tpm' = chab_tpm_avg.list[['RMT']]$tpm %>% 
+                                   rename_with(~gsub('rMT', 'RMT', .x)),
+                                 'zscores' = chab_z.df[,grepl(colnames(chab_z.df),
+                                                              pattern = 'Geneid|rMT')] %>% 
+                                   rename_with(~gsub('rMT', 'RMT', .x))),
                     'SBP' = c(chab_tpm_avg.list[['SBP']],

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxBUSCO.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_VvxBUSCO"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) 
# Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr', 
        lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0")
library('stringr')
library('purrr')
library('kableExtra')
library('plotly')
library('rmarkdown')
library('seqinr')
library(plotly)
```

```{r analyse_busco}
samplesheet <- read.delim("/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samplesheet.txt")

samplesheet_busco.df <- lapply(unique(samplesheet$assembly), 
                               function(assembly.file){
                                 # The tenth element in the file path string corresponds to the assembly names that I used.
                                 assembly <- strsplit(assembly.file, split = '/')[[1]][10]
                                 if(!is.na(assembly)){
                                   busco.file <- paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/busco/',
                                                        assembly)
                                   #Skip 2 in the read_tsv function as the first two lines are not part of the data frame.
                                   busco_api.df <- read_tsv(paste0(busco.file, 
                                                                   '_apicomplex/run_apicomplexa_odb10/full_table.tsv'),
                                                            skip = 2,
                                                            show_col_types = FALSE)
                                   busco_api.complete <- sum(busco_api.df$Status == "Complete")
                                   busco_api.fragment <- sum(busco_api.df$Status == "Fragmented")
                                   busco_api.duplicate <- sum(n_distinct(
                                     busco_api.df$`# Busco id`[busco_api.df$Status == "Duplicated"]
                                   ))

                                   busco_plasmo.df <- read_tsv(paste0(busco.file, 
                                                                      '_plasmodium/run_plasmodium_odb10/full_table.tsv'),
                                                               skip = 2,
                                                               show_col_types = FALSE)
                                   busco_plasmo.complete <- sum(busco_plasmo.df$Status == "Complete")
                                   busco_plasmo.fragment <- sum(busco_plasmo.df$Status == "Fragmented")
                                   busco_plasmo.duplicate <- sum(n_distinct(
                                     busco_plasmo.df$`# Busco id`[busco_plasmo.df$Status == "Duplicated"]
                                   ))

                                   data.frame(
                                     assembly = assembly.file,
                                     assembly_name = assembly,
                                     Experiment = samplesheet$Experiment[match(assembly.file, samplesheet$assembly)],
                                     num_busco_api_complete = busco_api.complete,
                                     perc_busco_api_complete = round(100*busco_api.complete/446,
                                                                     digits = 3),
                                     num_busco_api_fragment = busco_api.fragment,
                                     perc_busco_api_fragment = round(100*busco_api.fragment/446,
                                                                     digits = 3),
                                     num_busco_api_duplicate = busco_api.duplicate,
                                     perc_busco_api_duplicate = round(100*busco_api.duplicate/446,
                                                                      digits = 3),
                                     perc_busco_api_nonMissing = round(100*(
                                       busco_api.fragment+busco_api.complete+busco_api.duplicate)/446,
                                       digits = 3
                                     ),

                                     num_busco_plasmo_complete = busco_plasmo.complete,
                                     perc_busco_plasmo_complete = round(100*busco_plasmo.complete/3642,
                                                                        digits = 3),
                                     num_busco_plasmo_fragment = busco_plasmo.fragment,
                                     perc_busco_plasmo_fragment = round(100*busco_plasmo.fragment/3642,
                                                                        digits = 3),
                                     num_busco_plasmo_duplicate = busco_plasmo.duplicate,
                                     perc_busco_plasmo_duplicate = round(100*busco_plasmo.duplicate/3642,
                                                                         digits = 3),
                                     perc_busco_plasmo_nonMissing = round(100*(
                                       busco_plasmo.fragment+busco_plasmo.complete+busco_plasmo.duplicate)/3642,
                                       digits = 3
                                     )
                                   )
                                 } else { return(NULL) }
                               }) %>% bind_rows
```

```{r plotly}

#Get the number of pirs stats
assembly_stats.df <- read.table('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info_extra.txt')

colours_val <- unique(assembly_stats.df$expt_colour)
names(colours_val) <- unique(assembly_stats.df$experiment_num)
colours_val <- colours_val[match(sort(unique(assembly_stats.df$experiment_num)),
                                 names(colours_val))]

assembly_stats_busco <- inner_join(assembly_stats.df, 
                                   mutate(samplesheet_busco.df, 
                                          assembly_file = assembly), 
                                   by = 'assembly_file')
# write.table(assembly_stats_busco,
#             '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info_extra_busco.txt')

#Include '.+(?=0.95)' pattern for searching as Gunalan has multiple cdhitest threshold assemblies.
samplesheet_busco.df <- mutate(rowwise(samplesheet_busco.df), 
                               num_pir_predict = assembly_stats.df$num_uniq_mRNA_ORFs_over200aaLength[
                                 str_detect(assembly_stats.df$assembly, pattern = paste0(assembly_name, 
                                                                                         '.+(?=0.95)'))
                               ],
                               expt_num = unique(
                                 assembly_stats.df$experiment_num[
                                   str_detect(assembly_stats.df$assembly, pattern = paste0(assembly_name, 
                                                                                           '.+(?=0.95)'))
                                 ]),
                               num_transcripts = unique(
                                 assembly_stats.df$num_transcripts[
                                   str_detect(assembly_stats.df$assembly, pattern = paste0(assembly_name, 
                                                                                           '.+(?=0.95)'))
                                 ]),
                               num_mRNA_pir_over200aa_over1TPM = unique(
                                 assembly_stats.df$num_mRNA_pir_over200aa_over1TPM[
                                   str_detect(assembly_stats.df$assembly, pattern = paste0(assembly_name, 
                                                                                           '.+(?=0.95)'))
                                 ])
)

hline <- function(y = 0, color = "blue") {
  list(
    type = "line", 
    x0 = 0, 
    x1 = 1, 
    xref = "paper",
    y0 = y, 
    y1 = y, 
    line = list(color = 'black')
  )
}
vline <- function(x = 0, color = "red") {
  list(
    type = "line", 
    y0 = 0, 
    y1 = 1, 
    yref = "paper",
    x0 = x, 
    x1 = x, 
    line = list(color = color)
  )
}

plot_ly(samplesheet_busco.df,
        x = ~perc_busco_api_complete,
        y = ~perc_busco_api_nonMissing,
        text = ~assembly_name,
        color = ~Experiment,
        colors = rainbow(10),
        mode = 'markers') %>% 
  layout(shapes = list(hline(50)))

plot_ly(samplesheet_busco.df,
        x = ~perc_busco_plasmo_complete,
        y = ~perc_busco_plasmo_nonMissing,
        text = ~assembly_name,
        color = ~Experiment,
        colors = rainbow(10),
        mode = 'markers') %>% 
  layout(shapes = list(hline(50)))

plot_ly(samplesheet_busco.df,
        y = ~perc_busco_api_nonMissing,
        x = ~perc_busco_plasmo_nonMissing,
        text = ~assembly_name,
        color = ~expt_num,
        colors = colours_val,
        size = ~num_transcripts,
        # sizes = c(10,100),
        mode = 'markers') %>% 
  layout(shapes = list(hline(25), vline(25)))

plot_ly(samplesheet_busco.df,
        y = ~num_pir_predict,
        x = ~perc_busco_plasmo_nonMissing,
        text = ~assembly_name,
        color = ~expt_num,
        colors = colours_val,
        mode = 'markers')

```

```{r ggplot_busco_graph}

( plot_buscos <- ggplot(samplesheet_busco.df,
                      aes(x = perc_busco_plasmo_nonMissing,
                          y = perc_busco_api_nonMissing,
                          col = factor(expt_num, unique(sort(expt_num))),
                          size = num_pir_predict)) +
  geom_point(alpha = 0.75) +
  theme_classic() +
    theme(axis.text.x = element_text(size = 10),
          axis.text.y = element_text(size = 10),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12)) +
  ylim(0,100) +
  scale_color_manual(values = colours_val,
                     name = 'Experiment') +
    guides(color = guide_legend(override.aes = list(size=5),
                             ncol = 2)) +
  scale_size_continuous(name = expression(paste('Num predicted ',italic(pir),'s'))) +
  xlab('Percentage of Plasmodium BUSCOs') +
  ylab('Percentage of Apicomplexan BUSCOs') )

ggsave(plot = plot_buscos,
       '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/assemblies_buscoPlasmo-buscoApi-pirNum.pdf',
       device = 'pdf',
       width = 7,
       height = 7,
       units = "in")

ggplot(samplesheet_busco.df,
       aes(x = perc_busco_plasmo_nonMissing,
           y = num_mRNA_pir_over200aa_over1TPM,
           col = factor(expt_num, unique(sort(expt_num))),
           size = num_pir_predict)) +
  geom_point(alpha = 0.75) +
  theme_classic() +
  scale_color_manual(values = colours_val,
                     name = 'Experiment') +
  scale_size_continuous(name = 'Number of predicted pirs') +
  xlab('Percentage of Plasmodium BUSCOs') +
  ylab('Number of pir transcripts')

ggplot(samplesheet_busco.df,
       aes(x = perc_busco_plasmo_nonMissing,
           y = num_transcripts,
           col = factor(expt_num, unique(sort(expt_num))),
           size = num_pir_predict)) +
  geom_point(alpha = 0.75) +
  theme_classic() +
  scale_color_manual(values = colours_val,
                     name = 'Experiment') +
  scale_size_continuous(name = 'Number of predicted pirs') +
  xlab('Percentage of Plasmodium BUSCOs') +
  ylab('Number of transcripts')

```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_VivaxPirsInTheDeNovoAssemblies"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_blastn.outfmt6'
---

# Intro

This script is looks at the BLAST results of a Trinity _de novo_ assembled transcriptome.

# What this report shows:

*The level of counts / TPM of pirs that are detected by the Trinity de novo assembled transcriptome.
**e.g. What is the minimum expressed gene detected / What is the maximally expressed gene not detected.

# To-do:
*Check how they BLAST to the rest of the transcriptome, do pir de novo transcripts also BLAST to non-pir transcripts?
*Select the most informative data and poo-poo the rest.
*Save useful files made by this script as csv.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)

#Possible files for blast_output_file:
# '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_blastn.outfmt6'
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr', 
        lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0")
library('stringr')
library('purrr')
library('IRanges')
library('kableExtra')
library('plotly')
```

```{r function_ordering}
ordering <- function(to_order, order_vector, fromLast = TRUE){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = fromLast)
}
```

```{r function_plotRanges}
#' Plot genomic ranges from the IRanges object or GRanges object.
# from https://rdrr.io/github/JokingHero/rxtras/src/R/plot_wrappers.R

plotRanges <- function(x, xlim = x, main = deparse(substitute(x)),  col = "black", sep = 0.5, ...){
  if (!requireNamespace("IRanges", quietly = TRUE)) {
    stop("IRanges needed for this function to work. Please install it.",
         call. = FALSE)
  }
  height <- 1
  if (is(xlim, "Ranges"))
      xlim <- c(min(start(xlim)), max(end(xlim)))
  bins <- disjointBins(IRanges(start(x), end(x) + 1))
  plot.new()
  plot.window(xlim, c(0, max(bins) * (height + sep)))
  ybottom <- bins * (sep + height) - height
  rect(start(x) - 0.5, ybottom, end(x) + 0.5, ybottom + height, col = col, ...)
  title(main)
  axis(1)
}
```

```{r load_data}
pir.info <- read.csv('PlasmoDBv56_VvxPirs_P01_Sal1.csv', 
                     header = TRUE)
```

```{r read_blast_results}
# list.files('../blast', pattern = 'outfmt')

#Set a filter for bp length to assume is a valid match
filter_nident <- 100

outfmt6_headers <- 
  'qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore qcovs qcovhsp slen nident' %>% 
  strsplit(split = ' ') %>% 
  unlist

blast <- read.table(params$blast_output_file,
                    col.names = outfmt6_headers) %>% 
  mutate(
         sseqid = unlist(
           lapply(
             str_extract_all(sseqid, 
                             pattern = '.+(?=\\.1)'), 
             paste, 
             collapse = '_')
         ),
         scovhsp = round( (nident/slen * 100), digits = 2),
         scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
         evalue = format(evalue, scientific = TRUE)
  ) %>% 
  #Filter for the pir genes.
  filter(nident > filter_nident,
         sseqid %in% pir.info$Gene.ID) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

# Filter for the top hits for each de novo transcript by e-value.

blast_top <- group_by(blast, qseqid) %>% 
  slice_min(evalue, n = 1)

#Filter for the de novo transcripts which only match a single known pir.
# Also removing idents below 100 because these are probably poor matches.

blast_uniq <- filter(blast, num_genes_match == 1)

#sanity test
# all(blast$scovs >= blast$scovhsp)
```

Filtering matches by a minimum of `r filter_nident`bp to be a valid match.

# Number of hits for each de novo transcript with a BLAST hit.

Are there many de novo transcripts which match multiple different known transcripts? Are most of them unique?

```{r number_of_hits_per_deNovoTranscript}
blast %>% 
  group_by(qseqid) %>% 
  summarise(num = n_distinct(sseqid)) %>% 
  .$num %>% 
  hist(main = 'Histogram of BLAST hits per de novo transcript (50 max)',
       breaks = 50)
```

#Genes with the most number of uniquely mapping de novo transcripts.

```{r number_of_uniq_transcripts_per_gene}
blast %>% 
  filter(num_genes_match== 1) %>% 
  group_by(sseqid) %>% 
  summarise(num_of_uniq_deNovoTranscripts = n_distinct(qseqid),
            max_id = max(pident),
            min_id = min(pident),
            max_nident = max(nident),
            min_nident = min(nident),
            max_gap = max(gapopen),
            min_gap = min(gapopen),
            max_qcovh = max(qcovhsp),
            min_qcovh = min(qcovhsp),
            max_scovh = max(scovhsp),
            min_scovh = min(scovhsp)) %>% 
  arrange(desc(num_of_uniq_deNovoTranscripts)) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "400px")
```

# Coverage of the subject (known transcripts) and query (assembled de novo transcripts) sequences.

Using only the de novo transcripts that uniquely match one known pir transcript.

For our purposes the subject coverage is probably most important.

```{r scov_v_qcov}
fig <- plot_ly(
  blast_uniq,
  x = ~qcovhsp,
  y = ~scovhsp,
  text = ~paste(sseqid)
) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
fig
```

Number of unique genes found by BLAST: `r unique(blast$sseqid) %>% length`.

# Known transcripts with >95% coverage from one de novo transcript.

How many of the known transcripts are covered >=95% by de novo transcripts?

num_match_all and num_match_95 show how many other known transcripts these de novo transcripts match in total or >=95% respectively.

```{r transcripts_covering_most_of_known_transcripts}
transcripts_covered_by_denovo <- blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>%
  group_by(sseqid) %>% 
  filter(scovhsp >= 95) %>% 
  group_by(qseqid) %>%
  mutate(num_match_95 = n_distinct(sseqid)) %>% 
  arrange(sseqid)
transcripts_covered_by_denovo %>% 
  select(-qstart,-qend,-sstart,-send,-evalue,-bitscore,-qcovs,-slen,-sseqid,-nident,-scovs) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
```

Showing the BLAST alignments of these >=95% coverage de novo transcripts in a graph, and all the matches they have. 
The "n =" number after the transcript name shows how many other BLAST hits the de novo transcript had in total.
Size of each point shows the length of the alignment.

```{r transcripts_covering_most_of_known_transcripts_graph}
blast %>% 
  filter(qseqid %in% unique(transcripts_covered_by_denovo$qseqid)) %>%
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'bp', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

Alignments with >=95% known coverage but <50% de novo transcript coverage.

```{r low_qcovhsp}
transcripts_covered_by_denovo %>% 
  filter(qcovhsp < 50) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
```

# Transcripts matched by percentage coverage.

```{r transcript_coverage_by_percentage}

transcripts_covered_by_denovo_diffcov <- lapply(seq(0,100,by = 5),
                                                function(covrg) {
                                                  blast %>% 
                                                    group_by(sseqid) %>% 
                                                    filter(scovhsp >= covrg) %>%
                                                    .$sseqid %>% unique %>% length
                                                })
names(transcripts_covered_by_denovo_diffcov) <- seq(0,100,by = 5)
plot(y = unlist(transcripts_covered_by_denovo_diffcov), 
     x = seq(0,100,by = 5),
     ylim = c(0,nrow(pir.info)),
     main = 'For all de novo transcripts',
     ylab = 'Number of known transcripts covered',
     xlab = 'Percentage of coverage by individual transcripts',
     yaxt = 'none')
axis(2, at = seq(0,nrow(pir.info),by = 100))
# abline(a = 25, b = 0)

```

# Important instant details:

Filtered matches by a minimum of `r filter_nident`bp to be a valid match.

There are `r length(unique(transcripts_covered_by_denovo$sseqid))` known transcripts covered by >=95% of at least one individual _de novo_ transcript. There is a mean of `r mean(transcripts_covered_by_denovo$mismatch)` mismatches and `r mean(transcripts_covered_by_denovo$gapopen)` gaps and a median of `r median(transcripts_covered_by_denovo$mismatch)` mismatches and `r median(transcripts_covered_by_denovo$gapopen)` gaps in these alignments (Note that there are probably some de novo transcripts that align >=95% to the same known transcript). Of these `r length(unique(transcripts_covered_by_denovo$sseqid[transcripts_covered_by_denovo$num_match_all == 1]))`
 match only that known transcript and no others by the BLAST search. `r length(unique(filter(ungroup(transcripts_covered_by_denovo), qcovhsp < 50)$sseqid))` of the genes with >=95% known transcript alignments cover less than 50% of the de novo transcript.

See section [Known transcripts with >95% coverage from one de novo transcript.].

 With a threshold of 75% coverage `r transcripts_covered_by_denovo_diffcov['75']` known transcripts are represented by the de novo transcripts.

See section [Transcripts matched by percentage coverage.].

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis_SRX697844_SRR1571714_cdhit0.99_blastn.Rmd
Add Chapter 5 vivax de novo results folder

--- 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis.Rmd
@@
 ---
-title: "tl22-01_VivaxPirsInTheDeNovoAssemblies"
+title: "tl22-01_VivaxPirsInTheDeNovoAssemblies SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6"
 output: 
@@
 params:
-  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_blastn.outfmt6'
+  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6'
 ---
@@
 ```{r low_qcovhsp}
+if(transcripts_covered_by_denovo %>% 
+   filter(qcovhsp < 50) %>% 
+   nrow() > 0){
   transcripts_covered_by_denovo %>% 
@@
     kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
+}
+```
+
+# All the BLAST matches
+
+```{r transcripts_coverage_allHits_transcripts_graph}
+blast %>% 
+  group_by(qseqid) %>%
+  mutate(num_match_all = n_distinct(sseqid)) %>% 
+  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
+         over95 = ifelse(scovhsp >= 95, 
+                         '>=95% coverage', 
+                         '<95% coverage')) %>% 
+  plot_ly(x = ~qcovhsp,
+          y = ~scovhsp,
+          text = ~paste(sseqid, '\n', 
+                        length, 'aa', '\n',
+                        scovhsp, '% coverage of known'),
+          color = ~qseqid,
+          size = ~length,
+          symbol = ~over95,
+          symbols = c('triangle-up', 'circle'),
+          sizes = c(2,100),
+          legendgroup = ~qseqid) %>% 
+  layout(xaxis = list(title = "De novo transcript coverage"),
+         yaxis = list(title = "Known transcript coverage"))
+```
+
+# BLAST hit coverages for the top alignments.
+
+```{r transcripts_coverage_topBLAST_transcripts_graph}
+blast %>% 
+  group_by(qseqid) %>%
+  mutate(num_match_all = n_distinct(sseqid)) %>% 
+  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
+         over95 = ifelse(scovhsp >= 95, 
+                         '>=95% coverage', 
+                         '<95% coverage')) %>% 
+  group_by(sseqid) %>% 
+  slice_min(evalue, n = 1) %>% 
+  plot_ly(x = ~qcovhsp,
+          y = ~scovhsp,
+          text = ~paste(sseqid, '\n', 
+                        length, 'aa', '\n',
+                        scovhsp, '% coverage of known'),
+          color = ~qseqid,
+          size = ~length,
+          symbol = ~over95,
+          symbols = c('triangle-up', 'circle'),
+          sizes = c(2,100),
+          legendgroup = ~qseqid) %>% 
+  layout(xaxis = list(title = "De novo transcript coverage"),
+         yaxis = list(title = "Known transcript coverage"))
 ```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis_SRX697844_SRR1571714_scspades_blastpP01-Zhu16-Sal1vir.Rmd
Add Chapter 5 vivax de novo results folder

--- 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis_SRX697844_SRR1571714_cdhit0.99_blastn.Rmd
@@
 ---
-title: "tl22-01_VivaxPirsInTheDeNovoAssemblies SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6"
+title: "tl22-01_VivaxPirsInTheDeNovoAssemblies SRX697844_SRR1571714_getorfFind1_blastp-P01-Zhu16_Sal1vir"
 output: 
@@
 params:
-  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6'
+  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades_blastpP01-Zhu16-Sal1vir/SRX697844_SRR1571714/SRX697844_SRR1571714_scspades_blastpP01-Zhu16-Sal1vir_hmmerPfam-Plasmodium_Vir_blastp-P01-all_Sal1-vir_Zhu2016Vir.outfmt6'
 ---
@@
 
-This script is looks at the BLAST results of a Trinity _de novo_ assembled transcriptome.
+This script is looks at the BLAST results of the HMMer matches with the Plasmodium_Vir among the ORFs from the Plasmodium vivax RNAseq scspades assembly.
+BLASTed to the Plasmodium vivax P01 PlasmoDB v56 proteins, the Zhu 2016 de novo mRNA vir predicted getorfs, and the Sal1 virs.
 
@@
 
-*The level of counts / TPM of pirs that are detected by the Trinity de novo assembled transcriptome.
-**e.g. What is the minimum expressed gene detected / What is the maximally expressed gene not detected.
-
 # To-do:
-*Check how they BLAST to the rest of the transcriptome, do pir de novo transcripts also BLAST to non-pir transcripts?
-*Select the most informative data and poo-poo the rest.
-*Save useful files made by this script as csv.
 
@@
   mutate(
-         sseqid = unlist(
-           lapply(
-             str_extract_all(sseqid, 
-                             pattern = '.+(?=\\.1)'), 
-             paste, 
-             collapse = '_')
-         ),
+
          scovhsp = round( (nident/slen * 100), digits = 2),
@@
   #Filter for the pir genes.
+  # The Sal1 and Zhu genes are already only the virs so I grep for them first in ifelse.
+  # Then, for the P01, I extract the gene name and search in the pir.info list.
   filter(nident > filter_nident,
-         sseqid %in% pir.info$Gene.ID) %>%
+         ifelse(grepl(sseqid, pattern = 'PVX|comp'),
+                TRUE,
+                str_extract_all(sseqid, 
+                                pattern = '.+(?=\\.1)') %in% pir.info$Gene.ID)) %>%
   group_by(qseqid) %>% 
@@
 
+# BLAST hit coverages for the top alignments to the Zhu 2016 de novo virs.
+
+```{r transcripts_coverage_topBLAST_Zhu16transcripts_graph}
+blast %>% 
+  group_by(qseqid) %>%
+  mutate(num_match_all = n_distinct(sseqid)) %>% 
+  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
+         over95 = ifelse(scovhsp >= 95, 
+                         '>=95% coverage', 
+                         '<95% coverage')) %>% 
+  filter(grepl(sseqid, 
+               pattern = 'comp')) %>% 
+  group_by(sseqid) %>% 
+  slice_min(evalue, n = 1) %>% 
+  plot_ly(x = ~qcovhsp,
+          y = ~scovhsp,
+          text = ~paste(sseqid, '\n', 
+                        length, 'aa', '\n',
+                        scovhsp, '% coverage of known'),
+          color = ~qseqid,
+          size = ~length,
+          symbol = ~over95,
+          symbols = c('triangle-up', 'circle'),
+          sizes = c(2,100),
+          legendgroup = ~qseqid) %>% 
+  layout(xaxis = list(title = "De novo transcript coverage"),
+         yaxis = list(title = "Known transcript coverage"))
+```
+
 # Transcripts matched by percentage coverage.
@@
 There are `r length(unique(transcripts_covered_by_denovo$sseqid))` known transcripts covered by >=95% of at least one individual _de novo_ transcript.
+Of these `r length(unique(str_subset(transcripts_covered_by_denovo$sseqid, pattern = 'comp')))` are predicted virs from Zhu et al, 2016 and `r length(unique(str_subset(transcripts_covered_by_denovo$sseqid, pattern = 'PVX')))` are from Sal1 genome.
 There is a mean of `r mean(transcripts_covered_by_denovo$mismatch)` mismatches and `r mean(transcripts_covered_by_denovo$gapopen)` gaps and a median of `r median(transcripts_covered_by_denovo$mismatch)` mismatches and `r median(transcripts_covered_by_denovo$gapopen)` gaps in these alignments (Note that there are probably some de novo transcripts that align >=95% to the same known transcript).

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis_master.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_VivaxPirsInTheDeNovoAssemblies Master"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  assembly_name: 'SRX2884276_SRR5646671_scspades_0.95cdhitest'
  assembly_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/cdhitest/Rangel20_PvTCF/Rangel20_PvTCF_0.95_cdhitest.fasta'
  orf_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/Rangel20_PvTCF_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir.fasta'
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades/Rangel20_PvTCF/Rangel20_PvTCF_scspades_0.95cdhitest_hmmerPfam-Plasmodium_Vir_blast-VvxProt.outfmt6'
  new_table: FALSE
---

# Intro

This script analyses the results of a _de novo_ assembly of Plasmodium vivax RNAseq data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
# getwd()
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr', 
        lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0")
library('stringr')
library('purrr')
library('IRanges')
library('kableExtra')
library('plotly')
library('seqinr')
library('rmarkdown')
```

```{r function_ordering}
ordering <- function(to_order, order_vector, fromLast = TRUE){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = fromLast)
}
```

```{r load_data}
pir.info <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/PlasmoDBv56_VvxPirs_P01_Sal1.csv', 
                     header = TRUE)
```

```{r filter_nident}
#Set a filter for bp length to assume is a valid match
filter_nident <- 100
```

```{r assembly_name_expt}
#Extract the first two numbers of the name e.g. ERX2560125_ERR2541707
accession_name <- str_extract(params$assembly_name, 
                              pattern = '(SR|ER).{8,9}_(SR|ER).{8,9}(?=_)')
#If this works (not NA and only two elements) then find the expt name in the table
if(!is.na(accession_name) & length(unlist(strsplit(accession_name, '_'))) == 2){
  accession_table <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samples_to_cat.csv')
  expt_name <- accession_table[grepl(accession_table$fastq_1, pattern = accession_name),][,1]
} else {
  #Otherwise assume that the expt name is in the assembly_name
  expt_name <- str_extract(params$assembly_name, 
                           pattern = '[:upper:][:lower:]+[:digit:]{2}')
}
if(rlang::is_empty(expt_name)){
  stop(paste0('No experiment name for accession ', accession_name))
}
```

Filtering matches by a minimum of `r filter_nident`bp to be a valid match.

```{r read_blast_results}
# list.files('../blast', pattern = 'outfmt')

any_blast <- system(paste0('wc -l ', 
              params$blast_output_file), 
       intern = TRUE)
any_blast <- !grepl(any_blast, pattern = '^0 ')
if(any_blast){
outfmt6_headers <- 
  'qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore qcovs qcovhsp slen nident' %>% 
  strsplit(split = ' ') %>% 
  unlist

blast <- read.table(params$blast_output_file,
                    col.names = outfmt6_headers) %>% 
  mutate(
         scovhsp = round( (nident/slen * 100), digits = 2),
         scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
         evalue = format(evalue, scientific = TRUE)
  ) %>% 
  #Filter for the pir genes.
  # The Sal1 and Zhu genes are already only the virs so I grep for them first in ifelse.
  # Then, for the P01, I extract the gene name and search in the pir.info list.
  filter(nident > filter_nident,
         ifelse(grepl(sseqid, pattern = 'PVX|comp'),
                TRUE,
                str_extract_all(sseqid, 
                                pattern = '.+(?=\\.1)') %in% pir.info$Gene.ID)) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

# Filter for the top hits for each de novo transcript by e-value.

blast_top <- group_by(blast, qseqid) %>% 
  slice_min(evalue, n = 1)

#Filter for the de novo transcripts which only match a single known pir.
# Also removing idents below 100 because these are probably poor matches.

blast_uniq <- filter(blast, num_genes_match == 1)

#sanity test
# all(blast$scovs >= blast$scovhsp)
} else {
  blast <- 0
  print('There are no BLAST hits to the P01, Sal1 or Zhu16 de novo pirs')
}
```

# ORF analysis

```{r orf_fasta}
any_orfs <- system(paste0('wc -l ', params$orf_file), intern = TRUE)
any_orfs <- !grepl(any_orfs, pattern = '^0 ')
if(any_orfs){
  orf_fasta <- read.fasta(params$orf_file,
                          seqtype = c("AA"),
                          seqonly = FALSE)
} else {
  orf_fasta <- 0
  print("There are no pir gene ORFs detected by the criteria used.")
}
```

```{r orf_lengths, eval = any_orfs}

getLength(orf_fasta) %>% log %>% hist(breaks = 1000, 
                                      main = 'log ORF lengths with 150aa line')
abline(v = log(150), col = 'blue')

```

# Number of hits for each de novo transcript with a BLAST hit.

Are there many de novo transcripts which match multiple different known transcripts? Are most of them unique?

```{r number_of_hits_per_deNovoTranscript, eval = any_blast}
blast %>% 
  group_by(qseqid) %>% 
  summarise(num = n_distinct(sseqid)) %>% 
  .$num %>% 
  hist(main = 'Histogram of BLAST hits per de novo transcript (50 max)',
       breaks = 50)
```

#Genes with the most number of uniquely mapping de novo transcripts.

```{r number_of_uniq_transcripts_per_gene, eval = any_blast}
blast %>% 
  filter(num_genes_match== 1) %>% 
  group_by(sseqid) %>% 
  summarise(num_of_uniq_deNovoTranscripts = n_distinct(qseqid),
            max_id = max(pident),
            min_id = min(pident),
            max_nident = max(nident),
            min_nident = min(nident),
            max_gap = max(gapopen),
            min_gap = min(gapopen),
            max_qcovh = max(qcovhsp),
            min_qcovh = min(qcovhsp),
            max_scovh = max(scovhsp),
            min_scovh = min(scovhsp)) %>% 
  arrange(desc(num_of_uniq_deNovoTranscripts)) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "400px")
```

# Coverage of the subject (known transcripts) and query (assembled de novo transcripts) sequences.

Using only the de novo transcripts that uniquely match one known pir transcript.

For our purposes the subject coverage is probably most important.

```{r scov_v_qcov, eval = any_blast}
fig <- plot_ly(
  blast_uniq,
  x = ~qcovhsp,
  y = ~scovhsp,
  text = ~paste(sseqid)
) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
fig
```

Number of unique genes found by BLAST: `r ifelse(any_blast, unique(blast$sseqid) %>% length,0)`.

# Known transcripts with >95% coverage from one de novo transcript.

How many of the known transcripts are covered >=95% by de novo transcripts?

num_match_all and num_match_95 show how many other known transcripts these de novo transcripts match in total or >=95% respectively.

```{r transcripts_covering_most_of_known_transcripts, eval = any_blast}
transcripts_covered_by_denovo <- blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>%
  group_by(sseqid) %>% 
  filter(scovhsp >= 95) %>% 
  group_by(qseqid) %>%
  mutate(num_match_95 = n_distinct(sseqid)) %>% 
  arrange(sseqid)
transcripts_covered_by_denovo %>% 
  select(-qstart,-qend,-sstart,-send,-evalue,-bitscore,-qcovs,-slen,-sseqid,-nident,-scovs) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
```

Showing the BLAST alignments of these >=95% coverage de novo transcripts in a graph, and all the matches they have. 
The "n =" number after the transcript name shows how many other BLAST hits the de novo transcript had in total.
Size of each point shows the length of the alignment.

```{r transcripts_covering_most_of_known_transcripts_graph, eval = any_blast}
blast %>% 
  filter(qseqid %in% unique(transcripts_covered_by_denovo$qseqid)) %>%
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'bp', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

Alignments with >=95% known coverage but <50% de novo transcript coverage.

```{r low_qcovhsp, eval = any_blast}
if(transcripts_covered_by_denovo %>% 
   filter(qcovhsp < 50) %>% 
   nrow() > 0){
  transcripts_covered_by_denovo %>% 
    filter(qcovhsp < 50) %>% 
    kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
}
```

# All the BLAST matches

Fix: Colours for low numbers of hits.

```{r transcripts_coverage_allHits_transcripts_graph, eval = any_blast}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          legendgroup = ~qseqid) %>% 
  add_markers(text = ~paste(sseqid, '\n', 
                            length, 'aa', '\n',
                            scovhsp, '% coverage of known'),
              color = ~qseqid,
              size = ~length,
              symbol = ~over95,
              symbols = c('triangle-up', 'circle'),
              colors = "Set1",
              sizes = c(2,100)) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# BLAST hit coverages for the top alignments.

```{r transcripts_coverage_topBLAST_transcripts_graph, eval = any_blast}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  group_by(sseqid) %>% 
  slice_min(evalue, n = 1) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          legendgroup = ~qseqid) %>% 
  add_markers(text = ~paste(sseqid, '\n', 
                            length, 'aa', '\n',
                            scovhsp, '% coverage of known'),
              color = ~qseqid,
              size = ~length,
              symbol = ~over95,
              symbols = c('triangle-up', 'circle'),
              colors = "Set3",
              sizes = c(2,100)) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# Novel ORFs for ORFs not matching known virs.

Novel ORFs are described as those that don't match any known pir >=95% identity.

```{r orf_novel, eval = any_blast}
num_orf_novel_95 <- blast %>% 
  filter(scovhsp < 95) %>% 
  .$qseqid %>% 
  unique %>% 
  length

mRNA_novel_95 <- blast %>% 
  filter(scovhsp < 95) %>% 
  .$qseqid %>% 
  str_extract('(?<=NODE_)[[:digit:]]+(?=_)') %>% 
  unique

num_mRNA_novel_95 <-  length(mRNA_novel_95)

blast %>% 
  filter(scovhsp < 95) %>% 
  .$qseqid %>% 
  unique %>% 
  str_extract(pattern = '(?<=length_)[:digit:]+(?=_)') %>% 
  as.numeric %>% 
  hist(breaks = 50, main = "Novel ORF lengths (<95% id match known pirs)")
```

## Novel ORFs matched by percentage coverage threshold.

```{r orf_coverage_by_percentage, eval = any_blast}

orfs_covered_by_denovo_diffcov <- lapply(seq(0,100,by = 5),
                                                function(covrg) {
                                                  blast %>% 
                                                    group_by(sseqid) %>% 
                                                    filter(scovhsp < covrg) %>%
                                                    .$qseqid %>% unique %>% length
                                                })
names(orfs_covered_by_denovo_diffcov) <- seq(0,100,by = 5)
plot(y = unlist(orfs_covered_by_denovo_diffcov), 
     x = seq(0,100,by = 5),
     # ylim = c(0,max(unlist(orfs_covered_by_denovo_diffcov))),
     # main = 'For all de novo transcripts',
     ylab = 'Number of novel ORFs',
     xlab = 'Percentage of coverage matching a known pir',
     yaxt = 'none')
axis(2, at = seq(0,nrow(pir.info),by = 100))
# abline(a = 25, b = 0)

```

# BLAST hit coverages for the top alignments to the Zhu 2016 de novo virs.

```{r transcripts_coverage_topBLAST_Zhu16transcripts_graph, eval = any_blast}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  filter(grepl(sseqid, 
               pattern = 'comp')) %>% 
  group_by(sseqid) %>% 
  slice_min(evalue, n = 1) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# Transcripts matched by percentage coverage.

```{r transcript_coverage_by_percentage, eval = any_blast}

transcripts_covered_by_denovo_diffcov <- lapply(seq(0,100,by = 5),
                                                function(covrg) {
                                                  blast %>% 
                                                    group_by(sseqid) %>% 
                                                    filter(scovhsp >= covrg) %>%
                                                    .$sseqid %>% unique %>% length
                                                })
names(transcripts_covered_by_denovo_diffcov) <- seq(0,100,by = 5)
plot(y = unlist(transcripts_covered_by_denovo_diffcov), 
     x = seq(0,100,by = 5),
     ylim = c(0,nrow(pir.info)),
     main = 'For all de novo transcripts',
     ylab = 'Number of known transcripts covered',
     xlab = 'Percentage of coverage by individual transcripts',
     yaxt = 'none')
axis(2, at = seq(0,nrow(pir.info),by = 100))
# abline(a = 25, b = 0)

```

What data is worth saving? 
* Number of ORFs found.
* Number of ORFs over certain lengths.

```{r orfs_to_save}
number_transcripts <- system(paste0('cat ',
                                    params$assembly_file, 
                                    ' | grep ">" | wc -l'),
                             intern = TRUE)
if(any_orfs){
  number_orfs_unique_mRNA <- str_extract(getName(orf_fasta),
                                         pattern = '(?<=NODE_)[[:digit:]]+(?=_)') %>% 
    unique() %>% 
    length

  orf_lengths <- getLength(orf_fasta)

  orfs_unique_mRNA_Over200aaLength <- str_extract(getName(orf_fasta[orf_lengths >= 200]),
                                                  pattern = '(?<=NODE_)[[:digit:]]+(?=_)') %>% 
    unique()

  number_orfs_unique_mRNA_Over200aaLength <-  length(orfs_unique_mRNA_Over200aaLength)

  num_ORFs_over200aaLength = sum(orf_lengths>=200)
  num_ORFs_novel_nomatch95known = num_orf_novel_95
  num_uniq_mRNA_ORFs_predict = number_orfs_unique_mRNA
  num_uniq_mRNA_ORFs_over200aaLength = number_orfs_unique_mRNA_Over200aaLength

} else {
  number_orfs_unique_mRNA <- orf_lengths <- orfs_unique_mRNA_Over200aaLength <- number_orfs_unique_mRNA_Over200aaLength <- num_ORFs_over200aaLength <- num_ORFs_novel_nomatch95known <- num_uniq_mRNA_ORFs_predict <- num_uniq_mRNA_ORFs_over200aaLength <- 0
}
```

```{r blast_to_save}
if(any_blast){
  mRNA_novel_95_over200aa <- mRNA_novel_95[mRNA_novel_95 %in% orfs_unique_mRNA_Over200aaLength]

  num_mRNA_novel_95_over200aa <- length(mRNA_novel_95_over200aa)
  num_uniq_mRNA_ORF_novel_nomatch95known = num_mRNA_novel_95
  knwn_trnscrpt_cvg_over75 = transcripts_covered_by_denovo_diffcov[['75']]
  ancestral_over75 = blast %>% 
    group_by(qseqid) %>%
    mutate(num_match_all = n_distinct(sseqid)) %>%
    group_by(sseqid) %>% 
    filter(scovhsp >= 75, 
           grepl(sseqid, 
                 pattern = 'PVP01_1147700|PVX_113230')) %>% 
    nrow
  knwn_trnscrpt_cvg_over95 = length(unique(transcripts_covered_by_denovo$sseqid))
  zhu16_trnscrpt_cvg_over95 = length(unique(str_subset(transcripts_covered_by_denovo$sseqid, 
                                                       pattern = 'comp')))
  Sal1_trnscrpt_cvg_over95 = length(unique(str_subset(transcripts_covered_by_denovo$sseqid, 
                                                      pattern = 'PVX')))
  P01_trnscrpt_cvg_over95 = length(unique(str_subset(transcripts_covered_by_denovo$sseqid, 
                                                     pattern = 'PVP01')))
  ancestral_over95 = ifelse(
    any(grepl(transcripts_covered_by_denovo$sseqid,
              pattern = 'PVP01_1147700|PVX_113230')),
    paste(transcripts_covered_by_denovo$sseqid[
      grepl(transcripts_covered_by_denovo$sseqid,
            pattern = 'PVP01_1147700|PVX_113230')
    ], sep = '; '),
    0)
  mean_mismatches = mean(transcripts_covered_by_denovo$mismatch)
  mean_gaps = mean(transcripts_covered_by_denovo$gapopen)
  med_mismatches = median(transcripts_covered_by_denovo$mismatch)
  med_gaps = median(transcripts_covered_by_denovo$gapopen)
}else{
  mRNA_novel_95_over200aa <- num_mRNA_novel_95_over200aa <- num_uniq_mRNA_ORF_novel_nomatch95known <- knwn_trnscrpt_cvg_over75 <- ancestral_over75 <- knwn_trnscrpt_cvg_over95 <- zhu16_trnscrpt_cvg_over95 <- Sal1_trnscrpt_cvg_over95 <- P01_trnscrpt_cvg_over95 <- ancestral_over95 <- mean_mismatches <- mean_gaps <- med_mismatches <- med_gaps <- 0
}

number_orfs <- system(paste0('cat ',
                             params$orf_file,
                             ' | grep ">" | wc -l'), 
                      intern = TRUE)
```

```{r data_save}

df_save <- data.frame("experiment" = expt_name,
                      "assembly" = params$assembly_name,
                      "num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa" = num_mRNA_novel_95_over200aa,
                      "num_transcripts" = number_transcripts,
                      "num_ORFs_predict" = number_orfs,
                      "num_ORFs_over200aaLength" = num_ORFs_over200aaLength,
                      "num_ORFs_novel_nomatch95known" = num_ORFs_novel_nomatch95known,
                      "num_uniq_mRNA_ORFs_predict" = num_uniq_mRNA_ORFs_predict,
                      "num_uniq_mRNA_ORFs_over200aaLength" = num_uniq_mRNA_ORFs_over200aaLength,
                      "num_uniq_mRNA_ORF_novel_nomatch95known" = num_uniq_mRNA_ORF_novel_nomatch95known,
                      "knwn_trnscrpt_cvg_over75" = knwn_trnscrpt_cvg_over75,
                      "ancestral_over75" = ancestral_over75,
                      "knwn_trnscrpt_cvg_over95" = knwn_trnscrpt_cvg_over95,
                      "zhu16_trnscrpt_cvg_over95" = zhu16_trnscrpt_cvg_over95,
                      "Sal1_trnscrpt_cvg_over95" = Sal1_trnscrpt_cvg_over95,
                      "P01_trnscrpt_cvg_over95" = P01_trnscrpt_cvg_over95,
                      "ancestral_over95" = ancestral_over95,
                      "mean_mismatches" = mean_mismatches,
                      "mean_gaps" = mean_gaps,
                      "med_mismatches" = med_mismatches,
                      "med_gaps" = med_gaps,
                      "assembly_file" = params$assembly_file,
                      "orf_file" = params$orf_file,
                      "blast_file" = params$blast_output_file
)

if(!file.exists('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt') | params$new_table){
  #If table file doesn't exist then create it.
  write.table(df_save,
              file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt',
              sep = '\t',
              col.names = TRUE,
              append = FALSE,
              row.names = FALSE)
} else if (params$assembly_name %in% read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')$assembly){
  #If assembly_name is already in table then overwrite it.
  assembly_info <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')
  assembly_info[assembly_info$assembly %in% params$assembly_name,] <- df_save[1,]
  write.table(assembly_info,
              file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt',
              sep = '\t',
              row.names = FALSE)
} else {
  #Append to the existing table is both above are false
  write.table(df_save,
              file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt',
              sep = '\t',
              col.names = FALSE,
              append = TRUE,
              row.names = FALSE)
}

df_save %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "400px")

```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovo_getorfFind1_hmmerMatches_blast2VvxProt.Rmd
Add Chapter 5 vivax de novo results folder

--- 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis_SRX697844_SRR1571714_cdhit0.99_blastn.Rmd
@@
 ---
-title: "tl22-01_VivaxPirsInTheDeNovoAssemblies SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6"
+title: "tl22-01 BLASTing the HMMer matches from Plasmodium_Vir (using ORFs from getorf -find 1 option) to Plasmodium vivax P01 v56 proteins"
 output: 
@@
 params:
-  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6'
+  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades_getorfTableZero/SRX697844_SRR1571714/SRX697844_SRR1571714_scspades_getorfTableZero_hmmerPfam-Plasmodium_Vir_bast-VvxProt.outfmt6'
 ---
@@
 
-This script is looks at the BLAST results of a Trinity _de novo_ assembled transcriptome.
-
-# What this report shows:
-
-*The level of counts / TPM of pirs that are detected by the Trinity de novo assembled transcriptome.
-**e.g. What is the minimum expressed gene detected / What is the maximally expressed gene not detected.
+This script is looks at the BLAST results of the HMMer matches with the Plasmodium_Vir among the ORFs from the Plasmodium vivax RNAseq scspades assembly.
+BLASTed to the Plasmodium vivax P01 PlasmoDB v56 proteins.
 
 # To-do:
-*Check how they BLAST to the rest of the transcriptome, do pir de novo transcripts also BLAST to non-pir transcripts?
-*Select the most informative data and poo-poo the rest.
-*Save useful files made by this script as csv.
+* BLASTs to non-pirs?
 
@@
 options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
-
-#Possible files for blast_output_file:
-# '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_blastn.outfmt6'
 ```
@@
 ```{r read_blast_results}
-# list.files('../blast', pattern = 'outfmt')
-
 #Set a filter for bp length to assume is a valid match
@@
 
-blast <- read.table(params$blast_output_file,
+blast_all <- read.table(params$blast_output_file,
                         col.names = outfmt6_headers) %>% 
@@
   ) %>% 
+  filter(nident > filter_nident) %>%
+  group_by(qseqid) %>% 
+  mutate(num_genes_match = n_distinct(sseqid))
+
 #Filter for the pir genes.
-  filter(nident > filter_nident,
+blast <- filter(blast_all, 
                 sseqid %in% pir.info$Gene.ID) %>%
+  group_by(qseqid) %>% 
+  mutate(num_genes_match = n_distinct(sseqid))
+
+#Non-pir genes
+blast_nonpir <- filter(blast_all, 
+                       !sseqid %in% pir.info$Gene.ID) %>%
   group_by(qseqid) %>% 
@@
 
-# Coverage of the subject (known transcripts) and query (assembled de novo transcripts) sequences.
-
-Using only the de novo transcripts that uniquely match one known pir transcript.
-
-For our purposes the subject coverage is probably most important.
-
-```{r scov_v_qcov}
-fig <- plot_ly(
-  blast_uniq,
-  x = ~qcovhsp,
-  y = ~scovhsp,
-  text = ~paste(sseqid)
-) %>% 
-  layout(xaxis = list(title = "De novo transcript coverage"),
-         yaxis = list(title = "Known transcript coverage"))
-fig
-```
-
-Number of unique genes found by BLAST: `r unique(blast$sseqid) %>% length`.
-
 # Known transcripts with >95% coverage from one de novo transcript.
@@
 
+Note: PVP01_1147700 = ancestral
+
 ```{r transcripts_covering_most_of_known_transcripts_graph}
@@
           text = ~paste(sseqid, '\n', 
-                        length, 'bp', '\n',
+                        length, 'aa', '\n',
                         scovhsp, '% coverage of known'),
@@
      ylim = c(0,nrow(pir.info)),
-     main = 'For all de novo transcripts',
+     main = 'For pir de novo transcripts',
      ylab = 'Number of known transcripts covered',

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovo_hmmerMatches_blast2VvxProt.Rmd
Add Chapter 5 vivax de novo results folder

--- 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovo_getorfFind1_hmmerMatches_blast2VvxProt.Rmd
@@
 ---
-title: "tl22-01 BLASTing the HMMer matches from Plasmodium_Vir (using ORFs from getorf -find 1 option) to Plasmodium vivax P01 v56 proteins"
+title: "tl22-01 BLASTing the HMMer matches from Plasmodium_Vir to Plasmodium vivax P01 v56 proteins"
 output: 
@@
 params:
-  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades_getorfTableZero/SRX697844_SRR1571714/SRX697844_SRR1571714_scspades_getorfTableZero_hmmerPfam-Plasmodium_Vir_bast-VvxProt.outfmt6'
+  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_scspades_0.99cdhitest_hmmerMatches_blast-VvxProt.outfmt6'
 ---
@@
           text = ~paste(sseqid, '\n', 
-                        length, 'aa', '\n',
+                        length, 'bp', '\n',
                         scovhsp, '% coverage of known'),
@@
 ```{r low_qcovhsp}
-if(transcripts_covered_by_denovo %>% 
-   filter(qcovhsp < 50) %>% 
-   nrow() > 0){
 transcripts_covered_by_denovo %>% 
@@
   kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
-}
-```
-
-# All the BLAST matches
-
-```{r transcripts_coverage_allHits_transcripts_graph}
-blast %>% 
-  group_by(qseqid) %>%
-  mutate(num_match_all = n_distinct(sseqid)) %>% 
-  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
-         over95 = ifelse(scovhsp >= 95, 
-                         '>=95% coverage', 
-                         '<95% coverage')) %>% 
-  plot_ly(x = ~qcovhsp,
-          y = ~scovhsp,
-          text = ~paste(sseqid, '\n', 
-                        length, 'aa', '\n',
-                        scovhsp, '% coverage of known'),
-          color = ~qseqid,
-          size = ~length,
-          symbol = ~over95,
-          symbols = c('triangle-up', 'circle'),
-          sizes = c(2,100),
-          legendgroup = ~qseqid) %>% 
-  layout(xaxis = list(title = "De novo transcript coverage"),
-         yaxis = list(title = "Known transcript coverage"))
-```
-
-# BLAST hit coverages for the top alignments.
-
-```{r transcripts_coverage_topBLAST_transcripts_graph}
-blast %>% 
-  group_by(qseqid) %>%
-  mutate(num_match_all = n_distinct(sseqid)) %>% 
-  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
-         over95 = ifelse(scovhsp >= 95, 
-                         '>=95% coverage', 
-                         '<95% coverage')) %>% 
-  group_by(sseqid) %>% 
-  slice_min(evalue, n = 1) %>% 
-  plot_ly(x = ~qcovhsp,
-          y = ~scovhsp,
-          text = ~paste(sseqid, '\n', 
-                        length, 'aa', '\n',
-                        scovhsp, '% coverage of known'),
-          color = ~qseqid,
-          size = ~length,
-          symbol = ~over95,
-          symbols = c('triangle-up', 'circle'),
-          sizes = c(2,100),
-          legendgroup = ~qseqid) %>% 
-  layout(xaxis = list(title = "De novo transcript coverage"),
-         yaxis = list(title = "Known transcript coverage"))
 ```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxSalmonAlignment.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_VvxSalmonAlignment"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

# Intro

This script analyses the results of a _de novo_ assembly of Plasmodium vivax RNAseq data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = TRUE)
options(scipen = 999) 
# Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr', 
        lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0")
library('stringr')
library('purrr')
library('kableExtra')
library('plotly')
library('rmarkdown')
library('seqinr')
```

```{r function_ordering}
ordering <- function(to_order, order_vector, fromLast = TRUE){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = fromLast)
}
```

```{r function_uniquify}
#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
      name <- paste(vector_names[i],
                    df_num$rep[match(vector_names[i], df_num$names)],
                    sep = '_')
      #Increase rep counter
      df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
      new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r load_data}
pir.info <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/PlasmoDBv56_VvxPirs_P01_Sal1.csv', 
                     header = TRUE)
```

```{r de_novo_pir_list}
# pir_lists.dir <- "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/"
# pir_lists.files <- list.files(pir_lists.dir, 
#                               pattern = '0.95cdhitest_hmmerPfam-PlasmodiumVir.list',
#                               full.names = TRUE) %>% 
#   #Get those with file sizes above zero (no empty lists)
#   # file.info gets the file sizes
#   .[file.info(., extra_cols = FALSE)$size > 0]
# 
# # Read the pirs into a list (as a vector).
# # sapply through the file list, read in the table and subset with the automatic column name V1
# pir_list.vec <- unlist(
#   sapply(pir_lists.files, 
#          function(pir_list){
#            #str_removing the _1 at the end of the removal 
#            str_remove(read.table(pir_list)$V1, 
#                       pattern =  '_[[:digit:]]$')
#          })
# )

#Need to read in the >200aa ORF length  pirs
pir_list.vec <- str_extract(names(read.fasta('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta')),
                                  pattern = 'NODE.+(?=_[[:digit:]])')

print(paste0('Number of pirs total: ', length(pir_list.vec)))
```

```{r load_cluster_data}
node.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv') %>% 
  #Convert the cluster numbers to a factor as they are discrete categories.
  mutate(across(contains('lngth'), factor))
```

```{r define_subfam}
cluster_subfam.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/cluster_subfam_30pid25lngthI1.4.csv')

cluster_subfam_25.df <- head(cluster_subfam.df, 
                             n = 25)

# write.csv(cluster_subfam_25.df,
#           file = 'cluster_subfam_25_df.csv',
#           row.names = FALSE)

# rowwise makes mutate go row by row instead of vectorising across the whole X30pid25lngthI1.4 column
node_subfam.df <- rowwise(node.df) %>% 
  mutate(subfam_assignment = cluster_subfam.df$subfam[cluster_subfam.df$X30pid25lngthI1.4 %in% X30pid25lngthI1.4],
         pir_label = str_extract(label,
                                 pattern = 'NODE_.+(?=_[[:digit:]])'))

```

unassigned = No Sal1 sequences in the cluster for subfamily assignment.

```{r load_salmon}
#Get the full file paths of the salmon quant.sf files.
quant_files <- list.files(path = "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/salmon/all_files",
                          recursive = TRUE,
                          pattern = 'quant.sf',
                          full.names = TRUE)

#Load the samplesheet.txt file of metadata, add new column for salmon files
samplesheet <- read.delim("/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samplesheet.txt") %>% 
  mutate(salmon_quant = '')

#Make a table of the files and the sample metadata from samplesheet.txt
# for loop over all the quant_files that exist
for(quant_name in quant_files){
  #Looking for the sample prefix in the file path between 'all_files' and '{prefix}_quant'
  prefix <- str_extract(quant_name, 
                        pattern = '(?<=\\/all_files/).+(?=/.+_quant)')
  if(nrow(samplesheet[grep(samplesheet$prefix, 
                   pattern = prefix),]) > 1){
                     stop(paste0(prefix,
                                 " matches more than one row of the samplesheet"))
                   }
  samplesheet[grep(samplesheet$prefix, 
                   pattern = prefix),'salmon_quant'] <- quant_name
}

sum(grepl(samplesheet[,'salmon_quant'], pattern = 'camp')) == 247
```

Make a list element of all the quant values.

```{r list_expt_quant}
expt_quant <- lapply(unique(samplesheet$Experiment),
                     #For each experiment name
                     function(expt){
                       print(paste0('Starting ', expt))
                       #Get the experimental metadata by subsetting
                       expt_data <- samplesheet[samplesheet$Experiment == expt,]
                       if(!sum(grepl(expt_data$assembly, pattern = '.')) == 0){
                         #If sum of grepping for any character is not zero, do the analysis 
                         # i.e there are assembly names and probably assemblies/quant files
                         # For each individual assembly 
                         # (we can bind rows for samples that use the same assembly)
                         assembly_quants <- lapply(unique(expt_data$assembly), 
                                                   function(assembly){
                                                     #Get the prefix(es) that were quantified against this assembly
                                                     prefix <- expt_data[expt_data$assembly == assembly, 
                                                                         'prefix']
                                                     #Get the quant file name(s) and read it in
                                                     quant_file <- expt_data[expt_data$prefix %in% prefix,
                                                                             'salmon_quant']

                                                     #If the file name isn't zero length (i.e. there is no name)
                                                     if(!length(quant_file) == 0){
                                                       #If length is one then we can make the quant file for just one sample 
                                                       if(length(prefix) == 1){
                                                         quant <- read.delim(quant_file)
                                                         #Subset the Name and TPM columns and keep it as a data frame (drop = F)
                                                         quant.df <- quant[,c('Name','TPM'), drop = FALSE]

                                                         # Give a warning if the TPM values are all zero.
                                                         if(sum(as.double(quant$TPM)) == 0){
                                                           #Do not return a quant.df data.frame, making it nothing:
                                                           quant.df <- ''
                                                           print(paste0(prefix, 
                                                                        " associated file has recorded no TPM values"))
                                                           #empty return to end the lapply for this file
                                                           return()
                                                         }

                                                       } else {
                                                         #If longer than 1 so there's more than one quant file for this assembly
                                                         quant.rlist <- lapply(quant_file, read.delim)
                                                         #Use purrr reduce to bind the df in the list together with inner_join,
                                                         # combining by the 'Name' column which should be the same throughout.
                                                         quant.df <- purrr::reduce(quant.rlist, inner_join, by = 'Name')
                                                         quant.df <- quant.df[grepl(colnames(quant.df), pattern = 'Name|TPM')]

                                                         # Give a warning if the TPM values are all zero.
                                                         if(any(sapply(quant.df[grepl(colnames(quant.df), 
                                                                                      pattern = 'TPM')], 
                                                                       function(x) sum(as.double(x))) == 0)){
                                                           print(paste0("At least some of ", 
                                                                        expt, 
                                                                        " results have recorded no TPM values"))
                                                           return()
                                                         }

                                                       }

                                                       #Put the sample name (prefix) in the TPM column
                                                       colnames(quant.df)[grepl(colnames(quant.df), 
                                                                                pattern = 'TPM')] <- paste0(prefix,
                                                                                                            '_TPM')

                                                       # Mark the pir NODES as such (TRUE in the 'pir' column)
                                                       quant.df$pir <- quant.df$Name %in% pir_list.vec
                                                       print(paste0('Number of pirs = ', sum(quant.df$pir)))

                                                       #Assign the sub-families of the pirs
                                                       # Alter the label of the de novo pirs so they look the same
                                                       pir_labels <- str_extract(node_subfam.df$label,
                                                                                 pattern = 'NODE_.+(?=_[[:digit:]])')

                                                       quant.df <- rowwise(quant.df) %>% 
                                                         mutate(subfam = ifelse(Name %in% pir_labels,
                                                                                node_subfam.df$subfam_assignment[
                                                                                  pir_labels %in% Name
                                                                                ], 
                                                                                'noCluster'))

                                                       return(quant.df)
                                                     }
                                                   })

                         names(assembly_quants) <- str_extract(unique(expt_data$assembly),
                                                               pattern = '(?<=\\/cdhitest/).+(?=/.+fasta)')
                         print(paste0('Finished ', expt))
                         return(assembly_quants)  
                       }
                     }
)

names(expt_quant) <- unique(samplesheet$Experiment)
```

'noCluster' means that the pir sequence was not in the '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/gephi/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/nodes_mcl_50pid100lngthNoSame2Same.csv' file.
I think this means that they didn't have any BLAST similarity to other pir to the thresholds:
* 50% id.
* alignment >=100aa.

```{r save_quant}
lapply(names(expt_quant), 
       function(expt){
         lapply(names(expt_quant[[expt]]), 
                function(assembly){
                  if(!is.null(expt_quant[[expt]][[assembly]])){
                    #Make the directories to save the files in. Recursive = TRUE so that it will make parents t
                    # dir.create(paste( "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/quantFiles_R",
                    #                   expt, assembly, sep = '/'),
                    #            recursive = TRUE)
                    write.csv(expt_quant[[expt]][[assembly]],
                              file = paste( "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/quantFiles_R",
                                            expt, assembly,
                                            paste0('quant_', 
                                                   assembly,
                                                   '_subfam30pid25lngthI14.csv'),
                                            sep = '/'),
                              row.names = FALSE
                    )
                  }
                })
       })
```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxSalmonAnalysis.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_VvxSalmonAnalysis"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

# Intro

This script analyses the results of a _de novo_ assembly of Plasmodium vivax RNAseq data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE 
                      # warning = FALSE
                      )
options(scipen = 999) 
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
# Means that numbers are displayed normally not as 1e6 (1000000)
global_save_var <- FALSE
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('dplyr', 
        # lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0"
        )
library('readr')
library('stringr')
library('purrr')
library('kableExtra')
library('plotly')
library('rmarkdown')
library('data.table')
library('ggpubr')
library('seqinr')
```

```{r function_ordering}
ordering <- function(to_order, order_vector, fromLast = TRUE){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = fromLast)
}
```

```{r function_uniquify}
#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
      name <- paste(vector_names[i],
                    df_num$rep[match(vector_names[i], df_num$names)],
                    sep = '_')
      #Increase rep counter
      df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
      new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r read_quantR}

samplesheet <- read.delim("/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samplesheet.txt")

#Fix the missing helpful_assembly_name for Kim_U samples
samplesheet$helpful_assembly_name[grepl(samplesheet$library_name, pattern = 'U_{1,2}')] <- rep('Isolate_20', 2)

expt_num.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/expt_codes.csv')

expt_quant <- lapply(unique(samplesheet$Experiment),
                     function(expt){
                       #Get the experiment code.
                       expt_code <- expt_num.df$expt_code[match(expt, 
                                                                expt_num.df$expt)]
                         assembly.list <- lapply(unique(samplesheet$assembly[samplesheet$Experiment == expt]),
                                                 function(assembly_file){
                                                   assembly <- str_extract(assembly_file,
                                                                           pattern = '(?<=cdhitest\\/).+(?=\\/)')
                                                   #Get the file name by pasting the words together
                                                   quant_file <- paste(
                                                     "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/quantFiles_R",
                                                     expt, assembly, 
                                                     paste0('quant_', 
                                                            assembly,
                                                            '_subfam30pid25lngthI14.csv'
                                                     ),
                                                     sep = '/')
                                                   if(file_test("-f", quant_file)){
                                                     #If the file exists then we can read it in
                                                     quant_csv <- read.csv(quant_file, 
                                                                           header = TRUE)
                                                     return(quant_csv)
                                                   }
                                                 })
                         #Get the names for the list of assemblies, extract the important bits and replace the expt with the expt_code.
                         names(assembly.list) <- str_replace(
                           str_extract(unique(samplesheet$assembly[samplesheet$Experiment == expt]),
                                       pattern = '(?<=cdhitest\\/).+(?=\\/)'), 
                           pattern = expt,
                           replacement = expt_code)
                         return(assembly.list)
                     })

#Change the names of the experiment list to the experiment code.
names(expt_quant) <- expt_num.df$expt_code[match(unique(samplesheet$Experiment), 
                                                 expt_num.df$expt)]

#Remove null elements of list (no quant files)
expt_quant <- expt_quant[!unlist(lapply(expt_quant, 
                                        function(x) is.null(unlist(x))))]

```

# de novo transcript expression

```{r top25}
# Read in the top 25 clusters/sub-family designations
cluster_subfam_25.df <- read.csv(file = 'cluster_subfam_25_df.csv')
```

```{r top_of_each_subfam}
#Get consistent colour values so that each plot will use the same colour scheme.
# Rep each subfamily letter twice and paste '_2' onto the end.
select_subfam <- c('E', 'C_1', 'C_2', 'G', 'I', 'J', 'K_1', 'B_1',
                   'unassigned_1', 'unassigned_2', 'unassigned_3',
                   'other', 'noCluster') %>% ordering(c(.,'noCluster','other'))

cluster_subfam_25.df$subfamSelect <- ifelse(cluster_subfam_25.df$subfam %in% select_subfam, 
                                            cluster_subfam_25.df$subfam,
                                            'other')
cluster_subfam_select.df <- group_by(cluster_subfam_25.df, 
                                     subfamSelect) %>% 
  summarise(total_pirs = sum(total_pirs))

#I want 'none' to be coloured white, and 'other' grey, so I get a palette two colour too short and add them on.
top_subfam_col <- c(rep(safe_colorblind_palette, 
                        length.out = length(select_subfam) - 2), 'white', 'grey')

# top_subfam_col <- c('steelblue1',
#                     'orange', 'firebrick2',
#                     'yellow1',
#                     'springgreen4',
#                     'darkslateblue',
#                     'tomato3',
#                     'grey',
#                     'violetred', 'hotpink1', 'lightsalmon',
#                     'white','cyan3')

top_subfam_col <- c("#006ddb",
                    "#924900", "#db6d00",
                    "#ffff6d",
                    "seagreen3",
                    "#490092",

                    # "#920000",
                    'tomato3',

                    # "#004949",
                    'grey',

                    "#b66dff", "#ffb6db", "#ff6db6",
                    "white", "#009292")

pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
 "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
 "#920000","#924900","#db6d00","#24ff24","#ffff6d")
pie(rep(1,length(top_subfam_col)), col = top_subfam_col)
pie(rep(1,length(pal)), col = pal)

names(top_subfam_col) <- select_subfam

# write.csv(data.frame(subfam_assignment = names(top_subfam_col),
#                        colours = unname(top_subfam_col)),
#             file = 'top_subfam_col.csv',
#           row.names = FALSE)
```

How does the number of pir sub-families transcribed in each sample vary by BUSCO score?

```{r pir_diversity_v_BUSCOs}
buscos <- read.table('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info_extra_busco.txt')

subfam_numbers <- lapply(names(expt_quant), 
                         function(expt){
                           #expt <- 'E06'
                           #Doing two graphs, one for all pirs (above threshold) and one with the pir TPM normalised by the number in each sub-family/cluster.
                           expt_subfamNum <- lapply(
                             sort(names(expt_quant[[expt]])),
                             function(assembly){
                               #assembly <- 'E06_A'
                               print(assembly)
                               if(!is.null(expt_quant[[expt]][[assembly]])){
                                 # For this experiment+assembly, melt the pir part of the data frame so that we can calculate per sample the number of sub-families transcribed above 1 TPM.
                                 expt_quant[[expt]][[assembly]] %>% 
                                   filter(pir) %>% 
                                   reshape2::melt(.,
                                                  id.vars = c('pir','Name', 'subfam'),
                                                  measure.vars = str_subset(colnames(.), 
                                                                            pattern = 'TPM'),
                                                  value.name = 'tpm') %>% 
                                   group_by(variable, 
                                            subfam) %>% 
                                   summarise(max_tpm = max(tpm)) %>% 
                                   #Summarise should drop 'subfam' from grouping.
                                   filter(max_tpm >= 1) %>% 
                                   summarise(num_subfam = n_distinct(subfam)) %>% 
                                   mutate(expt = expt,
                                          assembly_name = assembly)
                               }
                             }) %>% 
                             bind_rows
                         }) %>% 
  bind_rows
#Warning "no non-missing arguments to max; returning -Inf" may indicate that there aren't any pirs present.
buscos <- mutate(rowwise(buscos), assembly_name = str_replace(assembly_name, experiment, experiment_num))
#Fix the messed up Gunalan names again
subfam_numbers <- mutate(subfam_numbers, assembly_name = str_replace(assembly_name, 
                                                                     'Guna19',
                                                                     'E07'))
busco_subfamNum.df <- inner_join(subfam_numbers, buscos, by = 'assembly_name')

expt_colors <- unique(buscos$expt_colour)
names(expt_colors) <- unique(buscos$experiment_num)
expt_colors <- expt_colors[sort(names(expt_colors))]

ggplot(busco_subfamNum.df,
       aes(x = perc_busco_plasmo_nonMissing,
           y = num_subfam,
           col = expt)) +
  geom_point(size = 5, alpha = .75) +
  theme_classic() +
  scale_color_manual(values = expt_colors[names(expt_colors) %in% busco_subfamNum.df$expt])

```

```{r barchart}

#Make samplesheet column for nicer sample names using other information
samplesheet_samplename <- 
  mutate(samplesheet, 
         Expt_code = expt_num.df$expt_code[match(Experiment, 
                                                 expt_num.df$expt)]) %>% 
  rowwise() %>% 
  mutate(sample_name = paste(c(Expt_code, 
                               Samples.combined, 
                               Sample_stage),
                             collapse = '_'),
         assembly_exptCode = str_replace_all(assembly, 
                                         pattern = Experiment, 
                                         replacement = Expt_code))

samplesheet_samplename$sample_name <- uniquify(samplesheet_samplename$sample_name, 
                                               first_one = TRUE) 
#Which type of bar chart am I making?
geom_col_pos <- 'fill'
# geom_col_pos <- 'stack'

# Make a stacked bar chart of the
lapply(names(expt_quant),
       function(expt){
         #expt <- 'E06'
         #Doing two graphs, one for all pirs (above threshold) and one with the pir TPM normalised by the number in each sub-family/cluster.
         subfampir.list <- lapply(
           c('totalTPM','absolute_subfam','normalised_subfam'),
           function(absolute_or_norm){
             #absolute_or_norm <- 'absolute_subfam'
             experiment_plot <- lapply(
               sort(names(expt_quant[[expt]])),
               function(assembly){
                 #assembly <- 'E06_A'
                 if(!is.null(expt_quant[[expt]][[assembly]])){

                   helpful_assembly_name <- unique(samplesheet_samplename$helpful_assembly_name[
                     str_detect(samplesheet_samplename$assembly_exptCode ,
                                assembly)
                   ])

                   #Set a size adjustment for when there are loads of graphs together and the axes need to be smaller
                   size_adjust <- ifelse(length(names(expt_quant[[expt]])) > 6,
                                         0.75,
                                         1)

                   subfam_top.df <- expt_quant[[expt]][[assembly]] %>%
                     filter(pir) %>%
                     #Include only the subfams in the top20
                     mutate(subfam_top = ifelse(subfam %in% select_subfam,
                                                subfam,
                                                'other')) %>%
                     reshape2::melt(.,
                                    id.vars = c('subfam_top','pir','Name'),
                                    measure.vars = str_subset(colnames(.),
                                                              pattern = 'TPM'),
                                    value.name = 'tpm') %>%
                     #Filtering step for TPM
                     filter(tpm >= 1) %>%
                     rowwise() %>%
                     mutate(sample_name = samplesheet_samplename$helpful_sample_name[
                       str_detect(samplesheet_samplename$prefix,
                                  str_remove(variable,
                                             pattern = '_TPM'))
                     ]) %>%
                     #Arranging by TPM here means that when factorised the transcripts will be ordered by tpm value
                     arrange(sample_name,subfam_top, tpm)

                   if(absolute_or_norm == 'normalised_subfam'){
                     #Normalise by the total number of pirs in each cluster
                     subfam_top.df <- mutate(subfam_top.df,
                                             number = cluster_subfam_select.df$total_pirs[
                                               match(subfam_top,
                                                     cluster_subfam_select.df$subfamSelect)
                                             ],
                                             tpm = tpm/cluster_subfam_select.df$total_pirs[
                                               match(subfam_top,
                                                     cluster_subfam_select.df$subfamSelect)
                                             ])
                   }

                   #If there are more rows than zero (i.e. there are pirs)
                   if(dim(subfam_top.df)[1] > 0){
                     plot <- ggplot(data = subfam_top.df,
                                    aes(x = factor(sample_name,
                                                   levels = ordering(sample_name,
                                                                     order_vector = c("^6hr", "before"))),
                                        y = tpm))

                     #For the 'totalTPM' graph I want the bar charts to be 'stacked' so the genes are added on top of each other and we can see the summed TPM values for each sample.
                     if(absolute_or_norm == 'totalTPM'){
                       geom_col_pos <- 'stack'
                       plot <- plot + geom_col(
                         position = geom_col_pos,
                         show.legend = FALSE
                       )
                     }else{
                       # For the graphs with the sub-families shown I want the bar charts to be 'fill' so that I'm visualising the ratios of the sub-family members.
                       geom_col_pos <- 'fill'
                       # colour and size so that each gene block has a small outline
                       plot <- plot + geom_col(
                         colour = 'black',
                         size = 0.02,
                         position = geom_col_pos,
                         show.legend = FALSE,
                         aes(fill = factor(subfam_top, 
                                           levels = names(top_subfam_col)))
                       )
                     }

                     plot <- plot + xlab('') + ylab('TPM') +

                       # facet_wrap(~subfam_top, ncol = 2, scales = "free_y") +
                       #Make scale_fill manual values depend on varying length of the palette so it can accomodate having more than 12 variables
                       scale_fill_manual(values = top_subfam_col) +
                       theme_classic() +
                       #Going to keep the axis text only if it's of interest, or the first plot. Must be in if{}else{} because ifelse breaks it.
                       theme(axis.text.x = element_text( size = 11*size_adjust),
                             axis.title.y = element_blank(),
                             axis.text.y = element_text(size = 15*size_adjust),
                             plot.title = element_text(size = 7*size_adjust)) +
                       ggtitle(paste0(expt,'_',assembly,'_',helpful_assembly_name))

                     if(geom_col_pos != 'fill'){
                       #If not using 'fill' then need a nicer y axis
                       plot <- plot +
                         scale_y_continuous(breaks = signif(
                           seq(0,
                               max(
                                 group_by(subfam_top.df,
                                          variable) %>%
                                   summarise(total_tpm = sum(tpm)) %>%
                                   .$total_tpm
                               ),
                               length.out = 5),
                           digits = 2)
                         )
                     }

                     return(plot)
                   } else {
                     return(NULL)
                   }

                 }
               })
           })

         names(subfampir.list) <- c('totalTPM','absolute_subfam','normalised_subfam')

         if(!is.null(unlist(subfampir.list))){
           # If list returned isn't just NULL

           #Go into the 'all_pirs' data to get the indexes for splitting the graphs into two groups.
           # This is to make the graphs look neater if there's more than ten datasets.
           # Need to do this on the 'all_pirs' data so that the two sets of graphs will be directly comparable. Doing this for
           #   each in turn may lead to different indexes due to different numbers of 'subfam_pirs'.

           #Get index of NULL values to remove. 
           notNull.index <- !unlist(lapply(subfampir.list[[1]],
                                           function(plot) is.null(plot)))
           #Remove the NULL elements of lists
           subfampir.list <- lapply(subfampir.list, function(x.list) x.list[notNull.index])
           assembly_names.notNull <- sort(names(expt_quant[[expt]]))[notNull.index]

           plot.list <- lapply(
             c("absolute_subfam",
               "normalised_subfam"),
             function(absolute_or_norm){
               twoplot.list <- lapply(
                 seq_along(subfampir.list[[1]]), 
                 function(subfampir.index){ 
                   apl <- subfampir.list[['totalTPM']][subfampir.index][[1]]
                   bpl <- subfampir.list[[absolute_or_norm]][subfampir.index][[1]]
                   assembly <- assembly_names.notNull[subfampir.index]
                   helpful_name <- unique(samplesheet_samplename$helpful_assembly_name[
                     str_detect(samplesheet_samplename$assembly_exptCode ,
                                assembly)
                   ])

                   if(global_save_var){
                     ggsave(
                       plot = cowplot::plot_grid(apl, bpl, ncol = 2),
                       filename = paste0(
                         '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/salmonQuant_plots/tl22-01_',
                         expt, '_', absolute_or_norm, '_', assembly,
                         '_quantSalmon_stacked_subfam30pid25lngthI14.pdf'
                       ),
                       width = 7,
                       height = 5,
                       units = "in",
                       dpi = 330
                     )
                   }

                   return(cowplot::plot_grid(apl, bpl, ncol = 2))
                 }
               )
             }
           )
         }
       })

```

Notes:
* Cheng19 samples particularly divergent?
* Some samples far more heterogeneous than others.
* Strain specific differences show overwhelming inclusion of some families over others.
** e.g. A few Kim19 show no other assembled sub-family beyond C_1.
** Kim17 isolate 2
* E_1s and C_1s tend to dominate expression (E_1 and C_1 are the largest groups).
* Most pir expression comes from those which cannot be straightforwardly characterised by existing sub-families.

* Kim samples mostly show major loss of pir transcription before and after chloroquine admin, although could this be explained by other factors? Also not seen for every sample.
* Muller19 Thai shows that the two sporozoite pirs are expressed, albeit at low quantities, neither in a subfam.

* Overall the proportions of sub-fams are very consistent! 
** Could this be an artefact of clustering and choosing the largest clusters to plot?

```{r legend}
#Just showing the legend without a plot to accompany the above.
legend.plot <- ggplot(data.frame(subfam = names(top_subfam_col), top_subfam_col), 
       aes(subfam, fill = subfam)) + 
  geom_bar() +
  scale_fill_manual(values = top_subfam_col) +
  guides(fill = guide_legend(title = 'Subfamily designation',
                             ncol = 1)) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+ # showing the point in the legend
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = c(0.5, 0.5), # move the legend to the center
        legend.title = element_blank(),
        legend.key = element_rect(fill='NA'),
        panel.grid = element_blank(),
        panel.border = element_rect(colour = "black", fill='white', size=1))

if(global_save_var){
  ggsave(
    plot = legend.plot,
    filename = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/salmonQuant_plots/tl22-01_legend_quantSalmon_stacked_subfam30pid25lngthI14.pdf',
    width = 7,
    height = 5,
    units = "in",
    dpi = 330
  )
}
```

```{r save_barchart_data}
expt_barchart.list <- lapply(names(expt_quant),
                           function(expt){
                             #expt <- 'E01'
                             #Doing two graphs, one for all pirs (above threshold) and one with the pir TPM normalised by the number in each sub-family/cluster.
                             subfampir.list <- 
                               lapply(sort(names(expt_quant[[expt]])),
                                      function(assembly){
                                        #assembly <- 'E01_SMRU1'
                                        experiment_plot <- if(!is.null(expt_quant[[expt]][[assembly]])){

                                          helpful_assembly_name <- unique(samplesheet_samplename$helpful_assembly_name[
                                            str_detect(samplesheet_samplename$assembly_exptCode ,
                                                       assembly)
                                          ])

                                          #Set a size adjustment for when there are loads of graphs together and the axes need to be smaller
                                          size_adjust <- ifelse(length(names(expt_quant[[expt]])) > 6,
                                                                0.75,
                                                                1)

                                          subfam_top.df <- expt_quant[[expt]][[assembly]] %>%
                                            filter(pir) %>%
                                            #Include only the subfams in the top20
                                            mutate(subfam_top = ifelse(subfam %in% select_subfam,
                                                                       subfam,
                                                                       'other')) %>%
                                            reshape2::melt(.,
                                                           id.vars = c('subfam_top','pir','Name'),
                                                           measure.vars = str_subset(colnames(.),
                                                                                     pattern = 'TPM'),
                                                           value.name = 'tpm',
                                                           variable.name = 'sample') %>%
                                            #Filtering step for TPM
                                            filter(tpm >= 1) %>%
                                            rowwise() %>%
                                            mutate(sample_name = samplesheet_samplename$helpful_sample_name[
                                              str_detect(samplesheet_samplename$prefix,
                                                         str_remove(sample,
                                                                    pattern = '_TPM'))
                                            ],
                                            assembly = assembly,
                                            expt = expt) %>%
                                            #Arranging by TPM here means that when factorised the transcripts will be ordered by tpm value
                                            arrange(sample_name,subfam_top, tpm)

                                          return(subfam_top.df)
                                        } 
                                      })
                             names(subfampir.list) <- sort(names(expt_quant[[expt]]))

                             subfampir.df <- bind_rows(subfampir.list[!unlist(lapply(subfampir.list, 
                                                                                     plyr::empty))])
                             return(subfampir.df)
                           })

names(expt_barchart.list) <- names(expt_quant)

expt_barchart.list <- expt_barchart.list[!unlist(lapply(expt_barchart.list,
                                                        plyr::empty))]

expt_barchart.df <- bind_rows(expt_barchart.list)

```

```{r highest_prop_subfam}

lapply(expt_barchart.list, 
       function(expt.df){
         expt.df %>% 
           group_by(subfam_top, 
                    assembly,
                    sample,
                    sample_name) %>%
           summarise(tpm_sum = sum(tpm))          %>%
           group_by(assembly, sample, sample_name) %>%
           slice_max(order_by = tpm_sum, n = 10)
       })

```

```{r normalised}
lapply(expt_barchart.list, 
       function(expt.df){
         subfam_top.df <- mutate(expt.df,
                                 number = cluster_subfam_25.df$total_pirs[
                                   match(subfam_top,
                                         cluster_subfam_25.df$subfam)
                                 ],
                                 tpm_norm = tpm/cluster_subfam_25.df$total_pirs[
                                   match(subfam_top,
                                         cluster_subfam_25.df$subfam)
                                 ]) %>% 
           group_by(subfam_top, 
                    assembly,
                    sample,
                    sample_name) %>%
           summarise(tpmNorm = sum(tpm_norm), 
                     tpm = sum(tpm),
                     num = n_distinct(Name))          %>%
           group_by(assembly, sample, sample_name) %>%
           slice_max(order_by = tpmNorm, n = 5)
       })
#other is NA right now - need to fix this.
```

Unassigned_3 seems to be enriched for transcription given its size.

Sub-families E and C_2 are large groups but still higher transcribers.

```{r heatmap_clustering}
heatmap.df <- lapply(names(expt_quant),
       function(expt){
         #expt <- 'E01'
           lapply(sort(names(expt_quant[[expt]])),
                  function(assembly){
                    #assembly <- 'E01_SMRU1'

                    #Filter for ext=pressed pirs
                    pir_tpmover1.df <- expt_quant[[expt]][[assembly]] %>% 
                      filter(pir) %>% rowwise %>% filter(
                             across(matches('TPM'), ~ any(.x > 1)))

                    if(nrow(pir_tpmover1.df) == 0){
                      # print('no pirs')
                      return()
                    }

                    helpful_assembly_name <- unique(samplesheet_samplename$helpful_assembly_name[
                      str_detect(samplesheet_samplename$assembly_exptCode ,
                                 assembly)
                    ])

                      #Calculate the TPM of each sub-family as a proportion of pir transcription
                      reshape2::melt(pir_tpmover1.df,
                                     id.vars = c('pir','Name', 'subfam'),
                                     measure.vars = str_subset(colnames(pir_tpmover1.df), 
                                                               pattern = 'TPM'),
                                     value.name = 'tpm',
                                     variable.name = 'sample') %>% 
                      group_by(sample, subfam) %>% 
                      summarise(subfam_tot = sum(tpm)) %>% 
                      mutate(subfam_prop = subfam_tot/sum(subfam_tot)*100) %>% 
                      mutate(sum_prop = sum(subfam_prop),
                             sample_name = paste0(expt, '_',
                                                  helpful_assembly_name, 
                                                  '_',
                                                  samplesheet_samplename$helpful_sample_name[
                                                    str_detect(samplesheet_samplename$prefix,
                                                               str_remove(sample,
                                                                          pattern = '_TPM'))
                                                  ])) %>% 
                      dcast(sample_name ~ subfam, value.var = 'subfam_prop') 
                  })
       }) %>% bind_rows

heatmap.df <- tibble::column_to_rownames(heatmap.df, 
                                 var = 'sample_name')
heatmap.df <- t(as.matrix(heatmap.df))

library('ComplexHeatmap')
heatmap.df[is.na(heatmap.df)] <- 0
Heatmap(heatmap.df, 
        cluster_rows = TRUE, 
        cluster_columns = TRUE,
        column_split = 10)

```

```{r node-df_tpmvalues}

# Add the TPM values to the node.df dataset alongside a threshold expressed/not expressed

# Get a list of all the de novo transcripts and their median TPM. 
#  Shouldn't be any with the same name unless by exceptional chance?
gene_quant <- lapply(names(expt_quant), 
                     function(expt.name) {
                       expt.list <- expt_quant[[expt.name]]
                       # print(names(expt.list))
                       lapply(names(expt.list), 
                              function(quant.name) {
                                quant.df <- expt.list[[quant.name]]
                                if(!is.null(quant.df)){
                                  #Melt the dataframe to calculate the median TPM value for each de novo transcript across all samples.
                                  quant.df %>% 
                                    reshape2::melt(.,
                                                   id.vars = c('pir','Name'),
                                                   measure.vars = str_subset(colnames(.), 
                                                                             pattern = 'TPM'),
                                                   value.name = 'tpm') %>% 
                                    group_by(Name, pir) %>% 
                                    summarise(med_tpm = median(tpm),
                                              max_tpm = max(tpm)) %>% 
                                    mutate(label_no_end = paste(quant.name, Name, sep = '_'),
                                           expt_assembly = paste(expt.name, quant.name, sep = '_'))
                                }
                              }) %>% 
                         bind_rows
                     }) %>% 
  bind_rows

# Need to add the assembly name to the transcript name (as in node.df) because some are the same, which I find surprising!
# Some are from the same assemblies but not all of them:
# gene_quant[gene_quant$Name == 'NODE_24950_length_245_cov_2.19474',]

```

```{r over1tpm_listFiles}
#Note that node_tpm.df includes none-quantified genes, like P01/Sal1/Zhu16 and the Rangel genes.
# Will use over or equal to 1 TPM in at least one sample (max_tpm > 1) as the threshold for active transcription.
expressed_pirTranscripts <- filter(gene_quant, pir, max_tpm>=1)

paste0('Number of transcribed pirs = ',
       nrow(expressed_pirTranscripts),
       ' and number of non-transcribed pirs = ',
       nrow(filter(gene_quant, pir)) - nrow(expressed_pirTranscripts))

G18G19_hmmerVir200aa.fa <- read.fasta('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta')

names_no_end <- str_remove(names(G18G19_hmmerVir200aa.fa), pattern = '_[[:digit:]]{1,2}$')

#These aren't all unique, I guess this doesn't impact anything later.

#Subset the fasta file by the pirs which are actively transcribed.
# Note that this may be greater than expected due to the fact that there may be multiple pir ORFs for one expressed pir transcript.
# Also it is smaller than the number of expressed_pirTranscripts because not all of the expressed_pirTranscripts have a >200aa ORF.
# Note that the pirs are defined as such based on the *_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir.list files (so before 200aa filtering).

G18G19_hmmerVir200aa_tpmOver1.fa <- G18G19_hmmerVir200aa.fa[
  names(G18G19_hmmerVir200aa.fa)[names_no_end %in% expressed_pirTranscripts$label_no_end]
  ]
print(paste0('Number of ORFs (>200aa ORF tpm over 1) = ',
      length(G18G19_hmmerVir200aa_tpmOver1.fa)))
print(paste0('Number of unique transcripts (>200aa ORF tpm over 1) = ',
      length(unique(str_extract(names(G18G19_hmmerVir200aa_tpmOver1.fa), 
                   pattern = '.+(?=_[[:digit:]]$)')))))

table(str_extract(names(G18G19_hmmerVir200aa_tpmOver1.fa), pattern ='.+(?=_NODE)'))

# write.fasta(G18G19_hmmerVir200aa_tpmOver1.fa,
#             names(G18G19_hmmerVir200aa_tpmOver1.fa),
#             file.out = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')

```

```{r Num-pir-transcript_v_Num-transcript-total}
assembly_info.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt') %>% 
  filter(!grepl(assembly, pattern = '0\\.9c|0\\.99'),
         !grepl(assembly, pattern = 'SRX5158513_SRR8347514|SRX5158512_SRR8347515'))  
#Filter out any none-0.95 assembly made. And potential Kim19_A duplicate?

#Order experiments by years
expt_num.vec <-ordering(unique(assembly_info.df$experiment), 
                                   order_vector = c(16:22))
#Get experiment numbers and add '0' to single digit numbers.
names(expt_num.vec) <-  paste0('E', 
                              sapply(seq_along(expt_num.vec), function(x){
                                ifelse(nchar(x)==1, 
                                       paste0('0', x),
                                       x)
                              }))

#Manually add Gunalan19 and _Boonkaew20 (in addition to Guna19 and Boonkaew20)
expt_num.vec <- c(expt_num.vec, E07 = 'Gunalan19', E08 = '_Boonkaew20')

#Get the relevant info from samplesheet otherwise loads of rows will be duplicated when joining.
samplesheet_small.df <- mutate(samplesheet, assembly_file = assembly) %>% 
  select(assembly_file, Cat.Sample.stage.s.) %>% unique

assembly_info_samplesheet.df <- left_join(assembly_info.df, samplesheet_small.df, by = 'assembly_file')

assembly_info_samplesheet.df <- mutate(assembly_info_samplesheet.df,
                                        experiment_num = names(expt_num.vec)[match(experiment,
                                                                                  expt_num.vec)])

#Get consistent shapes and colours for the graphs

shapes_val <- 1:length(unique(assembly_info_samplesheet.df$Cat.Sample.stage.s.))
names(shapes_val) <- unique(assembly_info_samplesheet.df$Cat.Sample.stage.s.)

colours_val <- safe_colorblind_palette[1:length(unique(names(expt_num.vec)))]
names(colours_val) <- unique(names(expt_num.vec))

expt_num.df <- data.frame(expt = expt_num.vec, 
                          expt_code = names(expt_num.vec), 
                          col = c(colours_val,colours_val['E07'],colours_val['E08']))
# write.csv(expt_num.df,
#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/expt_codes.csv',
#           row.names = FALSE)

( plot <- ggplot(assembly_info_samplesheet.df, 
       aes(x = num_transcripts,
           y = num_uniq_mRNA_ORFs_over200aaLength,
           col = experiment_num,
           shape = Cat.Sample.stage.s.
       )) +
  geom_point(size = 1, stroke = 2, alpha = .75) +
  scale_shape_manual(values=shapes_val,
                     name = 'Life cycle stage',
                     guide = guide_legend(ncol = 2)) +
  scale_color_manual(values = colours_val,
                     name = 'Experiment',
                     guide = guide_legend(ncol = 2)) +
  xlab('Number of assembled transcripts') +
  ylab(expression(paste('Number of putative ', italic('pir'), ' transcripts'))) +
  theme_classic() )

if(global_save_var){
  ggsave(plot = plot,
         '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/assemblies_numPirTranscripts_totalTranscripts.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}
```

ISSUE: Why is the number of sequences in '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta' larger than sum(assembly_info_samplesheet.df£$num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa)

Observations:
* Some drop-outs between the samples? Perhaps because one transcript was exclusively assembled from one sample.

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_analysingAssemblyStats.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_analysingAssemblyStats"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
# getwd()
```

```{r load_packages}
library(ggplot2)
library(stringr)
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
```

```{r read_data}
df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')
```

#Want to show the stages of development in this.

```{r expt_name}
 accession_table <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samples_to_cat.csv')

#Extract the first two numbers of the name e.g. ERX2560125_ERR2541707
df$Cat.Sample.stage.s. <- factor(
  sapply(
    df$assembly, 
    function(assembly_name){
      accession_name <- str_extract(assembly_name, 
                                    pattern = '(SR|ER).{8,9}_(SR|ER).{8,9}(?=_)')
      #If this works (not NA and only two elements) then find the expt name in the table
      if(!is.na(accession_name) & length(unlist(strsplit(accession_name, '_'))) == 2){
        cat_sample_stage <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$fastq_1, pattern = accession_name)]
      } else {
        #Get the first two elements of the name (separated by '_') - split then paste back together.
        expt_name <- paste(str_split_fixed(assembly_name, 
                                      pattern = '_', n = 3)[c(1,2)], collapse = '_')
        if(any(grepl(accession_table$Concatenated.file, 
                     pattern = expt_name))){
          #Otherwise this is probably a re-named concatenated file of samples, so
          # we need the 'Concatenated file' column.
          cat_sample_stage_all <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$Concatenated.file, 
                                                                            pattern = expt_name)]
        } else {
          #Instead of being sensible I had a different naming scheme for the un-concatenated Kim and Gunalan files.
          cat_sample_stage_all <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$file_prefix, 
                                                                            pattern = expt_name)]
        }
        cat_sample_stage <- unique(cat_sample_stage_all)
        if(rlang::is_empty(cat_sample_stage)){
          stop(paste0('No life cycle stage name for assembly ', assembly_name))
        } else if(!length(cat_sample_stage) == 1){
          cat_sample_stage <- 'Mixed.stages'
        }
      }
      return(cat_sample_stage)
    })
)
```

```{r numNovPir_v_totalTranscripts}
(plot <- ggplot(data = df, aes(x = num_transcripts, 
                               y = num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa,
                               col = experiment,
                               shape = tolower(Cat.Sample.stage.s.))) +
   geom_point() +
   scale_color_manual(values = safe_colorblind_palette) +
   scale_shape_manual(values=1:nlevels(df$Cat.Sample.stage.s.),
                      name = 'Life cycle stage') +
   xlab('Number of assembled transcripts') +
   ylab('Number of novel pir ORFs') +
   theme_classic() +
   theme(legend.text = element_text(size = 7),
         legend.title = element_text(size = 7),
         legend.box = 'horizontal'))
ggsave(plot = plot,
       '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/assemblies_numNovPir_v_totalTranscripts.pdf',
       device = 'pdf',
       width = 7,
       height = 5,
       units = "in")
```

```{r numPirORF_v_totalTranscripts}
ggplot(data = df, aes(x = num_transcripts, 
                      y = num_ORFs_predict,
                      col = experiment,
                      shape = tolower(Cat.Sample.stage.s.))) +
  geom_point() +  
  xlab('Number of assembled transcripts') +
  ylab('Number of predicted pir ORFs') +
  scale_color_manual(values = safe_colorblind_palette) +
  scale_shape_manual(values=1:nlevels(df$Cat.Sample.stage.s.),
                     name = 'Life cycle stage') +
  theme_classic() +
  theme(legend.text = element_text(size = 7),
        legend.title = element_text(size = 7),
        legend.box = 'horizontal')
```

```{r numNovel_v_numPirORF}
ggplot(data = df, aes(x = num_ORFs_predict, 
                      y = num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa,
                      col = experiment,
                      shape = tolower(Cat.Sample.stage.s.))) +
  geom_point() +
  xlab('Number of predicted pir ORFs') +
  ylab('Number of novel pir ORFs') +
  scale_color_manual(values = safe_colorblind_palette) +
  scale_shape_manual(values=1:nlevels(df$Cat.Sample.stage.s.),
                     name = 'Life cycle stage') +
  theme_classic() +
  theme(legend.text = element_text(size = 7),
        legend.title = element_text(size = 7),
        legend.box = 'horizontal')
```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_analysingNetworkStats.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_analysingNetworkStats"
output: html_document
---

Going to adapt this script to analyse multiple networks.

What do we want from the optimal network?:

* More sequences included among the larger clusters of the network.
* Sal1 sub-families are rarely mixed in the same clusters.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
# getwd()
```

```{r load_packages}
library(ggplot2)
library(stringr)
library(dplyr)
library(data.table)
library(seqinr)
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#6699CC", "#661100", "#888888")
global_save_var <- FALSE
```

```{r function_alternateReverse}
alternateReverse <- function(input.vec){
  #Alternate the indexes of a vector, with one alternation in reverse order.
  # Good for colour systems so the end of a gradient is shifted between very different colours.
  output.vec <- c(rbind(input.vec[seq_along(input.vec)%%2==1], 
                        rev(input.vec[seq_along(input.vec)%%2==0])))
  if(length(output.vec) != length(input.vec)){
    #If input.vec is odd length rbind will add an extra value at the end.
    # So the length will be longer, this removes the last, erroneous element.
    return(output.vec[-last(output.vec)])
  }else{
    return(output.vec)
  }
}
```

```{r function_uniquify}
#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
      name <- paste(vector_names[i],
                    df_num$rep[match(vector_names[i], df_num$names)],
                    sep = '_')
      #Increase rep counter
      df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
      new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r read_data}
accession_table <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samples_to_cat.csv')
assembly.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')
subfam.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-02_VirSubfamNetwork/pir_info/lopez13_vir_subfam.txt', header = FALSE)
colnames(subfam.df) <- c('Geneid','subfam')
expt.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/scripts/all-cat.list', header = FALSE, sep = '')
colnames(expt.df) <- c('Accession', 'expt')
expt_num.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/expt_codes.csv')
```

```{r MCL_parameters}
# list.files('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1', 
  #          pattern = 'mci.+list') %>% 
  # head %>% 
  # str_extract_all(pattern = '(?<=_)[[:digit:]]+(?=pid)|[[:digit:]]+(?=lngth)|(?<=mci\\.)I.+(?=\\.list)')

network_file_dir <- '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1'

#Search for all the MCI lists of the clustering, extract the relevant bits which shows the % id, alignment length and inflation factor.
network_parameters.list <- list.files(network_file_dir, 
           pattern = 'all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast.mci.I.+list') %>% 
  str_extract(pattern = 'I.+(?=\\.list)') %>% 
  # str_replace(pattern = 'NoSame2Same\\.mci\\.', replacement = '_') %>% 
  unique

#Get the MCI cluster values

mcl.list <- lapply(network_parameters.list,
                   function(network_parameters){
                     network_parameters.file <- list.files(network_file_dir,
                                                           pattern = paste0(
                                                             'all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast.mci.',
                                                             network_parameters, 
                                                             '.list'
                                                           ),
                                                           full.names = TRUE)
                     if(length(network_parameters.file) > 1){
                       stop(paste0('More than 1 possible file found for one set of network parameters for ',
                                   network_parameters))
                     }

        #Read_table saves the number of the line of the sequences (each line is a cluster) as a separate column
        network_parameters.table <- read.table(network_parameters.file)
        colnames(network_parameters.table) <- c('label', 'mcl_cluster')

        return(network_parameters.table)
       })
names(mcl.list) <- network_parameters.list
```

```{r get_MCLclusters}
fasta <- read.fasta(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')

inflation_values <- unique(str_extract(names(mcl.list),
                                       pattern = '(?<=I).+'))

# sequence.df <- lapply(names(fasta),
#        function(sequence.name){
#          if(grepl(sequence.name,
#                   pattern = 'P01|PVX|comp')){
#            #For the known pirs, which have P01, PVX or comp in the sequence name.
#            seq_source <- case_when(
#              grepl(sequence.name, pattern = 'P01') ~ 'P01',
#              grepl(sequence.name, pattern = 'PVX') ~ 'Sal1',
#              grepl(sequence.name, pattern = 'comp') ~ 'Zhu16'
#            )
#            seq_expt <- 'known'
#            #Get the subfam number for the Sal1 sequences.
#            seq_subfam <- ifelse(seq_source == 'Sal1',
#                                 subfam.df$subfam[match(sequence.name,
#                                                        subfam.df$Geneid)],
#                                 seq_source)
#            #Need to set the Sal1 sequences which don't have sub-family assignments to just Sal1
#            #  They are NA at this point.
#            seq_subfam <- ifelse(is.na(seq_subfam) && seq_source == 'Sal1',
#                   'Sal1',
#                   seq_subfam)
# 
#          }else{
#            #Get source name of de novo pirs from the first section of the sequence name
#            seq_source <- str_extract(sequence.name,
#                                      pattern = ".+(?=_NODE.+)")
#            seq_subfam <- 'assembled'
#            seq_expt <- expt.df$expt[match(seq_source, expt.df$Accession)]
#          }
# 
#          #Get Id (just a number) from the index of the sequence.
# 
#          info.df <- data.frame(Id = which(names(fasta) == sequence.name),
#                                label = sequence.name,
#                                source = seq_source,
#                                subfam = seq_subfam,
#                                expt = seq_expt)
# 
#          #Get the 10pid25lngthNoSame2Same.mci.IX values as this uses the most data and is the MCL clustering done with the most data.
# # 
# # mcl_10pid25lngth <- lapply(inflation_values,
# #                            function(infl){
# #                              mcl.df <- getElement(mcl.list,
# #                                                   paste0('10pid25lngthNoSame2Same.mci.I',
# #                                                          infl))
# #                            })
# # names(mcl_10pid25lngth) <- inflation_values
# 
#          #Go through the mcl.list to get the mcl cluster number for this sequence in each cluster.
#          # Make a dataframe to bind_cols with the previous information df.
#          sequence_mcl.df <- lapply(names(mcl.list),
#                 function(network_parameters){
#                   mcl.df <- getElement(mcl.list, network_parameters)
#                   #If it is in this network (i.e. has a BLAST alignment), then get cluster number
#                   # from 10pid25lngthNoSame2Same.mci.IX.
#                   # If not in the network, give a cluster number of zero
#                   # mcl_10pid25lngth_netpara <- getElement(mcl_10pid25lngth,
#                   #                                        str_extract(network_parameters,
#                   #                                                    pattern = '(?<=I).+'))
#                   mcl_cluster_num <- ifelse(sequence.name %in% mcl.df$label,
#                                             mcl.df$mcl_cluster[match(sequence.name,
#                                                                                        mcl.df$label)],
#                                             0)
#                   mcl_cluster_num.df <- data.frame(mcl_cluster_num)
#                   colnames(mcl_cluster_num.df) <- network_parameters
#                   return(mcl_cluster_num.df)
#                 }) %>% bind_cols
# 
#          return(bind_cols(info.df, sequence_mcl.df))
#        }) %>% bind_rows
#

# 
# write.table(sequence.df,
#             file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt')

sequence.df <- read.table(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt',
                          check.names = FALSE)

# # How many sequences have all zeros? Sequence.df should have same row num as fasta file length.
notInNetwork <- sequence.df[apply(sequence.df[,grepl(colnames(sequence.df), pattern  = '^I[[:digit:]]')], 
                     1, 
                     function(x) all(x == 0)),]$label
```

```{r blast_networks}
perc_ident.vec <- c(10,20,30,40,50,60,70,80,90)
length_aa.vec <- c(25,50,75,100,150,200,250,300)

# pid_lntgh.df <- lapply(perc_ident.vec, function(perc_ident){
#   lapply(length_aa.vec, function(length_aa){
#     #Read in BLAST file for these percentage id and length aa 
#     # Get a list of the unique sequences from the first two columns
#     blast.names <- read.delim(paste0(
#       '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_',
#       perc_ident,'pid',
#       length_aa,'lngthNoSame2Same.blast'
#     ),
#     header = FALSE) %>% 
#       .[,c(1,2)] %>% 
#       unlist() %>% 
#       unique
#     #lapply across the sequences in the fasta to record whether they are present in this network (i.e. in the BLAST file)
#     in_network.df <- lapply(names(fasta), 
#            function(seq.id){
#              in_network <- ifelse(seq.id %in% blast.names,
#                                   1,
#                                   0)
#              return(data.frame(label = seq.id, 
#                                in_network = in_network,
#                                pid = perc_ident,
#                                lngth = length_aa))
#            }) %>% bind_rows
#   })
# }) %>% bind_rows

```

```{r long_form_sequence.df}
# sequence.df.melt <- melt(sequence.df, 
#                          id.vars = c('Id','label','subfam','source','expt'),
#                          variable.name = 'network_parameters',
#                          value.name = 'cluster_num') %>% 
#   group_by(network_parameters, cluster_num) %>% 
#   mutate(cluster_size = n_distinct(label)) %>% 
#   group_by(network_parameters) %>% 
#   mutate(total_clusters = n_distinct(cluster_num),
#          total_clusters_over2 = n_distinct(cluster_num[cluster_size > 2]),
#          inflation = str_extract(network_parameters, pattern = '(?<=I).+'))
# 
# # all(sequence.df.melt$total_clusters >= sequence.df.melt$total_clusters_over2)
# 
# #Join the BLAST parameters and whether or not they're part of the network. 
# # Set the cluster_num to zero for not being in the network.
# sequence.df.melt <- inner_join(pid_lntgh.df, sequence.df.melt, by = 'label') %>% 
#   mutate(cluster_num = cluster_num*in_network) %>% 
#   group_by(pid, lngth, inflation) %>% 
#   mutate(num_inNetwork = sum(cluster_num > 0)) %>% 
#   select(-in_network)

# write.table(sequence.df.melt,
#             file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
sequence.df.melt <- read.table(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
```

```{r not_in_top100}
group_by(sequence.df.melt, pid, lngth, inflation) %>% 
  filter(cluster_num < 1 | cluster_num > 100) %>% 
  summarise(num = n_distinct(label)) %>% 
  arrange(num) 
```

```{r}
#Networks with large, un-subfam-mixed clusters
# Largest cluster is unmixed?
# Where is the highest I, J, B unmixed cluster?

cluster_subfam_mixing.df <- sequence.df.melt %>% 
    filter(source == 'Sal1',
           subfam %in% c('C','E','G','I','K','J','B'),
           cluster_num > 0) %>% 
  select(-Id, -source, -expt) %>% 
  group_by(across(-label)) %>% 
  #With the subfams I want to see which networks have the largest, unmixed, with all subfam in as few clusters as possible
  summarise(num_members = n_distinct(label)) %>% 
  group_by(pid, lngth, inflation, subfam) %>% 
  mutate(sf_total_groups = n_distinct(cluster_num)) %>% 
  group_by(across(-c(subfam, num_members, sf_total_groups))) %>% 
  summarise(subfam_members = paste(paste0(subfam, ' (', num_members, ' - ', sf_total_groups, ')'), collapse = ', '),
            max_prop = max(num_members)/sum(num_members)) %>% 
  arrange(desc(cluster_size), desc(num_inNetwork))
```

```{r plotly}
df <- group_by(cluster_subfam_mixing.df, pid, lngth, inflation, num_inNetwork) %>%
  mutate(num_clust = n_distinct(cluster_num)) %>% 
  filter(cluster_num %in% c(1:15),
         num_inNetwork > round(4468*0.9)) %>% 
  group_by(num_inNetwork, num_clust, pid, lngth, inflation, 
           total_clusters, total_clusters_over2) %>% 
  summarise(avg_maxProps = sum(max_prop)/n_distinct(cluster_num), 
            max_clustSize = max(cluster_size)) %>% 
  arrange(desc(avg_maxProps), num_clust)

library(plotly)
plot_ly(df, x = ~avg_maxProps, y = ~total_clusters, size = ~num_inNetwork,
        symbol = ~pid, color = ~lngth, text = ~paste0(pid,lngth,inflation))
ggplot(df,
       aes(x = avg_maxProps, 
              y = as.integer(total_clusters), 
              # size = inflation,
              shape = factor(lngth), 
              color = factor(inflation)
           )) +
  scale_color_manual(values = safe_colorblind_palette) +
  geom_point() +
  theme_classic() +
  guides(color = guide_legend(ncol = 2),
         size = guide_legend(ncol = 2),
         shape = guide_legend(ncol = 2)) +
  facet_wrap(facets = vars(pid))
```

```{r}
cluster_subfam_mixing.df %>% 
  filter(grepl(subfam_members, pattern = '^J'))

cluster_subfam_mixing.df %>% 
  filter(grepl(subfam_members, pattern = '^I'))

cluster_subfam_mixing.df %>% 
  filter(grepl(subfam_members, pattern = '^K')) %>% arrange(desc(num_inNetwork))

lapply(c(25, 50, 75, 100, 150),
       function(length){
         filter(cluster_subfam_mixing.df, 
                pid == 50,
                lngth == length,
                inflation == 1.4) %>% 
           select(!c(pid, lngth, inflation))
       })

filter(cluster_subfam_mixing.df, 
       pid == 30,
       lngth == 25,
       inflation == 1.4) %>% 
  select(!c(pid, lngth, inflation))

```

50pid, 50/25 lngth appears to be the best now.

```{r read_data}

# node.df <- filter(sequence.df.melt,
#        pid == 30,
#        lngth == c(25)) %>% 
#   mutate(network_parameters = paste0(pid,'pid',lngth,'lngthI',inflation)) %>% 
#   dcast(label + Id + source + subfam + expt ~ network_parameters, value.var = 'cluster_num')
# 
# write.csv(node.df,
#           '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv',
#           row.names = FALSE,
#           quote = FALSE)

node.df <- read.csv(
  '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv')

node.df[apply(node.df[,str_detect(colnames(node.df), pattern = 'pid.+lngth')], 
              1,
              function(x) all(x == 0)),] %>% 
  filter(!label %in% notInNetwork) %>% 
  select(label, subfam)

```

```{r define_subfam}

#Define the sub-families of the clusters using the assigned Sal1 members.
# This process works by going through the sub-families present in each cluster,
#   then defining the overall cluster based on the Sal1 present/not present.
#   Thenn filtering to ensure that there's only the one assignment for each cluster.

 #If there are no Sal1 sequences in the cluster say 'unassigned'

cluster_subfam.df <- node.df %>% 
  ungroup %>% 
 #Set non-Sal1 to 'unassigned', so that clusters only made up of them are named such.
  mutate(subfam = ifelse(subfam %in% c('assembled','P01','Zhu16'),
                         'unassigned',
                         subfam)) %>% 
  dplyr::group_by(X30pid25lngthI1.4, subfam) %>% 
  dplyr::summarise(num = n_distinct(label)) %>% 
  group_by(X30pid25lngthI1.4) %>% 
  mutate(total_subfam = n_distinct(subfam),
         total_pirs = sum(num)) %>% 
  group_by(X30pid25lngthI1.4, total_pirs) %>% 
  #Only keep 'unassigned'/'Sal1' if it is the only subfamily present (i.e. entire cluster is unassigned).
  # Unless only other subfam in cluster is unassigned/Sal1, then I want this to be unassigned.
  filter(!(subfam %in% c('unassigned', 'Sal1') & total_subfam > 1) | all(subfam %in% c('unassigned','Sal1'))) %>% 
  #Calculate the percentage the subfams constitute of each subfam to see if any groups are split between sub-families.
  mutate(subfam_perc = round(100*num/sum(num), digits = 2)) %>% 
  #Keep the subfam name if it makes up over 50% of the number of subfamily members in the cluster.
  # This will now become the subfam assignment of the cluster.
  # If there's no dominating subfam then assign it as mixed.
  mutate(mixed = ifelse(all(100*num/sum(num) <= 50), 'mixed', 'dom')) %>% 
  summarise(subfam_num = paste0(subfam, ' (n = ', num, ')', collapse = '; '),
            subfam_clust = case_when(
              #Need convert the genes in no clusters into it's own noCluster group. 
              # (not part of the BLAST network/no alignments)
              # convert to integer and character because being a factor makes it tricky
              as.integer(as.character(X30pid25lngthI1.4)) == 0 ~ 'noCluster',
              100*num/sum(num) > 50 ~ subfam,
              mixed == 'mixed' ~ 'mixed',
              TRUE ~ ''
            )
  ) %>% 
  #Remove the rows with an empty subfam_clust (grep for a letter of the alphabet) 
  filter(grepl(subfam_clust, pattern = '[[:alpha:]]')) %>% 
  #Use unique to remove the redundant rows this creates for the 'mixed' samples
  unique %>% 
  mutate(clust_num = X30pid25lngthI1.4)

#Uniquify for numbering the subfam
cluster_subfam.df$subfam <- uniquify(cluster_subfam.df$subfam_clust, 
                           first_one = TRUE)
# Create column subfam_assignment to make code later easier since node.melt already has 'subfam' column
cluster_subfam.df$subfam_assignment <- uniquify(cluster_subfam.df$subfam_clust, 
                           first_one = TRUE)

# write.csv(cluster_subfam.df,
#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/cluster_subfam_30pid25lngthI1.4.csv',
#           row.names = FALSE,
#           quote = FALSE)

```

```{r write_edge.csv}

# blast_30pid25lngth.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
#            header = FALSE)
# 
# node_labels.df <- blast_30pid25lngth.df %>%
#   rowwise() %>%
#   mutate(label1 = node.df$Id[node.df$label %in% V1],
#          label2 = node.df$Id[node.df$label %in% V2]) %>%
#   select(V1,V2,label1, label2)
# 
# node_labels_noDup.df <- node_labels.df[!duplicated(apply(node_labels.df,
#                                                          sort,
#                                                          MARGIN =1),
#                                                    MARGIN=2),]
# #Already appears to have same2same matches absent.
# 
# node_labels_noDup_toSave.df <- select(node_labels_noDup.df, label1, label2)
# colnames(node_labels_noDup_toSave.df) <- c('Source', 'Target')

# write.csv(node_labels_noDup_toSave.df,
#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/edges_30pid25lngthNoS2S.csv',
#           row.names = FALSE,
#           quote = FALSE)

```

#Want to show the stages of development in this.

```{r expt_name}

#Extract the first two numbers of the name e.g. ERX2560125_ERR2541707
stage_geog.df <- bind_rows(
  lapply(
    assembly.df$assembly, 
    function(assembly_name){
      accession_name <- str_extract(assembly_name, 
                                    pattern = '(SR|ER).{8,9}_(SR|ER).{8,9}(?=_)')
      #If this works (not NA and only two elements) then find the expt name in the table
      if(!is.na(accession_name) & length(unlist(strsplit(accession_name, '_'))) == 2){
        cat_sample_stage <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$fastq_1, pattern = accession_name)]
        geog_source <- accession_table$Geography_source[grepl(accession_table$fastq_1, pattern = accession_name)]
      } else {
        #Get the first two elements of the name (separated by '_') - split then paste back together.
        expt_name <- paste(str_split_fixed(assembly_name, 
                                           pattern = '_', n = 3)[c(1,2)], collapse = '_')
        if(any(grepl(accession_table$Concatenated.file, 
                     pattern = expt_name))){
          #Otherwise this is probably a re-named concatenated file of samples, so
          # we need the 'Concatenated file' column.
          cat_sample_stage_all <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$Concatenated.file, 
                                                                            pattern = expt_name)]
          geog_source_all <- accession_table$Geography_source[grepl(accession_table$Concatenated.file, 
                                                                    pattern = expt_name)]
        } else {
          #Instead of being sensible I had a different naming scheme for the un-concatenated Kim and Gunalan files.
          cat_sample_stage_all <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$file_prefix, 
                                                                            pattern = expt_name)]
          geog_source_all <- accession_table$Geography_source[grepl(accession_table$file_prefix, 
                                                                    pattern = expt_name)]
        }

        cat_sample_stage <- unique(cat_sample_stage_all)
        geog_source <- unique(geog_source_all)

        if(rlang::is_empty(cat_sample_stage) | rlang::is_empty(geog_source)){
          stop(paste0('No life cycle stage or geographical source name for assembly ', assembly_name))
        }

        if(!length(cat_sample_stage) == 1){
          cat_sample_stage <- 'Mixed.stages' 
        }

        if(!length(geog_source) == 1){ 
          geog_source <- 'Mixed.sources'
        }

      }
      return(data.frame(assembly = assembly_name, stage = cat_sample_stage, geog_source = geog_source))
    })
)
```

# Spread of experiments

```{r expt_makeup_of_clusters}

node.melt <- node.df %>% 
  setDT %>% 
  melt(measure.vars = patterns('lngth'), 
       variable.name = 'mcl', 
       value = 'clust_num') %>% 
  mutate(mcl = as.character(mcl),
         expt = ifelse(expt == 'known',
                       paste0(source,'_known'),
                       expt_num.df$expt_code[match(expt, expt_num.df$expt)])) %>% 
  inner_join(., cluster_subfam.df, by = 'clust_num')

safe_colorblind_palette_expt <- c(expt_num.df$col[1:9],
                                  hcl.colors(3, palette = "Lajolla")[1:3])
names(safe_colorblind_palette_expt) <- c(expt_num.df$expt_code[1:9],
                                         unique(node.melt$expt)[
                                           grepl(unique(node.melt$expt), 
                                                 pattern = 'known')
                                         ])

lapply(unique(node.melt$mcl), 
       function(mcl_name){

         node.melt.df <- filter(node.melt, 
                                mcl == paste(mcl_name)) %>% 
           group_by(clust_num, mcl) %>% 
           mutate(clust_size = paste0(clust_num, ' (n=', 
                                     n_distinct(label), ')')) %>% 
           group_by(expt, clust_num, clust_size) %>% 
           filter(clust_num %in% 1:20) %>% 
           dplyr::summarize(num = n_distinct(label)) %>% 
           arrange(clust_num)

         node.melt.df$clust_size <- factor(node.melt.df$clust_size,
                                           levels = unique(node.melt.df$clust_size))
         node.melt.df$expt <- factor(
           node.melt.df$expt,
           levels = names(safe_colorblind_palette_expt)[
             names(safe_colorblind_palette_expt) %in% unique(node.melt.df$expt)
           ]
         )

         lapply(c('withKnown','onlyDeNovo'), 
                function(include){

                  if(include == 'onlyDeNovo'){
                    node.melt.df <- filter(node.melt.df, 
                                           !grepl(expt, pattern = 'known'))
                  } else {
                    #Removing the '_known' from the P01 etc names so that the legend fits better.
                    node.melt.df <- mutate(node.melt.df, 
                                           expt = str_remove(expt, pattern = '_known'))
                    names(safe_colorblind_palette_expt) <- str_remove(names(safe_colorblind_palette_expt),
                                                                      pattern = '_known')
                  }

                  plot <- ggplot(data = node.melt.df,
                                 aes(x = clust_size, 
                                     y = num, 
                                     fill = expt)) +
                    geom_col(position = 'fill') +
                    xlab('Cluster number') +
                    ylab('Proportion of assembled pirs\nin each cluster') +
                    scale_fill_manual(values = safe_colorblind_palette_expt[
                      names(safe_colorblind_palette_expt) %in% unique(node.melt.df$expt)
                    ],
                    guide = guide_legend(title = 'Experiment') )+
                    ggtitle(paste(mcl_name)) +
                    theme_classic() +
                    theme(axis.text.x = element_text(angle = 90,
                                                     vjust = 0.5))

                  if(global_save_var){
                    ggsave(plot = plot,
                           paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_exptComposition_',
                                  include,'_',mcl_name,'.pdf'),
                           device = 'pdf',
                           width = 7,
                           height = 5,
                           units = "in")
                  }

                  return(plot)
                })
       })
```

```{r expt_composition}
# How many clusters to include? e.g. 10 is the top ten largest clusters.
# Uses clust_num to subset this, so including 0 will include noCluster.
num_clusters <- 10
clusters_include <- filter(cluster_subfam.df, 
                           clust_num %in% 1:10)$subfam_assignment

#Get standardised colours 
top_subfam_col <- read.csv('top_subfam_col.csv')

node.melt.df <- filter(node.melt, 
                                mcl == 'X30pid25lngthI1.4',
                       expt != 'E05') %>% 
  group_by(subfam_assignment, mcl) %>% 
  mutate(clust_size = paste0(subfam_assignment, ' (n=', 
                             n_distinct(label), ')'),
         clust_numbers = n_distinct(label) ) %>% 
  group_by(expt) %>% 
  mutate(expt = paste0(expt, ' (n=', 
                       n_distinct(label), ')')) %>% 
  group_by(expt, subfam_assignment, clust_size, clust_numbers) %>% 
  filter(subfam_assignment %in% clusters_include) %>% 
  dplyr::summarize(num = n_distinct(label)) %>% 
  arrange(desc(clust_numbers))

node.melt.df <- mutate(node.melt.df, 
                       is_known = str_detect(expt, pattern = '_known'),
                       expt = str_remove(expt, pattern = '_known'))

node.melt.df$clust_size <- factor(node.melt.df$clust_size,
                                  levels = unique(node.melt.df$clust_size))
# node.melt.df$expt <- factor(
#   node.melt.df$expt,
#   levels = names(safe_colorblind_palette_expt)[
#     names(safe_colorblind_palette_expt) %in% unique(node.melt.df$expt)
#   ]
# )

color_palette <- c(
  hcl.colors(num_clusters, 
             palette = 'Spectral')[1:(num_clusters/2)],
  hcl.colors(num_clusters/2, 
             palette = 'viridis')
)
# %>% 
#   alternateReverse()

colours_subfam <- inner_join(node.melt.df, 
                             top_subfam_col, 
                             by = 'subfam_assignment')
top_subfam_col.vec <- setNames(unique(colours_subfam$colours),
                           unique(colours_subfam$clust_size))

( plot <- ggplot(data = node.melt.df,
                 aes(x = expt, 
                     y = num, 
                     fill = clust_size)) +
    geom_col(position = 'fill') +
    xlab('Experiment') +
    ylab(expression(paste('Proportion of ', 
                          italic(pir), 
                          's from each cluster'))) +
    scale_fill_manual(
      values = c( 
        top_subfam_col.vec
      ),
      guide = guide_legend(title = 'Cluster')
    ) +
    theme_classic() +
    facet_grid(~ is_known, scales = 'free_x', space = 'free') +
    theme(axis.text.x = element_text(angle = 90,
                                     vjust = 0.5),
          strip.text.x = element_blank(),
          strip.background = element_blank())
  )

if(global_save_var){
  ggsave(plot = plot,
         paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_clustComp_Expt.pdf'),
         device = 'pdf',
         width = 7,
         height = 8,
         units = "in")
}
```

```{r geog_source}
geog.df <- filter(node.melt, 
                       mcl == 'X30pid25lngthI1.4') %>%
  #Get the geographic source from the stage_geog.df. Need to extract the part of the assembly name without '_scapdes_0.95cdhitest'
  mutate(geog_source = stage_geog.df$geog_source[match(source, 
                                                       str_extract(stage_geog.df$assembly, 
                                                                           pattern = '.+(?=_scspades)'))] ) %>% 
  group_by(clust_num, mcl) %>% 
  mutate(clust_size = paste0(clust_num, ' (n=', 
                             n_distinct(label), ')')) %>% 
  group_by(expt) %>% 
  mutate(expt_size = n_distinct(label)) %>% 
  group_by(clust_num, clust_size, geog_source, expt, expt_size) %>% 
  filter(clust_num %in% 1:20,
         !is.na(geog_source)) %>% 
  dplyr::summarize(num = n_distinct(label),
                   num_norm = n_distinct(label)/expt_size
                   ) %>%
  arrange(clust_num) %>% 
  unique()
#For some reason calculating num_norm means we need to use unique

geog.df$clust_size <- factor(geog.df$clust_size,
                                           levels = unique(geog.df$clust_size))

ggplot(data = filter(geog.df, !clust_num %in% c(1,18)),
                                 aes(x = clust_size, 
                                     y = num_norm, 
                                     fill = geog_source)) +
                    geom_col(position = 'stack', col = 'black') +
                    xlab('Cluster number') +
                    ylab('Number of assembled pirs\nin each cluster normalized by total expt num') +
                    # scale_fill_manual(values = safe_colorblind_palette_expt[
                    #   names(safe_colorblind_palette_expt) %in% unique(node.melt.df$expt)
                    # ],
                    # guide = guide_legend(title = 'Experiment') )+
                    # ggtitle(paste(mcl_name)) +
                    theme_classic() +
                    theme(axis.text.x = element_text(angle = 90,
                                                     vjust = 0.5))
```

Generally the assembled pirs are pretty well spread out throughout the top 10 clusters.

## What's the largest cluster made up of 100% from one experiment?

```{r largest_100pc_cluster}
lapply(unique(node.melt$mcl), 
       function(mcl_name){
filter(node.melt, 
                mcl == paste(mcl_name)) %>% 
  group_by(clust_num) %>% 
  mutate(clust_num_num = sum(n_distinct(label))) %>% 
  group_by(expt, clust_num, clust_num_num, mcl) %>% 
  dplyr::summarize(prop = 100*round(n_distinct(label)/clust_num_num, digits = 3)) %>% 
   # Why do I need to call unique instead of summarise() doing it properly?
   unique %>% 
  arrange(desc(prop), desc(clust_num_num)) %>% 
  .[1:20,]
       })
```

The largest clusters 100% from one experiment are only five nodes in size. Cluster 25 (n=9) is entirely P01.

```{r clusters_dominated_by_expt}
lapply(unique(node.melt$mcl), 
       function(mcl_name){
         filter(node.melt, 
                mcl == paste(mcl_name)) %>% 
  group_by(clust_num) %>% 
  mutate(clust_num_num = sum(n_distinct(label))) %>% 
  filter(clust_num_num > 10) %>% 
  group_by(expt, clust_num, clust_num_num) %>% 
  dplyr::summarize(prop = 100*round(n_distinct(label)/clust_num_num, digits = 3)) %>% 
   # Why do I need to call unique instead of summarise() doing it properly?
   unique %>% 
  arrange(desc(prop), desc(clust_num_num)) %>% 
  .[1:20,]})
```

There are 14 clusters over ten nodes in size represented >=50% by only one experiment. Most of these are from the 'known' set of pre-existing pirs (P01, Sal1, Zhu16 original assembly).

## How much of each expt is found in the top N clusters?

```{r expt_num_across_clusters, eval = FALSE}
lapply(unique(node.melt$mcl), 
       function(mcl_name){
         node_mcl.df <- filter(node.melt, 
                               mcl == paste(mcl_name),
                               !grepl(expt, pattern = 'known')) %>% 
           mutate(expt = factor(expt)) %>% 
           group_by(expt) %>% 
           mutate(expt_num = sum(n_distinct(label)))
         num_clust <- length(unique(node_mcl.df$clust_num))
         expt_prop_across_clusters <- lapply(c(0:9,seq(10,num_clust + 5,1)),
                                             function(cluster_name){
                                               # Need to convert expt to a factor so that complete() later will add the missing values for large cluster number.
                                               # Also need a special case for when using 0 clusters, to get the lines to start at 100%.
                                               if(cluster_name != 0){node_mcl.df <- filter(node_mcl.df,
                                                                                     !clust_num %in% 1:cluster_name)} 
                                                 node_mcl.df %>%  
                                                 group_by(expt) %>% 
                                                 mutate(expt_num_noTen = sum(n_distinct(label))) %>% 
                                                 summarize(prop = 100*round(expt_num_noTen/expt_num, digits = 3)) %>% 
                                                 unique %>% 
                                                 ungroup %>%  
                                                 tidyr::complete(expt, fill = list(prop = 0)) %>% 
                                                 mutate(num_clusters = cluster_name)
                                             }) %>% bind_rows()

         plot <- ggplot(data = expt_prop_across_clusters,
                aes(y = prop, x = num_clusters, col = expt, linetype = expt)) +
           geom_step() + 
           scale_color_manual(values = safe_colorblind_palette_expt[
             names(safe_colorblind_palette_expt) %in% unique(expt_prop_across_clusters$expt)
           ],
           labels = names(safe_colorblind_palette_expt)[
             names(safe_colorblind_palette_expt) %in% unique(expt_prop_across_clusters$expt)
           ],
           guide = guide_legend(ncol = 2)) +
           scale_linetype_manual(values = c(1,1,1,1,1,3,3,2,2),
                                 labels = names(safe_colorblind_palette_expt)[
                                   names(safe_colorblind_palette_expt) %in% unique(expt_prop_across_clusters$expt)
                                 ]) +
           theme_classic()

         if(global_save_var){
           ggsave(plot = plot,
                  paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_exptStep_'
                         ,mcl_name,'.pdf'),
                  device = 'pdf',
                  width = 7,
                  height = 5,
                  units = "in")
         }

         return(plot)
       })

```

# Spread of sub-family with clusters

Note that some clusters are missing because no Sal1 is present within them e.g. 15.

```{r subfam_makeup_of_clusters}

cbf_1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73", 
           "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
names(cbf_1) <- c("B","C","E","I","G","J","K", "unassigned")
cbf_1['unassigned']<- 'black'

lapply(unique(node.melt$mcl), 
       function(mcl_name){
         mcl_num_subfamNum.df <- filter(node.melt, 
                                        mcl == paste(mcl_name)) %>% 
           #Make clust_num numeric for correct ordering by number.
           mutate(clust_num = as.numeric(clust_num),
                  subfam = ifelse(subfam == 'Sal1', 
                                  'unassigned',
                                  subfam)) %>% 
           group_by(subfam, clust_num) %>% 
           filter(clust_num %in% 1:25, 
                  source == 'Sal1') %>%
           dplyr::summarize(num = n_distinct(label))

         plot <- ggplot(data = mcl_num_subfamNum.df,
                        aes(x = clust_num, y = num, fill = subfam)) +
           geom_col(position = 'stack') +
           scale_fill_manual(values = cbf_1,
                             guide = guide_legend(title = 'Sal1 Sub-family')) +
           #Set scale_x_cont here so that all cluster numbers are shown.
           scale_x_continuous(breaks = 1:25) +
           scale_y_continuous(n.breaks = 10) +
           xlab('Cluster number') +
           ylab('Number of pirs') +
           ggtitle(mcl_name) +
           theme_classic()

         if(global_save_var){
           ggsave(plot = plot,
                  paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_sal1Composition_',
                         mcl_name,'.pdf'),
                  device = 'pdf',
                  width = 7,
                  height = 5,
                  units = "in")
         }

           return(plot)
       })
```

```{r pir_sizes}

node_length.melt <- filter(node.melt, 
       mcl == 'X30pid25lngthI1.4') %>% 
  rowwise() %>% 
  mutate(orf_length = length(getElement(fasta, label)))

#Can remove all ORFs under 200aa to test how the short 'known' pirs impacts the results.
# Spoiler: They don't.
# %>% 
#   filter(orf_length >= 200)

#Select first 11 clusters and remove the short Zhu16 seqeunces.
# Order by median length and factor so that ggplot will show in this order.
node_length_clust1.11_noZhu16.melt <- filter(node_length.melt, 
                                             clust_num %in% 1:11, 
                                             expt != 'Zhu16_known') %>% 
  group_by(clust_num) %>% 
  mutate(orf_length_med = median(orf_length)) %>% 
  arrange(desc(orf_length_med)) %>% 
  ungroup() %>% 
  # ungroup() so that factor works properly and doesn't sort clust_num
  mutate(clust_num = factor(clust_num, 
                            levels = unique(clust_num)))

node_length_clust1.11_noZhu16.melt %>% 
  group_by(clust_num) %>% 
  summarise(med_length = median(orf_length))

(plot_orf <- ggplot(node_length_clust1.11_noZhu16.melt, 
                    aes(x = clust_num,
                        y= orf_length)) +
    geom_boxplot() +
    xlab('Cluster number') +
    ylab('Length of ORFs (aa)') +
    theme_classic() +
    theme(axis.text.x = element_text(size = 13),
          axis.text.y = element_text(size = 13)) )

if(global_save_var){
  ggsave(plot = plot_orf,
         '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_ORFlength.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}

```

```{r orfLengths}
ggplot(filter(node_length.melt, clust_num %in% c(1,12:20), 
              expt != 'Zhu16_known'), 
       aes(x = factor(clust_num),
           y= orf_length)) +
  geom_boxplot()
ggplot(filter(node_length.melt, clust_num %in% 1:11), 
       aes(x = factor(clust_num),
           y= orf_length)) +
  geom_boxplot() + facet_wrap(~expt)

ggplot(filter(node_length.melt, clust_num %in% 1:11), 
       aes(x = factor(expt),
           y= orf_length)) +
  geom_boxplot()

lapply(unique(node_length.melt$expt), function(expt){
  lengths <- node_length.melt$orf_length[node_length.melt$expt == expt]
  if(length(lengths) > 2){
    hist(lengths, 
         breaks = 100,
         main = expt,
         xlim = c(0,900))
  }
})
lapply(1:11, function(clust_num){
  lengths <- node_length.melt$orf_length[node_length.melt$clust_num == clust_num]
  if(length(lengths) > 2){
        hist(lengths, 
         breaks = 100,
         main = paste0('Cluster Number ', clust_num),
         xlim = c(0,900))
  }
})
```

Cluster 8 particularly long?

Looks like there's a little group of longer pirs for some experiments/P01/Sal1.
Like over 800aa - pretty much all in Cluster 1 (E). 
I wonder if this is the two groups in the Gephi network.

Cluster 7 and 9 low variability, potentially shorter?

```{r kwtest_length}
#K-W test of if there is a signif difference between the clusters for length

kruskal.test( node_length.melt$orf_length, factor(node_length.melt$clust_num))
kruskal.test(node_length_clust1.11_noZhu16.melt$orf_length,
             factor(node_length_clust1.11_noZhu16.melt$clust_num))

clust_comp <- combn(1:11, m = 2, simplify = FALSE)

#Do the tests for each combination, combione into a data frame and use p.adjust multiple testing adjustment.
( length_tests.df <- lapply(c('greater', 'less'), 
                            function(wilcox_text_alt){
                              lapply(clust_comp, function(comparison){
  test.stat <- wilcox.test(orf_length ~ clust_num,
                           data = filter(node_length.melt,
                                         expt != 'Zhu16_known'),
                           subset = clust_num %in% comparison,
                           alternative = wilcox_text_alt)
  return(data.frame(comparison_1 = comparison[1],
                    comparison_2 = comparison[2],
                     stat = test.stat$statistic,
                     pval = test.stat$p.value,
                    wilcox_text.alt = wilcox_text_alt))
                              }) %>% bind_rows
}) %>% 
  bind_rows %>% 
  mutate(padj = round(p.adjust(pval), digits = 5),
         pval = round(pval, digits = 5)) %>% 
  arrange(desc(padj)) )

length_med.df <- node_length.melt %>% 
  group_by(clust_num) %>% 
  summarise(orf_length_med = median(orf_length))

length_tests.mat <- mutate(length_tests.df,
                           orf_ratio = ifelse(padj <= 0.05, 
                                              length_med.df$orf_length_med[
                                                match(comparison_1, 
                                                      length_med.df$clust_num)
                                              ]/length_med.df$orf_length_med[
                                                match(comparison_2, 
                                                      length_med.df$clust_num)
                                              ], 
                                              NA)) %>% 
  dcast(.,
        comparison_1 ~ comparison_2,
        value.var = 'orf_ratio')
rownames(length_tests.mat) <- length_tests.mat[,1]
length_tests.mat <- length_tests.mat[,-1]

# nams <- unique(unlist(clust_comp))
# 
# #Need to add the first and last rows due to combn not using 1 as combination_2 and 11 as combination_1.
# one.mat <- matrix(nrow = nrow(length_tests.mat),
#        dimnames = list(rep(NA, length.out = nrow(length_tests.mat)), first(nams))) 
# length_tests.mat <- cbind(one.mat, length_tests.mat)
# 
# last.mat <- matrix(ncol = ncol(length_tests.mat),
#        dimnames = list(last(nams), rep(NA, 
#                                        length.out = ncol(length_tests.mat))))
# colnames(last.mat) <- names(length_tests.mat)
# length_tests.mat <- rbind(length_tests.mat, last.mat)
# 
# library(ComplexHeatmap)
# library(viridis)
# #Make matrix symmetrical
# length_tests.mat[
#   lower.tri(length_tests.mat)
# ] <- 1/(t(length_tests.mat)[
#   lower.tri(length_tests.mat)
# ])
# 
# colnames(length_tests.mat) <- rownames(length_tests.mat) <- nams
# 
# Heatmap(length_tests.mat, 
#         cluster_columns = FALSE, 
#         cluster_rows = FALSE,
#         na_col = 'black',
#         col = circlize::colorRamp2(c(0, 1, 1.4), c('blue', 'white', 'red')))

length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '^1$|3|8|10'))) %>%
  arrange(desc(padj))

length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '7'))) %>%
  arrange(desc(padj))
length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '9'))) %>%
  arrange(desc(padj))
```

```{r where_ancestral}

outfmt6_headers <- 
  'qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore qcovs qcovhsp slen nident' %>% 
  strsplit(split = ' ') %>% 
  unlist

  #Read in BLAST file 
blast_30pid25lngth.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
           col.names = outfmt6_headers) %>% 
  mutate(
    scovhsp = round( (nident/slen * 100), digits = 2),
    scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
    evalue = format(evalue, scientific = TRUE)
  )

#Ancestral: PVP01_1147700, PVX_113230
#Vir14:

node.melt %>% 
  filter(mcl == 'X30pid25lngthI1.4',
         label %in% c('PVP01_1147700', 'PVX_113230', 'PVX_108770'))

#All in cluster 4

ancestral_blast <- blast_30pid25lngth.df %>% 
  filter(qseqid %in% 
           c('PVP01_1147700', 'PVX_113230') | sseqid %in% 
           c('PVP01_1147700', 'PVX_113230'))
hist(ancestral_blast$pident, breaks = 100)
hist(ancestral_blast$scovhsp, breaks = 100)
hist(ancestral_blast$qcovhsp, breaks = 100)

ancestral_blast_filter <- filter(ancestral_blast,
                          qcovhsp > 50,
                          scovhsp > 50,
                          nident > 80) 
# ancestral_blast_filter <- filter(ancestral_blast, 
#                                  grepl(qseqid, pattern = 'Zhu') | grepl(sseqid, pattern = 'Zhu')) %>% 
#   arrange(pident)
ancestral_blasters <- unique(c(ancestral_blast_filter$qseqid, ancestral_blast_filter$sseqid))

node.df$ancestral_blast <- ifelse(node.df$label %in% ancestral_blasters,
                                  TRUE,
                                  FALSE)
filter(node.df, ancestral_blast) %>% 
  mutate(clust_num = X30pid25lngthI1.4) %>% 
  select(!contains('lngth'))

```

```{r}
vir14_blast <- blast_30pid25lngth.df %>% 
  filter(qseqid %in% 
           c('PVX_108770') | sseqid %in% 
           c('PVX_108770'))
hist(vir14_blast$pident, breaks = 100)
hist(vir14_blast$scovhsp, breaks = 100)
hist(vir14_blast$qcovhsp, breaks = 100)

vir14_blast_filter <- filter(vir14_blast,
                          qcovhsp > 50,
                          scovhsp > 50,
                          nident > 80) 
# vir14_blast_filter <- filter(vir14_blast, 
#                                  grepl(qseqid, pattern = 'Zhu') | grepl(sseqid, pattern = 'Zhu')) %>% 
#   arrange(pident)
vir14_blasters <- unique(c(vir14_blast_filter$qseqid, vir14_blast_filter$sseqid))

node.df$vir14_blast <- ifelse(node.df$label %in% vir14_blasters,
                                  TRUE,
                                  FALSE)
filter(node.df, vir14_blast) %>% 
  mutate(clust_num = X30pid25lngthI1.4) %>% 
  select(!contains('lngth'))

```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxBUSCO.Rmd
Add many required vivax files and change paths in vivax scripts

@@
 library('readr')
-library('dplyr', 
-        lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0")
+library('dplyr')
 library('stringr')
@@
 
-```{r analyse_busco}
-samplesheet <- read.delim("/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samplesheet.txt")
-
-samplesheet_busco.df <- lapply(unique(samplesheet$assembly), 
-                               function(assembly.file){
-                                 # The tenth element in the file path string corresponds to the assembly names that I used.
-                                 assembly <- strsplit(assembly.file, split = '/')[[1]][10]
-                                 if(!is.na(assembly)){
-                                   busco.file <- paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/busco/',
-                                                        assembly)
-                                   #Skip 2 in the read_tsv function as the first two lines are not part of the data frame.
-                                   busco_api.df <- read_tsv(paste0(busco.file, 
-                                                                   '_apicomplex/run_apicomplexa_odb10/full_table.tsv'),
-                                                            skip = 2,
-                                                            show_col_types = FALSE)
-                                   busco_api.complete <- sum(busco_api.df$Status == "Complete")
-                                   busco_api.fragment <- sum(busco_api.df$Status == "Fragmented")
-                                   busco_api.duplicate <- sum(n_distinct(
-                                     busco_api.df$`# Busco id`[busco_api.df$Status == "Duplicated"]
-                                   ))
-
-                                   busco_plasmo.df <- read_tsv(paste0(busco.file, 
-                                                                      '_plasmodium/run_plasmodium_odb10/full_table.tsv'),
-                                                               skip = 2,
-                                                               show_col_types = FALSE)
-                                   busco_plasmo.complete <- sum(busco_plasmo.df$Status == "Complete")
-                                   busco_plasmo.fragment <- sum(busco_plasmo.df$Status == "Fragmented")
-                                   busco_plasmo.duplicate <- sum(n_distinct(
-                                     busco_plasmo.df$`# Busco id`[busco_plasmo.df$Status == "Duplicated"]
-                                   ))
-
-                                   data.frame(
-                                     assembly = assembly.file,
-                                     assembly_name = assembly,
-                                     Experiment = samplesheet$Experiment[match(assembly.file, samplesheet$assembly)],
-                                     num_busco_api_complete = busco_api.complete,
-                                     perc_busco_api_complete = round(100*busco_api.complete/446,
-                                                                     digits = 3),
-                                     num_busco_api_fragment = busco_api.fragment,
-                                     perc_busco_api_fragment = round(100*busco_api.fragment/446,
-                                                                     digits = 3),
-                                     num_busco_api_duplicate = busco_api.duplicate,
-                                     perc_busco_api_duplicate = round(100*busco_api.duplicate/446,
-                                                                      digits = 3),
-                                     perc_busco_api_nonMissing = round(100*(
-                                       busco_api.fragment+busco_api.complete+busco_api.duplicate)/446,
-                                       digits = 3
-                                     ),
-
-                                     num_busco_plasmo_complete = busco_plasmo.complete,
-                                     perc_busco_plasmo_complete = round(100*busco_plasmo.complete/3642,
-                                                                        digits = 3),
-                                     num_busco_plasmo_fragment = busco_plasmo.fragment,
-                                     perc_busco_plasmo_fragment = round(100*busco_plasmo.fragment/3642,
-                                                                        digits = 3),
-                                     num_busco_plasmo_duplicate = busco_plasmo.duplicate,
-                                     perc_busco_plasmo_duplicate = round(100*busco_plasmo.duplicate/3642,
-                                                                         digits = 3),
-                                     perc_busco_plasmo_nonMissing = round(100*(
-                                       busco_plasmo.fragment+busco_plasmo.complete+busco_plasmo.duplicate)/3642,
-                                       digits = 3
-                                     )
-                                   )
-                                 } else { return(NULL) }
-                               }) %>% bind_rows
-```
-
-```{r plotly}
+```{r read_busco}
+samplesheet <- read.delim("samplesheet.txt")
 
 #Get the number of pirs stats
-assembly_stats.df <- read.table('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info_extra.txt')
+assembly_stats.df <- read.table('tl22-01_assembly_info_extra.txt')
 
@@
 
-assembly_stats_busco <- inner_join(assembly_stats.df, 
-                                   mutate(samplesheet_busco.df, 
-                                          assembly_file = assembly), 
-                                   by = 'assembly_file')
-# write.table(assembly_stats_busco,
-#             '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info_extra_busco.txt')
+assembly_stats_busco <- read.table('tl22-01_assembly_info_extra_busco.txt')
 
 #Include '.+(?=0.95)' pattern for searching as Gunalan has multiple cdhitest threshold assemblies.
-samplesheet_busco.df <- mutate(rowwise(samplesheet_busco.df), 
+samplesheet_busco.df <- mutate(rowwise(assembly_stats_busco), 
                                num_pir_predict = assembly_stats.df$num_uniq_mRNA_ORFs_over200aaLength[
@@
 )
+```
 
+```{r plotly}
 hline <- function(y = 0, color = "blue") {
@@
 ggsave(plot = plot_buscos,
-       '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/assemblies_buscoPlasmo-buscoApi-pirNum.pdf',
+       'plots/assemblies_buscoPlasmo-buscoApi-pirNum.pdf',
        device = 'pdf',

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxSalmonAlignment.Rmd
Add many required vivax files and change paths in vivax scripts

@@
 
-This script analyses the results of a _de novo_ assembly of Plasmodium vivax RNAseq data.
+This script reads the Salmon files and assigns them cluster numbers from the networks.
 
@@
 ```{r load_data}
-pir.info <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/PlasmoDBv56_VvxPirs_P01_Sal1.csv', 
+pir.info <- read.csv('PlasmoDBv56_VvxPirs_P01_Sal1.csv', 
                      header = TRUE)
@@
 ```{r de_novo_pir_list}
-# pir_lists.dir <- "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/"
-# pir_lists.files <- list.files(pir_lists.dir, 
-#                               pattern = '0.95cdhitest_hmmerPfam-PlasmodiumVir.list',
-#                               full.names = TRUE) %>% 
-#   #Get those with file sizes above zero (no empty lists)
-#   # file.info gets the file sizes
-#   .[file.info(., extra_cols = FALSE)$size > 0]
-# 
-# # Read the pirs into a list (as a vector).
-# # sapply through the file list, read in the table and subset with the automatic column name V1
-# pir_list.vec <- unlist(
-#   sapply(pir_lists.files, 
-#          function(pir_list){
-#            #str_removing the _1 at the end of the removal 
-#            str_remove(read.table(pir_list)$V1, 
-#                       pattern =  '_[[:digit:]]$')
-#          })
-# )
-
 #Need to read in the >200aa ORF length  pirs
-pir_list.vec <- str_extract(names(read.fasta('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta')),
+pir_list.vec <- str_extract(names(read.fasta('G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta')),
                                   pattern = 'NODE.+(?=_[[:digit:]])')
@@
 ```{r load_cluster_data}
-node.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv') %>% 
+node.df <- read.csv('sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv') %>% 
   #Convert the cluster numbers to a factor as they are discrete categories.
@@
 ```{r define_subfam}
-cluster_subfam.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/cluster_subfam_30pid25lngthI1.4.csv')
+cluster_subfam.df <- read.csv('cluster_subfam_30pid25lngthI1.4.csv')
 
@@
 #Get the full file paths of the salmon quant.sf files.
-quant_files <- list.files(path = "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/salmon/all_files",
+quant_files <- list.files(path = "salmon/all_files",
                           recursive = TRUE,
@@
 #Load the samplesheet.txt file of metadata, add new column for salmon files
-samplesheet <- read.delim("/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samplesheet.txt") %>% 
+samplesheet <- read.delim("samplesheet.txt") %>% 
   mutate(salmon_quant = '')
@@
 
-'noCluster' means that the pir sequence was not in the '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/gephi/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/nodes_mcl_50pid100lngthNoSame2Same.csv' file.
+'noCluster' means that the pir sequence was not in the 
+'nodes_mcl_50pid100lngthNoSame2Same.csv' file.
 I think this means that they didn't have any BLAST similarity to other pir to the thresholds:
@@
                     write.csv(expt_quant[[expt]][[assembly]],
-                              file = paste( "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/quantFiles_R",
+                              file = paste( "quantFiles_R",
                                             expt, assembly,

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxSalmonAnalysis.Rmd
Add many required vivax files and change paths in vivax scripts

@@
 # Load the required packages
-library('dplyr', 
-        # lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0"
-        )
+library('dplyr')
 library('readr')
@@
 
-samplesheet <- read.delim("/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samplesheet.txt")
+samplesheet <- read.delim("samplesheet.txt")
 
@@
 
-expt_num.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/expt_codes.csv')
+expt_num.df <- read.csv('expt_codes.csv')
 
@@
                                                    quant_file <- paste(
-                                                     "/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/quantFiles_R",
+                                                     "quantFiles_R",
                                                      expt, assembly, 
@@
 ```{r pir_diversity_v_BUSCOs}
-buscos <- read.table('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info_extra_busco.txt')
+buscos <- read.table('tl22-01_assembly_info_extra_busco.txt')
 
@@
                        filename = paste0(
-                         '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/salmonQuant_plots/tl22-01_',
+                         'plots/salmonQuant_plots/tl22-01_',
                          expt, '_', absolute_or_norm, '_', assembly,
@@
     plot = legend.plot,
-    filename = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/salmonQuant_plots/tl22-01_legend_quantSalmon_stacked_subfam30pid25lngthI14.pdf',
+    filename = 'plots/salmonQuant_plots/tl22-01_legend_quantSalmon_stacked_subfam30pid25lngthI14.pdf',
     width = 7,
@@
 
-G18G19_hmmerVir200aa.fa <- read.fasta('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta')
+G18G19_hmmerVir200aa.fa <- read.fasta('G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta')
 
@@
 #             names(G18G19_hmmerVir200aa_tpmOver1.fa),
-#             file.out = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')
+#             file.out = 'G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')
 
@@
 ```{r Num-pir-transcript_v_Num-transcript-total}
-assembly_info.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt') %>% 
+assembly_info.df <- read.delim('tl22-01_assembly_info.txt') %>% 
   filter(!grepl(assembly, pattern = '0\\.9c|0\\.99'),
@@
 # write.csv(expt_num.df,
-#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/expt_codes.csv',
+#           file = 'expt_codes.csv',
 #           row.names = FALSE)
@@
   ggsave(plot = plot,
-         '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/assemblies_numPirTranscripts_totalTranscripts.pdf',
+         'plots/assemblies_numPirTranscripts_totalTranscripts.pdf',
          device = 'pdf',
@@
 
-ISSUE: Why is the number of sequences in '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta' larger than sum(assembly_info_samplesheet.df£$num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa)
+ISSUE: Why is the number of sequences in 'G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta' larger than sum(assembly_info_samplesheet.df£$num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa)
 

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxTPMover1.Rmd
Add many required vivax files and change paths in vivax scripts

---
title: "tl22-01_VvxTPMover1"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

# Intro

This script analyses the results of a _de novo_ assembly of Plasmodium vivax RNAseq data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) 
# Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr')
library('stringr')
library('purrr')
library('kableExtra')
library('plotly')
library('rmarkdown')
library('seqinr')

safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
global_save_var <- FALSE
```

```{r function_ordering}
ordering <- function(to_order, order_vector, fromLast = TRUE){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = fromLast)
}
```

```{r function_uniquify}
#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
      name <- paste(vector_names[i],
                    df_num$rep[match(vector_names[i], df_num$names)],
                    sep = '_')
      #Increase rep counter
      df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
      new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r read_200aa_pirs}
G18G19_hmmerVir200aa.fa <- read.fasta('G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta')

#Need to read in the >200aa ORF length  pirs
pir_list.vec <- str_extract(
  names(G18G19_hmmerVir200aa.fa),
  pattern = 'NODE.+(?=_[[:digit:]])'
)
```

```{r load_salmon}
#Get the full file paths of the salmon quant.sf files.
quant_files <- list.files(path = "salmon/all_files",
                          recursive = TRUE,
                          pattern = 'quant.sf',
                          full.names = TRUE)

#Load the samplesheet.txt file of metadata, add new column for salmon files
samplesheet <- read.delim("samplesheet.txt") %>% 
  mutate(salmon_quant = '')

#Make a table of the files and the sample metadata from samplesheet.txt
# for loop over all the quant_files that exist
for(quant_name in quant_files){
  #Looking for the sample prefix in the file path between 'all_files' and '{prefix}_quant'
  prefix <- str_extract(quant_name, 
                        pattern = '(?<=\\/all_files/).+(?=/.+_quant)')
  if(nrow(samplesheet[grep(samplesheet$prefix, 
                   pattern = prefix),]) > 1){
                     stop(paste0(prefix,
                                 " matches more than one row of the samplesheet"))
                   }
  samplesheet[grep(samplesheet$prefix, 
                   pattern = prefix),'salmon_quant'] <- quant_name
}

sum(grepl(samplesheet[,'salmon_quant'], pattern = 'camp')) == 247
```

```{r list_expt_quant}
expt_quant <- lapply(unique(samplesheet$Experiment),
                     #For each experiment name
                     function(expt){
                       print(paste0('Starting ', expt))
                       #Get the experimental metadata by subsetting
                       expt_data <- samplesheet[samplesheet$Experiment == expt,]
                       if(!sum(grepl(expt_data$assembly, pattern = '.')) == 0){
                         #If sum of grepping for any character is not zero, do the analysis 
                         # i.e there are assembly names and probably assemblies/quant files
                         # For each individual assembly 
                         # (we can bind rows for samples that use the same assembly)
                         assembly_quants <- lapply(unique(expt_data$assembly), 
                                                   function(assembly){
                                                     #Get the prefix(es) that were quantified against this assembly
                                                     prefix <- expt_data[expt_data$assembly == assembly, 
                                                                         'prefix']
                                                     #Get the quant file name(s) and read it in
                                                     quant_file <- expt_data[expt_data$prefix %in% prefix,
                                                                             'salmon_quant']

                                                     #If the file name isn't zero length (i.e. there is no name)
                                                     if(!length(quant_file) == 0){
                                                       #If length is one then we can make the quant file for just one sample 
                                                       if(length(prefix) == 1){
                                                         quant <- read.delim(quant_file)
                                                         #Subset the Name and TPM columns and keep it as a data frame (drop = F)
                                                         quant.df <- quant[,c('Name','TPM'), drop = FALSE]

                                                         # Give a warning if the TPM values are all zero.
                                                         if(sum(as.double(quant$TPM)) == 0){
                                                           #Do not return a quant.df data.frame, making it nothing:
                                                           quant.df <- ''
                                                           print(paste0(prefix, 
                                                                        " associated file has recorded no TPM values"))
                                                           #empty return to end the lapply for this file
                                                           return()
                                                         }

                                                       } else {
                                                         #If longer than 1 so there's more than one quant file for this assembly
                                                         quant.rlist <- lapply(quant_file, read.delim)
                                                         #Use purrr reduce to bind the df in the list together with inner_join,
                                                         # combining by the 'Name' column which should be the same throughout.
                                                         quant.df <- purrr::reduce(quant.rlist, 
                                                                                   inner_join, 
                                                                                   by = 'Name')
                                                         quant.df <- quant.df[grepl(colnames(quant.df), 
                                                                                    pattern = 'Name|TPM')]

                                                         # Give a warning if the TPM values are all zero.
                                                         if(any(sapply(quant.df[grepl(colnames(quant.df), 
                                                                                      pattern = 'TPM')], 
                                                                       function(x) sum(as.double(x))) == 0)){
                                                           print(paste0("At least some of ", 
                                                                        expt, 
                                                                        " results have recorded no TPM values"))
                                                           return()
                                                         }

                                                       }

                                                       #Put the sample name (prefix) in the TPM column
                                                       colnames(quant.df)[grepl(colnames(quant.df), 
                                                                                pattern = 'TPM')] <- paste0(prefix,
                                                                                                            '_TPM')

                                                       # Mark the pir NODES as such (TRUE in the 'pir' column)
                                                       quant.df$pir <- quant.df$Name %in% pir_list.vec
                                                       print(paste0('Number of pirs = ', sum(quant.df$pir)))
                                                       return(quant.df)
                                                     }
                                                   })

                         names(assembly_quants) <- str_extract(unique(expt_data$assembly),
                                                               pattern = '(?<=\\/cdhitest/).+(?=/.+fasta)')
                         print(paste0('Finished ', expt))
                         return(assembly_quants)  
                       }
                     }
)

names(expt_quant) <- unique(samplesheet$Experiment)
```

```{r node-df_tpmvalues}

# Add the TPM values to the node.df dataset alongside a threshold expressed/not expressed

# Get a list of all the de novo transcripts and their max TPM. 
#  Shouldn't be any with the same name unless by exceptional chance?
pir_quant <- lapply(names(expt_quant), 
                     function(expt.name) {
                       expt.list <- expt_quant[[expt.name]]
                       # print(names(expt.list))
                       lapply(names(expt.list), 
                              function(quant.name) {
                                quant.df <- expt.list[[quant.name]]
                                if(!is.null(quant.df)){
                                  #Melt the dataframe to calculate the max TPM value for each de novo transcript
                                  # across all samples.
                                  #This may be faster without melting and using apply.
                                  filter(quant.df, pir) %>% 
                                    reshape2::melt(.,
                                                   id.vars = c('pir','Name'),
                                                   measure.vars = str_subset(colnames(.), 
                                                                             pattern = 'TPM'),
                                                   value.name = 'tpm') %>% 
                                    group_by(Name, pir) %>% 
                                    summarise(max_tpm = max(tpm)) %>% 
                                    mutate(label_no_end = paste(quant.name, Name, sep = '_'),
                                           expt_assembly = quant.name, 
                                           expt = expt.name)
                                }
                              }) %>% 
                         bind_rows
                     }) %>% 
  bind_rows

# sum(duplicated(pir_quant$Name))

write.csv(pir_quant,
          'pir_maxTPM.csv')

```

```{r pirRemoved}
pir_quant %>% 
  group_by(expt_assembly) %>% 
  summarise(num_over1TPM = sum(max_tpm >= 1),
            num_total = n_distinct(Name),
            perc_over1TPM = round(100*sum(max_tpm >= 1)/n_distinct(Name), digits = 2),
            med_max_over1tpm = median(max_tpm[max_tpm >= 1])) %>% 
  arrange(desc(perc_over1TPM))
```

```{r over1tpm_listFiles}
#Note that node_tpm.df includes none-quantified genes, like P01/Sal1/Zhu16 and the Rangel genes.
# Will use over or equal to 1 TPM in at least one sample (max_tpm > 1) as the threshold for active transcription.
expressed_pirTranscripts <- filter(pir_quant, max_tpm>=1)

paste0('Number of transcribed pirs = ',
       nrow(expressed_pirTranscripts),
       ' and number of non-transcribed pirs = ',
       nrow(pir_quant) - nrow(expressed_pirTranscripts))

names_no_end <- str_remove(names(G18G19_hmmerVir200aa.fa), 
                           pattern = '_[[:digit:]]{1,2}$')

#These aren't all unique, I guess this doesn't impact anything later.

#Subset the fasta file by the pirs which are actively transcribed.
# Note that this may be greater than expected due to the fact that there may be multiple pir ORFs for one expressed pir transcript.
# Also it is smaller than the number of expressed_pirTranscripts because not all of the expressed_pirTranscripts have a >200aa ORF.
# Note that the pirs are defined as such based on the *_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir.list files (so before 200aa filtering).

G18G19_hmmerVir200aa_tpmOver1.fa <- G18G19_hmmerVir200aa.fa[
  names(G18G19_hmmerVir200aa.fa)[names_no_end %in% expressed_pirTranscripts$label_no_end]
  ]

#Duplicates among these ORFs/transcripts.
# duplicates <- sapply(expressed_pirTranscripts$label_no_end,
#        function(x) names(G18G19_hmmerVir200aa.fa)[names_no_end %in% x] )
# duplicates[lapply(duplicates, length) > 1]

print(paste0('Number of ORFs (>200aa ORF tpm over 1) = ',
      length(G18G19_hmmerVir200aa_tpmOver1.fa)))
print(paste0('Number of unique transcripts (>200aa ORF tpm over 1) = ',
      length(unique(str_extract(names(G18G19_hmmerVir200aa_tpmOver1.fa), 
                   pattern = '.+(?=_[[:digit:]]$)')))))

table(str_extract(names(G18G19_hmmerVir200aa_tpmOver1.fa), pattern ='.+(?=_NODE)'))

#Save the fasta file of the over 1 TPM pirs.
# write.fasta(G18G19_hmmerVir200aa_tpmOver1.fa,
#             names(G18G19_hmmerVir200aa_tpmOver1.fa),
#             file.out = 'G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')

# sum(expressed_pirTranscripts$Name == 'NODE_2407_length_2160_cov_4.64561')

```

The percentage of overall transcripts and pir transcripts transcribed over 1 TPM for each sample.

```{r num_over_1}
# Count the number/percentage of de novo transcripts with an expression level over 1 TPM
over1.df <- lapply(names(expt_quant), 
       function(expt){
         expt_df.list <- lapply(names(expt_quant[[expt]]), 
                function(assembly){
                  if(!is.null(expt_quant[[expt]][[assembly]])){
                    df <- expt_quant[[expt]][[assembly]]
                    #reshape2::melt the data frame to put each sample (the measure.vars) on a separate row.
                    reshape2::melt(df, 
                         id.vars = c('Name','pir'),
                         measure.vars = str_subset(colnames(df), 
                                                   pattern = 'TPM'),
                         variable.name = 'sample',
                         value.name = 'tpm') %>% 
                      mutate(num_transcripts = n_distinct(Name),
                             num_pirs = n_distinct(Name[pir])) %>% 
                      group_by(sample, pir, 
                               num_pirs, num_transcripts
                      ) %>% 
                      summarise(over1_perc = unique(round(100*sum(tpm >= 1)/num_transcripts, 
                                                          digits = 2)),
                                over1_num = sum(tpm >= 1),
                                #For the pir percantage if it is over 100% (i.e. when this is
                                # calculating for all the transcripts total) then make it NA
                                over1_perc_pirs = ifelse(unique(sum(tpm >= 1)/num_pirs <= 1),
                                                         round(unique(100*sum(tpm >= 1)/num_pirs),
                                                               digits = 2),
                                                         NA),
                                total = n_distinct(Name)) %>% 
                      mutate(assembly = paste(assembly))
                  }
                })
         #Add the experiment as a row for binding all these together
         expt_df <- bind_rows(expt_df.list) %>% 
           mutate(expt = expt)
       }) %>% 
  #bind_rows together across all the experiments and assemblies.
  bind_rows

```

```{r assembly_info}

assembly_info.df <- read.delim('tl22-01_assembly_info.txt') %>% 
  filter(!grepl(assembly, pattern = '0\\.9c|0\\.99'),
         !grepl(assembly, pattern = 'SRX5158513_SRR8347514|SRX5158512_SRR8347515'))  
#Filter out any none-0.95 assembly made. And potential Kim19_A duplicate?

#Order experiments by years
expt_num.vec <-ordering(unique(assembly_info.df$experiment), 
                                   order_vector = c(16:22))
#Get experiment numbers and add '0' to single digit numbers.
names(expt_num.vec) <-  paste0('E', 
                              sapply(seq_along(expt_num.vec), function(x){
                                ifelse(nchar(x)==1, 
                                       paste0('0', x),
                                       x)
                              }))

#Manually add Gunalan19 and _Boonkaew20 (in addition to Guna19 and Boonkaew20)
expt_num.vec <- c(expt_num.vec, E07 = 'Gunalan19', E08 = '_Boonkaew20')

#Get the relevant info from samplesheet otherwise loads of rows will be duplicated when joining.
samplesheet_small.df <- mutate(samplesheet, assembly_file = assembly) %>% 
  select(assembly_file, Cat.Sample.stage.s.) %>% unique

assembly_info_samplesheet.df <- left_join(assembly_info.df, samplesheet_small.df, by = 'assembly_file')

assembly_info_samplesheet.df <- mutate(assembly_info_samplesheet.df,
                                        experiment_num = names(expt_num.vec)[match(experiment,
                                                                                  expt_num.vec)])

#Get consistent shapes and colours for the graphs

shapes_val <- 1:length(unique(assembly_info_samplesheet.df$Cat.Sample.stage.s.))
names(shapes_val) <- unique(assembly_info_samplesheet.df$Cat.Sample.stage.s.)

colours_val <- safe_colorblind_palette[1:length(unique(names(expt_num.vec)))]
names(colours_val) <- unique(names(expt_num.vec))

expt_num.df <- data.frame(expt = expt_num.vec, 
                          expt_code = names(expt_num.vec), 
                          col = c(colours_val,colours_val['E07'],colours_val['E08']))

#Get the stages of the life cycle used for the assembly (altering the Gunalan sample name in over1.df since they do not correspond).
over1.df$stage <- assembly_info_samplesheet.df$Cat.Sample.stage.s.[
  match(str_replace(over1.df$assembly, pattern = 'Gunalan19', replacement = 'Guna19'), 
        str_extract(assembly_info_samplesheet.df$assembly, 
                    pattern = '.+(?=_scspades_0.95cdhitest)'))
]

#Get the experiment numbers (altering the Gunalan sample name in over1.df since they do not correspond).
over1.df$experiment_num <- assembly_info_samplesheet.df$experiment_num[
  match(str_replace(over1.df$assembly, pattern = 'Gunalan19', replacement = 'Guna19'), 
        str_extract(assembly_info_samplesheet.df$assembly, 
                    pattern = '.+(?=_scspades_0.95cdhitest)'))
]

#Save the colour scheme
assembly_info_samplesheet.df$expt_colour <- colours_val[match(assembly_info_samplesheet.df$experiment_num,
                                                  names(colours_val))]

#Save number of pir transcripts over 1 TPM (altering the Gunalan sample name in over1.df since they do not correspond).
pir_over1.df <- filter(over1.df, pir)

assembly_info_samplesheet.df$num_mRNA_pir_over200aa_over1TPM <- pir_over1.df$over1_num[
  match(str_extract(assembly_info_samplesheet.df$assembly, 
                    pattern = '.+(?=_scspades_0.95cdhitest)'),
        str_replace(pir_over1.df$assembly, pattern = 'Gunalan19', replacement = 'Guna19'))
]

# Make NAs zero
assembly_info_samplesheet.df$num_mRNA_pir_over200aa_over1TPM <- ifelse(
  is.na(assembly_info_samplesheet.df$num_mRNA_pir_over200aa_over1TPM),
  0,
  assembly_info_samplesheet.df$num_mRNA_pir_over200aa_over1TPM 
)

write.table(assembly_info_samplesheet.df,
            'tl22-01_assembly_info_extra.txt')

```

```{r over1_plots}

(plot <- ggplot(data = filter(over1.df, !pir),
                aes(x = total, 
                    y = over1_perc,
                    col = experiment_num,
                    shape = stage)) +
    geom_point(size = 1, stroke = 2, alpha = .75) +
  scale_shape_manual(values=shapes_val,
                     name = 'Life cycle stage',
                     guide = guide_legend(ncol = 2)) +
  scale_color_manual(values = colours_val[!names(colours_val) %in% 'E10'],
                     name = 'Experiment',
                     guide = guide_legend(ncol = 2)) +
    xlab('Number of assembled transcripts') +
    ylab('Percentage of transcripts transcribed over 1 TPM') +
    theme_classic())

if(global_save_var){
  ggsave(plot = plot,
         'plots/assemblies_numTranscriptsOver1tpm_totalTranscripts.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}

(plot <- ggplot(data = filter(over1.df, pir),
       aes(x = total, 
           y = over1_perc_pirs,
           col = experiment_num,
           shape = stage)) +
    geom_point(size = 1, stroke = 2, alpha = .75) +
    #Subset the shapes/colours_val since not all the experiments/stages have any predicted pirs and are hence not included.
    scale_shape_manual(values=shapes_val[
      names(shapes_val) %in% unique(filter(over1.df, 
                                           pir)$stage)
    ],
    name = 'Life cycle stage',
    guide = guide_legend(ncol = 2)) +
    scale_color_manual(values = colours_val[
      names(colours_val) %in% unique(filter(over1.df, 
                                            pir)$experiment_num)
    ],
    name = 'Experiment',
    guide = guide_legend(ncol = 2)) +
   xlab(expression(paste('Number of assembled ',italic(pir),' transcripts'))) +
   ylab(expression(paste('Percentage of ',italic(pir),'s transcribed over 1 TPM'))) +
   theme_classic())

if(global_save_var){
  ggsave(plot = plot,
         'plots/assemblies_numPirOver1tpm_totalTranscripts.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}

```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_analysingAssemblyStats.Rmd
Add many required vivax files and change paths in vivax scripts

@@
 ```{r read_data}
-df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')
+df <- read.delim('tl22-01_assembly_info.txt')
 ```
@@
 ```{r expt_name}
- accession_table <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samples_to_cat.csv')
+ accession_table <- read.csv('samples_to_cat.csv')
 
@@
 ggsave(plot = plot,
-       '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/assemblies_numNovPir_v_totalTranscripts.pdf',
+       'plots/assemblies_numNovPir_v_totalTranscripts.pdf',
        device = 'pdf',

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_analysingNetworkStats.Rmd
Add many required vivax files and change paths in vivax scripts

@@
 ```{r read_data}
-accession_table <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samples_to_cat.csv')
-assembly.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')
-subfam.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-02_VirSubfamNetwork/pir_info/lopez13_vir_subfam.txt', header = FALSE)
+accession_table <- read.csv('samples_to_cat.csv')
+assembly.df <- read.delim('tl22-01_assembly_info.txt')
+subfam.df <- read.delim('lopez13_vir_subfam.txt', header = FALSE)
 colnames(subfam.df) <- c('Geneid','subfam')
-expt.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/scripts/all-cat.list', header = FALSE, sep = '')
+expt.df <- read.delim('all-cat.list', header = FALSE, sep = '')
 colnames(expt.df) <- c('Accession', 'expt')
-expt_num.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/expt_codes.csv')
+expt_num.df <- read.csv('expt_codes.csv')
 ```
@@
 ```{r MCL_parameters}
-# list.files('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1', 
+# list.files('hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1', 
   #          pattern = 'mci.+list') %>% 
@@
 
-network_file_dir <- '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1'
+network_file_dir <- 'networks'
 
@@
 ```{r get_MCLclusters}
-fasta <- read.fasta(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')
+fasta <- read.fasta(file = 'all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')
 
@@
 # write.table(sequence.df,
-#             file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt')
-
-sequence.df <- read.table(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt',
+#             file = 'networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt')
+
+sequence.df <- read.table(file = 'sequence-MCLclusters_tpmOver1.txt',
                           check.names = FALSE)
@@
 #     blast.names <- read.delim(paste0(
-#       '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_',
+#       'hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_',
 #       perc_ident,'pid',
@@
 # write.table(sequence.df.melt,
-#             file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
-sequence.df.melt <- read.table(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
+#             file = 'networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
+sequence.df.melt <- read.table(file = 'sequence-MCLclusters_tpmOver1_longForm.txt')
 ```
@@
 # write.csv(node.df,
-#           '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv',
+#           'networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv',
 #           row.names = FALSE,
@@
 node.df <- read.csv(
-  '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv')
+  'sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv')
 
@@
 # write.csv(cluster_subfam.df,
-#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/cluster_subfam_30pid25lngthI1.4.csv',
+#           file = 'networks/tpmOver1/cluster_subfam_30pid25lngthI1.4.csv',
 #           row.names = FALSE,
@@
 
-# blast_30pid25lngth.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
+# blast_30pid25lngth.df <- read.delim('hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
 #            header = FALSE)
@@
 # write.csv(node_labels_noDup_toSave.df,
-#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/edges_30pid25lngthNoS2S.csv',
+#           file = 'networks/tpmOver1/edges_30pid25lngthNoS2S.csv',
 #           row.names = FALSE,
@@
                     ggsave(plot = plot,
-                           paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_exptComposition_',
+                           paste0('plots/networks_exptComposition_',
                                   include,'_',mcl_name,'.pdf'),
@@
   ggsave(plot = plot,
-         paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_clustComp_Expt.pdf'),
+         paste0('plots/networks_clustComp_Expt.pdf'),
          device = 'pdf',
@@
            ggsave(plot = plot,
-                  paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_exptStep_'
+                  paste0('plots/networks_exptStep_'
                          ,mcl_name,'.pdf'),
@@
            mutate(clust_num = as.numeric(clust_num),
-                  subfam = ifelse(subfam == 'Sal1', 
+                  subfam = ifelse(subfam.x == 'Sal1', 
                                   'unassigned',
-                                  subfam)) %>% 
+                                  subfam.x)) %>% 
            group_by(subfam, clust_num) %>% 
@@
            ggsave(plot = plot,
-                  paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_sal1Composition_',
+                  paste0('plots/networks_sal1Composition_',
                          mcl_name,'.pdf'),
@@
 ```{r pir_sizes}
-
 node_length.melt <- filter(node.melt, 
@@
   ggsave(plot = plot_orf,
-         '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_ORFlength.pdf',
+         'plots/networks_ORFlength.pdf',
          device = 'pdf',
@@
 
-Cluster 8 particularly long?
-
-Looks like there's a little group of longer pirs for some experiments/P01/Sal1.
-Like over 800aa - pretty much all in Cluster 1 (E). 
-I wonder if this is the two groups in the Gephi network.
-
-Cluster 7 and 9 low variability, potentially shorter?
-
-```{r kwtest_length}
-#K-W test of if there is a signif difference between the clusters for length
-
-kruskal.test( node_length.melt$orf_length, factor(node_length.melt$clust_num))
-kruskal.test(node_length_clust1.11_noZhu16.melt$orf_length,
-             factor(node_length_clust1.11_noZhu16.melt$clust_num))
-
-clust_comp <- combn(1:11, m = 2, simplify = FALSE)
-
-#Do the tests for each combination, combione into a data frame and use p.adjust multiple testing adjustment.
-( length_tests.df <- lapply(c('greater', 'less'), 
-                            function(wilcox_text_alt){
-                              lapply(clust_comp, function(comparison){
-  test.stat <- wilcox.test(orf_length ~ clust_num,
-                           data = filter(node_length.melt,
-                                         expt != 'Zhu16_known'),
-                           subset = clust_num %in% comparison,
-                           alternative = wilcox_text_alt)
-  return(data.frame(comparison_1 = comparison[1],
-                    comparison_2 = comparison[2],
-                     stat = test.stat$statistic,
-                     pval = test.stat$p.value,
-                    wilcox_text.alt = wilcox_text_alt))
-                              }) %>% bind_rows
-}) %>% 
-  bind_rows %>% 
-  mutate(padj = round(p.adjust(pval), digits = 5),
-         pval = round(pval, digits = 5)) %>% 
-  arrange(desc(padj)) )
-
-length_med.df <- node_length.melt %>% 
-  group_by(clust_num) %>% 
-  summarise(orf_length_med = median(orf_length))
-
-length_tests.mat <- mutate(length_tests.df,
-                           orf_ratio = ifelse(padj <= 0.05, 
-                                              length_med.df$orf_length_med[
-                                                match(comparison_1, 
-                                                      length_med.df$clust_num)
-                                              ]/length_med.df$orf_length_med[
-                                                match(comparison_2, 
-                                                      length_med.df$clust_num)
-                                              ], 
-                                              NA)) %>% 
-  dcast(.,
-        comparison_1 ~ comparison_2,
-        value.var = 'orf_ratio')
-rownames(length_tests.mat) <- length_tests.mat[,1]
-length_tests.mat <- length_tests.mat[,-1]
-
-# nams <- unique(unlist(clust_comp))
-# 
-# #Need to add the first and last rows due to combn not using 1 as combination_2 and 11 as combination_1.
-# one.mat <- matrix(nrow = nrow(length_tests.mat),
-#        dimnames = list(rep(NA, length.out = nrow(length_tests.mat)), first(nams))) 
-# length_tests.mat <- cbind(one.mat, length_tests.mat)
-# 
-# last.mat <- matrix(ncol = ncol(length_tests.mat),
-#        dimnames = list(last(nams), rep(NA, 
-#                                        length.out = ncol(length_tests.mat))))
-# colnames(last.mat) <- names(length_tests.mat)
-# length_tests.mat <- rbind(length_tests.mat, last.mat)
-# 
-# library(ComplexHeatmap)
-# library(viridis)
-# #Make matrix symmetrical
-# length_tests.mat[
-#   lower.tri(length_tests.mat)
-# ] <- 1/(t(length_tests.mat)[
-#   lower.tri(length_tests.mat)
-# ])
-# 
-# colnames(length_tests.mat) <- rownames(length_tests.mat) <- nams
-# 
-# Heatmap(length_tests.mat, 
-#         cluster_columns = FALSE, 
-#         cluster_rows = FALSE,
-#         na_col = 'black',
-#         col = circlize::colorRamp2(c(0, 1, 1.4), c('blue', 'white', 'red')))
-
-length_tests.df %>%
-  filter(if_any(matches('comparison'),
-                ~grepl(.x,
-                       pattern = '^1$|3|8|10'))) %>%
-  arrange(desc(padj))
-
-length_tests.df %>%
-  filter(if_any(matches('comparison'),
-                ~grepl(.x,
-                       pattern = '7'))) %>%
-  arrange(desc(padj))
-length_tests.df %>%
-  filter(if_any(matches('comparison'),
-                ~grepl(.x,
-                       pattern = '9'))) %>%
-  arrange(desc(padj))
-```
-
 ```{r where_ancestral}
@@
   #Read in BLAST file 
-blast_30pid25lngth.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
+blast_30pid25lngth.df <- read.delim('all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
            col.names = outfmt6_headers) %>% 
@@
 
-```{r}
+```{r vir14}
 vir14_blast <- blast_30pid25lngth.df %>% 
